### 2016-05-12 code/rmarkdown/figures.Rmd
+ initial commit + directory structure + vignette from rapr package

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex2.tex]
---

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)

```

### 2016-05-12 code/rmarkdown/supporting-information.Rmd
+ initial commit + directory structure + vignette from rapr package

---
fontsize: 11pt
documentclass: article
bibliography: Endnote_lib.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble-latex.tex, preamble-latex3.tex]
---
## Supporting Information
All data and code is stored in an online repository (\url{www.github.com/paleo13/rapr-manuscript}) to replicate analyses in this study. See the \textit{README} file for instructions.

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)
```

### 2016-06-14 code/rmarkdown/article.Rmd
* version 1 + debugging parameters work

---
title: "rapr: Representative and Adequate Prioritisations in R"
author: |
  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
  | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: ["conservation", "biodiversity", "MILP", "reserve-selection"]
abstract: "A central aim in conservation is to maximise the long-term persistence of biodiversity. To fulfil this aim, reserve networks are used to safeguard biodiversity patterns (eg. species, populations) and processes (eg. evolutionary processes that underpin genetic variation). Reserve selection is often formulated as an optimisation problem to identify cost-effective prioritisations. However, most existing decision support tools are based on formulations that are well suited for preserving biodiversity patterns, but not biodiversity processes. To fill this gap in the conservation planning toolbox, we developed the \texttt{rapr R} package. This \texttt{R} package provides a toolkit to guide reserve selection using novel formulations of this problem. Here, we explore the functionality of this R package using a simulation study and two case-studies. We demonstrate how explicitly considering biodiversity processes can alter a prioritisation. In most cases, we found that only a few additional plannning units are required to sufficiently preserve of biodiversity processes."
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble-latex.tex
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
vignette: >
  %\VignetteIndexEntry{rapr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
# load .rda file
try(session::restore.session('data/final/results.rda'))

# set options
options(error=NULL)
```

<!--  title page -->
Short running title: Representative and adequate prioritisations \newline
Word count: XXXXX \newline
Number of references: XXXX \newline
Number of figures: XXXX \newline
Number of tables: XXXX \newline
Number of textboxes: XXXX \newline
Number of equations: XXXX \newline
\newpage

## Introduction
<!-- conservation and reserve networks -->
The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255]. To achieve this, conservation actions must preserve biodiversity patterns (eg. species, populations) and the processes that sustain them. One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r440]. Reserve networks buffer species from threatening processes [eg. urbanisation; @r255] and set the stage for direct management interventions [eg. captive breeding and reintroduction programs; @r290]. However, the resources available for conservation action are limited, and so reserve networks must be sited in places that satisfy conservation objectives for minimum cost [@r255]. To achieve this, reserve selection is often formulated as an optimisation problem and then solved to identify cost-effective candidate reserve systems [prioritisations; @r255].

<!-- biodiversity processes are important -->
To fulfil the overarching aims of conservation, reserve networks must preserve both ecological and evolutionary processes [@r3; @r255]. Ecological processes, such as predator-prey interactions, pollination, and decomposition, are required for biodiversity to persist over short time-scales. Typically, they operate over small geographic domains--with exceptions such as migration and refugial habitats--and can be preserved using suitably large planning units [@r446] that each contain a discrete unit of habitat [@r448]. On the other hand, evolutionary processes are required for biodiversity to persist over long time-scales, and they typically operate over large geographic domains. Protected areas must preserve adaptive evolutionary processes to foster resilience against environmental change [eg. climate change; @r247]. Protected areas must also preserve neutral evolutionary processes [@r5; @r67; @r63]. These processes are associated with the breakdown of gene flow between populations. They are important for maintaining genetic diversity, and avoiding inbreeding depression. In recent decades, a wealth of data on biodiversity processes has become available to conservation planners [eg. bioclimatic, genetic, and trait data; @r33; @r453; @481]. Yet this data is only rarely used to guide conservation planning [@r61]. This is due to that fact that existing reserve selection tools focus on preserving biodiversity patterns or processes--but not both.

<!-- parametrisation of biodiversity patterns and processes in conventional conservation planning tools -->
Many decision support tools have been developed to help identify cost-effective prioritisations [eg. \texttt{ ConsNet}, @r446; \texttt{ Marxan}, @r22; \texttt{ Zonation}, @r429]. Typically, these tools are used to deliver a prioritisation that preserves a set of target species (or features). Given a set of species, decision makers can use these tools to preserve biodiversity patterns by generating a prioritisation that secures an adequate proportion of each species' range. These tools, however, are not well-suited for biodiversity processes [but see @r203]. To preserve biodiversity processes, a prioritisation must preserve a representative sample of each species. For instance, to preserve predator-prey interactions, a prioritisation must preserve individuals from each predator and prey species in the same area. To preserve adaptive evolutionary processes, a prioritisation must preserve the adaptive landscape of each species--it must preserve individuals experiencing different selection pressures [@r63]. To preserve neutral evolutionary processes, a prioritisation must secure individuals descended from each of the genetic lineages that comprise each species [@r5]. To accommodate such information into a multi-species conservation planning exercise, species ranges must be partitioned into different groups [eg. using habitat data; @r2] as a pre-processing step to be compatible with existing tools. However, this approach assumes that the biodiversity processes that operate across a species' range can be split into discrete units. Yet data on biodiversity processes is often continuous and hyper-dimensional, and often cannot be reduced to a few categories without significant information loss [@r218].

<!-- aims of this paper -->
Today, one of the key issues in reserve selection is the lack of a unifying decision support tool that can accommodate data on biodiversity patterns and processes in a multi-species context. To fill this void, we present the \texttt{rapr R} package. This \texttt{R} package uses novel formulations of the reserve selection problem to provide decision makers with the tools to generate prioritisations that preserve biodiversity patterns and processes. We also explore the functionality of this \texttt{R} package using simulated species and a case-study conservation planning scenario.

## Problem formulations
The \texttt{rapr R} package uses two novel formulations of the reserve selection problem to identify cost-effective prioritisations. These formulations share many constraints and variables. For brevity, the variables used by both formulations will be defined. Biodiversity features are defined as the entity(s) that the prioritisation is required to preserve (eg. species, populations). Spatial attributes are defined as the variation across the species' range that the prioritisation is required to sample. They can be intrinsic (eg. genetic or trait variation) or extrinsic (eg. environmental variation) to the feature. These attributes should be related to the biodiversity processes that the decision maker aims to preserve.

Each attribute is conceptualised as a space. This space is termed an "attribute space". Each planning unit is thought to occupy a single point inside each space. For example, a decision maker may require a prioritisation that preserves populations along climatic gradients. To achieve this, the decision maker might use an "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this environmental space. By associating planning units with climatic data, they can be mapped from geographic space to this environmental attribute space.

Demand points are points that also exist in an attribute space. They are designated by the decision maker to indicate regions of the attribute space that should be preserved in the prioritisation (see below for discussion on how demand points can be generated for real-world datasets). The amount of variation in the attribute space that a prioritisation secures is a function of the distance between each demand point and each selected planning unit in the attribute space. The shorter the distances between the demand points and the planning units; the better the prioritisation is at securing the variation in the spatial attribute. To convert these amounts to a proportion--a meaningful unit for a decision maker--the distances between the selected planing units and the demand points are scaled by the distances between the demand points to the centroid of the demand points. In any attribute space there may exist points that are impossible (eg. mean annual rainfall -5 mm), do not occur in the study area (eg. mean annual temperature 30$^{\circ}$C in Antarctica). Additionally, there may be some regions that are desirable for some features and undesirable for others (eg. conditions known to be outside the physiological tolerance of a species). Thus a different set of demand points and weights are used for each attribute space and each feature. By placing demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritisations secure the feature in planning units with spatial attributes that are desirable for that feature.

To illustrate these concepts, we will briefly describe a conservation planning scenario example involving an attribute spaces and demand points. A decision maker may wish to develop a prioritisation for a single species. This species has four populations in the study area. However, the decision maker can only afford to preserve three population. The decision maker needs to preserve the adaptive landscape of the species, and preserve populations inhabiting different environmental conditions. To describe environmental variaiton, the decision maker obtained data on the environmental conditions (rainfall (ml) and temperature ($^{\circ}$C)) where each population was found. The decision maker then used this environmental data to construct a two-dimensional environmental attribute space. Next, the decision maker generated demand points as equi-distant points between the range of values where the populations were found [$\pm$ 20% to avoid edge effects; @r218]. By comparing the distribution of the demand points to the distribution of the populations in the attribute space, the decision maker can identify a suitable prioritisation (Figure 1). We can see that if the decision maker preserves both populations $A$ and $C$, they will effectively "double-up" on the same environmental characteristics, and in turn their waste resources. Instead, a more representative sample of the intra-specific variation could be preserved by securing populations $A$, $B$, and $D$. This example demonstrates how the inclusion of biodiversity processes can guide the reserve selection process.

The problem formulations used to guide reserve selection in the \texttt{rapr R} package are based on a combination of the \texttt{Marxan} reserve selection problem and the uncapacitated facility location problems [@r451]. All mathematical terms defined hereafter are described in Table S1 for convenience. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the variable. Define $F$ to be the set of features (indexed by $f$). Let $J$ be a set of planning units (indexed by $j$). Also, let $A_j$ denote the area, and $C_j$ denote the cost of preserving planning unit $j \in J$. To assess the extent to which each feature is secured in a given prioritisation, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$. The level of fragmentation associated with a prioritisation is parametrised as the net exposed boundary length. Let the shared edges between each planning unit $j \in J$ and $k \in J$ be $e_{jk}$.

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points. Demand points with greater weight $\lambda_{fsi}$ are more important, and the optimal solution will be likely to select planning units close to highly weighted demand points. As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.

Targets are used to ensure that prioritisations adequately preserve each species. Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient. Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$. Space-based targets ensure that a sufficient proportion of the intra-specific variation is secured. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $a \in A$. For convenience, these both types of targets are expressed as proportions in the \texttt{R} package.

The \texttt{R} package provides two formulations for reserve selection. They are based on the unreliable [@r451] and reliable uncapacitated facility location [@r16] problems. The key difference between them is that the reliable formulation explicitly considers the probability that the planning units are occupied when calculating the proportion of variation that a given solution secures, whereas the unreliable formulation does not.

### Unreliable formulation
In the unreliable formulation, the control variables are the $B$, $T_{s}$, and $\tau_{sa}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag{1a}\\
%
\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag{1b}\\
%
B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag{1c}\\
\end{align*}

The decision variables are the $X_j$ and $Y_{fsij}$ variables.

\begin{align*}
X_j
	&= \begin{cases}
		1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag{2a} \\
		0, & \parbox{25em}{otherwise} \\
	\end{cases} \\
%
Y_{fsij} &= \begin{cases}
		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag{2b} \\
		0, & \parbox{25em}{otherwise} \\
	\end{cases} \\
\end{align*}

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is assigned to a selected planning unit $J$ where $X_j = 1$. The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in  a given solution. Generally, demand points are assigned to the closest selected planning units (unless particularly low space-based targets are used).

The unreliable formulation (URAP) is a defined as a multi-objective optimisation problem.

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag{3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag{3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq \tau_{fs}  & \forall & 0 \leq f \leq F-1, \tag{3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag{3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag{3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag{3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

The objective function (3a) determines the utility of a given prioritisation: a combination of the total cost of a prioritisation and how fragmented it is. Constraints (3b) ensure that all the amount-based targets are met. Constraints (3c) ensure that all the space-based targets are met for each feature and each attribute space. For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted distance is then scaled by the inherent variation in the demand points ($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a proportion conceptually similar to the statistical $R^2$ value, and the constraints ensure that this proportion must be greater than or equal to the space-based target. Constraints (3d) ensure that only one planning unit is assigned to each demand point. Constraints (3e) ensure that demand points are only assigned to selected planning units. Constraints (3f) ensure that the $X$ and $Y$ variables are binary.

### Reliable formulation
The reliable formulation explicitly considers the probability that the planning units are inhabited. As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented. This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s). To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.

Backup levels levels are defined as $r$-levels [similar to failure levels in @r18]. The first backup $r$-level is used to calculate the level of representation when all of the selected planning units are occupied by all $f \in F$. For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level$=0$. This scenario essentially represents $Y_{fsij}$ in the unreliable formulation. The second backup $r$-level is used to assess the level of representation when the closest planning unit to each demand point $i$ is unoccupied. For this scenario, the second closest planning units are assigned at $r$-level$=1$. The third backup $r$-level is used to assess representation when the first two closest planning units are unoccupied. The third closest planning units are assigned at $r$-level$=2$. Continuing on, in this manner, the selected planning units in a prioritisation are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.

A final backup $r$-level when $r=R$ is used to assess the level of representation when the features $f \in F$ do not occupy any selected planning units in a prioritisation. Each demand point $i \in I$ for each $s \in S$ and $f \in F$ is assigned to an "imaginary" planning unit $j=J$ at $r=R$. The distance variables associated with this imaginary planning unit $d_{fsiJ}$ denote the loss of biological value associated with failing to secure a representative sample of feature $f$ in attribute space $s$. However, the $d$ variables are in distance units which are meaningless in this context. Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).

\begin{align*}
& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag{4} \\
& & & 0 \leq s \leq S-1\\
\end{align*}

Moderately-sized conservation planning problems often include several thousand planning units. It is currently not be feasible to solve this problem when considering all possible failure scenarios. As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$. For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level. Cui et al. [-@r16] found that $R=5$ yields similar solutions to $R=J$ when $J >> 5$. However, in most costs the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.

In the reliable formulation, the control variables are the $B$ (1a), $T_{s}$ (1b), $\tau_{sa}$ (1c), $R$, and $M$ variables.

\begin{align*}
R &= \parbox{25em}{number of failure levels} \tag{5a} \\
%
M &= \parbox{25em}{failure multiplier} \tag{5b}\\
%
\end{align*}

The decision variables are the $X_j$ (2a), $Y_{fsijr}$, $P_{fsijr}$ variables.

\begin{align*}
Y_{fsijr} &= \begin{cases}
		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag{6a} \\
		0, & \parbox{25em}{otherwise} \\
	\end{cases} \\
%
P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag{6a}\\
%
\end{align*}

The reliable formulation (RRAP) is a multi-objective optimisation problem.

\begin{align*}
& \text{(RRAP)} & \text{Min } & \text{(3a)}\\
%
& & \text{s.t. } & \text{(3b)}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} P_{fsijr} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag{7a}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7b}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq r \leq R\\
%
& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7c}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J\\
%
& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag{7d}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag{7e}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag{7f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J\\
%
& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag{7g}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J,\\
& & & & & 1 \leq r \leq R\\
%
& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag{7h}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J,\\
& & & & & 0 \leq r \leq R\\
\end{align*}

The objective function for the reliable formulation is the same as for the unreliable formation (3a). Similar to the unreliable formulation, constraints (3b) and (7a) ensure that the amount-based and space-based targets are met. Constraint (7b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$. Constraints (7d) ensure that only selected planning units are assigned to demand points $i \in I$. Constraints (7e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level. Constraints (7f--7g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information]. Constraints (7h) ensure that the $X$ and $Y$ variables are binary.

### Optimisation
The unreliable and reliable formulations are non-linear. However, the non-linear components can be linearised using existing techniques. First, the expression $X_j X_k$ in (3a) can be linearised using methods described by Beyer et al. [-@r426]. Second, the expression $P_{fsijr} Y_{jsijr}$ in (7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al. [-@r16]. Linearised versions of the problems can be solved using commercial exact algorithm solvers.

The \texttt{rapr} R package provides functions to express conservation planning data as an optimisation problems using linearised versions of the unreliable and reliable formulations. These optimisation problems can then be solved to generate prioritisations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}). Presently, academics can obtain a license at no cost from the Gurobi website. After installing the \texttt{Gurobi} software suite, users will need to install the \texttt{Gurobi R} package. This \texttt{R} package can be installed on recent Windows, Mac OSX, and Ubuntu Linux operating systems.

## Methods
To investigate the utility and beaviour of this problem, we conducted a simulation study and two case-studies. These studies involved generating prioritizations using only amount-based targets to represent prioritisations generated using conventional methods (eg. Marxan), and prioritizations using amount-based and space-based targets. All prioritisations were solved to within `r (general.params.LST[[MODE]]$MIPGap * 100)` % of optimality.

### Simulation study
We generated a study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid. We then simulated three species. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability. This species' distribution was based on a uniform distribution (eqn 8a). The second species was simulated to represent a species with an idealised distribution. It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8b). The third species represents a species with two distinct populations. This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8c). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8c. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were generated by calculating the centroids of each planning units and their weights were set as the probability that the planning units were occupied.

\begin{align*}
& P (\text{uniform species} | x, y}) &= 0.1 & \tag{8a}\\
& P (\text{normal species} | x, y}) &= XXXX & \tag{8b}\\
& P (\text{uniform species} | x, y}) &= XXX & \tag{8c}\\
\end{align*}

We generated four prioritisations for each species. First to represent prioritisations generated using conventional conservation planning methods, we generated prioritisations using only `` % amount targets. Second to show how the addition of geographic targets can affect a prioritisation, we  generated prioritisations using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets. Third to represent prioritisations using conventional planning methods that penalise for fragmentation, we generated prioritisations using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`. Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated prioritisations using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.

### Case-study 1
Here we investigated the potential benefits of using space-based targets to secure a representative proportion of a species' realised niche. We selected Queensland, Australia as the study area. This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats--such as rainforests, woodlands, and deserts. We obtained data for 19 bioclimatic variables across the region (at XXX resolution from XXXX; XXXX) and subject them to a principle components analysis to obtain the main gradients of environmental variation. We used the first two principle components (cumulatively explained `r ` % of the total variation) to characterise the environental variation across Queensland, Australia.

We used five bird species in this case study: blue-winged kookaburra (XXX), brown-backed honeyeater (XXX), brown falcon (XXX), and the southern cassory (XXXX)  . These bird species have different range-sizes and life-history strategies. For instance XXXXX is a generalist and occurs across most of Australia, whereas XXXX is a specialist and is only found in the rainforests of north-east Queensland. For each species, we mapped their distriubtion using the following method: we obtained observation data from the Atlas of Living Australia across the whole of Australia (using the \texttt{ALA4R R} package), spatially thinned the data to omit points within $10 km^2$ of each other to ameloriate the effects of sampling bias (using the developmental version of \texttt{spThin R} package; XXXX), and fit 85 % minimum convex polygons (using the \texttt{adehabitatHR R} package; XXX). We used this method because it is reproducible and moderately accurate. We generated `r ` demand points for each species using hyperbox kernels (implemented in the \texttt{hypervolume R} package; XXXX).

We generated two prioritisations. The first prioritisation was generated using `r ` % amount-based targets for all species. The second prioritisation was generated the same amount-based target with an additional ` r ` % niche-based for each species.

### Case-study 2

## Results
### Simulated species

### Case-study 1

### Case-study 2

## Implications and future directions
The \texttt{rapr} R package provides a unified approach to reserve selection. This R package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and processes. Additionally, the package contains funcionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves. Both the simulated and case-study species suggest that conservation planning exercises need to explicitly consider biodiversity processes during the reserve selection process to capture them.

One of the key advantages of the \texttt{rapr R} package is that it is general enough that any spatial variation could be considered an attribute space, regardless of whether this variation is intrinsic or extrinsic to the feature(s). For example, advances in genomic fields produced high resolution data on genetic information (eg. amplified fragment length polymorphisms, AFLPs; single nucleotide polymorphisms, SNPs). By using geostatistical analysis [eg. generalised dissimilarity modelling GDMs and gradient forests; @r370; @r452], this data has been used to generate maps describing the spatial distribution of genomic variation within a species [@r361; @r419]. These maps in turn could be used to construct a genomic attribute space, and in turn, could be used to generate prioritisations that secure a representative sample of genomic variation within a species. However, because the problem formulations used in this package are so general, the tools in this package could be misused, and generate poor quality prioritisations.

The degree to which a prioritisation truly secures a representative sample of a feature depends on the attribute spaces and distribution of demand points chosen by the decision maker. Ultimately, the space-based targets are set as a proportion based on the distribution of the demand-points. As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not be an effective prioritisation in reality. We therefore stress that decision makers must carefully consider which biodiversity processes need to preserved in the prioritisation, and what spatial data can be used to map these processes. Note that maximising one process can be detrimental to another. For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other. Whereas, to maximise the connecitivity of a prioritisation, reserves ineed to be closer to each other. To assist in the selection of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see the `make.DemandPoints` function). These routines essentially use the distribution of a feature in the attribute space to define a polygon. Demand points are then generated as random points within the polygon. A kernel is then fit to the distribution of the feature in the space [using @r231; @r450], and the demand points are weighted based on the estimated density of the feature at the demand points'.

The \texttt{rapr R} package could be further extended to identify more effective prioritisations. First, the formulation of fragmentation used in this package may be too simplistic in some cases (eg. exercises involving multiple species with different dispersal capabilities), and more realistic measures of fragmentation (eg. those used in Zonation) could be used to identify more effective prioritisations. Second, the problem does not consider temporal dynamics. Here, conservation actions are assumed to be implemented simultaneously in all selected planning units and assumed to remain implemented for all time. As a consequence, this \texttt{rapr R} package is not useful for scenarios where actions are implemented during multiple discrete periods in time (eg. actions are made adue to annual funding cycles), or scenarios involving threatening processes that vary across space and time [reviewed in @r11]. Future research may look into incorporating such elements into this \texttt{R} package.

To maximise the long-term persistence of biodiversity--the stated goal of conservation--decision makers need to identify prioritisations that preserve existing patterns of biodiversity and the processes that support them. To achieve this, conservation planners need a decision support tool that can explicitly accommodate biodiversity patterns and processes. Here, we developed the \texttt{rapr R} package to fill this void. By exploring the functiontality of this package using several simulated species, we found that including space-based targets can radically change a prioritisation for the simplest of species.

## Acknowledgements
JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.

## References

### 2016-06-14 code/rmarkdown/figures.Rmd
* version 1 + debugging parameters work

@@
 documentclass: article
-bibliography: Endnote_lib.bib
+bibliography: references.bib
 csl: reference-style.csl
@@
 
+# prelimary processing
+sim.pus.FPLY <- fortify(sim.pus)
+aus.FPLY <- fortify(aus.PLY)
+cs1.pus.FPLY <- fortify(cs1.pus)
+cs2.grid.FPLY <- fortify(cs2.grid.PLY)
+cs2.grid.sub.FPLY <- fortify(cs2.grid.sub.PLY)
+
+# store countries
+data(countriesHigh)
+countries.FPLY <- countriesHigh[
+	countriesHigh$ADMIN %in% c(
+		'Italy', 'Switzerland', 'France', 'Austria',
+		'Germany', 'Slovenia', 'Croatia', 'Hungary',
+		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
+	)
+,] %>% spFortify
+
 ```
+
+## Figures
+
+```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Example of an attribute space.
+This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (ml) values.
+Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found.
+Points represent demand points.
+In this space, populations closer to each other are considered more similar to each other."}
+## generate data
+# population data
+populations <- data.frame(
+	population=LETTERS[1:4],
+	temperature=c(22,29,21.6,34),
+	rainfall=c(6,12,5.5,6)
+)
+# demand point data
+make.dp <- function(x, n, pad.percent=0.2) {
+	padding<-diff(range(x))*pad.percent
+	return(seq(min(x)-padding, max(x)+padding, length.out=n))
+}
+demand.points <- expand.grid(
+	temperature=make.dp(populations$temperature, n=4, pad.percent=0.2),
+	rainfall=make.dp(populations$rainfall, n=4, pad.percent=0.2)
+)
+
+## plot data
+ggplot(aes(x=temperature, y=rainfall), data=populations) +
+	geom_point(data=demand.points, fill='gray70') +
+	geom_text(aes(label=population), size=6) +
+	theme_classic() +
+	xlab(expression('Temperature ('*degree*'C)')) +
+	ylab('Rainfall (mm)') +
+	theme(axis.line.x=element_line(), axis.line.y=element_line())
+```
+
+```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.5, fig.width=4.5, fig.cap='Simulation study.
+Squares denote planning units.
+Each panel shows the distribution for a different simulated species.
+The fill of each planning unit indicates the probability that a given species inhabits it.
+Panel (a) shows the uniformly distributed species, panel (b) shows the normally distributed species, and panel (c) shows the bimodally distributed species.'}
+# prepare data
+sim.pus.occ.FPLY <- ldply(
+	seq_len(nrow(sim.species)),
+	function(i) {
+		curr.DF <- sim.pus.FPLY
+		curr.DF$species <- sim.species$name[i]
+		curr.DF$Probability <- sim.dps[[i]]@weights[as.numeric(curr.DF$id)]
+		return(curr.DF)
+	}
+)
+sim.pus.occ.FPLY$species <- gsub('\n', ' ', as.character(sim.pus.occ.FPLY$species))
+sim.pus.occ.FPLY$species <- factor(as.character(sim.pus.occ.FPLY$species), levels=gsub('\n', ' ', sim.species$name)) 
+# make plot
+ggplot() +
+	geom_polygon(data=sim.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Probability), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 5, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(),
+		legend.position=c(1.15, 0.6)) +
+	facet_wrap(~ species, ncol=3)
+```
+
+```{r, echo=FALSE, fig.height=4.5, fig.width=4.0, fig.cap='Species used in the first case-study.
+Squares denote planning units.
+Each panel shows the distribution of a given species.
+For a given species, planning units inhabited by the species are shown in bright blue and those uninhabited are shown in dark blue.'}
+# prepare data
+cs1.pus.occ.FPLY <- ldply(
+	seq_along(cs1.params.LST[[MODE]]$common.names),
+	function(i) {
+		curr.DF <- cs1.pus.FPLY
+		curr.DF$species <- cs1.params.LST[[MODE]]$common.names[i]
+		curr.DF$Occupancy <- 'absent'
+		curr.DF$Occupancy[which(curr.DF$id %in% filter(cs1.ru@data@pu.species.probabilities, species==i)$pu)] <- 'present'
+		return(curr.DF)
+	}
+)
+cs1.pus.occ.FPLY$species <- factor(as.character(cs1.pus.occ.FPLY$species), levels=cs1.params.LST[[MODE]]$common.names)
+# make plots
+ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.045, 1.035), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+	) +
+	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ species, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=2.5, fig.width=3, fig.cap='Distribution of environmental variation across Queensland, Australia.
+Squares denote planning units.
+The panels show the first and second main gradients of climatic variation.'}
+# prepare data
+cs1.pca.FPTS <- ldply(
+	seq_len(cs1.params.LST[[MODE]]$n.pc.layers),
+	function(i) {
+		curr.DF <- cs1.bioclim.RAST[[i]] %>% 
+			aggregate(10) %>% 
+			rasterToPoints() %>% 
+			as.data.frame()
+		names(curr.DF)[3] <- 'Score'
+		curr.DF$pc <- paste0('Principal component ',i)
+		return(curr.DF)
+	}
+)
+# make plots
+p1 <- ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[1]), aes(x=x, y=y, fill=Score)) +
+	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+	) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ pc)
+p2 <- ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[2]), aes(x=x, y=y, fill=Score)) +
+	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+		) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ pc)
+# render plots
+grid.arrange(p1, p2, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=4.5, fig.width=4.5, fig.cap=paste0('Study area for the second case-study.
+Squares denote planning units.
+Panel (a) shows the planning units occupied by ',cs2.params.LST[[MODE]]$common.name,'.
+Panels (b--c) show the spatial distribution of the ordinated values that describe patterns of genetic variation associated this species.')}
+# prepare full grid data
+cs2.spp.grid.FPLY <- cs2.grid.FPLY
+cs2.spp.grid.FPLY$Title <- 'Occupancy'
+cs2.spp.grid.FPLY$Occupancy <- 'absent'
+cs2.spp.grid.FPLY$Occupancy[which(cs2.grid.PLY@data[[cs2.params.LST[[MODE]]$species.name]][as.numeric(cs2.spp.grid.FPLY$id)]==1)] <- 'present'
+
+# prepare subset grid data
+cs2.spp.grid.sub.FPLY <- cs2.grid.sub.FPLY
+cs2.spp.grid.sub.FPLY$nmds1 <- NA
+cs2.spp.grid.sub.FPLY$nmds2 <- NA
+cs2.nmds1 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d1')]]
+cs2.nmds2 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d2')]]
+
+curr.pos <- match(as.numeric(cs2.spp.grid.sub.FPLY$id), cs2.grid.sub.DF$id)
+cs2.spp.grid.sub.FPLY$nmds1 <- cs2.nmds1[curr.pos]
+cs2.spp.grid.sub.FPLY$nmds2 <- cs2.nmds2[curr.pos]
+
+cs2.spp.grid.sub.FPLY$Cost <- costs.MTX[match(cs2.grid.sub.PLY$id[as.numeric(cs2.spp.grid.sub.FPLY$id)], costs.MTX[,1]),2]
+
+cs2.spp.grid.sub.FPLY$Title1 <- 'Opportunity cost'
+cs2.spp.grid.sub.FPLY$Title2 <- 'Genetic NMDS 1'
+cs2.spp.grid.sub.FPLY$Title3 <- 'Genetic NMDS 2'
+
+# plot species data
+p1 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), alpha=0.8, color='grey10') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.position='none'
+	) +
+	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
+	coord_cartesian(xlim=buffered.range(cs2.grid.FPLY$long, 0), ylim=buffered.range(cs2.grid.FPLY$lat, 0)) +
+	facet_wrap(~ Title)
+# plot cost data
+p2 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Cost), alpha=0.8, color='grey10') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm'),
+		legend.title=element_blank()
+	) +
+	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
+	scale_fill_continuous(NULL) +
+	facet_wrap(~ Title1)
+# plot nmds axis 1
+p3 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds1), alpha=0.8, color='grey10') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm')
+	) + 
+	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
+	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred') +
+	facet_wrap(~ Title2)
+# plot nmds axis 2
+p4 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds2), alpha=0.8, color='grey10') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0),   legend.key.size=unit(3, 'mm')
+	) + 
+	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
+	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred') +
+	facet_wrap(~ Title3)
+# render plots
+grid.arrange(p1, p2, p3, p4, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap=paste0('Simulation study prioritisations.
+Squares denote planning units.
+The fill of each planning unit denotes the probability that it is occupied by a given species.
+The color the border of each planning unit indicates if it was selected for prioritisation.
+Each panel shows the prioritisation resulting from a given species and parameters.
+Each row of panels corresponds to a different species.
+The top row of panels (a,d,g,j ) shows prioritisations generated for the uniformly distributed species, middle row of panels (b,e,h,k) for the normally distributed species, and the bottom row of panels (c,f,i,l) for the bimodally distributed species.
+Each column of panels correpsonds to a different set of parameters used to generate the prioritisation.
+The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
+# prepare data
+sim.pus.sel.FPLY <- ldply(
+	seq_len(nrow(sim.spp.DF)),
+	function(i) {
+		curr.DF <- sim.pus.FPLY %>% mutate(species=sim.spp.DF$species[i], prioritisation=sim.spp.DF$prioritisation[i])
+		curr.selection <- selections(sim.prioritisations[[sim.spp.DF$species[i]]][[sim.spp.DF$prioritisation[i]]])[as.numeric(curr.DF$id)]
+		curr.DF$Selection <-  c('discarded','selected')[curr.selection+1]
+		return(curr.DF)
+	}
+)
+sim.pus.sel.FPLY$prioritisation <- factor(as.character(sim.pus.sel.FPLY$prioritisation),
+	levels=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'))
+# prepare plotting data
+sim.letter.DF2 <- mutate(sim.spp.DF, letter=paste0('(', letters[seq_along(species)],')'))
+sim.letter.DF2$long <- gCentroid(sim.pus[1,])@coords[1]
+sim.letter.DF2$lat <- gCentroid(sim.pus[1,])@coords[2]
+# make plot
+ggplot() +
+	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
+	geom_text(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', size=3) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
+		axis.text=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(),
+		axis.line=element_blank(), axis.line.x=element_blank(), axis.line.y=element_blank(),
+		plot.background=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		panel.background=element_blank(), panel.border=element_blank(), legend.position='none',
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank()) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
+	facet_grid(species ~ prioritisation)
+```
+
+```{r, echo=FALSE, fig.height=2.5, fig.width=4.5, fig.cap=paste0('Prioritisations for the niche-based case-study in Queensland, Australia.
+Squares represent planning units.
+Dark green planning units for selected for protection and pale green units were discarded.
+Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
+Panel (b) shows the planning units selected selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
+# prepare data
+cs1.pus.sel.FPLY <- ldply(
+	names(cs1.prioritisations),
+	function(x) {
+		curr.DF <- cs1.pus.FPLY
+		curr.DF$Prioritisation <- x
+		curr.DF$Solution <- 'discarded'
+		curr.DF$Solution[curr.DF$id %in% which(as.logical(selections(cs1.prioritisations[[x]])))] <- 'selected'
+		return(curr.DF)
+	}
+)
+cs1.pus.sel.FPLY$Prioritisation <- factor(as.character(cs1.pus.sel.FPLY$Prioritisation), levels=names(cs1.prioritisations))
+# make plots
+ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.03, 1.03), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+	) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ Prioritisation, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=5, fig.width=5, fig.cap=paste0('Prioritisations for the genetic case-study in the European Alps.
+Panel (a) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets.
+Panel (b) shows the planning units selected selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs2.params.LST[[MODE]]$genetic.target*100,' % genetic targets.
+The conventions for panels (a--b) are detailed in the caption for Figure 7.
+Panels (c--d) show the solutions in the genetic space.
+In these panels, each point is an average of the ordinations for individuals in each planning unit.
+Each point represents the typical genetic characeristics of individuals sampled in a given planning unit (based on an NMDS of the binary loci data).
+Planning units associated with points that are closer togeather contain individuals with more similar genetic characteristics than planning units that are further apart.
+Dark green planning units were selected for protection, and pale green planning units were discarded.')}
+# prepare spatial data
+cs2.sel.grid.sub.FPLY <- cs2.grid.sub.FPLY
+cs2.sel.grid.sub.FPLY$Title1 <-'Amount target'
+cs2.sel.grid.sub.FPLY$Title2 <- 'Amount & genetic target'
+
+cs2.sel.grid.sub.FPLY$Amount.selection <- 'discarded'
+cs2.sel.grid.sub.FPLY$Amount.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'
+
+cs2.sel.grid.sub.FPLY$Amount.genetic.selection <- 'discarded'
+cs2.sel.grid.sub.FPLY$Amount.genetic.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'
+
+# prepare genetic space plots
+cs2.sel.space.DF <- data.frame(cs2.genetic.AS@spaces[[1]]@planning.unit.points@coords)
+names(cs2.sel.space.DF) <- c('d1', 'd2')
+cs2.sel.space.DF$id <- cs2.genetic.AS@spaces[[1]]@planning.unit.points@ids
+
+cs2.sel.space.DF$Amount.selection <- 'discarded'
+cs2.sel.space.DF$Amount.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'
+
+cs2.sel.space.DF$Amount.genetic.selection <- 'discarded'
+cs2.sel.space.DF$Amount.genetic.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'
+
+# make plots
+p1 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.selection), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		axis.title=element_blank(), legend.position='none'
+	) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
+	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
+	facet_wrap(~ Title1)
+p2 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.genetic.selection), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm')
+	) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
+	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
+	facet_wrap(~ Title2)
+p3 <- ggplot() +
+	geom_point(data=cs2.sel.space.DF, aes(x=d1, y=d2, fill=Amount.selection), pch=21, size=3, color='black') +
+	theme_classic() + xlab('Genetic NMDS Axis 1') + ylab('Genetic NMDS Axis 2') +
+	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line()) +
+	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
+p4 <- ggplot() +
+	geom_point(data=cs2.sel.space.DF, aes(x=d1, y=d2, fill=Amount.genetic.selection), pch=21, size=3, color='black') +
+	theme_classic() + xlab('Genetic NMDS Axis 1') + ylab('Genetic NMDS Axis 2') +
+	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line()) +
+	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
+# render plots
+grid.arrange(p1,p2,p3,p4,ncol=2)
+
+```

### 2016-06-14 code/R/analysis/00-initialization.R

@@ -1,0 +1,71 @@
+#### set default error message
+# parse args
+args <- commandArgs(TRUE)
+if (length(args)>0) {
+	print(args)
+	if (grepl('MODE',args))
+		MODE <- strsplit(grep('MODE', args, value=TRUE), '=', fixed=TRUE)[[1]][[2]]
+}
+	
+#### Load pacakges
+# load bioconductor packages
+
+# load CRAN packages
+suppressMessages(library(stats))
+suppressMessages(library(Hmisc))
+suppressMessages(library(data.table))
+suppressMessages(library(grid))
+suppressMessages(library(gridExtra))
+suppressMessages(library(plotrix))
+suppressMessages(library(plyr))
+suppressMessages(library(dplyr))
+suppressMessages(library(tidyr))
+suppressMessages(library(pander))
+suppressMessages(library(vegan))
+suppressMessages(library(rgeos))
+suppressMessages(library(testthat))
+suppressMessages(library(parallel))
+suppressMessages(library(rworldxtra))
+suppressMessages(library(doParallel))
+suppressMessages(library(english))
+suppressMessages(library(session))
+suppressMessages(library(maptools))
+suppressMessages(library(RcppTOML))
+suppressMessages(library(knitr))
+suppressMessages(library(lazyWeave))
+suppressMessages(library(broom))
+suppressMessages(library(adehabitatHR))
+
+## load github packages
+# devtools::install_github('AtlasOfLivingAustralia/ALA4R')
+suppressMessages(library(ALA4R))
+# devtools::install_github('paleo13/spThin')
+suppressMessages(library(spThin))
+# devtools::install_github('paleo13/raspr')
+suppressMessages(library(rapr))
+# devtools::install_github('paleo13/structurer')
+suppressMessages(library(structurer))
+# devtools::install_github('paleo13/ggplot2')
+suppressMessages(library(ggplot2))
+
+# set pander options
+panderOptions('knitr.auto.asis', FALSE)
+
+# set seed for reproducibility
+set.seed(500)
+
+# set default select method
+select <- dplyr::select
+
+### Load functions
+for (x in dir(file.path('code', 'R', 'functions'), full.names=TRUE)) source(x)
+
+### set parameters
+if (!exists('MODE')) MODE <- 'debug'
+cat('MODE = ',MODE,'\n')
+general.params.LST <- parseTOML('code/parameters/general.toml')
+
+# save workspace
+save.session('data/intermediate/00-initialization.rda', compress='xz')
+
+ 

### 2016-06-14 code/R/analysis/01-load-data.R

@@ -1,0 +1,116 @@
+## load .rda
+session::restore.session('data/intermediate/00-initialization.rda')
+
+### load simulation data
+# all data is simulated using functions in the package - no data to load
+
+### load case-study 1 data
+# load in record bbox
+cs1.record.bbox <- shapefile('data/raw/record_bbox/record_bbox.shp') %>% writeWKT
+
+# load in environmental data
+cs1.bioclim.RAST <- stack('data/raw/BioClim_variables/pca.tif')
+cs1.pca.DF <- read.table('data/raw/BioClim_variables/pca.TXT', skip=94) %>% `names<-`(
+	c('Principle Component', 'Eigen Value', 'Variation explained (%)',
+	'Accumulative variation explained (%)')
+)
+
+### load case-study 2 data
+## compile spatial grid data
+# load in opportunity costs
+cs2.cost.RAST <- raster('data/raw/GRUMP_V1_Population_Density/grumpv1-popdensity.tif')
+
+# load in species population numbers
+cs2.spp.populations.DF <- fread('data/raw/Data_Meirmans_et_al_IntrabioDiv/NumberPopulations.csv', data.table=FALSE)
+
+# load grid cell centroids
+cs2.grid.DF <- fread(
+	'data/raw/Data_Meirmans_et_al_IntrabioDiv/README',
+	data.table=FALSE,
+	skip='cell\tLong\tLat'
+) %>% rename(
+		grid.longitude=Long,
+		grid.latitude=Lat
+) %>% mutate(
+	id=seq_along(grid.latitude)
+)
+
+# load in aflp data
+cs2.spp.aflp.paths <- dir(
+	'data/raw/Data_Meirmans_et_al_IntrabioDiv',
+	'^.*AFLP\\.dat$',
+	full.names=TRUE
+)
+
+cs2.spp.StructureData.LST <- llply(
+	cs2.spp.aflp.paths,
+	read.StructureData
+)
+
+# compile species occurence data
+# load in data
+cs2.spp.loc.paths <- dir(
+	'data/raw/Data_Meirmans_et_al_IntrabioDiv',
+	'^.*locations\\.txt$',
+	full.names=TRUE
+)
+cs2.spp.samples.DF <- ldply(
+	seq_along(cs2.spp.loc.paths),
+	.fun=function(i) {
+		x <- mutate(
[…]

### 2016-06-14 code/R/analysis/02-simulations.R

@@ -1,0 +1,136 @@
+## load .rda
+session::restore.session('data/intermediate/01-load-data.rda')
+
+## load parameters
+sim.params.LST <- parseTOML('code/parameters/simulations.toml')
+
+#### Simulate data
+# make planning units
+sim.pus <- sim.pus(as.integer(sim.params.LST[[MODE]]$number.planning.units))
+
+# simulate species distributions
+sim.spp <- lapply(
+	c('uniform', 'normal', 'bimodal'),
+	sim.species,
+  n=1,
+  x=sim.pus,
+  res=1
+)
+
+# generate coordinates for pus/demand points
+sim.pu.points <- PlanningUnitPoints(
+	coords=gCentroid(sim.pus, byid=TRUE)@coords,
+	ids=seq_len(nrow(sim.pus@data))
+)
+
+# create demand point objects
+sim.dps <- lapply(
+	sim.spp,
+	function(x) {
+		return(
+			DemandPoints(
+				sim.pu.points@coords,
+				c(extract(x, gCentroid(sim.pus, byid=TRUE)))
+			)
+		)
+	}
+)
+
+## create RapUnreliableOpts object
+# this stores parameters for the unreliable formulation problem (ie. BLM)
+sim.ro <- RapUnreliableOpts()
+
+## create RapData object
+# create data.frame with species info
+sim.species <- data.frame(name=c('Uniform\nspecies', 'Normal\nspecies', 'Bimodal\nspecies'))
+names(sim.dps) <- sim.species[[1]]
+
+## create data.frame with species and space targets
+# amount targets denoted with target=0)
+# space targets denoted with target=1)
+sim.targets <- data.frame(
+	species=rep(1:3, each=2),
+  target=rep(0:1),
+  proportion=rep(c(sim.params.LST[[MODE]]$amount.target, sim.params.LST[[MODE]]$space.target), 3)
+)
+
+# calculate probability of each species in each pu
+sim.pu.probabilities <- calcSpeciesAverageInPus(sim.pus, stack(sim.spp))
+
+## create AttributeSpace object
+# this stores the coordinates of the planning units in an attribute space
+# and the coordinates and weights of demand points in the space
+sim.attr.spaces <- AttributeSpaces(
+	list(
+		AttributeSpace(
+			planning.unit.points=sim.pu.points,
+			demand.points=sim.dps[[1]],
[…]

### 2016-06-14 code/R/analysis/03-case-study-1.R

@@ -1,0 +1,142 @@
+## load .rda
+session::restore.session('data/intermediate/01-load-data.rda')
+
+## load parameters
+cs1.params.LST <- parseTOML('code/parameters/case-study-1.toml')
+
+### Prepare data for prioritizations
+## prepare pca layers
+# select layers
+cs1.bioclim.RAST <- cs1.bioclim.RAST[[seq_len(cs1.params.LST[[MODE]]$n.pc.layers)]]
+
+# aggregate env layer to reduce computational burden
+cs1.bioclim.RAST <- aggregate(cs1.bioclim.RAST, cs1.params.LST[[MODE]]$environmental.layer.resolution/1000)
+
+## make planning units
+# download queensland layer
+aus.PLY <- getData('GADM', country='AUS', level=1, path='data/intermediate') %>% spTransform(cs1.bioclim.RAST@crs)
+qld.PLY <- aus.PLY[aus.PLY$HASC_1=='AU.QL',]  %>% gSimplify(1000)
+aus.PLY <- aus.PLY %>% gSimplify(1000)
+
+# create planning units over queensland
+cs1.pus <- blank.raster(qld.PLY, cs1.params.LST[[MODE]]$pu.size) %>% rasterToPolygons()
+cs1.pus@data$id <- seq_len(nrow(cs1.pus@data))
+cs1.pus@proj4string <- cs1.bioclim.RAST@crs
+
+# omit planning units that doesn't overlap with non-NA values in worldclim raster
+cs1.pus.RST <- rasterize(cs1.pus, y=cs1.bioclim.RAST, field='id')
+values.MTX <- zonal(cs1.bioclim.RAST, z=cs1.pus.RST)
+cs1.pus <- cs1.pus[which(rowSums(!is.finite(values.MTX[,-1]))==0),]
+values.MTX <- values.MTX[cs1.pus$id,]
+cs1.pus@data$id <- seq_len(nrow(cs1.pus@data))
+cs1.pus <- spChFIDs(cs1.pus, as.character(seq_len(nrow(cs1.pus@data))))
+
+# clip planning units to coastline
+cs1.pus <- gIntersection(cs1.pus, qld.PLY, byid=TRUE)
+cs1.pus <- SpatialPolygonsDataFrame(cs1.pus, data=data.frame(id=seq_along(cs1.pus@polygons), area=gArea(cs1.pus, byid=TRUE)/100000, cost=1, status=0L))
[…]

### 2016-06-14 code/R/analysis/04-case-study-2.R

@@ -1,0 +1,163 @@
+## load .rda
+session::restore.session('data/intermediate/01-load-data.rda')
+
+## load parameters
+cs2.params.LST <- parseTOML('code/parameters/case-study-2.toml')
+
+### prepare genetic attribute-spaces
+# subset to relevant species
+cs2.params.LST[[MODE]]$species.name <- gsub(' ', '_', cs2.params.LST[[MODE]]$species.name, fixed=TRUE)
+spp.pos <- which(unique(cs2.spp.samples.DF$species) %in% cs2.params.LST[[MODE]]$species.name)
+cs2.spp.StructureData.LST <- cs2.spp.StructureData.LST[spp.pos]
+cs2.spp.samples.DF <- cs2.spp.samples.DF[cs2.spp.samples.DF$species %in% cs2.params.LST[[MODE]]$species.name,]
+
+# run mds
+cs2.spp.nmds.LST <- llply(
+	seq_along(cs2.spp.StructureData.LST),
+	.fun=function(i) {
+		# manually classify loci as neutral or adaptive
+		curr.spp <- bayescanr::BayeScanData(
+			cs2.spp.StructureData.LST[[i]]@matrix,
+			primers=cs2.spp.StructureData.LST[[i]]@loci.names,
+			populations=rep('1', nrow(cs2.spp.StructureData.LST[[i]]@matrix)),
+			labels=cs2.spp.StructureData.LST[[i]]@sample.names
+		)
+		curr.nmds <- bayescanr::nmds(
+			curr.spp,
+			metric='gower',
+			max.stress=cs2.params.LST[[MODE]]$max.stress,
+			min.k=cs2.params.LST[[MODE]]$min.k,
+			max.k=cs2.params.LST[[MODE]]$max.k,
+			trymax=cs2.params.LST[[MODE]]$trymax
+		)
+	}
+)
+
+# store mds rotations for each sample
+cs2.spp.samples.DF <- ldply(
+	seq_along(unique(cs2.spp.samples.DF$species)),
+	.fun=function(i) {
+		cbind(
+			filter(cs2.spp.samples.DF, species==unique(cs2.spp.samples.DF$species)[i]),
+			`names<-`(
+				as.data.frame(cs2.spp.nmds.LST[[i]]$points),
+				paste0('genetic_d',seq_len(cs2.spp.nmds.LST[[i]]$ndim))
+			)
+		)
+	}
+)
+
[…]

### 2016-06-14 code/R/analysis/05-statistical-analysis.R

@@ -1,0 +1,9 @@
+## load .rda
+session::restore.session('data/intermediate/04-case-study-2.rda')
+load('data/intermediate/03-case-study-1.rda')
+load('data/intermediate/02-simulations.rda')
+
+
+## save workspace
+save.session('data/intermediate/05-statistical-analysis.rda', compress='xz')
+ 

### 2016-06-14 code/R/analysis/06-exports.R

@@ -1,0 +1,7 @@
+ ## load .rda
+session::restore.session('data/intermediate/05-statistical-analysis.rda')
+
+
+
+## save workspace
+save.session('data/final/results.rda', compress='xz')

### 2016-06-14 code/R/functions/data.R

@@ -1,0 +1,27 @@
+#' WGS1984 CRS
+#'
+#' This object stores the coordinate reference system (CRS) for latitude/longitude data (WGS1984).
+#'
+#' @seealso \code{\link[sp]{CRS}}
+#' @examples
+#' wgs1984 <- CRS('+init=epsg:4326')
+wgs1984 <- CRS('+init=epsg:4326')
+
+#' Europe Lambert Conformal Conic CRS
+#'
+#' This object stores the equal-area coordinate reference system (CRS) for the Europe Lambert Conformal Conic projection.
+#'
+#' @seealso \code{\link[sp]{CRS}}
+#' @examples
+#' europeEA <- CRS('+init=esri:102014')
+europeEA <- CRS('+init=esri:102014')
+
+#' Europe Equidistant Conic
+#'
+#' This object stores the equi-distant coordinate reference system (CRS) for the  projection.
+#'
+#' @seealso \code{\link[sp]{CRS}}
+#' @examples
+#' europeED <- CRS('+init=esri:102031')
+europeED <- CRS('+init=esri:102031')
+

### 2016-06-14 code/R/functions/extractResults.R

@@ -1,0 +1,46 @@
+#' Extract results from RapSolved object
+#'
+#' This function extracts results from a RapSolved object
+#'
+#' @param x \code{RapSolved} object.
+#' @export
+extractResults <- function(x) {
+	# extract score
+	score.DF <- data.frame(
+		Solution=rep(seq_along(x@results@summary$Score), each=length(x@data@species$name)),
+		Score=rep(x@results@summary$Score,each=length(x@data@species$name))
+	)
+	# extract amount held
+	amount.held.DF <- ldply(
+		seq_len(nrow(x@results@summary)),
+		function(r) {
+			data.frame(
+			Species=x@data@species$name,
+			amount.held=c(amount.held(x,y=r))
+		)
+	})
+	# extract space names
+	space.names <- sapply(x@data@attribute.spaces, 'slot', 'name')
+	# extract space held
+	space.held.DF <- ldply(
+		seq_len(nrow(x@results@summary)),
+		function(r) {
+		ldply(
+			x@data@species$name, 
+			function(y) {
+			# get space names
+			curr.DF <- as.data.frame(space.held(x, y=r, species=y))
+			names(curr.DF) <- space.names
+			# make data.frame
+			return(curr.DF)
+		})
+	})
+	# return data.frame
+	return(
+		cbind(
+			score.DF,
+			amount.held.DF,
+			space.held.DF
+		)
+	)
+}

### 2016-06-14 code/R/functions/make.genetic.AttributeSpace.R

@@ -1,0 +1,27 @@
+#' Make AttributeSpace using genetic data
+#'
+#' This function creates an \code{AttributeSpace} object with for a single species. 
+#'
+#' @param site.data \code{data.frame} with coordinates for planning units.
+#' @param species.data \code{data.frame} with coordinates of demand points for the species.
+#' @return \code{AttributeSpace}.
+#' @seealso \code{\link[rapr]{AttributeSpace}}, \code{\link[rapr]{AttributeSpace}}.
+#' @export
+make.genetic.AttributeSpace <- function(site.data,species.data,species) {
+	# extract ids of pus where the species is found
+	curr.ids<-which(rowSums(is.na(site.data))==0)
+	# return attribute space
+	return(
+		AttributeSpace(
+			planning.unit.points=PlanningUnitPoints(
+				coords=as.matrix(site.data[curr.ids,,drop=FALSE]),
+				ids=curr.ids
+			),
+			demand.points=DemandPoints(
+				coords=as.matrix(species.data),
+				weights=rep(1, nrow(species.data))
+			),
+			species=as.integer(species)
+		)
+	)
+}

### 2016-06-14 code/R/functions/make.targets.R

@@ -1,0 +1,48 @@
+#' Make target data using attribute spaces
+#'
+#' This function takes a set of species names and 
+#' attribute spaces and generates a table with targets.
+#'
+#' @param species \code{character} species names.
+#' @param genetic.space \code{AttributeSpace} object with genetic space data.
+#' @param amount.target \code{numeric} amount-based target
+#' @param genetic.target \code{numeric} genetic-based target
+#' @seealso \code{\link[rapr]{AttributeSpace}}.
+#' @export
+make.targets <- function(species, genetic.spaces, amount.target=0.2, genetic.target=0.2) {
+	# amount targets
+	amount.targets.DF <- data.frame(
+		species=seq_along(species),
+		target=rep(0, length(species)),
+		proportion=rep(amount.target, length(species)),
+		name=paste0('amount_',species)
+	)
+	# genetic spaces
+	genetic.targets.DF <- ldply(seq_along(genetic.spaces), .fun=function(i) {
+		# get species position
+		spp.pos <- which(sapply(genetic.spaces[[i]]@demand.points, function(y) {length(y@weights)})>2)
+		# return data.frame
+		spp.DF <- data.frame(
+			species=seq_along(species),
+			target=rep(i, length(species)),
+			proportion=replace(
+				rep(NA_real_, length(species)),
+				spp.pos,
+				genetic.target
+			),
+			name=replace(
+				rep('null',length(species)),
+				spp.pos,
+				paste0('genetic_',species[spp.pos])
+			)
+		)
+		return(spp.DF)
+	})
+	# return
+	return(
+		mutate(rbind(amount.targets.DF, genetic.targets.DF),
+		target=as.integer(target), species=as.integer(species))
+	)
+}
+
+

### 2016-06-14 code/R/functions/misc.R

@@ -1,0 +1,115 @@
+#' @include dependencies.R genetic.surrogates-internal.R
+
+#' Buffered range 
+#'
+#' This function returns the range of a vector that has been buffered by a percentage.
+#' 
+#' @param x \code{numeric} vector.
+#' @return \code{numeric} vector of two elements.
+#' @seealso \code{\link[base]{range}}.
+#' @examples
+#' buffered.range(c(0,0.5,1), 0.05)
+#' @export
+buffered.range <- function(x, percent=0.05) {
+	range(x)+(c(-1, 1) * ((diff(range(x))*percent)))
+}
+
+#' spFortify
+#' 
+#' This function is a convienience function designed to prepare a \code{Spatial*DataFrame} object
+#' for visualisation using ggplot2.
+#' @param x \code{\link[sp]{SpatialLinesDataFrame}}, \code{\link[sp]{SpatialPointsDataFrame}}, or \code{\link[sp]{SpatialPolygonsDataFrame}} object.
+#' @param columns \code{character} with column names to include in returned object.
+#' @return \code{data.frame}
+#' @seealso \code{\link[ggplot2]{fortify}}, \code{\link[sp]{SpatialLinesDataFrame}}, \code{\link[sp]{SpatialPointsDataFrame}} \code{\link[sp]{SpatialPolygonsDataFrame}}.
+#' @export
+spFortify <- function(x, columns=names(x@data)) {
+	if ('id' %in% names(x))
+		x@data <- rename(x@data, id2=id)
+	x$fid <- seq_len(nrow(x@data))
+	f <- fortify(x, region="fid")
+	f <- merge(f, x@data, by.x='id', by.y='fid')
+	f <- f[,which(names(f) != 'fid')]
+	return(f)
+}
+
+#' pretty.pval
+#' 
+#' This function returns the rounded version of a p-value for use in text along with the correct arithmatic operator.
+#' @param x \code{numeric} p-value
+#' @return \code{character} object.
+#' @example
+#' pretty.pval(0.04)
+#' pretty.pval(0.0000001)
+#' pretty.pval(0.565656)
+#' @export
+pretty.pval <- function(x) {
+	if (x > 1)
+		return('> 1')
+	if (x > 0.1)
+		return('> 0.1')
+	if (x > 0.05)
+		return('> 0.05')
+	if (x < 0.001)
+		return('< 0.001')
+	if (x < 0.01)
+		return('< 0.01')
[…]

### 2016-06-14 code/R/functions/species.prioritisation.R

@@ -1,0 +1,20 @@
+#' Generate prioritisation using specific targets
+#'
+#' This function is a wrapper for updating targets in a \code{RapUnsolved} object and solving it.
+#'
+#' @param x \code{RapUnsolved} object.
+#' @param amount.targets \code{numeric} amount-based targets to use.
+#' @param genetic.targets \code{numeric} genetic targets.
+#' @param ... arguments passed to solve.
+#' @return \code{RapSolved}
+species.prioritisation <- function(x, amount.targets, genetic.targets, ...) {
+	# init
+	genetic.pos <- grep('^genetic\\_.*$', x@data@targets$name)
+	# update targets
+	x@data@targets[which(x@data@targets$target==0),'proportion'] <- amount.targets
+	x@data@targets[genetic.pos,'proportion'] <- genetic.targets
+	# solve object
+	ret<-solve(x, ...)
+	return(ret)
+}
+

### 2016-06-15 code/rmarkdown/article.Rmd
+ fixed bug in makefile + finished figures + made progress in methods

@@
 
-## Methods
-To investigate the utility and beaviour of this problem, we conducted a simulation study and two case-studies.
+To understand the beaviour of this problem and showcase its utility, we conducted a simulation study and two case-studies.
 These studies involved generating prioritizations using only amount-based targets to represent prioritisations generated using conventional methods (eg.
-Marxan), and prioritizations using amount-based and space-based targets.
-All prioritisations were solved to within `r (general.params.LST[[MODE]]$MIPGap * 100)` % of optimality.
+Marxan), and prioritizations using amount-based and space-based targets using the unreliable formulation.
 
-### Simulation study
-We generated a study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid.
+## Simulation study
+### Methods
+We generated a study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Figure 2).
 We then simulated three species.
-The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability.
+The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Figure 2a).
 This species' distribution was based on a uniform distribution (eqn 8a).
-The second species was simulated to represent a species with an idealised distribution.
+The second species was simulated to represent a species with an idealised distribution (Figure 2b).
 It has a core range area, and is less likely to be found in areas that are distant from the core area.
 This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8b).
-The third species represents a species with two distinct populations.
+The third species represents a species with two distinct populations (Figure 2c).
 This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8c).
@@
 \begin{align*}
-& P (\text{uniform species} | x, y}) &= 0.1 & \tag{8a}\\
-& P (\text{normal species} | x, y}) &= XXXX & \tag{8b}\\
-& P (\text{uniform species} | x, y}) &= XXX & \tag{8c}\\
+& P (\text{uniform species} | x, y) &= 0.1 & \tag{8a}\\
+& P (\text{normal species} | x, y) &= XXXX & \tag{8b}\\
+& P (\text{uniform species} | x, y) &= XXX & \tag{8c}\\
 \end{align*}
@@
 We generated four prioritisations for each species.
-First to represent prioritisations generated using conventional conservation planning methods, we generated prioritisations using only `` % amount targets.
+First to represent prioritisations generated using conventional conservation planning methods, we generated prioritisations using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
 Second to show how the addition of geographic targets can affect a prioritisation, we  generated prioritisations using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets.
@@
 
-### Case-study 1
-Here we investigated the potential benefits of using space-based targets to secure a representative proportion of a species' realised niche.
+### Results
+
+## Case-study 1
+### Methods
+Here we investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species.
 We selected Queensland, Australia as the study area.
-This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats--such as rainforests, woodlands, and deserts.
-We obtained data for 19 bioclimatic variables across the region (at XXX resolution from XXXX; XXXX) and subject them to a principle components analysis to obtain the main gradients of environmental variation.
-We used the first two principle components (cumulatively explained `r ` % of the total variation) to characterise the environental variation across Queensland, Australia.
+This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats.
+We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subject them to a principal components analysis (using ArcMap 10.3.1).
+We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principle components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Figure 3).
 
-We used five bird species in this case study: blue-winged kookaburra (XXX), brown-backed honeyeater (XXX), brown falcon (XXX), and the southern cassory (XXXX)  .
-These bird species have different range-sizes and life-history strategies.
-For instance XXXXX is a generalist and occurs across most of Australia, whereas XXXX is a specialist and is only found in the rainforests of north-east Queensland.
-For each species, we mapped their distriubtion using the following method: we obtained observation data from the Atlas of Living Australia across the whole of Australia (using the \texttt{ALA4R R} package), spatially thinned the data to omit points within $10 km^2$ of each other to ameloriate the effects of sampling bias (using the developmental version of \texttt{spThin R} package; XXXX), and fit 85 % minimum convex polygons (using the \texttt{adehabitatHR R} package; XXX).
-We used this method because it is reproducible and moderately accurate.
-We generated `r ` demand points for each species using hyperbox kernels (implemented in the \texttt{hypervolume R} package; XXXX).
+We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), the brown falcon (_Falco berigora_), and the pale headed rosella (_Platycercus adscitus_).
+We used these bird species because they have different evolutionary histories and life-history strategies.
+We then mapped the extent of occurrence for each species using the following method: we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+We used this method because it is reproducible using freely available data and moderately accurate.
+
+We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Figure S2).
+To achieve this, for each species, we generated random points inside the species range and at each point extracted the principal component values at that location.
+We then fit hyperbox kernels to the distribution of these points to characterise the realised niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231].
+We then generated uniformly distributed points inside the species' kernels and estimated the density of the training points at the uniformly generated points.
+These uniformly distributed points and associated density estimated were used as demand point coordinates and weights (respectively).
 
 We generated two prioritisations.
-The first prioritisation was generated using `r ` % amount-based targets for all species.
-The second prioritisation was generated the same amount-based target with an additional ` r ` % niche-based for each species.
+The first prioritisation was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species.
+The second prioritisation was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based for each species.
 
-### Case-study 2
+### Results
 
-## Results
-### Simulated species
+## Case-study 2
+### Methods
+Here we used of space-based targets to generate a prioritisation to secure a representative sample of a species' intra-specific genetic variation.
+We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
+Although this dataset contains multiple plant species, we used data for the `r  cs2.params.LST[[MODE]]$common.name` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}).
+This species was chosen because it provides is locally wide-spread and has been shown to exhibit significant genetic structure [@r454].
+Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
+22.3 km $\times$ 25 km; Figure 5a).
+They visited every second grid cell, and if the `r cs2.params.LST[[MODE]]$common.name` was detected in a cell, samples were collected from three individuals.
+Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
+In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.
 
-### Case-study 1
+We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the individual-level presence/absence genetic data into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
+These continuous variables described the main theme of genetic variation within the species.
+We calculated the average of the values associated with individuals in each grid cell.
+These values were used to create a genetic attribute space (Figures 5c and 5d).
 
-### Case-study 2
+The grid cells were used as planning units.
+The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
+Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points.
+To ensure that the solutions did not prioritise particularly costly areas that could not be used for conservation action, we included opportunity cost data (Figure 5b).
+We obtained population density data [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
+
+We generated two prioritisations.
+The first prioritisation was generated using an `r cs2.params.LST[[MODE]]$amount.target*100` % amount-based target.
+The second prioritisation was generated using the same amount-based target and an additional `r cs2.params.LST[[MODE]]$genetic.target*100` % genetic-based target.
+
+### Results
 
@@
 The \texttt{rapr} R package provides a unified approach to reserve selection.
+One of the key advantages of the \texttt{rapr R} package is that it is general enough that any spatial variation could be considered an attribute space, regardless of whether this variation is intrinsic or extrinsic to the feature(s).
 This R package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and processes.
-Additionally, the package contains funcionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
+Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
 Both the simulated and case-study species suggest that conservation planning exercises need to explicitly consider biodiversity processes during the reserve selection process to capture them.
-
-One of the key advantages of the \texttt{rapr R} package is that it is general enough that any spatial variation could be considered an attribute space, regardless of whether this variation is intrinsic or extrinsic to the feature(s).
-For example, advances in genomic fields produced high resolution data on genetic information (eg.
-amplified fragment length polymorphisms, AFLPs; single nucleotide polymorphisms, SNPs).
-By using geostatistical analysis [eg.
-generalised dissimilarity modelling GDMs and gradient forests; @r370; @r452], this data has been used to generate maps describing the spatial distribution of genomic variation within a species [@r361; @r419].
-These maps in turn could be used to construct a genomic attribute space, and in turn, could be used to generate prioritisations that secure a representative sample of genomic variation within a species.
-However, because the problem formulations used in this package are so general, the tools in this package could be misused, and generate poor quality prioritisations.
 
@@
 
-The \texttt{rapr R} package could be further extended to identify more effective prioritisations.
-First, the formulation of fragmentation used in this package may be too simplistic in some cases (eg.
-exercises involving multiple species with different dispersal capabilities), and more realistic measures of fragmentation (eg.
-those used in Zonation) could be used to identify more effective prioritisations.
-Second, the problem does not consider temporal dynamics.
-Here, conservation actions are assumed to be implemented simultaneously in all selected planning units and assumed to remain implemented for all time.
-As a consequence, this \texttt{rapr R} package is not useful for scenarios where actions are implemented during multiple discrete periods in time (eg.
-actions are made adue to annual funding cycles), or scenarios involving threatening processes that vary across space and time [reviewed in @r11].
-Future research may look into incorporating such elements into this \texttt{R} package.
-
 To maximise the long-term persistence of biodiversity--the stated goal of conservation--decision makers need to identify prioritisations that preserve existing patterns of biodiversity and the processes that support them.
@@
 Here, we developed the \texttt{rapr R} package to fill this void.
-By exploring the functiontality of this package using several simulated species, we found that including space-based targets can radically change a prioritisation for the simplest of species.
+By exploring the functionality of this package using several simulated species, we found that including space-based targets can radically change a prioritisation for the simplest of species.
 

### 2016-06-15 code/rmarkdown/figures.Rmd
+ fixed bug in makefile + finished figures + made progress in methods

--- 2016-05-12 code/rmarkdown/figures.Rmd
@@
 
+# prelimary processing
+sim.pus.FPLY <- fortify(sim.pus)
+aus.FPLY <- fortify(aus.PLY)
+cs1.pus.FPLY <- fortify(cs1.pus)
+cs2.grid.FPLY <- fortify(cs2.grid.PLY)
+cs2.grid.sub.FPLY <- fortify(cs2.grid.sub.PLY)
+
+# store countries
+data(countriesHigh)
+countries.FPLY <- countriesHigh[
+	countriesHigh$ADMIN %in% c(
+		'Italy', 'Switzerland', 'France', 'Austria',
+		'Germany', 'Slovenia', 'Croatia', 'Hungary',
+		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
+	)
+,] %>% spFortify
+
 ```
+
+## Figures
+
+```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Example of an attribute space.
+This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (ml) values.
+Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found.
+Points represent demand points.
+In this space, populations closer to each other are considered more similar to each other."}
+## generate data
+# population data
+populations <- data.frame(
+	population=LETTERS[1:4],
+	temperature=c(22,29,21.6,34),
+	rainfall=c(6,12,5.5,6)
+)
+# demand point data
+make.dp <- function(x, n, pad.percent=0.2) {
+	padding<-diff(range(x))*pad.percent
+	return(seq(min(x)-padding, max(x)+padding, length.out=n))
+}
+demand.points <- expand.grid(
+	temperature=make.dp(populations$temperature, n=4, pad.percent=0.2),
+	rainfall=make.dp(populations$rainfall, n=4, pad.percent=0.2)
+)
+
+## plot data
+ggplot(aes(x=temperature, y=rainfall), data=populations) +
+	geom_point(data=demand.points, fill='gray70') +
+	geom_text(aes(label=population), size=6) +
+	theme_classic() +
+	xlab(expression('Temperature ('*degree*'C)')) +
+	ylab('Rainfall (mm)') +
+	theme(axis.line.x=element_line(), axis.line.y=element_line())
+```
+
+```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.5, fig.width=4.5, fig.cap='Simulation study.
+Squares denote planning units.
+Each panel shows the distribution for a different simulated species.
+The color of each planning unit indicates the probability that a given species inhabits it.
+Panel (a) shows the uniformly distributed species, panel (b) shows the normally distributed species, and panel (c) shows the bimodally distributed species.'}
+# prepare data
+sim.pus.occ.FPLY <- ldply(
+	seq_len(nrow(sim.species)),
+	function(i) {
+		curr.DF <- sim.pus.FPLY
+		curr.DF$species <- sim.species$name[i]
+		curr.DF$Probability <- sim.dps[[i]]@weights[as.numeric(curr.DF$id)]
+		return(curr.DF)
+	}
+)
+sim.pus.occ.FPLY$species <- gsub('\n', ' ', as.character(sim.pus.occ.FPLY$species))
+sim.pus.occ.FPLY$species <- factor(as.character(sim.pus.occ.FPLY$species), levels=gsub('\n', ' ', sim.species$name)) 
+# make letter data.frame
+sim.pus.occ.letters.DF <- data.frame(
+	x=gCentroid(sim.pus[1,])@coords[1],
+	y=gCentroid(sim.pus[1,])@coords[2],
+	species=unique(sim.pus.occ.FPLY$species),
+	letter=paste0('(', letters[seq_along(sim.species$name)], ')')
+)
+# make plot
+ggplot() +
+	geom_polygon(data=sim.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Probability), color='black') +
+	theme_classic() +
+	geom_text(data=sim.pus.occ.letters.DF, aes(x=x, y=y, label=letter), color='white') +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 5, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(),
+		legend.position=c(1.15, 0.6)) +
+	facet_wrap(~ species, ncol=3)
+```
+
+```{r, echo=FALSE, fig.height=2.5, fig.width=3, fig.cap='Distribution of environmental variation across Queensland, Australia.
+Squares denote planning units.
+The panels show the first and second main gradients of climatic variation.'}
+# prepare data
+cs1.pca.FPTS <- ldply(
+	seq_len(cs1.params.LST[[MODE]]$n.pc.layers),
+	function(i) {
+		curr.DF <- cs1.bioclim.RAST[[i]] %>% 
+			aggregate(10) %>% 
+			rasterToPoints() %>% 
+			as.data.frame()
+		names(curr.DF)[3] <- 'Score'
+		curr.DF$pc <- paste0('Principal component ',i)
+		return(curr.DF)
+	}
+)
+# make plots
+p1 <- ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[1]), aes(x=x, y=y, fill=Score)) +
+	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+	) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ pc)
+p2 <- ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[2]), aes(x=x, y=y, fill=Score)) +
+	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+		) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ pc)
+# render plots
+grid.arrange(p1, p2, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=4.5, fig.width=4.0, fig.cap='Extent of occurence for species used in the first case-study.
+Squares denote planning units.
+Each panel shows the distribution of a different species.
+For a given species, planning units that is inhabits are shown bright blue and those that it does not are shown in dark blue.'}
+# prepare data
+cs1.pus.occ.FPLY <- ldply(
+	seq_along(cs1.params.LST[[MODE]]$common.names),
+	function(i) {
+		curr.DF <- cs1.pus.FPLY
+		curr.DF$species <- cs1.params.LST[[MODE]]$common.names[i]
+		curr.DF$Occupancy <- 'absent'
+		curr.DF$Occupancy[which(curr.DF$id %in% filter(cs1.ru@data@pu.species.probabilities, species==i)$pu)] <- 'present'
+		return(curr.DF)
+	}
+)
+cs1.pus.occ.FPLY$species <- factor(as.character(cs1.pus.occ.FPLY$species), levels=cs1.params.LST[[MODE]]$common.names)
+# make letter data.frame
+cs1.pus.occ.letters.DF <- data.frame(
+	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.05)),
+	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.965)),
+	species=unique(cs1.pus.occ.FPLY$species),
+	letter=paste0('(', letters[seq_along(unique(cs1.pus.occ.FPLY$species))], ')')
+)
+# make plots
+ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black') +
+	geom_text(data=cs1.pus.occ.letters.DF, aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.045, 1.035), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+	) +
+	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ species, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=4.5, fig.width=4.5, fig.cap=paste0('Study area and data used in the second case-study.
+Squares denote planning units.
+Panel (a) shows all grid cells surveyed by the IntraBioDiv project.
+Grid cells occupied by ',cs2.params.LST[[MODE]]$common.name,' are shown in bright blue.
+The subsequent panels contain only the grid cells that were occupied by the species.
+Panel (b) shows the opportunity cost of each planning unit (estimated as the total human population density).
+Panels (b--c) show the spatial distribution of the ordinated genetic data.
+These values describe the typical genetic characteristics of individuals in each planning unit.
+Planning units with similar values/color have contain individuals with similar loci polymorphisms.')}
+# prepare full grid data
+cs2.spp.grid.FPLY <- cs2.grid.FPLY
+cs2.spp.grid.FPLY$Title <- 'Occupancy'
+cs2.spp.grid.FPLY$Occupancy <- 'absent'
+cs2.spp.grid.FPLY$Occupancy[which(cs2.grid.PLY@data[[cs2.params.LST[[MODE]]$species.name]][as.numeric(cs2.spp.grid.FPLY$id)]==1)] <- 'present'
+
+cs2.spp.grid.letters.DF <- data.frame(
+	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.05)),
+	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.95)),
+	letter=('(a)')
+)
+
+# prepare subset grid data
+cs2.spp.grid.sub.FPLY <- cs2.grid.sub.FPLY
+cs2.spp.grid.sub.FPLY$nmds1 <- NA
+cs2.spp.grid.sub.FPLY$nmds2 <- NA
+cs2.nmds1 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d1')]]
+cs2.nmds2 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d2')]]
+
+curr.pos <- match(as.numeric(cs2.spp.grid.sub.FPLY$id), cs2.grid.sub.DF$id)
+cs2.spp.grid.sub.FPLY$nmds1 <- cs2.nmds1[curr.pos]
+cs2.spp.grid.sub.FPLY$nmds2 <- cs2.nmds2[curr.pos]
+
+cs2.spp.grid.sub.FPLY$Cost <- costs.MTX[match(cs2.grid.sub.PLY$id[as.numeric(cs2.spp.grid.sub.FPLY$id)], costs.MTX[,1]),2]
+
+cs2.spp.grid.sub.FPLY$Title1 <- 'Opportunity cost'
+cs2.spp.grid.sub.FPLY$Title2 <- 'Genetic NMDS 1'
+cs2.spp.grid.sub.FPLY$Title3 <- 'Genetic NMDS 2'
+
+cs2.spp.grid.sub.letters.DF <- data.frame(
+	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.05)),
+	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.95)),
+	letter=c('(b)','(c)','(d)')
+)
+
+# plot species data
+p1 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), alpha=0.8, color='grey10') +
+	geom_text(data=cs2.spp.grid.letters.DF, aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.position='none'
+	) +
+	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
+	coord_cartesian(xlim=buffered.range(cs2.grid.FPLY$long, 0), ylim=buffered.range(cs2.grid.FPLY$lat, 0)) +
+	facet_wrap(~ Title)
+# plot cost data
+p2 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Cost), alpha=0.8, color='grey10') +
+	geom_text(data=cs2.spp.grid.sub.letters.DF[1,], aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm'),
+		legend.title=element_blank()
+	) +
+	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
+	scale_fill_continuous(NULL) +
+	facet_wrap(~ Title1)
+# plot nmds axis 1
+p3 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds1), alpha=0.8, color='grey10') +
+	geom_text(data=cs2.spp.grid.sub.letters.DF[2,], aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm')
+	) + 
+	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
+	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred') +
+	facet_wrap(~ Title2)
+# plot nmds axis 2
+p4 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds2), alpha=0.8, color='grey10') +
+	geom_text(data=cs2.spp.grid.sub.letters.DF[3,], aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0),  legend.key.size=unit(3, 'mm')
+	) + 
+	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
+	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred') +
+	facet_wrap(~ Title3)
+# render plots
+grid.arrange(p1, p2, p3, p4, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap=paste0('Simulation study prioritisations.
+Each panel shows a prioritisation generated for a single species using a set of parameters.
+Squares denote planning units.
+Dark green planning units were selected for protection and pale green planning units were discarded.
+Each row of panels corresponds to a different species.
+The top row of panels (a, d, g, j) shows prioritisations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
+Each column of panels correpsonds to a different set of parameters used to generate the prioritisation.
+The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
+# prepare data
+sim.pus.sel.FPLY <- ldply(
+	seq_len(nrow(sim.spp.DF)),
+	function(i) {
+		curr.DF <- sim.pus.FPLY %>% mutate(species=sim.spp.DF$species[i], prioritisation=sim.spp.DF$prioritisation[i])
+		curr.selection <- selections(sim.prioritisations[[sim.spp.DF$species[i]]][[sim.spp.DF$prioritisation[i]]])[as.numeric(curr.DF$id)]
+		curr.DF$Selection <-  c('discarded','selected')[curr.selection+1]
+		return(curr.DF)
+	}
+)
+sim.pus.sel.FPLY$prioritisation <- factor(as.character(sim.pus.sel.FPLY$prioritisation),
+	levels=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'))
+# prepare plotting data
+sim.letter.DF2 <- mutate(sim.spp.DF, letter=paste0('(', letters[seq_along(species)],')'))
+sim.letter.DF2$long <- gCentroid(sim.pus[1,])@coords[1]
+sim.letter.DF2$lat <- gCentroid(sim.pus[1,])@coords[2]
+# make plot
+ggplot() +
+	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
+	geom_text(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', size=3) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
+		axis.text=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(),
+		axis.line=element_blank(), axis.line.x=element_blank(), axis.line.y=element_blank(),
+		plot.background=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		panel.background=element_blank(), panel.border=element_blank(), legend.position='none',
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank()) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
+	facet_grid(species ~ prioritisation)
+```
+
+```{r, echo=FALSE, fig.height=2.5, fig.width=4.5, fig.cap=paste0('Prioritisations for the niche-based case-study in Queensland, Australia.
+Squares represent planning units.
+Dark green planning units for selected for protection and pale green units were discarded.
+Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
+Panel (b) shows the planning units selected selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
+# prepare data
+cs1.pus.sel.FPLY <- ldply(
+	names(cs1.prioritisations),
+	function(x) {
+		curr.DF <- cs1.pus.FPLY
+		curr.DF$Prioritisation <- x
+		curr.DF$Solution <- 'discarded'
+		curr.DF$Solution[curr.DF$id %in% which(as.logical(selections(cs1.prioritisations[[x]])))] <- 'selected'
+		return(curr.DF)
+	}
+)
+cs1.pus.sel.FPLY$Prioritisation <- factor(as.character(cs1.pus.sel.FPLY$Prioritisation), levels=names(cs1.prioritisations))
+# make letter data.frame
+cs1.pus.sel.letters.DF <- data.frame(
+	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.05)),
+	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.965)),
+	Prioritisation=names(cs1.prioritisations),
+	letter=paste0('(', letters[1:2], ')')
+)
+# make plots
+ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black') +
+	geom_text(data=cs1.pus.sel.letters.DF, aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.03, 1.03), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+	) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ Prioritisation, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=5, fig.width=5, fig.cap=paste0('Prioritisations for the genetic case-study in the European Alps.
+Panels (a--b) show prioritisations generated using different parameters.
+Squares denote planning units shown in geographic space.
+Panel (a) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets.
+Panel (b) shows the planning units selected selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs2.params.LST[[MODE]]$genetic.target*100,' % genetic targets.
+Panels (c--d) show the solutions in the genetic space.
+Each point corresponds to a planning unit.
+The coordinates of the points represent the typical genetic characeristics of individuals sampled in planning units (based on an NMDS of the binary loci data).
+Planning units associated with points that are closer togeather contain individuals with more similar genetic characteristics than planning units that are further apart.
+In all panels, dark green planning units were selected for protection, and pale green planning units were discarded.')}
+# prepare spatial data
+cs2.sel.grid.sub.FPLY <- cs2.grid.sub.FPLY
+cs2.sel.grid.sub.FPLY$Title1 <-'Amount target'
+cs2.sel.grid.sub.FPLY$Title2 <- 'Amount & genetic target'
+
+cs2.sel.grid.sub.FPLY$Amount.selection <- 'discarded'
+cs2.sel.grid.sub.FPLY$Amount.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'
+
+cs2.sel.grid.sub.FPLY$Amount.genetic.selection <- 'discarded'
+cs2.sel.grid.sub.FPLY$Amount.genetic.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'
+
+cs2.sel.grid.sub.letters.DF <- data.frame(
+	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.05)),
+	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.95)),
+	letter=c('(a)','(b)')
+)
+
+# prepare genetic space plots
+cs2.sel.space.DF <- data.frame(cs2.genetic.AS@spaces[[1]]@planning.unit.points@coords)
+names(cs2.sel.space.DF) <- c('d1', 'd2')
+cs2.sel.space.DF$id <- cs2.genetic.AS@spaces[[1]]@planning.unit.points@ids
+
+cs2.sel.space.DF$Amount.selection <- 'discarded'
+cs2.sel.space.DF$Amount.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'
+
+cs2.sel.space.DF$Amount.genetic.selection <- 'discarded'
+cs2.sel.space.DF$Amount.genetic.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'
+
+cs2.sel.space.letters.DF <- data.frame(
+	x=min(cs2.sel.space.DF$d1) + (diff(range(cs2.sel.space.DF$d1))*0.05),
+	y=min(cs2.sel.space.DF$d2) + (diff(range(cs2.sel.space.DF$d2))*1),
+	letter=c('(c)', '(d)')
+)
+
+# make plots
+p1 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.selection), color='black') +
+	geom_text(data=cs2.sel.grid.sub.letters.DF[1,], aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		axis.title=element_blank(), legend.position='none'
+	) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
+	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
+	facet_wrap(~ Title1)
+p2 <- ggplot() +
+	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.genetic.selection), color='black') +
+	geom_text(data=cs2.sel.grid.sub.letters.DF[2,], aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
+		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm')
+	) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
+	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
+	facet_wrap(~ Title2)
+p3 <- ggplot() +
+	geom_point(data=cs2.sel.space.DF, aes(x=d1, y=d2, fill=Amount.selection), pch=21, size=3, color='black') +
+	geom_text(data=cs2.sel.space.letters.DF[1,], aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() + xlab('Genetic NMDS Axis 1') + ylab('Genetic NMDS Axis 2') +
+	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line()) +
+	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
+p4 <- ggplot() +
+	geom_point(data=cs2.sel.space.DF, aes(x=d1, y=d2, fill=Amount.genetic.selection), pch=21, size=3, color='black') +
+	geom_text(data=cs2.sel.space.letters.DF[2,], aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() + xlab('Genetic NMDS Axis 1') + ylab('Genetic NMDS Axis 2') +
+	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line()) +
+	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
+# render plots
+grid.arrange(p1,p2,p3,p4,ncol=2)
+
+```

### 2016-06-16 code/rmarkdown/article.Rmd
+ article changes + increase threads

@@
 In recent decades, a wealth of data on biodiversity processes has become available to conservation planners [eg.
-bioclimatic, genetic, and trait data; @r33; @r453; @481].
+bioclimatic, genetic, and trait data; @r33; @r453; @r460].
 Yet this data is only rarely used to guide conservation planning [@r61].
@@
 Here we used of space-based targets to generate a prioritisation to secure a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods].
+We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r454 for further explanation of data collection methods].
 Although this dataset contains multiple plant species, we used data for the `r  cs2.params.LST[[MODE]]$common.name` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}).

### 2016-06-16 code/R/analysis/00-initialization.R

@@ -43,6 +43,8 @@ suppressMessages(library(ALA4R))
 suppressMessages(library(spThin))
 # devtools::install_github('paleo13/rapr')
 suppressMessages(library(rapr))
+# devtools::install_github('paleo13/bayescanr')
+suppressMessages(library(bayescanr))
 # devtools::install_github('paleo13/structurer')
 suppressMessages(library(structurer))
 # devtools::install_github('paleo13/ggplot2')

### 2016-06-16 code/R/analysis/03-case-study-1.R

@@ -83,7 +83,7 @@ names(cs1.spp.RST) <- cs1.params.LST[[MODE]]$common.names
 
 ## create RapUnsolved object
 # create template RapUnsolved with dummy attribute space data
-cs1.ru <- rap(pus=cs1.pus, species=cs1.spp.RST, spaces=list(cs1.bioclim.RAST), kernel.method='hypervolume',amount.target=cs1.params.LST[[MODE]]$amount.target, space.target=cs1.params.LST[[MODE]]$space.target, solve=FALSE, quantile=0.8, n.demand.points=20, Threads=general.params.LST[[MODE]]$threads, MIPGap=general.params.LST[[MODE]]$MIPGap, NumberSolutions=1, include.geographic.space=FALSE)
+cs1.ru <- rap(pus=cs1.pus, species=cs1.spp.RST, spaces=list(cs1.bioclim.RAST), kernel.method='hypervolume',amount.target=cs1.params.LST[[MODE]]$amount.target, space.target=cs1.params.LST[[MODE]]$space.target, solve=FALSE, quantile=0.95, n.demand.points=20, Threads=general.params.LST[[MODE]]$threads, MIPGap=general.params.LST[[MODE]]$MIPGap, NumberSolutions=1, include.geographic.space=FALSE)
 cs1.ru@data@species[[1]] <- cs1.params.LST[[MODE]]$common.names
 # create new attribute spaces
 cs1.spaces <- llply(seq_along(cs1.params.LST[[MODE]]$common.names), .fun=function(i) {

### 2016-06-21 code/R/analysis/03-case-study-1.R
+ fixed issue for third species where outlier environmental conditions are included

@@ -109,23 +109,16 @@ cs1.spaces <- llply(seq_along(cs1.params.LST[[MODE]]$common.names), .fun=functio
 	## extract coordinates in environmental space
 	curr.species.geo.points <- SpatialPoints(coords=randomPoints(cs1.spp.RST[[i]], n=200), proj4string=cs1.spp.RST[[i]]@crs)
 	curr.species.env.points <- extract(cs1.bioclim.RAST,curr.species.geo.points)
-	# omit outlying points in the randomly generated points
+	# put environmental coordinates on same scale as planning units
+	curr.species.env.points<-sweep(curr.species.env.points,MARGIN=2,FUN='-',curr.pu.coords.mean)
+	curr.species.env.points<-sweep(curr.species.env.points,MARGIN=2,FUN='/',curr.pu.coords.sd)
+	# omit outlying coordinates
 	curr.species.env.points.sp <- SpatialPoints(curr.species.env.points)
 	curr.species.env.points.mcp <- mcp(curr.species.env.points.sp, percent=cs1.params.LST[[MODE]]$mcp.percent)
 	curr.species.env.points <- curr.species.env.points[gIntersects(curr.species.env.points.sp,curr.species.env.points.mcp,byid=TRUE)[1,],]
-	# z-score coodinates
-	curr.species.env.points.mean <- apply(curr.species.env.points, 2, mean)
-	curr.species.env.points.sd <- apply(curr.species.env.points, 2, sd)
-	curr.species.env.points<-sweep(curr.species.env.points,MARGIN=2,FUN='-',curr.species.env.points.mean)
-	curr.species.env.points<-sweep(curr.species.env.points,MARGIN=2,FUN='/',curr.species.env.points.sd)
 	# generate demand points
-	raw.curr.species.dps <- make.DemandPoints(curr.species.env.points,  n=cs1.params.LST[[MODE]]$dp.number, quantile=cs1.params.LST[[MODE]]$dp.quantile, kernel.method='hypervolume', bandwidth=cs1.params.LST[[MODE]]$dp.bandwidth)
-	raw.curr.species.dps@coords<-sweep(raw.curr.species.dps@coords,MARGIN=2,FUN='*',curr.species.env.points.sd)
-	raw.curr.species.dps@coords<-sweep(raw.curr.species.dps@coords,MARGIN=2,FUN='+',curr.species.env.points.mean)
[…]

### 2016-06-21 code/R/analysis/03-case-study-1.R
+ bug fix in parallel extract function

@@ -138,7 +138,9 @@ cs1.prioritisations <- llply(
 		'Amount & niche targets'=list(cs1.params.LST[[MODE]]$amount.target,cs1.params.LST[[MODE]]$space.target)
 	),
 	function(y) {
-		update(cs1.ru, amount.target=y[[1]], space.target=y[[2]], solve=TRUE)
+		update(cs1.ru, amount.target=y[[1]], space.target=y[[2]],
+		Threads=general.params.LST[[MODE]]$threads, MIPGap=general.params.LST[[MODE]]$MIPGap,
+		solve=TRUE)
 	}
 )
 names(cs1.prioritisations) <- c('Amount targets','Amount & niche targets')

### 2016-06-21 code/R/functions/misc.R

@@ -126,7 +126,7 @@ parallel_extract <- function(x, y, threads=1, ...) {
 	# get extra args
 	args.LST <- list(...)
 	fids <- seq_along(y@polygons)
-	tids <- rep(seq_len(threads), ceiling(length(fids)/threads))
+	tids <- rep(seq_len(threads), each=ceiling(length(fids)/threads))
 	tids <- tids[seq_along(fids)]
 	fids <- split(fids, factor(tids))
 	## main proc

### 2016-06-21 code/rmarkdown/figures.Rmd
+ changed debug params to run quickly + fixed bug in simulation study's result figure + reordered figures to group them by case-study

@@
 
-```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.5, fig.width=4.5, fig.cap='Simulation study.
+```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.5, fig.width=4.5, fig.cap='Simulation study data set.
 Squares denote planning units.
@@
 	theme_classic() +
-	geom_text(data=sim.pus.occ.letters.DF, aes(x=x, y=y, label=letter), color='white') +
+	geom_text(data=sim.pus.occ.letters.DF, aes(x=x, y=y, label=letter), size=2, color='white') +
 	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 5, 0, 0), "lines"),
@@
 	facet_wrap(~ species, ncol=3)
+```
+
+```{r, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap=paste0('Simulation study prioritisations.
+Each panel shows a prioritisation generated for a single species using a set of parameters.
+Squares denote planning units.
+Dark green planning units were selected for protection and pale green planning units were discarded.
+Each row of panels corresponds to a different species.
+The top row of panels (a, d, g, j) shows prioritisations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
+Each column of panels correpsonds to a different set of parameters used to generate the prioritisation.
+The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
+# prepare data
+sim.pus.sel.FPLY <- ldply(
+	seq_len(nrow(sim.spp.DF)),
+	function(i) {
+		curr.DF <- sim.pus.FPLY %>% mutate(species=sim.spp.DF$species[i], prioritisation=sim.spp.DF$prioritisation[i])
+		curr.selection <- selections(sim.prioritisations[[as.character(sim.spp.DF$species[i])]][[as.character(sim.spp.DF$prioritisation[i])]])[as.numeric(curr.DF$id)]
+		curr.DF$Selection <-  c('discarded','selected')[curr.selection+1]
+		return(curr.DF)
+	}
+)
+sim.pus.sel.FPLY$prioritisation <- factor(as.character(sim.pus.sel.FPLY$prioritisation),
+	levels=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'))
+# prepare plotting data
+sim.letter.DF2 <- mutate(sim.spp.DF, letter=paste0('(', letters[seq_along(species)],')'))
+sim.letter.DF2$long <- gCentroid(sim.pus[1,])@coords[1]
+sim.letter.DF2$lat <- gCentroid(sim.pus[1,])@coords[2]
+# make plot
+ggplot() +
+	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
+	geom_text(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', size=1.8) +
+	theme_classic() +
+	theme(
+		axis.ticks=element_blank(), axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
+		axis.text=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(),
+		axis.line=element_blank(), axis.line.x=element_blank(), axis.line.y=element_blank(),
+		plot.background=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		panel.background=element_blank(), panel.border=element_blank(), legend.position='none',
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank()) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
+	facet_grid(species ~ prioritisation)
 ```
@@
 	facet_wrap(~ species, ncol=2)
+```
+
+```{r, echo=FALSE, fig.height=2.5, fig.width=4.5, fig.cap=paste0('Prioritisations for the niche-based case-study in Queensland, Australia.
+Squares represent planning units.
+Dark green planning units for selected for protection and pale green units were discarded.
+Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
+Panel (b) shows the planning units selected selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
+# prepare data
+cs1.pus.sel.FPLY <- ldply(
+	names(cs1.prioritisations),
+	function(x) {
+		curr.DF <- cs1.pus.FPLY
+		curr.DF$Prioritisation <- x
+		curr.DF$Solution <- 'discarded'
+		curr.DF$Solution[curr.DF$id %in% which(as.logical(selections(cs1.prioritisations[[x]])))] <- 'selected'
+		return(curr.DF)
+	}
+)
+cs1.pus.sel.FPLY$Prioritisation <- factor(as.character(cs1.pus.sel.FPLY$Prioritisation), levels=names(cs1.prioritisations))
+# make letter data.frame
+cs1.pus.sel.letters.DF <- data.frame(
+	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.05)),
+	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.965)),
+	Prioritisation=names(cs1.prioritisations),
+	letter=paste0('(', letters[1:2], ')')
+)
+# make plots
+ggplot() +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black') +
+	geom_text(data=cs1.pus.sel.letters.DF, aes(x=x, y=y, label=letter), color='black') +
+	theme_classic() +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white'),
+		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
+		legend.position=c(1.03, 1.03), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+	) +
+	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
+	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	facet_wrap(~ Prioritisation, ncol=2)
 ```
@@
 
-```{r, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap=paste0('Simulation study prioritisations.
-Each panel shows a prioritisation generated for a single species using a set of parameters.
-Squares denote planning units.
-Dark green planning units were selected for protection and pale green planning units were discarded.
-Each row of panels corresponds to a different species.
-The top row of panels (a, d, g, j) shows prioritisations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
-Each column of panels correpsonds to a different set of parameters used to generate the prioritisation.
-The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
-# prepare data
-sim.pus.sel.FPLY <- ldply(
-	seq_len(nrow(sim.spp.DF)),
-	function(i) {
-		curr.DF <- sim.pus.FPLY %>% mutate(species=sim.spp.DF$species[i], prioritisation=sim.spp.DF$prioritisation[i])
-		curr.selection <- selections(sim.prioritisations[[sim.spp.DF$species[i]]][[sim.spp.DF$prioritisation[i]]])[as.numeric(curr.DF$id)]
-		curr.DF$Selection <-  c('discarded','selected')[curr.selection+1]
-		return(curr.DF)
-	}
-)
-sim.pus.sel.FPLY$prioritisation <- factor(as.character(sim.pus.sel.FPLY$prioritisation),
-	levels=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'))
-# prepare plotting data
-sim.letter.DF2 <- mutate(sim.spp.DF, letter=paste0('(', letters[seq_along(species)],')'))
-sim.letter.DF2$long <- gCentroid(sim.pus[1,])@coords[1]
-sim.letter.DF2$lat <- gCentroid(sim.pus[1,])@coords[2]
-# make plot
-ggplot() +
-	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
-	geom_text(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', size=3) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
-		axis.text=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(),
-		axis.line=element_blank(), axis.line.x=element_blank(), axis.line.y=element_blank(),
-		plot.background=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		panel.background=element_blank(), panel.border=element_blank(), legend.position='none',
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank()) +
-	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
-	facet_grid(species ~ prioritisation)
-```
-
-```{r, echo=FALSE, fig.height=2.5, fig.width=4.5, fig.cap=paste0('Prioritisations for the niche-based case-study in Queensland, Australia.
-Squares represent planning units.
-Dark green planning units for selected for protection and pale green units were discarded.
-Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
-Panel (b) shows the planning units selected selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
-# prepare data
-cs1.pus.sel.FPLY <- ldply(
-	names(cs1.prioritisations),
-	function(x) {
-		curr.DF <- cs1.pus.FPLY
-		curr.DF$Prioritisation <- x
-		curr.DF$Solution <- 'discarded'
-		curr.DF$Solution[curr.DF$id %in% which(as.logical(selections(cs1.prioritisations[[x]])))] <- 'selected'
-		return(curr.DF)
-	}
-)
-cs1.pus.sel.FPLY$Prioritisation <- factor(as.character(cs1.pus.sel.FPLY$Prioritisation), levels=names(cs1.prioritisations))
-# make letter data.frame
-cs1.pus.sel.letters.DF <- data.frame(
-	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.05)),
-	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.965)),
-	Prioritisation=names(cs1.prioritisations),
-	letter=paste0('(', letters[1:2], ')')
-)
-# make plots
-ggplot() +
-	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black') +
-	geom_text(data=cs1.pus.sel.letters.DF, aes(x=x, y=y, label=letter), color='black') +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
-		legend.position=c(1.03, 1.03), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
-	) +
-	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
-	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
-	facet_wrap(~ Prioritisation, ncol=2)
-```
-
 ```{r, echo=FALSE, fig.height=5, fig.width=5, fig.cap=paste0('Prioritisations for the genetic case-study in the European Alps.

### 2016-06-23 code/rmarkdown/article.Rmd
+ major article edits to results and discussion + add in cs1 figs to si + fix bug in make clean

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: ["conservation", "biodiversity", "MILP", "reserve-selection"]
-abstract: "A central aim in conservation is to maximise the long-term persistence of biodiversity.
-To fulfil this aim, reserve networks are used to safeguard biodiversity patterns (eg.
+tags: ["conservation", "biodiversity", "MILP", "reserve-selection", "optimisation", "protected area"]
+abstract: "
+
+1.
+A central aim in conservation is to maximise the long-term persistence of biodiversity.
+To fulfill this aim, reserve networks are used to safeguard biodiversity patterns (eg.
 species, populations) and processes (eg.
-evolutionary processes that underpin genetic variation).
-Reserve selection is often formulated as an optimisation problem to identify cost-effective prioritisations.
-However, most existing decision support tools are based on formulations that are well suited for preserving biodiversity patterns, but not biodiversity processes.
-To fill this gap in the conservation planning toolbox, we developed the \texttt{rapr R} package.
-This \texttt{R} package provides a toolkit to guide reserve selection using novel formulations of this problem.
-Here, we explore the functionality of this R package using a simulation study and two case-studies.
-We demonstrate how explicitly considering biodiversity processes can alter a prioritisation.
-In most cases, we found that only a few additional plannning units are required to sufficiently preserve of biodiversity processes."
+evolutionary processes).
+Reserve selection is often formulated as an optimisation problem to identify cost-effective solutions.
+These formulations often use \"amount targets\" to specify the minimum amount habitat that must be preserved for each species.
+Although such targets are well suited for securing biodiversity patterns, they are not well suited for obtaining a representative sample of the biodiversity processes that play out across species' ranges.
+Currently, decision makers lack the tools to develop prioritisations that preserve both biodiversity patterns and processes.
+
+2.
+To address this issue, we developed the \\texttt{rapr R} package.
+This \\texttt{R} package provides a toolkit to guide reserve selection using explicit targets for biodiversity processes.
+Using a novel formulation of the reserve selection problem, users can set \"space targets\" to secure a representative sample of each species within an attribute space to preserve biodiversity processes (eg.
+an attribute space based on genetic data to preserve evolutionary processes).
+We explored the functionality of this \\texttt{R} package using a simulation study and two case studies.
+In each study, we generated prioritisations using amount targets--to represent conventional reserve selection methods--and compared them to prioritisations generated using both amount and space targets.
+
+3.
+We demonstrated how explicitly considering biodiversity processes in reserve selection can affect the resulting solution.
+The results from the simulation study suggest that most species could benefit from the explicit use of space-based targets--especially species with multi-model distributions in an attribute space.
+The results from both first case studies demonstrated that explicitly considering space-based targets for biodiversity processes can result in more effective prioritisations.
+
+4.
+The \\texttt{rapr R} package provides a unified framework for biodiversity patterns and species-specific processes in conservation planning.
+Overall, we found strong evidence that prioritisations developed  using amount targets may fail to secure  that can provide better outcomes than just considering biodiversity patterns.
+"
 output:
@@
 
-## Introduction
+# Introduction
 <!-- conservation and reserve networks -->
@@
 <!-- biodiversity processes are important -->
-To fulfil the overarching aims of conservation, reserve networks must preserve both ecological and evolutionary processes [@r3; @r255].
+To fulfill the overarching aims of conservation, reserve networks must preserve both ecological and evolutionary processes [@r3; @r255].
 Ecological processes, such as predator-prey interactions, pollination, and decomposition, are required for biodiversity to persist over short time-scales.
@@
 This \texttt{R} package uses novel formulations of the reserve selection problem to provide decision makers with the tools to generate prioritisations that preserve biodiversity patterns and processes.
-We also explore the functionality of this \texttt{R} package using simulated species and a case-study conservation planning scenario.
+We aim to define the concepts behind the problem formulations.
+Furthmore, we aim to explore the functionality of the \texttt{R} package by applying one of the formulations to a set of simulated species and two case studies.
 
-## Problem formulations
-The \texttt{rapr R} package uses two novel formulations of the reserve selection problem to identify cost-effective prioritisations.
-These formulations share many constraints and variables.
-For brevity, the variables used by both formulations will be defined.
+# Methods
+## PROBLEM FORMULATION
 Biodiversity features are defined as the entity(s) that the prioritisation is required to preserve (eg.
@@
 The decision maker then used this environmental data to construct a two-dimensional environmental attribute space.
-Next, the decision maker generated demand points as equi-distant points between the range of values where the populations were found [$\pm$ 20% to avoid edge effects; @r218].
-By comparing the distribution of the demand points to the distribution of the populations in the attribute space, the decision maker can identify a suitable prioritisation (Figure 1).
+Next, the decision maker generated demand points as equi-distant points between the range of values where the populations were found.
+By comparing the distribution of the demand points to the distribution of the populations in the attribute space, the decision maker can identify a suitable prioritisation (Fig.
+1).
 We can see that if the decision maker preserves both populations $A$ and $C$, they will effectively "double-up" on the same environmental characteristics, and in turn their waste resources.
@@
 
-The problem formulations used to guide reserve selection in the \texttt{rapr R} package are based on a combination of the \texttt{Marxan} reserve selection problem and the uncapacitated facility location problems [@r451].
-All mathematical terms defined hereafter are described in Table S1 for convenience.
+The formulations used to express the reserve selection problem in the \texttt{rapr R} package are based on a combination of the \texttt{Marxan} reserve selection problem and the uncapacitated facility location problems [@r451].
+Although the \texttt{rapr R} package provides two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in the Supporting Information.
+The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied when calculating the proportion of an attribute space sampled in a solution.
+All mathematical terms defined hereafter are described in Table S1.
 For convenience, the cardinality of sets will be denoted using the same symbol used to denote the variable.
+
 Define $F$ to be the set of features (indexed by $f$).
@@
 
-The \texttt{R} package provides two formulations for reserve selection.
-They are based on the unreliable [@r451] and reliable uncapacitated facility location [@r16] problems.
-The key difference between them is that the reliable formulation explicitly considers the probability that the planning units are occupied when calculating the proportion of variation that a given solution secures, whereas the unreliable formulation does not.
-
-### Unreliable formulation
-In the unreliable formulation, the control variables are the $B$, $T_{s}$, and $\tau_{sa}$ variables.
+The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{sa}$ variables.
 
@@
 This total weighted distance is then scaled by the inherent variation in the demand points ($\lambda_{fsi} \delta_{fsi}$).
-The resulting fraction yields a proportion conceptually similar to the statistical $R^2$ value, and the constraints ensure that this proportion must be greater than or equal to the space-based target.
+The resulting fraction yields a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
+The constraints ensure that this proportion must be equal to or greater than the space-based target.
 Constraints (3d) ensure that only one planning unit is assigned to each demand point.
@@
 
-### Reliable formulation
-The reliable formulation explicitly considers the probability that the planning units are inhabited.
-As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented.
-This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s).
-To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.
-
-Backup levels levels are defined as $r$-levels [similar to failure levels in @r18].
-The first backup $r$-level is used to calculate the level of representation when all of the selected planning units are occupied by all $f \in F$.
-For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level$=0$.
-This scenario essentially represents $Y_{fsij}$ in the unreliable formulation.
-The second backup $r$-level is used to assess the level of representation when the closest planning unit to each demand point $i$ is unoccupied.
-For this scenario, the second closest planning units are assigned at $r$-level$=1$.
-The third backup $r$-level is used to assess representation when the first two closest planning units are unoccupied.
-The third closest planning units are assigned at $r$-level$=2$.
-Continuing on, in this manner, the selected planning units in a prioritisation are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.
-
-A final backup $r$-level when $r=R$ is used to assess the level of representation when the features $f \in F$ do not occupy any selected planning units in a prioritisation.
-Each demand point $i \in I$ for each $s \in S$ and $f \in F$ is assigned to an "imaginary" planning unit $j=J$ at $r=R$.
-The distance variables associated with this imaginary planning unit $d_{fsiJ}$ denote the loss of biological value associated with failing to secure a representative sample of feature $f$ in attribute space $s$.
-However, the $d$ variables are in distance units which are meaningless in this context.
-Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).
-
-\begin{align*}
-& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag{4} \\
-& & & 0 \leq s \leq S-1\\
-\end{align*}
-
-Moderately-sized conservation planning problems often include several thousand planning units.
-It is currently not be feasible to solve this problem when considering all possible failure scenarios.
-As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$.
-For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level.
-Cui et al.
-[-@r16] found that $R=5$ yields similar solutions to $R=J$ when $J >> 5$.
-However, in most costs the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.
-
-In the reliable formulation, the control variables are the $B$ (1a), $T_{s}$ (1b), $\tau_{sa}$ (1c), $R$, and $M$ variables.
-
-\begin{align*}
-R &= \parbox{25em}{number of failure levels} \tag{5a} \\
-%
-M &= \parbox{25em}{failure multiplier} \tag{5b}\\
-%
-\end{align*}
-
-The decision variables are the $X_j$ (2a), $Y_{fsijr}$, $P_{fsijr}$ variables.
-
-\begin{align*}
-Y_{fsijr} &= \begin{cases}
-		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag{6a} \\
-		0, & \parbox{25em}{otherwise} \\
-	\end{cases} \\
-%
-P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag{6a}\\
-%
-\end{align*}
-
-The reliable formulation (RRAP) is a multi-objective optimisation problem.
-
-\begin{align*}
-& \text{(RRAP)} & \text{Min } & \text{(3a)}\\
-%
-& & \text{s.t. } & \text{(3b)}\\
-%
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} P_{fsijr} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag{7a}\\
-& & & & & 0 \leq s \leq S-1\\
-%
-& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7b}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq r \leq R\\
-%
-& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7c}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J\\
-%
-& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag{7d}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J-1\\
-%
-& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag{7e}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1\\
-%
-& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag{7f}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J\\
-%
-& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag{7g}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J,\\
-& & & & & 1 \leq r \leq R\\
-%
-& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag{7h}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J,\\
-& & & & & 0 \leq r \leq R\\
-\end{align*}
-
-The objective function for the reliable formulation is the same as for the unreliable formation (3a).
-Similar to the unreliable formulation, constraints (3b) and (7a) ensure that the amount-based and space-based targets are met.
-Constraint (7b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$.
-Constraints (7d) ensure that only selected planning units are assigned to demand points $i \in I$.
-Constraints (7e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level.
-Constraints (7f--7g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information].
-Constraints (7h) ensure that the $X$ and $Y$ variables are binary.
-
-### Optimisation
-The unreliable and reliable formulations are non-linear.
+## OPTIMISATION
+The unreliable formulation is non-linear.
 However, the non-linear components can be linearised using existing techniques.
-First, the expression $X_j X_k$ in (3a) can be linearised using methods described by Beyer et al.
+The expression $X_j X_k$ in (3a) can be linearised using methods described by Beyer et al.
 [-@r426].
-Second, the expression $P_{fsijr} Y_{jsijr}$ in (7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
-[-@r16].
 Linearised versions of the problems can be solved using commercial exact algorithm solvers.
-
-The \texttt{rapr} R package provides functions to express conservation planning data as an optimisation problems using linearised versions of the unreliable and reliable formulations.
+The \texttt{rapr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations.
 These optimisation problems can then be solved to generate prioritisations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).
@@
 After installing the \texttt{Gurobi} software suite, users will need to install the \texttt{Gurobi R} package.
-This \texttt{R} package can be installed on recent Windows, Mac OSX, and Ubuntu Linux operating systems.
 
-To understand the beaviour of this problem and showcase its utility, we conducted a simulation study and two case-studies.
-These studies involved generating prioritizations using only amount-based targets to represent prioritisations generated using conventional methods (eg.
-Marxan), and prioritizations using amount-based and space-based targets using the unreliable formulation.
+# Examples
 
-## Simulation study
+To understand the behaviour of the unreliable problem and showcase its value, we conducted a simulation study and two case studies.
+These studies involved generating solutions using only amount targets to represent prioritisations generated using conventional methods (eg.
+\texttt{Marxan}), and solutions using amount and space targets using the unreliable formulation.
+By comparing these solutions, we can guage the benefits of explicitly including space targets in reserve selection.
+All analyses were performed in \texttt{R} [version `r getRversion()`; @r192].
+
+## SIMULATION STUDY
 ### Methods
-We generated a study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Figure 2).
-We then simulated three species.
-The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Figure 2a).
+We simulated a hypothetical study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig.
+2).
+We then simulated three species across this study area.
+The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig.
+2a).
 This species' distribution was based on a uniform distribution (eqn 8a).
-The second species was simulated to represent a species with an idealised distribution (Figure 2b).
+The second species was simulated to represent a species with an idealised distribution (Fig.
+2b).
 It has a core range area, and is less likely to be found in areas that are distant from the core area.
 This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8b).
-The third species represents a species with two distinct populations (Figure 2c).
+The third species represents a species with two distinct populations (Fig.
+2c).
 This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8c).
@@
 
-We generated four prioritisations for each species.
-First to represent prioritisations generated using conventional conservation planning methods, we generated prioritisations using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-Second to show how the addition of geographic targets can affect a prioritisation, we  generated prioritisations using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets.
-Third to represent prioritisations using conventional planning methods that penalise for fragmentation, we generated prioritisations using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
-Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated prioritisations using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
+We generated four solutions for each species.
+First to represent solutions generated using conventional conservation planning methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
+Second to show how the addition of geographic targets can affect a prioritisation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets.
+Third to represent solutions using conventional planning methods that penalise for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
+Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated solutions using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
 
@@
 
-## Case-study 1
+The uniform species was simulated with a constant probability of occupancy across the study area (Fig.
+2a).
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount-based targets (Fig.
+3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area.
+This configuration is an artifact of the method used to solve this particular instance of the reserve selection problem.
+In fact, for this species, all prioritisations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
+In the absence of criteria to guarantee representativeness, reserve selection methods may or may not return solutions that secure a representative sample of the features.
+In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
+
+The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
+3b).
+Since all planning units have cost the same amount and have an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf.
+Fig 3a).
+Although the use of amount and space based targets has addressed adequacy and representativeness objectives (respectively), they have resulted in a solution that is a highly fragmented solution.
+
+The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig.
+3d).
+This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented.
+This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig.
+3c) contained the sample number of planning units as the solution with just the amount targets (Fig.
+3a).
+Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.
+
+The normally distributed species was simulated to contain a single core area where individuals are most prevalent and marginal areas where individuals are less likely to occur (Fig.
+2b).
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig.
+3e).
+Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy.
+Whilst this solution satisfies the adequacy objective for a prioritisation in a cost-effective manner--it fails to fulfill the representative objective for a prioritisation (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).
+
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species distribution (Fig.
+3f).
+This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution.
+By using amount and geographic targets in the reserve selection problem, we have obtained a solution that concentrates conservation effort in the range core and also the range margin.
+However, similar to the corresponding solution for the uniform species (Fig.
+3b), this solution is highly fragmented.
+
+By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig.
+3h).
+This solution contains `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units.
+Similar to the corresponding solution for the uniform species (Fig.
+3d), this solution contains more planning units than any other solutions for this species (cf.
+Figs.
+3e--3g).
+The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species'--even for species without significant structure.
+
+The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two isolated populations (Fig.
+2c).
+The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets assigned conservation effort to just one of the two populations (Fig.
+3i).
+While this solution secured an adequate proportion of the species' distribution (`r  round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
+The addition of boundary length modifiers to just amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig.
+3k).
+However, the addition of geographic space targets resolved these issues.
+
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets preserved individuals in both populations (Fig.
+3j).
+Although this solution required more planning units to fulfill both targets ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$n`$), this solution secured a representative sample of the bimodally distributed species (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % range sampled).
+Similar to corresponding solutions generated for both the uniformly (Fig.
+3b) and normally distributed species (Fig.
+3f), this solution was also highly fragmented and we could obtain a more well-connected solution at the expense of selecting additional planning units ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n`$; Fig.
+3l).
+The results for the bimodally distributed species suggest that species with highly structured populations or strong variation between individuals could benefit the most from prioritisations that are generated using space-based targets.
+
+## CASE STUDY 1
 ### Methods
@@
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subject them to a principal components analysis (using ArcMap 10.3.1).
-We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principle components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Figure 3).
+We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principle components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Fig.
+4).
 
@@
 We used these bird species because they have different evolutionary histories and life-history strategies.
-We then mapped the extent of occurrence for each species using the following method: we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+We then mapped the extent of occurrence for each species (Fig.
+5).
+To do this, we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
 We used this method because it is reproducible using freely available data and moderately accurate.
 
-We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Figure S2).
+We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Fig.
+S2).
 To achieve this, for each species, we generated random points inside the species range and at each point extracted the principal component values at that location.
@@
 
-We generated two prioritisations.
-The first prioritisation was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species.
-The second prioritisation was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based for each species.
+We generated two solutions.
+The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species.
+The second solution was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based for each species.
 
 ### Results
+Generally, the solution generated using amount targets preserved a representative sample of each the `r length(cs1.params.LST[[MODE]]$species.names)` bird species' niches--with exception to `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'a single')` bird species.
+Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realised niche of `r parsed.representative.space.held.names`.
+Yet it failed to achieve this for `r parsed.not.representative.space.held.names`.
 
-## Case-study 2
+## CASE STUDY 2
 ### Methods
 Here we used of space-based targets to generate a prioritisation to secure a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r454 for further explanation of data collection methods].
-Although this dataset contains multiple plant species, we used data for the `r  cs2.params.LST[[MODE]]$common.name` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}).
+We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@462; see @r461 and @r463 for further explanation of data collection methods].
+Althougthis dataset contains multiple plant species, we used data for the `r  cs2.params.LST[[MODE]]$common.name` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}).
 This species was chosen because it provides is locally wide-spread and has been shown to exhibit significant genetic structure [@r454].
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
-22.3 km $\times$ 25 km; Figure 5a).
+22.3 km $\times$ 25 km; Fig.
+7a).
 They visited every second grid cell, and if the `r cs2.params.LST[[MODE]]$common.name` was detected in a cell, samples were collected from three individuals.
@@
 We calculated the average of the values associated with individuals in each grid cell.
-These values were used to create a genetic attribute space (Figures 5c and 5d).
+These values were used to create a genetic attribute space (Figs.
+7c and 7d).
+To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances based on the grid cells' centroids [using the \texttt{ape R} package; @r464].
 
@@
 Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points.
-To ensure that the solutions did not prioritise particularly costly areas that could not be used for conservation action, we included opportunity cost data (Figure 5b).
+To ensure that the solutions did not prioritise particularly costly areas that could not be used for conservation action, we included opportunity cost data (Fig.
+7b).
 We obtained population density data [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
 
-We generated two prioritisations.
-The first prioritisation was generated using an `r cs2.params.LST[[MODE]]$amount.target*100` % amount-based target.
-The second prioritisation was generated using the same amount-based target and an additional `r cs2.params.LST[[MODE]]$genetic.target*100` % genetic-based target.
+We generated two solutions.
+The first solution was generated using an `r cs2.params.LST[[MODE]]$amount.target*100` % amount-based target.
+The second solution was generated using the same amount-based target and an additional `r cs2.params.LST[[MODE]]$genetic.target*100` % genetic-based target.
 
@@
 
-## Implications and future directions
+The binary genetic data was ordinated into a two-dimensional space that described the main different between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`).
+These values were then averaged to the planning unit level to describe the typical genetic characteristics of individuals in each planning unit (Figs.
+7c--7d).
+Planning units located near each were found to contain individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
+In terms of the typical genetic characteristics of the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
+This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.
+
+The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs.
+8a and 8c).
+This solution only sampled planning units that contained individuals belonging to one of the main two groups.
+On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups.
+Note that since these solutions were generated using opportunity cost data, the solution generated using only the amount target is essentially the minimum number of least costly areas needed to fulfill the amount target.
+The only difference between the two solutions is a single planning unit.
+By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation `r  round(cs2.spp.DF$genetic[1],2)` for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).
+
+# Implications and future directions
 The \texttt{rapr} R package provides a unified approach to reserve selection.
@@
 Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
-Both the simulated and case-study species suggest that conservation planning exercises need to explicitly consider biodiversity processes during the reserve selection process to capture them.
+Both the simulated and case study species suggest that conservation planning exercises need to explicitly consider biodiversity processes during the reserve selection process to capture them.
 
@@
 
-## Acknowledgements
+# Acknowledgements
 JOH is funded by an Australian Postgraduate Award (APA) scholarship.
@@
 
-## References
+# Data accessibility
+All data, code, and results are stored in an online repository (\url{www.github.com/paleo13/rapr-manuscript}) to replicate and valiate this study.
+
+# References

### 2016-06-23 code/rmarkdown/supporting-information.Rmd
+ major article edits to results and discussion + add in cs1 figs to si + fix bug in make clean

@@
 ---
-## Supporting Information
-All data and code is stored in an online repository (\url{www.github.com/paleo13/rapr-manuscript}) to replicate analyses in this study. See the \textit{README} file for instructions.
+# Supporting Information
 
@@
 ```
+
+## Reliable formulation
+The reliable formulation explicitly considers the probability that the planning units are inhabited.
+As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented.
+This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s).
+To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.
+
+Backup levels levels are defined as $r$-levels [similar to failure levels in @r18].
+The first backup $r$-level is used to calculate the level of representation when all of the selected planning units are occupied by all $f \in F$.
+For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level$=0$.
+This scenario essentially represents $Y_{fsij}$ in the unreliable formulation.
+The second backup $r$-level is used to assess the level of representation when the closest planning unit to each demand point $i$ is unoccupied.
+For this scenario, the second closest planning units are assigned at $r$-level$=1$.
+The third backup $r$-level is used to assess representation when the first two closest planning units are unoccupied.
+The third closest planning units are assigned at $r$-level$=2$.
+Continuing on, in this manner, the selected planning units in a prioritisation are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.
+
+A final backup $r$-level when $r=R$ is used to assess the level of representation when the features $f \in F$ do not occupy any selected planning units in a prioritisation.
+Each demand point $i \in I$ for each $s \in S$ and $f \in F$ is assigned to an "imaginary" planning unit $j=J$ at $r=R$.
+The distance variables associated with this imaginary planning unit $d_{fsiJ}$ denote the loss of biological value associated with failing to secure a representative sample of feature $f$ in attribute space $s$.
+However, the $d$ variables are in distance units which are meaningless in this context.
+Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).
+
+\begin{align*}
+& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag{4} \\
+& & & 0 \leq s \leq S-1\\
+\end{align*}
+
+Moderately-sized conservation planning problems often include several thousand planning units.
+It is currently not be feasible to solve this problem when considering all possible failure scenarios.
+As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$.
+For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level.
+Cui et al.
+[-@r16] found that $R=5$ yields similar solutions to $R=J$ when $J >> 5$.
+However, in most costs the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.
+
+The control variables for the reliable formulation are the $B$ (1a), $T_{s}$ (1b), $\tau_{sa}$ (1c), $R$, and $M$ variables.
+
+\begin{align*}
+R &= \parbox{25em}{number of failure levels} \tag{5a} \\
+%
+M &= \parbox{25em}{failure multiplier} \tag{5b}\\
+%
+\end{align*}
+
+The decision variables are the $X_j$ (2a), $Y_{fsijr}$, $P_{fsijr}$ variables.
+
+\begin{align*}
+Y_{fsijr} &= \begin{cases}
+		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag{6a} \\
+		0, & \parbox{25em}{otherwise} \\
+	\end{cases} \\
+%
+P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag{6a}\\
+%
+\end{align*}
+
+The reliable formulation (RRAP) is a multi-objective optimisation problem.
+
+\begin{align*}
+& \text{(RRAP)} & \text{Min } & \text{(3a)}\\
+%
+& & \text{s.t. } & \text{(3b)}\\
+%
+& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} P_{fsijr} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag{7a}\\
+& & & & & 0 \leq s \leq S-1\\
+%
+& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7b}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1,\\
+& & & & & 0 \leq r \leq R\\
+%
+& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7c}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1,\\
+& & & & & 0 \leq j \leq J\\
+%
+& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag{7d}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1,\\
+& & & & & 0 \leq j \leq J-1\\
+%
+& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag{7e}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1\\
+%
+& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag{7f}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1,\\
+& & & & & 0 \leq j \leq J\\
+%
+& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag{7g}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1,\\
+& & & & & 0 \leq j \leq J,\\
+& & & & & 1 \leq r \leq R\\
+%
+& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag{7h}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1,\\
+& & & & & 0 \leq j \leq J,\\
+& & & & & 0 \leq r \leq R\\
+\end{align*}
+
+The objective function for the reliable formulation is the same as for the unreliable formation (3a).
+Similar to the unreliable formulation, constraints (3b) and (7a) ensure that the amount-based and space-based targets are met.
+Constraint (7b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$.
+Constraints (7d) ensure that only selected planning units are assigned to demand points $i \in I$.
+Constraints (7e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level.
+Constraints (7f--7g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information].
+Constraints (7h) ensure that the $X$ and $Y$ variables are binary.
+
+The reliable formulation is non-linear.
+However, the non-linear components can be linearised.
+First--as discussed in the main text--the expression $X_j X_k$ in (3a) can be linearised using methods described by Beyer et al.
+[-@r426].
+Second, the expression $P_{fsijr} Y_{jsijr}$ in (7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
+[-@r16].
+
+## Figures 
+```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study.
+Each panel shows a the distribution of a solution in environmental space and how it samples the realised niche for a different species.
+The left column of panels shows the solution generated using amount targets.
+The right column of panels shows the solution generated using amount and niche targets.
+Each column of panels corresponds to a different species.
+Hexagons show the distribution of demand points.
+The color of each heaxgon denotes the weighted frequency of demand points inside it.
+Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution.'}
+## prepare data
+cs1.pu.niche.DF <- ddply(
+	cs1.spp.DF,
+	c('Species', 'Prioritisation'),
+	function(x) {
+		# init
+		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
+		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
+		curr.sel.pos <- which(curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@ids %in% which(as.logical(selections(curr.prioritisation))))
+		# extract planning units
+		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@coords %>% as.data.frame()
+		names(curr.DF) <- c('Niche1', 'Niche2')
+		curr.DF$Species <-x$Species[[1]]
+		curr.DF$Prioritisation <-x$Prioritisation[[1]]
+		curr.DF$Solution <- 'discarded'
+		curr.DF$Solution[curr.sel.pos] <- 'selected'
+		# return data
+		return(curr.DF)
+	}
+)
+cs1.pu.niche.DF <- filter(cs1.pu.niche.DF, Solution=='selected')
+
+cs1.dp.niche.DF <- ddply(
+	cs1.spp.DF,
+	c('Species', 'Prioritisation'),
+	function(x) {
+		# init
+		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
+		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
+		# extract demand points
+		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@coords %>% as.data.frame()
+		names(curr.DF) <- c('Niche1', 'Niche2')
+		curr.DF$weight <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@weights
+		# return data
+		return(curr.DF)
+	}
+)
+cs1.pu.niche.DF$Prioritisation <- factor(as.character(cs1.pu.niche.DF$Prioritisation), levels=names(cs1.prioritisations))
+cs1.dp.niche.DF$Prioritisation <- factor(as.character(cs1.dp.niche.DF$Prioritisation), levels=names(cs1.prioritisations))
+
+# make letter data.frame
+cs1.letters.DF <- expand.grid(
+	Species=levels(cs1.pu.niche.DF$Species),
+	Prioritisation=levels(cs1.pu.niche.DF$Prioritisation)
+) %>% arrange(Species, Prioritisation)
+cs1.letters.DF$letter <- paste0('(', letters[seq_len(nrow(cs1.letters.DF))], ')')
+cs1.letters.DF <- cbind(
+	cs1.letters.DF,
+	rbind(
+		cs1.pu.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2),
+		cs1.dp.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2)
+	) %>% summarise(
+		Niche1=min(Niche1)+(abs(diff(range(Niche1)))*0.01),
+		Niche2=min(Niche2)+(abs(diff(range(Niche2)))*0.95)
+	) %>% data.frame()
+)
+
+## make plots
+ggplot() +
+	stat_summary_hex(data=cs1.dp.niche.DF, aes(x=Niche1, y=Niche2, z=weight), bins=15, fun=function(z) {sum(z)}) +
+	geom_point(data=cs1.pu.niche.DF, aes(x=Niche1, y=Niche2), fill='#00441b', pch=21, color='black') +
+	geom_text(data=cs1.letters.DF, aes(x=Niche1, y=Niche2, label=letter), color='black') +
+	theme_classic() +
+	theme(
+		strip.background = element_rect(fill='grey20'),
+		strip.text = element_text(color='white')
+	) +
+	xlab('Niche (axis 1)') + ylab('Niche (axis 2)') +
+	scale_fill_gradientn(name='Total demand\npoint weight', colors=colorRampPalette(brewer.pal(6, 'PuBu')[-1])(100)) +
+	facet_grid(Species ~ Prioritisation)
+```
+
+## Tables
+
+## References

### 2016-06-23 code/R/analysis/00-initialization.R

@@ -12,6 +12,10 @@ if (length(args)>0) {
 
 # load CRAN packages
 suppressMessages(library(stats))
+suppressMessages(library(RColorBrewer))
+suppressMessages(library(hexbin))
+suppressMessages(library(fields))
+suppressMessages(library(ape))
 suppressMessages(library(Hmisc))
 suppressMessages(library(data.table))
 suppressMessages(library(grid))

### 2016-06-23 code/R/analysis/02-simulations.R

@@ -109,7 +109,7 @@ sim.prioritisations <- llply(
 			'Amount & space\ntargets and BLM'=c(sim.params.LST[[MODE]]$space.target, sim.params.LST[[MODE]]$blm)
 		),
 			function(j) {
-				sim.ru %>% spp.subset(x) %>% update(space.target=j[[1]], blm=j[[2]], solve=TRUE,
+				sim.ru %>% spp.subset(x) %>% update(space.target=j[[1]], BLM=j[[2]], solve=TRUE,
 				threads=general.params.LST[[MODE]]$threads, MIPGap=general.params.LST[[MODE]]$MIPGap)
 			}
 		)
@@ -124,8 +124,10 @@ sim.spp.DF <- ldply(
 		data.frame(
 			species=rep(as.character(sim.species[[1]][[i]]), 4),
 			prioritisation=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'),
-			amount.held=sapply(sim.prioritisations[[i]], function(s) {amount.held(s)[1]}),
-				space.held=sapply(sim.prioritisations[[i]], function(s) {space.held(s)[1]})
+			amount.held=sapply(sim.prioritisations[[i]], function(s) {amount.held(s)[1]})*100,
+			space.held=sapply(sim.prioritisations[[i]], function(s) {space.held(s)[1]})*100,
+			score=sapply(sim.prioritisations[[i]], function(s) {score(s)[1]}),
+			n=sapply(sim.prioritisations[[i]], function(s) {sum(selections(s))})
 		)
 	}
 )

### 2016-06-23 code/R/analysis/03-case-study-1.R

@@ -151,7 +151,9 @@ cs1.spp.DF <- ldply(
 	.fun=function(i) {
 		mutate(
 			extractResults(cs1.prioritisations[[i]]),
-			Prioritisation=names(cs1.prioritisations)[i]
+			Prioritisation=names(cs1.prioritisations)[i],
+			amount.held=amount.held*100,
+			niche=niche*100
 		)
 	}
 )

### 2016-06-23 code/R/analysis/04-case-study-2.R

@@ -153,7 +153,9 @@ cs2.spp.DF <- ldply(
 	.fun=function(i) {
 		mutate(
 			extractResults(cs2.prioritisations[[i]]),
-			Prioritisation=c('Amount','Genetic')[i]
+			Prioritisation=c('Amount','Genetic')[i],
+			amount.held=amount.held*100,
+			genetic=genetic*100
 		)
 	}
 )

### 2016-06-23 code/R/analysis/05-statistical-analysis.R

@@ -3,6 +3,55 @@ session::restore.session('data/intermediate/04-case-study-2.rda')
 load('data/intermediate/03-case-study-1.rda')
 load('data/intermediate/02-simulations.rda')
 
+## simulation study
+
+## case-study 1
+
+
+
+## case-study 2
+
+# generate inverse distance matrix
+cs2.pu.centroids <- gCentroid(cs2.grid.sub.PLY, byid=TRUE) %>% spTransform(CRS('+init=epsg:4326'))
+cs2.pu.inv.dists <- cs2.pu.centroids@coords %>% rdist.earth
+cs2.pu.inv.dists <- 1/cs2.pu.inv.dists
+diag(cs2.pu.inv.dists) <- 0
+
+# calculate Morans I
+cs1.nmds.MoransI <- llply(
+	unique(cs2.spp.samples.sub.DF$species),
+	function(x) {
+		curr.DF <- select(cs2.grid.sub.DF, contains(paste0(x,'_genetic_d')))
+		return(
+			llply(
+				seq_len(ncol(curr.DF)),
+				function(i) {
+					curr.col <- curr.DF[[i]]
+					curr.pos <- which(!is.na(curr.DF[[i]]))
+					return(
+						Moran.I(curr.DF[[i]][curr.pos], cs2.pu.inv.dists[curr.pos, curr.pos])
+					)
+				}
+			)
+		)
+	}
+)
+
+# calculate numbers and generate vectors for article
+amount.represented.species.names <- filter(cs1.spp.DF, Prioritisation=='Amount targets', niche >= cs1.params.LST[[MODE]]$'space.target')$Species
+amount.not.represented.species.names <- filter(cs1.spp.DF, Prioritisation=='Amount targets', niche<cs1.params.LST[[MODE]]$'space.target')$Species
+
+if (length(amount.represented.species.names)>1) {
+	parsed.representative.space.held.names <- paste0(paste(amount.represented.species.names[-length(amount.represented.species.names)], collapse=', '), ', and ', last(amount.represented.species.names))
+} else {
+	parsed.representative.space.held.names <- amount.represented.species.names
+}
+
+if (length(amount.not.represented.species.names)>1) {
+	parsed.not.representative.space.held.names  <- paste0(paste(amount.not.represented.species.names[-length(amount.not.represented.species.names)], collapse=', '), ', and ', last(amount.not.represented.species.names))
+} else {
[…]

### 2016-06-23 code/rmarkdown/article.Rmd
+ fix incorrect ref in article

@@
 Here we used of space-based targets to generate a prioritisation to secure a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@462; see @r461 and @r463 for further explanation of data collection methods].
+We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@r462; see @r461 and @r463 for further explanation of data collection methods].
 Althougthis dataset contains multiple plant species, we used data for the `r  cs2.params.LST[[MODE]]$common.name` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}).

### 2016-06-24 code/rmarkdown/article.Rmd
+ assorted improvments to article + added in table s1 + added code to allow analysis and ms compilation on different machines

@@
 ```{r, include=FALSE}
+## initialization
 # load .rda file
@@
 options(error=NULL)
+
+## preliminary processing
+for (i in seq_len(nlayers(cs1.bioclim.RAST)))
+	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])
+
 ```
@@
 Short running title: Representative and adequate prioritisations \newline
-Word count: XXXXX \newline
-Number of references: XXXX \newline
-Number of figures: XXXX \newline
-Number of tables: XXXX \newline
-Number of textboxes: XXXX \newline
-Number of equations: XXXX \newline
+Word count: 322 / 350 \newline
+Number of references: 34 \newline
+Number of figures: 8 \newline
+Number of tables: 0 \newline
+Number of textboxes: 0 \newline
+Number of equations: 8 \newline
 \newpage
@@
 <!-- conservation and reserve networks -->
-The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255].
+The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; r255].
 To achieve this, conservation actions must preserve biodiversity patterns (eg.
@@
 \begin{align*}
-T_s &= \text{amount target for feature $f$} \tag{1a}\\
+T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
 %
-\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag{1b}\\
+\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
 %
-B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag{1c}\\
+B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
 \end{align*}
@@
 	&= \begin{cases}
-		1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag{2a} \\
+		1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
 		0, & \parbox{25em}{otherwise} \\
@@
 Y_{fsij} &= \begin{cases}
-		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag{2b} \\
+		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
 		0, & \parbox{25em}{otherwise} \\
@@
 \begin{align*}
-& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag{3a} \\
+& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
 %
-& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag{3b}\\
+& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
 %
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq \tau_{fs}  & \forall & 0 \leq f \leq F-1, \tag{3c}\\
+& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq \tau_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
 & & & & & 0 \leq s \leq S-1\\
 %
-& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag{3d}\\
+& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
 & & & & & 0 \leq s \leq S-1, \\
@@
 %
-& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag{3e}\\
+& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
 & & & & & 0 \leq s \leq S-1, \\
@@
 %
-& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag{3f}\\
+& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 It has a core range area, and is less likely to be found in areas that are distant from the core area.
-This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8b).
+This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8c).
 The third species represents a species with two distinct populations (Fig.
 2c).
-This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8c).
-The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8c.
+This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8d).
+The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d.
 We used a geographic attribute space to showcase the behavior of the problem.
@@
 \begin{align*}
-& P (\text{uniform species} | x, y) &= 0.1 & \tag{8a}\\
-& P (\text{normal species} | x, y) &= XXXX & \tag{8b}\\
-& P (\text{uniform species} | x, y) &= XXX & \tag{8c}\\
+\\
+P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
+%
+f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
+%
+P \left( \text{normally distributed species} | \left( x , y \right)  \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
+%
+P \left( \text{bimodally distributed species} | \left(  x,y \right) \right) &= \text{Max} \begin{cases}
+f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
+\end{cases} \tag*{eqn 8d} \\
+%
 \end{align*}
@@
 Here we investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species.
+By preserving the populations in suitable habitats with different with environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63].
 We selected Queensland, Australia as the study area.
@@
 Yet it failed to achieve this for `r parsed.not.representative.space.held.names`.
+This result suggests that prioritisations generated using conventional methods may preserve a representative sample of most species realised niches (Figure S1).
+However, despite this, there may yet be species for which only a small fraction of their realised niche is preserved.
+By explictly using space targets, conservation planners can generate prioritisations that guaranteeably preserve a representative sample of species' niches.
 
@@
 # References
+
+\bibliography{references}

### 2016-06-24 code/rmarkdown/figures.Rmd
+ assorted improvments to article + added in table s1 + added code to allow analysis and ms compilation on different machines

@@
 
-# prelimary processing
+## prelimary processing
+# redefine file paths for rasters
+for (i in seq_len(nlayers(cs1.bioclim.RAST)))
+	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])
+
+# fortify polygons
 sim.pus.FPLY <- fortify(sim.pus)

### 2016-06-24 code/rmarkdown/supporting-information.Rmd
+ assorted improvments to article + added in table s1 + added code to allow analysis and ms compilation on different machines

@@
 options(error=NULL)
+
+## prelimary processing
+# redefine file paths for rasters
+for (i in seq_len(nlayers(cs1.bioclim.RAST)))
+	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])
+
 ```
@@
 
-## References
+```{r, echo=FALSE, results='asis'}
+# make table
+term.CHR <- c(
+	'$F$', 'set of biodiversity features (indexed by $f$)',
+	'$J$', 'set of planning units (indexed by $j$)',
+	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
+	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.',
+	'$S$', 'set of attribute spaces (indexed by $S$)',
+	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
+	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$.',
+	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$.',
+	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$.',
+	'$T_f$', 'amount target for feature $f$.',
+	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$.',
+	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0).',
+	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units.'
+)
+term.DF <- data.frame(
+	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
+	Description=term.CHR[seq(2, length(term.CHR), 2)]
+)
+term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')
+
+# render table
+curr.file <- tempfile(fileext='.tex')
+tmp <- latex(
+		term.DF,
+		file=curr.file, 
+		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
+		colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
+		caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritisation (URAP) problem.'
+	)
+readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
+	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
+	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+	cat()
+```
+
+# References

### 2016-06-24 code/R/analysis/05-statistical-analysis.R
+ bug fix in summarising results for default parameters

@@ -61,8 +61,10 @@ if (length(amount.not.represented.species.names)>1) {
 		}
 	)
 	parsed.not.representative.space.held.names <- paste0(paste(parsed.not.representative.space.held.names[-length(parsed.not.representative.space.held.names)], collapse=', '), ', and the ', last(parsed.not.representative.space.held.names))
-} else {
+} else if (length(amount.not.represented.species.names)==1) {
 	parsed.not.representative.space.held.names <- paste0(amount.not.represented.species.names, '(', round(filter(cs1.spp.DF, Prioritisation=='Amount targets', Species==amount.not.represented.species.names)$niche,2), ' %)')
+} else {
+	parsed.not.representative.space.held.names=c()
 }
 
 ## save workspace

### 2016-06-24 code/rmarkdown/article.Rmd
+ fix equations in main text and supp info

@@
 Short running title: Representative and adequate prioritisations \newline
-Word count: 322 / 350 \newline
+Abstract word count: 322 / 350 \newline
+Total word count: XXXXX / 8000 \newline
 Number of references: 34 \newline
@@
 <!-- conservation and reserve networks -->
-The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; r255].
+The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255].
 To achieve this, conservation actions must preserve biodiversity patterns (eg.
@@
 %
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq \tau_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
+& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
 & & & & & 0 \leq s \leq S-1\\
@@
 
-The objective function (3a) determines the utility of a given prioritisation: a combination of the total cost of a prioritisation and how fragmented it is.
-Constraints (3b) ensure that all the amount-based targets are met.
-Constraints (3c) ensure that all the space-based targets are met for each feature and each attribute space.
+The objective function (eqn 3a) determines the utility of a given prioritisation: a combination of the total cost of a prioritisation and how fragmented it is.
+Constraints (eqn 3b) ensure that all the amount-based targets are met.
+Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space.
 For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$).
@@
 The constraints ensure that this proportion must be equal to or greater than the space-based target.
-Constraints (3d) ensure that only one planning unit is assigned to each demand point.
-Constraints (3e) ensure that demand points are only assigned to selected planning units.
-Constraints (3f) ensure that the $X$ and $Y$ variables are binary.
+Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point.
+Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units.
+Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.
 
@@
 However, the non-linear components can be linearised using existing techniques.
-The expression $X_j X_k$ in (3a) can be linearised using methods described by Beyer et al.
+The expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al.
 [-@r426].

### 2016-06-24 code/rmarkdown/supporting-information.Rmd
+ fix equations in main text and supp info

@@
 \begin{align*}
-& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag{4} \\
+& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag*{eqn 4} \\
 & & & 0 \leq s \leq S-1\\
@@
 
-The control variables for the reliable formulation are the $B$ (1a), $T_{s}$ (1b), $\tau_{sa}$ (1c), $R$, and $M$ variables.
+The control variables for the reliable formulation are the $B$ (eqn 1a), $T_{s}$ (eqn 1b), $\tau_{sa}$ (eqn 1c), $R$, and $M$ variables.
 
 \begin{align*}
-R &= \parbox{25em}{number of failure levels} \tag{5a} \\
+R &= \parbox{25em}{number of failure levels} \tag*{eqn 5a} \\
 %
-M &= \parbox{25em}{failure multiplier} \tag{5b}\\
+M &= \parbox{25em}{failure multiplier} \tag*{eqn 5b}\\
 %
@@
 
-The decision variables are the $X_j$ (2a), $Y_{fsijr}$, $P_{fsijr}$ variables.
+The decision variables are the $X_j$ (eqn 2a), $Y_{fsijr}$, $P_{fsijr}$ variables.
 
@@
 Y_{fsijr} &= \begin{cases}
-		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag{6a} \\
+		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag*{eqn 6a} \\
 		0, & \parbox{25em}{otherwise} \\
@@
 %
-P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag{6a}\\
+P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag*{eqn 6a}\\
 %
@@
 %
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \left( \lambda_{fsi} d_{fsij} P_{fsijr} Y_{fsij} \right)^2}{\sum_{i=0}^{I-1} \left( \lambda_{fsi} \delta_{fsi} \right)^2 } \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag{7a}\\
+& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} P_{fsijr} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 7a}\\
 & & & & & 0 \leq s \leq S-1\\
 %
-& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7b}\\
+& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 7b}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag{7c}\\
+& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 7c}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag{7d}\\
+& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 7d}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag{7e}\\
+& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 7e}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag{7f}\\
+& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag*{eqn 7f}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag{7g}\\
+& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 7g}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag{7h}\\
+& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 7h}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 
-The objective function for the reliable formulation is the same as for the unreliable formation (3a).
-Similar to the unreliable formulation, constraints (3b) and (7a) ensure that the amount-based and space-based targets are met.
-Constraint (7b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$.
-Constraints (7d) ensure that only selected planning units are assigned to demand points $i \in I$.
-Constraints (7e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level.
-Constraints (7f--7g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information].
-Constraints (7h) ensure that the $X$ and $Y$ variables are binary.
+The objective function for the reliable formulation is the same as for the unreliable formation (eqn 3a).
+Similar to the unreliable formulation, constraints (eqn 3b) and (eqn 7a) ensure that the amount-based and space-based targets are met.
+Constraint (eqns 7b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$.
+Constraints (eqn 7d) ensure that only selected planning units are assigned to demand points $i \in I$.
+Constraints (eqn 7e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level.
+Constraints (eqns 7f--7g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information].
+Constraints (eqn 7h) ensure that the $X$ and $Y$ variables are binary.
 
@@
 However, the non-linear components can be linearised.
-First--as discussed in the main text--the expression $X_j X_k$ in (3a) can be linearised using methods described by Beyer et al.
+First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al.
 [-@r426].
-Second, the expression $P_{fsijr} Y_{jsijr}$ in (7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
+Second, the expression $P_{fsijr} Y_{jsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
 [-@r16].

### 2016-07-18 article/article.tex
+ assorted ms fixes including Rich's edits + results analysed + change to raptr

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rapr: Representative and Adequate Prioritisations in R},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{rapr: Representative and Adequate Prioritisations in R}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^1$,
Richard A. Fuller$^1$\\$^1$School of Biological Sciences, The University
of Queensland, Brisbane, QLD, Australia\\$^2$School of Geography,
Planning and Environmental Management, The University of Queensland,
Brisbane, QLD, Australia\\Correspondance should be addressed to
\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{18 July 2016}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,titling,natbib,pdfpages,lineno,booktabs}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% format toc
%\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}

% line numbers
\linenumbers

% format abstract
\renewcommand{\abstractname}{Summary}
\renewenvironment{abstract}
 {\small
  \begin{center}
  \bfseries \abstractname\vspace{-.5em}\vspace{0pt}
  \end{center}
  \list{} {%
   \setlength{\leftmargin}{2mm}
   \setlength{\rightmargin}{\leftmargin}%
  }%
  \item\relax}
{\endlist}

% format section headers
\titleformat*{\section}{\Large\bfseries}
\titleformat{\subsection}[display]
	{\large\sffamily\lsstyle}
	{\subsectiontitlename\ \thesubsection}
	{0.5em}{}
\titleformat*{\subsubsection}{\large\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

% define struts for tables
\newcommand\T{\rule{0pt}{2.6ex}} % top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut

% define command to put new lines in table cells
%\newcommand{\specialcell}[2][c]{%
%  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\begin{document}

\maketitle

\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  An underlying aim in spatial conservation planning is to maximise the
  long-term persistence of biodiversity. Yet current approaches
  overwhelming focus on ``adequacy'', the concept ensuring a target
  proportion of the distribution of a biodiversity feature (species or
  ecosystem) is reserved. Ultimately, to achieve long-term persistence,
  the processes the sustain biodiversity must also be conserved. Many
  biodiversity processes can be preserved by conserving a representative
  sample of each biodiversity feature. This idea has for decades been
  encapsulated in principle of ``representativeness''. For example,
  representing the full range of genetic diversity or ability to persist
  in a range of climatic conditions might be crucial for the long -term
  persistence of a species in a world where environmental change is
  accelerating. However, fully operationalising the principle of
  repressiveness in spatial conservation planning has proven
  methodologically challenging.
\item
  To address this issue, we developed the
  \texttt{rapr: Representative and Adequate Pioritisations in R}, a
  toolkit to guide reserve selection using explicit targets for
  representing biodiversity processes as well as meeting area protection
  targets. Using a novel formulation of the reserve selection problem,
  users set ``space targets'' to secure a representative sample of each
  species within an attribute space reflecting any biodiversity process
  (eg. an attribute space expressing variation in genetic
  characteristics between individuals in an area, or variation in
  climatic conditions between areas occupied by indivduals). We explored
  the functionality of this \texttt{R} package using a simulation study
  and two case studies. In each study, we generated prioritisations that
  aimed to preserve an amount of habitat for each species using amount
  target--to represent conventional reserve selection methods--and
  compared them with prioritisations generated using both habitat and
  space targets--to represent this method.
\item
  We show that explicitly considering biodiversity processes in reserve
  selection can substantially change the configuration of the resulting
  solution. Simulations suggest that including representativeness in
  conservation planning is particularly important where the biodiversity
  feature has multi-model distributions in an attribute space. Results
  from initial case studies show that explicitly considering space-based
  targets for biodiversity processes can result in prioritisations that
  do not necessarily require much more area than traditional approaches,
  but that could be much more effective in achieving long term
  biodiversity persistence.
\item
  The \texttt{rapr R} package provides a unified framework achieving
  spatial conservation prioritisations that are both adequate and
  representative, resulting in a greater chance of long term
  biodiversity persistence that area-based planning alone.
\end{enumerate}
\end{abstract}

Short running title: Representative and adequate prioritisations
\newline
Abstract word count: 370 / 350\\Total word count: XXXXX / 8000\\Number
of references: 34\\Number of figures: 8\\Number of tables: 0\\Number of
textboxes: 0\\Number of equations: 8\\\newpage

\section{Introduction}\label{introduction}

Perhaps the most fundamental aim of conservation is to maximise the
long-term persistence of biodiversity at organisational levels (McNeely
1994; Margules \& Pressey 2000). To achieve this, conservation actions
must preserve biodiversity patterns (eg. populations, species,
ecosystems), but also crucially the processes that sustain them. One of
the major tangible achievements of modern conservation has been the act
of setting aside areas for preservation (Sanderson \emph{et al.} 2015).
Reserve networks buffer species from gross threatening processes and set
the stage for enhanced management interventions (Gaston \emph{et al.}
\{2008\}). However, the resources available for conservation action are
limited, and so reserve networks must be sited in places that satisfy
conservation objectives for minimum cost (Margules \& Pressey 2000). To
achieve this, reserve selection is often formulated as an optimisation
problem and then solved to identify cost-effective candidate reserve
systems (termed prioritisations; Margules \& Pressey 2000). Many
decision support tools have been developed to help identify
cost-effective solutions (eg. \texttt{ ConsNet}, Ciarleglio \emph{et
al.} 2009;\texttt{ Marxan}, Ball \emph{et al.} 2009;\texttt{ Zonation},
Moilanen 2007). Typically, these tools are used to deliver a
prioritisation that preserves a set of species of interest to the
conservation planner (termed target species or features). Given a set of
species, decision makers can use these tools to preserve biodiversity
patterns by generating a prioritisation that secures an adequate
proportion of each species' range.

However, to achieve effective conservation, reserve networks must do
more than cover enough area. They must capture the processes that
sustain biodiversity over the long term; they must capture both
ecological and evolutionary processes (Crandall \emph{et al.} 2000;
Margules \& Pressey 2000). Ecological processes, such as predator-prey
interactions, pollination, and decomposition, are required for
biodiversity to persist over short time-scales. Typically, they operate
over small geographic domains--with exceptions such as migration and
refugial habitats--and can be preserved using suitably large planning
units (Ciarleglio \emph{et al.} 2009) that each contain a discrete unit
of habitat (Klein \emph{et al.} 2009). On the other hand, evolutionary
processes are required for biodiversity to persist over long
time-scales, and they typically operate over large geographic domains.
Protected areas must represent adaptive evolutionary processes to foster
resilience against environmental change (eg. climate change; Pyke \&
Fischer 2005). Protected areas must also preserve neutral evolutionary
processes (Moritz 1994, 1999, 2002), arising from restriction of gene
flow between populations. They are important for maintaining genetic
diversity, and avoiding inbreeding depression. In recent decades, a
wealth of data relevant to evolutionary processes has become available
to conservation planners (eg. bioclimatic, genetic, and trait data;
Hijmans \emph{et al.} 2005; Raymond \emph{et al.} 2015; Jones \emph{et
al.} 2009). Yet such data are rarely used to guide conservation planning
(Hendry \emph{et al.} 2010). This is perhaps in large part due to that
fact that existing reserve selection tools focus on capturing either
biodiversity patterns or processes--but not both.

Existing conservation planning tools are generally not well-suited for
representing biodiversity processes (but see Faith 2003). To preserve
biodiversity processes, a prioritisation must capture a representative
sample of each species. For instance, to preserve predator-prey
interactions, a prioritisation must preserve individuals from each
predator and prey species in the same area. To preserve adaptive
evolutionary processes, a prioritisation must preserve the adaptive
landscape of each species--populations experiencing different selection
pressures (Moritz 2002). To preserve neutral evolutionary processes, a
prioritisation must secure individuals descended from each of the
genetic lineages that comprise each species (Moritz 1994). Previous
attempts to accommodate such into multi-species conservation planning
have partitioned species' ranges into different groups, for example at
habitat discontinuities (e.g. Carvalho \emph{et al.} 2011) as a
pre-processing step to render a new set of ``pseudo-species''. However,
this approach assumes that the biodiversity processes that operate
across a species' range can be readily split into discrete units. Yet
data on biodiversity processes is often continuous and
hyper-dimensional, and often cannot be reduced to a few categories
without significant information loss (Faith \& Walker 1996).

One of the key issues in reserve selection is the lack of a unifying
decision support tool that can accommodate data on biodiversity patterns
and processes in a multi-species context. To begin to fill this gap, we
present the \texttt{rapr R} package. This \texttt{R} package uses novel
formulations of the reserve selection problem to provide decision makers
with the tools to generate prioritisations that preserve both
biodiversity patterns and processes. We aim to define the concepts
behind the problem formulations. Furthermore, we aim to explore the
functionality of the \texttt{rapr R} package by applying one of the
formulations to a set of simulated species and two case studies.

\section{Methods}\label{methods}

\subsection{PROBLEM FORMULATION}\label{problem-formulation}

Biodiversity features are defined as the entities that the
prioritisation is required to preserve (eg. species, ecosystems).
Spatial attributes are defined as the variation across the species'
range that the prioritisation is required to sample. They can be
intrinsic (eg. genetic or trait variation) or extrinsic (eg.
environmental variation) to the feature. These attributes should be
related to the biodiversity processes that the decision maker aims to
preserve.

Each attribute is conceptualised as a space. This space is termed an
``attribute space''. Each planning unit is thought to occupy a single
point inside each space. For example, a decision maker may require a
prioritisation that represents populations along climatic gradients. To
achieve this, the decision maker might use an ``climatic'' attribute
space with dimensions relating to mean annual temperature ($^{\circ}$C)
and precipitation (mm). Any given combination of temperature and
precipitation may be conceived as a point in this environmental space.
By associating planning units with climatic data, they can be mapped
from geographic space to this environmental attribute space.

Demand points are points that also exist in an attribute space. They are
designated by the decision maker to indicate regions of the attribute
space that the decision maker wishes to represent in the prioritisation
(see below for discussion on how demand points can be generated for
real-world datasets). The amount of variation in the attribute space
that a prioritisation secures is a function of the distance between each
demand point and each selected planning unit in the attribute space. The
shorter the distances between the demand points and the planning units;
the better the prioritisation is at securing the variation in the
spatial attribute. To convert these amounts to a proportion--a
meaningful unit for a decision maker--the distances between the selected
planing units and the demand points are scaled by the distances between
the demand points to the centroid of the demand points. In any attribute
space there may exist points that are impossible (eg. mean annual
rainfall -5 mm), or do not occur in the study area (eg. mean annual
temperature 30$^{\circ}$C in Antarctica). Additionally, there may be
some regions that are desirable for some features and undesirable for
others (eg. conditions known to be outside the physiological tolerance
of certain species). Thus a different set of demand points and weights
are used for each attribute space and each feature. By placing demand
points in desirable regions of an attribute space for a given feature,
the decision maker can ensure that prioritisations secure the feature in
planning units with spatial attributes that are desirable for that
feature.

To illustrate these concepts, consider an example conservation planning
scenario example involving an attribute space and demand points. We wish
to develop a prioritisation for a single species that has four
populations in the study area. However, we can only afford to preserve
three populations. We aim preserve the adaptive landscape of the
species, and to achieve this, we will preserve populations inhabiting
different environmental conditions. To describe environmental variation,
we obtain data on the environmental conditions (rainfall (mm) and
temperature ($^{\circ}$C)) where each population is found. These
environmental data are then used to construct a two-dimensional
environmental attribute space. Next, we generated demand points as
equi-distant points between the range of values where the populations
were found. By comparing the distribution of the demand points to the
distribution of the populations in the attribute space, we can identify
a suitable prioritisation (Fig. 1). We can see that preserving both
populations $A$ and $C$, effectively ``doubles-up'' on the same
environmental characteristics, resulting in considerable redundancy in
the solution. Instead, a more representative sample of the
intra-specific variation could be captured by securing populations $A$,
$B$, and $D$. This example demonstrates how the inclusion of
biodiversity processes can guide the reserve selection process.

The formulations used to express the reserve selection problem in the
\texttt{rapr R} package are based on a combination of the
\texttt{Marxan} reserve selection problem and the uncapacitated facility
location problems (Cornu{é}jols \emph{et al.} 1990). Although the
\texttt{rapr R} package provides two novel formulations, for brevity, we
will define the simpler formulation--referred to as the unreliable
formulation--below and define the more complex version--the reliable
formulation--in the Supporting Information S1. These formulations are
named after the unreliable and reliable facility location formulations
from which they are based upon (Cui \emph{et al.} 2010). The key
difference between these two formulations is that the reliable
formulation explicitly considers the probability that planning units are
occupied when calculating the proportion of an attribute space sampled
in a solution. All mathematical terms defined hereafter are described in
Table S1. For convenience, the cardinality of sets will be denoted using
the same symbol used to denote the variable.

Define $F$ to be the set of features one wishes to conserve (indexed by
$f$). Let $J$ be a set of planning units (indexed by $j$). Also, let
$A_j$ denote the area, and $C_j$ denote the cost of preserving planning
unit $j \in J$. To assess the extent to which each feature is secured in
a given prioritisation, let $q_{fj}$ denote the probability of feature
$f$ occupying planning unit $j$. The level of fragmentation associated
with a prioritisation is parametrised as the net exposed boundary
length. Let the shared edges between each planning unit $j \in J$ and
$k \in J$ be $e_{jk}$.

Let $S$ denote a set of attribute spaces (indexed by $s$). Each
$j \in J$ is associated with spatially explicit data that represent
coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a
set of demand points (indexed by $i$) for each feature $f \in F$ and
each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting
for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$
denote the distance between each demand point $i \in I$ and each
planning unit $j \in J$ for each feature $f \in F$ and attribute space
$s \in S$. To describe the inherent variation in the distribution of
demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote
the distance between each demand point $i \in I$ and the centroid of the
demand points. Demand points with greater weight $\lambda_{fsi}$ are
more important, and the optimal solution will be likely to select
planning units close to highly weighted demand points. As a consequence,
the decision maker will need to choose an appropriate weighting for each
demand point.

Targets are used to ensure that prioritisations adequately preserve each
species. Amount-based targets are used to ensure that the total amount
of habitat preserved is sufficient. Let $T_f$ denote the expected amount
of area that needs to be preserved for each feature $f \in F$.
Space-based targets ensure that a sufficient proportion of the
intra-specific variation is secured. Let $\tau_{fs}$ denote the
space-based targets for feature $f \in F$ and attribute space $a \in A$.
For convenience, these both types of targets are expressed as
proportions in the \texttt{R} package.

The control variables for the unreliable formulation are the $B$,
$T_{s}$, and $\tau_{sa}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

The decision variables are the $X_j$ and $Y_{fsij}$ variables.

\begin{align*}
X_j
    &= \begin{cases}
        1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
        0, & \parbox{25em}{otherwise} \\
    \end{cases} \\
%
Y_{fsij} &= \begin{cases}
        1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
        0, & \parbox{25em}{otherwise} \\
    \end{cases} \\
\end{align*}

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is
assigned to a selected planning unit $J$ where $X_j = 1$. The weighted
distance between the demand point and its assigned planning unit
$\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is
represented in a given solution. Generally, demand points are assigned
to the closest selected planning units (unless particularly low
space-based targets are used).

The unreliable formulation (URAP) is a defined as a multi-objective
optimisation problem.

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

The objective function (eqn 3a) determines the utility of a given
prioritisation: a combination of the total cost of a prioritisation and
how fragmented it is. Constraints (eqn 3b) ensure that all the
amount-based targets are met. Constraints (eqn 3c) ensure that all the
space-based targets are met for each feature and each attribute space.
For each feature and attribute space, the total weighted distance
between the demand points and their closest selected planning units is
calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted
distance is then scaled by the inherent variation in the demand points
($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a
proportion conceptually similar to the $R^2$ statistic used in $k$-means
clustering analysis. The constraints ensure that this proportion must be
equal to or greater than the space-based target. Constraints (eqn 3d)
ensure that only one planning unit is assigned to each demand point.
Constraints (eqn 3e) ensure that demand points are only assigned to
selected planning units. Constraints (eqn 3f) ensure that the $X$ and
$Y$ variables are binary.

\subsection{OPTIMISATION}\label{optimisation}

The unreliable formulation is non-linear. However, the non-linear
components can be linearised using existing techniques. The expression
$X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer
et al. (2016). Linearised versions of the problems can be solved using
commercial exact algorithm solvers. The \texttt{rapr} R package provides
functions to express conservation planning data as optimisation problems
using linearised versions of the unreliable and reliable formulations.
These optimisation problems can then be solved to generate
prioritisations using the commercial \texttt{Gurobi} software suite
(\url{http://www.gurobi.com}). Presently, academics can obtain a license
at no cost from the Gurobi website. After installing the \texttt{Gurobi}
software suite, users will need to install the \texttt{Gurobi R}
package.

\section{Examples}\label{examples}

To understand the behaviour of the unreliable problem and showcase its
value, we conducted a simulation study and two case studies. These
studies involved generating solutions using only amount targets to
represent prioritisations generated using conventional methods (eg.
\texttt{Marxan}), and solutions using amount and space targets using the
unreliable formulation. By comparing these solutions, we can guage the
benefits of explicitly including space targets in reserve selection. All
analyses were performed in \texttt{R} (version 3.3.0; {R Core Team}
2016).

\subsection{SIMULATION STUDY}\label{simulation-study}

\subsubsection{Methods}\label{methods-1}

We simulated a hypothetical study area with planning units arranged in
$10 \times 10$ square grid (Fig. 2). We then simulated three species
across this study area. The first species was simulated to represent a
hyper-generalist--occurring in all planning units with equal probability
(Fig. 2a). This species' distribution was based on a uniform
distribution (eqn 8a). The second species was simulated to represent a
species with an idealised distribution (Fig. 2b). It has a core range
area, and is less likely to be found in areas that are distant from the
core area. This species' distribution was simulated using the
probability density function of a single multivariate normal
distribution (eqn 8c). The third species represents a species with two
distinct populations (Fig. 2c). This species' distribution was simulated
to depict a bimodal distribution, based on the combination of the
probability density functions of two multivariate normal distributions
(eqn 8d). The probability that each species inhabited a given planning
unit was calculated using the $X, Y$ coordinates of the planning unit
and eqns 8a--8d. We used a geographic attribute space to showcase the
behavior of the problem. For each species, demand points were generated
by calculating the centroids of each planning units and their weights
were set as the probability that the planning units were occupied.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 8d} \\
%
\end{align*}

We generated four solutions for each species. First to represent
solutions generated using conventional conservation planning methods, we
generated solutions using only 20 \% amount targets. Second to show how
the addition of geographic targets can affect a prioritisation, we
generated solutions using 20 \% amount targets and 75 \% geographic
targets. Third to represent solutions using conventional planning
methods that penalise for fragmentation, we generated solutions using 20
\% amount targets and a boundary length modifier of 1. Fourth to
illustrate the combined affect using geographic targets and penalising
fragmentation, we generated solutions using using 20 \% amount targets
and 75 \% geographic targets and a boundary length modifier of 1.

\subsubsection{Results}\label{results}

The uniform species was simulated to occur with a constant probability
of occupancy across the study area (Fig. 2a). The solution generated
using 20 \% amount-based targets (Fig. 3a) selected 20 planning units
near the southern end of study area. This configuration is an artifact
of the method used to solve this particular instance of the reserve
selection problem. In fact, for this species, all prioritisations
containing 20 planning units are optimal when considering only 20 \%
amount targets. In the absence of criteria to guarantee
representativeness, reserve selection methods may or may not return
solutions that secure a representative sample of the features. In this
particular case, the solution does not secure a representative sample of
the species' range (-23.64 \% sampled).

The addition of a 75 \% geographic target resulted in a solution that
secured a representative sample of the uniform species' range (90.67 \%
sampled; Fig. 3b). Since all planning units have equal cost and an equal
chance of being occupied, this solution has the same number of planning
units as the solution generated using only amount targets (cf.~Fig 3a).
Although the use of amount and space based targets has addressed
adequacy and representativeness objectives (respectively), they have
resulted in a highly fragmented solution.

The addition of a positive boundary length modifier (BLM) parameter
resulted in a well-connected solution that secured an adequate
proportion (20 \% ) and representative sample (75.76 \% sampled) of the
species geographic distribution (Fig. 3d). This solution contains 20
planning units--a few more than the previously discussed solutions--to
ensure that the solution secures a representative proportion of the
species geographic distribution in a configuration that is not highly
fragmented. This result suggests that the combination of space targets
and boundary length modifiers may yield solutions that contain more
planning units, since the solution generated using amount targets and a
boundary length modifier (Fig. 3c) contained the sample number of
planning units as the solution with just the amount targets (Fig. 3a).
Overall, these results show that under the simplest of conditions, the
reliance on just amount targets can cause an under-specified reserve
selection problem that is unlikely to return a suitable solution for
implementation.

The normally distributed species was simulated to contain a single core
area where individuals are most prevalent and marginal areas where
individuals are less likely to occur (Fig. 2b). The solution generated
using 20 \% amount targets contained 10 planning units and concentrated
conservation efforts in the core area (Fig. 3e). Since all planning
units have equal costs and areas, this solution contains the planning
units with the highest probabilities of occupancy. Whilst this solution
satisfies the adequacy objective for a prioritisation in a
cost-effective manner--it fails to fulfill the representative objective
for a prioritisation (59.09 \% distribution sampled).

The solution generated using 20 \% amount targets and 75 \% geographic
targets resulted in a solution that is both adequate and representative
in terms of the uniform species distribution (Fig. 3f). This solution
used 11 planning units to secure 20.02 \% and sample 76.31 \% of the
normal species' distribution. By using amount and geographic targets in
the reserve selection problem, we have obtained a solution that
concentrates conservation effort in the range core and also the range
margin. However, similar to the corresponding solution for the uniform
species (Fig. 3b), this solution is highly fragmented.

By using 20 \% amount targets and 75 \% geographic targets and a
boundary length modifier parameter ($BLM = 1$), we obtained a solution
that fulfills adequacy, representativeness and connectivity objectives
(Fig. 3h). This solution contains 12 planning units. Similar to the
corresponding solution for the uniform species (Fig. 3d), this solution
contains more planning units than any other solutions for this species
(cf.~Figs. 3e--3g). The results for the normally distributed species
suggest that the space targets can result in solutions that secure a
more representative sample of the species'--even for species without
significant structure.

The bimodally distributed species was simulated to represent a species
with a highly structured distribution, and it contains two distinct
populations (Fig. 2c). The solution generated using just 20 \% amount
targets assigned conservation effort to just one of the two populations
(Fig. 3i). While this solution secured an adequate proportion of the
species' distribution (20.18 \% secured), it did not sample a
representative proportion of the range (12 \% sampled). The addition of
boundary length modifiers to just amount targets resulted in a solution
which sampled even less of the species distribution (25.19 \% sampled;
Fig. 3k). However, the addition of geographic space targets resolved
these issues.

The solution generated using 20 \% amount and 75 \% geographic targets
included planning units from both populations (Fig. 3j). Although this
solution required more planning units to fulfill both targets ($n=9$),
this solution secured a representative sample of the bimodally
distributed species (80.95 \% range sampled). Similar to corresponding
solutions generated for both the uniformly (Fig. 3b) and normally
distributed species (Fig. 3f), this solution was also highly fragmented
and we could obtain a more well-connected solution at the expense of
selecting additional planning units ($n=9$; Fig. 3l). The results for
the bimodally distributed species suggest that species with highly
structured populations or strong variation between individuals could
benefit the most from prioritisations that are generated using
space-based targets.

\subsection{CASE STUDY 1}\label{case-study-1}

\subsubsection{Methods}\label{methods-2}

We investigated how space-based targets can be used in a multi-species
planning context to generate a prioritisation that sufficiently
preserves the realised niche for several species. By preserving the
populations in suitable habitats with different with environmental
conditions, conservation planners can preserve the species' adaptive
landscape and foster resilience against environmental change (Moritz
2002). We selected Queensland, Australia as the study area. This region
is ideal for exploring the potential of niche-based targets because it
contains a broad range of different habitats. We obtained data for 19
bioclimatic variables across the region (at $30^{\prime \prime}$
resolution from \url{www.worldclim.org}; Hijmans \emph{et al.} 2005) and
subjected them to a principal components analysis (using ArcMap 10.3.1).
We used the first two principal components (cumulatively explained 99.5
\% of the total variation) to characterise the environmental variation
across the study area (Fig. 4).

We selected four bird species that span a range of different
evolutionary histories, distributions and ecologies: blue-winged
kookaburra (\emph{Dacelo leachii}), brown-backed honeyeater
(\emph{Ramsayornis modestus}), brown falcon (\emph{Falco berigora}), and
pale headed rosella (\emph{Platycercus adscitus}). We then mapped the
extent of occurrence for each species (Fig. 5). To do this, we obtained
observation data from the Atlas of Living Australia across the whole of
Australia (using the \texttt{ALA4R R} package; Raymond \emph{et al.}
2015), spatially thinned the data to omit points within 10 km$^2$ of
each other to ameliorate the effects of sampling bias (using the
developmental version of the \texttt{spThin R} package;
\url{www.github.com/mlammens/spThin}; Aiello-Lammens \emph{et al.}
2015), and fit 85 \% minimum convex polygons (using the
\texttt{adehabitatHR R} package; Calenge 2006). We used this method
because it is trivially reproducible using freely available data.

We generated 500 demand points for each species (Fig. S2). To achieve
this, for each species, we generated random points inside the species
range and at each point extracted the principal component values at that
location. We then fit hyperbox kernels to the distribution of these
points to characterise the realised niche of each species (using a
manually chosen bandwidth of 0.2 and a 0.5 quantile to map the core
parts of the species' niches; implemented in the \texttt{hypervolume R}
package; Blonder \emph{et al.} 2014). We then generated uniformly
distributed points inside the species' kernels and estimated the density
of the training points at the uniformly generated points. These
uniformly distributed points and associated density estimated were used
as demand point coordinates and weights (respectively).

We generated two solutions. The first solution was generated using 20 \%
amount-based targets for all species. The second solution was generated
the same amount-based target with an additional 75 \% niche-based for
each species.

\subsubsection{Results}\label{results-1}

Generally, the solution generated using amount targets preserved a
representative sample of each the four bird species' niches--with one
excpetion. Specifically, this solution sampled over 75 \% of the
realised niche of blue-winged kookaburra (89.54 \%), brown falcon (92.31
\%), and the pale-headed rosella (76.86 \%). Yet it failed to achieve
this for Brown-backed honeyeater(30.93 \%). This result suggests that
prioritisations generated using conventional methods may preserve a
representative sample of most species realised niches (Figure S1).
However, despite this, there may yet be species for which only a small
fraction of their realised niche is preserved. By explictly using space
targets, conservation planners can generate prioritisations that are
guaranteed to capture a representative sample of species' niches.

\subsection{CASE STUDY 2}\label{case-study-2}

\subsubsection{Methods}\label{methods-3}

Here we used space-based targets to generate a prioritisation securing a
representative sample of a species' intra-specific genetic variation. We
used species occurrence and multilocus AFLP data collected by the
international IntraBioDiv project in the European Alps (Meirmans
\emph{et al.} 2011; see Alvarez \emph{et al.} 2009 for further
explanation of data collection methods). Althougthis dataset contains
multiple plant species, we used data for the betony-leaved rampion
(\textit{Phyteuma betonicifolium}), a wide-spread species with
significant genetic structure (Aiello-Lammens \emph{et al.} 2015).
Members of the IntraBioDiv project collected data using a $20^{\prime}$
longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25
km; Fig. 7a). They visited every second grid cell, and if the
betony-leaved rampion was detected in a cell, samples were collected
from three individuals. Samples were genotyped using amplified fragment
length polymorphisms (AFLP; Raymond \emph{et al.} 2015), and used to
construct matrices denoting the presence/absence of polymorphisms at
loci. In total, 131 individuals were genotyped at 138 markers.

We used non-metric multi-dimensional scaling (NMDS; using Gower
distances to accommodate sparsity; Gower 1971; implemented the
\texttt{cluster R} package; Maechler \emph{et al.} 2015) to ordinate the
presence (or absence) of locus-specific alleles wthin individuals into
two continuous variables (implemented in the \texttt{vegan R} package;
Oksanen \emph{et al.} 2015). These continuous variables described the
main axes of genetic variation within the species. We calculated the
average of the values associated with individuals in each grid cell.
These values were used to create a genetic attribute space (Figs. 7c and
7d). To assess spatial auto-correlation, we calculated Moran's I
auto-correlation index for each NMDS axis using inverse great circle
distances based on the grid cells' centroids (using the \texttt{ape R}
package; Paradis \emph{et al.} 2004).

The grid cells were used as planning units. The grid-cell averaged
ordinations were used to describe the typical genetic characteristics of
individuals in the planning unit. Since the number of planning units was
relatively small, we used the same grid-cell averaged ordinations as
demand points. To ensure that the solutions did not prioritise
particularly costly areas, we included opportunity cost data (Fig. 7b).
We obtained population density data (1 km resolution from the Global
Rural-Urban Mapping Project; \texttt{GRUMP V1}; CIESEN, Columbia
University \emph{et al.} 2011) and estimated the total population
density inside each grid cell.

We generated two solutions. The first solution was generated using an 10
\% amount-based target. The second solution was generated using the same
amount-based target and an additional 95 \% genetic-based target.

\subsubsection{Results}\label{results-2}

The binary genetic data was ordinated into a two-dimensional space that
described the main different between individuals (stress = 0.17). These
values were then averaged to the planning unit level to describe the
typical genetic characteristics of individuals in each planning unit
(Figs. 7c--7d). Planning units located near each were found to contain
individuals with similar genetic characteristics (Moran's I: NMDS axis
1, $I = 0.4$, $P < 0.001$; Moran's I: NMDS axis 2, $I = 0.32$,
$P < 0.001$). In terms of the average genetic characteristics of
individuals in the planning units, they tended to cluster into two main
groups, with evidence of within-group structure inside the larger group
(Fig 8c). This analysis supports previous work by Alvarez \emph{et al.}
(2009) who also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a
representative portion of the species' genetic variation (15.54 \%
sampled; Figs. 8a and 8c). This solution only sampled planning units
that contained individuals belonging to one of the main two groups. On
the other hand, the solution generated using additional genetic targets
selected planning units belonging to both of the main groups. Note that
since these solutions were generated using opportunity cost data, the
solution generated using only the amount target is essentially the
minimum number of least costly areas needed to fulfill the amount
target. The only difference between the two solutions is a single
planning unit. By swapping this single planning unit, the solution
generated using amount and genetic targets was able to preserve a
representative portion of the species' genetic variation 15.54 for only
a minor increase in cost (99.41 total cost compared to 99.66 total
cost).

\section{Implications and future
directions}\label{implications-and-future-directions}

The \texttt{rapr} R package provides a unified approach to reserve
selection. One of the key advantages of this package is that it is
general enough to incorporate any spatial variable as an attribute
space. This package can accommodate intrinsic or extrinsic variation to
the feature(s). For example, adaptation processes could be secured using
environmental variation or genetic variation among loci under selection.
Ecological processes, such as predator-prey interactions, could be
secured by capturing intra-specific trophic variation. Different
attribute spaces with different targets can be also specified for
different species. As long as the variation can be described using
Euclidean distances, the \texttt{R} package can be used to obtain a
representative sample (note that data may require transformation to
conform to this assumption). This \textbackslash{}texttt\{R\} package
provides decision makers with the tools to generate prioritisations that
secure both biodiversity patterns and the processes that maintain
biodiversity. Additionally, the package contains functionality to
accommodate uncertainty in the distribution of features, and also
identify suitably connected reserves. Both the simulated and case study
species suggest that conservation planning exercises need to explicitly
consider biodiversity processes during the reserve selection process to
ensure they are captured.

The degree to which a prioritisation truly secures a representative
sample of a feature depends on the attribute spaces and distribution of
demand points chosen by the decision maker. Ultimately, the space-based
targets are set as a proportion based on the distribution of the
demand-points. As a consequence, if the decision maker uses an
inappropriate set of spatial variables to construct an attribute space,
or an inappropriate set of demand points, then the optimal solution will
not be an effective prioritisation. We therefore stress that decision
makers must carefully consider which biodiversity processes need to be
reflected in the prioritisation, and which spatial data can be used to
represent these processes. Note that maximising one process can be
detrimental to another. For instance, to maximise the geographic spread
of a prioritisation, reserves need to be further away from each other,
yet to maximise the connectivity of a prioritisation, reserves need to
be closer to each other. To assist in the selection of appropriate
demand points, the \texttt{R} package provides several routines for
generating demand points (see the \texttt{make.DemandPoints} function).
These routines essentially use the distribution of a feature in the
attribute space to define a polygon. Demand points are then generated as
random points within the polygon. A kernel is then fit to the
distribution of the feature in the space (using Blonder \emph{et al.}
2014; Duong 2015), and the demand points are weighted based on the
estimated density of the feature at the demand points.

The formulation requires spatially comprehensive data to map attribute
spaces to planning units. For instance, in both case-studies, all
planning units occupied by the species were associated with
values/coordinates in the attribute spaces. However, most real-world
data sets are patchy--some planning units will be occupied by species
for which attribute space data is not available. To use such patchy data
with the \textbackslash{}texttt\{rapr R\} package, the gaps in the data
must first be filled in as a pre-processing step. Spatially explicit
models could be used to estimate values in locations that are missing
from a data set (eg. krigging or generalised dissimilarity models;
Oliver \& Webster 1990; Ferrier 2002). This approach has been
successfully applied to a range of biological data sets (Thomassen
\emph{et al.} 2010).

To maximise the long-term persistence of biodiversity, decision makers
need to identify prioritisations that preserve existing patterns of
biodiversity and the processes that support them. To achieve this,
conservation planners need a decision support tool that can explicitly
accommodate biodiversity patterns and processes. Here, we developed the
\texttt{rapr R} package to fill this void. By exploring the
functionality of this package using several simulated species, we found
that including space-based targets can radically change a prioritisation
for the simplest of species.

\section{Acknowledgements}\label{acknowledgements}

JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
has an Australian Research Council Future Fellowship. This work was
supported by the Centre of Excellence for Environmental Decisions (CEED)
and the Landscape Ecology and Conservation Group (LEC) at The University
of Queensland.

\section{Data accessibility}\label{data-accessibility}

All data, code, and results are stored in an online repository
(\url{www.github.com/paleo13/rapr-manuscript}) to permit replication and
validation of this study.

\section{References}\label{references}

\bibliography{references}

Aiello-Lammens, M.E., Boria, R.A., Radosavljevic, A., Vilela, B. \&
Anderson, R.P. (2015). spThin: an R package for spatial thinning of
species occurrence records for use in ecological niche models.
\emph{Ecography}, \textbf{38}, 541--545.

Alvarez, N., Thiel-Egenter, C., Tribsch, A., Holderegger, R., Manel, S.,
Sch{ö}nswetter, P., Taberlet, P., Brodbeck, S., Gaudeul, M., Gielly, L.,
K{ü}pfer, P., Mansion, G., Negrini, R., Paun, O., Pellecchia, M., Rioux,
D., Sch{ü}pfer, F., Van Loo, M., Winkler, M., Gugerli, F. \& IntraBioDiv
Consortium. (2009). History or ecology? Substrate type as a major driver
of patial genetic structure in Alpine plants. \emph{Ecology Letters},
\textbf{12}, 632--640.

Ball, I., Possingham, H. \& Watts, M.E. (2009). Marxan and relatives:
software for spatial conservation prioritisation. \emph{Spatial
conservation prioritisation: Quantitative methods \& computational
tools} (eds A. Moilanen, K.A. Wilson \& H. Possingham), pp. 185--189.
Oxford University Press, Oxford, UK.

Beyer, H.L., Dujardin, Y., Watts, M.E. \& Possingham, H.P. (2016).
Solving conservation planning problems with integer linear programming.
\emph{Ecological Modelling}, \textbf{328}, 14--22.

Blonder, B., Lamanna, C., Violle, C. \& Enquist, B.J. (2014). The
\emph{n}-dimensional hypervolume. \emph{Global Ecology and
Biogeography}, \textbf{23}, 595--609.

CIESEN, Columbia University, International Food Policy Research
Institute, The World Bank \& Centro Internacional de Agricultura
Tropical. (2011). Global rural-urban mapping project, Version 1 (GRUMP
v1): Urban extents grid.

Calenge, C. (2006). The package adehabitat for the R software: tool for
the analysis of space and habitat use by animals. \emph{Ecological
Modelling}, \textbf{197}, 1035.

Carvalho, S.B., Brito, J.C., Crespo, E.J. \& Possingham, H.P. (2011).
Incorporating evolutionary processes into conservation planning using
species distribution data: a case study with the western Mediterranean
herpetofauna. \emph{Diversity \& Distributions}, \textbf{17}, 408--421.

Ciarleglio, M., Wesley Barnes, J. \& Sarkar, S. (2009). ConsNet: new
software for the selection of conservation area networks with spatial
and multi-criteria analyses. \emph{Ecography}, \textbf{32}, 205--209.

Cornu{é}jols, G., Nemhauser, G.L. \& Wolsey, L.A. (1990). The
uncapacitated facility location problem. \emph{Discrete Location Theory}
(eds P.B. Mirchandani \& R.L. Francis), pp. 119--171. Wiley, New York.

Crandall, K.A., Bininda-Emonds, O.R.P., Mace, G.M. \& Wayne, R.K.
(2000). Considering evolutionary processes in conservation biology.
\emph{Trends in Ecology \& Evolution}, \textbf{15}, 290--295.

Cui, T.T., Ouyang, Y.F. \& Shen, Z.J.M. (2010). Reliable facility
location design under the risk of disruptions. \emph{Operations
Research}, \textbf{58}, 998--1011.

Duong, T. (2015). \emph{ks: Kernel Smoothing, version 1.9.4}.

Faith, D.P. (2003). Environmental diversity (ED) as surrogate
information for species-level biodiversity. \emph{Ecography},
\textbf{26}, 374--379.

Faith, D.P. \& Walker, P.A. (1996). Environmental diversity: on the
best-possible use of surrogate data for assessing the relative
biodiversity of sets of areas. \emph{Biodiversity \& Conservation},
\textbf{5}, 399--415.

Ferrier, S. (2002). Mapping Spatial Pattern in Biodiversity for Regional
Conservation Planning: Where to from Here? \emph{Systematic Biology},
\textbf{51}, 331--363.

Gaston, K.J., Jackson, S.E., Cantu-Salazar, L. \& Cruz-Pinon, G.
(\{2008\}). The Ecological Performance of Protected Areas. \emph{ANNUAL
REVIEW OF ECOLOGY EVOLUTION AND SYSTEMATICS, Annual Review of Ecology
Evolution and Systematics}, \textbf{39}, 93--113.

Gower, J.C. (1971). A general coefficient of similarity and some of its
properties. \emph{Biometrics}, \textbf{27}, 857--871.

Hendry, A.P., Lohmann, L.G., Conti, E., Cracraft, J., Crandall, K.A.,
Faith, D.P., Hauser, C., Joly, C.A., Kogure, K., Larigauderie, A.,
Magallon, S., Moritz, C., Tillier, S., Zardoya, R., Prieur-Richard,
A.H., Walther, B.A., Yahara, T. \& Donoghue, M.J. (2010). Evolutionary
biology in biodiversity science, conservation, and policy: A call to
action. \emph{Evolution}, \textbf{64}, 1517--1528.

Hijmans, R.J., Cameron, S.E., Parra, J.L., Jones, P.G. \& Jarvis, A.
(2005). Very high resolution interpolated climate surfaces for global
land areas. \emph{International Journal of Climatology}, \textbf{25},
1965--1978.

Jones, K.E., Bielby, J., Cardillo, M., Fritz, S.A., O'Dell, J., Orme,
C.D.L., Safi, K., Sechrest, W., Boakes, E.H., Carbone, C., Connolly, C.,
Cutts, M.J., Foster, J.K., Grenyer, R., Habib, M., Plaster, C.A., Price,
S.A., Rigby, E.A., Rist, J., Teacher, A., Bininda-Emonds, O.R.P.,
Gittleman, J.L., Mace, G.M. \& Purvis, A. (2009). PanTHERIA: a
species-level database of life history, ecology, and geography of extant
and recently extinct mammals. \emph{Ecology}, \textbf{90}, 2648--2648.

Klein, C., Wilson, K., Watts, M., Stein, J., Berry, S., Carwardine, J.,
Smith, M.S., Mackey, B. \& Possingham, H. (2009). Incorporating
ecological and evolutionary processes into continental-scale
conservation planning. \emph{Ecological applications}, \textbf{19},
206--217.

Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M. \& Hornik, K.
(2015). \emph{cluster: Cluster Analysis Basics and Extensions}.

Margules, C.R. \& Pressey, R.L. (2000). Systematic conservation
planning. \emph{Nature}, \textbf{405}, 243--253.

McNeely, J.A. (1994). Protected areas for the 21st-century: Working to
provide benefits to society. \emph{Biodiversity and Conservation},
\textbf{3}, 390--405.

Meirmans, P., Goudet, J., IntraBioDiv Consortium \& Gaggiotti, O.
(2011). Ecology and life history affect different aspects of the
population structure of 27 high-alpine plants. \emph{Molecular Ecology},
\textbf{20}, 3144--3155.

Moilanen, A. (2007). Landscape Zonation, benefit functions and
target-based planning: unifying reserve selection strategies.
\emph{Biological Conservation}, \textbf{134}, 571--579.

Moritz, C. (1999). Conservation units and translocations: strategies for
conserving evolutionary processes. \emph{Hereditas}, \textbf{130},
217--228.

Moritz, C. (1994). Defining evolutionarily significant units for
conservation. \emph{Trends in Ecology \& Evolution}, \textbf{9},
373--375.

Moritz, C. (2002). Strategies to protect biological diversity and the
evolutionary processes that sustain it. \emph{Systematic Biology},
\textbf{51}, 238--254.

Oksanen, J., Blanchet, F.G., Kindt, R., Legendre, P., Minchin, P.R.,
O'Hara, R.B., Simpson, G.L., Solymos, P., Stevens, M.H.H. \& Wagner, H.
(2015). \emph{vegan: community ecology package}.

Oliver, M.A. \& Webster, R. (1990). Kriging: a method of interpolation
for geographical information systems. \emph{International Journal of
Geographical Information Systems}, \textbf{4}, 313--332.

Paradis, E., Claude, J. \& Strimmer, K. (2004). APE: analyses of
phylogenetics and evolution in R language. \emph{Bioinformatics},
\textbf{20}, 289--290.

Pyke, C.R. \& Fischer, D.T. (2005). Selection of bioclimatically
representative biological reserve systems under climate change.
\emph{Biological Conservation}, \textbf{121}, 429--441.

{R Core Team}. (2016). R: A language and environment for statistical
computing.

Raymond, B., VanDerWal, J. \& Belbin, L. (2015). \emph{ALA4R: Atlas of
Living Australia (ALA) Data and Resources in R}.

Sanderson, E.W., Segan, D.B. \& Watson, J.E.M. (2015). Global status of
and prospects for protection of terrestrial geophysical diversity.
\emph{Conservation Biology}, \textbf{29}, 649--656.

Thomassen, H.A., Buermann, W., Milá, B., Graham, C.H., Cameron, S.E.,
Schneider, C.J., Pollinger, J.P., Saatchi, S., Wayne, R.K. \& Smith,
T.B. (2010). Modeling environmentally associated morphological and
genetic variation in a rainforest bird, and its application to
conservation prioritization. \emph{Evolutionary Applications},
\textbf{3}, 1--16.

\end{document}

### 2016-07-18 article/figures.tex
+ assorted ms fixes including Rich's edits + results analysed + change to raptr

--- 2016-07-18 article/article.tex
@@
             pdfauthor={},
-            pdftitle={rapr: Representative and Adequate Prioritisations in R},
+            pdftitle={},
             colorlinks=true,
@@
 \setlength{\droptitle}{-2em}
-  \title{rapr: Representative and Adequate Prioritisations in R}
-  \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^1$,
-Richard A. Fuller$^1$\\$^1$School of Biological Sciences, The University
-of Queensland, Brisbane, QLD, Australia\\$^2$School of Geography,
-Planning and Environmental Management, The University of Queensland,
-Brisbane, QLD, Australia\\Correspondance should be addressed to
-\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
-  \preauthor{\centering\large\emph}
-  \postauthor{\par}
-  \predate{\centering\large\emph}
-  \postdate{\par}
-  \date{18 July 2016}
+  \title{}
+  \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
 
@@
 %  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
+% insert S before figure and table caption
+\pagenumbering{gobble}
 
@@
 
-\begin{abstract}
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  An underlying aim in spatial conservation planning is to maximise the
-  long-term persistence of biodiversity. Yet current approaches
-  overwhelming focus on ``adequacy'', the concept ensuring a target
-  proportion of the distribution of a biodiversity feature (species or
-  ecosystem) is reserved. Ultimately, to achieve long-term persistence,
-  the processes the sustain biodiversity must also be conserved. Many
-  biodiversity processes can be preserved by conserving a representative
-  sample of each biodiversity feature. This idea has for decades been
-  encapsulated in principle of ``representativeness''. For example,
-  representing the full range of genetic diversity or ability to persist
-  in a range of climatic conditions might be crucial for the long -term
-  persistence of a species in a world where environmental change is
-  accelerating. However, fully operationalising the principle of
-  repressiveness in spatial conservation planning has proven
-  methodologically challenging.
-\item
-  To address this issue, we developed the
-  \texttt{rapr: Representative and Adequate Pioritisations in R}, a
-  toolkit to guide reserve selection using explicit targets for
-  representing biodiversity processes as well as meeting area protection
-  targets. Using a novel formulation of the reserve selection problem,
-  users set ``space targets'' to secure a representative sample of each
-  species within an attribute space reflecting any biodiversity process
-  (eg. an attribute space expressing variation in genetic
-  characteristics between individuals in an area, or variation in
-  climatic conditions between areas occupied by indivduals). We explored
-  the functionality of this \texttt{R} package using a simulation study
-  and two case studies. In each study, we generated prioritisations that
-  aimed to preserve an amount of habitat for each species using amount
-  target--to represent conventional reserve selection methods--and
-  compared them with prioritisations generated using both habitat and
-  space targets--to represent this method.
-\item
-  We show that explicitly considering biodiversity processes in reserve
-  selection can substantially change the configuration of the resulting
-  solution. Simulations suggest that including representativeness in
-  conservation planning is particularly important where the biodiversity
-  feature has multi-model distributions in an attribute space. Results
-  from initial case studies show that explicitly considering space-based
-  targets for biodiversity processes can result in prioritisations that
-  do not necessarily require much more area than traditional approaches,
-  but that could be much more effective in achieving long term
-  biodiversity persistence.
-\item
-  The \texttt{rapr R} package provides a unified framework achieving
-  spatial conservation prioritisations that are both adequate and
-  representative, resulting in a greater chance of long term
-  biodiversity persistence that area-based planning alone.
-\end{enumerate}
-\end{abstract}
+\subsection{Figures}\label{figures}
 
-Short running title: Representative and adequate prioritisations
-\newline
-Abstract word count: 370 / 350\\Total word count: XXXXX / 8000\\Number
-of references: 34\\Number of figures: 8\\Number of tables: 0\\Number of
-textboxes: 0\\Number of equations: 8\\\newpage
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{Example of an attribute space. This environmental attribute
+space has dimensions relating to annual temperature ($^{\circ}$C) and
+rainfall (mm) values. Letters denote the environmental conditions
+associated with the geographic locations where four hypothetical
+populations are found. Points represent demand points. In this space,
+populations closer to each other are considered more similar to each
+other.}
+\end{figure}
 
-\section{Introduction}\label{introduction}
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{Simulated distributions of three species. Squares denote
+planning units. The color of each planning unit indicates the
+probability that each species inhabits it. Panel (a) shows the uniformly
+distributed species, panel (b) shows the normally distributed species,
+and panel (c) shows the bimodally distributed species.}
+\end{figure}
 
-Perhaps the most fundamental aim of conservation is to maximise the
-long-term persistence of biodiversity at organisational levels (McNeely
-1994; Margules \& Pressey 2000). To achieve this, conservation actions
-must preserve biodiversity patterns (eg. populations, species,
-ecosystems), but also crucially the processes that sustain them. One of
-the major tangible achievements of modern conservation has been the act
-of setting aside areas for preservation (Sanderson \emph{et al.} 2015).
-Reserve networks buffer species from gross threatening processes and set
-the stage for enhanced management interventions (Gaston \emph{et al.}
-\{2008\}). However, the resources available for conservation action are
-limited, and so reserve networks must be sited in places that satisfy
-conservation objectives for minimum cost (Margules \& Pressey 2000). To
-achieve this, reserve selection is often formulated as an optimisation
-problem and then solved to identify cost-effective candidate reserve
-systems (termed prioritisations; Margules \& Pressey 2000). Many
-decision support tools have been developed to help identify
-cost-effective solutions (eg. \texttt{ ConsNet}, Ciarleglio \emph{et
-al.} 2009;\texttt{ Marxan}, Ball \emph{et al.} 2009;\texttt{ Zonation},
-Moilanen 2007). Typically, these tools are used to deliver a
-prioritisation that preserves a set of species of interest to the
-conservation planner (termed target species or features). Given a set of
-species, decision makers can use these tools to preserve biodiversity
-patterns by generating a prioritisation that secures an adequate
-proportion of each species' range.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{Simulation study prioritisations. Each panel shows a
+prioritisation generated for a single species using a set of parameters.
+Squares denote planning units, with dark green showing those planning
+units selected for protection in a particular prioritisation. The top
+row of panels (a, d, g, j) shows prioritisations generated for the
+uniformly distributed species, middle row of panels (b, e, h, k) for the
+normally distributed species, and the bottom row of panels (c, f, i, l)
+for the bimodally distributed species. Each column of panels corresponds
+to a different set of parameters used to generate the prioritisation.
+The left column of panels (a--c) shows prioritizations generated using
+only 20 \% amount targets, the middle-left column of panels (d--f) for
+prioritisations generated using 20 \% amount and 75 \% space targets,
+the middle-right column of panels (g--i) for prioritisations generated
+using 20 \% amount based targets and a boundary length modifier (BLM) of
+1, and the right column of panels (j--l) for prioritisations generated
+using 20 \% amount and 75 \% space targets and also a boundary length
+modifier of 1.}
+\end{figure}
 
-However, to achieve effective conservation, reserve networks must do
-more than cover enough area. They must capture the processes that
-sustain biodiversity over the long term; they must capture both
-ecological and evolutionary processes (Crandall \emph{et al.} 2000;
-Margules \& Pressey 2000). Ecological processes, such as predator-prey
-interactions, pollination, and decomposition, are required for
-biodiversity to persist over short time-scales. Typically, they operate
-over small geographic domains--with exceptions such as migration and
-refugial habitats--and can be preserved using suitably large planning
-units (Ciarleglio \emph{et al.} 2009) that each contain a discrete unit
-of habitat (Klein \emph{et al.} 2009). On the other hand, evolutionary
-processes are required for biodiversity to persist over long
-time-scales, and they typically operate over large geographic domains.
-Protected areas must represent adaptive evolutionary processes to foster
-resilience against environmental change (eg. climate change; Pyke \&
-Fischer 2005). Protected areas must also preserve neutral evolutionary
-processes (Moritz 1994, 1999, 2002), arising from restriction of gene
-flow between populations. They are important for maintaining genetic
-diversity, and avoiding inbreeding depression. In recent decades, a
-wealth of data relevant to evolutionary processes has become available
-to conservation planners (eg. bioclimatic, genetic, and trait data;
-Hijmans \emph{et al.} 2005; Raymond \emph{et al.} 2015; Jones \emph{et
-al.} 2009). Yet such data are rarely used to guide conservation planning
-(Hendry \emph{et al.} 2010). This is perhaps in large part due to that
-fact that existing reserve selection tools focus on capturing either
-biodiversity patterns or processes--but not both.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{Distribution of environmental variation across Queensland,
+Australia. Squares denote planning units. The panels show the first and
+second main gradients of climatic variation.}
+\end{figure}
 
-Existing conservation planning tools are generally not well-suited for
-representing biodiversity processes (but see Faith 2003). To preserve
-biodiversity processes, a prioritisation must capture a representative
-sample of each species. For instance, to preserve predator-prey
-interactions, a prioritisation must preserve individuals from each
-predator and prey species in the same area. To preserve adaptive
-evolutionary processes, a prioritisation must preserve the adaptive
-landscape of each species--populations experiencing different selection
-pressures (Moritz 2002). To preserve neutral evolutionary processes, a
-prioritisation must secure individuals descended from each of the
-genetic lineages that comprise each species (Moritz 1994). Previous
-attempts to accommodate such into multi-species conservation planning
-have partitioned species' ranges into different groups, for example at
-habitat discontinuities (e.g. Carvalho \emph{et al.} 2011) as a
-pre-processing step to render a new set of ``pseudo-species''. However,
-this approach assumes that the biodiversity processes that operate
-across a species' range can be readily split into discrete units. Yet
-data on biodiversity processes is often continuous and
-hyper-dimensional, and often cannot be reduced to a few categories
-without significant information loss (Faith \& Walker 1996).
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{Distribution of the species used in the first case-study.
+Squares denote planning units. Each panel shows the distribution of a
+different species.}
+\end{figure}
 
-One of the key issues in reserve selection is the lack of a unifying
-decision support tool that can accommodate data on biodiversity patterns
-and processes in a multi-species context. To begin to fill this gap, we
-present the \texttt{rapr R} package. This \texttt{R} package uses novel
-formulations of the reserve selection problem to provide decision makers
-with the tools to generate prioritisations that preserve both
-biodiversity patterns and processes. We aim to define the concepts
-behind the problem formulations. Furthermore, we aim to explore the
-functionality of the \texttt{rapr R} package by applying one of the
-formulations to a set of simulated species and two case studies.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{Prioritisations for the niche-based case-study in Queensland,
+Australia. Squares represent planning units. Dark green planning units
+for selected for protection and pale green units were discarded. Panel
+(a) shows the planning units selected when using 20 \% amount targets.
+Panel (b) shows the planning units selected selected when using 20 \%
+amount targets and 75 \% niche targets}
+\end{figure}
 
-\section{Methods}\label{methods}
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-9-1.pdf}
+\caption{Study area and data used in the second case-study. Squares
+denote planning units. Panel (a) shows all grid cells surveyed by the
+IntraBioDiv project. Grid cells occupied by Betony-leaved Rampion are
+shown in bright blue. The subsequent panels contain only the grid cells
+that were occupied by the species. Panel (b) shows the opportunity cost
+of each planning unit (estimated as the total human population density).
+Panels (c--d) show the spatial distribution of the ordinated genetic
+data. These values describe the typical genetic characteristics of
+individuals in each planning unit. Planning units with similar
+values/colors contain individuals with similar loci polymorphisms.}
+\end{figure}
 
-\subsection{PROBLEM FORMULATION}\label{problem-formulation}
-
-Biodiversity features are defined as the entities that the
-prioritisation is required to preserve (eg. species, ecosystems).
-Spatial attributes are defined as the variation across the species'
-range that the prioritisation is required to sample. They can be
-intrinsic (eg. genetic or trait variation) or extrinsic (eg.
-environmental variation) to the feature. These attributes should be
-related to the biodiversity processes that the decision maker aims to
-preserve.
-
-Each attribute is conceptualised as a space. This space is termed an
-``attribute space''. Each planning unit is thought to occupy a single
-point inside each space. For example, a decision maker may require a
-prioritisation that represents populations along climatic gradients. To
-achieve this, the decision maker might use an ``climatic'' attribute
-space with dimensions relating to mean annual temperature ($^{\circ}$C)
-and precipitation (mm). Any given combination of temperature and
-precipitation may be conceived as a point in this environmental space.
-By associating planning units with climatic data, they can be mapped
-from geographic space to this environmental attribute space.
-
-Demand points are points that also exist in an attribute space. They are
-designated by the decision maker to indicate regions of the attribute
-space that the decision maker wishes to represent in the prioritisation
-(see below for discussion on how demand points can be generated for
-real-world datasets). The amount of variation in the attribute space
-that a prioritisation secures is a function of the distance between each
-demand point and each selected planning unit in the attribute space. The
-shorter the distances between the demand points and the planning units;
-the better the prioritisation is at securing the variation in the
-spatial attribute. To convert these amounts to a proportion--a
-meaningful unit for a decision maker--the distances between the selected
-planing units and the demand points are scaled by the distances between
-the demand points to the centroid of the demand points. In any attribute
-space there may exist points that are impossible (eg. mean annual
-rainfall -5 mm), or do not occur in the study area (eg. mean annual
-temperature 30$^{\circ}$C in Antarctica). Additionally, there may be
-some regions that are desirable for some features and undesirable for
-others (eg. conditions known to be outside the physiological tolerance
-of certain species). Thus a different set of demand points and weights
-are used for each attribute space and each feature. By placing demand
-points in desirable regions of an attribute space for a given feature,
-the decision maker can ensure that prioritisations secure the feature in
-planning units with spatial attributes that are desirable for that
-feature.
-
-To illustrate these concepts, consider an example conservation planning
-scenario example involving an attribute space and demand points. We wish
-to develop a prioritisation for a single species that has four
-populations in the study area. However, we can only afford to preserve
-three populations. We aim preserve the adaptive landscape of the
-species, and to achieve this, we will preserve populations inhabiting
-different environmental conditions. To describe environmental variation,
-we obtain data on the environmental conditions (rainfall (mm) and
-temperature ($^{\circ}$C)) where each population is found. These
-environmental data are then used to construct a two-dimensional
-environmental attribute space. Next, we generated demand points as
-equi-distant points between the range of values where the populations
-were found. By comparing the distribution of the demand points to the
-distribution of the populations in the attribute space, we can identify
-a suitable prioritisation (Fig. 1). We can see that preserving both
-populations $A$ and $C$, effectively ``doubles-up'' on the same
-environmental characteristics, resulting in considerable redundancy in
-the solution. Instead, a more representative sample of the
-intra-specific variation could be captured by securing populations $A$,
-$B$, and $D$. This example demonstrates how the inclusion of
-biodiversity processes can guide the reserve selection process.
-
-The formulations used to express the reserve selection problem in the
-\texttt{rapr R} package are based on a combination of the
-\texttt{Marxan} reserve selection problem and the uncapacitated facility
-location problems (Cornu{é}jols \emph{et al.} 1990). Although the
-\texttt{rapr R} package provides two novel formulations, for brevity, we
-will define the simpler formulation--referred to as the unreliable
-formulation--below and define the more complex version--the reliable
-formulation--in the Supporting Information S1. These formulations are
-named after the unreliable and reliable facility location formulations
-from which they are based upon (Cui \emph{et al.} 2010). The key
-difference between these two formulations is that the reliable
-formulation explicitly considers the probability that planning units are
-occupied when calculating the proportion of an attribute space sampled
-in a solution. All mathematical terms defined hereafter are described in
-Table S1. For convenience, the cardinality of sets will be denoted using
-the same symbol used to denote the variable.
-
-Define $F$ to be the set of features one wishes to conserve (indexed by
-$f$). Let $J$ be a set of planning units (indexed by $j$). Also, let
-$A_j$ denote the area, and $C_j$ denote the cost of preserving planning
-unit $j \in J$. To assess the extent to which each feature is secured in
-a given prioritisation, let $q_{fj}$ denote the probability of feature
-$f$ occupying planning unit $j$. The level of fragmentation associated
-with a prioritisation is parametrised as the net exposed boundary
-length. Let the shared edges between each planning unit $j \in J$ and
-$k \in J$ be $e_{jk}$.
-
-Let $S$ denote a set of attribute spaces (indexed by $s$). Each
-$j \in J$ is associated with spatially explicit data that represent
-coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a
-set of demand points (indexed by $i$) for each feature $f \in F$ and
-each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting
-for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$
-denote the distance between each demand point $i \in I$ and each
-planning unit $j \in J$ for each feature $f \in F$ and attribute space
-$s \in S$. To describe the inherent variation in the distribution of
-demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote
-the distance between each demand point $i \in I$ and the centroid of the
-demand points. Demand points with greater weight $\lambda_{fsi}$ are
-more important, and the optimal solution will be likely to select
-planning units close to highly weighted demand points. As a consequence,
-the decision maker will need to choose an appropriate weighting for each
-demand point.
-
-Targets are used to ensure that prioritisations adequately preserve each
-species. Amount-based targets are used to ensure that the total amount
-of habitat preserved is sufficient. Let $T_f$ denote the expected amount
-of area that needs to be preserved for each feature $f \in F$.
-Space-based targets ensure that a sufficient proportion of the
-intra-specific variation is secured. Let $\tau_{fs}$ denote the
-space-based targets for feature $f \in F$ and attribute space $a \in A$.
-For convenience, these both types of targets are expressed as
-proportions in the \texttt{R} package.
-
-The control variables for the unreliable formulation are the $B$,
-$T_{s}$, and $\tau_{sa}$ variables.
-
-\begin{align*}
-T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
-%
-\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
-%
-B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
-\end{align*}
-
-The decision variables are the $X_j$ and $Y_{fsij}$ variables.
-
-\begin{align*}
-X_j
-    &= \begin{cases}
-        1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
-        0, & \parbox{25em}{otherwise} \\
-    \end{cases} \\
-%
-Y_{fsij} &= \begin{cases}
-        1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
-        0, & \parbox{25em}{otherwise} \\
-    \end{cases} \\
-\end{align*}
-
-Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is
-assigned to a selected planning unit $J$ where $X_j = 1$. The weighted
-distance between the demand point and its assigned planning unit
-$\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is
-represented in a given solution. Generally, demand points are assigned
-to the closest selected planning units (unless particularly low
-space-based targets are used).
-
-The unreliable formulation (URAP) is a defined as a multi-objective
-optimisation problem.
-
-\begin{align*}
-& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
-%
-& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
-%
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
-& & & & & 0 \leq s \leq S-1\\
-%
-& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
-& & & & & 0 \leq s \leq S-1, \\
-& & & & & 0 \leq i \leq I-1\\
-%
-& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
-& & & & & 0 \leq s \leq S-1, \\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J-1\\
-%
-& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1\\
-%
-\end{align*}
-
-The objective function (eqn 3a) determines the utility of a given
-prioritisation: a combination of the total cost of a prioritisation and
-how fragmented it is. Constraints (eqn 3b) ensure that all the
-amount-based targets are met. Constraints (eqn 3c) ensure that all the
-space-based targets are met for each feature and each attribute space.
-For each feature and attribute space, the total weighted distance
-between the demand points and their closest selected planning units is
-calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted
-distance is then scaled by the inherent variation in the demand points
-($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a
-proportion conceptually similar to the $R^2$ statistic used in $k$-means
-clustering analysis. The constraints ensure that this proportion must be
-equal to or greater than the space-based target. Constraints (eqn 3d)
-ensure that only one planning unit is assigned to each demand point.
-Constraints (eqn 3e) ensure that demand points are only assigned to
-selected planning units. Constraints (eqn 3f) ensure that the $X$ and
-$Y$ variables are binary.
-
-\subsection{OPTIMISATION}\label{optimisation}
-
-The unreliable formulation is non-linear. However, the non-linear
-components can be linearised using existing techniques. The expression
-$X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer
-et al. (2016). Linearised versions of the problems can be solved using
-commercial exact algorithm solvers. The \texttt{rapr} R package provides
-functions to express conservation planning data as optimisation problems
-using linearised versions of the unreliable and reliable formulations.
-These optimisation problems can then be solved to generate
-prioritisations using the commercial \texttt{Gurobi} software suite
-(\url{http://www.gurobi.com}). Presently, academics can obtain a license
-at no cost from the Gurobi website. After installing the \texttt{Gurobi}
-software suite, users will need to install the \texttt{Gurobi R}
-package.
-
-\section{Examples}\label{examples}
-
-To understand the behaviour of the unreliable problem and showcase its
-value, we conducted a simulation study and two case studies. These
-studies involved generating solutions using only amount targets to
-represent prioritisations generated using conventional methods (eg.
-\texttt{Marxan}), and solutions using amount and space targets using the
-unreliable formulation. By comparing these solutions, we can guage the
-benefits of explicitly including space targets in reserve selection. All
-analyses were performed in \texttt{R} (version 3.3.0; {R Core Team}
-2016).
-
-\subsection{SIMULATION STUDY}\label{simulation-study}
-
-\subsubsection{Methods}\label{methods-1}
-
-We simulated a hypothetical study area with planning units arranged in
-$10 \times 10$ square grid (Fig. 2). We then simulated three species
-across this study area. The first species was simulated to represent a
-hyper-generalist--occurring in all planning units with equal probability
-(Fig. 2a). This species' distribution was based on a uniform
-distribution (eqn 8a). The second species was simulated to represent a
-species with an idealised distribution (Fig. 2b). It has a core range
-area, and is less likely to be found in areas that are distant from the
-core area. This species' distribution was simulated using the
-probability density function of a single multivariate normal
-distribution (eqn 8c). The third species represents a species with two
-distinct populations (Fig. 2c). This species' distribution was simulated
-to depict a bimodal distribution, based on the combination of the
-probability density functions of two multivariate normal distributions
-(eqn 8d). The probability that each species inhabited a given planning
-unit was calculated using the $X, Y$ coordinates of the planning unit
-and eqns 8a--8d. We used a geographic attribute space to showcase the
-behavior of the problem. For each species, demand points were generated
-by calculating the centroids of each planning units and their weights
-were set as the probability that the planning units were occupied.
-
-\begin{align*}
-\\
-P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
-%
-f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
-%
-P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
-%
-P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
-f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
-\end{cases} \tag*{eqn 8d} \\
-%
-\end{align*}
-
-We generated four solutions for each species. First to represent
-solutions generated using conventional conservation planning methods, we
-generated solutions using only 20 \% amount targets. Second to show how
-the addition of geographic targets can affect a prioritisation, we
-generated solutions using 20 \% amount targets and 75 \% geographic
-targets. Third to represent solutions using conventional planning
-methods that penalise for fragmentation, we generated solutions using 20
-\% amount targets and a boundary length modifier of 1. Fourth to
-illustrate the combined affect using geographic targets and penalising
-fragmentation, we generated solutions using using 20 \% amount targets
-and 75 \% geographic targets and a boundary length modifier of 1.
-
-\subsubsection{Results}\label{results}
-
-The uniform species was simulated to occur with a constant probability
-of occupancy across the study area (Fig. 2a). The solution generated
-using 20 \% amount-based targets (Fig. 3a) selected 20 planning units
-near the southern end of study area. This configuration is an artifact
-of the method used to solve this particular instance of the reserve
-selection problem. In fact, for this species, all prioritisations
-containing 20 planning units are optimal when considering only 20 \%
-amount targets. In the absence of criteria to guarantee
-representativeness, reserve selection methods may or may not return
-solutions that secure a representative sample of the features. In this
-particular case, the solution does not secure a representative sample of
-the species' range (-23.64 \% sampled).
-
-The addition of a 75 \% geographic target resulted in a solution that
-secured a representative sample of the uniform species' range (90.67 \%
-sampled; Fig. 3b). Since all planning units have equal cost and an equal
-chance of being occupied, this solution has the same number of planning
-units as the solution generated using only amount targets (cf.~Fig 3a).
-Although the use of amount and space based targets has addressed
-adequacy and representativeness objectives (respectively), they have
-resulted in a highly fragmented solution.
-
-The addition of a positive boundary length modifier (BLM) parameter
-resulted in a well-connected solution that secured an adequate
-proportion (20 \% ) and representative sample (75.76 \% sampled) of the
-species geographic distribution (Fig. 3d). This solution contains 20
-planning units--a few more than the previously discussed solutions--to
-ensure that the solution secures a representative proportion of the
-species geographic distribution in a configuration that is not highly
-fragmented. This result suggests that the combination of space targets
-and boundary length modifiers may yield solutions that contain more
-planning units, since the solution generated using amount targets and a
-boundary length modifier (Fig. 3c) contained the sample number of
-planning units as the solution with just the amount targets (Fig. 3a).
-Overall, these results show that under the simplest of conditions, the
-reliance on just amount targets can cause an under-specified reserve
-selection problem that is unlikely to return a suitable solution for
-implementation.
-
-The normally distributed species was simulated to contain a single core
-area where individuals are most prevalent and marginal areas where
-individuals are less likely to occur (Fig. 2b). The solution generated
-using 20 \% amount targets contained 10 planning units and concentrated
-conservation efforts in the core area (Fig. 3e). Since all planning
-units have equal costs and areas, this solution contains the planning
-units with the highest probabilities of occupancy. Whilst this solution
-satisfies the adequacy objective for a prioritisation in a
-cost-effective manner--it fails to fulfill the representative objective
-for a prioritisation (59.09 \% distribution sampled).
-
-The solution generated using 20 \% amount targets and 75 \% geographic
-targets resulted in a solution that is both adequate and representative
-in terms of the uniform species distribution (Fig. 3f). This solution
-used 11 planning units to secure 20.02 \% and sample 76.31 \% of the
-normal species' distribution. By using amount and geographic targets in
-the reserve selection problem, we have obtained a solution that
-concentrates conservation effort in the range core and also the range
-margin. However, similar to the corresponding solution for the uniform
-species (Fig. 3b), this solution is highly fragmented.
-
-By using 20 \% amount targets and 75 \% geographic targets and a
-boundary length modifier parameter ($BLM = 1$), we obtained a solution
-that fulfills adequacy, representativeness and connectivity objectives
-(Fig. 3h). This solution contains 12 planning units. Similar to the
-corresponding solution for the uniform species (Fig. 3d), this solution
-contains more planning units than any other solutions for this species
-(cf.~Figs. 3e--3g). The results for the normally distributed species
-suggest that the space targets can result in solutions that secure a
-more representative sample of the species'--even for species without
-significant structure.
-
-The bimodally distributed species was simulated to represent a species
-with a highly structured distribution, and it contains two distinct
-populations (Fig. 2c). The solution generated using just 20 \% amount
-targets assigned conservation effort to just one of the two populations
-(Fig. 3i). While this solution secured an adequate proportion of the
-species' distribution (20.18 \% secured), it did not sample a
-representative proportion of the range (12 \% sampled). The addition of
-boundary length modifiers to just amount targets resulted in a solution
-which sampled even less of the species distribution (25.19 \% sampled;
-Fig. 3k). However, the addition of geographic space targets resolved
-these issues.
-
-The solution generated using 20 \% amount and 75 \% geographic targets
-included planning units from both populations (Fig. 3j). Although this
-solution required more planning units to fulfill both targets ($n=9$),
-this solution secured a representative sample of the bimodally
-distributed species (80.95 \% range sampled). Similar to corresponding
-solutions generated for both the uniformly (Fig. 3b) and normally
-distributed species (Fig. 3f), this solution was also highly fragmented
-and we could obtain a more well-connected solution at the expense of
-selecting additional planning units ($n=9$; Fig. 3l). The results for
-the bimodally distributed species suggest that species with highly
-structured populations or strong variation between individuals could
-benefit the most from prioritisations that are generated using
-space-based targets.
-
-\subsection{CASE STUDY 1}\label{case-study-1}
-
-\subsubsection{Methods}\label{methods-2}
-
-We investigated how space-based targets can be used in a multi-species
-planning context to generate a prioritisation that sufficiently
-preserves the realised niche for several species. By preserving the
-populations in suitable habitats with different with environmental
-conditions, conservation planners can preserve the species' adaptive
-landscape and foster resilience against environmental change (Moritz
-2002). We selected Queensland, Australia as the study area. This region
-is ideal for exploring the potential of niche-based targets because it
-contains a broad range of different habitats. We obtained data for 19
-bioclimatic variables across the region (at $30^{\prime \prime}$
-resolution from \url{www.worldclim.org}; Hijmans \emph{et al.} 2005) and
-subjected them to a principal components analysis (using ArcMap 10.3.1).
-We used the first two principal components (cumulatively explained 99.5
-\% of the total variation) to characterise the environmental variation
-across the study area (Fig. 4).
-
-We selected four bird species that span a range of different
-evolutionary histories, distributions and ecologies: blue-winged
-kookaburra (\emph{Dacelo leachii}), brown-backed honeyeater
-(\emph{Ramsayornis modestus}), brown falcon (\emph{Falco berigora}), and
-pale headed rosella (\emph{Platycercus adscitus}). We then mapped the
-extent of occurrence for each species (Fig. 5). To do this, we obtained
-observation data from the Atlas of Living Australia across the whole of
-Australia (using the \texttt{ALA4R R} package; Raymond \emph{et al.}
-2015), spatially thinned the data to omit points within 10 km$^2$ of
-each other to ameliorate the effects of sampling bias (using the
-developmental version of the \texttt{spThin R} package;
-\url{www.github.com/mlammens/spThin}; Aiello-Lammens \emph{et al.}
-2015), and fit 85 \% minimum convex polygons (using the
-\texttt{adehabitatHR R} package; Calenge 2006). We used this method
-because it is trivially reproducible using freely available data.
-
-We generated 500 demand points for each species (Fig. S2). To achieve
-this, for each species, we generated random points inside the species
-range and at each point extracted the principal component values at that
-location. We then fit hyperbox kernels to the distribution of these
-points to characterise the realised niche of each species (using a
-manually chosen bandwidth of 0.2 and a 0.5 quantile to map the core
-parts of the species' niches; implemented in the \texttt{hypervolume R}
-package; Blonder \emph{et al.} 2014). We then generated uniformly
-distributed points inside the species' kernels and estimated the density
-of the training points at the uniformly generated points. These
-uniformly distributed points and associated density estimated were used
-as demand point coordinates and weights (respectively).
-
-We generated two solutions. The first solution was generated using 20 \%
-amount-based targets for all species. The second solution was generated
-the same amount-based target with an additional 75 \% niche-based for
-each species.
-
-\subsubsection{Results}\label{results-1}
-
-Generally, the solution generated using amount targets preserved a
-representative sample of each the four bird species' niches--with one
-excpetion. Specifically, this solution sampled over 75 \% of the
-realised niche of blue-winged kookaburra (89.54 \%), brown falcon (92.31
-\%), and the pale-headed rosella (76.86 \%). Yet it failed to achieve
-this for Brown-backed honeyeater(30.93 \%). This result suggests that
-prioritisations generated using conventional methods may preserve a
-representative sample of most species realised niches (Figure S1).
-However, despite this, there may yet be species for which only a small
-fraction of their realised niche is preserved. By explictly using space
-targets, conservation planners can generate prioritisations that are
-guaranteed to capture a representative sample of species' niches.
-
-\subsection{CASE STUDY 2}\label{case-study-2}
-
-\subsubsection{Methods}\label{methods-3}
-
-Here we used space-based targets to generate a prioritisation securing a
-representative sample of a species' intra-specific genetic variation. We
-used species occurrence and multilocus AFLP data collected by the
-international IntraBioDiv project in the European Alps (Meirmans
-\emph{et al.} 2011; see Alvarez \emph{et al.} 2009 for further
-explanation of data collection methods). Althougthis dataset contains
-multiple plant species, we used data for the betony-leaved rampion
-(\textit{Phyteuma betonicifolium}), a wide-spread species with
-significant genetic structure (Aiello-Lammens \emph{et al.} 2015).
-Members of the IntraBioDiv project collected data using a $20^{\prime}$
-longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25
-km; Fig. 7a). They visited every second grid cell, and if the
-betony-leaved rampion was detected in a cell, samples were collected
-from three individuals. Samples were genotyped using amplified fragment
-length polymorphisms (AFLP; Raymond \emph{et al.} 2015), and used to
-construct matrices denoting the presence/absence of polymorphisms at
-loci. In total, 131 individuals were genotyped at 138 markers.
-
-We used non-metric multi-dimensional scaling (NMDS; using Gower
-distances to accommodate sparsity; Gower 1971; implemented the
-\texttt{cluster R} package; Maechler \emph{et al.} 2015) to ordinate the
-presence (or absence) of locus-specific alleles wthin individuals into
-two continuous variables (implemented in the \texttt{vegan R} package;
-Oksanen \emph{et al.} 2015). These continuous variables described the
-main axes of genetic variation within the species. We calculated the
-average of the values associated with individuals in each grid cell.
-These values were used to create a genetic attribute space (Figs. 7c and
-7d). To assess spatial auto-correlation, we calculated Moran's I
-auto-correlation index for each NMDS axis using inverse great circle
-distances based on the grid cells' centroids (using the \texttt{ape R}
-package; Paradis \emph{et al.} 2004).
-
-The grid cells were used as planning units. The grid-cell averaged
-ordinations were used to describe the typical genetic characteristics of
-individuals in the planning unit. Since the number of planning units was
-relatively small, we used the same grid-cell averaged ordinations as
-demand points. To ensure that the solutions did not prioritise
-particularly costly areas, we included opportunity cost data (Fig. 7b).
-We obtained population density data (1 km resolution from the Global
-Rural-Urban Mapping Project; \texttt{GRUMP V1}; CIESEN, Columbia
-University \emph{et al.} 2011) and estimated the total population
-density inside each grid cell.
-
-We generated two solutions. The first solution was generated using an 10
-\% amount-based target. The second solution was generated using the same
-amount-based target and an additional 95 \% genetic-based target.
-
-\subsubsection{Results}\label{results-2}
-
-The binary genetic data was ordinated into a two-dimensional space that
-described the main different between individuals (stress = 0.17). These
-values were then averaged to the planning unit level to describe the
-typical genetic characteristics of individuals in each planning unit
-(Figs. 7c--7d). Planning units located near each were found to contain
-individuals with similar genetic characteristics (Moran's I: NMDS axis
-1, $I = 0.4$, $P < 0.001$; Moran's I: NMDS axis 2, $I = 0.32$,
-$P < 0.001$). In terms of the average genetic characteristics of
-individuals in the planning units, they tended to cluster into two main
-groups, with evidence of within-group structure inside the larger group
-(Fig 8c). This analysis supports previous work by Alvarez \emph{et al.}
-(2009) who also found evidence of genetic structure within this species.
-
-The solution generated using just the amount target failed to preserve a
-representative portion of the species' genetic variation (15.54 \%
-sampled; Figs. 8a and 8c). This solution only sampled planning units
-that contained individuals belonging to one of the main two groups. On
-the other hand, the solution generated using additional genetic targets
-selected planning units belonging to both of the main groups. Note that
-since these solutions were generated using opportunity cost data, the
-solution generated using only the amount target is essentially the
-minimum number of least costly areas needed to fulfill the amount
-target. The only difference between the two solutions is a single
-planning unit. By swapping this single planning unit, the solution
-generated using amount and genetic targets was able to preserve a
-representative portion of the species' genetic variation 15.54 for only
-a minor increase in cost (99.41 total cost compared to 99.66 total
-cost).
-
-\section{Implications and future
-directions}\label{implications-and-future-directions}
-
-The \texttt{rapr} R package provides a unified approach to reserve
-selection. One of the key advantages of this package is that it is
-general enough to incorporate any spatial variable as an attribute
-space. This package can accommodate intrinsic or extrinsic variation to
-the feature(s). For example, adaptation processes could be secured using
-environmental variation or genetic variation among loci under selection.
-Ecological processes, such as predator-prey interactions, could be
-secured by capturing intra-specific trophic variation. Different
-attribute spaces with different targets can be also specified for
-different species. As long as the variation can be described using
-Euclidean distances, the \texttt{R} package can be used to obtain a
-representative sample (note that data may require transformation to
-conform to this assumption). This \textbackslash{}texttt\{R\} package
-provides decision makers with the tools to generate prioritisations that
-secure both biodiversity patterns and the processes that maintain
-biodiversity. Additionally, the package contains functionality to
-accommodate uncertainty in the distribution of features, and also
-identify suitably connected reserves. Both the simulated and case study
-species suggest that conservation planning exercises need to explicitly
-consider biodiversity processes during the reserve selection process to
-ensure they are captured.
-
-The degree to which a prioritisation truly secures a representative
-sample of a feature depends on the attribute spaces and distribution of
-demand points chosen by the decision maker. Ultimately, the space-based
-targets are set as a proportion based on the distribution of the
-demand-points. As a consequence, if the decision maker uses an
-inappropriate set of spatial variables to construct an attribute space,
-or an inappropriate set of demand points, then the optimal solution will
-not be an effective prioritisation. We therefore stress that decision
-makers must carefully consider which biodiversity processes need to be
-reflected in the prioritisation, and which spatial data can be used to
-represent these processes. Note that maximising one process can be
-detrimental to another. For instance, to maximise the geographic spread
-of a prioritisation, reserves need to be further away from each other,
-yet to maximise the connectivity of a prioritisation, reserves need to
-be closer to each other. To assist in the selection of appropriate
-demand points, the \texttt{R} package provides several routines for
-generating demand points (see the \texttt{make.DemandPoints} function).
-These routines essentially use the distribution of a feature in the
-attribute space to define a polygon. Demand points are then generated as
-random points within the polygon. A kernel is then fit to the
-distribution of the feature in the space (using Blonder \emph{et al.}
-2014; Duong 2015), and the demand points are weighted based on the
-estimated density of the feature at the demand points.
-
-The formulation requires spatially comprehensive data to map attribute
-spaces to planning units. For instance, in both case-studies, all
-planning units occupied by the species were associated with
-values/coordinates in the attribute spaces. However, most real-world
-data sets are patchy--some planning units will be occupied by species
-for which attribute space data is not available. To use such patchy data
-with the \textbackslash{}texttt\{rapr R\} package, the gaps in the data
-must first be filled in as a pre-processing step. Spatially explicit
-models could be used to estimate values in locations that are missing
-from a data set (eg. krigging or generalised dissimilarity models;
-Oliver \& Webster 1990; Ferrier 2002). This approach has been
-successfully applied to a range of biological data sets (Thomassen
-\emph{et al.} 2010).
-
-To maximise the long-term persistence of biodiversity, decision makers
-need to identify prioritisations that preserve existing patterns of
-biodiversity and the processes that support them. To achieve this,
-conservation planners need a decision support tool that can explicitly
-accommodate biodiversity patterns and processes. Here, we developed the
-\texttt{rapr R} package to fill this void. By exploring the
-functionality of this package using several simulated species, we found
-that including space-based targets can radically change a prioritisation
-for the simplest of species.
-
-\section{Acknowledgements}\label{acknowledgements}
-
-JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
-has an Australian Research Council Future Fellowship. This work was
-supported by the Centre of Excellence for Environmental Decisions (CEED)
-and the Landscape Ecology and Conservation Group (LEC) at The University
-of Queensland.
-
-\section{Data accessibility}\label{data-accessibility}
-
-All data, code, and results are stored in an online repository
-(\url{www.github.com/paleo13/rapr-manuscript}) to permit replication and
-validation of this study.
-
-\section{References}\label{references}
-
-\bibliography{references}
-
-Aiello-Lammens, M.E., Boria, R.A., Radosavljevic, A., Vilela, B. \&
-Anderson, R.P. (2015). spThin: an R package for spatial thinning of
-species occurrence records for use in ecological niche models.
-\emph{Ecography}, \textbf{38}, 541--545.
-
-Alvarez, N., Thiel-Egenter, C., Tribsch, A., Holderegger, R., Manel, S.,
-Sch{ö}nswetter, P., Taberlet, P., Brodbeck, S., Gaudeul, M., Gielly, L.,
-K{ü}pfer, P., Mansion, G., Negrini, R., Paun, O., Pellecchia, M., Rioux,
-D., Sch{ü}pfer, F., Van Loo, M., Winkler, M., Gugerli, F. \& IntraBioDiv
-Consortium. (2009). History or ecology? Substrate type as a major driver
-of patial genetic structure in Alpine plants. \emph{Ecology Letters},
-\textbf{12}, 632--640.
-
-Ball, I., Possingham, H. \& Watts, M.E. (2009). Marxan and relatives:
-software for spatial conservation prioritisation. \emph{Spatial
-conservation prioritisation: Quantitative methods \& computational
-tools} (eds A. Moilanen, K.A. Wilson \& H. Possingham), pp. 185--189.
-Oxford University Press, Oxford, UK.
-
-Beyer, H.L., Dujardin, Y., Watts, M.E. \& Possingham, H.P. (2016).
-Solving conservation planning problems with integer linear programming.
-\emph{Ecological Modelling}, \textbf{328}, 14--22.
-
-Blonder, B., Lamanna, C., Violle, C. \& Enquist, B.J. (2014). The
-\emph{n}-dimensional hypervolume. \emph{Global Ecology and
-Biogeography}, \textbf{23}, 595--609.
-
-CIESEN, Columbia University, International Food Policy Research
-Institute, The World Bank \& Centro Internacional de Agricultura
-Tropical. (2011). Global rural-urban mapping project, Version 1 (GRUMP
-v1): Urban extents grid.
-
-Calenge, C. (2006). The package adehabitat for the R software: tool for
-the analysis of space and habitat use by animals. \emph{Ecological
-Modelling}, \textbf{197}, 1035.
-
-Carvalho, S.B., Brito, J.C., Crespo, E.J. \& Possingham, H.P. (2011).
-Incorporating evolutionary processes into conservation planning using
-species distribution data: a case study with the western Mediterranean
-herpetofauna. \emph{Diversity \& Distributions}, \textbf{17}, 408--421.
-
-Ciarleglio, M., Wesley Barnes, J. \& Sarkar, S. (2009). ConsNet: new
-software for the selection of conservation area networks with spatial
-and multi-criteria analyses. \emph{Ecography}, \textbf{32}, 205--209.
-
-Cornu{é}jols, G., Nemhauser, G.L. \& Wolsey, L.A. (1990). The
-uncapacitated facility location problem. \emph{Discrete Location Theory}
-(eds P.B. Mirchandani \& R.L. Francis), pp. 119--171. Wiley, New York.
-
-Crandall, K.A., Bininda-Emonds, O.R.P., Mace, G.M. \& Wayne, R.K.
-(2000). Considering evolutionary processes in conservation biology.
-\emph{Trends in Ecology \& Evolution}, \textbf{15}, 290--295.
-
-Cui, T.T., Ouyang, Y.F. \& Shen, Z.J.M. (2010). Reliable facility
-location design under the risk of disruptions. \emph{Operations
-Research}, \textbf{58}, 998--1011.
-
-Duong, T. (2015). \emph{ks: Kernel Smoothing, version 1.9.4}.
-
-Faith, D.P. (2003). Environmental diversity (ED) as surrogate
-information for species-level biodiversity. \emph{Ecography},
-\textbf{26}, 374--379.
-
-Faith, D.P. \& Walker, P.A. (1996). Environmental diversity: on the
-best-possible use of surrogate data for assessing the relative
-biodiversity of sets of areas. \emph{Biodiversity \& Conservation},
-\textbf{5}, 399--415.
-
-Ferrier, S. (2002). Mapping Spatial Pattern in Biodiversity for Regional
-Conservation Planning: Where to from Here? \emph{Systematic Biology},
-\textbf{51}, 331--363.
-
-Gaston, K.J., Jackson, S.E., Cantu-Salazar, L. \& Cruz-Pinon, G.
-(\{2008\}). The Ecological Performance of Protected Areas. \emph{ANNUAL
-REVIEW OF ECOLOGY EVOLUTION AND SYSTEMATICS, Annual Review of Ecology
-Evolution and Systematics}, \textbf{39}, 93--113.
-
-Gower, J.C. (1971). A general coefficient of similarity and some of its
-properties. \emph{Biometrics}, \textbf{27}, 857--871.
-
-Hendry, A.P., Lohmann, L.G., Conti, E., Cracraft, J., Crandall, K.A.,
-Faith, D.P., Hauser, C., Joly, C.A., Kogure, K., Larigauderie, A.,
-Magallon, S., Moritz, C., Tillier, S., Zardoya, R., Prieur-Richard,
-A.H., Walther, B.A., Yahara, T. \& Donoghue, M.J. (2010). Evolutionary
-biology in biodiversity science, conservation, and policy: A call to
-action. \emph{Evolution}, \textbf{64}, 1517--1528.
-
-Hijmans, R.J., Cameron, S.E., Parra, J.L., Jones, P.G. \& Jarvis, A.
-(2005). Very high resolution interpolated climate surfaces for global
-land areas. \emph{International Journal of Climatology}, \textbf{25},
-1965--1978.
-
-Jones, K.E., Bielby, J., Cardillo, M., Fritz, S.A., O'Dell, J., Orme,
-C.D.L., Safi, K., Sechrest, W., Boakes, E.H., Carbone, C., Connolly, C.,
-Cutts, M.J., Foster, J.K., Grenyer, R., Habib, M., Plaster, C.A., Price,
-S.A., Rigby, E.A., Rist, J., Teacher, A., Bininda-Emonds, O.R.P.,
-Gittleman, J.L., Mace, G.M. \& Purvis, A. (2009). PanTHERIA: a
-species-level database of life history, ecology, and geography of extant
-and recently extinct mammals. \emph{Ecology}, \textbf{90}, 2648--2648.
-
-Klein, C., Wilson, K., Watts, M., Stein, J., Berry, S., Carwardine, J.,
-Smith, M.S., Mackey, B. \& Possingham, H. (2009). Incorporating
-ecological and evolutionary processes into continental-scale
-conservation planning. \emph{Ecological applications}, \textbf{19},
-206--217.
-
-Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M. \& Hornik, K.
-(2015). \emph{cluster: Cluster Analysis Basics and Extensions}.
-
-Margules, C.R. \& Pressey, R.L. (2000). Systematic conservation
-planning. \emph{Nature}, \textbf{405}, 243--253.
-
-McNeely, J.A. (1994). Protected areas for the 21st-century: Working to
-provide benefits to society. \emph{Biodiversity and Conservation},
-\textbf{3}, 390--405.
-
-Meirmans, P., Goudet, J., IntraBioDiv Consortium \& Gaggiotti, O.
-(2011). Ecology and life history affect different aspects of the
-population structure of 27 high-alpine plants. \emph{Molecular Ecology},
-\textbf{20}, 3144--3155.
-
-Moilanen, A. (2007). Landscape Zonation, benefit functions and
-target-based planning: unifying reserve selection strategies.
-\emph{Biological Conservation}, \textbf{134}, 571--579.
-
-Moritz, C. (1999). Conservation units and translocations: strategies for
-conserving evolutionary processes. \emph{Hereditas}, \textbf{130},
-217--228.
-
-Moritz, C. (1994). Defining evolutionarily significant units for
-conservation. \emph{Trends in Ecology \& Evolution}, \textbf{9},
-373--375.
-
-Moritz, C. (2002). Strategies to protect biological diversity and the
-evolutionary processes that sustain it. \emph{Systematic Biology},
-\textbf{51}, 238--254.
-
-Oksanen, J., Blanchet, F.G., Kindt, R., Legendre, P., Minchin, P.R.,
-O'Hara, R.B., Simpson, G.L., Solymos, P., Stevens, M.H.H. \& Wagner, H.
-(2015). \emph{vegan: community ecology package}.
-
-Oliver, M.A. \& Webster, R. (1990). Kriging: a method of interpolation
-for geographical information systems. \emph{International Journal of
-Geographical Information Systems}, \textbf{4}, 313--332.
-
-Paradis, E., Claude, J. \& Strimmer, K. (2004). APE: analyses of
-phylogenetics and evolution in R language. \emph{Bioinformatics},
-\textbf{20}, 289--290.
-
-Pyke, C.R. \& Fischer, D.T. (2005). Selection of bioclimatically
-representative biological reserve systems under climate change.
-\emph{Biological Conservation}, \textbf{121}, 429--441.
-
-{R Core Team}. (2016). R: A language and environment for statistical
-computing.
-
-Raymond, B., VanDerWal, J. \& Belbin, L. (2015). \emph{ALA4R: Atlas of
-Living Australia (ALA) Data and Resources in R}.
-
-Sanderson, E.W., Segan, D.B. \& Watson, J.E.M. (2015). Global status of
-and prospects for protection of terrestrial geophysical diversity.
-\emph{Conservation Biology}, \textbf{29}, 649--656.
-
-Thomassen, H.A., Buermann, W., Milá, B., Graham, C.H., Cameron, S.E.,
-Schneider, C.J., Pollinger, J.P., Saatchi, S., Wayne, R.K. \& Smith,
-T.B. (2010). Modeling environmentally associated morphological and
-genetic variation in a rainforest bird, and its application to
-conservation prioritization. \emph{Evolutionary Applications},
-\textbf{3}, 1--16.
+\begin{figure}[htbp]
+\centering
+\includegraphics{figures_files/figure-latex/unnamed-chunk-10-1.pdf}
+\caption{Prioritisations for the genetic case-study in the European
+Alps. Panels (a--b) show prioritisations generated using different
+parameters. Squares denote planning units shown in geographic space.
+Panel (a) shows the planning units selected when using 10 \% amount
+targets. Panel (b) shows the planning units selected when using 10 \%
+amount targets and 95 \% genetic targets. Panels (c--d) show the
+solutions in the genetic space. Each point corresponds to a planning
+unit. The coordinates of the points represent the typical genetic
+characeristics of individuals sampled in that planning unit (based on an
+NMDS of the binary loci data). Planning units associated with points
+that are closer togeather contain individuals with more similar genetic
+characteristics than planning units that are further apart. In all
+panels, dark green planning units were selected for protection, and pale
+green planning units were discarded.}
+\end{figure}
 

### 2016-07-18 article/supporting-information.tex
+ assorted ms fixes including Rich's edits + results analysed + change to raptr

--- 2016-06-24 code/rmarkdown/supporting-information.Rmd
@@
----
-fontsize: 11pt
-documentclass: article
-bibliography: references.bib
-csl: reference-style.csl
-output:
-  rmarkdown::pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-    includes:
-      in_header: [preamble-latex.tex, preamble-latex3.tex]
----
-# Supporting Information
+\documentclass[11pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
 
-```{r, include=FALSE}
-# set wd
-knitr::opts_knit$set(root.dir=normalizePath('../..'))
-```
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-```{r, include=FALSE}
-## initialization
-# load .rda file
-try(session::restore.session('data/final/results.rda'))
+%%% Change title format to be more compact
+\usepackage{titling}
 
-# set options
-options(error=NULL)
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
 
-## prelimary processing
-# redefine file paths for rasters
-for (i in seq_len(nlayers(cs1.bioclim.RAST)))
-	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])
+\setlength{\droptitle}{-2em}
+  \title{}
+  \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+  \author{}
+  \preauthor{}\postauthor{}
+  \date{}
+  \predate{}\postdate{}
 
-```
+% load packages
+\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,titling,natbib,pdfpages,lineno,booktabs}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage[utf8]{inputenc}
+\usepackage[doublespacing]{setspace}
 
-## Reliable formulation
-The reliable formulation explicitly considers the probability that the planning units are inhabited.
-As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented.
-This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s).
-To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.
+% format captions
+\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}
 
-Backup levels levels are defined as $r$-levels [similar to failure levels in @r18].
-The first backup $r$-level is used to calculate the level of representation when all of the selected planning units are occupied by all $f \in F$.
-For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level$=0$.
-This scenario essentially represents $Y_{fsij}$ in the unreliable formulation.
-The second backup $r$-level is used to assess the level of representation when the closest planning unit to each demand point $i$ is unoccupied.
-For this scenario, the second closest planning units are assigned at $r$-level$=1$.
-The third backup $r$-level is used to assess representation when the first two closest planning units are unoccupied.
-The third closest planning units are assigned at $r$-level$=2$.
-Continuing on, in this manner, the selected planning units in a prioritisation are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.
+% format toc
+%\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
+% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
+% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
+% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}
 
-A final backup $r$-level when $r=R$ is used to assess the level of representation when the features $f \in F$ do not occupy any selected planning units in a prioritisation.
-Each demand point $i \in I$ for each $s \in S$ and $f \in F$ is assigned to an "imaginary" planning unit $j=J$ at $r=R$.
-The distance variables associated with this imaginary planning unit $d_{fsiJ}$ denote the loss of biological value associated with failing to secure a representative sample of feature $f$ in attribute space $s$.
-However, the $d$ variables are in distance units which are meaningless in this context.
-Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).
+% line numbers
+\linenumbers
+
+% format abstract
+\renewcommand{\abstractname}{Summary}
+\renewenvironment{abstract}
+ {\small
+  \begin{center}
+  \bfseries \abstractname\vspace{-.5em}\vspace{0pt}
+  \end{center}
+  \list{} {%
+   \setlength{\leftmargin}{2mm}
+   \setlength{\rightmargin}{\leftmargin}%
+  }%
+  \item\relax}
+{\endlist}
+
+% format section headers
+\titleformat*{\section}{\Large\bfseries}
+\titleformat{\subsection}[display]
+	{\large\sffamily\lsstyle}
+	{\subsectiontitlename\ \thesubsection}
+	{0.5em}{}
+\titleformat*{\subsubsection}{\large\itshape}
+
+% make figures static
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+	\expandafter\origfigure\expandafter[H]
+} {
+	\endorigfigure
+}
+
+% define struts for tables
+\newcommand\T{\rule{0pt}{2.6ex}} % top strut
+\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut
+
+% define command to put new lines in table cells
+%\newcommand{\specialcell}[2][c]{%
+%  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
+% insert S before figure and table caption
+\pagenumbering{gobble}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+
+\begin{document}
+
+\maketitle
+
+\section{Supporting Information}\label{supporting-information}
+
+\subsection{Reliable formulation}\label{reliable-formulation}
+
+The reliable formulation explicitly considers the probability that the
+planning units are inhabited. As a consequence, it may deliver
+prioritisations that will sufficiently represent an attribute space even
+if the features do not inhabit several of the planning units when the
+prioritisation is implemented. This behaviour is achieved by siting
+back-up planning units near selected planning units with low occupancy
+probabilities in the attribute space(s). To ensure that prioritisations
+are robust against multiple planning units being uninhabited, the
+problem assigns planning units at multiple backup levels.
+
+Backup levels levels are defined as $r$-levels (similar to failure
+levels in Snyder \& Daskin 2005). The first backup $r$-level is used to
+calculate the level of representation when all of the selected planning
+units are occupied by all $f \in F$. For this scenario, the closest
+selected planning unit to each demand point $i$ for attribute space $s$
+is assigned at $r$-level$=0$. This scenario essentially represents
+$Y_{fsij}$ in the unreliable formulation. The second backup $r$-level is
+used to assess the level of representation when the closest planning
+unit to each demand point $i$ is unoccupied. For this scenario, the
+second closest planning units are assigned at $r$-level$=1$. The third
+backup $r$-level is used to assess representation when the first two
+closest planning units are unoccupied. The third closest planning units
+are assigned at $r$-level$=2$. Continuing on, in this manner, the
+selected planning units in a prioritisation are assigned to each demand
+point $i \in I$, attribute space $s \in S$, and each feature $f \in F$
+at an $r$-level.
+
+A final backup $r$-level when $r=R$ is used to assess the level of
+representation when the features $f \in F$ do not occupy any selected
+planning units in a prioritisation. Each demand point $i \in I$ for each
+$s \in S$ and $f \in F$ is assigned to an ``imaginary'' planning unit
+$j=J$ at $r=R$. The distance variables associated with this imaginary
+planning unit $d_{fsiJ}$ denote the loss of biological value associated
+with failing to secure a representative sample of feature $f$ in
+attribute space $s$. However, the $d$ variables are in distance units
+which are meaningless in this context. Thus these variables are
+calculated using a failure multiplier ($M$) and the maximum distance
+between the planning units and the demand points for $f \in F$,
+$s \in S$ (4).
 
@@
 
-Moderately-sized conservation planning problems often include several thousand planning units.
-It is currently not be feasible to solve this problem when considering all possible failure scenarios.
-As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$.
-For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level.
-Cui et al.
-[-@r16] found that $R=5$ yields similar solutions to $R=J$ when $J >> 5$.
-However, in most costs the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.
+Moderately-sized conservation planning problems often include several
+thousand planning units. It is currently not be feasible to solve this
+problem when considering all possible failure scenarios. As a
+consequence, the $R$ variable can be any $1 \leq R \leq J-1$. For
+instance, when $R=3$ only 2 backup levels are considered in addition to
+the final backup level. Cui et al. (2010) found that $R=5$ yields
+similar solutions to $R=J$ when $J >> 5$. However, in most costs the
+decision maker will likely be limited to $R=1$ to obtain prioritisations
+in a feasible amount of time.
 
-The control variables for the reliable formulation are the $B$ (eqn 1a), $T_{s}$ (eqn 1b), $\tau_{sa}$ (eqn 1c), $R$, and $M$ variables.
+The control variables for the reliable formulation are the $B$ (eqn 1a),
+$T_{s}$ (eqn 1b), $\tau_{sa}$ (eqn 1c), $R$, and $M$ variables.
 
@@
 
-The decision variables are the $X_j$ (eqn 2a), $Y_{fsijr}$, $P_{fsijr}$ variables.
+The decision variables are the $X_j$ (eqn 2a), $Y_{fsijr}$, $P_{fsijr}$
+variables.
 
@@
 
-The reliable formulation (RRAP) is a multi-objective optimisation problem.
+The reliable formulation (RRAP) is a multi-objective optimisation
+problem.
 
@@
 
-The objective function for the reliable formulation is the same as for the unreliable formation (eqn 3a).
-Similar to the unreliable formulation, constraints (eqn 3b) and (eqn 7a) ensure that the amount-based and space-based targets are met.
-Constraint (eqns 7b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$.
-Constraints (eqn 7d) ensure that only selected planning units are assigned to demand points $i \in I$.
-Constraints (eqn 7e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level.
-Constraints (eqns 7f--7g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information].
-Constraints (eqn 7h) ensure that the $X$ and $Y$ variables are binary.
+The objective function for the reliable formulation is the same as for
+the unreliable formation (eqn 3a). Similar to the unreliable
+formulation, constraints (eqn 3b) and (eqn 7a) ensure that the
+amount-based and space-based targets are met. Constraint (eqns 7b--7c)
+ensure that each planning unit is only assigned to one backup $r$-level
+for $i \in I$. Constraints (eqn 7d) ensure that only selected planning
+units are assigned to demand points $i \in I$. Constraints (eqn 7e)
+ensure that the imaginary planning unit is always assigned to the
+highest backup $r$-level. Constraints (eqns 7f--7g) determine the
+probability that planning unit $j$ will be used to sample demand point
+$i \in I$ for $s \in S$ and $f \in F$ (see Cui \emph{et al.} 2010 for
+more information). Constraints (eqn 7h) ensure that the $X$ and $Y$
+variables are binary.
 
-The reliable formulation is non-linear.
-However, the non-linear components can be linearised.
-First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al.
-[-@r426].
-Second, the expression $P_{fsijr} Y_{jsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
-[-@r16].
+The reliable formulation is non-linear. However, the non-linear
+components can be linearised. First--as discussed in the main text--the
+expression $X_j X_k$ in (eqn 3a) can be linearised using methods
+described by Beyer et al. (2016). Second, the expression
+$P_{fsijr} Y_{jsijr}$ in (eqn 7a) can be linearised using techniques
+described by Sherali and Alameddine (1992) as implemented in Cui et al.
+(2010).
 
-## Figures 
-```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study.
-Each panel shows a the distribution of a solution in environmental space and how it samples the realised niche for a different species.
-The left column of panels shows the solution generated using amount targets.
-The right column of panels shows the solution generated using amount and niche targets.
-Each column of panels corresponds to a different species.
-Hexagons show the distribution of demand points.
-The color of each heaxgon denotes the weighted frequency of demand points inside it.
-Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution.'}
-## prepare data
-cs1.pu.niche.DF <- ddply(
-	cs1.spp.DF,
-	c('Species', 'Prioritisation'),
-	function(x) {
-		# init
-		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
-		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
-		curr.sel.pos <- which(curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@ids %in% which(as.logical(selections(curr.prioritisation))))
-		# extract planning units
-		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@coords %>% as.data.frame()
-		names(curr.DF) <- c('Niche1', 'Niche2')
-		curr.DF$Species <-x$Species[[1]]
-		curr.DF$Prioritisation <-x$Prioritisation[[1]]
-		curr.DF$Solution <- 'discarded'
-		curr.DF$Solution[curr.sel.pos] <- 'selected'
-		# return data
-		return(curr.DF)
-	}
-)
-cs1.pu.niche.DF <- filter(cs1.pu.niche.DF, Solution=='selected')
+\subsection{Figures}\label{figures}
 
-cs1.dp.niche.DF <- ddply(
-	cs1.spp.DF,
-	c('Species', 'Prioritisation'),
-	function(x) {
-		# init
-		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
-		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
-		# extract demand points
-		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@coords %>% as.data.frame()
-		names(curr.DF) <- c('Niche1', 'Niche2')
-		curr.DF$weight <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@weights
-		# return data
-		return(curr.DF)
-	}
-)
-cs1.pu.niche.DF$Prioritisation <- factor(as.character(cs1.pu.niche.DF$Prioritisation), levels=names(cs1.prioritisations))
-cs1.dp.niche.DF$Prioritisation <- factor(as.character(cs1.dp.niche.DF$Prioritisation), levels=names(cs1.prioritisations))
+\begin{figure}[htbp]
+\centering
+\includegraphics{supporting-information_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{Attribute spaces used in the first case-study. Each panel shows
+a the distribution of a solution in environmental space and how it
+samples the realised niche for a different species. The left column of
+panels shows the solution generated using amount targets. The right
+column of panels shows the solution generated using amount and niche
+targets. Each column of panels corresponds to a different species.
+Hexagons show the distribution of demand points. The color of each
+heaxgon denotes the weighted frequency of demand points inside it.
+Points represent the environmental conditions associated with planning
+units inside the species geographic range that were selected for
+preservation in a given solution.}
+\end{figure}
 
-# make letter data.frame
-cs1.letters.DF <- expand.grid(
-	Species=levels(cs1.pu.niche.DF$Species),
-	Prioritisation=levels(cs1.pu.niche.DF$Prioritisation)
-) %>% arrange(Species, Prioritisation)
-cs1.letters.DF$letter <- paste0('(', letters[seq_len(nrow(cs1.letters.DF))], ')')
-cs1.letters.DF <- cbind(
-	cs1.letters.DF,
-	rbind(
-		cs1.pu.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2),
-		cs1.dp.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2)
-	) %>% summarise(
-		Niche1=min(Niche1)+(abs(diff(range(Niche1)))*0.01),
-		Niche2=min(Niche2)+(abs(diff(range(Niche2)))*0.95)
-	) %>% data.frame()
-)
+\subsection{Tables}\label{tables}
 
-## make plots
-ggplot() +
-	stat_summary_hex(data=cs1.dp.niche.DF, aes(x=Niche1, y=Niche2, z=weight), bins=15, fun=function(z) {sum(z)}) +
-	geom_point(data=cs1.pu.niche.DF, aes(x=Niche1, y=Niche2), fill='#00441b', pch=21, color='black') +
-	geom_text(data=cs1.letters.DF, aes(x=Niche1, y=Niche2, label=letter), color='black') +
-	theme_classic() +
-	theme(
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white')
-	) +
-	xlab('Niche (axis 1)') + ylab('Niche (axis 2)') +
-	scale_fill_gradientn(name='Total demand\npoint weight', colors=colorRampPalette(brewer.pal(6, 'PuBu')[-1])(100)) +
-	facet_grid(Species ~ Prioritisation)
-```
+\begin{table}[H]
+\caption{Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritisation (URAP) problem.\label{term.DF}} 
+\begin{center}
+\begin{tabular}{cl}
+\toprule[1pt]
+\multicolumn{1}{c}{Symbol}&\multicolumn{1}{c}{Description}\tabularnewline
+\midrule
+$F$&\parbox[t][][t]{15cm}{set of biodiversity features (indexed by $f$)}\tabularnewline
+$J$&\parbox[t][][t]{15cm}{set of planning units (indexed by $j$)}\tabularnewline
+$q_{fj}$&\parbox[t][][t]{15cm}{probability of feature $f$ occupying planning unit $j$}\tabularnewline
+$e_{jk}$&\parbox[t][][t]{15cm}{shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.}\tabularnewline
+$S$&\parbox[t][][t]{15cm}{set of attribute spaces (indexed by $S$)}\tabularnewline
+$I_{fsi}$&\parbox[t][][t]{15cm}{set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$}\tabularnewline
+$\lambda_{fsi}$&\parbox[t][][t]{15cm}{set of weights for demand point $i$ for feature $f$ in attribute space $s$.}\tabularnewline
+$d_{fsij}$&\parbox[t][][t]{15cm}{distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$.}\tabularnewline
+$\delta_{fsi}$&\parbox[t][][t]{15cm}{the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$.}\tabularnewline
+$T_f$&\parbox[t][][t]{15cm}{amount target for feature $f$.}\tabularnewline
+$\tau_{fs}$&\parbox[t][][t]{15cm}{space-based target for feature $f$ in attribute space $s$.}\tabularnewline
+$X_j$&\parbox[t][][t]{15cm}{binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0).}\tabularnewline
+$Y_{fsij}$&\parbox[t][][t]{15cm}{binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units.}\tabularnewline
+\bottomrule[1pt]
+\end{tabular}\end{center}
 
-## Tables
+\end{table}
 
-```{r, echo=FALSE, results='asis'}
-# make table
-term.CHR <- c(
-	'$F$', 'set of biodiversity features (indexed by $f$)',
-	'$J$', 'set of planning units (indexed by $j$)',
-	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
-	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.',
-	'$S$', 'set of attribute spaces (indexed by $S$)',
-	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
-	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$.',
-	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$.',
-	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$.',
-	'$T_f$', 'amount target for feature $f$.',
-	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$.',
-	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0).',
-	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units.'
-)
-term.DF <- data.frame(
-	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
-	Description=term.CHR[seq(2, length(term.CHR), 2)]
-)
-term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')
+\section{References}\label{references}
 
-# render table
-curr.file <- tempfile(fileext='.tex')
-tmp <- latex(
-		term.DF,
-		file=curr.file, 
-		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
-		colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
-		caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritisation (URAP) problem.'
-	)
-readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
-	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
-	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
-	cat()
-```
+Beyer, H.L., Dujardin, Y., Watts, M.E. \& Possingham, H.P. (2016).
+Solving conservation planning problems with integer linear programming.
+\emph{Ecological Modelling}, \textbf{328}, 14--22.
 
-# References
+Cui, T.T., Ouyang, Y.F. \& Shen, Z.J.M. (2010). Reliable facility
+location design under the risk of disruptions. \emph{Operations
+Research}, \textbf{58}, 998--1011.
+
+Sherali, H. \& Alameddine, A. (1992). A new reformulation-linearization
+technique for bilinear programming problems. \emph{Journal of Global
+Optimization}, \textbf{2}, 379--410.
+
+Snyder, L.V. \& Daskin, M.S. (2005). Reliability models for facility
+location: the expected failure cost case. \emph{Transportation Science},
+\textbf{39}, 400--416.
+
+\end{document}

### 2016-07-18 code/rmarkdown/article.Rmd
+ assorted ms fixes including Rich's edits + results analysed + change to raptr

@@
 ---
-title: "rapr: Representative and Adequate Prioritisations in R"
+title: "raptr: Representative and Adequate Prioritisation Toolkit in R"
 author: |
@@
 1.
-A central aim in conservation is to maximise the long-term persistence of biodiversity.
-To fulfill this aim, reserve networks are used to safeguard biodiversity patterns (eg.
-species, populations) and processes (eg.
-evolutionary processes).
-Reserve selection is often formulated as an optimisation problem to identify cost-effective solutions.
-These formulations often use \"amount targets\" to specify the minimum amount habitat that must be preserved for each species.
-Although such targets are well suited for securing biodiversity patterns, they are not well suited for obtaining a representative sample of the biodiversity processes that play out across species' ranges.
-Currently, decision makers lack the tools to develop prioritisations that preserve both biodiversity patterns and processes.
+An underlying aim in conservation planning is to maximise the long-term persistence of biodiversity.
+To fulfill this aim, the processes the sustain biodiversity must be conserved.
+Many biodiversity processes can be preserved by conserving a representative sample of each biodiversity feature (species or ecosystem).
+This idea has for decades been encapsulated in the principle of \"representativeness\".
+For example, representing the range of genetic diversity or ability to persist in a range of climatic conditions might be crucial for the long-term persistence of a species in a world where environmental change is accelerating.
+Yet current approaches overwhelmingly focus on \"adequacy\", the concept of ensuring a target proportion of the distribution of a biodiversity feature is reserved.
 
 2.
-To address this issue, we developed the \\texttt{rapr R} package.
-This \\texttt{R} package provides a toolkit to guide reserve selection using explicit targets for biodiversity processes.
-Using a novel formulation of the reserve selection problem, users can set \"space targets\" to secure a representative sample of each species within an attribute space to preserve biodiversity processes (eg.
-an attribute space based on genetic data to preserve evolutionary processes).
-We explored the functionality of this \\texttt{R} package using a simulation study and two case studies.
-In each study, we generated prioritisations using amount targets--to represent conventional reserve selection methods--and compared them to prioritisations generated using both amount and space targets.
+To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R}, a toolkit to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets.
+Users set \"space targets\" to secure a representative sample of each feature within an attribute space that reflects a biodiversity process (eg.
+an attribute space expressing variation in genetic characteristics between individuals in an area, or variation in climatic conditions between inhabited areas).
+We explored the functionality of this package using a simulated study and two case studies.
+In each study, we generated solutions that secured a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compared them with solutions generated using both amount and space targets.
 
 3.
-We demonstrated how explicitly considering biodiversity processes in reserve selection can affect the resulting solution.
-The results from the simulation study suggest that most species could benefit from the explicit use of space-based targets--especially species with multi-model distributions in an attribute space.
-The results from both first case studies demonstrated that explicitly considering space-based targets for biodiversity processes can result in more effective prioritisations.
+We show that explicitly considering biodiversity processes in reserve selection can substantially change the solution.
+The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multi-model distributions in an attribute space.
+The case studies show that explicitly considering space-based targets for biodiversity processes can return solutions that could be far more effective in achieving long term biodiversity persistence and do not require much more area than traditional approaches.
 
 4.
-The \\texttt{rapr R} package provides a unified framework for biodiversity patterns and species-specific processes in conservation planning.
-Overall, we found strong evidence that prioritisations developed  using amount targets may fail to secure  that can provide better outcomes than just considering biodiversity patterns.
+The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritisations that are both adequate and representative, resulting in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
+
 "
@@
 csl: reference-style.csl
-vignette: >
-  %\VignetteIndexEntry{rapr}
-  %\VignetteEngine{knitr::rmarkdown}
-  \usepackage[utf8]{inputenc}
 ---
@@
 for (i in seq_len(nlayers(cs1.bioclim.RAST)))
-	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])
+	cs1.bioclim.RAST@layers[[i]]@file@name <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])
 
@@
 Short running title: Representative and adequate prioritisations \newline
-Abstract word count: 322 / 350 \newline
-Total word count: XXXXX / 8000 \newline
-Number of references: 34 \newline
-Number of figures: 8 \newline
-Number of tables: 0 \newline
-Number of textboxes: 0 \newline
-Number of equations: 8 \newline
+Abstract word count: 350 / 350  
+Total word count: XXXXX / 8000  
+Number of references: 37
+Number of figures: 8
+Number of tables: 0  
+Number of textboxes: 0  
+Number of equations: 8  
 \newpage
@@
 <!-- conservation and reserve networks -->
-The overarching aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255].
+Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity at organisational levels [@r423; @r255].
 To achieve this, conservation actions must preserve biodiversity patterns (eg.
-species, populations) and the processes that sustain them.
+populations, species, ecosystems), but also crucially the processes that sustain them.
 One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r440].
-Reserve networks buffer species from threatening processes [eg.
-urbanisation; @r255] and set the stage for direct management interventions [eg.
-captive breeding and reintroduction programs; @r290].
+Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465].
 However, the resources available for conservation action are limited, and so reserve networks must be sited in places that satisfy conservation objectives for minimum cost [@r255].
-To achieve this, reserve selection is often formulated as an optimisation problem and then solved to identify cost-effective candidate reserve systems [prioritisations; @r255].
+To achieve this, reserve selection is often formulated as an optimisation problem and then solved to identify cost-effective candidate reserve systems [termed prioritisations; @r255].
+Many decision support tools have been developed to help identify cost-effective solutions [eg.
+\texttt{ ConsNet}, @r446; \texttt{ Marxan}, @r22; \texttt{ Zonation}, @r429].
+Typically, these tools are used to deliver a prioritisation that preserves a set of species of interest to the conservation planner (termed target species or features).
+Given a set of species, decision makers can use these tools to preserve biodiversity patterns by generating a prioritisation that secures an adequate proportion of each species' range.
 
 <!-- biodiversity processes are important -->
-To fulfill the overarching aims of conservation, reserve networks must preserve both ecological and evolutionary processes [@r3; @r255].
+However, to achieve effective conservation, reserve networks must do more than cover enough area.
+They must capture the processes that sustain biodiversity over the long term; they must capture both ecological and evolutionary processes [@r3; @r255].
 Ecological processes, such as predator-prey interactions, pollination, and decomposition, are required for biodiversity to persist over short time-scales.
@@
 On the other hand, evolutionary processes are required for biodiversity to persist over long time-scales, and they typically operate over large geographic domains.
-Protected areas must preserve adaptive evolutionary processes to foster resilience against environmental change [eg.
+Protected areas must represent adaptive evolutionary processes to foster resilience against environmental change [eg.
 climate change; @r247].
-Protected areas must also preserve neutral evolutionary processes [@r5; @r67; @r63].
-These processes are associated with the breakdown of gene flow between populations.
+Protected areas must also preserve neutral evolutionary processes [@r5; @r67; @r63], arising from restriction of gene flow between populations.
 They are important for maintaining genetic diversity, and avoiding inbreeding depression.
-In recent decades, a wealth of data on biodiversity processes has become available to conservation planners [eg.
+In recent decades, a wealth of data relevant to evolutionary processes has become available to conservation planners [eg.
 bioclimatic, genetic, and trait data; @r33; @r453; @r460].
-Yet this data is only rarely used to guide conservation planning [@r61].
-This is due to that fact that existing reserve selection tools focus on preserving biodiversity patterns or processes--but not both.
+Yet such data are rarely used to guide conservation planning [@r61].
+This is perhaps in large part due to that fact that existing reserve selection tools focus on capturing either biodiversity patterns or processes--but not both.
 
 <!-- parametrisation of biodiversity patterns and processes in conventional conservation planning tools -->
-Many decision support tools have been developed to help identify cost-effective prioritisations [eg.
-\texttt{ ConsNet}, @r446; \texttt{ Marxan}, @r22; \texttt{ Zonation}, @r429].
-Typically, these tools are used to deliver a prioritisation that preserves a set of target species (or features).
-Given a set of species, decision makers can use these tools to preserve biodiversity patterns by generating a prioritisation that secures an adequate proportion of each species' range.
-These tools, however, are not well-suited for biodiversity processes [but see @r203].
-To preserve biodiversity processes, a prioritisation must preserve a representative sample of each species.
+Existing conservation planning tools are generally not well-suited for representing biodiversity processes [but see @r203].
+To preserve biodiversity processes, a prioritisation must capture a representative sample of each species.
 For instance, to preserve predator-prey interactions, a prioritisation must preserve individuals from each predator and prey species in the same area.
-To preserve adaptive evolutionary processes, a prioritisation must preserve the adaptive landscape of each species--it must preserve individuals experiencing different selection pressures [@r63].
+To preserve adaptive evolutionary processes, a prioritisation must preserve the adaptive landscape of each species--populations experiencing different selection pressures [@r63].
 To preserve neutral evolutionary processes, a prioritisation must secure individuals descended from each of the genetic lineages that comprise each species [@r5].
-To accommodate such information into a multi-species conservation planning exercise, species ranges must be partitioned into different groups [eg.
-using habitat data; @r2] as a pre-processing step to be compatible with existing tools.
-However, this approach assumes that the biodiversity processes that operate across a species' range can be split into discrete units.
+Previous attempts to accommodate such into multi-species conservation planning have partitioned species' ranges into different groups, for example at habitat discontinuities [e.g.
+@r2] as a pre-processing step to render a new set of "pseudo-species".
+However, this approach assumes that the biodiversity processes that operate across a species' range can be readily split into discrete units.
 Yet data on biodiversity processes is often continuous and hyper-dimensional, and often cannot be reduced to a few categories without significant information loss [@r218].
@@
 <!-- aims of this paper -->
-Today, one of the key issues in reserve selection is the lack of a unifying decision support tool that can accommodate data on biodiversity patterns and processes in a multi-species context.
-To fill this void, we present the \texttt{rapr R} package.
-This \texttt{R} package uses novel formulations of the reserve selection problem to provide decision makers with the tools to generate prioritisations that preserve biodiversity patterns and processes.
+One of the key issues in reserve selection is the lack of a unifying decision support tool that can accommodate data on biodiversity patterns and processes in a multi-species context.
+To begin to fill this gap, we present the \texttt{raptr R} package.
+This \texttt{R} package uses novel formulations of the reserve selection problem to provide decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes.
 We aim to define the concepts behind the problem formulations.
-Furthmore, we aim to explore the functionality of the \texttt{R} package by applying one of the formulations to a set of simulated species and two case studies.
+Furthermore, we aim to explore the functionality of the \texttt{raptr R} package by applying one of the formulations to a set of simulated species and two case studies.
 
@@
 ## PROBLEM FORMULATION
-Biodiversity features are defined as the entity(s) that the prioritisation is required to preserve (eg.
-species, populations).
+Biodiversity features are defined as the entities that the prioritisation is required to preserve (eg.
+species, ecosystems).
 Spatial attributes are defined as the variation across the species' range that the prioritisation is required to sample.
@@
 Each planning unit is thought to occupy a single point inside each space.
-For example, a decision maker may require a prioritisation that preserves populations along climatic gradients.
+For example, a decision maker may require a prioritisation that represents populations along climatic gradients.
 To achieve this, the decision maker might use an "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm).
@@
 Demand points are points that also exist in an attribute space.
-They are designated by the decision maker to indicate regions of the attribute space that should be preserved in the prioritisation (see below for discussion on how demand points can be generated for real-world datasets).
+They are designated by the decision maker to indicate regions of the attribute space that the decision maker wishes to represent in the prioritisation (see below for discussion on how demand points can be generated for real-world datasets).
 The amount of variation in the attribute space that a prioritisation secures is a function of the distance between each demand point and each selected planning unit in the attribute space.
@@
 In any attribute space there may exist points that are impossible (eg.
-mean annual rainfall -5 mm), do not occur in the study area (eg.
+mean annual rainfall -5 mm), or do not occur in the study area (eg.
 mean annual temperature 30$^{\circ}$C in Antarctica).
 Additionally, there may be some regions that are desirable for some features and undesirable for others (eg.
-conditions known to be outside the physiological tolerance of a species).
+conditions known to be outside the physiological tolerance of certain species).
 Thus a different set of demand points and weights are used for each attribute space and each feature.
@@
 
-To illustrate these concepts, we will briefly describe a conservation planning scenario example involving an attribute spaces and demand points.
-A decision maker may wish to develop a prioritisation for a single species.
-This species has four populations in the study area.
-However, the decision maker can only afford to preserve three population.
-The decision maker needs to preserve the adaptive landscape of the species, and preserve populations inhabiting different environmental conditions.
-To describe environmental variaiton, the decision maker obtained data on the environmental conditions (rainfall (ml) and temperature ($^{\circ}$C)) where each population was found.
-The decision maker then used this environmental data to construct a two-dimensional environmental attribute space.
-Next, the decision maker generated demand points as equi-distant points between the range of values where the populations were found.
-By comparing the distribution of the demand points to the distribution of the populations in the attribute space, the decision maker can identify a suitable prioritisation (Fig.
+To illustrate these concepts, consider an example conservation planning scenario example involving an attribute space and demand points.
+We wish to develop a prioritisation for a single species that has four populations in the study area.
+However, we can only afford to preserve three populations.
+We aim preserve the adaptive landscape of the species, and to achieve this, we will preserve populations inhabiting different environmental conditions.
+To describe environmental variation, we obtain data on the environmental conditions (rainfall (mm) and temperature ($^{\circ}$C)) where each population is found.
+These environmental data are then used to construct a two-dimensional environmental attribute space.
+Next, we generated demand points as equi-distant points between the range of values where the populations were found.
+By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify a suitable prioritisation (Fig.
 1).
-We can see that if the decision maker preserves both populations $A$ and $C$, they will effectively "double-up" on the same environmental characteristics, and in turn their waste resources.
-Instead, a more representative sample of the intra-specific variation could be preserved by securing populations $A$, $B$, and $D$.
+We can see that preserving both populations $A$ and $C$, effectively "doubles-up" on the same environmental characteristics, resulting in considerable redundancy in the solution.
+Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$, $B$, and $D$.
 This example demonstrates how the inclusion of biodiversity processes can guide the reserve selection process.
 
-The formulations used to express the reserve selection problem in the \texttt{rapr R} package are based on a combination of the \texttt{Marxan} reserve selection problem and the uncapacitated facility location problems [@r451].
-Although the \texttt{rapr R} package provides two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in the Supporting Information.
+The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan} reserve selection problem and the uncapacitated facility location problems [@r451].
+Although the \texttt{raptr R} package provides two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in the Supporting Information S1.
+These formulations are named after the unreliable and reliable facility location formulations from which they are based upon [@r16].
 The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied when calculating the proportion of an attribute space sampled in a solution.
@@
 
-Define $F$ to be the set of features (indexed by $f$).
+Define $F$ to be the set of features one wishes to conserve (indexed by $f$).
 Let $J$ be a set of planning units (indexed by $j$).
@@
 Linearised versions of the problems can be solved using commercial exact algorithm solvers.
-The \texttt{rapr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations.
+The \texttt{raptr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations.
 These optimisation problems can then be solved to generate prioritisations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).
@@
 
-The uniform species was simulated with a constant probability of occupancy across the study area (Fig.
+The uniform species was simulated to occur with a constant probability of occupancy across the study area (Fig.
 2a).
@@
 3b).
-Since all planning units have cost the same amount and have an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf.
+Since all planning units have equal cost and an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf.
 Fig 3a).
-Although the use of amount and space based targets has addressed adequacy and representativeness objectives (respectively), they have resulted in a solution that is a highly fragmented solution.
+Although the use of amount and space based targets has addressed adequacy and representativeness objectives (respectively), they have resulted in a highly fragmented solution.
 
@@
 
-The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two isolated populations (Fig.
+The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig.
 2c).
@@
 
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets preserved individuals in both populations (Fig.
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets included planning units from both populations (Fig.
 3j).
@@
 ### Methods
-Here we investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species.
+We investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species.
 By preserving the populations in suitable habitats with different with environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63].
@@
 This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats.
-We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subject them to a principal components analysis (using ArcMap 10.3.1).
-We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principle components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Fig.
+We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
+We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Fig.
 4).
 
-We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), the brown falcon (_Falco berigora_), and the pale headed rosella (_Platycercus adscitus_).
-We used these bird species because they have different evolutionary histories and life-history strategies.
+We selected four bird species that span a range of different evolutionary histories, distributions and ecologies: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_).
 We then mapped the extent of occurrence for each species (Fig.
@@
 To do this, we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
-We used this method because it is reproducible using freely available data and moderately accurate.
+We used this method because it is trivially reproducible using freely available data.
 
@@
 ### Results
-Generally, the solution generated using amount targets preserved a representative sample of each the `r length(cs1.params.LST[[MODE]]$species.names)` bird species' niches--with exception to `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'a single')` bird species.
+Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches--with `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'one')` excpetion.
 Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realised niche of `r parsed.representative.space.held.names`.
@@
 However, despite this, there may yet be species for which only a small fraction of their realised niche is preserved.
-By explictly using space targets, conservation planners can generate prioritisations that guaranteeably preserve a representative sample of species' niches.
+By explictly using space targets, conservation planners can generate prioritisations that are guaranteed to capture a representative sample of species' niches.
 
@@
 ### Methods
-Here we used of space-based targets to generate a prioritisation to secure a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and genomic data collected by the international IntraBioDiv project in the European Alps [@r462; see @r461 and @r463 for further explanation of data collection methods].
-Althougthis dataset contains multiple plant species, we used data for the `r  cs2.params.LST[[MODE]]$common.name` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}).
-This species was chosen because it provides is locally wide-spread and has been shown to exhibit significant genetic structure [@r454].
+Here we used space-based targets to generate a prioritisation securing a representative sample of a species' intra-specific genetic variation.
+We used species occurrence and multilocus AFLP data collected by the international IntraBioDiv project in the European Alps [@r462; see @r461 and @r463 for further explanation of data collection methods].
+Althougthis dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r454].
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 7a).
-They visited every second grid cell, and if the `r cs2.params.LST[[MODE]]$common.name` was detected in a cell, samples were collected from three individuals.
+They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals.
 Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
@@
 
-We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the individual-level presence/absence genetic data into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
-These continuous variables described the main theme of genetic variation within the species.
+We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles wthin individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
+These continuous variables described the main axes of genetic variation within the species.
 We calculated the average of the values associated with individuals in each grid cell.
@@
 Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points.
-To ensure that the solutions did not prioritise particularly costly areas that could not be used for conservation action, we included opportunity cost data (Fig.
+To ensure that the solutions did not prioritise particularly costly areas, we included opportunity cost data (Fig.
 7b).
-We obtained population density data [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
+We obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
 
@@
 Planning units located near each were found to contain individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
-In terms of the typical genetic characteristics of the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
+In terms of the average genetic characteristics of individuals in the 	planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
 This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.
@@
 # Implications and future directions
-The \texttt{rapr} R package provides a unified approach to reserve selection.
-One of the key advantages of the \texttt{rapr R} package is that it is general enough that any spatial variation could be considered an attribute space, regardless of whether this variation is intrinsic or extrinsic to the feature(s).
-This R package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and processes.
+The \texttt{raptr} R package provides a unified approach to reserve selection.
+One of the key advantages of this package is that it is general enough to incorporate any spatial variable as an attribute space.
+This package can accommodate intrinsic or extrinsic variation to the feature(s).
+For example, adaptation processes could be secured using environmental variation or genetic variation among loci under selection.
+Ecological processes, such as predator-prey interactions, could be secured by capturing intra-specific trophic variation.
+Different attribute spaces with different targets can be also specified for different species.
+As long as the variation can be described using Euclidean distances, the \texttt{R} package can be used to obtain a representative sample (note that data may require transformation to conform to this assumption).
+This \\texttt{R} package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and the processes that maintain biodiversity.
 Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
-Both the simulated and case study species suggest that conservation planning exercises need to explicitly consider biodiversity processes during the reserve selection process to capture them.
+Both the simulated and case study species suggest that conservation planning exercises need to explicitly consider biodiversity processes during the reserve selection process to ensure they are captured.
 
@@
 Ultimately, the space-based targets are set as a proportion based on the distribution of the demand-points.
-As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not be an effective prioritisation in reality.
-We therefore stress that decision makers must carefully consider which biodiversity processes need to preserved in the prioritisation, and what spatial data can be used to map these processes.
+As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not be an effective prioritisation.
+We therefore stress that decision makers must carefully consider which biodiversity processes need to be reflected in the prioritisation, and which spatial data can be used to represent these processes.
 Note that maximising one process can be detrimental to another.
-For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other.
-Whereas, to maximise the connecitivity of a prioritisation, reserves ineed to be closer to each other.
+For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other, yet to maximise the connectivity of a prioritisation, reserves need to be closer to each other.
 To assist in the selection of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see the `make.DemandPoints` function).
@@
 Demand points are then generated as random points within the polygon.
-A kernel is then fit to the distribution of the feature in the space [using @r231; @r450], and the demand points are weighted based on the estimated density of the feature at the demand points'.
+A kernel is then fit to the distribution of the feature in the space [using @r231; @r450], and the demand points are weighted based on the estimated density of the feature at the demand points.
 
-To maximise the long-term persistence of biodiversity--the stated goal of conservation--decision makers need to identify prioritisations that preserve existing patterns of biodiversity and the processes that support them.
+The formulation requires spatially comprehensive data to map attribute spaces to planning units.
+For instance, in both case-studies, all planning units occupied by the species were associated with values/coordinates in the attribute spaces.
+However, most real-world data sets are patchy--some planning units will be occupied by species for which attribute space data is not available.
+To use such patchy data with the \\texttt{raptr R} package, the gaps in the data must first be filled in as a pre-processing step.
+Spatially explicit models could be used to estimate values in locations that are missing from a data set [eg.
+krigging or generalised dissimilarity models; @r467; @r466].
+This approach has been successfully applied to a range of biological data sets [@r468].
+
+To maximise the long-term persistence of biodiversity, decision makers need to identify prioritisations that preserve existing patterns of biodiversity and the processes that support them.
 To achieve this, conservation planners need a decision support tool that can explicitly accommodate biodiversity patterns and processes.
-Here, we developed the \texttt{rapr R} package to fill this void.
+Here, we developed the \texttt{raptr R} package to fill this void.
 By exploring the functionality of this package using several simulated species, we found that including space-based targets can radically change a prioritisation for the simplest of species.
@@
 # Data accessibility
-All data, code, and results are stored in an online repository (\url{www.github.com/paleo13/rapr-manuscript}) to replicate and valiate this study.
+All data, code, and results are stored in an online repository (\url{www.github.com/paleo13/raptr-manuscript}) to permit replication and validation of this study.
 

### 2016-07-18 code/rmarkdown/figures.Rmd
+ assorted ms fixes including Rich's edits + results analysed + change to raptr

@@
 for (i in seq_len(nlayers(cs1.bioclim.RAST)))
-	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'rapr-manuscript')[[1]][[2]])
+	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])
 
@@
 ```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Example of an attribute space.
-This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (ml) values.
+This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm) values.
 Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found.
@@
 
-```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.5, fig.width=4.5, fig.cap='Simulation study data set.
+```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.5, fig.width=4.5, fig.cap='Simulated distributions of three species.
 Squares denote planning units.
-Each panel shows the distribution for a different simulated species.
-The color of each planning unit indicates the probability that a given species inhabits it.
+The color of each planning unit indicates the probability that each species inhabits it.
 Panel (a) shows the uniformly distributed species, panel (b) shows the normally distributed species, and panel (c) shows the bimodally distributed species.'}
@@
 Each panel shows a prioritisation generated for a single species using a set of parameters.
-Squares denote planning units.
-Dark green planning units were selected for protection and pale green planning units were discarded.
-Each row of panels corresponds to a different species.
+Squares denote planning units, with dark green showing those planning units selected for protection in a particular prioritisation.
 The top row of panels (a, d, g, j) shows prioritisations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
-Each column of panels correpsonds to a different set of parameters used to generate the prioritisation.
+Each column of panels corresponds to a different set of parameters used to generate the prioritisation.
 The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
@@
 
-```{r, echo=FALSE, fig.height=4.5, fig.width=4.0, fig.cap='Extent of occurence for species used in the first case-study.
+```{r, echo=FALSE, fig.height=4.5, fig.width=4.0, fig.cap='Distribution of the species used in the first case-study.
 Squares denote planning units.
-Each panel shows the distribution of a different species.
-For a given species, planning units that is inhabits are shown bright blue and those that it does not are shown in dark blue.'}
+Each panel shows the distribution of a different species.'}
 # prepare data
@@
 Panel (b) shows the opportunity cost of each planning unit (estimated as the total human population density).
-Panels (b--c) show the spatial distribution of the ordinated genetic data.
+Panels (c--d) show the spatial distribution of the ordinated genetic data.
 These values describe the typical genetic characteristics of individuals in each planning unit.
-Planning units with similar values/color have contain individuals with similar loci polymorphisms.')}
+Planning units with similar values/colors contain individuals with similar loci polymorphisms.')}
 # prepare full grid data
@@
 Panel (a) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets.
-Panel (b) shows the planning units selected selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs2.params.LST[[MODE]]$genetic.target*100,' % genetic targets.
+Panel (b) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs2.params.LST[[MODE]]$genetic.target*100,' % genetic targets.
 Panels (c--d) show the solutions in the genetic space.
 Each point corresponds to a planning unit.
-The coordinates of the points represent the typical genetic characeristics of individuals sampled in planning units (based on an NMDS of the binary loci data).
+The coordinates of the points represent the typical genetic characeristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data).
 Planning units associated with points that are closer togeather contain individuals with more similar genetic characteristics than planning units that are further apart.
@@
 cs2.sel.space.letters.DF <- data.frame(
-	x=min(cs2.sel.space.DF$d1) + (diff(range(cs2.sel.space.DF$d1))*0.05),
-	y=min(cs2.sel.space.DF$d2) + (diff(range(cs2.sel.space.DF$d2))*1),
+	x=min(cs2.sel.space.DF$d1) + (diff(range(cs2.sel.space.DF$d1))*0.3),
+	y=min(cs2.sel.space.DF$d2) + (diff(range(cs2.sel.space.DF$d2))*0.95),
 	letter=c('(c)', '(d)')

### 2016-07-18 code/R/analysis/00-initialization.R

@@ -45,8 +45,8 @@ suppressMessages(library(adehabitatHR))
 suppressMessages(library(ALA4R))
 # devtools::install_github('paleo13/spThin')
 suppressMessages(library(spThin))
-# devtools::install_github('paleo13/rapr')
-suppressMessages(library(rapr))
+# devtools::install_github('paleo13/raptr')
+suppressMessages(library(raptr))
 # devtools::install_github('paleo13/bayescanr')
 suppressMessages(library(bayescanr))
 # devtools::install_github('paleo13/structurer')

### 2016-07-18 code/R/functions/make.genetic.AttributeSpace.R

@@ -5,7 +5,7 @@
 #' @param site.data \code{data.frame} with coordinates for planning units.
 #' @param species.data \code{data.frame} with coordinates of demand points for the species.
 #' @return \code{AttributeSpace}.
-#' @seealso \code{\link[rapr]{AttributeSpace}}, \code{\link[rapr]{AttributeSpace}}.
+#' @seealso \code{\link[raptr]{AttributeSpace}}, \code{\link[rapr]{AttributeSpace}}.
 #' @export
 make.genetic.AttributeSpace <- function(site.data,species.data,species) {
 	# extract ids of pus where the species is found

### 2016-07-18 code/R/functions/make.genetic.AttributeSpace.R
+ change to raptr

@@ -5,7 +5,7 @@
 #' @param site.data \code{data.frame} with coordinates for planning units.
 #' @param species.data \code{data.frame} with coordinates of demand points for the species.
 #' @return \code{AttributeSpace}.
-#' @seealso \code{\link[raptr]{AttributeSpace}}, \code{\link[rapr]{AttributeSpace}}.
+#' @seealso \code{\link[raptr]{AttributeSpace}}.
 #' @export
 make.genetic.AttributeSpace <- function(site.data,species.data,species) {
 	# extract ids of pus where the species is found

### 2016-07-18 code/R/functions/make.targets.R

@@ -7,7 +7,7 @@
 #' @param genetic.space \code{AttributeSpace} object with genetic space data.
 #' @param amount.target \code{numeric} amount-based target
 #' @param genetic.target \code{numeric} genetic-based target
-#' @seealso \code{\link[rapr]{AttributeSpace}}.
+#' @seealso \code{\link[raptr]{AttributeSpace}}.
 #' @export
 make.targets <- function(species, genetic.spaces, amount.target=0.2, genetic.target=0.2) {
 	# amount targets

### 2016-07-22 code/rmarkdown/article.Rmd
+ article fixes + readme updates

@@
 2.
-To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R}, a toolkit to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets.
+To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets.
 Users set \"space targets\" to secure a representative sample of each feature within an attribute space that reflects a biodiversity process (eg.
@@
 <!-- conservation and reserve networks -->
-Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity at organisational levels [@r423; @r255].
+Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255].
 To achieve this, conservation actions must preserve biodiversity patterns (eg.
@@
 Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465].
-However, the resources available for conservation action are limited, and so reserve networks must be sited in places that satisfy conservation objectives for minimum cost [@r255].
+However, the resources available for conservation action are limited, and so reserve networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].
 To achieve this, reserve selection is often formulated as an optimisation problem and then solved to identify cost-effective candidate reserve systems [termed prioritisations; @r255].
@@
 \texttt{ ConsNet}, @r446; \texttt{ Marxan}, @r22; \texttt{ Zonation}, @r429].
-Typically, these tools are used to deliver a prioritisation that preserves a set of species of interest to the conservation planner (termed target species or features).
+Typically, these tools are used to deliver a prioritisation that secures a set of species of interest to the conservation planner (termed target species or features).
 Given a set of species, decision makers can use these tools to preserve biodiversity patterns by generating a prioritisation that secures an adequate proportion of each species' range.
+However, to achieve effective conservation, reserve networks must do more than cover enough area.
 
 <!-- biodiversity processes are important -->
-However, to achieve effective conservation, reserve networks must do more than cover enough area.
-They must capture the processes that sustain biodiversity over the long term; they must capture both ecological and evolutionary processes [@r3; @r255].
+Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255].
 Ecological processes, such as predator-prey interactions, pollination, and decomposition, are required for biodiversity to persist over short time-scales.
@@
 Yet such data are rarely used to guide conservation planning [@r61].
-This is perhaps in large part due to that fact that existing reserve selection tools focus on capturing either biodiversity patterns or processes--but not both.
+This is perhaps in large part due to the fact that existing reserve selection tools focus on capturing either biodiversity patterns or processes--but not both.
 
 <!-- parametrisation of biodiversity patterns and processes in conventional conservation planning tools -->
-Existing conservation planning tools are generally not well-suited for representing biodiversity processes [but see @r203].
+Existing conservation planning tools are generally not well-suited for representing biodiversity patterns and processes.
 To preserve biodiversity processes, a prioritisation must capture a representative sample of each species.
@@
 To preserve neutral evolutionary processes, a prioritisation must secure individuals descended from each of the genetic lineages that comprise each species [@r5].
-Previous attempts to accommodate such into multi-species conservation planning have partitioned species' ranges into different groups, for example at habitat discontinuities [e.g.
+Planning tools that focus on such biodiversity processes [eg.
+\texttt{Diversity}; @r203] are unable to accomodate species-specific targets to ensure that prioritisations secure biodiversity features.
+Similarly, planning tools that focus on biodiversity patterns [eg.
+\texttt{Marxan}] cannot easily accomodate biodiversity processes.
+Previous attempts to accommodate processes into multi-species conservation planning have partitioned species' ranges into different groups, for example at habitat discontinuities [e.g.
 @r2] as a pre-processing step to render a new set of "pseudo-species".
 However, this approach assumes that the biodiversity processes that operate across a species' range can be readily split into discrete units.
-Yet data on biodiversity processes is often continuous and hyper-dimensional, and often cannot be reduced to a few categories without significant information loss [@r218].
+Yet data on biodiversity processes is often continuous, hierarchical, and hyper-dimensional, and so often cannot be reduced to a few categories without significant information loss [@r218].
 

### 2016-07-25 code/rmarkdown/article.Rmd
+ assorted manuscript fixes

@@
     includes:
-      in_header: preamble-latex.tex
+      in_header: preamble.tex
 fontsize: 11pt
@@
 Short running title: Representative and adequate prioritisations \newline
-Abstract word count: 350 / 350  
-Total word count: XXXXX / 8000  
+Abstract word count: 348 / 350  
+Total word count: 5481 / 7000
 Number of references: 37
@@
 We aim preserve the adaptive landscape of the species, and to achieve this, we will preserve populations inhabiting different environmental conditions.
-To describe environmental variation, we obtain data on the environmental conditions (rainfall (mm) and temperature ($^{\circ}$C)) where each population is found.
+To describe environmental variation, we obtained data on the environmental conditions (rainfall (mm) and temperature ($^{\circ}$C)) where each population is found.
 These environmental data are then used to construct a two-dimensional environmental attribute space.
@@
 The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan} reserve selection problem and the uncapacitated facility location problems [@r451].
-Although the \texttt{raptr R} package provides two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in the Supporting Information S1.
+Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in the Supporting Information S1.
 These formulations are named after the unreliable and reliable facility location formulations from which they are based upon [@r16].
@@
 Presently, academics can obtain a license at no cost from the Gurobi website.
-After installing the \texttt{Gurobi} software suite, users will need to install the \texttt{Gurobi R} package.
 
@@
 In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
+Note that the proportion here is negative because it performs worse than simply citing a planning unit in the centroid of the species' demand points.
 
@@
 Here we used space-based targets to generate a prioritisation securing a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and multilocus AFLP data collected by the international IntraBioDiv project in the European Alps [@r462; see @r461 and @r463 for further explanation of data collection methods].
-Althougthis dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r454].
+We used species occurrence and multilocus AFLP data collected by the international IntraBioDiv project in the European Alps [@r462; see @r461 and @r463 for further explanation of data collection].
+Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r454] as a simple case study.
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 The only difference between the two solutions is a single planning unit.
-By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation `r round(cs2.spp.DF$genetic[1],2)` for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).
+By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs.
+8b and 8d) for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).
 
@@
 Ecological processes, such as predator-prey interactions, could be secured by capturing intra-specific trophic variation.
-Different attribute spaces with different targets can be also specified for different species.
-As long as the variation can be described using Euclidean distances, the \texttt{R} package can be used to obtain a representative sample (note that data may require transformation to conform to this assumption).
-This \\texttt{R} package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and the processes that maintain biodiversity.
+Additionally, different attribute spaces with different targets can be specified for different species.
+As long as the variation can be described using Euclidean distances, the \texttt{R} package can be used to obtain a representative sample (note that data may require transformation or dimension reduction to meet this assumption).
+This \texttt{R} package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and the processes that maintain biodiversity.
 Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
@@
 The formulation requires spatially comprehensive data to map attribute spaces to planning units.
-For instance, in both case-studies, all planning units occupied by the species were associated with values/coordinates in the attribute spaces.
+For instance, in both case-studies, all planning units occupied by the species were associated with coordinates in the attribute spaces.
 However, most real-world data sets are patchy--some planning units will be occupied by species for which attribute space data is not available.
-To use such patchy data with the \\texttt{raptr R} package, the gaps in the data must first be filled in as a pre-processing step.
-Spatially explicit models could be used to estimate values in locations that are missing from a data set [eg.
-krigging or generalised dissimilarity models; @r467; @r466].
-This approach has been successfully applied to a range of biological data sets [@r468].
+To use such patchy data with the \texttt{raptr R} package, the gaps in the data must first be filled in.
+Spatially explicit models [eg.
+krigging or generalised dissimilarity models; @r467; @r466] could be used to estimate data for locations that are missing from the data set as a pre-processing step.
+This approach has been successfully applied to a range of biological data sets [eg.
+@r468].
 
@@
 Here, we developed the \texttt{raptr R} package to fill this void.
-By exploring the functionality of this package using several simulated species, we found that including space-based targets can radically change a prioritisation for the simplest of species.
+By exploring the functionality of this package using several simulated and case study species, we found that including space-based targets can result in substantially better solutions.
 
@@
 # Data accessibility
+The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Netowrk (\url{https://cran.r-project.org/}).
 All data, code, and results are stored in an online repository (\url{www.github.com/paleo13/raptr-manuscript}) to permit replication and validation of this study.

### 2016-07-25 code/rmarkdown/figures.Rmd
+ assorted manuscript fixes

@@
     includes:
-      in_header: [preamble-latex.tex, preamble-latex2.tex]
+      in_header: [preamble.tex, figures-preamble.tex]
 ---

### 2016-07-25 code/rmarkdown/supporting-information.Rmd
+ assorted manuscript fixes

@@
     includes:
-      in_header: [preamble-latex.tex, preamble-latex3.tex]
+      in_header: [preamble.tex, si-preamble.tex]
 ---

### 2016-07-25 code/rmarkdown/article.Rmd
+ include infra-specific refs

@@
 In recent decades, a wealth of data relevant to evolutionary processes has become available to conservation planners [eg.
-bioclimatic, genetic, and trait data; @r33; @r453; @r460].
+bioclimatic, genetic, phylogenetic, and trait data; @r33; @r453; @r460].
+This boom has resulted in significant progress towards evolutionary processes into conservation planning at the infra- and inter-specific level [eg.
+@r68; @r470; @r469].
 Yet such data are rarely used to guide conservation planning [@r61].
@@
 <!-- parametrisation of biodiversity patterns and processes in conventional conservation planning tools -->
-Existing conservation planning tools are generally not well-suited for representing biodiversity patterns and processes.
-To preserve biodiversity processes, a prioritisation must capture a representative sample of each species.
+Existing conservation planning tools are generally not well-suited for representing inttra-specific biodiversity patterns and processes.
+To preserve the biodiversity processes that underpin a given species, a prioritisation must capture a representative sample of each species.
 For instance, to preserve predator-prey interactions, a prioritisation must preserve individuals from each predator and prey species in the same area.
@@
 Planning tools that focus on such biodiversity processes [eg.
-\texttt{Diversity}; @r203] are unable to accomodate species-specific targets to ensure that prioritisations secure biodiversity features.
+\texttt{Diversity}; @r203] are unable to accommodate species-specific targets to ensure that prioritisations secure biodiversity features.
 Similarly, planning tools that focus on biodiversity patterns [eg.

### 2016-08-17 code/R/analysis/00-initialization.R
+ fixed bug where checkpoint only used in first script

@@ -6,17 +6,23 @@ if (length(args)>0) {
 	if (grepl('MODE',args))
 		MODE <- strsplit(grep('MODE', args, value=TRUE), '=', fixed=TRUE)[[1]][[2]]
 }
-	
+
+### set parameters
+if (!exists('MODE')) MODE <- 'debug'
+cat('MODE = ',MODE,'\n')
+
 #### Load pacakges
 # load bioconductor packages
 # set checkpoint
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
-library(checkpoint)
+if (!'RcppTOML' %in% installed.packages()[,'Package']) install.packages('RcppTOML')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-checkpoint('2016-07-25', R.version='3.3.1')
+general.params.LST <- parseTOML('code/parameters/general.toml')
+checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version)
 
 # load CRAN packages
 library(stats)
+library(checkpoint)
 library(RColorBrewer)
 library(hexbin)
 library(fields)
@@ -59,7 +65,7 @@ library(spThin)
 
 # install raptr
 if (!'raptr' %in% installed.packages()[,'Package']) {
-	withr::with_libpaths(.libPaths()[1], install.packages(c('adehabitatLT', 'adehabitatHS', 'deldir', 'R.utils', 'geometry', 'KernSmooth', 'misc3d', 'multicool', 'fastcluster', 'rgdal', 'raster', 'PBSmapping', 'RJSONIO', 'R.methodsS3', 'R.oo')))
+	withr::with_libpaths(.libPaths()[1], install.packages(c('adehabitatLT', 'adehabitatHS', 'deldir', 'R.utils', 'geometry', 'KernSmooth', 'misc3d', 'multicool', 'fastcluster', 'rgdal', 'raster', 'PBSmapping', 'RJSONIO', 'R.methodsS3', 'R.oo', 'RgoogleMaps')))
 	devtools::install_github('paleo13/raptr', dependencies=TRUE)
 }
 library(raptr)
@@ -71,6 +77,16 @@ library(bayescanr)
 if (!'structurer' %in% installed.packages()[,'Package'])
 	withr::with_libpaths(.libPaths()[1], devtools::install_github('paleo13/structurer', dependencies=TRUE))
[…]

### 2016-08-17 code/R/analysis/01-load-data.R

@@ -1,5 +1,6 @@
 ## load .rda
 session::restore.session('data/intermediate/00-initialization.rda')
+checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ### load simulation data
 # all data is simulated using functions in the package - no data to load

### 2016-08-17 code/R/analysis/02-simulations.R

@@ -1,5 +1,6 @@
 ## load .rda
 session::restore.session('data/intermediate/01-load-data.rda')
+checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## load parameters
 sim.params.LST <- parseTOML('code/parameters/simulations.toml')

### 2016-08-17 code/R/analysis/03-case-study-1.R

@@ -1,5 +1,6 @@
 ## load .rda
 session::restore.session('data/intermediate/01-load-data.rda')
+checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## load parameters
 cs1.params.LST <- parseTOML('code/parameters/case-study-1.toml')

### 2016-08-17 code/R/analysis/04-case-study-2.R

@@ -1,5 +1,6 @@
 ## load .rda
 session::restore.session('data/intermediate/01-load-data.rda')
+checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## load parameters
 cs2.params.LST <- parseTOML('code/parameters/case-study-2.toml')

### 2016-08-17 code/R/analysis/05-statistical-analysis.R

@@ -2,15 +2,13 @@
 session::restore.session('data/intermediate/04-case-study-2.rda')
 load('data/intermediate/03-case-study-1.rda')
 load('data/intermediate/02-simulations.rda')
+checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## simulation study
 
 ## case-study 1
 
-
-
 ## case-study 2
-
 # generate inverse distance matrix
 cs2.pu.centroids <- gCentroid(cs2.grid.sub.PLY, byid=TRUE) %>% spTransform(CRS('+init=epsg:4326'))
 cs2.pu.inv.dists <- cs2.pu.centroids@coords %>% rdist.earth

### 2016-08-17 code/R/analysis/06-exports.R

@@ -1,5 +1,6 @@
  ## load .rda
 session::restore.session('data/intermediate/05-statistical-analysis.rda')
+checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 
 

### 2016-08-17 code/R/analysis/00-initialization.R
+ fix more bug

@@ -17,8 +17,10 @@ cat('MODE = ',MODE,'\n')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!'RcppTOML' %in% installed.packages()[,'Package']) install.packages('RcppTOML')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-general.params.LST <- parseTOML('code/parameters/general.toml')
-checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version)
+general.params.LST <- Rcpp::parseTOML('code/parameters/general.toml')
+checkpoint::checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version)
+if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
+if (!'RcppTOML' %in% installed.packages()[,'Package']) install.packages('RcppTOML')
 
 # load CRAN packages
 library(stats)

### 2016-08-17 code/R/analysis/00-initialization.R
+ fix bug in 00*.R

@@ -17,7 +17,7 @@ cat('MODE = ',MODE,'\n')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!'RcppTOML' %in% installed.packages()[,'Package']) install.packages('RcppTOML')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-general.params.LST <- Rcpp::parseTOML('code/parameters/general.toml')
+general.params.LST <- RcppTOML::parseTOML('code/parameters/general.toml')
 checkpoint::checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version)
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!'RcppTOML' %in% installed.packages()[,'Package']) install.packages('RcppTOML')

### 2016-08-17 code/R/analysis/00-initialization.R
+ add raptr dependencies

@@ -67,7 +67,7 @@ library(spThin)
 
 # install raptr
 if (!'raptr' %in% installed.packages()[,'Package']) {
-	withr::with_libpaths(.libPaths()[1], install.packages(c('adehabitatLT', 'adehabitatHS', 'deldir', 'R.utils', 'geometry', 'KernSmooth', 'misc3d', 'multicool', 'fastcluster', 'rgdal', 'raster', 'PBSmapping', 'RJSONIO', 'R.methodsS3', 'R.oo', 'RgoogleMaps')))
+	withr::with_libpaths(.libPaths()[1], install.packages(c('adehabitatLT', 'adehabitatHS', 'deldir', 'R.utils', 'geometry', 'KernSmooth', 'misc3d', 'multicool', 'fastcluster', 'rgdal', 'raster', 'PBSmapping', 'RJSONIO', 'R.methodsS3', 'R.oo', 'RgoogleMaps', 'png', 'gdalUtils', 'rgl', 'shape')))
 	devtools::install_github('paleo13/raptr', dependencies=TRUE)
 }
 library(raptr)

### 2016-08-17 code/R/analysis/00-initialization.R
+ actually use checkpoint this time

@@ -7,20 +7,18 @@ if (length(args)>0) {
 		MODE <- strsplit(grep('MODE', args, value=TRUE), '=', fixed=TRUE)[[1]][[2]]
 }
 
-### set parameters
-if (!exists('MODE')) MODE <- 'debug'
-cat('MODE = ',MODE,'\n')
-
 #### Load pacakges
 # load bioconductor packages
 # set checkpoint
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
-if (!'RcppTOML' %in% installed.packages()[,'Package']) install.packages('RcppTOML')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-general.params.LST <- RcppTOML::parseTOML('code/parameters/general.toml')
-checkpoint::checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version)
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
-if (!'RcppTOML' %in% installed.packages()[,'Package']) install.packages('RcppTOML')
+
+### set parameters
+if (!exists('MODE')) MODE <- 'debug'
+cat('MODE = ',MODE,'\n')
+general.params.LST <- RcppTOML::parseTOML('code/parameters/general.toml')
 
 # load CRAN packages
 library(stats)

### 2016-08-17 code/R/analysis/01-load-data.R

@@ -1,6 +1,6 @@
 ## load .rda
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/00-initialization.rda')
-checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ### load simulation data
 # all data is simulated using functions in the package - no data to load

### 2016-08-17 code/R/analysis/02-simulations.R

@@ -1,6 +1,6 @@
 ## load .rda
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/01-load-data.rda')
-checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## load parameters
 sim.params.LST <- parseTOML('code/parameters/simulations.toml')

### 2016-08-17 code/R/analysis/03-case-study-1.R

@@ -1,6 +1,6 @@
 ## load .rda
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/01-load-data.rda')
-checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## load parameters
 cs1.params.LST <- parseTOML('code/parameters/case-study-1.toml')
@@ -49,13 +49,13 @@ cs1.subset.records.LST <- llply(cs1.records.LST, function(x) {
 	# extract valid records
 	x <- dplyr::filter(
 		x$data,
-		speciesOutsideExpertRange==FALSE,
 		inferredDuplicateRecord==FALSE,
 		coordinateUncertaintyInMetres<=10000,
 		is.finite(longitude),
 		is.finite(latitude)
 	)
 	if ('speciesOutsideExpertRange' %in% names(x)) x <- filter(x, speciesOutsideExpertRange==FALSE)
+	if ('habitatMismatch' %in% names(x)) x <- filter(x, habitatMismatch==FALSE)
 	# subset to first n records (used for debugging parameters)
 	if (nrow(x) > cs1.params.LST[[MODE]]$max.records) x <- x[seq_len(cs1.params.LST[[MODE]]$max.records),]
 	# return subsetted data

### 2016-08-17 code/R/analysis/04-case-study-2.R

@@ -1,6 +1,6 @@
 ## load .rda
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/01-load-data.rda')
-checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## load parameters
 cs2.params.LST <- parseTOML('code/parameters/case-study-2.toml')

### 2016-08-17 code/R/analysis/05-statistical-analysis.R

@@ -1,8 +1,8 @@
 ## load .rda
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/04-case-study-2.rda')
 load('data/intermediate/03-case-study-1.rda')
 load('data/intermediate/02-simulations.rda')
-checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
 
 ## simulation study
 

### 2016-08-17 code/R/analysis/06-exports.R

@@ -1,8 +1,6 @@
  ## load .rda
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/05-statistical-analysis.rda')
-checkpoint(general.params.LST[[MODE]]$checkpoint_date, R.version=general.params.LST[[MODE]]$checkpoint_R_version, scanForPackages=FALSE)
-
-
 
 ## save workspace
 save.session('data/final/results.rda', compress='xz')

### 2016-08-17 code/rmarkdown/article.Rmd
+ use checkpoint in rmd files

@@
 # load .rda file
+checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
 try(session::restore.session('data/final/results.rda'))

### 2016-08-17 code/R/analysis/00-initialization.R
+ ensure that rcpptoml is installed before trying to load toml files

@@ -15,11 +15,6 @@ if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
 checkpoint::checkpoint('2016-07-25', R.version='3.3.1')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 
-### set parameters
-if (!exists('MODE')) MODE <- 'debug'
-cat('MODE = ',MODE,'\n')
-general.params.LST <- RcppTOML::parseTOML('code/parameters/general.toml')
-
 # load CRAN packages
 library(stats)
 library(checkpoint)
@@ -91,6 +86,11 @@ if (!'gurobi' %in% installed.packages()[,'Package']) {
 # manually install custom fork of ggplot2 for plotting
 devtools::install_github('paleo13/ggplot2', force=TRUE)
 library(ggplot2)
+
+### set parameters
+if (!exists('MODE')) MODE <- 'debug'
+cat('MODE = ',MODE,'\n')
+general.params.LST <- RcppTOML::parseTOML('code/parameters/general.toml')
 
 ## misc settings
 # set pander options

### 2016-09-28 code/rmarkdown/article.Rmd
+  updated refs and makefile

@@
 <!-- parametrisation of biodiversity patterns and processes in conventional conservation planning tools -->
-Existing conservation planning tools are generally not well-suited for representing inttra-specific biodiversity patterns and processes.
+Existing conservation planning tools are generally not well-suited for representing intra-specific biodiversity patterns and processes.
 To preserve the biodiversity processes that underpin a given species, a prioritisation must capture a representative sample of each species.
-For instance, to preserve predator-prey interactions, a prioritisation must preserve individuals from each predator and prey species in the same area.
+For instance, to preserve predator-prey interactions, a prioritisation must preserve individuals from each species in the same areas [@r471].
 To preserve adaptive evolutionary processes, a prioritisation must preserve the adaptive landscape of each species--populations experiencing different selection pressures [@r63].

### 2016-09-30 code/rmarkdown/article.Rmd
+ assorted changes

@@
 checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
-try(session::restore.session('data/final/results.rda'))
-
-# set options
-options(error=NULL)
+session::restore.session('data/final/results.rda')
 
@@
 # Data accessibility
-The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Netowrk (\url{https://cran.r-project.org/}).
+The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://cran.r-project.org/}).
 All data, code, and results are stored in an online repository (\url{www.github.com/paleo13/raptr-manuscript}) to permit replication and validation of this study.

### 2016-09-30 code/R/analysis/00-initialization.R
+ updated cmds to render ms

@@ -52,6 +52,7 @@ library(english)
 library(knitr)
 library(lazyWeave)
 library(broom)
+library(rmarkdown)
 
 
 ## load github packages

### 2016-09-30 code/rmarkdown/article.Rmd
+ make cmd fixes

@@
 The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://cran.r-project.org/}).
-All data, code, and results are stored in an online repository (\url{www.github.com/paleo13/raptr-manuscript}) to permit replication and validation of this study.
+All data, code, and results are stored in an online repository (\url{https://www.github.com/paleo13/raptr-manuscript}) to permit replication and validation of this study.
 

### 2016-09-30 code/rmarkdown/article.Rmd
+ reference fixes + added kate temp files to gitignore
+ update r pkg refs

@@
 In recent decades, a wealth of data relevant to evolutionary processes has become available to conservation planners [eg.
-bioclimatic, genetic, phylogenetic, and trait data; @r33; @r453; @r460].
+bioclimatic, genetic, phylogenetic, and trait data; @r33; @r460; @r453].
 This boom has resulted in significant progress towards evolutionary processes into conservation planning at the infra- and inter-specific level [eg.
@@
 \texttt{Diversity}; @r203] are unable to accommodate species-specific targets to ensure that prioritisations secure biodiversity features.
-Similarly, planning tools that focus on biodiversity patterns [eg.
-\texttt{Marxan}] cannot easily accomodate biodiversity processes.
+Similarly, planning tools that focus on biodiversity patterns (eg.
+\texttt{Marxan}) cannot easily accomodate biodiversity processes.
 Previous attempts to accommodate processes into multi-species conservation planning have partitioned species' ranges into different groups, for example at habitat discontinuities [e.g.

### 2016-09-30 code/rmarkdown/article.Rmd
+ attempts to manipulate refs

@@
 Here we used space-based targets to generate a prioritisation securing a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and multilocus AFLP data collected by the international IntraBioDiv project in the European Alps [@r462; see @r461 and @r463 for further explanation of data collection].
-Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r454] as a simple case study.
+We used species occurrence and multilocus AFLP data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
+Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study.
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals.
-Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r453], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
+Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
 In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.

### 2016-10-12 code/rmarkdown/article.Rmd
+ try using aligned environment
+ attempt 2
+ aligned attempt 2
+ attempt 3
+ assorted improvements after jonathons comments

@@
 An underlying aim in conservation planning is to maximise the long-term persistence of biodiversity.
-To fulfill this aim, the processes the sustain biodiversity must be conserved.
+To fulfill this aim, the processes that sustain biodiversity must be conserved.
 Many biodiversity processes can be preserved by conserving a representative sample of each biodiversity feature (species or ecosystem).
@@
 2.
-To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets.
-Users set \"space targets\" to secure a representative sample of each feature within an attribute space that reflects a biodiversity process (eg.
-an attribute space expressing variation in genetic characteristics between individuals in an area, or variation in climatic conditions between inhabited areas).
+To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using conventional area protection targets as well as explicit targets for representing intra-specific biodiversity processes.
+Users set \"amount targets\" to preserve an adequate proportion of habitat for each species.
+They also set \"representation targets\" to secure a representative sample of variation within features reflect a biodiversity process (eg.
+genetic characteristics between individuals in areas, or variation in climatic conditions between inhabited areas).
 We explored the functionality of this package using a simulated study and two case studies.
@@
 We show that explicitly considering biodiversity processes in reserve selection can substantially change the solution.
-The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multi-model distributions in an attribute space.
-The case studies show that explicitly considering space-based targets for biodiversity processes can return solutions that could be far more effective in achieving long term biodiversity persistence and do not require much more area than traditional approaches.
+The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multi-modal distributions in an attribute space.
+The case studies show that explicitly considering representation-based targets for biodiversity processes can return solutions that could be far more effective in achieving long term biodiversity persistence and do not require much more area to be reserved than traditional approaches.
 
@@
 # Introduction
-<!-- conservation and reserve networks -->
+<!-- conservation and reserve selection -->
 Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255].
@@
 Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465].
-However, the resources available for conservation action are limited, and so reserve networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].
-To achieve this, reserve selection is often formulated as an optimisation problem and then solved to identify cost-effective candidate reserve systems [termed prioritisations; @r255].
-Many decision support tools have been developed to help identify cost-effective solutions [eg.
-\texttt{ ConsNet}, @r446; \texttt{ Marxan}, @r22; \texttt{ Zonation}, @r429].
-Typically, these tools are used to deliver a prioritisation that secures a set of species of interest to the conservation planner (termed target species or features).
-Given a set of species, decision makers can use these tools to preserve biodiversity patterns by generating a prioritisation that secures an adequate proportion of each species' range.
-However, to achieve effective conservation, reserve networks must do more than cover enough area.
+Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].
+However, despite the importance of biodiversity processes, current approaches for incorporating them into reserve selection are limited.
 
 <!-- biodiversity processes are important -->
-Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255].
-Ecological processes, such as predator-prey interactions, pollination, and decomposition, are required for biodiversity to persist over short time-scales.
-Typically, they operate over small geographic domains--with exceptions such as migration and refugial habitats--and can be preserved using suitably large planning units [@r446] that each contain a discrete unit of habitat [@r448].
-On the other hand, evolutionary processes are required for biodiversity to persist over long time-scales, and they typically operate over large geographic domains.
-Protected areas must represent adaptive evolutionary processes to foster resilience against environmental change [eg.
-climate change; @r247].
-Protected areas must also preserve neutral evolutionary processes [@r5; @r67; @r63], arising from restriction of gene flow between populations.
-They are important for maintaining genetic diversity, and avoiding inbreeding depression.
-In recent decades, a wealth of data relevant to evolutionary processes has become available to conservation planners [eg.
-bioclimatic, genetic, phylogenetic, and trait data; @r33; @r460; @r453].
-This boom has resulted in significant progress towards evolutionary processes into conservation planning at the infra- and inter-specific level [eg.
-@r68; @r470; @r469].
-Yet such data are rarely used to guide conservation planning [@r61].
-This is perhaps in large part due to the fact that existing reserve selection tools focus on capturing either biodiversity patterns or processes--but not both.
+Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255; @r474].
+Ecological processes are important for biodiversity to persist over short time-scales.
+For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@rXXX].
+Additionally, migratory processes  allow many species to access suitable habitat for breeding [@rXXX].
+On the other hand, evolutionary processes are important for biodiversity to persist over long time-scales.
+Adaptive evolutionary processes foster resilience against environmental change [@r3; @r43], and neutral evolutionary processes maintain genetic diversity [@r5; @r67; @r63].
+In recent decades, a wealth of data relevant to biodiversity processes has become available to conservation planners [eg.
+bioclimatic, distribution, genetic, phylogenetic, and trait data; @rXXX; @r33; @r460; @r453].
+This boom has resulted in significant progress towards incorporating both ecological and evolutionary processes into conservation planning [@r473; @r474; @r448; @r2; @r68; @r470; @r469; @r247; @r488; @r468; @r407].
 
-<!-- parametrisation of biodiversity patterns and processes in conventional conservation planning tools -->
-Existing conservation planning tools are generally not well-suited for representing intra-specific biodiversity patterns and processes.
-To preserve the biodiversity processes that underpin a given species, a prioritisation must capture a representative sample of each species.
-For instance, to preserve predator-prey interactions, a prioritisation must preserve individuals from each species in the same areas [@r471].
-To preserve adaptive evolutionary processes, a prioritisation must preserve the adaptive landscape of each species--populations experiencing different selection pressures [@r63].
-To preserve neutral evolutionary processes, a prioritisation must secure individuals descended from each of the genetic lineages that comprise each species [@r5].
-Planning tools that focus on such biodiversity processes [eg.
-\texttt{Diversity}; @r203] are unable to accommodate species-specific targets to ensure that prioritisations secure biodiversity features.
-Similarly, planning tools that focus on biodiversity patterns (eg.
-\texttt{Marxan}) cannot easily accomodate biodiversity processes.
-Previous attempts to accommodate processes into multi-species conservation planning have partitioned species' ranges into different groups, for example at habitat discontinuities [e.g.
-@r2] as a pre-processing step to render a new set of "pseudo-species".
-However, this approach assumes that the biodiversity processes that operate across a species' range can be readily split into discrete units.
-Yet data on biodiversity processes is often continuous, hierarchical, and hyper-dimensional, and so often cannot be reduced to a few categories without significant information loss [@r218].
+<!-- current methods for preserving biodiversity processes -->
+Many decision support tools have been developed to help identify cost-effective solutions [eg.
+\texttt{ConsNet}, @r446; \texttt{Marxan}, @r22; \texttt{Zonation}, @r429].
+Typically, these tools are used to deliver a prioritisation that secures a set of species of interest to the conservation planner (termed target species or features).
+Given a set of species, decision makers use these tools to generate a prioritisation that secures an adequate proportion of each species' range.
+To accommodate biodiversity processes, current methods involve partitioning species' ranges into different groups as a pre-processing step to render a new set of "pseudo-species" [eg.
+at habitat discontinuities; @r2].
+However, this approach suffers two main limitations.
+First, splitting species' ranges into discrete groups is not a trivial task.
+Data on biodiversity processes are often continuous, hierarchical, and hyper-dimensional, and so may not be able to be reduced to a few groups without significant information loss [@r218].
+Second, by splitting the proxy variable into groups, the decision maker cannot directly assess how well a prioritisation actually samples variation in the original proxy variable.
+Simply setting targets for representing each of the pseudo-species does not necessarily guarantee that the original proxy variable been well sampled--and by extension that the biodiversity process has been preserved.
 
 <!-- aims of this paper -->
-One of the key issues in reserve selection is the lack of a unifying decision support tool that can accommodate data on biodiversity patterns and processes in a multi-species context.
+One of the key issues in reserve selection is the lack of a unifying decision support tool that can accommodate data on biodiversity patterns and processes.
 To begin to fill this gap, we present the \texttt{raptr R} package.
 This \texttt{R} package uses novel formulations of the reserve selection problem to provide decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes.
-We aim to define the concepts behind the problem formulations.
-Furthermore, we aim to explore the functionality of the \texttt{raptr R} package by applying one of the formulations to a set of simulated species and two case studies.
+Here, we aim to define the concepts that underpin this \texttt{R} package and explore its functionality.
 
@@
 7c--7d).
-Planning units located near each were found to contain individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
+Planning units located near each other were found to contain individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
 In terms of the average genetic characteristics of individuals in the 	planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).

### 2016-10-12 code/rmarkdown/article.Rmd
+ actually check new version of article into vc

@@
 Many biodiversity processes can be preserved by conserving a representative sample of each biodiversity feature (species or ecosystem).
-This idea has for decades been encapsulated in the principle of \"representativeness\".
-For example, representing the range of genetic diversity or ability to persist in a range of climatic conditions might be crucial for the long-term persistence of a species in a world where environmental change is accelerating.
-Yet current approaches overwhelmingly focus on \"adequacy\", the concept of ensuring a target proportion of the distribution of a biodiversity feature is reserved.
+For example, representing ability for individuals to persist in a range of climatic conditions might be crucial for the long-term persistence of a species in a world where environmental change is accelerating.
+Yet current approaches overwhelmingly focus on securing a target proportion of the features' geographic range.
 
 2.
-To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using conventional area protection targets as well as explicit targets for representing intra-specific biodiversity processes.
-Users set \"amount targets\" to preserve an adequate proportion of habitat for each species.
-They also set \"representation targets\" to secure a representative sample of variation within features reflect a biodiversity process (eg.
-genetic characteristics between individuals in areas, or variation in climatic conditions between inhabited areas).
-We explored the functionality of this package using a simulated study and two case studies.
+To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets.
+Users set \"space targets\" to secure a representative sample of each feature within an attribute space that reflects a biodiversity process (eg.
+an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
+Users also set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribtuion.
+We demonstrated the functionality of this package using a simulated study and two case studies.
 In each study, we generated solutions that secured a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compared them with solutions generated using both amount and space targets.
@@
 We show that explicitly considering biodiversity processes in reserve selection can substantially change the solution.
-The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multi-modal distributions in an attribute space.
-The case studies show that explicitly considering representation-based targets for biodiversity processes can return solutions that could be far more effective in achieving long term biodiversity persistence and do not require much more area to be reserved than traditional approaches.
+The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multi-model distributions in an attribute space.
+The case studies show that explicitly considering space-based targets for biodiversity processes can return solutions that could be far more effective in achieving long term biodiversity persistence and do not require a much greater area reserved than traditional approaches.
 
@@
 Short running title: Representative and adequate prioritisations \newline
-Abstract word count: 348 / 350  
-Total word count: 5481 / 7000  
-Number of references: 37  
-Number of figures: 8  
+Abstract word count: XXX / XXX  
+Total word count: XXX / XXX  
+Number of references: XXX  
+Number of figures: XXX  
 Number of tables: 0  
 Number of textboxes: 0  
-Number of equations: 8  
+Number of equations: XXX  
 \newpage
 
-# Introduction
 <!-- conservation and reserve selection -->
@@
 Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].
-However, despite the importance of biodiversity processes, current approaches for incorporating them into reserve selection are limited.
+However, despite the importance of biodiversity processes, they are only rarely used to guide conservation planning.
 
 <!-- biodiversity processes are important -->
-Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255; @r474].
+Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255; @r9].
 Ecological processes are important for biodiversity to persist over short time-scales.
-For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@rXXX].
-Additionally, migratory processes  allow many species to access suitable habitat for breeding [@rXXX].
+For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@r476].
+Additionally, migratory processes allow many species to access suitable habitat throughout the year [r477].
 On the other hand, evolutionary processes are important for biodiversity to persist over long time-scales.
 Adaptive evolutionary processes foster resilience against environmental change [@r3; @r43], and neutral evolutionary processes maintain genetic diversity [@r5; @r67; @r63].
-In recent decades, a wealth of data relevant to biodiversity processes has become available to conservation planners [eg.
-bioclimatic, distribution, genetic, phylogenetic, and trait data; @rXXX; @r33; @r460; @r453].
-This boom has resulted in significant progress towards incorporating both ecological and evolutionary processes into conservation planning [@r473; @r474; @r448; @r2; @r68; @r470; @r469; @r247; @r488; @r468; @r407].
+In recent decades, a wealth of data relevant to biodiversity processes has become available [eg.
+bioclimatic, distribution, genetic, phylogenetic, and trait data; @r406; @r33; @r460; @r453].
+This is sparked significant interest in using this data to make better conservation decisions [@r473; @r9; @r448; @r2; @r68; @r470; @r469; @r247; @r488; @r468; @r407; @r61; @r475].
 
 <!-- current methods for preserving biodiversity processes -->
-Many decision support tools have been developed to help identify cost-effective solutions [eg.
-\texttt{ConsNet}, @r446; \texttt{Marxan}, @r22; \texttt{Zonation}, @r429].
-Typically, these tools are used to deliver a prioritisation that secures a set of species of interest to the conservation planner (termed target species or features).
-Given a set of species, decision makers use these tools to generate a prioritisation that secures an adequate proportion of each species' range.
-To accommodate biodiversity processes, current methods involve partitioning species' ranges into different groups as a pre-processing step to render a new set of "pseudo-species" [eg.
-at habitat discontinuities; @r2].
-However, this approach suffers two main limitations.
-First, splitting species' ranges into discrete groups is not a trivial task.
-Data on biodiversity processes are often continuous, hierarchical, and hyper-dimensional, and so may not be able to be reduced to a few groups without significant information loss [@r218].
-Second, by splitting the proxy variable into groups, the decision maker cannot directly assess how well a prioritisation actually samples variation in the original proxy variable.
-Simply setting targets for representing each of the pseudo-species does not necessarily guarantee that the original proxy variable been well sampled--and by extension that the biodiversity process has been preserved.
+Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Marxan} and \texttt{Zonation}; @r22; @r429].
+The cornerstone of these tools is that decision makers can set targets for a proportion of the geographic range of species, or ecosystems (termed features).
+To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step [@r9; @r448; @r247].
+These groups are partitioned based on a proxy variable that relates to a biodiversity process of interest [@r473].
+For instance, species ranges' have been partitioned according to habitat discontinuities to capture adaptive processes [@r2].
+However, by partitioning features into different groups, it is not clear how well a given solution samples the original proxy variable, and ultimately, how well the solution captures the biodiversity process of interest.
+To overcome this limitation, Faith and Walker [-@r218] developed a decision support tool (\texttt{DIVERSITY}) to explicitly sample proxy variables to secure biodiversity processes.
+However, this tool is not widely used today, in part, because it cannot accommodate explicit targets on biodiversity patterns for multiple features.
 
 <!-- aims of this paper -->
-One of the key issues in reserve selection is the lack of a unifying decision support tool that can accommodate data on biodiversity patterns and processes.
-To begin to fill this gap, we present the \texttt{raptr R} package.
-This \texttt{R} package uses novel formulations of the reserve selection problem to provide decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes.
-Here, we aim to define the concepts that underpin this \texttt{R} package and explore its functionality.
+One of the key issues in reserve selection is the lack of unified a decision support tool that uses explicit and separate targets for biodiversity patterns and processes.
+To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{MARXAN} into new optimisation problems, and implemented them in the \texttt{raptr R} package.
+This \texttt{R} package provides decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes.
+Here, we aim to the provide an in-depth understanding of its inner workings and explore its functionality.
 
@@
 species, ecosystems).
-Spatial attributes are defined as the variation across the species' range that the prioritisation is required to sample.
+Spatial attributes are defined as the variation across the feature' range that the prioritisation is required to sample.
 They can be intrinsic (eg.
@@
 
-Each attribute is conceptualised as a space.
-This space is termed an "attribute space".
-Each planning unit is thought to occupy a single point inside each space.
+A set of attributes is conceptualised as a space (termed an "attribute space").
+Each planning unit may be thought to exist as a single point inside a given space.
 For example, a decision maker may require a prioritisation that represents populations along climatic gradients.
@@
 By associating planning units with climatic data, they can be mapped from geographic space to this environmental attribute space.
+In addition to planning units, attribute spaces also contain demand points.
 
-Demand points are points that also exist in an attribute space.
-They are designated by the decision maker to indicate regions of the attribute space that the decision maker wishes to represent in the prioritisation (see below for discussion on how demand points can be generated for real-world datasets).
-The amount of variation in the attribute space that a prioritisation secures is a function of the distance between each demand point and each selected planning unit in the attribute space.
+Demand points are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritisation (see below for discussion on how demand points can be generated for real-world data sets).
+The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space.
 The shorter the distances between the demand points and the planning units; the better the prioritisation is at securing the variation in the spatial attribute.
-To convert these amounts to a proportion--a meaningful unit for a decision maker--the distances between the selected planing units and the demand points are scaled by the distances between the demand points to the centroid of the demand points.
+To calculate the proportion of variation secured in a given solution, the distances are scaled by the distance between the demand points to their centroid (equivalent to calculating R$^2$ values for K-means clustering).
 In any attribute space there may exist points that are impossible (eg.
@@
 Thus a different set of demand points and weights are used for each attribute space and each feature.
-By placing demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritisations secure the feature in planning units with spatial attributes that are desirable for that feature.
+By using demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritisations secure the feature in planning units with spatial attributes that are desirable for that feature.
 
-To illustrate these concepts, consider an example conservation planning scenario example involving an attribute space and demand points.
-We wish to develop a prioritisation for a single species that has four populations in the study area.
-However, we can only afford to preserve three populations.
-We aim preserve the adaptive landscape of the species, and to achieve this, we will preserve populations inhabiting different environmental conditions.
-To describe environmental variation, we obtained data on the environmental conditions (rainfall (mm) and temperature ($^{\circ}$C)) where each population is found.
-These environmental data are then used to construct a two-dimensional environmental attribute space.
-Next, we generated demand points as equi-distant points between the range of values where the populations were found.
-By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify a suitable prioritisation (Fig.
+To illustrate these concepts, consider the following example: we wish to develop a prioritisation for a single species that has four populations.
+However, we can only afford to preserve three of the four populations.
+To ensure that we preserve a representative sample, we want to want preserve populations inhabiting different environmental conditions.
+We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population.
+We used this data to construct a two-dimensional environmental attribute space.
+Next, we generated demand points as equi-distant points inside this space.
+By comparing the distributions of demand points and populations in the attribute space, we can identify a suitable prioritisation (Fig.
 1).
-We can see that preserving both populations $A$ and $C$, effectively "doubles-up" on the same environmental characteristics, resulting in considerable redundancy in the solution.
+We can see that a solution that prioritises both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and would contain in considerable redundancy.
 Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$, $B$, and $D$.
-This example demonstrates how the inclusion of biodiversity processes can guide the reserve selection process.
+This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide reserve selection process .
 
-The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan} reserve selection problem and the uncapacitated facility location problems [@r451].
+The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan}, \texttt{DIVERSITY}, and the uncapacitated facility location problems [@r451].
 Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in the Supporting Information S1.
@@
 
-Targets are used to ensure that prioritisations adequately preserve each species.
+Targets are used to ensure that prioritisations adequately preserve each feature.
 Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient.
 Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$.
-Space-based targets ensure that a sufficient proportion of the intra-specific variation is secured.
+Space-based targets ensure that a representative sample of each feature is secured.
 Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $a \in A$.
-For convenience, these both types of targets are expressed as proportions in the \texttt{R} package.
+For convenience, all targets are expressed as proportions in the \texttt{R} package.
 
@@
 %
-\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
+\tau_{sa} &= \parbox{25em}{representation target for feature $f$ in attribute space $a$ according to the corresponding demand points} \tag*{eqn 1b}\\
 %
@@
 Y_{fsij} &= \begin{cases}
-		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
+		1, & \parbox{25em}{if demand point $i$ is used to represent planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
 		0, & \parbox{25em}{otherwise} \\
@@
 Linearised versions of the problems can be solved using commercial exact algorithm solvers.
-The \texttt{raptr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations.
+The \texttt{raptr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations (see Supporting Information S1 for formulation).
 These optimisation problems can then be solved to generate prioritisations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).
@@
 ### Methods
-We simulated a hypothetical study area with planning units arranged in $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig.
+We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig.
 2).
@@
 
-### Results
+### Uniformly distributed species
 
@@
 3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area.
-This configuration is an artifact of the method used to solve this particular instance of the reserve selection problem.
+This specific configuration is an artifact of the \texttt{Gurobi} solver used to solve the problem.
 In fact, for this species, all prioritisations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
@@
 In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-Note that the proportion here is negative because it performs worse than simply citing a planning unit in the centroid of the species' demand points.
+Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points.
 
@@
 Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.
+
+### Normally distributed species
 
@@
 3e--3g).
-The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species'--even for species without significant structure.
+The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species'--even for species without significant spatial structure.
+
+### Bimodally distributed species
 
@@
 We investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species.
-By preserving the populations in suitable habitats with different with environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63].
+By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63].
 We selected Queensland, Australia as the study area.
@@
 To do this, we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
-We used this method because it is trivially reproducible using freely available data.
+We used this method because it is entirely reproducible using freely available data.
 
@@
 S2).
-To achieve this, for each species, we generated random points inside the species range and at each point extracted the principal component values at that location.
-We then fit hyperbox kernels to the distribution of these points to characterise the realised niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231].
+We used the following method to ensure that the demand points reflected the core parts of the species' realised niches.
+For each species, we generated random points inside the species range and at each point extracted the principal component values at that location.
+We then fit hyperbox kernels to the distribution of principal component values to characterise the realised niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231].
 We then generated uniformly distributed points inside the species' kernels and estimated the density of the training points at the uniformly generated points.
@@
 The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species.
-The second solution was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based for each species.
+The second solution was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.
 
@@
 Yet it failed to achieve this for `r parsed.not.representative.space.held.names`.
-This result suggests that prioritisations generated using conventional methods may preserve a representative sample of most species realised niches (Figure S1).
-However, despite this, there may yet be species for which only a small fraction of their realised niche is preserved.
-By explictly using space targets, conservation planners can generate prioritisations that are guaranteed to capture a representative sample of species' niches.
+This result demonstrates that though solution generated using conventional methods may preserve a representative sample of some features (Figure S1), only through the use of explicit targets can decision makers be sure that solutions secure a representative sample for all features.
 
@@
 
-The binary genetic data was ordinated into a two-dimensional space that described the main different between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`).
-These values were then averaged to the planning unit level to describe the typical genetic characteristics of individuals in each planning unit (Figs.
+A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`; Figs.
 7c--7d).
-Planning units located near each other were found to contain individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
+Planning units located near each were contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
 In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
@@
 Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
-Both the simulated and case study species suggest that conservation planning exercises need to explicitly consider biodiversity processes during the reserve selection process to ensure they are captured.
+Both the simulated and case study species suggest that explicitly considering proxies for biodiversity processes can significantly alter prioritisations.
 
@@
 Ultimately, the space-based targets are set as a proportion based on the distribution of the demand-points.
-As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not be an effective prioritisation.
+As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points , then the optimal solution will not be an effective prioritisation.
 We therefore stress that decision makers must carefully consider which biodiversity processes need to be reflected in the prioritisation, and which spatial data can be used to represent these processes.
@@
 RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Excellence for Environmental Decisions (CEED) and the Landscape Ecology and Conservation Group (LEC) at The University of Queensland.
+This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
 
@@
 The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://cran.r-project.org/}).
-All data, code, and results are stored in an online repository (\url{https://www.github.com/paleo13/raptr-manuscript}) to permit replication and validation of this study.
+All data, code, and results are stored in an online repository (\url{https://www.github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
 

### 2016-10-13 code/rmarkdown/article.Rmd
+ assorted fixes

@@
 For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@r476].
-Additionally, migratory processes allow many species to access suitable habitat throughout the year [r477].
+Additionally, migratory processes allow many species to access suitable habitat throughout the year [@r477].
 On the other hand, evolutionary processes are important for biodiversity to persist over long time-scales.
@@
 bioclimatic, distribution, genetic, phylogenetic, and trait data; @r406; @r33; @r460; @r453].
-This is sparked significant interest in using this data to make better conservation decisions [@r473; @r9; @r448; @r2; @r68; @r470; @r469; @r247; @r488; @r468; @r407; @r61; @r475].
+This is sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r407; @r61; @r2; @r470; @r475; @r469].
 
 <!-- current methods for preserving biodiversity processes -->
-Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Marxan} and \texttt{Zonation}; @r22; @r429].
+Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Marxan} and \texttt{Zonation}; @r429; @r22].
 The cornerstone of these tools is that decision makers can set targets for a proportion of the geographic range of species, or ecosystems (termed features).
-To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step [@r9; @r448; @r247].
+To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step [@r9; @r247; @r448].
 These groups are partitioned based on a proxy variable that relates to a biodiversity process of interest [@r473].
@@
 <!-- aims of this paper -->
-One of the key issues in reserve selection is the lack of unified a decision support tool that uses explicit and separate targets for biodiversity patterns and processes.
-To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{MARXAN} into new optimisation problems, and implemented them in the \texttt{raptr R} package.
+One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets for biodiversity patterns and processes.
+To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{MARXAN} into new formulations of the reserve selection problem, and implemented them in the \texttt{raptr R} package.
 This \texttt{R} package provides decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes.
-Here, we aim to the provide an in-depth understanding of its inner workings and explore its functionality.
+Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore the functionality.
 
 # Methods
-## PROBLEM FORMULATION
+## OPTIMISATION PROBLEM FORMULATION
 Biodiversity features are defined as the entities that the prioritisation is required to preserve (eg.
@@
 Demand points are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritisation (see below for discussion on how demand points can be generated for real-world data sets).
-The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space.
+The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space (see below for mathematical expressions).
 The shorter the distances between the demand points and the planning units; the better the prioritisation is at securing the variation in the spatial attribute.
@@
 The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan}, \texttt{DIVERSITY}, and the uncapacitated facility location problems [@r451].
-Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in the Supporting Information S1.
+Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
 These formulations are named after the unreliable and reliable facility location formulations from which they are based upon [@r16].
@@
 Linearised versions of the problems can be solved using commercial exact algorithm solvers.
-The \texttt{raptr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations (see Supporting Information S1 for formulation).
+The \texttt{raptr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations (see Appendix S1 for formulation).
 These optimisation problems can then be solved to generate prioritisations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).

### 2016-10-13 code/rmarkdown/article.Rmd
+ minor ref fix

@@
 bioclimatic, distribution, genetic, phylogenetic, and trait data; @r406; @r33; @r460; @r453].
-This is sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r407; @r61; @r2; @r470; @r475; @r469].
+This is sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r475; @r469].
 

### 2016-11-02 code/rmarkdown/article.Rmd
+ use raptr version on cran

@@
 # load .rda file
-checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/final/results.rda')

### 2016-11-02 code/rmarkdown/figures.Rmd
+ use raptr version on cran

--- 2016-07-25 code/rmarkdown/figures.Rmd
@@
 # load .rda file
+checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
 try(session::restore.session('data/final/results.rda'))

### 2016-11-02 code/rmarkdown/supporting-information.Rmd
+ assorted fixes
+ use raptr version on cran

--- 2016-07-25 code/rmarkdown/supporting-information.Rmd
@@
 
-## Reliable formulation
+# Appendix S1: Reliable formulation
 The reliable formulation explicitly considers the probability that the planning units are inhabited.

### 2016-11-02 code/R/analysis/00-initialization.R

@@ -12,7 +12,7 @@ if (length(args)>0) {
 # set checkpoint
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-checkpoint::checkpoint('2016-07-25', R.version='3.3.1')
+checkpoint::checkpoint('2016-11-01', R.version='3.3.1')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 
 # load CRAN packages
@@ -37,6 +37,7 @@ library(tidyr)
 library(testthat)
 
 
+library(raptr)
 library(rgeos)
 library(rworldxtra)
 library(maptools)
@@ -64,22 +65,15 @@ library(ALA4R)
 
 
 if (!'spThin' %in% installed.packages()[,'Package'])
-	withr::with_libpaths(.libPaths()[1], devtools::install_github('paleo13/spThin', dependencies=NA))
+	withr::with_libpaths(.libPaths()[1], devtools::install_github('jeffreyhanson/spThin', dependencies=NA))
 library(spThin)
 
-# install raptr
-if (!'raptr' %in% installed.packages()[,'Package']) {
-	withr::with_libpaths(.libPaths()[1], install.packages(c('adehabitatLT', 'adehabitatHS', 'deldir', 'R.utils', 'geometry', 'KernSmooth', 'misc3d', 'multicool', 'fastcluster', 'rgdal', 'raster', 'PBSmapping', 'RJSONIO', 'R.methodsS3', 'R.oo', 'RgoogleMaps', 'png', 'gdalUtils', 'rgl', 'shape')))
-	withr::with_libpaths(.libPaths()[1], devtools::install_github('paleo13/raptr', dependencies=NA))
-}
-library(raptr)
-
 if (!'bayescanr' %in% installed.packages()[,'Package'])
-	withr::with_libpaths(.libPaths()[1], devtools::install_github('paleo13/bayescanr', dependencies=NA))
+	withr::with_libpaths(.libPaths()[1], devtools::install_github('jeffreyhanson/bayescanr', dependencies=NA))
 library(bayescanr)
 
 if (!'structurer' %in% installed.packages()[,'Package'])
-	withr::with_libpaths(.libPaths()[1], devtools::install_github('paleo13/structurer', dependencies=NA))
[…]

### 2016-11-02 code/R/analysis/01-load-data.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/00-initialization.rda')
 
 ### load simulation data

### 2016-11-02 code/R/analysis/02-simulations.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/01-load-data.rda')
 
 ## load parameters

### 2016-11-02 code/R/analysis/05-statistical-analysis.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/04-case-study-2.rda')
 load('data/intermediate/03-case-study-1.rda')
 load('data/intermediate/02-simulations.rda')

### 2016-11-02 code/R/analysis/06-exports.R

@@ -1,5 +1,5 @@
  ## load .rda
-checkpoint::checkpoint('2016-07-25', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/05-statistical-analysis.rda')
 
 ## save workspace

### 2016-11-02 code/R/analysis/00-initialization.R
+ add in Rcurl dep for ALA4R

@@ -58,7 +58,7 @@ library(rmarkdown)
 
 ## load github packages
 if (!'ALA4R' %in% installed.packages()[,'Package']) {
-	withr::with_libpaths(.libPaths()[1], install.packages('bitops'))
+  withr::with_libpaths(.libPaths()[1], c(install.packages('bitops', 'RCurl')))
 	withr::with_libpaths(.libPaths()[1], devtools::install_github('AtlasOfLivingAustralia/ALA4R', dependencies=NA))
 }
 library(ALA4R)

### 2016-11-02 code/R/analysis/00-initialization.R
+ explicitly call methods

@@ -102,8 +102,10 @@ panderOptions('knitr.auto.asis', FALSE)
 # set seed for reproducibility
 set.seed(500)
 
-# set default select method
+# set default method
 select <- dplyr::select
+rename <- dplyr::rename
+extract <- raster::extract
 
 ### Load functions
 for (x in dir(file.path('code', 'R', 'functions'), full.names=TRUE)) source(x)

### 2016-11-02 code/R/analysis/02-simulations.R

@@ -127,7 +127,7 @@ sim.spp.DF <- ldply(
 			prioritisation=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'),
 			amount.held=sapply(sim.prioritisations[[i]], function(s) {amount.held(s)[1]})*100,
 			space.held=sapply(sim.prioritisations[[i]], function(s) {space.held(s)[1]})*100,
-			score=sapply(sim.prioritisations[[i]], function(s) {score(s)[1]}),
+			score=sapply(sim.prioritisations[[i]], function(s) {raptr:::score.RapSolved(s)[1]}), # need to call method explicitly due to conflicts in NAMESPACE
 			n=sapply(sim.prioritisations[[i]], function(s) {sum(selections(s))})
 		)
 	}

### 2016-11-04 code/rmarkdown/article.Rmd
+ edits to ms + increase blm value

@@
 For example, representing ability for individuals to persist in a range of climatic conditions might be crucial for the long-term persistence of a species in a world where environmental change is accelerating.
-Yet current approaches overwhelmingly focus on securing a target proportion of the features' geographic range.
+Yet current approaches overwhelmingly focus on securing a target proportion of features' geographic ranges; little attention is directed towards targeting a representative sample of them.
 
@@
 an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
-Users also set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribtuion.
+Users also set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribution.
 We demonstrated the functionality of this package using a simulated study and two case studies.
-In each study, we generated solutions that secured a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compared them with solutions generated using both amount and space targets.
+In each study, we generated solutions that secured a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compared them to solutions generated using both amount and space targets.
 
@@
 We show that explicitly considering biodiversity processes in reserve selection can substantially change the solution.
-The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multi-model distributions in an attribute space.
-The case studies show that explicitly considering space-based targets for biodiversity processes can return solutions that could be far more effective in achieving long term biodiversity persistence and do not require a much greater area reserved than traditional approaches.
+The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multimodal distributions in an attribute space.
+The case studies show that explicitly considering space-based targets for biodiversity processes can result in solutions that could be far more effective in achieving long term biodiversity persistence and do not require a much greater area reserved than traditional approaches.
 
@@
 In recent decades, a wealth of data relevant to biodiversity processes has become available [eg.
-bioclimatic, distribution, genetic, phylogenetic, and trait data; @r406; @r33; @r460; @r453].
-This is sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r475; @r469].
+bioclimatic, distribution, genetic, phylogenetic, and trait data; @r33; @r460; @r478; @r453].
+This has sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r475; @r469].
 
@@
 Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Marxan} and \texttt{Zonation}; @r429; @r22].
-The cornerstone of these tools is that decision makers can set targets for a proportion of the geographic range of species, or ecosystems (termed features).
+The cornerstone of these tools is that they allow decision makers to set targets to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest).
 To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step [@r9; @r247; @r448].
@@
 However, by partitioning features into different groups, it is not clear how well a given solution samples the original proxy variable, and ultimately, how well the solution captures the biodiversity process of interest.
-To overcome this limitation, Faith and Walker [-@r218] developed a decision support tool (\texttt{DIVERSITY}) to explicitly sample proxy variables to secure biodiversity processes.
-However, this tool is not widely used today, in part, because it cannot accommodate explicit targets on biodiversity patterns for multiple features.
+This limitation is also present in some of the earliest reserve selection methods [eg.
+@r295], and to overcome it, Faith and Walker [-@r218] developed a decision support tool (\texttt{DIVERSITY}) to explicitly sample proxy variables to secure biodiversity processes.
+However, this tool is not widely used today, in part, because it cannot accommodate explicit targets on biodiversity patterns for multiple features (unlike and \texttt{Marxan} and \texttt{Zonation}).
 
@@
 One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets for biodiversity patterns and processes.
-To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{MARXAN} into new formulations of the reserve selection problem, and implemented them in the \texttt{raptr R} package.
+To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implemented them in the \texttt{raptr R} package.
 This \texttt{R} package provides decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes.
-Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore the functionality.
+Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.
 
@@
 However, we can only afford to preserve three of the four populations.
-To ensure that we preserve a representative sample, we want to want preserve populations inhabiting different environmental conditions.
+To ensure that we preserve a representative sample, we want to preserve populations inhabiting different environmental conditions.
 We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population.
@@
 Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$, $B$, and $D$.
-This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide reserve selection process .
+This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide the reserve selection process.
 
@@
 Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
-These formulations are named after the unreliable and reliable facility location formulations from which they are based upon [@r16].
-The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied when calculating the proportion of an attribute space sampled in a solution.
+These formulations are based on unreliable and reliable facility location problems and have been named accordingly [@r16].
+The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating the proportion of an attribute space sampled in a solution.
 All mathematical terms defined hereafter are described in Table S1.
-For convenience, the cardinality of sets will be denoted using the same symbol used to denote the variable.
+For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.
 
@@
 To assess the extent to which each feature is secured in a given prioritisation, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$.
-The level of fragmentation associated with a prioritisation is parametrised as the net exposed boundary length.
+The level of fragmentation associated with a prioritisation is parametrised as the total exposed boundary length (as in \texttt{Marxan}).
 Let the shared edges between each planning unit $j \in J$ and $k \in J$ be $e_{jk}$.
@@
 To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.
-Demand points with greater weight $\lambda_{fsi}$ are more important, and the optimal solution will be likely to select planning units close to highly weighted demand points.
+Demand points with greater weight $\lambda_{fsi}$ are more important, and the solutions that select planning units close to highly weighted demand points will be more favorable.
 As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.
@@
 Space-based targets ensure that a representative sample of each feature is secured.
-Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $a \in A$.
+Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
 For convenience, all targets are expressed as proportions in the \texttt{R} package.
 
-The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{sa}$ variables.
+The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.
 
@@
 %
-\tau_{sa} &= \parbox{25em}{representation target for feature $f$ in attribute space $a$ according to the corresponding demand points} \tag*{eqn 1b}\\
+\tau_{fs} &= \parbox{25em}{representation target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
 %
-B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
+B &= \text{boundary length modifier (BLM) to penalise fragmented solutions} \tag*{eqn 1c}\\
 \end{align*}
@@
 The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution.
-Generally, demand points are assigned to the closest selected planning units (unless particularly low space-based targets are used).
+Generally, demand points are represented by their closest selected planning units (unless particularly low space-based targets are used).
 
@@
 This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig.
-3c) contained the sample number of planning units as the solution with just the amount targets (Fig.
+3c) contained the same number of planning units as the solution with just the amount targets (Fig.
 3a).
@@
 
-The normally distributed species was simulated to contain a single core area where individuals are most prevalent and marginal areas where individuals are less likely to occur (Fig.
+The normally distributed species was simulated to contain a single core area where individuals were most prevalent, and marginal areas where individuals were less likely to occur (Fig.
 2b).
@@
 
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species distribution (Fig.
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig.
 3f).
@@
 3e--3g).
-The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species'--even for species without significant spatial structure.
+The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species--even for species without significant spatial structure.
 
@@
 While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-The addition of boundary length modifiers to just amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig.
+The addition of boundary length modifiers to amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig.
 3k).
@@
 Here we used space-based targets to generate a prioritisation securing a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and multilocus AFLP data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
+We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
 Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study.
@@
 
-We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles wthin individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
+We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
 These continuous variables described the main axes of genetic variation within the species.
@@
 7c--7d).
-Planning units located near each were contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
+Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
 In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
@@
 8a and 8c).
-This solution only sampled planning units that contained individuals belonging to one of the main two groups.
+This solution only sampled planning units that contained individuals belonging to one of the two main groups.
 On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups.
@@
 This package can accommodate intrinsic or extrinsic variation to the feature(s).
-For example, adaptation processes could be secured using environmental variation or genetic variation among loci under selection.
+For example, adaptation processes could be secured using environmental variation or genetic variation.
 Ecological processes, such as predator-prey interactions, could be secured by capturing intra-specific trophic variation.
@@
 The degree to which a prioritisation truly secures a representative sample of a feature depends on the attribute spaces and distribution of demand points chosen by the decision maker.
-Ultimately, the space-based targets are set as a proportion based on the distribution of the demand-points.
-As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points , then the optimal solution will not be an effective prioritisation.
+Ultimately, the space-based targets are set as a proportion based on the distribution of the demand points.
+As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points , then the optimal solution will not be effective.
 We therefore stress that decision makers must carefully consider which biodiversity processes need to be reflected in the prioritisation, and which spatial data can be used to represent these processes.
 Note that maximising one process can be detrimental to another.
-For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other, yet to maximise the connectivity of a prioritisation, reserves need to be closer to each other.
+For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other, yet to maximise connectivity, reserves may need to be closer together.
 To assist in the selection of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see the `make.DemandPoints` function).
-These routines essentially use the distribution of a feature in the attribute space to define a polygon.
-Demand points are then generated as random points within the polygon.
-A kernel is then fit to the distribution of the feature in the space [using @r231; @r450], and the demand points are weighted based on the estimated density of the feature at the demand points.
+These routines use kernels to automatically generate demand points based on the distribution of features in spatial data sets [using @r231; @r450].
 

### 2016-11-04 code/rmarkdown/supporting-information.Rmd
+ edits to ms + increase blm value

@@
 
-# Appendix S1: Reliable formulation
-The reliable formulation explicitly considers the probability that the planning units are inhabited.
+## Appendix S1
+The reliable formulation explicitly considers the probability that planning units are inhabited by features.
 As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented.
@@
 The distance variables associated with this imaginary planning unit $d_{fsiJ}$ denote the loss of biological value associated with failing to secure a representative sample of feature $f$ in attribute space $s$.
-However, the $d$ variables are in distance units which are meaningless in this context.
+However, the $d$ variables are in distance units which effectively meaningless in this context.
 Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).
@@
 Cui et al.
-[-@r16] found that $R=5$ yields similar solutions to $R=J$ when $J >> 5$.
-However, in most costs the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.
+[-@r16] found that $R=5$ yields similar solutions to $R >> 5$.
+However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.
 
-The control variables for the reliable formulation are the $B$ (eqn 1a), $T_{s}$ (eqn 1b), $\tau_{sa}$ (eqn 1c), $R$, and $M$ variables.
+The control variables for the reliable formulation are the $B$ (eqn 1a), $T_{s}$ (eqn 1b), $\tau_{fs}$ (eqn 1c), $R$, and $M$ variables.
 
@@
 [-@r426].
-Second, the expression $P_{fsijr} Y_{jsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
+Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
 [-@r16].
@@
 	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
-	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$.',
-	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$.',
-	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$.',
-	'$T_f$', 'amount target for feature $f$.',
-	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$.',
-	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0).',
-	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units.'
+	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$',
+	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$',
+	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$',
+	'$T_f$', 'amount target for feature $f$',
+	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$',
+	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)',
+	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units'
 )

### 2016-11-11 code/rmarkdown/article.Rmd
+ added feedback after talk with Hugh

@@
 An underlying aim in conservation planning is to maximise the long-term persistence of biodiversity.
-To fulfill this aim, the processes that sustain biodiversity must be conserved.
-Many biodiversity processes can be preserved by conserving a representative sample of each biodiversity feature (species or ecosystem).
-For example, representing ability for individuals to persist in a range of climatic conditions might be crucial for the long-term persistence of a species in a world where environmental change is accelerating.
-Yet current approaches overwhelmingly focus on securing a target proportion of features' geographic ranges; little attention is directed towards targeting a representative sample of them.
+To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved.
+One way to conserve such processes at the feature level (eg.
+species, ecosystem) is to preserve a representative sample of the physical attributes that underpin them across the feature's geographic distribution.
+For example, preserving the adaptive processes currently acting on a species might be crucial for its long-term persistence in a world where environmental change is accelerating.
+By conserving individuals with the ability to persist in a range of climatic conditions--physical attributes associated with adaptation--protected areas can foster evolutionary processes.
+Yet current approaches overwhelmingly focus on securing a target proportion of each features' geographic range; little attention is directed towards targeting a representative sample of them.
 
 2.
-To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as area protection targets.
-Users set \"space targets\" to secure a representative sample of each feature within an attribute space that reflects a biodiversity process (eg.
-an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
-Users also set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribution.
-We demonstrated the functionality of this package using a simulated study and two case studies.
-In each study, we generated solutions that secured a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compared them to solutions generated using both amount and space targets.
+To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as more typical area protection targets.
+Users set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribution.
+Additionally, users set \"space targets\" to ensure that solutions also secure a representative sample of an evolutionary or ecologically relevant attribute (space) across a feature's geographic distribution (eg.
+an attribute expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
+We demonstrate the functionality of this package using simulations and two case studies.
+We generate solutions that secure a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compare them with solutions generated using both amount and space targets.
 
 3.
-We show that explicitly considering biodiversity processes in reserve selection can substantially change the solution.
-The simulations suggest that including representativeness in conservation planning is important where biodiversity features have multimodal distributions in an attribute space.
-The case studies show that explicitly considering space-based targets for biodiversity processes can result in solutions that could be far more effective in achieving long term biodiversity persistence and do not require a much greater area reserved than traditional approaches.
+We show that markedly different solutions emerge when explicitly considering within feature representation in reserve selection.
+Our simulations suggest that including representativeness in conservation planning is important where biodiversity features have multimodal distributions in an attribute space.
+The case studies show that explicitly setting space-based targets can result solutions that preserve a much more representative sample of features.
+Additionally, the case studies suggest that securing a representative sample of features may not require a much  greater area of reservation than traditional approaches.
 
@@
 bioclimatic, distribution, genetic, phylogenetic, and trait data; @r33; @r460; @r478; @r453].
-This has sparked significant interest in using this data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r475; @r469].
+This has sparked significant interest in using these data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r479; @r475; @r469].
 
 <!-- current methods for preserving biodiversity processes -->
-Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Marxan} and \texttt{Zonation}; @r429; @r22].
-The cornerstone of these tools is that they allow decision makers to set targets to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest).
-To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step [@r9; @r247; @r448].
-These groups are partitioned based on a proxy variable that relates to a biodiversity process of interest [@r473].
-For instance, species ranges' have been partitioned according to habitat discontinuities to capture adaptive processes [@r2].
-However, by partitioning features into different groups, it is not clear how well a given solution samples the original proxy variable, and ultimately, how well the solution captures the biodiversity process of interest.
-This limitation is also present in some of the earliest reserve selection methods [eg.
-@r295], and to overcome it, Faith and Walker [-@r218] developed a decision support tool (\texttt{DIVERSITY}) to explicitly sample proxy variables to secure biodiversity processes.
-However, this tool is not widely used today, in part, because it cannot accommodate explicit targets on biodiversity patterns for multiple features (unlike and \texttt{Marxan} and \texttt{Zonation}).
+Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22].
+These tools allow decision makers to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest).
+To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step.
+The features are partitioned based on an attribute variable that relates to the biodiversity process of interest [@r9; @r247; @r448].
+For instance, species distributions can be divided according to habitat discontinuities to capture adaptive processes [@r2].
+However, using this method is challenging because biodiversity data are often continuous and multi-dimensional.
+
+<!-- issues with partitioning features -->
+After partitioning features based on continuous variables and obtaining a solution, it is not clear how well the solution samples the original continuous variables.
+For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution samples 10 % of each sub-feature, how much of the original continuous variation does the solution sample?
+This is not immediately clear.
+If the solution sites reserves evenly along the original variable then it may sample far more than 10 % of the original variation.
+Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning continuous geographic variation based on political boundaries; @r483], then it may sample much less of the original variation.
+This limitation has not arisen recently.
+In fact, it dates back to some of the earliest reserve selection methods [eg.
+@r295].
+
+<!-- diversity method for conservation -->
+During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritisations that secure a representative sample of continuous environmental variation across a study area.
+Although this method has been pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
+Recent work has built on this reserve selection method by solving problems with more advanced optimisation routines [@r482].
+However, \texttt{DIVERSITY}-based reserve selection methods are not widely used today, in part, because they do not accommodate explicit amount-based targets for multiple features (_cf_.
+\texttt{Marxan} and \texttt{Zonation}).
 
 <!-- aims of this paper -->
-One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets for biodiversity patterns and processes.
-To begin to fill this gap, we united the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implemented them in the \texttt{raptr R} package.
-This \texttt{R} package provides decision makers with the tools to generate prioritisations that preserve both biodiversity patterns and processes.
+One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets to obtain solutions that secure an adequate amount of habitat for each feature and a representative sample of each feature.
+To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
+This \texttt{R} package provides decision makers with the tools to generate prioritisations based on data that relates to biodiversity patterns and processes.
 Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.
@@
 Spatial attributes are defined as the variation across the feature' range that the prioritisation is required to sample.
-They can be intrinsic (eg.
-genetic or trait variation) or extrinsic (eg.
-environmental variation) to the feature.
-These attributes should be related to the biodiversity processes that the decision maker aims to preserve.
+They can be intrinsic variation (eg.
+genetic or trait) variation that is extrinsic to the feature (eg.
+environmental conditions).
+There should, however, be a reasonable underlying hypothesis that relates this variation to the biodiversity processes that the decision maker aims to preserve.
 
@@
 
-Demand points are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritisation (see below for discussion on how demand points can be generated for real-world data sets).
+Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritisation (see below for discussion on how demand points can be generated for real-world data sets).
 The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space (see below for mathematical expressions).
@@
 1).
-We can see that a solution that prioritises both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and would contain in considerable redundancy.
-Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$, $B$, and $D$.
+We can see that a solution that prioritises both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and constitute considerable redundancy if both selected.
+Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$ (or $C$), $B$, and $D$.
 This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide the reserve selection process.
 
-The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on a combination of the \texttt{Marxan}, \texttt{DIVERSITY}, and the uncapacitated facility location problems [@r451].
+The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and the uncapacitated facility location problems [@r451].
 Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
@@
 \texttt{Marxan}), and solutions using amount and space targets using the unreliable formulation.
-By comparing these solutions, we can guage the benefits of explicitly including space targets in reserve selection.
+By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection.
 All analyses were performed in \texttt{R} [version `r getRversion()`; @r192].
@@
 Yet it failed to achieve this for `r parsed.not.representative.space.held.names`.
-This result demonstrates that though solution generated using conventional methods may preserve a representative sample of some features (Figure S1), only through the use of explicit targets can decision makers be sure that solutions secure a representative sample for all features.
+This result demonstrates that though solutions generated using conventional methods may preserve a representative sample of some features (Figure S1), only through the use of explicit targets can decision makers be sure that solutions secure a representative sample for all features.
 
@@
 To maximise the long-term persistence of biodiversity, decision makers need to identify prioritisations that preserve existing patterns of biodiversity and the processes that support them.
-To achieve this, conservation planners need a decision support tool that can explicitly accommodate biodiversity patterns and processes.
+To achieve this, conservation planners need a decision support tool that can explicitly accommodate data on biodiversity patterns and processes.
 Here, we developed the \texttt{raptr R} package to fill this void.
-By exploring the functionality of this package using several simulated and case study species, we found that including space-based targets can result in substantially better solutions.
+By exploring the functionality of this package using several simulated and case study species, we found that including space-based targets can result in substantially different solutions.
 
@@
 RAF has an Australian Research Council Future Fellowship.
-This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
+This work was supported by the Centre for Biodiversity and Conservation Science (CBCS).
 

### 2016-11-11 code/rmarkdown/article.Rmd
+ further changes based on hugh's comments

@@
 4.
-The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritisations that are both adequate and representative, resulting in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
+The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritisations that secure an adequate and representative sample of features.
+Prioritisations that fulfill that , resulting in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
 
@@
 
-<!-- conservation and reserve selection -->
 Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255].
@@
 Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].
-However, despite the importance of biodiversity processes, they are only rarely used to guide conservation planning.
 
-<!-- biodiversity processes are important -->
-Prioritisations must capture the ecological and evolutionary processes that sustain biodiversity [@r3; @r255; @r9].
-Ecological processes are important for biodiversity to persist over short time-scales.
-For instance, predator-prey processes can mediate ecosystem structure and function through top-down control [@r476].
-Additionally, migratory processes allow many species to access suitable habitat throughout the year [@r477].
-On the other hand, evolutionary processes are important for biodiversity to persist over long time-scales.
-Adaptive evolutionary processes foster resilience against environmental change [@r3; @r43], and neutral evolutionary processes maintain genetic diversity [@r5; @r67; @r63].
-In recent decades, a wealth of data relevant to biodiversity processes has become available [eg.
-bioclimatic, distribution, genetic, phylogenetic, and trait data; @r33; @r460; @r478; @r453].
-This has sparked significant interest in using these data to make better conservation decisions [@r68; @r9; @r473; @r247; @r448; @r468; @r2; @r407; @r61; @r470; @r479; @r475; @r469].
-
-<!-- current methods for preserving biodiversity processes -->
 Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22].
 These tools allow decision makers to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest).
-To incorporate some information on biodiversity processes, current methods involve partitioning the features into different groups as a pre-processing step.
-The features are partitioned based on an attribute variable that relates to the biodiversity process of interest [@r9; @r247; @r448].
-For instance, species distributions can be divided according to habitat discontinuities to capture adaptive processes [@r2].
+One method to incorporate data on the ecological and environmental processes that sustain features is to conserve a representative sample of the physical attributes that underpin these processes.
+To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448].
+For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritisations that promote adaptive processes [@r2].
 However, using this method is challenging because biodiversity data are often continuous and multi-dimensional.
 
-<!-- issues with partitioning features -->
-After partitioning features based on continuous variables and obtaining a solution, it is not clear how well the solution samples the original continuous variables.
-For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution samples 10 % of each sub-feature, how much of the original continuous variation does the solution sample?
-This is not immediately clear.
-If the solution sites reserves evenly along the original variable then it may sample far more than 10 % of the original variation.
+Simply partitioning features based on a continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable.
+For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution preserve 10 % of the geographic range of each sub-feature, how much of the original continuous variation does the solution represent?
+The answer is not immediately clear.
+If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation.
 Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning continuous geographic variation based on political boundaries; @r483], then it may sample much less of the original variation.
+Conventional reserve selection problems cannot differentiate between either of these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce fragmentation.
 This limitation has not arisen recently.
@@
 
-<!-- diversity method for conservation -->
 During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritisations that secure a representative sample of continuous environmental variation across a study area.
-Although this method has been pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
+Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
 Recent work has built on this reserve selection method by solving problems with more advanced optimisation routines [@r482].
@@
 
-<!-- aims of this paper -->
 One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets to obtain solutions that secure an adequate amount of habitat for each feature and a representative sample of each feature.
 To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
-This \texttt{R} package provides decision makers with the tools to generate prioritisations based on data that relates to biodiversity patterns and processes.
+This \texttt{R} package provides decision makers with the tools to generate prioritisations based on data that relate to biodiversity patterns and processes.
 Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.
@@
 A set of attributes is conceptualised as a space (termed an "attribute space").
-Each planning unit may be thought to exist as a single point inside a given space.
+Each planning unit may be thought to exist as a single point inside a given space ().
 For example, a decision maker may require a prioritisation that represents populations along climatic gradients.

### 2016-11-18 code/rmarkdown/article.Rmd
added in Richs, Hughs, and Jonathan's comments

@@
 author: |
-  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^1$, Richard A. Fuller$^1$
+  | Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^{1,3}$, Richard A. Fuller$^1$
   | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
   | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
+  | $^3$The Nature Conservancy, South Brisbane, QLD, Australia
   | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
@@
 By conserving individuals with the ability to persist in a range of climatic conditions--physical attributes associated with adaptation--protected areas can foster evolutionary processes.
-Yet current approaches overwhelmingly focus on securing a target proportion of each features' geographic range; little attention is directed towards targeting a representative sample of them.
+Despite this, current approaches overwhelmingly focus on securing a target proportion of each features' geographic range and little attention has been directed towards targeting a representative sample of them.
 
 2.
-To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing biodiversity processes as well as more typical area protection targets.
+To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
 Users set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribution.
-Additionally, users set \"space targets\" to ensure that solutions also secure a representative sample of an evolutionary or ecologically relevant attribute (space) across a feature's geographic distribution (eg.
-an attribute expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
+Additionally, users set attribute \"space targets\" to ensure that reserve systems also secure a representative sample of evolutionary or ecologically relevant attributes (space) across a feature's geographic distribution (eg.
+an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
 We demonstrate the functionality of this package using simulations and two case studies.
-We generate solutions that secure a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compare them with solutions generated using both amount and space targets.
+We generate solutions that secure a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compare them with solutions generated using both amount and attribute space targets.
 
@@
 Our simulations suggest that including representativeness in conservation planning is important where biodiversity features have multimodal distributions in an attribute space.
-The case studies show that explicitly setting space-based targets can result solutions that preserve a much more representative sample of features.
+The case studies show that explicitly setting space-based targets can result in solutions that preserve a much more representative sample of features.
 Additionally, the case studies suggest that securing a representative sample of features may not require a much  greater area of reservation than traditional approaches.
@@
 The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritisations that secure an adequate and representative sample of features.
-Prioritisations that fulfill that , resulting in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
+Prioritisations that achieve this are likely to result in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
 
@@
 populations, species, ecosystems), but also crucially the processes that sustain them.
-One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r440].
+One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r484; @r440].
 Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465].
@@
 Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22].
-These tools allow decision makers to obtain solutions that secure a proportion of the geographic range of biological features (populations, species, or ecosystems of conservation interest).
-One method to incorporate data on the ecological and environmental processes that sustain features is to conserve a representative sample of the physical attributes that underpin these processes.
+Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of conservation by setting targets.
+One method to incorporate data on the ecological and environmental processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes.
 To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448].
@@
 
-Simply partitioning features based on a continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable.
+Simply partitioning features based on continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable.
 For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution preserve 10 % of the geographic range of each sub-feature, how much of the original continuous variation does the solution represent?
@@
 If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation.
-Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning continuous geographic variation based on political boundaries; @r483], then it may sample much less of the original variation.
-Conventional reserve selection problems cannot differentiate between either of these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce fragmentation.
-This limitation has not arisen recently.
+Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning continuous geographic variation based on political boundaries; @r483], then it might sample much less of the original variation.
+Conventional reserve selection problems cannot differentiate between these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce cost or fragmentation.
+This limitation has been known of quite some time.
 In fact, it dates back to some of the earliest reserve selection methods [eg.
@@
 Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
-Recent work has built on this reserve selection method by solving problems with more advanced optimisation routines [@r482].
-However, \texttt{DIVERSITY}-based reserve selection methods are not widely used today, in part, because they do not accommodate explicit amount-based targets for multiple features (_cf_.
-\texttt{Marxan} and \texttt{Zonation}).
+Recent work has built on this reserve selection method by solving problems with more advanced optimisation algorithms [@r482].
+However, \texttt{DIVERSITY} has limited utility for species-based conservation planning, in part, because it cannot accommodate multiple biological features.
+Additionally, unlike \texttt{Marxan} and \texttt{Zonation}, users cannot set intuitive targets to obtain cost effective solutions.
 
-One of the key issues in reserve selection is the lack of a unified decision support tool that uses explicit and separate targets to obtain solutions that secure an adequate amount of habitat for each feature and a representative sample of each feature.
+Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of each feature.
 To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
@@
 # Methods
-## OPTIMISATION PROBLEM FORMULATION
+## PROBLEM FORMULATION
 Biodiversity features are defined as the entities that the prioritisation is required to preserve (eg.
@@
 A set of attributes is conceptualised as a space (termed an "attribute space").
-Each planning unit may be thought to exist as a single point inside a given space ().
 For example, a decision maker may require a prioritisation that represents populations along climatic gradients.
-To achieve this, the decision maker might use an "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm).
+To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm).
 Any given combination of temperature and precipitation may be conceived as a point in this environmental space.
-By associating planning units with climatic data, they can be mapped from geographic space to this environmental attribute space.
+Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given space.
+By associating the planning units with climatic data, and calculating a representative statistic for each planning unit (eg.
+mean), they can be mapped from geographic space to this environmental attribute space.
 In addition to planning units, attribute spaces also contain demand points.
@@
 
-The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and the uncapacitated facility location problems [@r451].
+The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
 Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
@@
 [-@r426].
-Linearised versions of the problems can be solved using commercial exact algorithm solvers.
+Linearised versions of the problems can be solved using commercial exact algorithms.
 The \texttt{raptr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations (see Appendix S1 for formulation).
@@
 The \texttt{raptr} R package provides a unified approach to reserve selection.
-One of the key advantages of this package is that it is general enough to incorporate any spatial variable as an attribute space.
-This package can accommodate intrinsic or extrinsic variation to the feature(s).
-For example, adaptation processes could be secured using environmental variation or genetic variation.
-Ecological processes, such as predator-prey interactions, could be secured by capturing intra-specific trophic variation.
-Additionally, different attribute spaces with different targets can be specified for different species.
-As long as the variation can be described using Euclidean distances, the \texttt{R} package can be used to obtain a representative sample (note that data may require transformation or dimension reduction to meet this assumption).
-This \texttt{R} package provides decision makers with the tools to generate prioritisations that secure both biodiversity patterns and the processes that maintain biodiversity.
+This \texttt{R} package provides decision makers with the tools to generate prioritisations that secure intra-specific as well as as inter-specific biodiversity patterns.
 Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
-Both the simulated and case study species suggest that explicitly considering proxies for biodiversity processes can significantly alter prioritisations.
+Both the simulated and case study species suggest that explicitly considering representation as well as amount targets can significantly alter prioritisations.
 
-The degree to which a prioritisation truly secures a representative sample of a feature depends on the attribute spaces and distribution of demand points chosen by the decision maker.
+One of the key advantages of this package is that it is general enough to incorporate any spatially explicit variable as an attribute space.
+This package can accommodate intrinsic or extrinsic variation in the feature(s).
+For example, adaptation processes could be secured using environmental variation [eg.
+@r247].
+Predator-prey processes could be conserved by capturing the overlapping distributions of predators and each of their prey species [eg.
+@r485].
+Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserve so there is equitable access to parks by people [eg.
+@r483].
+As long as the variation can be described using Euclidean distances--which could be achieved through transformation or dimension reduction techniques--the \texttt{R} package can be used to obtain a representative sample.
+
+The degree to which a prioritisation truly secures a representative sample of a feature depends on the attribute spaces and the distribution of demand points chosen by the decision maker.
 Ultimately, the space-based targets are set as a proportion based on the distribution of the demand points.
-As a consequence, if the decision maker uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points , then the optimal solution will not be effective.
-We therefore stress that decision makers must carefully consider which biodiversity processes need to be reflected in the prioritisation, and which spatial data can be used to represent these processes.
-Note that maximising one process can be detrimental to another.
+As a consequence, if the planner uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not effectively represent the decision maker's objectives.
+We therefore stress that planners must carefully consider which biodiversity processes need to be reflected in the prioritisation, and which spatial data can be used to represent these processes.
+Note that maximising for one attribute space can result in trading off against another.
 For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other, yet to maximise connectivity, reserves may need to be closer together.
-To assist in the selection of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see the `make.DemandPoints` function).
+Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space.
+To assist the generation of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see the `make.DemandPoints` function).
 These routines use kernels to automatically generate demand points based on the distribution of features in spatial data sets [using @r231; @r450].
@@
 The formulation requires spatially comprehensive data to map attribute spaces to planning units.
-For instance, in both case-studies, all planning units occupied by the species were associated with coordinates in the attribute spaces.
-However, most real-world data sets are patchy--some planning units will be occupied by species for which attribute space data is not available.
-To use such patchy data with the \texttt{raptr R} package, the gaps in the data must first be filled in.
-Spatially explicit models [eg.
-krigging or generalised dissimilarity models; @r467; @r466] could be used to estimate data for locations that are missing from the data set as a pre-processing step.
+However, many real-world data sets are patchy.
+For example, due to constrained resources, genetic data may have only been collected in some planning units.
+To use such patchy data with the \texttt{raptr R} package, the planning units that are missing data must either be omitted (as was done in the second case-study), or their missing values must be estimated.
+Missing data could be estimated using spatially explicit models [eg.
+krigging or generalised dissimilarity models; @r467; @r466].
 This approach has been successfully applied to a range of biological data sets [eg.
@@
 To maximise the long-term persistence of biodiversity, decision makers need to identify prioritisations that preserve existing patterns of biodiversity and the processes that support them.
-To achieve this, conservation planners need a decision support tool that can explicitly accommodate data on biodiversity patterns and processes.
+To achieve this, conservation planners need a decision support tool that can deliver cost-effective prioritisations that secure an adequate amount of a representative sample of biodiversity features.
 Here, we developed the \texttt{raptr R} package to fill this void.
@@
 JOH is funded by an Australian Postgraduate Award (APA) scholarship.
-RAF has an Australian Research Council Future Fellowship.
+RAF and HPP have Australian Research Council Fellowships.
 This work was supported by the Centre for Biodiversity and Conservation Science (CBCS).

### 2016-11-23 code/rmarkdown/article.Rmd
+ add in readmes + compile manuscript into single pdf and single docx

@@
 ---
-title: "raptr: Representative and Adequate Prioritisation Toolkit in R"
+title: "raptr: Representative and Adequate Prioritization Toolkit in R"
 author: |
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: ["conservation", "biodiversity", "MILP", "reserve-selection", "optimisation", "protected area"]
+tags: ["conservation", "biodiversity", "MILP", "reserve-selection", "optimization", "protected area"]
 abstract: "
@@
 1.
-An underlying aim in conservation planning is to maximise the long-term persistence of biodiversity.
+An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity.
 To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved.
@@
 2.
-To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritisation Toolkit in R} to guide reserve selection using explicit targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
-Users set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the feature's geographic distribution.
-Additionally, users set attribute \"space targets\" to ensure that reserve systems also secure a representative sample of evolutionary or ecologically relevant attributes (space) across a feature's geographic distribution (eg.
+To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritization Toolkit in R} to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
+Users set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the features' geographic distributions.
+Additionally, users set attribute \"space targets\" to ensure that reserve systems also secure a representative sample of ecologically or evolutionary relevant attributes (space) across the features' geographic distributions (eg.
 an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
@@
 4.
-The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritisations that secure an adequate and representative sample of features.
-Prioritisations that achieve this are likely to result in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
+The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritizations that secure an adequate and representative sample of features.
+Prioritizations that achieve this are likely to result in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
 
@@
 
-```{r, include=FALSE}
-# set wd
-knitr::opts_knit$set(root.dir=normalizePath('../..'))
-```
-
-```{r, include=FALSE}
-## initialization
-# load .rda file
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
-session::restore.session('data/final/results.rda')
-
-## preliminary processing
-for (i in seq_len(nlayers(cs1.bioclim.RAST)))
-  cs1.bioclim.RAST@layers[[i]]@file@name <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])
-
-```
-
-<!--  title page -->
-Short running title: Representative and adequate prioritisations \newline
+Short running title: Representative and adequate prioritizations \newline
 Abstract word count: XXX / XXX  
@@
 Number of equations: XXX
-\newpage
 
-Perhaps the most fundamental aim of conservation is to maximise the long-term persistence of biodiversity [@r423; @r255].
-To achieve this, conservation actions must preserve biodiversity patterns (eg.
-populations, species, ecosystems), but also crucially the processes that sustain them.
-One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r484; @r440].
-Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465].
-Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].
+\clearpage 
+\input{./text.tex}
 
-Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22].
-Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of conservation by setting targets.
-One method to incorporate data on the ecological and environmental processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes.
-To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448].
-For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritisations that promote adaptive processes [@r2].
-However, using this method is challenging because biodiversity data are often continuous and multi-dimensional.
+\clearpage 
+\input{./figures.tex}
 
-Simply partitioning features based on continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable.
-For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution preserve 10 % of the geographic range of each sub-feature, how much of the original continuous variation does the solution represent?
-The answer is not immediately clear.
-If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation.
-Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning continuous geographic variation based on political boundaries; @r483], then it might sample much less of the original variation.
-Conventional reserve selection problems cannot differentiate between these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce cost or fragmentation.
-This limitation has been known of quite some time.
-In fact, it dates back to some of the earliest reserve selection methods [eg.
-@r295].
-
-During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritisations that secure a representative sample of continuous environmental variation across a study area.
-Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
-Recent work has built on this reserve selection method by solving problems with more advanced optimisation algorithms [@r482].
-However, \texttt{DIVERSITY} has limited utility for species-based conservation planning, in part, because it cannot accommodate multiple biological features.
-Additionally, unlike \texttt{Marxan} and \texttt{Zonation}, users cannot set intuitive targets to obtain cost effective solutions.
-
-Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of each feature.
-To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
-This \texttt{R} package provides decision makers with the tools to generate prioritisations based on data that relate to biodiversity patterns and processes.
-Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.
-
-# Methods
-## PROBLEM FORMULATION
-Biodiversity features are defined as the entities that the prioritisation is required to preserve (eg.
-species, ecosystems).
-Spatial attributes are defined as the variation across the feature' range that the prioritisation is required to sample.
-They can be intrinsic variation (eg.
-genetic or trait) variation that is extrinsic to the feature (eg.
-environmental conditions).
-There should, however, be a reasonable underlying hypothesis that relates this variation to the biodiversity processes that the decision maker aims to preserve.
-
-A set of attributes is conceptualised as a space (termed an "attribute space").
-For example, a decision maker may require a prioritisation that represents populations along climatic gradients.
-To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm).
-Any given combination of temperature and precipitation may be conceived as a point in this environmental space.
-Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given space.
-By associating the planning units with climatic data, and calculating a representative statistic for each planning unit (eg.
-mean), they can be mapped from geographic space to this environmental attribute space.
-In addition to planning units, attribute spaces also contain demand points.
-
-Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritisation (see below for discussion on how demand points can be generated for real-world data sets).
-The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space (see below for mathematical expressions).
-The shorter the distances between the demand points and the planning units; the better the prioritisation is at securing the variation in the spatial attribute.
-To calculate the proportion of variation secured in a given solution, the distances are scaled by the distance between the demand points to their centroid (equivalent to calculating R$^2$ values for K-means clustering).
-In any attribute space there may exist points that are impossible (eg.
-mean annual rainfall -5 mm), or do not occur in the study area (eg.
-mean annual temperature 30$^{\circ}$C in Antarctica).
-Additionally, there may be some regions that are desirable for some features and undesirable for others (eg.
-conditions known to be outside the physiological tolerance of certain species).
-Thus a different set of demand points and weights are used for each attribute space and each feature.
-By using demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritisations secure the feature in planning units with spatial attributes that are desirable for that feature.
-
-To illustrate these concepts, consider the following example: we wish to develop a prioritisation for a single species that has four populations.
-However, we can only afford to preserve three of the four populations.
-To ensure that we preserve a representative sample, we want to preserve populations inhabiting different environmental conditions.
-We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population.
-We used this data to construct a two-dimensional environmental attribute space.
-Next, we generated demand points as equi-distant points inside this space.
-By comparing the distributions of demand points and populations in the attribute space, we can identify a suitable prioritisation (Fig.
-1).
-We can see that a solution that prioritises both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and constitute considerable redundancy if both selected.
-Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$ (or $C$), $B$, and $D$.
-This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide the reserve selection process.
-
-The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
-Although the \texttt{raptr R} package utlises two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
-These formulations are based on unreliable and reliable facility location problems and have been named accordingly [@r16].
-The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating the proportion of an attribute space sampled in a solution.
-All mathematical terms defined hereafter are described in Table S1.
-For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.
-
-Define $F$ to be the set of features one wishes to conserve (indexed by $f$).
-Let $J$ be a set of planning units (indexed by $j$).
-Also, let $A_j$ denote the area, and $C_j$ denote the cost of preserving planning unit $j \in J$.
-To assess the extent to which each feature is secured in a given prioritisation, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$.
-The level of fragmentation associated with a prioritisation is parametrised as the total exposed boundary length (as in \texttt{Marxan}).
-Let the shared edges between each planning unit $j \in J$ and $k \in J$ be $e_{jk}$.
-
-Let $S$ denote a set of attribute spaces (indexed by $s$).
-Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$.
-Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$.
-Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$.
-Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$.
-To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.
-Demand points with greater weight $\lambda_{fsi}$ are more important, and the solutions that select planning units close to highly weighted demand points will be more favorable.
-As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.
-
-Targets are used to ensure that prioritisations adequately preserve each feature.
-Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient.
-Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$.
-Space-based targets ensure that a representative sample of each feature is secured.
-Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
-For convenience, all targets are expressed as proportions in the \texttt{R} package.
-
-The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.
-
-\begin{align*}
-T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
-%
-\tau_{fs} &= \parbox{25em}{representation target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
-%
-B &= \text{boundary length modifier (BLM) to penalise fragmented solutions} \tag*{eqn 1c}\\
-\end{align*}
-
-The decision variables are the $X_j$ and $Y_{fsij}$ variables.
-
-\begin{align*}
-X_j
-  &= \begin{cases}
-    1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
-    0, & \parbox{25em}{otherwise} \\
-  \end{cases} \\
-%
-Y_{fsij} &= \begin{cases}
-    1, & \parbox{25em}{if demand point $i$ is used to represent planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
-    0, & \parbox{25em}{otherwise} \\
-  \end{cases} \\
-\end{align*}
-
-Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is assigned to a selected planning unit $J$ where $X_j = 1$.
-The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution.
-Generally, demand points are represented by their closest selected planning units (unless particularly low space-based targets are used).
-
-The unreliable formulation (URAP) is a defined as a multi-objective optimisation problem.
-
-\begin{align*}
-& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
-%
-& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
-%
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
-& & & & & 0 \leq s \leq S-1\\
-%
-& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
-& & & & & 0 \leq s \leq S-1, \\
-& & & & & 0 \leq i \leq I-1\\
-%
-& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
-& & & & & 0 \leq s \leq S-1, \\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J-1\\
-%
-& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1\\
-%
-\end{align*}
-
-The objective function (eqn 3a) determines the utility of a given prioritisation: a combination of the total cost of a prioritisation and how fragmented it is.
-Constraints (eqn 3b) ensure that all the amount-based targets are met.
-Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space.
-For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$).
-This total weighted distance is then scaled by the inherent variation in the demand points ($\lambda_{fsi} \delta_{fsi}$).
-The resulting fraction yields a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
-The constraints ensure that this proportion must be equal to or greater than the space-based target.
-Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point.
-Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units.
-Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.
-
-## OPTIMISATION
-The unreliable formulation is non-linear.
-However, the non-linear components can be linearised using existing techniques.
-The expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al.
-[-@r426].
-Linearised versions of the problems can be solved using commercial exact algorithms.
-The \texttt{raptr} R package provides functions to express conservation planning data as optimisation problems using linearised versions of the unreliable and reliable formulations (see Appendix S1 for formulation).
-These optimisation problems can then be solved to generate prioritisations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).
-Presently, academics can obtain a license at no cost from the Gurobi website.
-
-# Examples
-
-To understand the behaviour of the unreliable problem and showcase its value, we conducted a simulation study and two case studies.
-These studies involved generating solutions using only amount targets to represent prioritisations generated using conventional methods (eg.
-\texttt{Marxan}), and solutions using amount and space targets using the unreliable formulation.
-By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection.
-All analyses were performed in \texttt{R} [version `r getRversion()`; @r192].
-
-## SIMULATION STUDY
-### Methods
-We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig.
-2).
-We then simulated three species across this study area.
-The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig.
-2a).
-This species' distribution was based on a uniform distribution (eqn 8a).
-The second species was simulated to represent a species with an idealised distribution (Fig.
-2b).
-It has a core range area, and is less likely to be found in areas that are distant from the core area.
-This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8c).
-The third species represents a species with two distinct populations (Fig.
-2c).
-This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8d).
-The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d.
-We used a geographic attribute space to showcase the behavior of the problem.
-For each species, demand points were generated by calculating the centroids of each planning units and their weights were set as the probability that the planning units were occupied.
-
-\begin{align*}
-\\
-P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
-%
-f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
-%
-P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
-%
-P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
-f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
-\end{cases} \tag*{eqn 8d} \\
-%
-\end{align*}
-
-We generated four solutions for each species.
-First to represent solutions generated using conventional conservation planning methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-Second to show how the addition of geographic targets can affect a prioritisation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets.
-Third to represent solutions using conventional planning methods that penalise for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
-Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated solutions using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
-
-### Uniformly distributed species
-
-The uniform species was simulated to occur with a constant probability of occupancy across the study area (Fig.
-2a).
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount-based targets (Fig.
-3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area.
-This specific configuration is an artifact of the \texttt{Gurobi} solver used to solve the problem.
-In fact, for this species, all prioritisations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-In the absence of criteria to guarantee representativeness, reserve selection methods may or may not return solutions that secure a representative sample of the features.
-In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points.
-
-The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
-3b).
-Since all planning units have equal cost and an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf.
-Fig 3a).
-Although the use of amount and space based targets has addressed adequacy and representativeness objectives (respectively), they have resulted in a highly fragmented solution.
-
-The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig.
-3d).
-This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented.
-This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig.
-3c) contained the same number of planning units as the solution with just the amount targets (Fig.
-3a).
-Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.
-
-### Normally distributed species
-
-The normally distributed species was simulated to contain a single core area where individuals were most prevalent, and marginal areas where individuals were less likely to occur (Fig.
-2b).
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig.
-3e).
-Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy.
-Whilst this solution satisfies the adequacy objective for a prioritisation in a cost-effective manner--it fails to fulfill the representative objective for a prioritisation (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).
-
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig.
-3f).
-This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution.
-By using amount and geographic targets in the reserve selection problem, we have obtained a solution that concentrates conservation effort in the range core and also the range margin.
-However, similar to the corresponding solution for the uniform species (Fig.
-3b), this solution is highly fragmented.
-
-By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig.
-3h).
-This solution contains `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units.
-Similar to the corresponding solution for the uniform species (Fig.
-3d), this solution contains more planning units than any other solutions for this species (cf.
-Figs.
-3e--3g).
-The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species--even for species without significant spatial structure.
-
-### Bimodally distributed species
-
-The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig.
-2c).
-The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets assigned conservation effort to just one of the two populations (Fig.
-3i).
-While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-The addition of boundary length modifiers to amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig.
-3k).
-However, the addition of geographic space targets resolved these issues.
-
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets included planning units from both populations (Fig.
-3j).
-Although this solution required more planning units to fulfill both targets ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$n`$), this solution secured a representative sample of the bimodally distributed species (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % range sampled).
-Similar to corresponding solutions generated for both the uniformly (Fig.
-3b) and normally distributed species (Fig.
-3f), this solution was also highly fragmented and we could obtain a more well-connected solution at the expense of selecting additional planning units ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n`$; Fig.
-3l).
-The results for the bimodally distributed species suggest that species with highly structured populations or strong variation between individuals could benefit the most from prioritisations that are generated using space-based targets.
-
-## CASE STUDY 1
-### Methods
-We investigated how space-based targets can be used in a multi-species planning context to generate a prioritisation that sufficiently preserves the realised niche for several species.
-By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63].
-We selected Queensland, Australia as the study area.
-This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats.
-We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
-We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterise the environmental variation across the study area (Fig.
-4).
-
-We selected four bird species that span a range of different evolutionary histories, distributions and ecologies: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_).
-We then mapped the extent of occurrence for each species (Fig.
-5).
-To do this, we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
-We used this method because it is entirely reproducible using freely available data.
-
-We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Fig.
-S2).
-We used the following method to ensure that the demand points reflected the core parts of the species' realised niches.
-For each species, we generated random points inside the species range and at each point extracted the principal component values at that location.
-We then fit hyperbox kernels to the distribution of principal component values to characterise the realised niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231].
-We then generated uniformly distributed points inside the species' kernels and estimated the density of the training points at the uniformly generated points.
-These uniformly distributed points and associated density estimated were used as demand point coordinates and weights (respectively).
-
-We generated two solutions.
-The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species.
-The second solution was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.
-
-### Results
-Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches--with `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'one')` excpetion.
-Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realised niche of `r parsed.representative.space.held.names`.
-Yet it failed to achieve this for `r parsed.not.representative.space.held.names`.
-This result demonstrates that though solutions generated using conventional methods may preserve a representative sample of some features (Figure S1), only through the use of explicit targets can decision makers be sure that solutions secure a representative sample for all features.
-
-## CASE STUDY 2
-### Methods
-Here we used space-based targets to generate a prioritisation securing a representative sample of a species' intra-specific genetic variation.
-We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
-Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study.
-Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
-22.3 km $\times$ 25 km; Fig.
-7a).
-They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals.
-Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
-In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.
-
-We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
-These continuous variables described the main axes of genetic variation within the species.
-We calculated the average of the values associated with individuals in each grid cell.
-These values were used to create a genetic attribute space (Figs.
-7c and 7d).
-To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances based on the grid cells' centroids [using the \texttt{ape R} package; @r464].
-
-The grid cells were used as planning units.
-The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
-Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points.
-To ensure that the solutions did not prioritise particularly costly areas, we included opportunity cost data (Fig.
-7b).
-We obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
-
-We generated two solutions.
-The first solution was generated using an `r cs2.params.LST[[MODE]]$amount.target*100` % amount-based target.
-The second solution was generated using the same amount-based target and an additional `r cs2.params.LST[[MODE]]$genetic.target*100` % genetic-based target.
-
-### Results
-
-A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`; Figs.
-7c--7d).
-Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
-In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
-This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.
-
-The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs.
-8a and 8c).
-This solution only sampled planning units that contained individuals belonging to one of the two main groups.
-On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups.
-Note that since these solutions were generated using opportunity cost data, the solution generated using only the amount target is essentially the minimum number of least costly areas needed to fulfill the amount target.
-The only difference between the two solutions is a single planning unit.
-By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs.
-8b and 8d) for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).
-
-# Implications and future directions
-The \texttt{raptr} R package provides a unified approach to reserve selection.
-This \texttt{R} package provides decision makers with the tools to generate prioritisations that secure intra-specific as well as as inter-specific biodiversity patterns.
-Additionally, the package contains functionality to accommodate uncertainty in the distribution of features, and also identify suitably connected reserves.
-Both the simulated and case study species suggest that explicitly considering representation as well as amount targets can significantly alter prioritisations.
-
-One of the key advantages of this package is that it is general enough to incorporate any spatially explicit variable as an attribute space.
-This package can accommodate intrinsic or extrinsic variation in the feature(s).
-For example, adaptation processes could be secured using environmental variation [eg.
-@r247].
-Predator-prey processes could be conserved by capturing the overlapping distributions of predators and each of their prey species [eg.
-@r485].
-Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserve so there is equitable access to parks by people [eg.
-@r483].
-As long as the variation can be described using Euclidean distances--which could be achieved through transformation or dimension reduction techniques--the \texttt{R} package can be used to obtain a representative sample.
-
-The degree to which a prioritisation truly secures a representative sample of a feature depends on the attribute spaces and the distribution of demand points chosen by the decision maker.
-Ultimately, the space-based targets are set as a proportion based on the distribution of the demand points.
-As a consequence, if the planner uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not effectively represent the decision maker's objectives.
-We therefore stress that planners must carefully consider which biodiversity processes need to be reflected in the prioritisation, and which spatial data can be used to represent these processes.
-Note that maximising for one attribute space can result in trading off against another.
-For instance, to maximise the geographic spread of a prioritisation, reserves need to be further away from each other, yet to maximise connectivity, reserves may need to be closer together.
-Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space.
-To assist the generation of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see the `make.DemandPoints` function).
-These routines use kernels to automatically generate demand points based on the distribution of features in spatial data sets [using @r231; @r450].
-
-The formulation requires spatially comprehensive data to map attribute spaces to planning units.
-However, many real-world data sets are patchy.
-For example, due to constrained resources, genetic data may have only been collected in some planning units.
-To use such patchy data with the \texttt{raptr R} package, the planning units that are missing data must either be omitted (as was done in the second case-study), or their missing values must be estimated.
-Missing data could be estimated using spatially explicit models [eg.
-krigging or generalised dissimilarity models; @r467; @r466].
-This approach has been successfully applied to a range of biological data sets [eg.
-@r468].
-
-To maximise the long-term persistence of biodiversity, decision makers need to identify prioritisations that preserve existing patterns of biodiversity and the processes that support them.
-To achieve this, conservation planners need a decision support tool that can deliver cost-effective prioritisations that secure an adequate amount of a representative sample of biodiversity features.
-Here, we developed the \texttt{raptr R} package to fill this void.
-By exploring the functionality of this package using several simulated and case study species, we found that including space-based targets can result in substantially different solutions.
-
-# Acknowledgements
-JOH is funded by an Australian Postgraduate Award (APA) scholarship.
-RAF and HPP have Australian Research Council Fellowships.
-This work was supported by the Centre for Biodiversity and Conservation Science (CBCS).
-
-# Data accessibility
-The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://cran.r-project.org/}).
-All data, code, and results are stored in an online repository (\url{https://www.github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
-
-# References
-
-\bibliography{references}
+\clearpage 
+\input{./supporting-information.tex}

### 2016-11-23 code/rmarkdown/cover-letter.tex
+ add in readmes + compile manuscript into single pdf and single docx

\documentclass{letter}
\usepackage[a4paper, left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  urlcolor=blue
}
\signature{\vspace{-20pt} Jeffrey Hanson}
\address{School of Biological Sciences\\The University of Queensland\\Brisbane, QLD, Australia}

\longindentation=-1pt
\begin{document}
\begin{letter}{Department of Animal and Plant Sciences\\University of Sheffield\\Sheffield S10 2TN, UK}
\opening{Dear Professor Rob Freckleton,}

Biodiversity is in crisis. To combat further declines, the most fundamental aim of conservation is to maximize the long-term persistence of biodiversity. To achieve this, conservation actions must preserve biodiversity patterns (eg. populations, species, ecosystems), but also crucially the ecological and evolutionary processes that sustain them. Protected area networks buffer species from gross threatening processes and set the stage for enhanced management interventions. As you are aware, a wealth of data relevant to biodiversity processes has become available to the global scientific community over the last decade (eg. genetic, environmental, and morphological data). However, currently available decision support tools only have a limited capacity to use such data to guide conservation planning.

To begin to fill this gap, our manuscript entitled \textit{"raptr: Representative and Adequate Prioritization Toolkit in R"} describes a new conservation reserve planning tool that explicitly accommodates data pertaining to biodiversity patterns and biodiversity processes. This tool is available as an \texttt{R} package on \textit{The Comprehensive R Archive Network} (\url{https://CRAN.R-project.org/package=raptr}). By applying the tool to a simulation study and two case studies, we showcase its behavior and compare it with conventional methods for reserve selection. In each of these studies, we found that explicitly targeting a representative sample of a species--for instance in terms of its genetic variation to conserve evolutionary processes--resulted in a different set of conservation priorities.

We chose to publish our work in \textit{Methods in Ecology and Evolution} because of its focus on novel statistical and computational methodologies. We desire to communicate our findings to the global community of ecologists and conservation biologists that follow this journal. The content of our manuscript has not been published elsewhere has not been published elsewhere, nor has it been submitted for publication to another journal. All authors approved the manuscript and this submission. In support of the journal's commitment to reproducible research, we have stored all data, code, and results in an online repository to permit replication and validation of this study (\url{https://www.github.com/jeffreyhanson/raptr-manuscript}).

\closing{Thank you for your consideration,}

\end{letter}
\end{document}

### 2016-11-23 code/rmarkdown/figures.Rmd
+ edits to ms + increase blm value
+ add in readmes + compile manuscript into single pdf and single docx

@@
 ---
-fontsize: 11pt
-documentclass: article
 bibliography: references.bib
@@
     keep_tex: yes
-    includes:
-      in_header: [preamble.tex, figures-preamble.tex]
 ---
@@
 
-## Figures
+# Figures
 
@@
 
-```{r, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap=paste0('Simulation study prioritisations.
-Each panel shows a prioritisation generated for a single species using a set of parameters.
+```{r, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap=paste0('Simulation study prioritizations.
+Each panel shows a prioritization generated for a single species using a set of parameters.
 Squares denote planning units, with dark green showing those planning units selected for protection in a particular prioritisation.
-The top row of panels (a, d, g, j) shows prioritisations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
+The top row of panels (a, d, g, j) shows prioritizations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
 Each column of panels corresponds to a different set of parameters used to generate the prioritisation.
-The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritisations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
+The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
 # prepare data
@@
 
-```{r, echo=FALSE, fig.height=2.5, fig.width=4.5, fig.cap=paste0('Prioritisations for the niche-based case-study in Queensland, Australia.
+```{r, echo=FALSE, fig.height=2.5, fig.width=4.5, fig.cap=paste0('Prioritizations for the niche-based case-study in Queensland, Australia.
 Squares represent planning units.
@@
 Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
-Panel (b) shows the planning units selected selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
+Panel (b) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
 # prepare data
@@
 Panel (a) shows all grid cells surveyed by the IntraBioDiv project.
-Grid cells occupied by ',cs2.params.LST[[MODE]]$common.name,' are shown in bright blue.
+Grid cells occupied by ',tolower(cs2.params.LST[[MODE]]$common.name),' are shown in bright blue.
 The subsequent panels contain only the grid cells that were occupied by the species.
@@
 
-```{r, echo=FALSE, fig.height=5, fig.width=5, fig.cap=paste0('Prioritisations for the genetic case-study in the European Alps.
+```{r, echo=FALSE, fig.height=5, fig.width=5, fig.cap=paste0('Prioritizations for the genetic case-study in the European Alps.
 Panels (a--b) show prioritisations generated using different parameters.
@@
 grid.arrange(p1,p2,p3,p4,ncol=2)
-
 ```

### 2016-11-23 code/rmarkdown/supporting-information.Rmd
+ add in readmes + compile manuscript into single pdf and single docx

@@
     keep_tex: yes
-    includes:
-      in_header: [preamble.tex, si-preamble.tex]
 ---
@@
 ```
+
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
@@
 The reliable formulation explicitly considers the probability that planning units are inhabited by features.
-As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented.
-This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s).
-To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.
+As a consequence, it may deliver prioritizations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritization is implemented.
+This behavior is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s).
+To ensure that prioritizations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.
 
@@
 The third closest planning units are assigned at $r$-level$=2$.
-Continuing on, in this manner, the selected planning units in a prioritisation are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.
+Continuing on, in this manner, the selected planning units in a prioritization are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.
 
@@
 [-@r16] found that $R=5$ yields similar solutions to $R >> 5$.
-However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.
+However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritizations in a feasible amount of time.
 
@@
 The reliable formulation is non-linear.
-However, the non-linear components can be linearised.
-First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al.
+However, the non-linear components can be linearized.
+First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al.
 [-@r426].
-Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
+Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 7a) can be linearized using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
 [-@r16].
@@
 ```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study.
-Each panel shows a the distribution of a solution in environmental space and how it samples the realised niche for a different species.
+Each panel shows a the distribution of a solution in environmental space and how it samples the realized niche for a different species.
 The left column of panels shows the solution generated using amount targets.
@@
 	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
-	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.',
+	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.',
 	'$S$', 'set of attribute spaces (indexed by $S$)',
@@
 	colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
-	caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritisation (URAP) problem.'
+	caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritization (URAP) problem.'
 )
@@
 
-# References
+## References

### 2016-11-23 code/rmarkdown/text.Rmd
+ add in readmes + compile manuscript into single pdf and single docx

---
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
---

```{r, include=FALSE}
# set wd
knitr::opts_knit$set(root.dir=normalizePath('../..'))
```

```{r, include=FALSE}
## initialization
# load .rda file
checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
session::restore.session('data/final/results.rda')

## preliminary processing
for (i in seq_len(nlayers(cs1.bioclim.RAST)))
  cs1.bioclim.RAST@layers[[i]]@file@name <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])

```

Perhaps the most fundamental aim of conservation is to maximize the long-term persistence of biodiversity [@r423; @r255]. To achieve this, conservation actions must preserve biodiversity patterns (eg. populations, species, ecosystems), but also crucially the processes that sustain them. One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r484; @r440]. Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465]. Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using this method is challenging because biodiversity data are often continuous and multi-dimensional.

Simply partitioning features based on continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable. For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution captures 10 % of the geographic range of each sub-feature, how much of the original continuous variation does the solution represent? The answer is not immediately clear. If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation. Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning features based on political boundaries; @r483], then it might sample much less of the original variation. Conventional reserve selection problems cannot differentiate between these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce cost or fragmentation. This limitation has been known of quite some time. In fact, it dates back to some of the earliest reserve selection methods [eg. @r295].

During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of environmental variation across a study area. Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range. Recent work has built on this reserve selection method by solving problems with more advanced optimization algorithms [@r482]. However, \texttt{DIVERSITY} has limited utility for species-based conservation planning, in part, because it cannot accommodate multiple biological features--unlike more commonly used tools such as \texttt{Marxan} and \texttt{Zonation}. This fundamental limitation means that \texttt{DIVERSITY} is not useful for most conservation planning scenarios.

Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of each feature. To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package. This \texttt{R} package provides decision makers with the tools to generate prioritizations based on data that relate to biodiversity patterns and processes. Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.

# Methods
## PROBLEM FORMULATION
Biodiversity features are defined as the entities that the prioritization is required to preserve (eg. species, ecosystems). Spatial attributes are defined as the variation across the feature' range that the prioritization is required to sample. They can be intrinsic variation (eg. genetic or trait) variation that is extrinsic to the feature (eg. environmental conditions). There should, however, be a reasonable underlying hypothesis that relates this variation to the biodiversity processes that the decision maker aims to preserve.

A set of attributes is conceptualized as a space (termed an "attribute space"). For example, a decision maker may require a prioritization that represents populations along climatic gradients. To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this environmental space. Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given attribute space. By associating the planning units with climatic data, and calculating a representative statistic for each planning unit (eg. mean), they can be mapped from geographic space to this environmental attribute space. In addition to planning units, attribute spaces also contain demand points.

Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets). The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space (see below for mathematical formulations). The shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the variation in the spatial attribute. To calculate the proportion of variation secured in a given solution, the distances are scaled by the distance between the demand points to their centroid (equivalent to calculating R$^2$ values for K-means clustering). In any attribute space there may exist points that are impossible (eg. mean annual rainfall -5 mm), or do not occur in the study area (eg. mean annual temperature 30$^{\circ}$C in Antarctica). Additionally, there may be some regions that are desirable for some features and undesirable for others (eg. conditions known to be outside the physiological tolerance of certain species). Thus a different set of demand points and weights are used for each attribute space and each feature. By siting demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritizations secure the feature in planning units with spatial attributes that are desirable or representative for that feature.

To illustrate these concepts, consider the following example: we wish to develop a prioritization for a single species that has four populations. However, we can only afford to preserve three of the four populations. To ensure that we preserve a representative sample, we want to preserve populations inhabiting different environmental conditions. We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population. We used this data to construct a two-dimensional environmental attribute space. Next, we generated demand points as equi-distant points inside this space. By comparing the distributions of demand points and populations in the attribute space, we can identify a suitable prioritization (Fig. 1). We can see that a solution that prioritizes both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and constitute considerable redundancy if both selected. Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$ (or $C$), $B$, and $D$. This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide the reserve selection process.

The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451]. Although the \texttt{raptr R} package utilizes two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. These formulations are based on unreliable and reliable facility location problems and have been named accordingly [@r16]. The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating the proportion of an attribute space sampled in a solution. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.

Define $F$ to be the set of features one wishes to conserve (indexed by $f$). Let $J$ be a set of planning units (indexed by $j$). Also, let $A_j$ denote the area, and $C_j$ denote the cost of preserving planning unit $j \in J$. To assess the extent to which each feature is secured in a given prioritization, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$. The level of fragmentation associated with a prioritization is parameterized as the total exposed boundary length (as in \texttt{Marxan}). Let the shared edges between each planning unit $j \in J$ and $k \in J$ be $e_{jk}$.

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points. Demand points with greater weight $\lambda_{fsi}$ are more important, and the solutions that select planning units close to highly weighted demand points will be more favorable. As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.

Targets are used to ensure that prioritizations adequately preserve each feature. Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient. Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$. Space-based targets ensure that a representative sample of each feature is secured. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. For convenience, all targets are expressed as proportions in the \texttt{R} package.

The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{fs} &= \parbox{25em}{representation target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM) to penalize fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

The decision variables are the $X_j$ and $Y_{fsij}$ variables.

\begin{align*}
X_j
  &= \begin{cases}
    1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
    0, & \parbox{25em}{otherwise} \\
  \end{cases} \\
%
Y_{fsij} &= \begin{cases}
    1, & \parbox{25em}{if demand point $i$ is used to represent planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
    0, & \parbox{25em}{otherwise} \\
  \end{cases} \\
\end{align*}

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is assigned to a selected planning unit $J$ where $X_j = 1$. The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution. Generally, demand points are represented by their closest selected planning units (unless particularly low space-based targets are used).

The unreliable formulation (URAP) is a defined as a multi-objective optimization problem.

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

The objective function (eqn 3a) determines the utility of a given prioritization: a combination of the total cost of a prioritization and how fragmented it is. Constraints (eqn 3b) ensure that all the amount-based targets are met. Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space. For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted distance is then scaled by the inherent variation in the demand points ($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis. The constraints ensure that this proportion must be equal to or greater than the space-based target. Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point. Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units. Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.

## Optimization
The unreliable formulation is non-linear. However, the non-linear components can be linearized using existing techniques. The expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al. [-@r426]. Linearized versions of the problems can be solved using commercial exact algorithms. The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation). These optimization problems can then be solved to generate prioritizations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}). Presently, academics can obtain a license at no cost from the Gurobi website.

# Examples

To understand the behavior of the unreliable problem and showcase its value, we conducted a simulation study and two case studies. These studies involved generating solutions using only amount targets to represent prioritizations generated using conventional methods (eg. \texttt{Marxan}), and solutions using amount and space targets using the unreliable formulation. By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection. All analyses were performed in \texttt{R} [version `r getRversion()`; @r192].

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig. 2). We then simulated three species across this study area. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig. 2a). This species' distribution was based on a uniform distribution (eqn 8a). The second species was simulated to represent a species with an idealized distribution (Fig. 2b). It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8c). The third species represents a species with two distinct populations (Fig. 2c). This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8d). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were generated by calculating the centroids of each planning units and their weights were set as the probability that the planning units were occupied.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 8d} \\
%
\end{align*}

We generated four solutions for each species. First to represent solutions generated using conventional conservation planning methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets. Second to show how the addition of geographic targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets. Third to represent solutions using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`. Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated solutions using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.

### Uniformly distributed species

The uniform species was simulated to occur with a constant probability of occupancy across the study area (Fig. 2a). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount-based targets (Fig. 3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area. This specific configuration is an artifact of the \texttt{Gurobi} solver used to solve the problem. In fact, for this species, all prioritizations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets. In the absence of criteria to guarantee representativeness, reserve selection methods may or may not return solutions that secure a representative sample of the features. In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points.

The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig. 3b). Since all planning units have equal cost and an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf. Fig 3a). Although the use of amount and space based targets has addressed adequacy and representativeness objectives (respectively), they have resulted in a highly fragmented solution.

The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig. 3d). This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented. This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig. 3c) contained the same number of planning units as the solution with just the amount targets (Fig. 3a). Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.

### Normally distributed species

The normally distributed species was simulated to contain a single core area where individuals were most prevalent, and marginal areas where individuals were less likely to occur (Fig. 2b). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig. 3e). Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy. Whilst this solution satisfies the adequacy objective for a prioritization in a cost-effective manner--it fails to fulfill the representative objective for a prioritization (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig. 3f). This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution. By using amount and geographic targets in the reserve selection problem, we have obtained a solution that concentrates conservation effort in the range core and also the range margin. However, similar to the corresponding solution for the uniform species (Fig. 3b), this solution is highly fragmented.

By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig. 3h). This solution contains `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units. Similar to the corresponding solution for the uniform species (Fig. 3d), this solution contains more planning units than any other solutions for this species (cf. Figs. 3e--3g). The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species--even for species without significant spatial structure.

### Bimodally distributed species

The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig. 2c). The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets assigned conservation effort to just one of the two populations (Fig. 3i). While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). The addition of boundary length modifiers to amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig. 3k). However, the addition of geographic space targets resolved these issues.

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets included planning units from both populations (Fig. 3j). Although this solution required more planning units to fulfill both targets ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$n`$), this solution secured a representative sample of the bimodally distributed species (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % range sampled). Similar to corresponding solutions generated for both the uniformly (Fig. 3b) and normally distributed species (Fig. 3f), this solution was also highly fragmented and we could obtain a more well-connected solution at the expense of selecting additional planning units ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n`$; Fig. 3l). The results for the bimodally distributed species suggest that species with highly structured populations or strong variation between individuals could benefit the most from prioritizations that are generated using space-based targets.

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the realized niche for several species. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63]. We selected Queensland, Australia as the study area. This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterize the environmental variation across the study area (Fig. 4).

We selected four bird species that span a range of different ecological requirements, evolutionary histories, distributions: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_). We then mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Fig. S2). We used the following method to ensure that the demand points reflected the core parts of the species' realized niches. For each species, we generated random points inside the species range and at each point extracted the principal component values at that location. We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231]. We then generated uniformly distributed points inside the species' kernels and estimated the density of the training points at the uniformly generated points. These uniformly distributed points and associated density estimated were used as demand point coordinates and weights (respectively).

We generated two solutions. The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species. The second solution was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.

### Results
Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches--with `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'one')` excpetion. Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realized niche of `r parsed.representative.space.held.names`. Yet it failed to achieve this for `r parsed.not.representative.space.held.names`. This result demonstrates that though solutions generated using conventional methods may preserve a representative sample of some features (Figure S1), only through the use of explicit targets can decision makers be sure that solutions secure a representative sample for all features.

## CASE STUDY 2
### Methods
Here we used space-based targets to generate a prioritization securing a representative sample of a species' intra-specific genetic variation. We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462]. Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study. Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Fig. 7a). They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals. Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci. In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We calculated the average of the values associated with individuals in each grid cell. These values were used to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances based on the grid cells' centroids [using the \texttt{ape R} package; @r464].

The grid cells were used as planning units. The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit. Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points. To ensure that the solutions did not prioritize particularly costly areas, we included opportunity cost data (Fig. 7b). We obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.

We generated two solutions. The first solution was generated using an `r cs2.params.LST[[MODE]]$amount.target*100` % amount-based target. The second solution was generated using the same amount-based target and an additional `r cs2.params.LST[[MODE]]$genetic.target*100` % genetic-based target.

### Results

A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`; Figs. 7c--7d). Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$). In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c). This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs. 8a and 8c). This solution only sampled planning units that contained individuals belonging to one of the two main groups. On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups. Note that since these solutions were generated using opportunity cost data, the solution generated using only the amount target is essentially the minimum number of least costly areas needed to fulfill the amount target. The only difference between the two solutions is a single planning unit. By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs. 8b and 8d) for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).

# Implications and future directions
The \texttt{raptr R} package provides a unified approach to reserve selection. Conservation planners can use this \texttt{R} package to generate prioritizations that secure intra-specific as well as as inter-specific biodiversity patterns. Additionally, the package contains functionality to accommodate uncertainty in the distribution of features using the reliable formulation (defined in Appendix S1), and ensure that solutions are well-connected. Both the simulated and case study species suggest that explicitly considering representation as well as amount targets can significantly alter prioritizations.

One of the key advantages of this package is that it is general enough to incorporate any spatially explicit variable as an attribute space. This package can accommodate intrinsic or extrinsic variation in the feature(s). For example, adaptation processes could be secured using environmental variation [eg. @r247]. Predator-prey processes could be conserved by capturing the overlapping distributions of predators and each of their prey species [eg. @r471; @r485]. Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserves so there is equitable access to parks by people [eg. @r483]. As long as the variation can be described using Euclidean distances--which could be achieved by transforming or applying dimension reduction techniques--the \texttt{R} package can be used to obtain a representative sample.

The degree to which a prioritization truly secures a representative sample of a feature depends on the attribute spaces and the distribution of demand points chosen by the decision maker. Ultimately, the space-based targets are set as a proportion based on the distribution of the demand points. As a consequence, if the planner uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not effectively represent the decision maker's objectives. We therefore stress that planners must carefully consider which biodiversity processes need to be reflected in the prioritization, and which spatial data can be used to represent these processes. Note that maximizing for one attribute space can result in trading off against another. For instance, to maximize the geographic spread of a prioritization, reserves need to be further away from each other, yet to maximize connectivity, reserves may need to be closer together. Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space. To assist the generation of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see documentation for the `make.DemandPoints` function). These routines use kernels to automatically generate demand points based on the distribution of features in spatial data sets [using @r231; @r450].

The formulation requires spatially comprehensive data to map attribute spaces to planning units. However, many real-world data sets are patchy. For example, due to constrained resources, genetic data may have only been collected in some planning units. To use such patchy data with the \texttt{raptr R} package, the planning units that are missing data must either be omitted (as we did in the second case-study), or their missing values can be estimated. Missing data could be estimated using spatially explicit models [eg. krigging or generalized dissimilarity models; @r467; @r466]. This approach has been successfully applied to a range of biological data sets [eg. @r468].

To maximize the long-term persistence of biodiversity, decision makers need to identify prioritizations that preserve existing patterns of biodiversity and the processes that support them. To achieve this, conservation planners need a decision support tool that can deliver cost-effective prioritizations that secure an adequate amount of a representative sample of biodiversity features. Here, we developed the \texttt{raptr R} package to fill this void. By exploring the functionality of this package using several simulated and case study species, we found that including space-based targets can result in substantially different solutions.

# Acknowledgements
JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF and HPP have Australian Research Council Fellowships. This work was supported by the Centre for Biodiversity and Conservation Science (CBCS).

# Data accessibility
The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}). All data, code, and results are stored in an online repository (\url{https://www.github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.

# References

\bibliography{references}

### 2016-11-23 code/rmarkdown/figures.Rmd
+ resize figures to journal specifications

@@
 
+column.width <- 2.75591
+
 ```
@@
 
-```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.5, fig.width=4.5, fig.cap='Simulated distributions of three species.
+```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.6, fig.width=column.width, fig.cap='Simulated distributions of three species.
 Squares denote planning units.
@@
 sim.pus.occ.letters.DF <- data.frame(
-	x=gCentroid(sim.pus[1,])@coords[1],
-	y=gCentroid(sim.pus[1,])@coords[2],
+	x=min(sim.pus.occ.FPLY$long) + 1.5,
+	y=max(sim.pus.occ.FPLY$lat) - 1.5,
 	species=unique(sim.pus.occ.FPLY$species),
@@
 	theme_classic() +
-	geom_text(data=sim.pus.occ.letters.DF, aes(x=x, y=y, label=letter), size=2, color='white') +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 5, 0, 0), "lines"),
+	geom_label(data=sim.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=2) +
+	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(),
 		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(),
-		legend.position=c(1.15, 0.6)) +
+		strip.text = element_text(color='white', size=7),
+    axis.title = element_blank(),
+    legend.title = element_text(color='black', size=7),
+		legend.text = element_text(color='black', size=7),
+		legend.margin = unit(0, 'cm'),
+		legend.position='bottom') +
 	facet_wrap(~ species, ncol=3)
@@
 
-```{r, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap=paste0('Simulation study prioritizations.
+```{r, echo=FALSE, fig.height=3.5, fig.width=2*column.width, fig.cap=paste0('Simulation study prioritizations.
 Each panel shows a prioritization generated for a single species using a set of parameters.
@@
 # prepare plotting data
-sim.letter.DF2 <- mutate(sim.spp.DF, letter=paste0('(', letters[seq_along(species)],')'))
-sim.letter.DF2$long <- gCentroid(sim.pus[1,])@coords[1]
-sim.letter.DF2$lat <- gCentroid(sim.pus[1,])@coords[2]
+sim.letter.DF2 <- sim.spp.DF %>%
+  mutate(letter=paste0('(', letters[seq_along(species)],')'),
+         long=min(sim.pus.sel.FPLY$long) + 1.6,
+         lat=max(sim.pus.sel.FPLY$lat) - 1.5)
 # make plot
@@
 	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
-	geom_text(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', size=1.8) +
+	geom_label(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', fill=alpha('white', 0.8), size=5) +
 	theme_classic() +
@@
 
-```{r, echo=FALSE, fig.height=2.5, fig.width=3, fig.cap='Distribution of environmental variation across Queensland, Australia.
+```{r, echo=FALSE, fig.height=1.6, fig.width=column.width, fig.cap='Distribution of environmental variation across Queensland, Australia.
 Squares denote planning units.
@@
 		names(curr.DF)[3] <- 'Score'
-		curr.DF$pc <- paste0('Principal component ',i)
+		curr.DF$pc <- paste0('PC ',i)
 		return(curr.DF)
@@
 		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
-		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
 	) +
 	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[1])$Score, n=3, min.n=2)) +
 	facet_wrap(~ pc)
@@
 		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
-		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
 		) +
 	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[2])$Score, n=2, min.n=2)) +
 	facet_wrap(~ pc)
@@
 
-```{r, echo=FALSE, fig.height=4.5, fig.width=4.0, fig.cap='Distribution of the species used in the first case-study.
+```{r, echo=FALSE, fig.height=2.35, fig.width=column.width, fig.cap='Distribution of the species used in the first case-study.
 Squares denote planning units.
-Each panel shows the distribution of a different species.'}
+Each panel shows the distribution of a different species.
+Planning units occupied by a species are shown in bright blue.'}
 # prepare data
@@
 cs1.pus.occ.letters.DF <- data.frame(
-	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.05)),
-	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.965)),
+	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.06)),
+	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.90)),
 	species=unique(cs1.pus.occ.FPLY$species),
@@
 	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black') +
-	geom_text(data=cs1.pus.occ.letters.DF, aes(x=x, y=y, label=letter), color='black') +
+	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black', size=0.05) +
+	geom_label(data=cs1.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
 		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
-		legend.position=c(1.045, 1.035), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+		strip.text = element_text(color='white', size=8),
+		axis.title=element_blank(), legend.position='none'
 	) +
@@
 
-```{r, echo=FALSE, fig.height=2.5, fig.width=4.5, fig.cap=paste0('Prioritizations for the niche-based case-study in Queensland, Australia.
+```{r, echo=FALSE, fig.height=1.175, fig.width=column.width, fig.cap=paste0('Prioritizations for the niche-based case-study in Queensland, Australia.
 Squares represent planning units.
@@
 cs1.pus.sel.letters.DF <- data.frame(
-	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.05)),
-	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.965)),
+	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.06)),
+	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.90)),
 	Prioritisation=names(cs1.prioritisations),
@@
 	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black') +
-	geom_text(data=cs1.pus.sel.letters.DF, aes(x=x, y=y, label=letter), color='black') +
+	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black', size=0.1) +
+	geom_label(data=cs1.pus.sel.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
 		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
-		legend.position=c(1.03, 1.03), legend.justification=c(1,1), legend.key.size=unit(3, 'mm')
+		strip.text = element_text(color='white', size=8),
+		axis.title=element_blank(), legend.position='none'
 	) +
@@
 
-```{r, echo=FALSE, fig.height=4.5, fig.width=4.5, fig.cap=paste0('Study area and data used in the second case-study.
+```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Study area and data used in the second case-study.
 Squares denote planning units.
@@
 cs2.spp.grid.letters.DF <- data.frame(
-	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.05)),
-	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.95)),
+	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.06)),
+	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.90)),
 	letter=('(a)')
@@
 cs2.spp.grid.sub.letters.DF <- data.frame(
-	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.05)),
-	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.95)),
+	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
+	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.90)),
 	letter=c('(b)','(c)','(d)')
@@
 	geom_polygon(data=cs2.spp.grid.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), alpha=0.8, color='grey10') +
-	geom_text(data=cs2.spp.grid.letters.DF, aes(x=x, y=y, label=letter), color='black') +
+	geom_label(data=cs2.spp.grid.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
 	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Cost), alpha=0.8, color='grey10') +
-	geom_text(data=cs2.spp.grid.sub.letters.DF[1,], aes(x=x, y=y, label=letter), color='black') +
+	geom_label(data=cs2.spp.grid.sub.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
 		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm'),
-		legend.title=element_blank()
+		legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
+		legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
 	) +
 	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
-	scale_fill_continuous(NULL) +
+	scale_fill_continuous(name=element_blank(), breaks=pretty(cs2.spp.grid.sub.FPLY$Cost, n=3, min.n=2)) +
 	facet_wrap(~ Title1)
@@
 	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds1), alpha=0.8, color='grey10') +
-	geom_text(data=cs2.spp.grid.sub.letters.DF[2,], aes(x=x, y=y, label=letter), color='black') +
+	geom_label(data=cs2.spp.grid.sub.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
     axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm')
+    legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
+    legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
 	) + 
 	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
-	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred') +
+	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred', breaks=c(0, 0.3)) +
 	facet_wrap(~ Title2)
@@
 	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds2), alpha=0.8, color='grey10') +
-	geom_text(data=cs2.spp.grid.sub.letters.DF[3,], aes(x=x, y=y, label=letter), color='black') +
+	geom_label(data=cs2.spp.grid.sub.letters.DF[3,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
     axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-		legend.position=c(1, 0), legend.justification=c(1,0),  legend.key.size=unit(3, 'mm')
+    legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
+    legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
 	) + 
 	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
-	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred') +
+	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred', breaks=pretty(cs2.spp.grid.sub.FPLY$nmds2, n=2, min.n=2)) +
 	facet_wrap(~ Title3)
@@
 
-```{r, echo=FALSE, fig.height=5, fig.width=5, fig.cap=paste0('Prioritizations for the genetic case-study in the European Alps.
+```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Prioritizations for the genetic case-study in the European Alps.
 Panels (a--b) show prioritisations generated using different parameters.
@@
 cs2.sel.grid.sub.letters.DF <- data.frame(
-	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.05)),
-	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.95)),
+	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
+	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.90)),
 	letter=c('(a)','(b)')
@@
 cs2.sel.space.letters.DF <- data.frame(
-	x=min(cs2.sel.space.DF$d1) + (diff(range(cs2.sel.space.DF$d1))*0.3),
+	x=min(cs2.sel.space.DF$d1) + (diff(range(cs2.sel.space.DF$d1))*0.85),
 	y=min(cs2.sel.space.DF$d2) + (diff(range(cs2.sel.space.DF$d2))*0.95),
@@
 	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.selection), color='black') +
-	geom_text(data=cs2.sel.grid.sub.letters.DF[1,], aes(x=x, y=y, label=letter), color='black') +
+	geom_label(data=cs2.sel.grid.sub.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
 		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
+		strip.text = element_text(color='white', size=8),
 		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
@@
 	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.genetic.selection), color='black') +
-	geom_text(data=cs2.sel.grid.sub.letters.DF[2,], aes(x=x, y=y, label=letter), color='black') +
+	geom_label(data=cs2.sel.grid.sub.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
 	theme_classic() +
@@
 		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-		legend.position=c(1, 0), legend.justification=c(1,0), legend.key.size=unit(3, 'mm')
+		strip.text = element_text(color='white', size=8),
+		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'), legend.position='none'
 	) +
@@
 p3 <- ggplot() +
-	geom_point(data=cs2.sel.space.DF, aes(x=d1, y=d2, fill=Amount.selection), pch=21, size=3, color='black') +
-	geom_text(data=cs2.sel.space.letters.DF[1,], aes(x=x, y=y, label=letter), color='black') +
-	theme_classic() + xlab('Genetic NMDS Axis 1') + ylab('Genetic NMDS Axis 2') +
-	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line()) +
+	geom_point(data=arrange(cs2.sel.space.DF, Amount.selection), aes(x=d1, y=d2, fill=Amount.selection), pch=21, size=3, color='black') +
+	geom_label(data=cs2.sel.space.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
+	theme_classic() + xlab('NMDS 1') + ylab('NMDS 2') +
+	scale_x_continuous(expand=c(0.08, 0)) + scale_y_continuous(expand=c(0.08, 0)) + 
+	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line(),
+    axis.text=element_text(size=8), axis.title=element_text(size=8), plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
 	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
 p4 <- ggplot() +
-	geom_point(data=cs2.sel.space.DF, aes(x=d1, y=d2, fill=Amount.genetic.selection), pch=21, size=3, color='black') +
-	geom_text(data=cs2.sel.space.letters.DF[2,], aes(x=x, y=y, label=letter), color='black') +
-	theme_classic() + xlab('Genetic NMDS Axis 1') + ylab('Genetic NMDS Axis 2') +
-	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line()) +
+	geom_point(data=arrange(cs2.sel.space.DF, Amount.genetic.selection), aes(x=d1, y=d2, fill=Amount.genetic.selection), pch=21, size=3, color='black') +
+	geom_label(data=cs2.sel.space.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
+	theme_classic() + xlab('NMDS 1') + ylab('NMDS 2') +
+	scale_x_continuous(expand=c(0.08, 0)) + scale_y_continuous(expand=c(0.08, 0)) + 
+  theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line(),
+    axis.text=element_text(size=8), axis.title=element_text(size=8), plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
 	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))

### 2016-11-26 code/rmarkdown/figures.Rmd
+ update checkpoint date to use raptr 0.0.3

@@
 # load .rda file
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
 try(session::restore.session('data/final/results.rda'))

### 2016-11-26 code/rmarkdown/text.Rmd
+ update checkpoint date to use raptr 0.0.3

@@
 # load .rda file
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/final/results.rda')

### 2016-11-26 code/R/analysis/00-initialization.R

@@ -12,7 +12,7 @@ if (length(args)>0) {
 # set checkpoint
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1')
+checkpoint::checkpoint('2016-11-26', R.version='3.3.1')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 
 # load CRAN packages
@@ -36,7 +36,6 @@ library(dplyr)
 library(tidyr)
 library(testthat)
 
-
 library(raptr)
 library(rgeos)
 library(rworldxtra)
@@ -54,7 +53,6 @@ library(knitr)
 library(lazyWeave)
 library(broom)
 library(rmarkdown)
-
 
 ## load github packages
 if (!'ALA4R' %in% installed.packages()[,'Package']) {

### 2016-11-26 code/R/analysis/01-load-data.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/00-initialization.rda')
 
 ### load simulation data

### 2016-11-26 code/R/analysis/02-simulations.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/01-load-data.rda')
 
 ## load parameters

### 2016-11-26 code/R/analysis/05-statistical-analysis.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/04-case-study-2.rda')
 load('data/intermediate/03-case-study-1.rda')
 load('data/intermediate/02-simulations.rda')

### 2016-11-26 code/R/analysis/06-exports.R

@@ -1,5 +1,5 @@
  ## load .rda
-checkpoint::checkpoint('2016-11-01', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
 session::restore.session('data/intermediate/05-statistical-analysis.rda')
 
 ## save workspace

### 2016-11-26 code/rmarkdown/article.Rmd
+ minor fixes to get article to compile

@@
 documentclass: article
+subparagraph: true
 bibliography: references.bib

### 2016-11-27 code/rmarkdown/article.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

@@
 1.
-An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity.
+An underlying aim in conservation is to maximize the long-term persistence of biodiversity.
 To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved.
@@
 species, ecosystem) is to preserve a representative sample of the physical attributes that underpin them across the feature's geographic distribution.
-For example, preserving the adaptive processes currently acting on a species might be crucial for its long-term persistence in a world where environmental change is accelerating.
-By conserving individuals with the ability to persist in a range of climatic conditions--physical attributes associated with adaptation--protected areas can foster evolutionary processes.
-Despite this, current approaches overwhelmingly focus on securing a target proportion of each features' geographic range and little attention has been directed towards targeting a representative sample of them.
+For example, the long-term persistence of a species may depend on its ability to adapt to new climatic conditions.
+By conserving individuals with the ability to persist in a range of conditions--physical attributes associated with adaptation--protected areas can foster evolutionary processes.
+Despite this, current reserve selection methods overwhelmingly focus on securing an adequate proportion of features' geographic ranges and little attention has been directed towards capturing a representative sample.
 
 2.
-To address this issue, we developed the \\texttt{raptr: Representative and Adequate Prioritization Toolkit in R} to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
-Users set \"amount targets\"--similar to conventional reserve selection methods--to secure a target proportion of the features' geographic distributions.
-Additionally, users set attribute \"space targets\" to ensure that reserve systems also secure a representative sample of ecologically or evolutionary relevant attributes (space) across the features' geographic distributions (eg.
-an attribute space expressing variation in genetic characteristics between individuals, or variation in climatic conditions between areas).
+To address this issue, we developed the \\texttt{raptr R} package to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
+Users set "amount targets"--similar to conventional methods--to secure a target proportion of the features' geographic distributions.
+Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
+climatic or genetic variation).
 We demonstrate the functionality of this package using simulations and two case studies.
-We generate solutions that secure a proportion of the species' distributions using amount targets--representing conventional reserve selection--and compare them with solutions generated using both amount and attribute space targets.
+We generate solutions that secure a proportion of the species' geographic distributions using amount targets--representing conventional methods--and compare them with solutions generated using amount and space targets.
 
 3.
-We show that markedly different solutions emerge when explicitly considering within feature representation in reserve selection.
-Our simulations suggest that including representativeness in conservation planning is important where biodiversity features have multimodal distributions in an attribute space.
-The case studies show that explicitly setting space-based targets can result in solutions that preserve a much more representative sample of features.
-Additionally, the case studies suggest that securing a representative sample of features may not require a much greater area of reservation than traditional approaches.
+We show that different solutions emerge when explicitly considering within feature representation in reserve selection.
+Our simulations suggest that targeting representativeness is particularly important for biodiversity features that have multimodal distributions in an attribute space.
+The case studies show that setting space targets can result in solutions that conserve a much more representative sample of features--and this can be achieved with only a slight increase in reserve size.
 
 4.
-The \\texttt{raptr R} package provides a unified framework for achieving spatial conservation prioritizations that secure an adequate and representative sample of features.
-Prioritizations that achieve this are likely to result in a greater chance of achieving long-term biodiversity persistence than area-based planning alone.
+The \\texttt{raptr R} package provides a toolkit for making spatial conservation prioritizations that secure an adequate and representative sample of features.
+By targeting representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
 
@@
 
-Short running title: Representative and adequate prioritizations \newline
-Abstract word count: XXX / XXX  
-Total word count: XXX / XXX  
-Number of references: XXX  
-Number of figures: XXX  
+Short running title: Representative and adequate prioritizations  
+Abstract word count: 350 / 350  
+Total word count: 6977 / 7000  
+Number of references: 42  
+Number of figures: 8  
 Number of tables: 0  
 Number of textboxes: 0  
-Number of equations: XXX
+Number of equations: 4
 

### 2016-11-27 code/rmarkdown/figures.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

@@
 ```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Example of an attribute space.
-This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm) values.
+This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm).
 Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found.
 Points represent demand points.
-In this space, populations closer to each other are considered more similar to each other."}
+In this space, populations closer to each other have similar environmental conditions."}
 ## generate data
@@
 
-```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.6, fig.width=column.width, fig.cap='Simulated distributions of three species.
+```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.6, fig.width=column.width, fig.cap='Distributions of three simulated species.
 Squares denote planning units.
-The color of each planning unit indicates the probability that each species inhabits it.
-Panel (a) shows the uniformly distributed species, panel (b) shows the normally distributed species, and panel (c) shows the bimodally distributed species.'}
+The color of each planning unit indicates the probability that species inhabit it.'}
 # prepare data
@@
 
-```{r, echo=FALSE, fig.height=3.5, fig.width=2*column.width, fig.cap=paste0('Simulation study prioritizations.
+```{r, echo=FALSE, fig.height=3.5, fig.width=2*column.width, fig.cap=paste0('Prioritizations for the simulation study.
 Each panel shows a prioritization generated for a single species using a set of parameters.
-Squares denote planning units, with dark green showing those planning units selected for protection in a particular prioritisation.
+Squares denote planning units.
+Dark green units show planning units selected for protection.
 The top row of panels (a, d, g, j) shows prioritizations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
@@
 Squares denote planning units.
-The panels show the first and second main gradients of climatic variation.'}
+Panels show the two main gradients of climatic variation.'}
 # prepare data
@@
 ```{r, echo=FALSE, fig.height=2.35, fig.width=column.width, fig.cap='Distribution of the species used in the first case-study.
+Each panel shows a different species.
 Squares denote planning units.
-Each panel shows the distribution of a different species.
 Planning units occupied by a species are shown in bright blue.'}
@@
 
-```{r, echo=FALSE, fig.height=1.175, fig.width=column.width, fig.cap=paste0('Prioritizations for the niche-based case-study in Queensland, Australia.
+```{r, echo=FALSE, fig.height=1.175, fig.width=column.width, fig.cap=paste0('Prioritizations for the first case study.
 Squares represent planning units.
-Dark green planning units for selected for protection and pale green units were discarded.
+Dark green planning units were selected for protection.
 Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
@@
 
-```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Study area and data used in the second case-study.
+```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Data used for the second case-study.
 Squares denote planning units.
@@
 
-```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Prioritizations for the genetic case-study in the European Alps.
+```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Prioritizations for the second case study.
 Panels (a--b) show prioritisations generated using different parameters.
 Squares denote planning units shown in geographic space.
+Dark green planning units were selected for protection.
 Panel (a) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets.
@@
 The coordinates of the points represent the typical genetic characeristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data).
-Planning units associated with points that are closer togeather contain individuals with more similar genetic characteristics than planning units that are further apart.
-In all panels, dark green planning units were selected for protection, and pale green planning units were discarded.')}
+Planning units associated with points that are closer togeather contain individuals with more similar genetic characteristics than planning units that are further apart.')}
 # prepare spatial data

### 2016-11-27 code/rmarkdown/supporting-information.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

@@
 However, the $d$ variables are in distance units which effectively meaningless in this context.
-Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).
+Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (eqn 5).
 
 \begin{align*}
-& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag*{eqn 4} \\
+& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag*{eqn 5} \\
 & & & 0 \leq s \leq S-1\\
@@
 \begin{align*}
-R &= \parbox{25em}{number of failure levels} \tag*{eqn 5a} \\
+R &= \parbox{25em}{number of failure levels} \tag*{eqn 6a} \\
 %
-M &= \parbox{25em}{failure multiplier} \tag*{eqn 5b}\\
+M &= \parbox{25em}{failure multiplier} \tag*{eqn 6b}\\
 %
@@
 Y_{fsijr} &= \begin{cases}
-		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag*{eqn 6a} \\
+		1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag*{eqn 7a} \\
 		0, & \parbox{25em}{otherwise} \\
@@
 %
-P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag*{eqn 6a}\\
+P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag*{eqn 7b}\\
 %
@@
 %
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} P_{fsijr} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 7a}\\
+& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} P_{fsijr} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 8a}\\
 & & & & & 0 \leq s \leq S-1\\
 %
-& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 7b}\\
+& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 8b}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 7c}\\
+& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 8c}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 7d}\\
+& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 8d}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 7e}\\
+& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 8e}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag*{eqn 7f}\\
+& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag*{eqn 8f}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 7g}\\
+& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 8g}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 %
-& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 7h}\\
+& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 8h}\\
 & & & & & 0 \leq s \leq S-1,\\
@@
 The objective function for the reliable formulation is the same as for the unreliable formation (eqn 3a).
-Similar to the unreliable formulation, constraints (eqn 3b) and (eqn 7a) ensure that the amount-based and space-based targets are met.
-Constraint (eqns 7b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$.
-Constraints (eqn 7d) ensure that only selected planning units are assigned to demand points $i \in I$.
-Constraints (eqn 7e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level.
-Constraints (eqns 7f--7g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information].
-Constraints (eqn 7h) ensure that the $X$ and $Y$ variables are binary.
+Similar to the unreliable formulation, constraints (eqn 3b) and (eqn 8a) ensure that the amount-based and space-based targets are met.
+Constraint (eqns 8b--7c) ensure that each planning unit is only assigned to one backup $r$-level for $i \in I$.
+Constraints (eqn 8d) ensure that only selected planning units are assigned to demand points $i \in I$.
+Constraints (eqn 8e) ensure that the imaginary planning unit is always assigned to the highest backup $r$-level.
+Constraints (eqns 8f--8g) determine the probability that planning unit $j$ will be used to sample demand point $i \in I$ for $s \in S$ and $f \in F$ [see @r16 for more information].
+Constraints (eqn 8h) ensure that the $X$ and $Y$ variables are binary.
 
@@
 [-@r426].
-Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 7a) can be linearized using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
+Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 8a) can be linearized using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al.
 [-@r16].

### 2016-11-27 code/rmarkdown/text.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

@@
 2a).
-This species' distribution was based on a uniform distribution (eqn 8a).
+This species' distribution was based on a uniform distribution (eqn 4a).
 The second species was simulated to represent a species with an idealized distribution (Fig.
@@
 It has a core range area, and is less likely to be found in areas that are distant from the core area.
-This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 8c).
+This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 4c).
 The third species represents a species with two distinct populations (Fig.
 2c).
-This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 8d).
+This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 4d).
 The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d.
 We used a geographic attribute space to showcase the behavior of the problem.
-For each species, demand points were generated by calculating the centroids of each planning units and their weights were set as the probability that the planning units were occupied.
+For each species, demand points were set as the planing units' centroids and were weighted according to the units' probability of occupancy.
 
@@
 \\
-P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
+P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 4a} \\
 %
-f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
+f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 4b} \\
 %
-P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
+P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4c} \\
 %
@@
 f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
-\end{cases} \tag*{eqn 8d} \\
+\end{cases} \tag*{eqn 4d} \\
 %
@@
 We generated four solutions for each species.
-First to represent solutions generated using conventional conservation planning methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-Second to show how the addition of geographic targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets.
-Third to represent solutions using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
+First to represent solutions generated using conventional methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
+Second to show how the addition of representation targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets.
+Third to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
 Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated solutions using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
@@
 3a).
-Overall, these results show that under the simplest of conditions, the reliance on just amount targets can cause an under-specified reserve selection problem that is unlikely to return a suitable solution for implementation.
+Overall, these results show that under the simplest of conditions, using just amount targets results in an under-specified reserve selection problem that is unlikely to return an effective solution.
 
@@
 This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution.
-By using amount and geographic targets in the reserve selection problem, we have obtained a solution that concentrates conservation effort in the range core and also the range margin.
+By using amount and geographic targets in the reserve selection problem, we obtained a solution that concentrates conservation effort in the range core and also the range margin.
 However, similar to the corresponding solution for the uniform species (Fig.
@@
 2c).
-The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets assigned conservation effort to just one of the two populations (Fig.
+The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets prioritized individuals in just one of the two populations (Fig.
 3i).
@@
 ### Methods
-We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the realized niche for several species.
+We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches.
 By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63].
-We selected Queensland, Australia as the study area.
-This region is ideal for exploring the potential of niche-based targets because it contains a broad range of different habitats.
+We selected Queensland, Australia as the study area because it contains a broad range of different habitats.
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
-We used the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components (cumulatively explained `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation) to characterize the environmental variation across the study area (Fig.
+We used the scores from the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation; Fig.
 4).
 
-We selected four bird species that span a range of different ecological requirements, evolutionary histories, distributions: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_).
-We then mapped the extent of occurrence for each species (Fig.
+We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_).
+These species span a range of different habitat requirements.
+We mapped the extent of occurrence for each species (Fig.
 5).
-To do this, we obtained observation data from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
 We used this method because it is entirely reproducible using freely available data.
@@
 S2).
-We used the following method to ensure that the demand points reflected the core parts of the species' realized niches.
-For each species, we generated random points inside the species range and at each point extracted the principal component values at that location.
+To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method.
+For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location.
 We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231].
-We then generated uniformly distributed points inside the species' kernels and estimated the density of the training points at the uniformly generated points.
-These uniformly distributed points and associated density estimated were used as demand point coordinates and weights (respectively).
+We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations.
+These uniformly distributed points and associated density estimations used as demand point coordinates and weights (respectively).
 
 We generated two solutions.
-The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount-based targets for all species.
-The second solution was generated the same amount-based target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.
+The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount targets for all species.
+The second solution was generated the same amount target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.
 
 ### Results
-Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches--with `r ifelse(length(amount.not.represented.species.names)>1, as.character(as.english(length(amount.represented.species.names))), 'one')` excpetion.
-Specifically, this solution sampled over `r cs1.params.LST[[MODE]]$space.target*100` % of the realized niche of `r parsed.representative.space.held.names`.
-Yet it failed to achieve this for `r parsed.not.representative.space.held.names`.
-This result demonstrates that though solutions generated using conventional methods may preserve a representative sample of some features (Figure S1), only through the use of explicit targets can decision makers be sure that solutions secure a representative sample for all features.
+Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches.
+(Figure S1).
+This solution secured over `r cs1.params.LST[[MODE]]$space.target*100` % of the realized niche of `r parsed.representative.space.held.names`.
+Yet it failed to achieve this for the `r parsed.not.representative.space.held.names`.
+This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.
 
@@
 We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
-Although this dataset contains multiple plant species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}), a wide-spread species with significant genetic structure [@r462] as a simple case study.
+Although this dataset contains multiple species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}) as a simple case study, because it has been found to exhibit significant genetic structure [@r462] as a simple case study.
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 7a).
-They visited every second grid cell, and if the `r tolower(cs2.params.LST[[MODE]]$common.name)` was detected in a cell, samples were collected from three individuals.
+They visited every second grid cell, and if the species was detected in a cell, samples were collected from three individuals.
 Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
@@
 These continuous variables described the main axes of genetic variation within the species.
-We calculated the average of the values associated with individuals in each grid cell.
-These values were used to create a genetic attribute space (Figs.
+We averaged together values corresponding to samples collected in the same grid cell, and used them to create a genetic attribute space (Figs.
 7c and 7d).
-To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances based on the grid cells' centroids [using the \texttt{ape R} package; @r464].
+To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using the grid cells' centroids and inverse great circle distances [using the \texttt{ape R} package; @r464].
 
-The grid cells were used as planning units.
+The grid cells were used as planning units for generating prioritizations.
 The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
 Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points.
-To ensure that the solutions did not prioritize particularly costly areas, we included opportunity cost data (Fig.
+To ensure that the solutions did not prioritize particularly costly areas, we obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
+We used these values to represent opportunity cost (Fig.
 7b).
-We obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
 
@@
 8a and 8c).
-This solution only sampled planning units that contained individuals belonging to one of the two main groups.
-On the other hand, the solution generated using additional genetic targets selected planning units belonging to both of the main groups.
-Note that since these solutions were generated using opportunity cost data, the solution generated using only the amount target is essentially the minimum number of least costly areas needed to fulfill the amount target.
-The only difference between the two solutions is a single planning unit.
-By swapping this single planning unit, the solution generated using amount and genetic targets was able to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs.
-8b and 8d) for only a minor increase in cost (`r round(cs2.spp.DF$Score[1],2)` total cost compared to `r round(cs2.spp.DF$Score[2],2)` total cost).
+This solution only conserved individuals in one of the two main genetic groups.
+In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target.
+Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs.
+8b and 8d).
+Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2.spp.DF$Score[2],2)` compared to `r round(cs2.spp.DF$Score[1],2)` total cost).
 
@@
 
-One of the key advantages of this package is that it is general enough to incorporate any spatially explicit variable as an attribute space.
+One of the key advantages of this package is that it is general enough to incorporate almost any spatially explicit variable.
 This package can accommodate intrinsic or extrinsic variation in the feature(s).
 For example, adaptation processes could be secured using environmental variation [eg.
-@r247].
-Predator-prey processes could be conserved by capturing the overlapping distributions of predators and each of their prey species [eg.
+@r247], and trophic processes could be conserved by capturing the overlapping distributions of predator and prey species [eg.
 @r471; @r485].
@@
 @r483].
-As long as the variation can be described using Euclidean distances--which could be achieved by transforming or applying dimension reduction techniques--the \texttt{R} package can be used to obtain a representative sample.
+As long as the variation can be described using Euclidean distances--which could be achieved through transformation or dimension reduction--the \texttt{R} package can be used to obtain a representative sample.
 
-The degree to which a prioritization truly secures a representative sample of a feature depends on the attribute spaces and the distribution of demand points chosen by the decision maker.
-Ultimately, the space-based targets are set as a proportion based on the distribution of the demand points.
-As a consequence, if the planner uses an inappropriate set of spatial variables to construct an attribute space, or an inappropriate set of demand points, then the optimal solution will not effectively represent the decision maker's objectives.
-We therefore stress that planners must carefully consider which biodiversity processes need to be reflected in the prioritization, and which spatial data can be used to represent these processes.
-Note that maximizing for one attribute space can result in trading off against another.
-For instance, to maximize the geographic spread of a prioritization, reserves need to be further away from each other, yet to maximize connectivity, reserves may need to be closer together.
+The degree to which a prioritization truly secures a representative sample of a feature depends on the attribute space(s) and the distribution of demand points chosen by the conservation planner.
+The optimal solution will not effectively represent the decision maker's objectives if an inapropriate set of spatial variables or demand points are used to represent the attribute space.
 Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space.
-To assist the generation of appropriate demand points, the \texttt{R} package provides several routines for generating demand points (see documentation for the `make.DemandPoints` function).
-These routines use kernels to automatically generate demand points based on the distribution of features in spatial data sets [using @r231; @r450].
+To assist the generation of appropriate demand points, the \texttt{R} package provides several routines that use the distribution of the features in spatial data sets to make demand points (see documentation for the `make.DemandPoints` function).
 
-The formulation requires spatially comprehensive data to map attribute spaces to planning units.
+Attribute data need to be spatially comprehensive to map planning units to attribute spaces.
 However, many real-world data sets are patchy.
-For example, due to constrained resources, genetic data may have only been collected in some planning units.
-To use such patchy data with the \texttt{raptr R} package, the planning units that are missing data must either be omitted (as we did in the second case-study), or their missing values can be estimated.
-Missing data could be estimated using spatially explicit models [eg.
+For example, due to constrained resources, genetic data may only be available for some planning units.
+To use such patchy data, planners could omit the planning units that are missing data, or estimate the missing data using spatially explicit models [eg.
 krigging or generalized dissimilarity models; @r467; @r466].

### 2016-11-27 code/R/analysis/05-statistical-analysis.R

@@ -36,8 +36,8 @@ cs1.nmds.MoransI <- llply(
 )
 
 # calculate numbers and generate vectors for article
-amount.represented.species.names <- filter(cs1.spp.DF, Prioritisation=='Amount targets', niche >= (cs1.params.LST[[MODE]]$'space.target'*100))$Species
-amount.not.represented.species.names <- filter(cs1.spp.DF, Prioritisation=='Amount targets', niche<(cs1.params.LST[[MODE]]$'space.target'*100))$Species
+amount.represented.species.names <- tolower(filter(cs1.spp.DF, Prioritisation=='Amount targets', niche >= (cs1.params.LST[[MODE]]$'space.target'*100))$Species)
+amount.not.represented.species.names <- tolower(filter(cs1.spp.DF, Prioritisation=='Amount targets', niche<(cs1.params.LST[[MODE]]$'space.target'*100))$Species)
 
 if (length(amount.represented.species.names)>1) {
 	parsed.representative.space.held.names <- sapply(

### 2016-11-28 code/rmarkdown/article.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

@@
 abstract: "
-
 1.
@@
 To address this issue, we developed the \\texttt{raptr R} package to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
-Users set "amount targets"--similar to conventional methods--to secure a target proportion of the features' geographic distributions.
-Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
+Users set \"amount targets\"--similar to conventional methods--to secure a sufficient proportion of the features' geographic distributions.
+Additionally, users set \"space targets\" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
 climatic or genetic variation).
 We demonstrate the functionality of this package using simulations and two case studies.
-We generate solutions that secure a proportion of the species' geographic distributions using amount targets--representing conventional methods--and compare them with solutions generated using amount and space targets.
+We generate solutions using just amount targets--representing conventional methods--and compare them with solutions generated using amount and space targets.
 
@@
 By targeting representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
-
 "
@@
 Short running title: Representative and adequate prioritizations  
-Abstract word count: 350 / 350  
-Total word count: 6977 / 7000  
+Abstract word count: 345 / 350  
+Total word count: 6999 / 7000  
 Number of references: 42  

### 2016-11-28 code/rmarkdown/figures.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

@@
 
-```{r, echo=FALSE, fig.height=2.35, fig.width=column.width, fig.cap='Distribution of the species used in the first case-study.
+```{r, echo=FALSE, fig.height=2.35, fig.width=column.width, fig.cap='Distribution of the species used in the first case study.
 Each panel shows a different species.
@@
 Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
-Panel (b) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets')}
+Panel (b) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets.')}
 # prepare data
@@
 
-```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Data used for the second case-study.
+```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Data used for the second case study.
 Squares denote planning units.
@@
 Grid cells occupied by ',tolower(cs2.params.LST[[MODE]]$common.name),' are shown in bright blue.
-The subsequent panels contain only the grid cells that were occupied by the species.
+The subsequent panels contain only show occupied grid cells.
 Panel (b) shows the opportunity cost of each planning unit (estimated as the total human population density).
@@
 ```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Prioritizations for the second case study.
-Panels (a--b) show prioritisations generated using different parameters.
+Panels (a--b) show prioritizations generated using different parameters.
 Squares denote planning units shown in geographic space.
@@
 Each point corresponds to a planning unit.
-The coordinates of the points represent the typical genetic characeristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data).
-Planning units associated with points that are closer togeather contain individuals with more similar genetic characteristics than planning units that are further apart.')}
+The coordinates of the points represent the typical genetic characteristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data).
+Planning units associated with points that are closer together contain individuals with more similar genetic characteristics than planning units that are further apart.')}
 # prepare spatial data

### 2016-11-28 code/rmarkdown/text.Rmd
+ add in document statistics for article + reduce length of abstract, main text and figures to within limits

@@
 [-@r426].
-Linearized versions of the problems can be solved using commercial exact algorithms.
-The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation).
-These optimization problems can then be solved to generate prioritizations using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).
+The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).
 Presently, academics can obtain a license at no cost from the Gurobi website.
@@
 
-To understand the behavior of the unreliable problem and showcase its value, we conducted a simulation study and two case studies.
+To understand and showcase the behavior of the unreliable problem, we conducted a simulation study and two case studies.
 These studies involved generating solutions using only amount targets to represent prioritizations generated using conventional methods (eg.
@@
 3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area.
-This specific configuration is an artifact of the \texttt{Gurobi} solver used to solve the problem.
+This specific configuration is an artifact of the method used to solve the problem.
 In fact, for this species, all prioritizations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-In the absence of criteria to guarantee representativeness, reserve selection methods may or may not return solutions that secure a representative sample of the features.
+In the absence of additional criteria, reserve selection methods may or may not return solutions that secure a representative sample of the features.
 In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).

### 2016-11-28 code/R/analysis/05-statistical-analysis.R

@@ -36,8 +36,8 @@ cs1.nmds.MoransI <- llply(
 )
 
 # calculate numbers and generate vectors for article
-amount.represented.species.names <- tolower(filter(cs1.spp.DF, Prioritisation=='Amount targets', niche >= (cs1.params.LST[[MODE]]$'space.target'*100))$Species)
-amount.not.represented.species.names <- tolower(filter(cs1.spp.DF, Prioritisation=='Amount targets', niche<(cs1.params.LST[[MODE]]$'space.target'*100))$Species)
+amount.represented.species.names <- filter(cs1.spp.DF, Prioritisation=='Amount targets', niche >= (cs1.params.LST[[MODE]]$'space.target'*100))$Species
+amount.not.represented.species.names <- filter(cs1.spp.DF, Prioritisation=='Amount targets', niche<(cs1.params.LST[[MODE]]$'space.target'*100))$Species
 
 if (length(amount.represented.species.names)>1) {
 	parsed.representative.space.held.names <- sapply(
@@ -64,6 +64,9 @@ if (length(amount.not.represented.species.names)>1) {
 } else {
 	parsed.not.representative.space.held.names=c()
 }
+
+amount.represented.species.names <- tolower(amount.represented.species.names)
+amount.not.represented.species.names <- tolower(amount.not.represented.species.names)
 
 ## save workspace
 save.session('data/intermediate/05-statistical-analysis.rda', compress='xz')

### 2016-11-28 code/rmarkdown/article.Rmd
+ fix eqn numbering + more small changes to wording

@@
 1.
-An underlying aim in conservation is to maximize the long-term persistence of biodiversity.
+An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity.
 To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved.
@@
 Short running title: Representative and adequate prioritizations  
-Abstract word count: 345 / 350  
-Total word count: 6999 / 7000  
+Abstract word count: 346 / 350  
+Total word count: 6974 / 7000  
 Number of references: 42  

### 2016-11-28 code/rmarkdown/text.Rmd
+ fix eqn numbering + more small changes to wording

@@
 This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 4d).
-The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 8a--8d.
+The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 4a--4d.
 We used a geographic attribute space to showcase the behavior of the problem.
@@
 3d).
-This solution contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--a few more than the previously discussed solutions--to ensure that the solution secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented.
-This result suggests that the combination of space targets and boundary length modifiers may yield solutions that contain more planning units, since the solution generated using amount targets and a boundary length modifier (Fig.
-3c) contained the same number of planning units as the solution with just the amount targets (Fig.
-3a).
+Although this solution also contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--the same amount as previously discussed solutions--it secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented.
 Overall, these results show that under the simplest of conditions, using just amount targets results in an under-specified reserve selection problem that is unlikely to return an effective solution.
@@
 This solution contains `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units.
-Similar to the corresponding solution for the uniform species (Fig.
-3d), this solution contains more planning units than any other solutions for this species (cf.
+Unlike the solution for the uniform species generated using the same parameters, however, this solution contains more planning units than any other solutions for this species (cf.
 Figs.
 3e--3g).
-The results for the normally distributed species suggest that the space targets can result in solutions that secure a more representative sample of the species--even for species without significant spatial structure.
+These results suggests that the combination of space targets and boundary length modifiers may require solutions that contain more planning units.
+They also also suggest that space targets can result in solutions that secure a more representative sample of the species--even for species without significant spatial structure.
 
@@
 Although this solution required more planning units to fulfill both targets ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$n`$), this solution secured a representative sample of the bimodally distributed species (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % range sampled).
-Similar to corresponding solutions generated for both the uniformly (Fig.
-3b) and normally distributed species (Fig.
+Similar to corresponding solution generated for the normally distributed species (Fig.
 3f), this solution was also highly fragmented and we could obtain a more well-connected solution at the expense of selecting additional planning units ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n`$; Fig.
@@
 @r471; @r485].
+Note that maximizing one process can result in trading off against another (eg.
+geographic spread vs.
+connectivity).
 Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserves so there is equitable access to parks by people [eg.

### 2016-11-28 code/rmarkdown/article.Rmd
+ add tweetable abstract

@@
 Short running title: Representative and adequate prioritizations
+Tweetable abstract: Make smarter spatial conservation decisions that foster biodiversity processes using raptr R package
 Abstract word count: 346 / 350  

### 2016-11-28 code/rmarkdown/article.Rmd
+ add spaces after tweetable abstract for new line
+ add spaces after short title
+ move abstract onto seperate page

--- 2016-11-28 code/rmarkdown/article.Rmd
@@
 tags: ["conservation", "biodiversity", "MILP", "reserve-selection", "optimization", "protected area"]
-abstract: "
-1.
-An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity.
-To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved.
-One way to conserve such processes at the feature level (eg.
-species, ecosystem) is to preserve a representative sample of the physical attributes that underpin them across the feature's geographic distribution.
-For example, the long-term persistence of a species may depend on its ability to adapt to new climatic conditions.
-By conserving individuals with the ability to persist in a range of conditions--physical attributes associated with adaptation--protected areas can foster evolutionary processes.
-Despite this, current reserve selection methods overwhelmingly focus on securing an adequate proportion of features' geographic ranges and little attention has been directed towards capturing a representative sample.
-
-2.
-To address this issue, we developed the \\texttt{raptr R} package to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
-Users set \"amount targets\"--similar to conventional methods--to secure a sufficient proportion of the features' geographic distributions.
-Additionally, users set \"space targets\" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
-climatic or genetic variation).
-We demonstrate the functionality of this package using simulations and two case studies.
-We generate solutions using just amount targets--representing conventional methods--and compare them with solutions generated using amount and space targets.
-
-3.
-We show that different solutions emerge when explicitly considering within feature representation in reserve selection.
-Our simulations suggest that targeting representativeness is particularly important for biodiversity features that have multimodal distributions in an attribute space.
-The case studies show that setting space targets can result in solutions that conserve a much more representative sample of features--and this can be achieved with only a slight increase in reserve size.
-
-4.
-The \\texttt{raptr R} package provides a toolkit for making spatial conservation prioritizations that secure an adequate and representative sample of features.
-By targeting representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
-"
 output:
@@
 Short running title: Representative and adequate prioritizations  
+Short abstract: Make smarter spatial conservation decisions that foster biodiversity processes using raptr R package  
 Abstract word count: 346 / 350  
@@
 \clearpage
+
+1.
+An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity.
+To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved.
+One way to conserve such processes at the feature level (eg.
+species, ecosystem) is to preserve a representative sample of the physical attributes that underpin them across the feature's geographic distribution.
+For example, the long-term persistence of a species may depend on its ability to adapt to new climatic conditions.
+By conserving individuals with the ability to persist in a range of conditions--physical attributes associated with adaptation--protected areas can foster evolutionary processes.
+Despite this, current reserve selection methods overwhelmingly focus on securing an adequate proportion of features' geographic ranges and little attention has been directed towards capturing a representative sample.
+
+2.
+To address this issue, we developed the \\texttt{raptr R} package to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
+Users set "amount targets"--similar to conventional methods--to secure a sufficient proportion of the features' geographic distributions.
+Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
+climatic or genetic variation).
+We demonstrate the functionality of this package using simulations and two case studies.
+We generate solutions using just amount targets--representing conventional methods--and compare them with solutions generated using amount and space targets.
+
+3.
+We show that different solutions emerge when explicitly considering within feature representation in reserve selection.
+Our simulations suggest that targeting representativeness is particularly important for biodiversity features that have multimodal distributions in an attribute space.
+The case studies show that setting space targets can result in solutions that conserve a much more representative sample of features--and this can be achieved with only a slight increase in reserve size.
+
+4.
+The \\texttt{raptr R} package provides a toolkit for making spatial conservation prioritizations that secure an adequate and representative sample of features.
+By targeting representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
+
+\clearpage 
 \input{./text.tex}

### 2016-11-29 code/rmarkdown/article.Rmd
+ finalise ms format for mee submission

@@
 2.
-To address this issue, we developed the \\texttt{raptr R} package to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
+To address this issue, we developed the \texttt{raptr R} package to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
 Users set "amount targets"--similar to conventional methods--to secure a sufficient proportion of the features' geographic distributions.
@@
 4.
-The \\texttt{raptr R} package provides a toolkit for making spatial conservation prioritizations that secure an adequate and representative sample of features.
+The \texttt{raptr R} package provides a toolkit for making spatial conservation prioritizations that secure an adequate and representative sample of features.
 By targeting representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
@@
 \input{./figures.tex}
-
-\clearpage 
-\input{./supporting-information.tex}

### 2016-11-29 code/rmarkdown/supporting-information.Rmd
+ finalise ms format for mee submission

@@
 ---
+title: "Supporting Information"
+fontsize: 11pt
+documentclass: article
+subparagraph: true
 bibliography: references.bib
@@
     keep_tex: yes
+    includes:
+      in_header: preamble.tex
 ---
-# Supporting Information
 

### 2016-11-29 code/rmarkdown/article.Rmd
+ update formatting

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-tags: ["conservation", "biodiversity", "MILP", "reserve-selection", "optimization", "protected area"]
 output:
@@
 
-Short running title: Representative and adequate prioritizations  
-Short abstract: Make smarter spatial conservation decisions that foster biodiversity processes using raptr R package  
 Abstract word count: 346 / 350  
@@
 Number of equations: 4  
+Short running title: Representative and adequate prioritizations  
+Short abstract: Make smarter spatial conservation decisions using the raptr R package https.github.com/jeffreyhanson/raptr  
+Keywords: conservation, biodiversity, MILP, reserve-selection, optimization, protected area
 
 \clearpage
+
+# Summary
 

### 2016-11-29 code/rmarkdown/text.Rmd
+ update formatting

@@
 ```
+
+# Introduction
 
@@
 [-@r426].
-The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{http://www.gurobi.com}).
+The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
 Presently, academics can obtain a license at no cost from the Gurobi website.
@@
 5).
-To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{www.github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
 We used this method because it is entirely reproducible using freely available data.
@@
 The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
-All data, code, and results are stored in an online repository (\url{https://www.github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
+All data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
 

### 2016-11-29 code/rmarkdown/cover-letter.tex
+ fix typo in cover letter
+ remove the github repo from the cover letter because this will be made available once the repo has been accepted

@@
 We desire to communicate our findings to the global community of ecologists and conservation biologists that follow this journal.
-The content of our manuscript has not been published elsewhere has not been published elsewhere, nor has it been submitted for publication to another journal.
+The content of our manuscript has not been published elsewhere, nor has it been submitted for publication to another journal.
 All authors approved the manuscript and this submission.
-In support of the journal's commitment to reproducible research, we have stored all data, code, and results in an online repository to permit replication and validation of this study (\url{https://www.github.com/jeffreyhanson/raptr-manuscript}).
 

### 2017-02-07 code/rmarkdown/article.Rmd
address some of the reviewers comments add in benchmark analysis

@@
 
-Abstract word count: 346 / 350  
-Total word count: 6974 / 7000  
+Abstract word count: XXXX / 350  
+Total word count: XXXX / 7000  
 Number of references: 42  
@@
 Number of equations: 4  
-Short running title: Representative and adequate prioritizations  
-Short abstract: Make smarter spatial conservation decisions using the raptr R package https.github.com/jeffreyhanson/raptr  
+Short running title: 
+Short abstract: Capture biodiversity processes and patterns in prioritizations using raptr https://github.com/jeffreyhanson/raptr  
 Keywords: conservation, biodiversity, MILP, reserve-selection, optimization, protected area
@@
 One way to conserve such processes at the feature level (eg.
-species, ecosystem) is to preserve a representative sample of the physical attributes that underpin them across the feature's geographic distribution.
+species, ecosystem) is to preserve a sample of the feature (eg.
+individuals, areas) that is representative according to variation in the intrinsic or extrinsic physical attributes that underpin the process of interest.
 For example, the long-term persistence of a species may depend on its ability to adapt to new climatic conditions.
-By conserving individuals with the ability to persist in a range of conditions--physical attributes associated with adaptation--protected areas can foster evolutionary processes.
-Despite this, current reserve selection methods overwhelmingly focus on securing an adequate proportion of features' geographic ranges and little attention has been directed towards capturing a representative sample.
+By conserving individuals with traits that are advantageous to their local environment--physical attributes associated with adaptation--protected areas can foster the process of evolutionary adaptation.
+Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of individuals or habitat for each feature.
+Little attention has been directed towards capturing a representative sample of each feature.
 
@@
 The \texttt{raptr R} package provides a toolkit for making spatial conservation prioritizations that secure an adequate and representative sample of features.
-By targeting representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
+By targeting within-feature representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
 

### 2017-02-07 code/rmarkdown/text.Rmd
+ fix typos
address some of the reviewers comments add in benchmark analysis

@@
 If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation.
-Whereas, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning features based on political boundaries; @r483], then it might sample much less of the original variation.
+Alternatively, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning features based on political boundaries; @r483], then it might sample much less of the original variation.
 Conventional reserve selection problems cannot differentiate between these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce cost or fragmentation.
@@
 
-Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of each feature.
+Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of the variation in each feature.
 To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
@@
 species, ecosystems).
-Spatial attributes are defined as the variation across the feature' range that the prioritization is required to sample.
+Here, attributes are defined as the variation across the feature' range that the prioritization is required to sample.
 They can be intrinsic variation (eg.
@@
 
-The formulations used to express the reserve selection problem in the \texttt{raptr R} package are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
-Although the \texttt{raptr R} package utilizes two novel formulations, for brevity, we will define the simpler formulation--referred to as the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
-These formulations are based on unreliable and reliable facility location problems and have been named accordingly [@r16].
-The key difference between these two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating the proportion of an attribute space sampled in a solution.
+The \texttt{raptr R} package utilizes two novel formulations of reserve selection problem to generate prioritizations.
+These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
+Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the unreliable and reliable formulations.
+The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a solution samples a feature's attribute space.
+Whereas, the unreliable formulation assumes that features occupy the planning units in which they are found in with 100 % certainty when calculating how well a solution samples a feature's attribute space.
+For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
 All mathematical terms defined hereafter are described in Table S1.
@@
 The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution.
-Generally, demand points are represented by their closest selected planning units (unless particularly low space-based targets are used).
+Generally, demand points are represented by their closest selected planning units (unless extremely small space-based targets are used, such as 0.000001, and so the solution is no longer affected by the space-based target).
 
@@
 The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
-Presently, academics can obtain a license at no cost from the Gurobi website.
+Presently, academics can obtain a license at no cost from the \texttt{Gurobi} website.
 
@@
 We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches.
-By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience against environmental change [@r63].
-We selected Queensland, Australia as the study area because it contains a broad range of different habitats.
+By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63].
+We selected Queensland, Australia as the study area.
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
@@
 For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location.
-We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species [using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches; implemented in the \texttt{hypervolume R} package; @r231].
+We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231].
 We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations.
-These uniformly distributed points and associated density estimations used as demand point coordinates and weights (respectively).
+These uniformly distributed points and associated density estimations were used as demand point coordinates and weights (respectively).
 
@@
 We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
-Although this dataset contains multiple species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}) as a simple case study, because it has been found to exhibit significant genetic structure [@r462] as a simple case study.
+Although this dataset contains multiple species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}) as a simple case study, because it has been found to exhibit significant genetic structure [@r462].
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 
-We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
+We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented using the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
 These continuous variables described the main axes of genetic variation within the species.
@@
 7c and 7d).
-To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using the grid cells' centroids and inverse great circle distances [using the \texttt{ape R} package; @r464].
+To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled inverse great circle distances between the grid cells' centroids as weights [using the \texttt{ape R} package; @r464].
 
 The grid cells were used as planning units for generating prioritizations.
+As previously mentioned, data were not collected in every grid cell, and so the planning units are arranged in a checkerboard pattern.
 The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
@@
 To use such patchy data, planners could omit the planning units that are missing data, or estimate the missing data using spatially explicit models [eg.
-krigging or generalized dissimilarity models; @r467; @r466].
+krigging or generalized dissimilarity models; @r467; @r486].
 This approach has been successfully applied to a range of biological data sets [eg.

### 2017-02-07 code/R/analysis/00-initialization.R

@@ -12,7 +12,7 @@ if (length(args)>0) {
 # set checkpoint
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-checkpoint::checkpoint('2016-11-26', R.version='3.3.1')
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 
 # load CRAN packages
@@ -45,6 +45,7 @@ library(adehabitatHR)
 library(vegan)
 library(fields)
 library(ape)
+library(RandomFields)
 
 library(Hmisc)
 library(pander)

### 2017-02-07 code/R/analysis/01-load-data.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/00-initialization.rda')
 
 ### load simulation data

### 2017-02-07 code/R/analysis/03-case-study-1.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/01-load-data.rda')
 
 ## load parameters

### 2017-02-07 code/R/analysis/05-statistical-analysis.R

@@ -1,5 +1,5 @@
 ## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/04-case-study-2.rda')
 load('data/intermediate/03-case-study-1.rda')
 load('data/intermediate/02-simulations.rda')
@@ -27,7 +27,7 @@ cs1.nmds.MoransI <- llply(
 					curr.col <- curr.DF[[i]]
 					curr.pos <- which(!is.na(curr.DF[[i]]))
 					return(
-						Moran.I(curr.DF[[i]][curr.pos], cs2.pu.inv.dists[curr.pos, curr.pos])
+						Moran.I(curr.DF[[i]][curr.pos], cs2.pu.inv.dists[curr.pos, curr.pos], scale=TRUE)
 					)
 				}
 			)

### 2017-02-07 code/R/analysis/06-benchmark-analysis.R

@@ -1,0 +1,76 @@
+## load .rda
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
+session::restore.session('data/intermediate/05-statistical-analysis.rda')
+
+## load parameeters
+benchmark.params.LST <- parseTOML('code/parameters/benchmark.toml')
+
+## run benchmark analysis for unreliable formulation
+unreliable.benchmark.params.DF <- expand.grid(number.features=benchmark.params.LST[[MODE]][['unreliable.number.of.features']],
+  number.planning.units=benchmark.params.LST[[MODE]][['unreliable.number.of.planning.units']],
+  replicate=seq_len(benchmark.params.LST[[MODE]][['replicates']]))
+
+unreliable.benchmark.DF <- ldply(sample.int(nrow(unreliable.benchmark.params.DF)), function(i) {
+  # simulate data
+  source('code/R/functions/simulate.R')
+  curr.rd <- simulate.problem.data(
+    number.features=unreliable.benchmark.params.DF[i,1],
+    number.planning.units=unreliable.benchmark.params.DF[i,2],
+    amount.target=benchmark.params.LST[[MODE]]$amount.target,
+    space.target=benchmark.params.LST[[MODE]]$space.target,
+    probability.of.occupancy=benchmark.params.LST[[MODE]]$occupancy.probability
+  )
+  curr.go <- GurobiOpts(MIPGap=general.params.LST[[MODE]][['MIPGap']],
+    Threads=general.params.LST[[MODE]][['threads']])
+  
+  ldply(benchmark.params.LST[[MODE]]$blm, function(j) {
+    curr.ru <- RapUnsolved(RapUnreliableOpts(BLM=j), curr.rd)
+    curr.time <- system.time({curr.rs <- solve(curr.ru, curr.go)})
+    data.frame(
+      number.features=unreliable.benchmark.params.DF[i,1],
+      number.planning.units=unreliable.benchmark.params.DF[i,2],
+      blm=j,
+      replicate=unreliable.benchmark.params.DF[i,3],
+      time=curr.time[['elapsed']]
+    )
+  })
+})  
+  
+## run benchmark analysis for reliable formulation
+reliable.benchmark.params.DF <- expand.grid(number.features=benchmark.params.LST[[MODE]][['reliable.number.of.features']],
[…]

### 2017-02-07 code/R/analysis/06-exports.R

@@ -1,6 +1,0 @@
- ## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
-session::restore.session('data/intermediate/05-statistical-analysis.rda')
-
-## save workspace
-save.session('data/final/results.rda', compress='xz')

### 2017-02-07 code/R/analysis/07-exports.R

@@ -1,0 +1,6 @@
+ ## load .rda
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
+session::restore.session('data/intermediate/06-benchmark-analysis.rda')
+
+## save workspace
+save.session('data/final/results.rda', compress='xz')

### 2017-02-07 code/R/functions/simulate.R

@@ -1,0 +1,154 @@
+#' Simulate data
+#'
+#' Simulate spatially auto-correlated data using random fields.
+#'
+#' @param x \code{\link[raster]{RasterLayer-class}} object to use as
+#    a template.
+#'
+#' @param n \code{integer} number of species to simulate.
+#'
+#' @param model \code{\link[RandomFields]{RP}} model object
+#'   to use for simulating data.
+#'
+#' @param transform \code{function} to transform values output
+#'   from the random fields simulation.
+#'
+#' @param ... additional arguments passed to
+#'   \code{\link[RandomFields]{RFsimulate}}.
+#'
+#' @return \code{\link[raster]{RasterStack-class}} object with a
+#'   layer for each species.
+#' 
+#' @seealso \code{\link[RandomFields]{RFsimulate}},
+#'   \code{\link{simulate_cost}}, \code{\link{simulate_species}}.
+#'
+simulate_data <- function(x, n, model, transform=identity, ...) {
+  # assert valid arguments
+  if (!'RandomFields' %in% rownames(installed.packages()))
+    stop('the "RandomFields" package needs to be installed to simulate data')
+  assertthat::assert_that(
+    inherits(x, 'RasterLayer'),
+    assertthat::is.number(n),
+    is.finite(raster::cellStats(x, 'max')),
+    inherits(model, 'RMmodel'),
+    inherits(transform, 'function'))
+  # generate values for rasters
+  coords <- as(x, 'SpatialPoints')@coords
+  mtx <- RandomFields::RFsimulate(model=model, x=coords[,1], y=coords[,2],
+                                  n=n, spConform=FALSE, ...)
+  # convert to matrix if not a matrix
+  if (!inherits(mtx, 'matrix'))
+    mtx <- matrix(mtx, ncol=1)
+  # generate populate rasters with values
+  stk <- raster::stack(lapply(seq_len(ncol(mtx)), function(i) {
+    r <- x
+    r[raster::Which(!is.na(r))] <- transform(mtx[,i])
+    r
+  }))
+  # return raster stack with simulated distributions
+  return(stk)
+}
+
+#' Simulate species habitat suitabilities
+#'
+#' Generates a random set of species using random field models. By default,
+#' the output will contain values between zero and one.
+#' 
+#' @inheritParams simulate_data
+#'
+#' @return \code{\link[raster]{RasterStack-class}} object.
[…]

### 2017-02-07 code/rmarkdown/supporting-information.Rmd
+ update formatting
address some of the reviewers comments add in benchmark analysis
updated benchmark analysis

@@
 
-## Appendix S1
+# Appendix S1
 The reliable formulation explicitly considers the probability that planning units are inhabited by features.
@@
 
-## Figures 
+# Appendix S2
+
+We conducted a benchmark analysis to understand how much time would be required to solve different sized conservation planning problems.
+Additionally, we also aimed to understand how the time required to solve a problem varied depending on which formulation was used (unreliable or reliable), and if fragmentation constraints were used.
+
+## METHODS
+
+We generated simulated data sets conservation planning problems using varying number of features `r paste(benchmark.params.LST$number.of.features, collapse=',')` and planning units (paste(benchmark.params.LST$number.of.planning.units, collapse=',')).
+For a given number of planning units and features, the data was simulated using the following steps.
+The planning units were simulated as squares that were arranged in a grid or near-grid like fashion.
+The units' costs were set to one.
+The spatial distribution of features were generated using binary spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @rXXX].
+If features were associated with less than two occupied planning units, these features were set to occupy two randomly selected planning units.
+All occupied planning units were set to have a probability of occupancy of `r benchmark.params.LST[[MODE]]$occupancy.probability`.
+Similar to the simulated data sets used in the main text, a geographic attribute space was used for each feature.
+The features' demand points were set as the centroids of the planning units that they occupied, and were weighted by the units' probability of occupancy.
+All problems used amount-based targets of `r benchmark.params.LST[[MODE]]$amount.target*100` % and space-based targets of `r benchmark.params.LST[[MODE]]$space.target*100`
+
+We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifer, and using a boundary length modifier of `r max(benchmark.params.LST[[MODE]]$blm`.
+To assess variation in solve times, we generated and solved `r benchmark.params.LST[[MODE]]$replicates` random data sets per combination of number of features and planning units.
+All conservation planning problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) and a an Intel(R) Xeon(R) CPU E5-2630 (version 3; 2.40 GHz).
+To provide comparable estimates for the typical user, problems were solved using `r as.english(general.params.LST[[MODE]]$threads)` threads.
+
+## RESULTS
+
+```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap='insert caption here.'}
+ggplot(benchmark.results.DF, aes(x=number.planning.units, y=time, color=number.features)) +
+  geom_line() +
+  facet_wrap(~ settings)
+```
+
+# Figures 
 ```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study.
@@
 
-## Tables
+# Tables
 

### 2017-02-07 code/R/analysis/06-benchmark-analysis.R

@@ -5,17 +5,19 @@ session::restore.session('data/intermediate/05-statistical-analysis.rda')
 ## load parameeters
 benchmark.params.LST <- parseTOML('code/parameters/benchmark.toml')
 
-## run benchmark analysis for unreliable formulation
-unreliable.benchmark.params.DF <- expand.grid(number.features=benchmark.params.LST[[MODE]][['unreliable.number.of.features']],
-  number.planning.units=benchmark.params.LST[[MODE]][['unreliable.number.of.planning.units']],
+## set benchmark parameter combinations
+benchmark.params.DF1 <- expand.grid(number.features=benchmark.params.LST[[MODE]][['number.of.features']],
+  number.planning.units=benchmark.params.LST[[MODE]][['number.of.planning.units']],
   replicate=seq_len(benchmark.params.LST[[MODE]][['replicates']]))
 
-unreliable.benchmark.DF <- ldply(sample.int(nrow(unreliable.benchmark.params.DF)), function(i) {
+benchmark.params.DF2 <- expand.grid(formulation=c('unreliable', 'reliable'), blm=benchmark.params.LST[[MODE]]$blm )
+  
+## run benchmark analysis
+benchmark.results.DF <- ldply(sample.int(nrow(benchmark.params.DF1)), function(i) {
   # simulate data
-  source('code/R/functions/simulate.R')
   curr.rd <- simulate.problem.data(
-    number.features=unreliable.benchmark.params.DF[i,1],
-    number.planning.units=unreliable.benchmark.params.DF[i,2],
+    number.features=benchmark.params.DF1[i,1],
+    number.planning.units=benchmark.params.DF1[i,2],
     amount.target=benchmark.params.LST[[MODE]]$amount.target,
     space.target=benchmark.params.LST[[MODE]]$space.target,
     probability.of.occupancy=benchmark.params.LST[[MODE]]$occupancy.probability
@@ -23,54 +25,24 @@ unreliable.benchmark.DF <- ldply(sample.int(nrow(unreliable.benchmark.params.DF)
   curr.go <- GurobiOpts(MIPGap=general.params.LST[[MODE]][['MIPGap']],
     Threads=general.params.LST[[MODE]][['threads']])
   
-  ldply(benchmark.params.LST[[MODE]]$blm, function(j) {
-    curr.ru <- RapUnsolved(RapUnreliableOpts(BLM=j), curr.rd)
[…]

### 2017-02-07 code/R/functions/simulate.R

@@ -78,15 +78,21 @@ simulate_species <- function(x, n=1, model=RandomFields::RMgauss(),
 #' @return \code{RapData}.
 simulate.problem.data <- function(number.planning.units, number.features, probability.of.occupancy,
                                   amount.target=0.2, space.target=0.8) {
-  # simulate planning units
+  # simulate landscape
   curr.landscape <- raster::raster(ncol=ceiling(sqrt(number.planning.units)), nrow=ceiling(sqrt(number.planning.units))) %>%
     setValues(1) %>%  `extent<-`(c(0,1,0,1))
-  curr.pus <- rasterToPolygons(curr.landscape, n=4)
-  curr.pus <- curr.pus[seq_len(number.planning.units),]
   
   # simulate species
   curr.spp <- simulate_species(curr.landscape, n=number.features, 
                 model=RPbernoulli(RMgauss()), transform=identity)
+  
+  # change extents so that polygon boundarys are not too small
+  extent(curr.landscape) <- c(0, ceiling(sqrt(number.planning.units)), 0, ceiling(sqrt(number.planning.units)))
+  extent(curr.spp) <- extent(curr.landscape)
+  
+  # simulate pus
+  curr.pus <- rasterToPolygons(curr.landscape, n=4)
+  curr.pus <- curr.pus[seq_len(number.planning.units),]
 
   # find out which species are in which pus
   pu.pts <- gCentroid(curr.pus, byid=TRUE)

### 2017-02-09 code/rmarkdown/figures.Rmd
+ fix typos
update text

@@
 # load .rda file
-checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 try(session::restore.session('data/final/results.rda'))
@@
 	population=LETTERS[1:4],
-	temperature=c(22,29,21.6,34),
-	rainfall=c(6,12,5.5,6)
+	temperature=c(35,34,31,21.6),
+	rainfall=c(4.5,5,9.5,5.5)   
 )
+
 # demand point data
@@
 ## plot data
-ggplot(aes(x=temperature, y=rainfall), data=populations) +
+ggplot(aes(x=rainfall, y=temperature), data=populations) +
 	geom_point(data=demand.points, fill='gray70') +
@@
 	theme_classic() +
-	xlab(expression('Temperature ('*degree*'C)')) +
-	ylab('Rainfall (mm)') +
+	ylab(expression('Temperature ('*degree*'C)')) +
+	xlab('Rainfall (mm)') +
 	theme(axis.line.x=element_line(), axis.line.y=element_line())
@@
 The top row of panels (a, d, g, j) shows prioritizations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
-Each column of panels corresponds to a different set of parameters used to generate the prioritisation.
+Each column of panels corresponds to a different set of parameters used to generate the prioritization.
 The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}

### 2017-02-09 code/rmarkdown/text.Rmd
update text

@@
 # load .rda file
-checkpoint::checkpoint('2016-11-26', R.version='3.3.1', scanForPackages=FALSE)
+checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/final/results.rda')
@@
 conditions known to be outside the physiological tolerance of certain species).
-Thus a different set of demand points and weights are used for each attribute space and each feature.
+Thus a different set of demand points and weights may be appropriate for different attribute spaces and features.
 By siting demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritizations secure the feature in planning units with spatial attributes that are desirable or representative for that feature.
@@
 1).
-We can see that a solution that prioritizes both populations _A_ and _C_, effectively "doubles-up" on the same environmental characteristics and constitute considerable redundancy if both selected.
-Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$ (or $C$), $B$, and $D$.
-This example demonstrates how attribute spaces and demand points relate to biodiversity processes, and can be used to guide the reserve selection process.
+We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same environmental characteristics and constitute considerable redundancy if both selected.
+Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$ (or $B$), $C$, and $D$.
+However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, demand points could be sited in conditions that reflect this goal (eg.
+the top two rows of demand points in Fig.
+1).
+Given this new set of demand points, populations _A_, _B_, and _C_ would be prioritized.
+As we have demonstrated, demand points provide a flexible means to guide the reserve selection process.
 
@@
 Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$.
-Space-based targets ensure that a representative sample of each feature is secured.
+Space-based targets ensure that a representative or desirable sample of each feature is secured.
 Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
@@
 %
-\tau_{fs} &= \parbox{25em}{representation target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
+\tau_{fs} &= \parbox{25em}{space-based target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
 %
@@
 The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution.
-Generally, demand points are represented by their closest selected planning units (unless extremely small space-based targets are used, such as 0.000001, and so the solution is no longer affected by the space-based target).
+Generally, demand points are represented by their closest selected planning units (unless extremely small space-based targets are used, such as 0.000001, and so the solution effectively unconstrained by the space-based target).
 
@@
 First to represent solutions generated using conventional methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-Second to show how the addition of representation targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets.
+Second to show how the addition of space-based targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % space targets to ensure that the solution samples the geographic spread of the species (hereafter referred to as geographic spread targets).
 Third to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
-Fourth to illustrate the combined affect using geographic targets and penalising fragmentation, we generated solutions using using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
+Fourth to illustrate the combined affect using space-based targets and penalising fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic spread targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
 
@@
 
-The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
+The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic spread target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
 3b).
@@
 
-The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic distribution (Fig.
+The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic spread (Fig.
 3d).
-Although this solution also contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--the same amount as previously discussed solutions--it secures a representative proportion of the species geographic distribution in a configuration that is not highly fragmented.
+Although this solution also contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--the same amount as previously discussed solutions--it secures a representative proportion of the species geographic spread in a configuration that is not highly fragmented.
 Overall, these results show that under the simplest of conditions, using just amount targets results in an under-specified reserve selection problem that is unlikely to return an effective solution.
@@
 
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig.
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig.
 3f).
 This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution.
-By using amount and geographic targets in the reserve selection problem, we obtained a solution that concentrates conservation effort in the range core and also the range margin.
+By using amount and geographic spread targets in the reserve selection problem, we obtained a solution that concentrates conservation effort in the range core and also the range margin.
 However, similar to the corresponding solution for the uniform species (Fig.
@@
 
-By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig.
+By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig.
 3h).
@@
 3k).
-However, the addition of geographic space targets resolved these issues.
+However, the addition of geographic spread targets resolved these issues.
 
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic targets included planning units from both populations (Fig.
+The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets included planning units from both populations (Fig.
 3j).
@@
 
-The degree to which a prioritization truly secures a representative sample of a feature depends on the attribute space(s) and the distribution of demand points chosen by the conservation planner.
-The optimal solution will not effectively represent the decision maker's objectives if an inapropriate set of spatial variables or demand points are used to represent the attribute space.
-Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space.
-To assist the generation of appropriate demand points, the \texttt{R} package provides several routines that use the distribution of the features in spatial data sets to make demand points (see documentation for the `make.DemandPoints` function).
+The degree to which a prioritization truly secures a representative sample of a feature depends on (1) the attribute space(s), (2) the distribution of demand points chosen by the conservation planner, and (3) the space target used.
+The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space.
+Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this).
+As with conventional amount-based targets, planners will have to ensure that space targets are high enough to fulfill conservation objectives [eg.
+when setting targets to conserve intra-specific genetic variation; @r487].
+We encourage conservation planners to consult with experts to ensure that appropriate space targets are used.
 
@@
 @r468].
+We caution that inaccuracy and uncertainty in data can negatively impact prioritizations [@r488], and recommend that planners ensure that data are reliable or accommodate uncertainty into the planning process (eg.
+by using the reliable formulation).
 

### 2017-02-09 code/rmarkdown/supporting-information.Rmd
update si

@@
 
-We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifer, and using a boundary length modifier of `r max(benchmark.params.LST[[MODE]]$blm`.
+We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifer, and using a boundary length modifier of `r max(benchmark.params.LST[[MODE]]$blm)`.
 To assess variation in solve times, we generated and solved `r benchmark.params.LST[[MODE]]$replicates` random data sets per combination of number of features and planning units.

### 2017-02-09 code/rmarkdown/supporting-information.Rmd
update si and code

@@
 ```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap='insert caption here.'}
-ggplot(benchmark.results.DF, aes(x=number.planning.units, y=time, color=number.features)) +
-  geom_line() +
-  facet_wrap(~ settings)
+plot(1)
 ```

### 2017-02-09 code/R/analysis/05-benchmark-analysis.R

@@ -72,5 +72,5 @@ benchmark.DF <- unreliable.benchmark.DF %>%
   rbind(reliable.benchmark.DF %>% mutate(formulation='reliable'))
   
 ## save workspace
-save.session('data/intermediate/06-benchmark-analysis.rda', compress='xz')
+save.session('data/intermediate/05-benchmark-analysis.rda', compress='xz')
 

### 2017-02-09 code/R/analysis/06-statistical-analysis.R

@@ -66,6 +66,4 @@ amount.represented.species.names <- tolower(amount.represented.species.names)
 amount.not.represented.species.names <- tolower(amount.not.represented.species.names)
 
 ## save workspace
-save.session('data/intermediate/05-statistical-analysis.rda', compress='xz')
- 
-
+save.session('data/intermediate/06-statistical-analysis.rda', compress='xz')

### 2017-02-09 code/R/analysis/07-exports.R

@@ -1,6 +1,6 @@
  ## load .rda
 checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/06-benchmark-analysis.rda')
+session::restore.session('data/intermediate/06-statistical-analysis.rda')
 
 ## save workspace
 save.session('data/final/results.rda', compress='xz')

### 2017-02-09 code/rmarkdown/figures.Rmd
address more comments

@@
 
+# remove small islands from australia basemap
+aus.sub.PLY <- disaggregate(aus.PLY)
+aus.sub.PLY <- SpatialPolygonsDataFrame(aus.sub.PLY, data=data.frame(size=gArea(aus.sub.PLY, byid=TRUE)))
+aus.sub.PLY <- aus.sub.PLY[aus.sub.PLY$size >= quantile(aus.sub.PLY$size, 0.9, names=FALSE),]
+
 # fortify polygons
 sim.pus.FPLY <- fortify(sim.pus)
-aus.FPLY <- fortify(aus.PLY)
+aus.FPLY <- fortify(aus.sub.PLY)
 cs1.pus.FPLY <- fortify(cs1.pus)
@@
 	temperature=c(35,34,31,21.6),
-	rainfall=c(4.5,5,9.5,5.5)   
+	rainfall=c(4.5,5.3,9.5,5.5)   
 )
@@
 
-```{r, echo=FALSE, fig.height=1.6, fig.width=column.width, fig.cap='Distribution of environmental variation across Queensland, Australia.
+```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap='Distribution of environmental variation across Queensland, Australia.
 Squares denote planning units.
@@
 	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[1]), aes(x=x, y=y, fill=Score)) +
-	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
+	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black', size=0.05) +
 	theme_classic() +
@@
 	) +
-	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
   scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[1])$Score, n=3, min.n=2)) +
@@
 	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[2]), aes(x=x, y=y, fill=Score)) +
-	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black') +
+	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black', size=0.05) +
 	theme_classic() +
@@
 		) +
-	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
 	scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[2])$Score, n=2, min.n=2)) +
@@
 
-```{r, echo=FALSE, fig.height=2.35, fig.width=column.width, fig.cap='Distribution of the species used in the first case study.
+```{r, echo=FALSE, fig.height=3.8, fig.width=column.width, fig.cap='Distribution of the species used in the first case study.
 Each panel shows a different species.
@@
 )
+
 cs1.pus.occ.FPLY$species <- factor(as.character(cs1.pus.occ.FPLY$species), levels=cs1.params.LST[[MODE]]$common.names)
@@
 cs1.pus.occ.letters.DF <- data.frame(
-	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.06)),
-	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.90)),
+	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.08)),
+	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.95)),
 	species=unique(cs1.pus.occ.FPLY$species),
@@
 ggplot() +
-	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70', size=0.05) +
 	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black', size=0.05) +
@@
 		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=8),
+		strip.text = element_text(color='white', size=7),
 		axis.title=element_blank(), legend.position='none'
@@
 	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
-	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
 	facet_wrap(~ species, ncol=2)
@@
 
-```{r, echo=FALSE, fig.height=1.175, fig.width=column.width, fig.cap=paste0('Prioritizations for the first case study.
+```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap=paste0('Prioritizations for the first case study.
 Squares represent planning units.
@@
 cs1.pus.sel.letters.DF <- data.frame(
-	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.06)),
-	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.90)),
+	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.08)),
+	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.95)),
 	Prioritisation=names(cs1.prioritisations),
@@
 ggplot() +
-	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black', size=0.1) +
+	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70', size=0.05) +
+	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black', size=0.05) +
 	geom_label(data=cs1.pus.sel.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
@@
 	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
-	coord_cartesian(xlim=cs1.pus@bbox[1,], ylim=cs1.pus@bbox[2,]) +
+	coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
 	facet_wrap(~ Prioritisation, ncol=2)
@@
 
-```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Data used for the second case study.
+```{r, echo=FALSE, fig.height=2.2, fig.width=column.width, fig.cap=paste0('Data used for the second case study.
 Squares denote planning units.
@@
 cs2.spp.grid.letters.DF <- data.frame(
-	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.06)),
-	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.90)),
+	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.07)),
+	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.875)),
 	letter=('(a)')
@@
 	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
-	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.90)),
+	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.89)),
 	letter=c('(b)','(c)','(d)')
@@
 		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-		legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
+		legend.position=c(0.25, 0.58), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'), legend.text=element_text(size=6.0),
 		legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
@@
     axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-    legend.position=c(0.25, 0.55), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'),
+    legend.position=c(0.25, 0.58), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'), legend.text=element_text(size=6.0),
     legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
@@
 	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
-	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.90)),
+	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.9)),
 	letter=c('(a)','(b)')

### 2017-02-09 code/rmarkdown/supporting-information.Rmd
address more comments

@@
 ```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap='insert caption here.'}
-plot(1)
+# prepare data
+plot.DF <- benchmark.results.DF %>%
+  mutate(blm = paste0('BLM = ',blm))
+
+# make plot
+ggplot(plot.DF, aes(x=number.planning.units, y=time, color=number.features)) +
+  geom_boxplot() +
+  facet_grid(formulation ~ blm) +
+  theme_classic() +
+  xlab('Number of planning units) + ylab('Solve time (s)') +
+  theme(axis.line.x=element_line(), axis.line.y=element_line())
 ```

### 2017-02-09 code/rmarkdown/text.Rmd
address more comments

@@
 For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
-However, using this method is challenging because biodiversity data are often continuous and multi-dimensional.
+However, using this method is challenging because biodiversity often cannot be divided into operational groups for conservation planning without loss of information [@r489; @r490; @r218].
 
@@
 Conventional reserve selection problems cannot differentiate between these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce cost or fragmentation.
-This limitation has been known of quite some time.
+This limitation has been known of quite some time [@r218].
 In fact, it dates back to some of the earliest reserve selection methods [eg.

### 2017-02-09 code/rmarkdown/supporting-information.Rmd
benchmark analysis results figure

--- 2017-02-09 code/rmarkdown/supporting-information.Rmd
@@
 
-```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap='insert caption here.'}
-plot(1)
+```{r, echo=FALSE, message=FALSE, fig.height=5.0, fig.width=4.5, fig.cap='Benchmark analysis.
+Boxplots show the amount of time required to solve a problem under a given set of parameters.
+Note that times are shown on a log_{10} scale.'}
+# prepare data
+plot.DF <- benchmark.DF %>%
+  mutate(blm = paste0('BLM = ',blm)) %>% 
+  mutate(number.features=factor(number.features)) %>%
+  mutate(formulation = paste0(formulation, ' formulation')) %>%
+  mutate(number.planning.units=factor(number.planning.units))
+# make plot
+ggplot(plot.DF, aes(x=number.planning.units, y=time, color=number.features)) +
+  geom_boxplot() +
+  facet_grid(blm ~ formulation ) +
+  scale_y_continuous(trans='log10') +
+  xlab('Number of planning units') + ylab('Solve time (s)') +
+  labs(color='Number of features') +
+  theme(axis.line.x=element_line(), axis.line.y=element_line(),
+    strip.background = element_rect(fill='grey20'),
+    strip.text = element_text(color='white', size=12),
+    legend.position='bottom')
 ```

### 2017-02-10 code/rmarkdown/supporting-information.Rmd
assorted changes to ms

@@
 
-Moderately-sized conservation planning problems often include several thousand planning units.
-It is currently not be feasible to solve this problem when considering all possible failure scenarios.
+Conservation planning problems often use more than several hundred planning units.
+It is currently not be feasible to solve this problem when considering all possible failure scenarios within a reasonable amount of time (see Appendix S2).
 As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$.
 For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level.
-Cui et al.
-[-@r16] found that $R=5$ yields similar solutions to $R >> 5$.
+Cui _et al._ [-@r16] found that $R=5$ yields similar solutions to $R >> 5$.
 However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritizations in a feasible amount of time.
@@
 
-We conducted a benchmark analysis to understand how much time would be required to solve different sized conservation planning problems.
-Additionally, we also aimed to understand how the time required to solve a problem varied depending on which formulation was used (unreliable or reliable), and if fragmentation constraints were used.
+We conducted a benchmark analysis to understand how much time would be required to build and solve different sized conservation planning problems.
+Additionally, we also aimed to understand how the time required to execute a problem varied depending on which formulation was used (unreliable or reliable), and if fragmentation constraints were used.
 
@@
 
-We generated simulated data sets conservation planning problems using varying number of features `r paste(benchmark.params.LST$number.of.features, collapse=',')` and planning units (paste(benchmark.params.LST$number.of.planning.units, collapse=',')).
+We generated simulated conservation planning data sets using varying number of features `r paste(benchmark.params.LST$number.of.features, collapse=',')` and planning units (paste(benchmark.params.LST$number.of.planning.units, collapse=',')).
 For a given number of planning units and features, the data was simulated using the following steps.
-The planning units were simulated as squares that were arranged in a grid or near-grid like fashion.
-The units' costs were set to one.
-The spatial distribution of features were generated using binary spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @rXXX].
+First, the planning units were simulated as squares that were arranged in a grid.
+All planning unit costs were set to one.
+The spatial distribution of features were generated using binary, spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @rXXX].
 If features were associated with less than two occupied planning units, these features were set to occupy two randomly selected planning units.
@@
 We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifer, and using a boundary length modifier of `r max(benchmark.params.LST[[MODE]]$blm)`.
-To assess variation in solve times, we generated and solved `r benchmark.params.LST[[MODE]]$replicates` random data sets per combination of number of features and planning units.
-All conservation planning problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) and a an Intel(R) Xeon(R) CPU E5-2630 (version 3; 2.40 GHz).
-To provide comparable estimates for the typical user, problems were solved using `r as.english(general.params.LST[[MODE]]$threads)` threads.
+To assess variation in execution times, we generated and solved `r benchmark.params.LST[[MODE]]$replicates` random data sets per combination of number of features and planning units.
+All conservation planning problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) on a system with 64 GB RAM and an Intel(R) Xeon(R) CPU (E5-2630 v3; 2.40 GHz).
+All problems were solved using a single thread.
 
-## RESULTS
+## RESULTS AND DISCUSSION
+
+The benchmark analysis revealed that moderately-sized conservation problems can be solved in a realtively short period of time (Figure 1).
+As predicted, problems were solved much quicker when using the unreliable formulation than when using the reliable formulation.
+Additionally, the solve times varied substiantially for different data sets with the same proporties (ie.
+number of features and planning units).
+This result suggests that it may difficult to estimate exactly how long a given problem will take to solve.
 
@@
 ```
+
+We were not able to explore how long large-sized problems would take to solve because our computer system did not have enough memory.
+One method to reduce the problem size is to set a maximum distance between demand points and planning units such that demand points cannot be represented by planning units that that are too far away.
+This method could be applied using functions provided in the \texttt{raptr R} package.
+Future research could involve implementing more advanced methods to solve the optimization problem [eg.
+as described in @r16; @r15].
 

### 2017-02-10 code/rmarkdown/text.Rmd
assorted changes to ms

@@
 By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection.
-All analyses were performed in \texttt{R} [version `r getRversion()`; @r192].
+We have also conducted a benchmark analysis to estimate how long it would take to solve various sized problem (Appendix S2).
+All analyses were performed in \texttt{R} [version `r getRversion()`; @r192] and all optimization problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.
 

### 2017-02-11 code/rmarkdown/figures.Rmd
address reviewers comments reduce some of the length

@@
 
-```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Example of an attribute space.
+```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Attribute space example.
 This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm).
@@
 Points represent demand points.
-In this space, populations closer to each other have similar environmental conditions."}
+In this space, populations close to each other inhabit similar environmental conditions."}
 ## generate data
@@
 Squares denote planning units.
-The color of each planning unit indicates the probability that species inhabit it.'}
+Colors indicate probability of occupancy.'}
 # prepare data
@@
 Squares denote planning units.
-Dark green units show planning units selected for protection.
-The top row of panels (a, d, g, j) shows prioritizations generated for the uniformly distributed species, middle row of panels (b, e, h, k) for the normally distributed species, and the bottom row of panels (c, f, i, l) for the bimodally distributed species.
-Each column of panels corresponds to a different set of parameters used to generate the prioritization.
-The left column of panels (a--c) shows prioritizations generated using only ',(sim.params.LST[[MODE]]$amount.target*100),' % amount targets, the middle-left column of panels (d--f) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets, the middle-right column of panels (g--i) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount based targets and a boundary length modifier (BLM) of ',sim.params.LST[[MODE]]$blm,', and the right column of panels (j--l) for prioritizations generated using ',(sim.params.LST[[MODE]]$amount.target*100),' % amount and ',(sim.params.LST[[MODE]]$space.target*100),' % space targets and also a boundary length modifier of ',(sim.params.LST[[MODE]]$blm),'.')}
+Planning units shown in dark green units were selected for protection.
+Each row of panels show prioritizations generated for a different species.
+Each column of panels corresponds to a different set of parameters used to generate the prioritization.')}
 # prepare data
@@
 
-```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap='Distribution of environmental variation across Queensland, Australia.
+```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap='Distribution of climatic variation across Queensland, Australia.
 Squares denote planning units.
@@
 ```{r, echo=FALSE, fig.height=3.8, fig.width=column.width, fig.cap='Distribution of the species used in the first case study.
-Each panel shows a different species.
-Squares denote planning units.
-Planning units occupied by a species are shown in bright blue.'}
+Each panel shows the geographic range for a different species.
+Squares denote planning units.'}
 # prepare data
@@
 ```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap=paste0('Prioritizations for the first case study.
-Squares represent planning units.
-Dark green planning units were selected for protection.
+See Fig.
+3 caption for conventions.
 Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
@@
 Panel (b) shows the opportunity cost of each planning unit (estimated as the total human population density).
-Panels (c--d) show the spatial distribution of the ordinated genetic data.
+Panels (c--d) show the spatial distribution of the ordinations describing genetic variation.
 These values describe the typical genetic characteristics of individuals in each planning unit.
@@
 Panels (a--b) show prioritizations generated using different parameters.
-Squares denote planning units shown in geographic space.
-Dark green planning units were selected for protection.
+See Fig.
+3 caption for conventions.
 Panel (a) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets.

### 2017-02-11 code/rmarkdown/supporting-information.Rmd
address reviewers comments reduce some of the length

@@
 We were not able to explore how long large-sized problems would take to solve because our computer system did not have enough memory.
-One method to reduce the problem size is to set a maximum distance between demand points and planning units such that demand points cannot be represented by planning units that that are too far away.
-This method could be applied using functions provided in the \texttt{raptr R} package.
-Future research could involve implementing more advanced methods to solve the optimization problem [eg.
+One method to reduce the problem size would be to simply use less demand points.
+Another method to reduce problem size would be to set a maximum distance between demand points and planning units such that demand points cannot be represented by planning units that that are too far away.
+Both of these method could be applied using functions provided in the \texttt{raptr R} package.
+Future research is needed to explore these methods and understand how the impact solution quality.
+Additionally, future research could involve implementing more advanced methods to solve the optimization problem [eg.
 as described in @r16; @r15].

### 2017-02-11 code/rmarkdown/text.Rmd
address reviewers comments reduce some of the length

@@
 However, using this method is challenging because biodiversity often cannot be divided into operational groups for conservation planning without loss of information [@r489; @r490; @r218].
+This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [@r489].
 
-Simply partitioning features based on continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable.
-For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution captures 10 % of the geographic range of each sub-feature, how much of the original continuous variation does the solution represent?
-The answer is not immediately clear.
-If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation.
-Alternatively, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning features based on political boundaries; @r483], then it might sample much less of the original variation.
-Conventional reserve selection problems cannot differentiate between these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce cost or fragmentation.
-This limitation has been known of quite some time [@r218].
-In fact, it dates back to some of the earliest reserve selection methods [eg.
-@r295].
-
-During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of environmental variation across a study area.
+To overcome this limitation, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of the environmental variation found across a study area.
 Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
 Recent work has built on this reserve selection method by solving problems with more advanced optimization algorithms [@r482].
-However, \texttt{DIVERSITY} has limited utility for species-based conservation planning, in part, because it cannot accommodate multiple biological features--unlike more commonly used tools such as \texttt{Marxan} and \texttt{Zonation}.
-This fundamental limitation means that \texttt{DIVERSITY} is not useful for most conservation planning scenarios.
+However, existing \texttt{DIVERSITY}-based methods have limited utility for species-based conservation planning, in part, because they cannot accommodate multiple biological features and do not provide options to specify specific targets pertaining to the amount of habitat required to conserve species (unlike \texttt{Marxan} and \texttt{Zonation}).
+As a consequence, such methods are not useful for most conservation planning exercises [@r255].
 
@@
 species, ecosystems).
-Here, attributes are defined as the variation across the feature' range that the prioritization is required to sample.
-They can be intrinsic variation (eg.
-genetic or trait) variation that is extrinsic to the feature (eg.
-environmental conditions).
-There should, however, be a reasonable underlying hypothesis that relates this variation to the biodiversity processes that the decision maker aims to preserve.
+Here, attributes are defined as the variation across the features' ranges that the prioritization is required to sample.
+These attributes can be intrinsic (eg.
+genetic or phenotypic) or extrinsic (eg.
+environmental conditions) to the feature.
+There should, however, be a reasonable underlying hypothesis that relates the attribute to the biodiversity processes that needs to be conserved.
 
@@
 To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm).
-Any given combination of temperature and precipitation may be conceived as a point in this environmental space.
+Any given combination of temperature and precipitation may be conceived as a point in this climatic space.
 Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given attribute space.
-By associating the planning units with climatic data, and calculating a representative statistic for each planning unit (eg.
-mean), they can be mapped from geographic space to this environmental attribute space.
-In addition to planning units, attribute spaces also contain demand points.
+By associating the planning units with climatic data, and calculating a descriptive statistic for each planning unit (eg.
+mean), they can be mapped from geographic space to this climatic attribute space.
+In addition to points representing planning units, attribute spaces also contain demand points.
 
 Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets).
-The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space (see below for mathematical formulations).
-The shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the variation in the spatial attribute.
-To calculate the proportion of variation secured in a given solution, the distances are scaled by the distance between the demand points to their centroid (equivalent to calculating R$^2$ values for K-means clustering).
+For instance, by siting demand points throughout an attribute space, planners can obtain solutions that conserve a representative sample of an attribute space.
+Alternatively, by siting demand points in specific regions of an attribute space, planners can obtain solutions that target specific samples of an attribute space.
+For a given set of demand points, the shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the desired variation in the attribute space.
 In any attribute space there may exist points that are impossible (eg.
@@
 conditions known to be outside the physiological tolerance of certain species).
-Thus a different set of demand points and weights may be appropriate for different attribute spaces and features.
-By siting demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritizations secure the feature in planning units with spatial attributes that are desirable or representative for that feature.
+Thus a different set of demand points may be appropriate for different attribute spaces and features.
 
 To illustrate these concepts, consider the following example: we wish to develop a prioritization for a single species that has four populations.
-However, we can only afford to preserve three of the four populations.
-To ensure that we preserve a representative sample, we want to preserve populations inhabiting different environmental conditions.
-We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population.
-We used this data to construct a two-dimensional environmental attribute space.
+Since we can only afford to preserve three of the four populations, our objective is to conserve the most representative sample possible.
+To achieve this goal, we obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population.
+We used this data to construct a two-dimensional climatic attribute space.
 Next, we generated demand points as equi-distant points inside this space.
-By comparing the distributions of demand points and populations in the attribute space, we can identify a suitable prioritization (Fig.
+By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify the optimal solution (Fig.
 1).
-We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same environmental characteristics and constitute considerable redundancy if both selected.
-Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$ (or $B$), $C$, and $D$.
-However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, demand points could be sited in conditions that reflect this goal (eg.
+We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same climatic characteristics and constitutes considerable redundancy if both selected.
+Instead, a more representative sample of the intra-specific variation could be captured by securing populations _A_ (or _B_), _C_, and _D_.
+However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, we could site demand points in environmental conditions with temperatures over 30 $^{\circ}$C (ie.
 the top two rows of demand points in Fig.
@@
 Given this new set of demand points, populations _A_, _B_, and _C_ would be prioritized.
-As we have demonstrated, demand points provide a flexible means to guide the reserve selection process.
+Since demand points can be sited and weighted in any configuration, they provide a flexible means to guide the reserve selection process.
 
@@
 These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
-Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the unreliable and reliable formulations.
-The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a solution samples a feature's attribute space.
-Whereas, the unreliable formulation assumes that features occupy the planning units in which they are found in with 100 % certainty when calculating how well a solution samples a feature's attribute space.
+Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations.
+The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a given solution samples a feature's attribute space.
+On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with 100 % certainty when performing these calculations.
 For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
@@
 
-Targets are used to ensure that prioritizations adequately preserve each feature.
-Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient.
-Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$.
-Space-based targets ensure that a representative or desirable sample of each feature is secured.
+Targets are used to ensure that prioritizations sufficiently conserve each feature.
+Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}).
+Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$.
+Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved.
+Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid.
+This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [@r491; page 106].
 Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
-For convenience, all targets are expressed as proportions in the \texttt{R} package.
-
 The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.
@@
 %
-\tau_{fs} &= \parbox{25em}{space-based target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
+\tau_{fs} &= \parbox{25em}{space target for feature $f$ in attribute space $s$} \tag*{eqn 1b}\\
 %
@@
 
-Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is assigned to a selected planning unit $J$ where $X_j = 1$.
-The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution.
-Generally, demand points are represented by their closest selected planning units (unless extremely small space-based targets are used, such as 0.000001, and so the solution effectively unconstrained by the space-based target).
+Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie.
+a $j \in J$ where $X_j = 1$).
+The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$).
+Generally--unless small space targets are used, eg.
+0.000001, so that the solution is effectively unconstrained by the space target--demand points are represented by their closest selected planning units.
+In poorer quality solutions, demand points will be represented by planning units that have larger distances between them.
+As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.
 
@@
 Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space.
-For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$).
-This total weighted distance is then scaled by the inherent variation in the demand points ($\lambda_{fsi} \delta_{fsi}$).
-The resulting fraction yields a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
-The constraints ensure that this proportion must be equal to or greater than the space-based target.
+For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}$).
+This total weighted distance is then scaled by the inherent variation in the demand points ($\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}$).
+As previously mentioned, the resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
+The constraints (eqn 3c) ensure that the proportion of variation in the demand points secured in the solution must be equal to or greater than the space-based target.
 Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point.
@@
 [-@r426].
-The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
+The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for reliable formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
 Presently, academics can obtain a license at no cost from the \texttt{Gurobi} website.
@@
 To understand and showcase the behavior of the unreliable problem, we conducted a simulation study and two case studies.
-These studies involved generating solutions using only amount targets to represent prioritizations generated using conventional methods (eg.
-\texttt{Marxan}), and solutions using amount and space targets using the unreliable formulation.
+These studies involved generating solutions using only amount targets--equivalent to solutions generated using conventional methods (eg.
+\texttt{Marxan})--and solutions using amount and space targets using the unreliable formulation.
 By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection.
-We have also conducted a benchmark analysis to estimate how long it would take to solve various sized problem (Appendix S2).
+We have also conducted a benchmark analysis to estimate how long it would take to solve various sized problems (Appendix S2).
 All analyses were performed in \texttt{R} [version `r getRversion()`; @r192] and all optimization problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.
@@
 2).
-We then simulated three species across this study area.
+We then simulated three species across the study area.
 The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig.
@@
 It has a core range area, and is less likely to be found in areas that are distant from the core area.
-This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 4c).
-The third species represents a species with two distinct populations (Fig.
+This species' distribution was simulated using the density function of a single multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
+The third species was simulated to represent a species with two distinct populations (Fig.
 2c).
-This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 4d).
-The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 4a--4d.
+It was simulated using the maximum of two density of two multivariate normal distributions (eqn 4c).
+The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 4a--4c.
 We used a geographic attribute space to showcase the behavior of the problem.
@@
 \\
-P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 4a} \\
+P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
 %
-f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 4b} \\
-%
-P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4c} \\
+P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
 %
 P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
-f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
-\end{cases} \tag*{eqn 4d} \\
+\matchcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
+\end{cases} \tag*{eqn 4c} \\
 %
@@
 
-The uniform species was simulated to occur with a constant probability of occupancy across the study area (Fig.
+The uniform species occurred in all planning units with equality probability (Fig.
 2a).
@@
 In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points.
+Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points (equivalent to the $R^2$ of a regression model that explains less variation in the response variable than the mean of the response variable).
 
@@
 Fig 3a).
-Although the use of amount and space based targets has addressed adequacy and representativeness objectives (respectively), they have resulted in a highly fragmented solution.
+Although the use of amount and space targets has yielded a solution that conserves a sample of the species that contains an adequate amount of habitat and a representative sample of the species' geographic spread, they have yielded a solution that is also highly fragmented solution.
 
@@
 Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy.
-Whilst this solution satisfies the adequacy objective for a prioritization in a cost-effective manner--it fails to fulfill the representative objective for a prioritization (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).
+Whilst this solution conserves a sufficient amount of habitat in a cost-effective manner--it fails to secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).
 
@@
 However, similar to the corresponding solution for the uniform species (Fig.
-3b), this solution is highly fragmented.
+3b), this solution is also highly fragmented.
 
@@
 when setting targets to conserve intra-specific genetic variation; @r487].
-We encourage conservation planners to consult with experts to ensure that appropriate space targets are used.
+We encourage conservation planners to consult with experts to identify suitable targets.
 
@@
 # Data accessibility
-The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
+The \texttt{raptr R} package can be downloaded from The Comprehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
 All data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.

### 2017-02-13 code/rmarkdown/article.Rmd
update text
address more of the reviewers' concerns

@@
 
-Abstract word count: XXXX / 350  
+Abstract word count: 347 / 350  
 Total word count: XXXX / 7000  
@@
 Short running title: 
-Short abstract: Capture biodiversity processes and patterns in prioritizations using raptr https://github.com/jeffreyhanson/raptr  
-Keywords: conservation, biodiversity, MILP, reserve-selection, optimization, protected area
+Short abstract: Prioritize biodiversity processes and patterns using raptr https://github.com/jeffreyhanson/raptr  
+Keywords: conservation, biodiversity, mixed integer linear programming, optimization, protected areas
 
@@
 species, ecosystem) is to preserve a sample of the feature (eg.
-individuals, areas) that is representative according to variation in the intrinsic or extrinsic physical attributes that underpin the process of interest.
-For example, the long-term persistence of a species may depend on its ability to adapt to new climatic conditions.
-By conserving individuals with traits that are advantageous to their local environment--physical attributes associated with adaptation--protected areas can foster the process of evolutionary adaptation.
-Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of individuals or habitat for each feature.
+individuals, areas) that is representative in terms of the intrinsic or extrinsic physical attributes that underpin the process of interest.
+For example, by conserving a sample of populations with local adaptations that is representative in terms of the full variation found in the species--physical attributes associated with adaptation--protected areas can foster adaptative processes by ensuring these adaptations are not lost.
+Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature.
 Little attention has been directed towards capturing a representative sample of each feature.
@@
 2.
-To address this issue, we developed the \texttt{raptr R} package to guide reserve selection using targets for representing spatially explicit variables that underpin biodiversity processes as well as more typical area protection targets.
-Users set "amount targets"--similar to conventional methods--to secure a sufficient proportion of the features' geographic distributions.
+To address this issue, we developed the \texttt{raptr R} package to help guide reserve selection.
+Users set "amount targets"--similar to conventional methods--to ensure that solutions secure a sufficient proportion of area or habitat for each feature.
 Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
-climatic or genetic variation).
+environmental or genetic variation).
 We demonstrate the functionality of this package using simulations and two case studies.
-We generate solutions using just amount targets--representing conventional methods--and compare them with solutions generated using amount and space targets.
+In these studies, we generated solutions using amount targets--representing conventional methods--and compared them with solutions generated using amount and space targets.
 
 3.
-We show that different solutions emerge when explicitly considering within feature representation in reserve selection.
-Our simulations suggest that targeting representativeness is particularly important for biodiversity features that have multimodal distributions in an attribute space.
-The case studies show that setting space targets can result in solutions that conserve a much more representative sample of features--and this can be achieved with only a slight increase in reserve size.
+Our results demonstrate that different solutions emerge when targeting a representative sample of each feature.
+We show that using these targets is important for features that have multimodal distributions in the process-related attributes (eg.
+species with multimodal niches).
+We also found that solutions could conserve a far more representative sample with only a slight increase in reserve size.
 
 4.
-The \texttt{raptr R} package provides a toolkit for making spatial conservation prioritizations that secure an adequate and representative sample of features.
-By targeting within-feature representativeness, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
+The \texttt{raptr R} package provides a toolkit for making prioritizations that secure an adequate and representative sample of each feature.
+By using solutions that secure a representative sample of each feature, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
 

### 2017-02-13 code/rmarkdown/text.Rmd
address more of the reviewers' concerns

@@
 ## Optimization
-The unreliable formulation is non-linear.
-However, the non-linear components can be linearized using existing techniques.
-The expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al.
-[-@r426].
-The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for reliable formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
-Presently, academics can obtain a license at no cost from the \texttt{Gurobi} website.
+Although the reserve selection problems presented here are non-linear (see Appendix S1 for reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [@r16].
+The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
+Presently, academics can obtain a license at no cost.
 
@@
 
-To understand and showcase the behavior of the unreliable problem, we conducted a simulation study and two case studies.
-These studies involved generating solutions using only amount targets--equivalent to solutions generated using conventional methods (eg.
-\texttt{Marxan})--and solutions using amount and space targets using the unreliable formulation.
-By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection.
-We have also conducted a benchmark analysis to estimate how long it would take to solve various sized problems (Appendix S2).
-All analyses were performed in \texttt{R} [version `r getRversion()`; @r192] and all optimization problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.
+To showcase the behavior of the unreliable formulation, we conducted a simulation study and two case studies.
+To understand how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2).
+We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.
 
@@
 ### Methods
-We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig.
+We simulated a hypothetical study area with square planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ grid (Fig.
 2).
-We then simulated three species across the study area.
-The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig.
-2a).
-This species' distribution was based on a uniform distribution (eqn 4a).
-The second species was simulated to represent a species with an idealized distribution (Fig.
+We then simulated three species to inhabit this study area.
+Firstly, we simulated a hyper-generalist species (hereafter referred to as the "uniformly distributed species").
+It all planning units with equal probability (Fig.
+2a; eqn 4a).
+Secondly, we simulated a species with simple habitat requirements (hereafter referred to as the "normally distributed species"; Fig.
 2b).
-It has a core range area, and is less likely to be found in areas that are distant from the core area.
-This species' distribution was simulated using the density function of a single multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
-The third species was simulated to represent a species with two distinct populations (Fig.
+This species was most likely to be found in planning units in the centre of the study area, and was simulated using density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
+Thirdly, we simulated a species two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig.
 2c).
 It was simulated using the maximum of two density of two multivariate normal distributions (eqn 4c).
-The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 4a--4c.
-We used a geographic attribute space to showcase the behavior of the problem.
-For each species, demand points were set as the planing units' centroids and were weighted according to the units' probability of occupancy.
+Planning unit occupancy was calculated using the $(X, Y)$ coordinates of the units' centroids and the corresponding equations.
+We used a geographic attribute space to provide an intuitive visualization of the solutions.
+Demand points were set as the planning units' centroids and were weighted according to the units' probability of occupancy.
 
@@
 We generated four solutions for each species.
-First to represent solutions generated using conventional methods, we generated solutions using only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-Second to show how the addition of space-based targets can affect a prioritization, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % space targets to ensure that the solution samples the geographic spread of the species (hereafter referred to as geographic spread targets).
-Third to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
-Fourth to illustrate the combined affect using space-based targets and penalising fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic spread targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
+Firstly, to represent solutions generated using conventional methods (eg.
+\texttt{Marxan}), we generated solutions using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
+Secondly, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets).
+Thirdly, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
+Fourthly, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic spread targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
 
@@
 
-The uniform species occurred in all planning units with equality probability (Fig.
+The solution generated for the uniformly distributed species using an amount target prioritized `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area (Fig.
 2a).
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount-based targets (Fig.
-3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area.
-This specific configuration is an artifact of the method used to solve the problem.
-In fact, for this species, all prioritizations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-In the absence of additional criteria, reserve selection methods may or may not return solutions that secure a representative sample of the features.
-In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points (equivalent to the $R^2$ of a regression model that explains less variation in the response variable than the mean of the response variable).
+For this species, all solutions containing this many planning units are optimal when using a `r sim.params.LST[[MODE]]$amount.target*100` % amount target to constrain the reserve selection problem.
+We obtained this particular solution due to artifacts in the solver.
+This solution captured a small amount of the geographic spread of the species (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
+In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the centre of the species' distribution.
 
-The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic spread target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
+By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
 3b).
-Since all planning units have equal cost and an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf.
-Fig 3a).
-Although the use of amount and space targets has yielded a solution that conserves a sample of the species that contains an adequate amount of habitat and a representative sample of the species' geographic spread, they have yielded a solution that is also highly fragmented solution.
-
-The addition of a positive boundary length modifier (BLM) parameter resulted in a well-connected solution that secured an adequate proportion (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % ) and representative sample (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % sampled) of the species geographic spread (Fig.
+Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented.
+However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well-connected solution that met all of the objectives (Fig.
 3d).
-Although this solution also contains `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units--the same amount as previously discussed solutions--it secures a representative proportion of the species geographic spread in a configuration that is not highly fragmented.
-Overall, these results show that under the simplest of conditions, using just amount targets results in an under-specified reserve selection problem that is unlikely to return an effective solution.
+This solution prioritized the same number of planning units as the previous solutions even though it is far superior.
+As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this objective.
 
@@
 
-The normally distributed species was simulated to contain a single core area where individuals were most prevalent, and marginal areas where individuals were less likely to occur (Fig.
-2b).
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig.
+The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core range area (Fig.
 3e).
-Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy.
-Whilst this solution conserves a sufficient amount of habitat in a cost-effective manner--it fails to secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).
-
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig.
+While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution.
+By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
 3f).
-This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution.
-By using amount and geographic spread targets in the reserve selection problem, we obtained a solution that concentrates conservation effort in the range core and also the range margin.
-However, similar to the corresponding solution for the uniform species (Fig.
-3b), this solution is also highly fragmented.
-
-By using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets and a boundary length modifier parameter ($BLM = 1$), we obtained a solution that fulfills adequacy, representativeness and connectivity objectives (Fig.
+Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (Fig.
 3h).
-This solution contains `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n` planning units.
-Unlike the solution for the uniform species generated using the same parameters, however, this solution contains more planning units than any other solutions for this species (cf.
-Figs.
-3e--3g).
-These results suggests that the combination of space targets and boundary length modifiers may require solutions that contain more planning units.
-They also also suggest that space targets can result in solutions that secure a more representative sample of the species--even for species without significant spatial structure.
+However, unlike the solutions for the uniformly distributed species, the solution generated using an amount target, a space target, and fragmentation penalties required more planning units than the other solutions.
+These results suggest that solutions may need to prioritize more planning units to meet additional conservation objectives.
 
@@
 
-The bimodally distributed species was simulated to represent a species with a highly structured distribution, and it contains two distinct populations (Fig.
-2c).
-The solution generated using just `r sim.params.LST[[MODE]]$amount.target*100` % amount targets prioritized individuals in just one of the two populations (Fig.
+The solution generated for the bimodally distributed species using an amount target only conserved individuals belonging to one of the two populations (Fig.
 3i).
-While this solution secured an adequate proportion of the species' distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$amount.held,2)` % secured), it did not sample a representative proportion of the range (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-The addition of boundary length modifiers to amount targets resulted in a solution which sampled even less of the species distribution (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig.
+As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
+The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig.
 3k).
-However, the addition of geographic spread targets resolved these issues.
-
-The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets included planning units from both populations (Fig.
-3j).
-Although this solution required more planning units to fulfill both targets ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$n`$), this solution secured a representative sample of the bimodally distributed species (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % range sampled).
-Similar to corresponding solution generated for the normally distributed species (Fig.
-3f), this solution was also highly fragmented and we could obtain a more well-connected solution at the expense of selecting additional planning units ($n=`r filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$n`$; Fig.
+However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % geographic spread secured; Fig.
 3l).
-The results for the bimodally distributed species suggest that species with highly structured populations or strong variation between individuals could benefit the most from prioritizations that are generated using space-based targets.
+This finding suggests that species with large intra-specific variation could benefit the most from prioritizations generated using space-based targets.
 
@@
 The \texttt{raptr R} package provides a unified approach to reserve selection.
-Conservation planners can use this \texttt{R} package to generate prioritizations that secure intra-specific as well as as inter-specific biodiversity patterns.
-Additionally, the package contains functionality to accommodate uncertainty in the distribution of features using the reliable formulation (defined in Appendix S1), and ensure that solutions are well-connected.
-Both the simulated and case study species suggest that explicitly considering representation as well as amount targets can significantly alter prioritizations.
+Conservation planners can use this \texttt{R} package to generate prioritizations that secure intra- and inter-specific biodiversity patterns.
+Both the simulations and case studies show that explicitly targeting a representative sample of each feature in a conservation planning exercise can substantially alter prioritizations.
+Additionally, we found that prioritizations often needed to secure more habitat to conserve a representative sample of each feature.
+Although it may not be practical to conserve substantially more habitat than the amount specified using amount-based targets [eg.
+10 % geographic range; @r255], we did find that even small increases in reserve size could yield large gains.
 
@@
 @r468].
-We caution that inaccuracy and uncertainty in data can negatively impact prioritizations [@r488], and recommend that planners ensure that data are reliable or accommodate uncertainty into the planning process (eg.
-by using the reliable formulation).
+We caution that inaccuracy and uncertainty in data can negatively impact prioritizations [@r488], and recommend that planners ensure that data are sufficiently reliable or use the reliable formulation (described in Appendix S1) to accommodate uncertainty into the reserve selection process.
 
 To maximize the long-term persistence of biodiversity, decision makers need to identify prioritizations that preserve existing patterns of biodiversity and the processes that support them.
-To achieve this, conservation planners need a decision support tool that can deliver cost-effective prioritizations that secure an adequate amount of a representative sample of biodiversity features.
-Here, we developed the \texttt{raptr R} package to fill this void.
-By exploring the functionality of this package using several simulated and case study species, we found that including space-based targets can result in substantially different solutions.
+It is becoming increasingly apparent that simply protecting a large area will fail to achieve this [@r492].
+Here, we developed the \texttt{raptr R} package to provide conservation planners with the tools to deliver cost-effective prioritizations that secure an adequate amount of a representative sample of biodiversity features.
+By exploring the functionality of this software package, we found that explicitly targeting a representative sample of each feature can result in substantially different solutions.
 

### 2017-03-15 code/R/analysis/00-initialization.R
manually specify cran snapshot

@@ -12,6 +12,7 @@ if (length(args)>0) {
 # set checkpoint
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
+checkpoint::setSnapshot('2016-11-26')
 checkpoint::checkpoint('2016-11-26', R.version='3.3.2')
 if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
 
@@ -87,6 +88,9 @@ if (!'gurobi' %in% installed.packages()[,'Package']) {
 devtools::install_github('jeffreyhanson/ggplot2', force=TRUE)
 library(ggplot2)
 
+# install bioconductor packages
+
+
 ### set parameters
 if (!exists('MODE')) MODE <- 'debug'
 cat('MODE = ',MODE,'\n')
@@ -110,5 +114,3 @@ for (x in dir(file.path('code', 'R', 'functions'), full.names=TRUE)) source(x)
 
 # save workspace
 save.session('data/intermediate/00-initialization.rda', compress='xz')
-
- 

### 2017-03-15 code/R/analysis/05-benchmark-analysis.R
update benchmark anlysis

@@ -2,75 +2,51 @@
 checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
 session::restore.session('data/intermediate/00-initialization.rda')
 
-## load parameeters
+## load parameters
 benchmark.params.LST <- parseTOML('code/parameters/benchmark.toml')
 
-## run benchmark analysis for unreliable formulation
-unreliable.benchmark.params.DF <- expand.grid(number.features=benchmark.params.LST[[MODE]][['unreliable.number.of.features']],
-  number.planning.units=benchmark.params.LST[[MODE]][['unreliable.number.of.planning.units']],
-  replicate=seq_len(benchmark.params.LST[[MODE]][['replicates']]))
+# load simulation functions
+source('code/R/functions/simulate.R')
 
-unreliable.benchmark.DF <- ldply(sample.int(nrow(unreliable.benchmark.params.DF)), function(i) {
+## compile benchmark parameters
+benchmark.params.DF <- ldply(seq_along(benchmark.params.LST[[MODE]][['number.of.features']]), function(i) {
+  expand.grid(number.features=benchmark.params.LST[[MODE]][['number.of.features']][[i]],
+    number.planning.units=benchmark.params.LST[[MODE]][['number.of.planning.units']][[i]],
+    formulation = benchmark.params.LST[[MODE]][['formulation']],
+    blm = benchmark.params.LST[[MODE]][['blm']],
+    replicate = seq_len(benchmark.params.LST[[MODE]][['replicates']]))
+  }
+)
+
+benchmark.DF <- ldply(benchmark.params.DF, function(i) {
   # simulate data
-  source('code/R/functions/simulate.R')
   curr.rd <- simulate.problem.data(
-    number.features=unreliable.benchmark.params.DF[i,1],
-    number.planning.units=unreliable.benchmark.params.DF[i,2],
+    number.features=benchmark.params.DF[i,1],
+    number.planning.units=benchmark.params.DF[i,2],
     amount.target=benchmark.params.LST[[MODE]]$amount.target,
     space.target=benchmark.params.LST[[MODE]]$space.target,
     probability.of.occupancy=benchmark.params.LST[[MODE]]$occupancy.probability
   )
+  # create unsolved problem
[…]

### 2017-03-15 code/R/functions/simulate.R

@@ -18,7 +18,7 @@
 #'
 #' @return \code{\link[raster]{RasterStack-class}} object with a
 #'   layer for each species.
-#' 
+#'
 #' @seealso \code{\link[RandomFields]{RFsimulate}},
 #'   \code{\link{simulate_cost}}, \code{\link{simulate_species}}.
 #'
@@ -53,7 +53,7 @@ simulate_data <- function(x, n, model, transform=identity, ...) {
 #'
 #' Generates a random set of species using random field models. By default,
 #' the output will contain values between zero and one.
-#' 
+#'
 #' @inheritParams simulate_data
 #'
 #' @return \code{\link[raster]{RasterStack-class}} object.
@@ -67,7 +67,7 @@ simulate_species <- function(x, n=1, model=RandomFields::RMgauss(),
 
 #' Simulate problem data
 #'
-#' Simulates data for a conservation planning problem. 
+#' Simulates data for a conservation planning problem.
 #'
 #' @param number.planning.units \code{integer} number of planning units in problem.
 #' @param number.features \code{integer} number of features in problem.
@@ -81,15 +81,15 @@ simulate.problem.data <- function(number.planning.units, number.features, probab
   # simulate landscape
   curr.landscape <- raster::raster(ncol=ceiling(sqrt(number.planning.units)), nrow=ceiling(sqrt(number.planning.units))) %>%
     setValues(1) %>%  `extent<-`(c(0,1,0,1))
-  
+
   # simulate species
-  curr.spp <- simulate_species(curr.landscape, n=number.features, 
+  curr.spp <- simulate_species(curr.landscape, n=number.features,
                 model=RPbernoulli(RMgauss()), transform=identity)
-  
+
   # change extents so that polygon boundarys are not too small
   extent(curr.landscape) <- c(0, ceiling(sqrt(number.planning.units)), 0, ceiling(sqrt(number.planning.units)))
   extent(curr.spp) <- extent(curr.landscape)
-  
+
   # simulate pus
   curr.pus <- rasterToPolygons(curr.landscape, n=4)
   curr.pus <- curr.pus[seq_len(number.planning.units),]
@@ -102,7 +102,7 @@ simulate.problem.data <- function(number.planning.units, number.features, probab
   for (j in which(apply(mtx, 2, sum) < 2)) {
     mtx[sample.int(nrow(mtx), size=2),j] <- probability.of.occupancy
   }
-  
+
[…]

### 2017-03-15 code/R/analysis/05-benchmark-analysis.R
fx bugs in benchmark

@@ -12,13 +12,13 @@ source('code/R/functions/simulate.R')
 benchmark.params.DF <- ldply(seq_along(benchmark.params.LST[[MODE]][['number.of.features']]), function(i) {
   expand.grid(number.features=benchmark.params.LST[[MODE]][['number.of.features']][[i]],
     number.planning.units=benchmark.params.LST[[MODE]][['number.of.planning.units']][[i]],
-    formulation = benchmark.params.LST[[MODE]][['formulation']],
-    blm = benchmark.params.LST[[MODE]][['blm']],
-    replicate = seq_len(benchmark.params.LST[[MODE]][['replicates']]))
+    formulation=benchmark.params.LST[[MODE]][['formulation']][[i]],
+    blm=benchmark.params.LST[[MODE]][['blm']][[i]],
+    replicate=seq_len(benchmark.params.LST[[MODE]][['replicates']]))
   }
 )
 
-benchmark.DF <- ldply(benchmark.params.DF, function(i) {
+benchmark.DF <- ldply(sample.int(nrow(benchmark.params.DF)), function(i) {
   # simulate data
   curr.rd <- simulate.problem.data(
     number.features=benchmark.params.DF[i,1],
@@ -35,6 +35,8 @@ benchmark.DF <- ldply(benchmark.params.DF, function(i) {
   } else {
     curr.ru <- RapUnsolved(RapReliableOpts(BLM=benchmark.params.DF[i,4]), curr.rd)
   }
+  # print problem object for logging
+  print(curr.ru)
   # solve problems
   curr.time <- system.time({curr.rs <- solve(curr.ru, curr.go)})
   # return results

### 2017-03-18 code/rmarkdown/article.Rmd
fix bug in benchmark fix Jonathon's affiliation in ms

@@
   | $^1$School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
-  | $^2$School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
+  | $^2$School of Earth and Environmental Sciences, The University of Queensland, Brisbane, QLD, Australia
   | $^3$The Nature Conservancy, South Brisbane, QLD, Australia
-  | Correspondance should be addressed to jeffrey.hanson@uqconnect.edu.au
+  | Correspondence should be addressed to jeffrey.hanson@uqconnect.edu.au
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 individuals, areas) that is representative in terms of the intrinsic or extrinsic physical attributes that underpin the process of interest.
-For example, by conserving a sample of populations with local adaptations that is representative in terms of the full variation found in the species--physical attributes associated with adaptation--protected areas can foster adaptative processes by ensuring these adaptations are not lost.
+For example, by conserving a sample of populations with local adaptations that is representative in terms of the full variation found in the species--physical attributes associated with adaptation--protected areas can foster adaptive processes by ensuring these adaptations are not lost.
 Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature.

### 2017-03-18 code/R/analysis/05-benchmark-analysis.R

@@ -29,7 +29,8 @@ benchmark.DF <- ldply(sample.int(nrow(benchmark.params.DF)), function(i) {
   )
   # create unsolved problem
   curr.go <- GurobiOpts(MIPGap=general.params.LST[[MODE]][['MIPGap']],
-    Threads=general.params.LST[[MODE]][['threads']])
+    Threads=general.params.LST[[MODE]][['threads']],
+    TimeLimit = benchmark.params.LST[[MODE]]$time.limit)
   if (benchmark.params.DF[i,3] == "unreliable") {
     curr.ru <- RapUnsolved(RapUnreliableOpts(BLM=benchmark.params.DF[i,4]), curr.rd)
   } else {

### 2017-04-06 code/rmarkdown/figures.Rmd
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references

@@
 
-```{r, include=FALSE}
+```{r, include = FALSE}
 # set wd
-knitr::opts_knit$set(root.dir=normalizePath('../..'))
+knitr::opts_knit$set(root.dir = normalizePath("../.."))
 ```
 
-```{r, include=FALSE}
+```{r, include = FALSE}
 ## initialization
 # load .rda file
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-try(session::restore.session('data/final/results.rda'))
-
-# set options
-options(error=NULL)
+try(session::restore.session("data/final/results.rda"))
 
@@
 # redefine file paths for rasters
-for (i in seq_len(nlayers(cs1.bioclim.RAST)))
-	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])
+for (i in seq_len(raster::nlayers(cs1_bioclim_raster)))
+  cs1_bioclim_raster@layers[[i]]@file@name  <- paste0(getwd(),
+    strsplit(cs1_bioclim_raster@layers[[i]]@file@name,
+             "raptr-manuscript")[[1]][[2]])
 
 # remove small islands from australia basemap
-aus.sub.PLY <- disaggregate(aus.PLY)
-aus.sub.PLY <- SpatialPolygonsDataFrame(aus.sub.PLY, data=data.frame(size=gArea(aus.sub.PLY, byid=TRUE)))
-aus.sub.PLY <- aus.sub.PLY[aus.sub.PLY$size >= quantile(aus.sub.PLY$size, 0.9, names=FALSE),]
+aus_sub_polygons <- sp::disaggregate(aus_polygons)
+aus_sub_polygons <- aus_sub_polygons %>%
+                    sp::SpatialPolygonsDataFrame(
+                      data = data.frame(size = rgeos::gArea(aus_sub_polygons,
+                                                            byid = TRUE)))
+aus_sub_polygons <- aus_sub_polygons[aus_sub_polygons$size >=
+  quantile(aus_sub_polygons$size, 0.9, names=FALSE),]
 
 # fortify polygons
-sim.pus.FPLY <- fortify(sim.pus)
-aus.FPLY <- fortify(aus.sub.PLY)
-cs1.pus.FPLY <- fortify(cs1.pus)
-cs2.grid.FPLY <- fortify(cs2.grid.PLY)
-cs2.grid.sub.FPLY <- fortify(cs2.grid.sub.PLY)
+sim_pus_fpolygons <- ggplot2::fortify(sim_pus)
+aus_fpolygons <- ggplot2::fortify(aus_sub_polygons)
+cs1_pus_fpolygons <- ggplot2::fortify(cs1_pus)
+cs2_grid_fpolygons <- ggplot2::fortify(cs2_grid_polygons)
+cs2_grid_sub_fpolygons <- ggplot2::fortify(cs2_grid_sub_polygons)
 
 # store countries
-data(countriesHigh)
-countries.FPLY <- countriesHigh[
+data(countriesHigh, package = "rworldxtra")
+countries_fpolygons <- countriesHigh[
   countriesHigh$ADMIN %in% c(
-		'Italy', 'Switzerland', 'France', 'Austria',
-		'Germany', 'Slovenia', 'Croatia', 'Hungary',
-		'Monaco', 'Germany', 'Slovakia', 'Czech Republic'
-	)
-,] %>% spFortify
+    "Italy", "Switzerland", "France", "Austria",
+    "Germany", "Slovenia", "Croatia", "Hungary",
+    "Monaco", "Germany", "Slovakia", "Czech Republic"
+  ),] %>% sp_fortify
 
-column.width <- 2.75591
+column_width <- 2.75591
 
@@
 
-```{r, echo=FALSE, message=FALSE, fig.height=2, fig.width=2, fig.cap="Attribute space example.
+```{r, echo = FALSE, message = FALSE, fig.height = 2, fig.width = 2, fig.cap = "Attribute space example.
 This environmental attribute space has dimensions relating to annual temperature ($^{\\circ}$C) and rainfall (mm).
@@
 # population data
-populations <- data.frame(
-	population=LETTERS[1:4],
-	temperature=c(35,34,31,21.6),
-	rainfall=c(4.5,5.3,9.5,5.5)   
-)
+populations <- data.frame(population = LETTERS[1:4],
+                          temperature = c(35, 34, 31, 21.6),
+                          rainfall = c(4.5, 5.3, 9.5, 5.5))
 
 # demand point data
-make.dp <- function(x, n, pad.percent=0.2) {
-	padding<-diff(range(x))*pad.percent
-	return(seq(min(x)-padding, max(x)+padding, length.out=n))
+make_dp <- function(x, n, pad.percent = 0.2) {
+  padding <- diff(range(x)) * pad.percent
+  return(seq(min(x) - padding, max(x) + padding, length.out = n))
 }
 demand.points <- expand.grid(
-	temperature=make.dp(populations$temperature, n=4, pad.percent=0.2),
-	rainfall=make.dp(populations$rainfall, n=4, pad.percent=0.2)
-)
+  temperature = make_dp(populations$temperature, n = 4, pad.percent = 0.2),
+  rainfall = make_dp(populations$rainfall, n = 4, pad.percent = 0.2))
 
 ## plot data
-ggplot(aes(x=rainfall, y=temperature), data=populations) +
-	geom_point(data=demand.points, fill='gray70') +
-	geom_text(aes(label=population), size=6) +
-	theme_classic() +
-	ylab(expression('Temperature ('*degree*'C)')) +
-	xlab('Rainfall (mm)') +
-	theme(axis.line.x=element_line(), axis.line.y=element_line())
+ggplot2::ggplot(ggplot2::aes(x = rainfall, y = temperature),
+                data = populations) +
+  ggplot2::geom_point(data = demand.points, fill = "gray70") +
+  ggplot2::geom_text(ggplot2::aes(label = population), size = 6) +
+  ggplot2::theme_classic() +
+  ggplot2::ylab(expression("Temperature ("*degree*"C)")) +
+  ggplot2::xlab("Rainfall (mm)") +
+  ggplot2::theme(axis.line.x = ggplot2::element_line(),
+                 axis.line.y = ggplot2::element_line())
 ```
 
-```{r, echo=FALSE, messages=FALSE, results='hide', fig.height=1.6, fig.width=column.width, fig.cap='Distributions of three simulated species.
+```{r, echo = FALSE, messages = FALSE, results = "hide", fig.height = 1.6, fig.width = column_width, fig.cap = "Distributions of three simulated species.
 Squares denote planning units.
-Colors indicate probability of occupancy.'}
+Colors indicate probability of occupancy."}
 # prepare data
-sim.pus.occ.FPLY <- ldply(
-	seq_len(nrow(sim.species)),
+sim_pus_occ_fpolygons <- plyr::ldply(
+  seq_len(nrow(sim_species)),
   function(i) {
-		curr.DF <- sim.pus.FPLY
-		curr.DF$species <- sim.species$name[i]
-		curr.DF$Probability <- sim.dps[[i]]@weights[as.numeric(curr.DF$id)]
-		return(curr.DF)
+    curr_data <- sim_pus_fpolygons
+    curr_data$species <- sim_species$name[i]
+    curr_data$Probability <- sim_demand_points[[i]]@weights[
+      as.numeric(curr_data$id)]
+    return(curr_data)
   }
 )
-sim.pus.occ.FPLY$species <- gsub('\n', ' ', as.character(sim.pus.occ.FPLY$species))
-sim.pus.occ.FPLY$species <- factor(as.character(sim.pus.occ.FPLY$species), levels=gsub('\n', ' ', sim.species$name)) 
+sim_pus_occ_fpolygons$species <- gsub("\n", " ",
+  as.character(sim_pus_occ_fpolygons$species))
+sim_pus_occ_fpolygons$species <- factor(as.character(
+  sim_pus_occ_fpolygons$species), levels = gsub("\n", " ", sim_species$name))
 # make letter data.frame
-sim.pus.occ.letters.DF <- data.frame(
-	x=min(sim.pus.occ.FPLY$long) + 1.5,
-	y=max(sim.pus.occ.FPLY$lat) - 1.5,
-	species=unique(sim.pus.occ.FPLY$species),
-	letter=paste0('(', letters[seq_along(sim.species$name)], ')')
-)
+sim_pus_occ_letters <- data.frame(
+  x = min(sim_pus_occ_fpolygons$long) + 1.5,
+  y = max(sim_pus_occ_fpolygons$lat) - 1.5,
+  species = unique(sim_pus_occ_fpolygons$species),
+  letter=paste0("(", letters[seq_along(sim_species$name)], ")"))
 # make plot
-ggplot() +
-	geom_polygon(data=sim.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Probability), color='black') +
-	theme_classic() +
-	geom_label(data=sim.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=2) +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=7),
-    axis.title = element_blank(),
-    legend.title = element_text(color='black', size=7),
-		legend.text = element_text(color='black', size=7),
-		legend.margin = unit(0, 'cm'),
-		legend.position='bottom') +
-	facet_wrap(~ species, ncol=3)
+ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = sim_pus_occ_fpolygons,
+                        ggplot2::aes(x = long, y = lat,  group = group,
+                                     fill = Probability),
+                        color="black") +
+  ggplot2::theme_classic() +
+  ggplot2::geom_label(data = sim_pus_occ_letters,
+                     ggplot2::aes(label = letter, x = x, y =y ), color="black",
+                     fill = scales::alpha("white", 0.8), size = 2) +
+  ggplot2::theme(axis.ticks = ggplot2::element_blank(),
+                axis.text = ggplot2::element_blank(),
+                axis.line = ggplot2::element_blank(),
+                strip.background = ggplot2::element_rect(fill = "grey20"),
+                strip.text = ggplot2::element_text(color = "white", size = 7),
+                axis.title = ggplot2::element_blank(),
+                legend.text = ggplot2::element_text(color = "black", size = 7),
+                legend.margin = ggplot2::unit(0, "cm"),
+                legend.position = "bottom") +
+  ggplot2::facet_wrap(~ species, ncol = 3)
 ```
 
-```{r, echo=FALSE, fig.height=3.5, fig.width=2*column.width, fig.cap=paste0('Prioritizations for the simulation study.
+```{r, echo = FALSE, fig.height = 3.5, fig.width = 2 * column_width, fig.cap = paste0("Prioritizations for the simulation study.
 Each panel shows a prioritization generated for a single species using a set of parameters.
 Squares denote planning units.
-Planning units shown in dark green units were selected for protection.
+Dark green planning units were selected for protection.
 Each row of panels show prioritizations generated for a different species.
-Each column of panels corresponds to a different set of parameters used to generate the prioritization.')}
+Each column of panels corresponds to a different set of parameters used to generate the prioritization.")}
 # prepare data
-sim.pus.sel.FPLY <- ldply(
-	seq_len(nrow(sim.spp.DF)),
+sim_pus_sel_fpolygons <- plyr::ldply(seq_len(nrow(sim_spp_results)), function(i) {
+    curr_data <- sim_pus_fpolygons %>%
+                 dplyr::mutate(species = sim_spp_results$species[i],
+                               prioritisation =
+                                sim_spp_results$prioritisation[i])
+    curr_selection <- sim_prioritisations %>%
+                      `[[`(as.character(sim_spp_results$species[i])) %>%
+                      `[[`(as.character(sim_spp_results$prioritisation[i])) %>%
+                      raptr::selections() %>%
+                      `[`(as.numeric(curr_data$id))
+    curr_data$Selection <-  c("discarded","selected")[curr_selection+1]
+    return(curr_data)
+})
+sim_pus_sel_fpolygons$prioritisation <- factor(
+  as.character(sim_pus_sel_fpolygons$prioritisation),
+  levels = c("Amount\ntargets", "Amount &\nspace targets",
+             "Amount target\nand BLM", "Amount & space\ntargets and BLM"))
+# prepare plotting data
+sim_letter_data2 <- sim_spp_results %>%
+  dplyr::mutate(letter = paste0("(", letters[seq_along(species)],")"),
+                long = min(sim_pus_sel_fpolygons$long) + 1.6,
+                lat = max(sim_pus_sel_fpolygons$lat) - 1.5)
+# make plot
+ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = sim_pus_sel_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                     fill = Selection),
+                         color="black") +
+  ggplot2::geom_label(data = sim_letter_data2,
+                      ggplot2::aes(label = letter, x = long, y = lat),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size=5) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.ticks.x = ggplot2::element_blank(),
+    axis.ticks.y = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.text.x = ggplot2::element_blank(),
+    axis.text.y = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    axis.line.x = ggplot2::element_blank(),
+    axis.line.y = ggplot2::element_blank(),
+    plot.background = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    panel.background = ggplot2::element_blank(),
+    panel.border = ggplot2::element_blank(),
+    legend.position="none",
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white"),
+    axis.title = ggplot2::element_blank()) +
+  ggplot2::scale_fill_manual(name="Solution",
+                             values = c("discarded"="#f7fcfd",
+                                        "selected"="#00441b")) +
+  ggplot2::facet_grid(species ~ prioritisation)
+```
+
+```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = "Two main gradients of climatic variation across Queensland, Australia. Polygons denote planning units."}
+# prepare data
+cs1_pca_fpoints <- plyr::ldply(seq_len(cs1_parameters$number_pc_layers),
   function(i) {
-		curr.DF <- sim.pus.FPLY %>% mutate(species=sim.spp.DF$species[i], prioritisation=sim.spp.DF$prioritisation[i])
-		curr.selection <- selections(sim.prioritisations[[as.character(sim.spp.DF$species[i])]][[as.character(sim.spp.DF$prioritisation[i])]])[as.numeric(curr.DF$id)]
-		curr.DF$Selection <-  c('discarded','selected')[curr.selection+1]
-		return(curr.DF)
+    curr_data <- cs1_bioclim_raster[[i]] %>%
+                     raster::aggregate(10) %>%
+                     raster::rasterToPoints() %>%
+                     as.data.frame()
+    names(curr_data)[3] <- "Score"
+    curr_data$pc <- paste0("PC ",i)
+    return(curr_data)
+})
+# make plots
+p1 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = aus_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill = "grey85", color = "grey70") +
+  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
+                                            pc == unique(pc)[1]),
+                       ggplot2::aes(x = x, y = y, fill = Score)) +
+  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill = NA, color = "black", size = 0.05) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white"),
+    axis.title = ggplot2::element_blank(),
+    legend.background = ggplot2::element_rect(color = "transparent",
+                                              fill = "transparent"),
+    legend.position = c(1.1, 1.07),
+    legend.justification = c(1,1),
+    legend.key.height = ggplot2::unit(1.5, "mm"),
+    legend.key.width = ggplot2::unit(2, "mm")) +
+  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
+                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
+                           expand = FALSE) +
+  ggplot2::scale_fill_continuous(breaks =
+    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[1])$Score, n = 3,
+          min.n = 2)) +
+  ggplot2::facet_wrap(~ pc)
+p2 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = aus_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill="grey85", color="grey70") +
+  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
+                                            pc == unique(pc)[2]),
+                       ggplot2::aes(x = x, y = y, fill = Score)) +
+  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill = NA, color = "black", size = 0.05) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white"),
+    axis.title = ggplot2::element_blank(),
+    legend.background = ggplot2::element_rect(color="transparent",
+                                              fill="transparent"),
+    legend.position = c(1.1, 1.07),
+    legend.justification = c(1,1),
+    legend.key.height = ggplot2::unit(1.5, "mm"),
+    legend.key.width = ggplot2::unit(2, "mm")) +
+  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
+                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
+                           expand = FALSE) +
+  ggplot2::scale_fill_continuous(breaks =
+    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[2])$Score, n = 2,
+           min.n = 2)) +
+  ggplot2::facet_wrap(~ pc)
+# render plots
+gridExtra::grid.arrange(p1, p2, ncol = 2)
+```
+
+```{r, echo = FALSE, fig.height = 3.8, fig.width = column_width, fig.cap = "Distribution of the species used in the first case study.
+Polygons denote planning units.
+Planning units occupied by a given species are shown in light blue."}
+# prepare data
+cs1_pus_occ_fpolygons <- plyr::ldply(
+  seq_along(cs1_parameters$common_names),
+  function(i) {
+    curr_data <- cs1_pus_fpolygons
+    curr_data$species <- cs1_parameters$common_names[i]
+    curr_data$Occupancy <- "absent"
+    curr_pus <- cs1_ru@data@pu.species.probabilities %>%
+                dplyr::filter(species == i) %>%
+                `[[`("pu")
+    curr_data$Occupancy[which(curr_data$id %in% curr_pus)] <- "present"
+    return(curr_data)
+})
+
+cs1_pus_occ_fpolygons$species <- factor(
+    as.character(cs1_pus_occ_fpolygons$species),
+    levels = cs1_parameters$common_names)
+# make letter data.frame
+cs1_pus_occ_letters_data <- data.frame(
+  x = unname(cs1_pus@bbox[1,1] + (diff(cs1_pus@bbox[1, ]) * 0.08)),
+  y = unname(cs1_pus@bbox[2,1] + (diff(cs1_pus@bbox[2, ]) * 0.95)),
+  species=unique(cs1_pus_occ_fpolygons$species),
+  letter=paste0("(", letters[seq_along(unique(cs1_pus_occ_fpolygons$species))],
+                ")")
+)
+# make plots
+ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = aus_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill = "grey85", color = "grey70", size = 0.05) +
+  ggplot2::geom_polygon(data = cs1_pus_occ_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                     fill = Occupancy),
+                        color = "black", size = 0.05) +
+  ggplot2::geom_label(data = cs1_pus_occ_letters_data,
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8), size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white", size = 7),
+    axis.title = ggplot2::element_blank(),
+    legend.position = "none") +
+  ggplot2::scale_fill_manual(name = "Occupancy",
+                             values = c("absent"="#132B43",
+                                        "present"="#56B1F7")) +
+  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
+                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
+                           expand = FALSE) +
+  ggplot2::facet_wrap(~ species, ncol = 2)
+```
+
+```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = paste0("Prioritizations for the first case study.
+Polygons denote planning units.
+Dark green planning units were selected for protection.
+Panel (a) shows the solution generated when using ",cs1_parameters$amount_target*100," % amount targets.
+Panel (b) shows the solution when using ", cs1_parameters$amount_target*100, " % amount targets and ", cs1_parameters$space_target*100, " % niche targets.")}
+# prepare data
+cs1_pus_sel_fpolygons <- plyr::ldply(
+  names(cs1_prioritisations),
+  function(x) {
+    curr_data <- cs1_pus_fpolygons
+    curr_data$Prioritisation <- x
+    curr_data$Solution <- "discarded"
+    curr_pus <- which(as.logical(raptr::selections(cs1_prioritisations[[x]])))
+    curr_data$Solution[curr_data$id %in% curr_pus] <- "selected"
+    return(curr_data)
   }
 )
-sim.pus.sel.FPLY$prioritisation <- factor(as.character(sim.pus.sel.FPLY$prioritisation),
-	levels=c('Amount\ntargets', 'Amount &\nspace targets', 'Amount target\nand BLM', 'Amount & space\ntargets and BLM'))
-# prepare plotting data
-sim.letter.DF2 <- sim.spp.DF %>%
-  mutate(letter=paste0('(', letters[seq_along(species)],')'),
-         long=min(sim.pus.sel.FPLY$long) + 1.6,
-         lat=max(sim.pus.sel.FPLY$lat) - 1.5)
-# make plot
-ggplot() +
-	geom_polygon(data=sim.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Selection), color='black') +
-	geom_label(data=sim.letter.DF2, aes(label=letter, x=long, y=lat), color='black', fill=alpha('white', 0.8), size=5) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
-		axis.text=element_blank(), axis.text.x=element_blank(), axis.text.y=element_blank(),
-		axis.line=element_blank(), axis.line.x=element_blank(), axis.line.y=element_blank(),
-		plot.background=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		panel.background=element_blank(), panel.border=element_blank(), legend.position='none',
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank()) +
-	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
-	facet_grid(species ~ prioritisation)
+cs1_pus_sel_fpolygons$Prioritisation <- factor(as.character(
+  cs1_pus_sel_fpolygons$Prioritisation),
+  levels = names(cs1_prioritisations))
+# make letter data.frame
+cs1_pus_sel_letters_data <- data.frame(
+  x = unname(cs1_pus@bbox[1,1] + (diff(cs1_pus@bbox[1, ]) * 0.08)),
+  y = unname(cs1_pus@bbox[2,1] + (diff(cs1_pus@bbox[2, ]) * 0.95)),
+  Prioritisation = names(cs1_prioritisations),
+  letter = paste0("(", letters[1:2], ")"))
+# make plots
+ggplot2::ggplot() +
+  ggplot2::geom_polygon(data=aus_fpolygons,
+                        ggplot2::aes(x=long, y=lat, group=group),
+                        fill = "grey85", color = "grey70", size = 0.05) +
+  ggplot2::geom_polygon(data = cs1_pus_sel_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                    fill = Solution),
+                        color = "black", size = 0.05) +
+  ggplot2::geom_label(data = cs1_pus_sel_letters_data,
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                                                             size=3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white", size = 8),
+    axis.title = ggplot2::element_blank(),
+    legend.position = "none") +
+  ggplot2::scale_fill_manual(name = "Solution",
+                             values = c("discarded"="#f7fcfd",
+                                        "selected"="#00441b")) +
+  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
+                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
+                           expand = FALSE) +
+  ggplot2::facet_wrap(~ Prioritisation, ncol = 2)
 ```
 
-```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap='Distribution of climatic variation across Queensland, Australia.
+```{r, echo = FALSE, fig.height = 2.2, fig.width = column_width, fig.cap = paste0("Data used for the second case study.
 Squares denote planning units.
-Panels show the two main gradients of climatic variation.'}
-# prepare data
-cs1.pca.FPTS <- ldply(
-	seq_len(cs1.params.LST[[MODE]]$n.pc.layers),
-	function(i) {
-		curr.DF <- cs1.bioclim.RAST[[i]] %>% 
-			aggregate(10) %>% 
-			rasterToPoints() %>% 
-			as.data.frame()
-		names(curr.DF)[3] <- 'Score'
-		curr.DF$pc <- paste0('PC ',i)
-		return(curr.DF)
-	}
-)
-# make plots
-p1 <- ggplot() +
-	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[1]), aes(x=x, y=y, fill=Score)) +
-	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black', size=0.05) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
-		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
-	) +
-  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
-  scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[1])$Score, n=3, min.n=2)) +
-	facet_wrap(~ pc)
-p2 <- ggplot() +
-	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_raster(data=filter(cs1.pca.FPTS, pc==unique(pc)[2]), aes(x=x, y=y, fill=Score)) +
-	geom_polygon(data=cs1.pus.FPLY, aes(x=long, y=lat, group=group), fill=NA, color='black', size=0.05) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.background=element_rect(color='transparent', fill='transparent'),
-		legend.position=c(1.1, 1.07), legend.justification=c(1,1), legend.key.height=unit(1.5, 'mm'), legend.key.width=unit(2, 'mm')
-		) +
-  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
-	scale_fill_continuous(breaks=pretty(filter(cs1.pca.FPTS, pc==unique(pc)[2])$Score, n=2, min.n=2)) +
-	facet_wrap(~ pc)
+Panel (a) shows all grid cells surveyed by the IntraBioDiv project.
+Grid cells occupied by the ",tolower(cs2_parameters$common_name)," are shown in bright blue.
+The subsequent panels contain only show occupied grid cells.
+Panel (b) shows the acquisition cost of each planning unit (estimated as the total human population density).
+Panels (c--d) show the spatial distribution of the ordinations describing genetic variation.
+These values describe the typical genetic characteristics of individuals in each planning unit.
+Planning units with similar values/colors contain individuals with similar loci polymorphisms.")}
+# prepare full grid data
+cs2_spp_grid_fpolygons <- cs2_grid_fpolygons
+cs2_spp_grid_fpolygons$Title <- "Occupancy"
+cs2_spp_grid_fpolygons$Occupancy <- "absent"
+status <- cs2_grid_polygons@data[[cs2_parameters$species_name]][as.numeric(
+        cs2_spp_grid_fpolygons$id)]
+cs2_spp_grid_fpolygons$Occupancy[which(status==1)] <- "present"
+
+cs2_spp_grid_letters_data <- data.frame(
+  x = unname(cs2_grid_polygons@bbox[1, 1] +
+      (diff(cs2_grid_polygons@bbox[1, ]) * 0.07)),
+  y = unname(cs2_grid_polygons@bbox[2, 1] +
+      (diff(cs2_grid_polygons@bbox[2, ]) * 0.875)),
+  letter = ("(a)"))
+
+# prepare subset grid data
+cs2_spp_grid_sub_fpolygons <- cs2_grid_sub_fpolygons
+cs2_spp_grid_sub_fpolygons$nmds1 <- NA
+cs2_spp_grid_sub_fpolygons$nmds2 <- NA
+cs2.nmds1 <- cs2_grid_sub_data[[paste0(cs2_parameters$species_name,
+                                       "_genetic_d1")]]
+cs2.nmds2 <- cs2_grid_sub_data[[paste0(cs2_parameters$species_name,
+                                       "_genetic_d2")]]
+
+curr_pos <- match(as.numeric(cs2_spp_grid_sub_fpolygons$id),
+                  cs2_grid_sub_data$id)
+cs2_spp_grid_sub_fpolygons$nmds1 <- cs2.nmds1[curr_pos]
+cs2_spp_grid_sub_fpolygons$nmds2 <- cs2.nmds2[curr_pos]
+
+ids <- cs2_grid_sub_polygons$id[as.numeric(cs2_spp_grid_sub_fpolygons$id)]
+cs2_spp_grid_sub_fpolygons$Cost <- costs_matrix[match(ids,
+                                                      costs_matrix[, 1]), 2]
+
+cs2_spp_grid_sub_fpolygons$Title1 <- "Opportunity cost"
+cs2_spp_grid_sub_fpolygons$Title2 <- "Genetic NMDS 1"
+cs2_spp_grid_sub_fpolygons$Title3 <- "Genetic NMDS 2"
+
+cs2_spp_grid_sub_letters_data <- data.frame(
+  x = unname(cs2_grid_sub_polygons@bbox[1, 1] +
+      (diff(cs2_grid_sub_polygons@bbox[1, ]) * 0.06)),
+  y = unname(cs2_grid_sub_polygons@bbox[2, 1] +
+      (diff(cs2_grid_sub_polygons@bbox[2, ]) * 0.89)),
+  letter = c("(b)", "(c)", "(d)"))
+
+# plot species data
+p1 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = countries_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill = "grey85", color = "grey70") +
+  ggplot2::geom_polygon(data = cs2_spp_grid_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                     fill = Occupancy),
+                        alpha = 0.8, color = "grey10") +
+  ggplot2::geom_label(data = cs2_spp_grid_letters_data,
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white"),
+    axis.title = ggplot2::element_blank(),
+    legend.position = "none") +
+  ggplot2::scale_fill_manual(name = "Occupancy",
+                             values = c("absent"="#132B43",
+                                        "present"="#56B1F7")) +
+  ggplot2::coord_cartesian(xlim = buffered_range(cs2_grid_fpolygons$long, 0),
+                           ylim = buffered_range(cs2_grid_fpolygons$lat, 0)) +
+  ggplot2::facet_wrap(~ Title)
+# plot cost data
+p2 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = countries_fpolygons,
+                        ggplot2::aes(x =long, y = lat, group = group),
+                        fill = "grey85", color = "grey70") +
+  ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                     fill = Cost),
+                        alpha = 0.8, color = "grey10") +
+  ggplot2::geom_label(data = cs2_spp_grid_sub_letters_data[1, ],
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill="grey20"),
+    strip.text = ggplot2::element_text(color="white"),
+    axis.title = ggplot2::element_blank(),
+    legend.background = ggplot2::element_rect(color = "black", fill = "white"),
+    legend.position = c(0.25, 0.58),
+    legend.key.height = ggplot2::unit(2.5, "mm"),
+    legend.key.width = ggplot2::unit(3.25, "mm"),
+    legend.text = ggplot2::element_text(size = 6.0),
+    legend.title = ggplot2::element_blank(),
+    legend.direction = "horizontal",
+    legend.justification = c(0,1),
+    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
+                                                                  "mm"))) +
+  ggplot2::coord_cartesian(
+    xlim = buffered_range(cs2_grid_sub_fpolygons$long, 0),
+    ylim = buffered_range(cs2_grid_sub_fpolygons$lat, 0)) +
+    ggplot2::scale_fill_continuous(name = ggplot2::element_blank(),
+      breaks = pretty(cs2_spp_grid_sub_fpolygons$Cost, n = 3, min.n = 2)) +
+  ggplot2::facet_wrap(~ Title1)
+# plot nmds axis 1
+p3 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = countries_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill = "grey85", color = "grey70") +
+  ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                     fill = nmds1),
+                         alpha=0.8, color="grey10") +
+  ggplot2::geom_label(data = cs2_spp_grid_sub_letters_data[2, ],
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill="grey20"),
+    strip.text = ggplot2::element_text(color="white"),
+    axis.title = ggplot2::element_blank(),
+    legend.background = ggplot2::element_rect(color="black", fill="white"),
+    legend.position = c(0.25, 0.58),
+    legend.key.height = ggplot2::unit(2.5, "mm"),
+    legend.key.width = ggplot2::unit(3.25, "mm"),
+    legend.text = ggplot2::element_text(size=6.0),
+    legend.title = ggplot2::element_blank(),
+    legend.direction = "horizontal",
+    legend.justification = c(0,1),
+    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
+                                                                  "mm"))) +
+  ggplot2::coord_cartesian(
+      xlim = buffered_range(cs2_grid_sub_fpolygons$long, 0),
+      ylim = buffered_range(cs2_grid_sub_fpolygons$lat, 0)) +
+  ggplot2::scale_fill_gradient2(name = ggplot2::element_blank(),
+                                low = "darkblue", mid = "white",
+                                high = "darkred", breaks = c(0, 0.3)) +
+  ggplot2::facet_wrap(~ Title2)
+# plot nmds axis 2
+p4 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = countries_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill="grey85", color="grey70") +
+  ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                     fill = nmds2),
+                         alpha=0.8, color="grey10") +
+  ggplot2::geom_label(data = cs2_spp_grid_sub_letters_data[3,],
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill="grey20"),
+    strip.text = ggplot2::element_text(color="white"),
+    axis.title = ggplot2::element_blank(),
+    legend.background = ggplot2::element_rect(color="black", fill="white"),
+    legend.position = c(0.25, 0.58),
+    legend.key.height = ggplot2::unit(2.5, "mm"),
+    legend.key.width = ggplot2::unit(3.25, "mm"),
+    legend.text = ggplot2::element_text(size=6.0),
+    legend.title = ggplot2::element_blank(),
+    legend.direction = "horizontal",
+    legend.justification = c(0,1),
+    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
+                                                                   "mm"))) +
+  ggplot2::coord_cartesian(
+    xlim = buffered_range(cs2_grid_sub_fpolygons$long, 0),
+    ylim = buffered_range(cs2_grid_sub_fpolygons$lat, 0)) +
+  ggplot2::scale_fill_gradient2(name = ggplot2::element_blank(),
+                                low = "darkblue", mid = "white",
+                                high = "darkred",
+                                breaks = pretty(
+                                  cs2_spp_grid_sub_fpolygons$nmds2, n = 2,
+                                  min.n = 2)) +
+  ggplot2::facet_wrap(~ Title3)
 # render plots
-grid.arrange(p1, p2, ncol=2)
+gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
 ```
 
-```{r, echo=FALSE, fig.height=3.8, fig.width=column.width, fig.cap='Distribution of the species used in the first case study.
-Each panel shows the geographic range for a different species.
-Squares denote planning units.'}
-# prepare data
-cs1.pus.occ.FPLY <- ldply(
-	seq_along(cs1.params.LST[[MODE]]$common.names),
-	function(i) {
-		curr.DF <- cs1.pus.FPLY
-		curr.DF$species <- cs1.params.LST[[MODE]]$common.names[i]
-		curr.DF$Occupancy <- 'absent'
-		curr.DF$Occupancy[which(curr.DF$id %in% filter(cs1.ru@data@pu.species.probabilities, species==i)$pu)] <- 'present'
-		return(curr.DF)
-	}
-)
-
-cs1.pus.occ.FPLY$species <- factor(as.character(cs1.pus.occ.FPLY$species), levels=cs1.params.LST[[MODE]]$common.names)
-# make letter data.frame
-cs1.pus.occ.letters.DF <- data.frame(
-	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.08)),
-	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.95)),
-	species=unique(cs1.pus.occ.FPLY$species),
-	letter=paste0('(', letters[seq_along(unique(cs1.pus.occ.FPLY$species))], ')')
-)
-# make plots
-ggplot() +
-	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70', size=0.05) +
-	geom_polygon(data=cs1.pus.occ.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), color='black', size=0.05) +
-	geom_label(data=cs1.pus.occ.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=7),
-		axis.title=element_blank(), legend.position='none'
-	) +
-	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
-  coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
-	facet_wrap(~ species, ncol=2)
-```
-
-```{r, echo=FALSE, fig.height=1.9, fig.width=column.width, fig.cap=paste0('Prioritizations for the first case study.
-See Fig.
-3 caption for conventions.
-Panel (a) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets.
-Panel (b) shows the planning units selected when using ',cs1.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs1.params.LST[[MODE]]$space.target*100,' % niche targets.')}
-# prepare data
-cs1.pus.sel.FPLY <- ldply(
-	names(cs1.prioritisations),
-	function(x) {
-		curr.DF <- cs1.pus.FPLY
-		curr.DF$Prioritisation <- x
-		curr.DF$Solution <- 'discarded'
-		curr.DF$Solution[curr.DF$id %in% which(as.logical(selections(cs1.prioritisations[[x]])))] <- 'selected'
-		return(curr.DF)
-	}
-)
-cs1.pus.sel.FPLY$Prioritisation <- factor(as.character(cs1.pus.sel.FPLY$Prioritisation), levels=names(cs1.prioritisations))
-# make letter data.frame
-cs1.pus.sel.letters.DF <- data.frame(
-	x=unname(cs1.pus@bbox[1,1] + (diff(cs1.pus@bbox[1,])*0.08)),
-	y=unname(cs1.pus@bbox[2,1] + (diff(cs1.pus@bbox[2,])*0.95)),
-	Prioritisation=names(cs1.prioritisations),
-	letter=paste0('(', letters[1:2], ')')
-)
-# make plots
-ggplot() +
-	geom_polygon(data=aus.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70', size=0.05) +
-	geom_polygon(data=cs1.pus.sel.FPLY, aes(x=long, y=lat, group=group, fill=Solution), color='black', size=0.05) +
-	geom_label(data=cs1.pus.sel.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=8),
-		axis.title=element_blank(), legend.position='none'
-	) +
-	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b')) +
-	coord_cartesian(xlim=buffered.range(cs1.pus@bbox[1,], 0.025), ylim=buffered.range(cs1.pus@bbox[2,], 0.025), expand=FALSE) +
-	facet_wrap(~ Prioritisation, ncol=2)
-```
-
-```{r, echo=FALSE, fig.height=2.2, fig.width=column.width, fig.cap=paste0('Data used for the second case study.
-Squares denote planning units.
-Panel (a) shows all grid cells surveyed by the IntraBioDiv project.
-Grid cells occupied by ',tolower(cs2.params.LST[[MODE]]$common.name),' are shown in bright blue.
-The subsequent panels contain only show occupied grid cells.
-Panel (b) shows the opportunity cost of each planning unit (estimated as the total human population density).
-Panels (c--d) show the spatial distribution of the ordinations describing genetic variation.
-These values describe the typical genetic characteristics of individuals in each planning unit.
-Planning units with similar values/colors contain individuals with similar loci polymorphisms.')}
-# prepare full grid data
-cs2.spp.grid.FPLY <- cs2.grid.FPLY
-cs2.spp.grid.FPLY$Title <- 'Occupancy'
-cs2.spp.grid.FPLY$Occupancy <- 'absent'
-cs2.spp.grid.FPLY$Occupancy[which(cs2.grid.PLY@data[[cs2.params.LST[[MODE]]$species.name]][as.numeric(cs2.spp.grid.FPLY$id)]==1)] <- 'present'
-
-cs2.spp.grid.letters.DF <- data.frame(
-	x=unname(cs2.grid.PLY@bbox[1,1] + (diff(cs2.grid.PLY@bbox[1,])*0.07)),
-	y=unname(cs2.grid.PLY@bbox[2,1] + (diff(cs2.grid.PLY@bbox[2,])*0.875)),
-	letter=('(a)')
-)
-
-# prepare subset grid data
-cs2.spp.grid.sub.FPLY <- cs2.grid.sub.FPLY
-cs2.spp.grid.sub.FPLY$nmds1 <- NA
-cs2.spp.grid.sub.FPLY$nmds2 <- NA
-cs2.nmds1 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d1')]]
-cs2.nmds2 <- cs2.grid.sub.DF[[paste0(cs2.params.LST[[MODE]]$species.name,'_genetic_d2')]]
-
-curr.pos <- match(as.numeric(cs2.spp.grid.sub.FPLY$id), cs2.grid.sub.DF$id)
-cs2.spp.grid.sub.FPLY$nmds1 <- cs2.nmds1[curr.pos]
-cs2.spp.grid.sub.FPLY$nmds2 <- cs2.nmds2[curr.pos]
-
-cs2.spp.grid.sub.FPLY$Cost <- costs.MTX[match(cs2.grid.sub.PLY$id[as.numeric(cs2.spp.grid.sub.FPLY$id)], costs.MTX[,1]),2]
-
-cs2.spp.grid.sub.FPLY$Title1 <- 'Opportunity cost'
-cs2.spp.grid.sub.FPLY$Title2 <- 'Genetic NMDS 1'
-cs2.spp.grid.sub.FPLY$Title3 <- 'Genetic NMDS 2'
-
-cs2.spp.grid.sub.letters.DF <- data.frame(
-	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
-	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.89)),
-	letter=c('(b)','(c)','(d)')
-)
-
-# plot species data
-p1 <- ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs2.spp.grid.FPLY, aes(x=long, y=lat, group=group, fill=Occupancy), alpha=0.8, color='grey10') +
-	geom_label(data=cs2.spp.grid.letters.DF, aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.position='none'
-	) +
-	scale_fill_manual(name='Occupancy', values=c('absent'='#132B43', 'present'='#56B1F7')) +
-	coord_cartesian(xlim=buffered.range(cs2.grid.FPLY$long, 0), ylim=buffered.range(cs2.grid.FPLY$lat, 0)) +
-	facet_wrap(~ Title)
-# plot cost data
-p2 <- ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Cost), alpha=0.8, color='grey10') +
-	geom_label(data=cs2.spp.grid.sub.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-		legend.position=c(0.25, 0.58), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'), legend.text=element_text(size=6.0),
-		legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
-	) +
-	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
-	scale_fill_continuous(name=element_blank(), breaks=pretty(cs2.spp.grid.sub.FPLY$Cost, n=3, min.n=2)) +
-	facet_wrap(~ Title1)
-# plot nmds axis 1
-p3 <- ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
-	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds1), alpha=0.8, color='grey10') +
-	geom_label(data=cs2.spp.grid.sub.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-    axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-    legend.position=c(0.25, 0.58), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'), legend.text=element_text(size=6.0),
-    legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
-	) + 
-	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
-	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred', breaks=c(0, 0.3)) +
-	facet_wrap(~ Title2)
-# plot nmds axis 2
-p4 <- ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),fill='grey85', color='grey70') +
-	geom_polygon(data=cs2.spp.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=nmds2), alpha=0.8, color='grey10') +
-	geom_label(data=cs2.spp.grid.sub.letters.DF[3,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(
-		axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white'),
-    axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-    legend.position=c(0.25, 0.58), legend.key.height=unit(2.5, 'mm'), legend.key.width=unit(3.25, 'mm'), legend.text=element_text(size=6.0),
-    legend.title=element_blank(), legend.direction='horizontal', legend.justification=c(0,1), legend.background=element_rect(size=unit(0.25, 'mm'))
-	) + 
-	coord_cartesian(xlim=buffered.range(cs2.grid.sub.FPLY$long, 0), ylim=buffered.range(cs2.grid.sub.FPLY$lat, 0)) +
-	scale_fill_gradient2(name=element_blank(), low='darkblue', mid='white', high='darkred', breaks=pretty(cs2.spp.grid.sub.FPLY$nmds2, n=2, min.n=2)) +
-	facet_wrap(~ Title3)
-# render plots
-grid.arrange(p1, p2, p3, p4, ncol=2)
-```
-
-```{r, echo=FALSE, fig.height=2.5, fig.width=column.width, fig.cap=paste0('Prioritizations for the second case study.
+```{r, echo = FALSE, fig.height = 2.5, fig.width = column_width, fig.cap = paste0("Prioritizations for the second case study.
 Panels (a--b) show prioritizations generated using different parameters.
-See Fig.
-3 caption for conventions.
-Panel (a) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets.
-Panel (b) shows the planning units selected when using ',cs2.params.LST[[MODE]]$amount.target*100,' % amount targets and ',cs2.params.LST[[MODE]]$genetic.target*100,' % genetic targets.
+Polygons denote planning units.
+Dark green planning units were selected for protection.
+Panel (a) shows the planning units selected when using ",cs2_parameters$amount_target * 100," % amount targets.
+Panel (b) shows the planning units selected when using ",cs2_parameters$amount.target*100," % amount targets and ",cs2_parameters$genetic_target * 100," % genetic targets.
 Panels (c--d) show the solutions in the genetic space.
@@
 The coordinates of the points represent the typical genetic characteristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data).
-Planning units associated with points that are closer together contain individuals with more similar genetic characteristics than planning units that are further apart.')}
+Planning units associated with points that are closer together contain individuals with more similar genetic characteristics than planning units that are further apart.")}
 # prepare spatial data
-cs2.sel.grid.sub.FPLY <- cs2.grid.sub.FPLY
-cs2.sel.grid.sub.FPLY$Title1 <-'Amount target'
-cs2.sel.grid.sub.FPLY$Title2 <- 'Amount & genetic target'
+cs2_sel_grid_sub_fpolygons <- cs2_grid_sub_fpolygons
+cs2_sel_grid_sub_fpolygons$Title1 <-"Amount target"
+cs2_sel_grid_sub_fpolygons$Title2 <- "Amount & genetic target"
 
-cs2.sel.grid.sub.FPLY$Amount.selection <- 'discarded'
-cs2.sel.grid.sub.FPLY$Amount.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'
+cs2_sel_grid_sub_fpolygons$Amount_selection <- "discarded"
+sel_1 <- as.logical(raptr::selections(cs2_prioritisations[[1]]))
+rows_1 <- cs2_sel_grid_sub_fpolygons$id %in% which(sel_1)
+cs2_sel_grid_sub_fpolygons$Amount_selection[rows_1] <- "selected"
 
-cs2.sel.grid.sub.FPLY$Amount.genetic.selection <- 'discarded'
-cs2.sel.grid.sub.FPLY$Amount.genetic.selection[cs2.sel.grid.sub.FPLY$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'
+cs2_sel_grid_sub_fpolygons$Amount_genetic_selection <- "discarded"
+sel_2 <- as.logical(raptr::selections(cs2_prioritisations[[2]]))
+rows_2 <- cs2_sel_grid_sub_fpolygons$id %in% which(sel_2)
+cs2_sel_grid_sub_fpolygons$Amount_genetic_selection[rows_2] <- "selected"
 
-cs2.sel.grid.sub.letters.DF <- data.frame(
-	x=unname(cs2.grid.sub.PLY@bbox[1,1] + (diff(cs2.grid.sub.PLY@bbox[1,])*0.06)),
-	y=unname(cs2.grid.sub.PLY@bbox[2,1] + (diff(cs2.grid.sub.PLY@bbox[2,])*0.9)),
-	letter=c('(a)','(b)')
-)
+cs2_sel_grid_sub_letters_data <- data.frame(
+  x = unname(cs2_grid_sub_polygons@bbox[1, 1] +
+      (diff(cs2_grid_sub_polygons@bbox[1, ]) * 0.06)),
+  y = unname(cs2_grid_sub_polygons@bbox[2,1] +
+      (diff(cs2_grid_sub_polygons@bbox[2, ]) * 0.9)),
+  letter = c("(a)", "(b)"))
 
 # prepare genetic space plots
-cs2.sel.space.DF <- data.frame(cs2.genetic.AS@spaces[[1]]@planning.unit.points@coords)
-names(cs2.sel.space.DF) <- c('d1', 'd2')
-cs2.sel.space.DF$id <- cs2.genetic.AS@spaces[[1]]@planning.unit.points@ids
+cs2_sel_space_data <- cs2_genetic_attribute_spaces %>%
+                      slot("spaces") %>%
+                      `[[`(1) %>%
+                      slot("planning.unit.points") %>%
+                      slot("coords") %>%
+                      data.frame() %>%
+                      magrittr::set_names(c("d1", "d2")) %>%
+                      dplyr::mutate(id = cs2_genetic_attribute_spaces %>%
+                                         slot("spaces") %>%
+                                         `[[`(1) %>%
+                                         slot("planning.unit.points") %>%
+                                         slot("ids"))
 
-cs2.sel.space.DF$Amount.selection <- 'discarded'
-cs2.sel.space.DF$Amount.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[1]])))] <- 'selected'
+cs2_sel_space_data$Amount_selection <- "discarded"
+sel_1 <- as.logical(raptr::selections(cs2_prioritisations[[1]]))
+rows_1 <- cs2_sel_space_data$id %in% which(sel_1)
+cs2_sel_space_data$Amount_selection[rows_1] <- "selected"
 
-cs2.sel.space.DF$Amount.genetic.selection <- 'discarded'
-cs2.sel.space.DF$Amount.genetic.selection[cs2.sel.space.DF$id %in% which(as.logical(selections(cs2.prioritisations[[2]])))] <- 'selected'
+cs2_sel_space_data$Amount_genetic_selection <- "discarded"
+sel_2 <- as.logical(raptr::selections(cs2_prioritisations[[2]]))
+rows_2 <- cs2_sel_space_data$id %in% which(sel_2)
+cs2_sel_space_data$Amount_genetic_selection[rows_2] <- "selected"
 
-cs2.sel.space.letters.DF <- data.frame(
-	x=min(cs2.sel.space.DF$d1) + (diff(range(cs2.sel.space.DF$d1))*0.85),
-	y=min(cs2.sel.space.DF$d2) + (diff(range(cs2.sel.space.DF$d2))*0.95),
-	letter=c('(c)', '(d)')
-)
+cs2_sel_space_letters_data <- data.frame(
+  x = min(cs2_sel_space_data$d1) +
+      (diff(range(cs2_sel_space_data$d1)) * 0.85),
+  y = min(cs2_sel_space_data$d2) +
+      (diff(range(cs2_sel_space_data$d2)) * 0.95),
+  letter = c("(c)", "(d)"))
 
 # make plots
-p1 <- ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.selection), color='black') +
-	geom_label(data=cs2.sel.grid.sub.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=8),
-		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'),
-		axis.title=element_blank(), legend.position='none'
-	) +
-	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
-	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
-	facet_wrap(~ Title1)
-p2 <- ggplot() +
-	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group), fill='grey85', color='grey70') +
-	geom_polygon(data=cs2.sel.grid.sub.FPLY, aes(x=long, y=lat, group=group, fill=Amount.genetic.selection), color='black') +
-	geom_label(data=cs2.sel.grid.sub.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() +
-	theme(axis.ticks=element_blank(), axis.text=element_blank(), axis.line=element_blank(), plot.margin = unit(c(0, 0, 0, 0), "lines"),
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white', size=8),
-		axis.title=element_blank(), legend.background=element_rect(color='black', fill='white'), legend.position='none'
-	) +
-	scale_fill_manual(name='Solution', values=c('discarded'='#f7fcfd', 'selected'='#00441b', 'uninhabited'='grey65')) +
-	coord_cartesian(xlim=cs2.grid.sub.PLY@bbox[1,], ylim=cs2.grid.sub.PLY@bbox[2,]) +
-	facet_wrap(~ Title2)
-p3 <- ggplot() +
-	geom_point(data=arrange(cs2.sel.space.DF, Amount.selection), aes(x=d1, y=d2, fill=Amount.selection), pch=21, size=3, color='black') +
-	geom_label(data=cs2.sel.space.letters.DF[1,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() + xlab('NMDS 1') + ylab('NMDS 2') +
-	scale_x_continuous(expand=c(0.08, 0)) + scale_y_continuous(expand=c(0.08, 0)) + 
-	theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line(),
-    axis.text=element_text(size=8), axis.title=element_text(size=8), plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
-	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
-p4 <- ggplot() +
-	geom_point(data=arrange(cs2.sel.space.DF, Amount.genetic.selection), aes(x=d1, y=d2, fill=Amount.genetic.selection), pch=21, size=3, color='black') +
-	geom_label(data=cs2.sel.space.letters.DF[2,], aes(label=letter, x=x, y=y), color='black', fill=alpha('white', 0.8), size=3) +
-	theme_classic() + xlab('NMDS 1') + ylab('NMDS 2') +
-	scale_x_continuous(expand=c(0.08, 0)) + scale_y_continuous(expand=c(0.08, 0)) + 
-  theme(legend.position='none', axis.line.x=element_line(), axis.line.y=element_line(),
-    axis.text=element_text(size=8), axis.title=element_text(size=8), plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
-	scale_fill_manual(name='Solution', values=c('discarded'='transparent', 'selected'='#00441b'))
+p1 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = countries_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group),
+                        fill = "grey85", color = "grey70") +
+  ggplot2::geom_polygon(data = cs2_sel_grid_sub_fpolygons,
+                        ggplot2::aes(x = long, y = lat, group = group,
+                                     fill = Amount_selection), color="black") +
+  ggplot2::geom_label(data = cs2_sel_grid_sub_letters_data[1, ],
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white", size = 8),
+    axis.title = ggplot2::element_blank(),
+    legend.background = ggplot2::element_rect(color = "black", fill = "white"),
+    axis.title = ggplot2::element_blank(),
+    legend.position="none") +
+  ggplot2::scale_fill_manual(name = "Solution",
+                             values = c("discarded" = "#f7fcfd",
+                                        "selected" = "#00441b",
+                                        "uninhabited" = "grey65")) +
+  ggplot2::coord_cartesian(
+      xlim = cs2_grid_sub_polygons@bbox[1, ],
+      ylim = cs2_grid_sub_polygons@bbox[2, ]) +
+  ggplot2::facet_wrap(~ Title1)
+p2 <- ggplot2::ggplot() +
+  ggplot2::geom_polygon(data = countries_fpolygons,
+                        ggplot2::aes(x=long, y=lat, group=group),
+                        fill="grey85", color="grey70") +
+  ggplot2::geom_polygon(data = cs2_sel_grid_sub_fpolygons,
+                        ggplot2::aes(x=long, y=lat, group=group,
+                                     fill = Amount_genetic_selection),
+                                     color="black") +
+  ggplot2::geom_label(data = cs2_sel_grid_sub_letters_data[2, ],
+                      ggplot2::aes(label = letter, x =x , y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(
+    axis.ticks = ggplot2::element_blank(),
+    axis.text = ggplot2::element_blank(),
+    axis.line = ggplot2::element_blank(),
+    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white", size = 8),
+    axis.title = ggplot2::element_blank(),
+    legend.background = ggplot2::element_rect(color = "black", fill = "white"),
+    legend.position = "none") +
+  ggplot2::scale_fill_manual(name = "Solution",
+                             values = c("discarded" = "#f7fcfd",
+                                        "selected" = "#00441b",
+                                        "uninhabited" = "grey65")) +
+  ggplot2::coord_cartesian(
+    xlim = cs2_grid_sub_polygons@bbox[1, ],
+    ylim = cs2_grid_sub_polygons@bbox[2, ]) +
+  ggplot2::facet_wrap(~ Title2)
+p3 <- ggplot2::ggplot() +
+  ggplot2::geom_point(data = dplyr::arrange(cs2_sel_space_data,
+                                            Amount_selection),
+                      ggplot2::aes(x = d1, y = d2, fill = Amount_selection),
+                      pch = 21, size = 3, color = "black") +
+  ggplot2::geom_label(data = cs2_sel_space_letters_data[1,],
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::xlab("NMDS 1") +
+  ggplot2::ylab("NMDS 2") +
+  ggplot2::scale_x_continuous(expand = c(0.08, 0)) +
+  ggplot2::scale_y_continuous(expand = c(0.08, 0)) +
+  ggplot2::theme(
+      legend.position="none",
+      axis.line.x = ggplot2::element_line(),
+      axis.line.y = ggplot2::element_line(),
+      axis.text = ggplot2::element_text(size = 8),
+      axis.title = ggplot2::element_text(size = 8),
+      plot.margin = ggplot2::unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
+  ggplot2::scale_fill_manual(name = "Solution",
+                             values = c("discarded" = "transparent",
+                                        "selected" = "#00441b"))
+p4 <- ggplot2::ggplot() +
+  ggplot2::geom_point(data = dplyr::arrange(cs2_sel_space_data,
+                                            Amount_genetic_selection),
+                      ggplot2::aes(x = d1, y = d2,
+                                   fill = Amount_genetic_selection),
+                       pch = 21, size = 3, color = "black") +
+  ggplot2::geom_label(data = cs2_sel_space_letters_data[2,],
+                      ggplot2::aes(label = letter, x = x, y = y),
+                      color = "black", fill = scales::alpha("white", 0.8),
+                      size = 3) +
+  ggplot2::theme_classic() +
+  ggplot2::xlab("NMDS 1") +
+  ggplot2::ylab("NMDS 2") +
+  ggplot2::scale_x_continuous(expand = c(0.08, 0)) +
+  ggplot2::scale_y_continuous(expand = c(0.08, 0)) +
+  ggplot2::theme(
+    legend.position="none",
+    axis.line.x = ggplot2::element_line(),
+    axis.line.y = ggplot2::element_line(),
+    axis.text = ggplot2::element_text(size = 8),
+    axis.title = ggplot2::element_text(size = 8),
+    plot.margin = ggplot2::unit(c(0.5, 0.5, 0.5, 0.5), "pt")) +
+  ggplot2::scale_fill_manual(name = "Solution",
+                             values = c("discarded" = "transparent",
+                                        "selected" = "#00441b"))
 # render plots
-grid.arrange(p1,p2,p3,p4,ncol=2)
+gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
 ```

### 2017-04-06 code/rmarkdown/supporting-information.Rmd
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references

@@
 # set wd
-knitr::opts_knit$set(root.dir=normalizePath('../..'))
+knitr::opts_knit$set(root.dir = normalizePath("../.."))
 ```
@@
 
-```{r, include=FALSE}
+```{r, include = FALSE}
 ## initialization
-# load .rda file
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-try(session::restore.session('data/final/results.rda'))
+# load session
+try(session::restore.session("data/final/results.rda"))
 
-# set options
-options(error=NULL)
-
-## prelimary processing
+## preliminary processing
 # redefine file paths for rasters
-for (i in seq_len(nlayers(cs1.bioclim.RAST)))
-	cs1.bioclim.RAST@layers[[i]]@file@name  <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])
+for (i in seq_len(raster::nlayers(cs1_bioclim_raster)))
+  cs1_bioclim_raster@layers[[i]]@file@name  <- paste0(getwd(),
+    strsplit(cs1_bioclim_raster@layers[[i]]@file@name,
+             "raptr-manuscript")[[1]][[2]])
 
@@
 
-The reliable formulation (RRAP) is a multi-objective optimisation problem.
+The reliable formulation (RRAP) is a multi-objective optimization problem.
 
@@
 
-We generated simulated conservation planning data sets using varying number of features `r paste(benchmark.params.LST$number.of.features, collapse=',')` and planning units (paste(benchmark.params.LST$number.of.planning.units, collapse=',')).
+We generated simulated conservation planning data sets using varying number of features (`r paste(benchmark_parameters$number_of_features, collapse = ",")`) and planning units (`r paste(benchmark_parameters$number_of_planning_units, collapse = ",")`).
 For a given number of planning units and features, the data was simulated using the following steps.
@@
 All planning unit costs were set to one.
-The spatial distribution of features were generated using binary, spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @rXXX].
+The spatial distribution of features were generated using binary, spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @r494; @r493].
 If features were associated with less than two occupied planning units, these features were set to occupy two randomly selected planning units.
-All occupied planning units were set to have a probability of occupancy of `r benchmark.params.LST[[MODE]]$occupancy.probability`.
+All occupied planning units were set to have a probability of occupancy of `r benchmark_parameters$occupancy_probability`.
 Similar to the simulated data sets used in the main text, a geographic attribute space was used for each feature.
 The features' demand points were set as the centroids of the planning units that they occupied, and were weighted by the units' probability of occupancy.
-All problems used amount-based targets of `r benchmark.params.LST[[MODE]]$amount.target*100` % and space-based targets of `r benchmark.params.LST[[MODE]]$space.target*100`
+All problems used amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100`
 
-We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifer, and using a boundary length modifier of `r max(benchmark.params.LST[[MODE]]$blm)`.
-To assess variation in execution times, we generated and solved `r benchmark.params.LST[[MODE]]$replicates` random data sets per combination of number of features and planning units.
-All conservation planning problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) on a system with 64 GB RAM and an Intel(R) Xeon(R) CPU (E5-2630 v3; 2.40 GHz).
+We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifier, and using a boundary length modifier of `r max(benchmark_parameters$blm)`.
+To assess variation in execution times, we generated and solved `r benchmark_parameters$replicates` random data sets per combination of number of features and planning units.
+All conservation planning problems were solved to within `r general_parameters$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) on a system with 64 GB RAM and an Intel(R) Xeon (R) CPU (E5-2630 v3; 2.40 GHz).
 All problems were solved using a single thread.
@@
 
-The benchmark analysis revealed that moderately-sized conservation problems can be solved in a realtively short period of time (Figure 1).
+The benchmark analysis revealed that moderately-sized conservation problems can be solved in a relatively short period of time (Figure S2).
 As predicted, problems were solved much quicker when using the unreliable formulation than when using the reliable formulation.
-Additionally, the solve times varied substiantially for different data sets with the same proporties (ie.
+Additionally, the solve times varied substantially for different data sets with the same properties (ie.
 number of features and planning units).
 This result suggests that it may difficult to estimate exactly how long a given problem will take to solve.
-
-```{r, echo=FALSE, message=FALSE, fig.height=5.0, fig.width=4.5, fig.cap='Benchmark analysis.
-Boxplots show the amount of time required to solve a problem under a given set of parameters.
-Note that times are shown on a log_{10} scale.'}
-# prepare data
-plot.DF <- benchmark.DF %>%
-  mutate(blm = paste0('BLM = ',blm)) %>% 
-  mutate(number.features=factor(number.features)) %>%
-  mutate(formulation = paste0(formulation, ' formulation')) %>%
-  mutate(number.planning.units=factor(number.planning.units))
-# make plot
-ggplot(plot.DF, aes(x=number.planning.units, y=time, color=number.features)) +
-  geom_boxplot() +
-  facet_grid(blm ~ formulation ) +
-  scale_y_continuous(trans='log10') +
-  xlab('Number of planning units') + ylab('Solve time (s)') +
-  labs(color='Number of features') +
-  theme(axis.line.x=element_line(), axis.line.y=element_line(),
-    strip.background = element_rect(fill='grey20'),
-    strip.text = element_text(color='white', size=12),
-    legend.position='bottom')
-```
 
@@
 Future research is needed to explore these methods and understand how the impact solution quality.
-Additionally, future research could involve implementing more advanced methods to solve the optimization problem [eg.
+Additionally, future research could involve implementing more advanced methods to solve the optimization problems [eg.
 as described in @r16; @r15].
 
+\clearpage
+
 # Figures
-```{r, echo=FALSE, fig.height=8, fig.width=8.5, fig.cap='Attribute spaces used in the first case-study.
+```{r, echo = FALSE, fig.height = 8, fig.width = 8.5, fig.cap = "Attribute spaces used in the first case-study.
 Each panel shows a the distribution of a solution in environmental space and how it samples the realized niche for a different species.
@@
 Hexagons show the distribution of demand points.
-The color of each heaxgon denotes the weighted frequency of demand points inside it.
-Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution.'}
+The color of each hexagon denotes the weighted frequency of demand points inside it.
+Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution."}
 ## prepare data
-cs1.pu.niche.DF <- ddply(
-	cs1.spp.DF,
-	c('Species', 'Prioritisation'),
+cs1_pu_niche_data <- plyr::ddply(cs1_spp_results,
+  c("Species", "Prioritisation"),
   function(x) {
     # init
-		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
-		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
-		curr.sel.pos <- which(curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@ids %in% which(as.logical(selections(curr.prioritisation))))
+    curr_prioritisation <- cs1_prioritisations[[as.character(x$Prioritisation)]]
+    curr_spp_pos <- match(x$Species[1], curr_prioritisation@data@species[[1]])
+    curr_ids <- curr_prioritisation %>%
+                slot("data") %>%
+                slot("attribute.spaces") %>%
+                `[[`(1) %>%
+                slot("spaces") %>%
+                `[[`(curr_spp_pos) %>%
+                slot("planning.unit.points") %>%
+                slot("ids")
+    curr_sel_pos <- which(curr_ids %in% which(as.logical(
+                          raptr::selections(curr_prioritisation))))
     # extract planning units
-		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@planning.unit.points@coords %>% as.data.frame()
-		names(curr.DF) <- c('Niche1', 'Niche2')
-		curr.DF$Species <-x$Species[[1]]
-		curr.DF$Prioritisation <-x$Prioritisation[[1]]
-		curr.DF$Solution <- 'discarded'
-		curr.DF$Solution[curr.sel.pos] <- 'selected'
+    curr_data <- curr_prioritisation %>%
+                slot("data") %>%
+                slot("attribute.spaces") %>%
+                `[[`(1) %>%
+                slot("spaces") %>%
+                `[[`(curr_spp_pos) %>%
+                slot("planning.unit.points") %>%
+                slot("coords") %>%
+                as.data.frame() %>%
+                magrittr::set_names(c("Niche1", "Niche2")) %>%
+                dplyr::mutate(Species = x[["Species"]][[1]],
+                              Prioritisation = x[["Prioritisation"]][[1]],
+                              Solution <- "discarded")
+    curr_data$Solution[curr_sel_pos] <- "selected"
     # return data
-		return(curr.DF)
-	}
-)
-cs1.pu.niche.DF <- filter(cs1.pu.niche.DF, Solution=='selected')
+    return(curr_data)
+})
+cs1_pu_niche_data %<>% dplyr::filter(Solution == "selected")
 
-cs1.dp.niche.DF <- ddply(
-	cs1.spp.DF,
-	c('Species', 'Prioritisation'),
+cs1_dp_niche_data <- plyr::ddply(
+  cs1_spp_results,
+  c("Species", "Prioritisation"),
   function(x) {
     # init
-		curr.prioritisation <- cs1.prioritisations[[as.character(x$Prioritisation)]]
-		curr.spp.pos <- match(x$Species[1], curr.prioritisation@data@species[[1]])
+    curr_prioritisation <- cs1_prioritisations[[as.character(x$Prioritisation)]]
+    curr_spp_pos <- match(x$Species[1], curr_prioritisation@data@species[[1]])
     # extract demand points
-		curr.DF <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@coords %>% as.data.frame()
-		names(curr.DF) <- c('Niche1', 'Niche2')
-		curr.DF$weight <- curr.prioritisation@data@attribute.spaces[[1]]@spaces[[curr.spp.pos]]@demand.points@weights
+    curr_data <- curr_prioritisation %>%
+                  slot("data") %>%
+                  slot("attribute.spaces") %>%
+                  `[[`(1) %>%
+                  slot("spaces") %>%
+                  `[[`(curr_spp_pos) %>%
+                  slot("demand.points") %>%
+                  slot("coords") %>%
+                  as.data.frame() %>%
+                  magrittr::set_names(c("Niche1", "Niche2"))
+    curr_data$weight <- curr_prioritisation %>%
+                        slot("data") %>%
+                        slot("attribute.spaces") %>%
+                        `[[`(1) %>%
+                        slot("spaces") %>%
+                        `[[`(curr_spp_pos) %>%
+                        slot("demand.points") %>%
+                        slot("weights")
     # return data
-		return(curr.DF)
-	}
-)
-cs1.pu.niche.DF$Prioritisation <- factor(as.character(cs1.pu.niche.DF$Prioritisation), levels=names(cs1.prioritisations))
-cs1.dp.niche.DF$Prioritisation <- factor(as.character(cs1.dp.niche.DF$Prioritisation), levels=names(cs1.prioritisations))
+    return(curr_data)
+})
+cs1_pu_niche_data$Prioritisation <- factor(as.character(
+                                      cs1_pu_niche_data$Prioritisation),
+                                      levels = names(cs1_prioritisations))
+cs1_dp_niche_data$Prioritisation <- factor(as.character(
+                                      cs1_dp_niche_data$Prioritisation),
+                                      levels = names(cs1_prioritisations))
 
 # make letter data.frame
-cs1.letters.DF <- expand.grid(
-	Species=levels(cs1.pu.niche.DF$Species),
-	Prioritisation=levels(cs1.pu.niche.DF$Prioritisation)
-) %>% arrange(Species, Prioritisation)
-cs1.letters.DF$letter <- paste0('(', letters[seq_len(nrow(cs1.letters.DF))], ')')
-cs1.letters.DF <- cbind(
-	cs1.letters.DF,
-	rbind(
-		cs1.pu.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2),
-		cs1.dp.niche.DF %>% select(Species, Prioritisation, Niche1, Niche2)
-	) %>% summarise(
-		Niche1=min(Niche1)+(abs(diff(range(Niche1)))*0.01),
-		Niche2=min(Niche2)+(abs(diff(range(Niche2)))*0.95)
-	) %>% data.frame()
-)
+cs1_letters_data <- expand.grid(Species = levels(cs1_pu_niche_data$Species),
+                                Prioritisation = levels(
+                                  cs1_pu_niche_data$Prioritisation)) %>%
+                    dplyr::arrange(Species, Prioritisation) %>%
+                    dplyr::mutate(letter = paste0("(",
+                      letters[seq_len(nrow(.))], ")"))
+
+cs1_letters_data2 <- cs1_pu_niche_data %>%
+                     dplyr::select(Species, Prioritisation, Niche1, Niche2) %>%
+                     rbind(cs1_dp_niche_data %>% dplyr::select(Species,
+                                                               Prioritisation,
+                                                               Niche1,
+                                                               Niche2)) %>%
+                     dplyr::summarise(Niche1 = min(Niche1) +
+                                               (abs(diff(range(Niche1))) *
+                                               0.01),
+                                      Niche2 = min(Niche2) +
+                                               (abs(diff(range(Niche2))) *
+                                               0.95)) %>%
+                     data.frame()
+
+cs1_letters_data %<>% cbind(cs1_letters_data2)
 
 ## make plots
-ggplot() +
-	stat_summary_hex(data=cs1.dp.niche.DF, aes(x=Niche1, y=Niche2, z=weight), bins=15, fun=function(z) {sum(z)}) +
-	geom_point(data=cs1.pu.niche.DF, aes(x=Niche1, y=Niche2), fill='#00441b', pch=21, color='black') +
-	geom_text(data=cs1.letters.DF, aes(x=Niche1, y=Niche2, label=letter), color='black') +
-	theme_classic() +
-	theme(
-		strip.background = element_rect(fill='grey20'),
-		strip.text = element_text(color='white')
-	) +
-	xlab('Niche (axis 1)') + ylab('Niche (axis 2)') +
-	scale_fill_gradientn(name='Total demand\npoint weight', colors=colorRampPalette(brewer.pal(6, 'PuBu')[-1])(100)) +
-	facet_grid(Species ~ Prioritisation)
+ggplot2::ggplot() +
+  ggplot2::stat_summary_hex(data = cs1_dp_niche_data,
+                            ggplot2::aes(x = Niche1, y = Niche2, z = weight),
+                            bins = 15, fun = function(z) sum(z)) +
+  ggplot2::geom_point(data = cs1_pu_niche_data,
+                      ggplot2::aes(x = Niche1, y = Niche2),
+                      fill = "#00441b", pch = 21, color = "black") +
+  ggplot2::geom_text(data = cs1_letters_data,
+                     ggplot2::aes(x = Niche1, y = Niche2, label = letter),
+                     color = "black") +
+  ggplot2::theme_classic() +
+  ggplot2::theme(strip.background = ggplot2::element_rect(fill = "grey20"),
+                 strip.text = ggplot2::element_text(color = "white")) +
+  ggplot2::xlab("Niche (axis 1)") +
+  ggplot2::ylab("Niche (axis 2)") +
+  ggplot2::scale_fill_gradientn(name = "Total demand\npoint weight",
+                                colors = colorRampPalette(
+                                  RColorBrewer::brewer.pal(6,
+                                                          "PuBu")[-1])(100)) +
+  ggplot2::facet_grid(Species ~ Prioritisation)
 ```
+
+```{r, echo = FALSE, message = FALSE, fig.height = 5.0, fig.width = 4.5, fig.cap = "Time required to solve different sized problems using different parameters and formulations.
+Points represent average times and bars show standard errors.
+Time is shown on a log$_{10}$ scale."}
+# prepare data
+plot_data <- benchmark_results %>%
+  dplyr::mutate(blm = paste0("BLM = ", blm)) %>%
+  dplyr::mutate(number_features = factor(number_features)) %>%
+  dplyr::mutate(formulation = paste0(formulation, " formulation")) %>%
+  dplyr::mutate(number_planning_units = factor(number_planning_units))
+# make plot
+ggplot2::ggplot(plot_data,
+                ggplot2::aes(x = number_planning_units, y = time,
+                             color = number_features)) +
+  ggplot2::stat_summary(fun.y = mean, geom = "point",
+                        position = ggplot2::position_dodge(0.95)) +
+  ggplot2::stat_summary(fun.data = ggplot2::mean_se, geom = "errorbar",
+                        position = ggplot2::position_dodge(0.95)) +
+  ggplot2::facet_grid(blm ~ formulation ) +
+  ggplot2::scale_y_continuous(trans = "log10") +
+  ggplot2::xlab("Number of planning units") +
+  ggplot2::ylab("Solve time (s)") +
+  ggplot2::labs(color = "Number of features") +
+  ggplot2::theme(
+    axis.line.x = ggplot2::element_line(),
+    axis.line.y = ggplot2::element_line(),
+    strip.background = ggplot2::element_rect(fill = "grey20"),
+    strip.text = ggplot2::element_text(color = "white"),
+    legend.position = "bottom")
+```
+
+\clearpage
 
@@
 
-```{r, echo=FALSE, results='asis'}
+```{r, echo = FALSE, results = "asis"}
 # make table
-term.CHR <- c(
-	'$F$', 'set of biodiversity features (indexed by $f$)',
-	'$J$', 'set of planning units (indexed by $j$)',
-	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
-	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.',
-	'$S$', 'set of attribute spaces (indexed by $S$)',
-	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
-	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$',
-	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$',
-	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$',
-	'$T_f$', 'amount target for feature $f$',
-	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$',
-	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)',
-	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units'
+terms <- c(
+  "$F$", "set of biodiversity features (indexed by $f$)",
+  "$J$", "set of planning units (indexed by $j$)",
+  "$q_{fj}$", "probability of feature $f$ occupying planning unit $j$",
+  "$e_{jk}$", "shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.",
+  "$S$", "set of attribute spaces (indexed by $S$)",
+  "$I_{fsi}$", "set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$",
+  "$\\lambda_{fsi}$", "set of weights for demand point $i$ for feature $f$ in attribute space $s$",
+  "$d_{fsij}$", "distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$",
+  "$\\delta_{fsi}$", "the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$",
+  "$T_f$", "amount target for feature $f$",
+  "$\\tau_{fs}$", "space-based target for feature $f$ in attribute space $s$",
+  "$X_j$", "binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)",
+  "$Y_{fsij}$", "binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determining the amount of the attribute space sampled by the selected planning units"
 )
-term.DF <- data.frame(
-	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
-	Description=term.CHR[seq(2, length(term.CHR), 2)]
+term_data <- data.frame(
+  Symbol = terms[seq(1, length(terms), 2)],
+  Description=terms[seq(2, length(terms), 2)]
 )
-term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')
+term_data$Description <- paste0("\\parbox[t][][t]{15cm}{",
+                                term_data$Description, "}")
 
 # render table
-curr.file <- tempfile(fileext='.tex')
-tmp <- latex(
-	term.DF,
-	file=curr.file, 
-	digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
-	colheads=c('Symbol', 'Description'), col.just=c('c', 'l'),
-	caption='Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritization (URAP) problem.'
-)
-readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
-	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
-	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
+curr_file <- tempfile(fileext = ".tex")
+tmp <- Hmisc::latex(term_data,
+                    file = curr_file,
+                    digits = 2, rowname = NULL, first.hline.double = FALSE,
+                    booktabs = TRUE, here = TRUE,
+                    colheads = c("Symbol", "Description"),
+                    col.just = c("c", "l"),
+                    caption = paste0("Symbols and descriptions of terms used ",
+                                     "in the formulation of the unreliable ",
+                                     "representative and adequate ",
+                                     "prioritization (URAP) problem."))
+curr_file %>%
+  readLines() %>%
+  `[`(-1) %>%
+  paste0(collapse = "\n") %>%
+  gsub(pattern = "toprule", replacement = "toprule[1pt]", fixed = TRUE) %>%
+  gsub(pattern = "bottomrule", replacement = "bottomrule[1pt]",
+       fixed = TRUE) %>%
   cat()
@@
 
+\clearpage
+
 ## References

### 2017-04-06 code/rmarkdown/text.Rmd
update benchmark anlysis
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references

@@
 
-```{r, include=FALSE}
+```{r, include = FALSE}
 # set wd
-knitr::opts_knit$set(root.dir=normalizePath('../..'))
+knitr::opts_knit$set(root.dir = normalizePath("../.."))
 ```
 
-```{r, include=FALSE}
+```{r, include = FALSE}
 ## initialization
-# load .rda file
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/final/results.rda')
+# load session
+session::restore.session("data/final/results.rda")
 
 ## preliminary processing
-for (i in seq_len(nlayers(cs1.bioclim.RAST)))
-  cs1.bioclim.RAST@layers[[i]]@file@name <- paste0(getwd(), strsplit(cs1.bioclim.RAST@layers[[i]]@file@name, 'raptr-manuscript')[[1]][[2]])
+for (i in seq_len(nlayers(cs1_bioclim_raster)))
+  cs1_bioclim_raster@layers[[i]]@file@name <- paste0(getwd(),
+    strsplit(cs1_bioclim_raster@layers[[i]]@file@name,
+             "raptr-manuscript")[[1]][[2]])
 
@@
 The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$).
-Generally--unless small space targets are used, eg.
-0.000001, so that the solution is effectively unconstrained by the space target--demand points are represented by their closest selected planning units.
+Generally--unless near zero space targets are used so that the problem is effectively unconstrained by the target--demand points are represented by their closest selected planning units.
 In poorer quality solutions, demand points will be represented by planning units that have larger distances between them.
@@
 To understand how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2).
-We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.
+We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality.
 
@@
 ### Methods
-We simulated a hypothetical study area with square planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ grid (Fig.
+We simulated a hypothetical study area with square planning units arranged in a $`r sqrt(sim_parameters$number_planning_units)` \times `r sqrt(sim_parameters$number_planning_units)`$ grid (Fig.
 2).
@@
 Firstly, we simulated a hyper-generalist species (hereafter referred to as the "uniformly distributed species").
-It all planning units with equal probability (Fig.
+It occupied all planning units with equal probability (Fig.
 2a; eqn 4a).
@@
 2b).
-This species was most likely to be found in planning units in the centre of the study area, and was simulated using density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
+This species was most likely to be found in planning units nearest to the center of the study area.
+It was simulated using the density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
 Thirdly, we simulated a species two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig.
 2c).
-It was simulated using the maximum of two density of two multivariate normal distributions (eqn 4c).
-Planning unit occupancy was calculated using the $(X, Y)$ coordinates of the units' centroids and the corresponding equations.
+It was simulated using the maximum density of two multivariate normal distributions (eqn 4c).
+For a given species, planning unit occupancy was calculated using the $(X, Y)$ coordinates of the units' centroids and the relevant equation.
 We used a geographic attribute space to provide an intuitive visualization of the solutions.
@@
 %
-P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
+P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
 %
 P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
-\matchcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
+\mathcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
 \end{cases} \tag*{eqn 4c} \\
@@
 Firstly, to represent solutions generated using conventional methods (eg.
-\texttt{Marxan}), we generated solutions using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets.
-Secondly, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets).
-Thirdly, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
-Fourthly, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim.params.LST[[MODE]]$amount.target*100)` % amount targets and `r (sim.params.LST[[MODE]]$space.target*100)` % geographic spread targets and a boundary length modifier of `r sim.params.LST[[MODE]]$blm`.
+\texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets.
+Secondly, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets).
+Thirdly, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`.
+Fourthly, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % geographic spread targets, and boundary length modifiers of `r sim_parameters$blm`.
 
@@
 
-The solution generated for the uniformly distributed species using an amount target prioritized `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area (Fig.
+The solution generated for the uniformly distributed species using an amount target prioritized `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` planning units near the southern end of study area (Fig.
 2a).
-For this species, all solutions containing this many planning units are optimal when using a `r sim.params.LST[[MODE]]$amount.target*100` % amount target to constrain the reserve selection problem.
-We obtained this particular solution due to artifacts in the solver.
-This solution captured a small amount of the geographic spread of the species (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the centre of the species' distribution.
+For this species, all solutions containing this number of planning units are optimal when using a `r sim_parameters$amount_target*100` % amount target.
+We obtained this particular solution due to artifacts in the solver (eg.
+seed for the random number generator, order of variables in the problem matrix).
+This solution captured a small amount of the geographic spread of the species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled).
+In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the center of the species' distribution.
 
-By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
+By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
 3b).
 Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented.
-However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well-connected solution that met all of the objectives (Fig.
+However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (Fig.
 3d).
 This solution prioritized the same number of planning units as the previous solutions even though it is far superior.
-As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this objective.
+As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.
 
@@
 3e).
-While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution.
-By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig.
+While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution.
+By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
 3f).
@@
 3i).
-As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled).
-The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount target\nand BLM')$space.held,2)` % sampled; Fig.
+As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled).
+The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount target\nand BLM")$space_held, 2)` % sampled; Fig.
 3k).
-However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(filter(sim.spp.DF, species=='Bimodal\nspecies', prioritisation=='Amount & space\ntargets and BLM')$space.held,2)` % geographic spread secured; Fig.
+However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % geographic spread secured; Fig.
 3l).
@@
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
-We used the scores from the first `r as.character(as.english(cs1.params.LST[[MODE]]$n.pc.layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1.pca.DF[[4]][cs1.params.LST[[MODE]]$n.pc.layers],1)` % of the total variation; Fig.
+We used the scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig.
 4).
@@
 5).
-To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1.params.LST[[MODE]]$thin.distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1.params.LST[[MODE]]$mcp.percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
 We used this method because it is entirely reproducible using freely available data.
 
-We generated `r cs1.params.LST[[MODE]]$dp.number` demand points for each species (Fig.
+We generated `r cs1_parameters$dp_number` demand points for each species (Fig.
 S2).
@@
 For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location.
-We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1.params.LST[[MODE]]$dp.bandwidth[1]` and a `r cs1.params.LST[[MODE]]$dp.quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231].
+We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1_parameters$dp_bandwidth[1]` and a `r cs1_parameters$dp_quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231].
 We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations.
@@
 
-We generated two solutions.
-The first solution was generated using `r cs1.params.LST[[MODE]]$amount.target*100` % amount targets for all species.
-The second solution was generated the same amount target with an additional `r cs1.params.LST[[MODE]]$space.target*100` % niche-based target for each species.
+We generated two multi-species prioritizations.
+The first solution was generated using `r cs1_parameters$amount_target*100` % amount targets for each species.
+The second solution was generated the same amount targets with an additional `r cs1_parameters$space_target*100` % niche-based target for each species.
 
 ### Results
-Generally, the solution generated using amount targets preserved a representative sample of each the `r as.character(as.english(length(cs1.params.LST[[MODE]]$species.names)))` bird species' niches.
+Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches.
 (Figure S1).
-This solution secured over `r cs1.params.LST[[MODE]]$space.target*100` % of the realized niche of `r parsed.representative.space.held.names`.
-Yet it failed to achieve this for the `r parsed.not.representative.space.held.names`.
+This solution secured over `r cs1_parameters$space_target*100` % of the realized niche of `r parsed_representative_space_held_names`.
+Yet it failed to achieve this for the `r parsed_not_representative_space_held_names`.
 This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.
@@
 We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
-Although this dataset contains multiple species, we used data for the `r tolower(cs2.params.LST[[MODE]]$common.name)` (\textit{`r gsub('_', ' ',cs2.params.LST[[MODE]]$species.name,fixed=TRUE)`}) as a simple case study, because it has been found to exhibit significant genetic structure [@r462].
+Although this dataset contains multiple species, we used data for the `r tolower(cs2_parameters$common_name)` (\textit{`r gsub('_', ' ',cs2_parameters$species_name,fixed=TRUE)`}) as a simple case study, because exhibits significant inter-population genetic structure [@r462].
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
-In total, `r structurer::n.samples(cs2.spp.StructureData.LST[[1]])` individuals were genotyped at `r structurer::n.loci(cs2.spp.StructureData.LST[[1]])` markers.
+In total, `r structurer::n.samples(cs2_spp_structure_data[[1]])` individuals were genotyped at `r structurer::n.loci(cs2_spp_structure_data[[1]])` markers.
 
@@
 7c and 7d).
-To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled inverse great circle distances between the grid cells' centroids as weights [using the \texttt{ape R} package; @r464].
+To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled  great circle distances between the grid cells' centroids [using the \texttt{ape R} package; @r464].
 
 The grid cells were used as planning units for generating prioritizations.
-As previously mentioned, data were not collected in every grid cell, and so the planning units are arranged in a checkerboard pattern.
+Since data were not collected in every grid cell, and the planning units were arranged in a checkerboard pattern.
 The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
@@
 We generated two solutions.
-The first solution was generated using an `r cs2.params.LST[[MODE]]$amount.target*100` % amount-based target.
-The second solution was generated using the same amount-based target and an additional `r cs2.params.LST[[MODE]]$genetic.target*100` % genetic-based target.
+The first solution was generated using an `r cs2_parameters$amount_target*100` % amount-based target.
+The second solution was generated using the same amount-based target and an additional `r cs2_parameters$genetic_target*100` % genetic-based target.
 
@@
 
-A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2.spp.nmds.LST[[1]]$stress,2)`; Figs.
+A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2_spp_nmds[[1]]$stress,2)`; Figs.
 7c--7d).
-Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1.nmds.MoransI[[1]][[1]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1.nmds.MoransI[[1]][[2]]$observed,2)`$, $P `r pvalString(cs1.nmds.MoransI[[1]][[2]]$p.value)`$).
+Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1_nmds_morans_i[[1]][[1]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1_nmds_morans_i[[1]][[2]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[2]]$p.value)`$).
 In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
@@
 
-The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2.spp.DF$genetic[1],2)` % sampled; Figs.
+The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2_spp_results$genetic[1], 2)` % sampled; Figs.
 8a and 8c).
@@
 In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target.
-Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2.spp.DF$genetic[2],2)` % sampled; Figs.
+Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs.
 8b and 8d).
-Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2.spp.DF$Score[2],2)` compared to `r round(cs2.spp.DF$Score[1],2)` total cost).
+Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2_spp_results$Score[2],2)` compared to `r round(cs2_spp_results$Score[1],2)` total cost).
 

### 2017-04-06 code/R/analysis/00-initialization.R

@@ -1,116 +1,34 @@
-#### set default error message
-# parse args
+## Set parameters
 args <- commandArgs(TRUE)
-if (length(args)>0) {
-	print(args)
-	if (grepl('MODE',args))
-		MODE <- strsplit(grep('MODE', args, value=TRUE), '=', fixed=TRUE)[[1]][[2]]
+if (length(args) > 0) {
+  print(args)
+  if (grepl("MODE", args))
+    MODE <- strsplit(grep("MODE", args, value = TRUE), "=",
+                          fixed = TRUE)[[1]][[2]]
 }
 
-#### Load pacakges
-# load bioconductor packages
-# set checkpoint
-if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
-if (!file.exists('~/.checkpoint')) dir.create('~/.checkpoint')
-checkpoint::setSnapshot('2016-11-26')
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2')
-if (!'checkpoint' %in% installed.packages()[,'Package']) install.packages('checkpoint')
+if (!exists("MODE")) MODE <- "debug"
+cat("MODE = ", MODE, "\n")
 
-# load CRAN packages
-library(stats)
-library(checkpoint)
-library(session)
-library(data.table)
-library(RcppTOML)
+## Load packages
+library(magrittr)
+library(raster)
+library(sp)
 
-library(parallel)
-library(doParallel)
+## Load parameters
+general_parameters <- RcppTOML::parseTOML("code/parameters/general.toml") %>%
+                      `[[`(MODE)
 
-library(RColorBrewer)
-library(hexbin)
-library(grid)
-library(gridExtra)
-library(plotrix)
 
-library(plyr)
-library(dplyr)
-library(tidyr)
-library(testthat)
-
-library(raptr)
-library(rgeos)
-library(rworldxtra)
-library(maptools)
-library(adehabitatHR)
-
-library(vegan)
-library(fields)
-library(ape)
-library(RandomFields)
-
-library(Hmisc)
-library(pander)
-library(english)
-library(knitr)
-library(lazyWeave)
-library(broom)
-library(rmarkdown)
-
-## load github packages
-if (!'ALA4R' %in% installed.packages()[,'Package']) {
-  withr::with_libpaths(.libPaths()[1], c(install.packages('bitops', 'RCurl')))
[…]

### 2017-04-06 code/R/analysis/01-load-data.R

@@ -1,117 +1,111 @@
-## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/00-initialization.rda')
-
-### load simulation data
-# all data is simulated using functions in the package - no data to load
+## load session
+session::restore.session("data/intermediate/00-initialization.rda")
 
 ### load case-study 1 data
 # load in record bbox
-cs1.record.bbox <- shapefile('data/raw/record_bbox/record_bbox.shp') %>% writeWKT
+cs1_record_bbox <- "data/raw/record_bbox/record_bbox.shp" %>%
+                    raster::shapefile() %>%
+                    rgeos::writeWKT()
 
 # load in environmental data
-cs1.bioclim.RAST <- stack('data/raw/BioClim_variables/pca.tif')
-cs1.pca.DF <- read.table('data/raw/BioClim_variables/pca.TXT', skip=94) %>% `names<-`(
-	c('Principle Component', 'Eigen Value', 'Variation explained (%)',
-	'Accumulative variation explained (%)')
-)
+cs1_bioclim_raster <- raster::stack("data/raw/BioClim_variables/pca.tif")
+cs1_pca_data <- "data/raw/BioClim_variables/pca.TXT" %>%
+              read.table(skip = 94) %>%
+              magrittr::set_names(c("Principle Component", "Eigen Value",
+                                    "Variation explained (%)",
+                                    "Accumulative variation explained (%)"))
 
 ### load case-study 2 data
 ## compile spatial grid data
 # load in opportunity costs
-cs2.cost.RAST <- raster('data/raw/GRUMP_V1_Population_Density/grumpv1-popdensity.tif')
+cs2_cost_raster <- paste0("data/raw/GRUMP_V1_Population_Density/",
+                          "grumpv1-popdensity.tif") %>%
+                  raster::raster()
 
 # load in species population numbers
-cs2.spp.populations.DF <- fread('data/raw/Data_Meirmans_et_al_IntrabioDiv/NumberPopulations.csv', data.table=FALSE)
+cs2_spp_populations_data <- paste0("data/raw/Data_Meirmans_et_al_IntrabioDiv",
[…]

### 2017-04-06 code/R/analysis/02-simulations.R

@@ -1,139 +1,111 @@
-## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/01-load-data.rda')
+## load session
+session::restore.session("data/intermediate/01-load-data.rda")
 
 ## load parameters
-sim.params.LST <- parseTOML('code/parameters/simulations.toml')
+sim_parameters <- RcppTOML::parseTOML("code/parameters/simulations.toml") %>%
+                  `[[`(MODE)
 
 #### Simulate data
 # make planning units
-sim.pus <- sim.pus(as.integer(sim.params.LST[[MODE]]$number.planning.units))
+sim_pus <- raptr::sim.pus(as.integer(sim_parameters$number_planning_units))
 
 # simulate species distributions
-sim.spp <- lapply(
-	c('uniform', 'normal', 'bimodal'),
-	sim.species,
-  n=1,
-  x=sim.pus,
-  res=1
-)
+sim_spp <- lapply(c("uniform", "normal", "bimodal"), raptr::sim.species, n = 1,
+                  x = sim_pus, res = 1)
 
 # generate coordinates for pus/demand points
-sim.pu.points <- PlanningUnitPoints(
-	coords=gCentroid(sim.pus, byid=TRUE)@coords,
-	ids=seq_len(nrow(sim.pus@data))
-)
+sim_pu_points <- sim_pus %>%
+                 rgeos::gCentroid(byid = TRUE) %>%
+                 slot("coords") %>%
+                 raptr::PlanningUnitPoints(ids = seq_len(nrow(sim_pus@data)))
+
 
 # create demand point objects
-sim.dps <- lapply(
-	sim.spp,
-	function(x) {
-		return(
-			DemandPoints(
-				sim.pu.points@coords,
-				c(extract(x, gCentroid(sim.pus, byid=TRUE)))
-			)
-		)
-	}
-)
+sim_demand_points <- lapply(sim_spp, function(x) raptr::DemandPoints(
+  sim_pu_points@coords, c(extract(x, rgeos::gCentroid(sim_pus, byid = TRUE)))))
 
 ## create RapUnreliableOpts object
 # this stores parameters for the unreliable formulation problem (ie. BLM)
-sim.ro <- RapUnreliableOpts()
+sim_ro <- raptr::RapUnreliableOpts()
 
 ## create RapData object
 # create data.frame with species info
[…]

### 2017-04-06 code/R/analysis/03-case-study-1.R

@@ -1,164 +1,207 @@
-## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/01-load-data.rda')
+## load session
+session::restore.session("data/intermediate/01-load-data.rda")
 
 ## load parameters
-cs1.params.LST <- parseTOML('code/parameters/case-study-1.toml')
+cs1_parameters <- RcppTOML::parseTOML("code/parameters/case-study-1.toml") %>%
+                  `[[`(MODE)
 
 ### Prepare data for prioritizations
 ## prepare pca layers
 # select layers
-cs1.bioclim.RAST <- cs1.bioclim.RAST[[seq_len(cs1.params.LST[[MODE]]$n.pc.layers)]]
+cs1_bioclim_raster <- cs1_bioclim_raster[[cs1_parameters$number_pc_layers %>%
+                                          seq_len()]]
 
 # aggregate env layer to reduce computational burden
-cs1.bioclim.RAST <- aggregate(cs1.bioclim.RAST, cs1.params.LST[[MODE]]$environmental.layer.resolution/1000)
+cs1_bioclim_raster %<>% raster::aggregate(
+  cs1_parameters$environmental_layer_resolution / 1000)
 
 ## make planning units
 # download queensland layer
-aus.PLY <- getData('GADM', country='AUS', level=1, path='data/intermediate') %>% spTransform(cs1.bioclim.RAST@crs)
-qld.PLY <- aus.PLY[aus.PLY$HASC_1=='AU.QL',]  %>% gSimplify(1000)
-aus.PLY <- aus.PLY %>% gSimplify(1000)
+aus_polygons <- raster::getData("GADM", country = "AUS", level = 1,
+                                path = "data/intermediate") %>%
+                                sp::spTransform(cs1_bioclim_raster@crs)
+qld_polygons <- aus_polygons[aus_polygons$HASC_1 == "AU.QL", ]  %>%
+                rgeos::gSimplify(1000)
+aus_polygons %<>% rgeos::gSimplify(1000)
 
 # create planning units over queensland
-cs1.pus <- blank.raster(qld.PLY, cs1.params.LST[[MODE]]$pu.size) %>% rasterToPolygons()
-cs1.pus@data$id <- seq_len(nrow(cs1.pus@data))
[…]

### 2017-04-06 code/R/analysis/04-case-study-2.R

@@ -1,166 +1,157 @@
-## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/01-load-data.rda')
+## load session
+session::restore.session("data/intermediate/01-load-data.rda")
 
 ## load parameters
-cs2.params.LST <- parseTOML('code/parameters/case-study-2.toml')
+cs2_parameters <- RcppTOML::parseTOML("code/parameters/case-study-2.toml") %>%
+                  `[[`(MODE)
 
 ### prepare genetic attribute-spaces
 # subset to relevant species
-cs2.params.LST[[MODE]]$species.name <- gsub(' ', '_', cs2.params.LST[[MODE]]$species.name, fixed=TRUE)
-spp.pos <- which(unique(cs2.spp.samples.DF$species) %in% cs2.params.LST[[MODE]]$species.name)
-cs2.spp.StructureData.LST <- cs2.spp.StructureData.LST[spp.pos]
-cs2.spp.samples.DF <- cs2.spp.samples.DF[cs2.spp.samples.DF$species %in% cs2.params.LST[[MODE]]$species.name,]
+cs2_parameters$species_name <- gsub(" ", "_", cs2_parameters$species_name,
+                                    fixed = TRUE)
+spp_pos <- which(unique(cs2_spp_samples_data$species) %in%
+                 cs2_parameters$species_name)
+cs2_spp_structure_data <- cs2_spp_structure_data[spp_pos]
+cs2_spp_samples_data <- cs2_spp_samples_data[cs2_spp_samples_data$species %in%
+                                             cs2_parameters$species_name, ]
 
 # run mds
-cs2.spp.nmds.LST <- llply(
-	seq_along(cs2.spp.StructureData.LST),
-	.fun=function(i) {
-		# manually classify loci as neutral or adaptive
-		curr.spp <- bayescanr::BayeScanData(
-			cs2.spp.StructureData.LST[[i]]@matrix,
-			primers=cs2.spp.StructureData.LST[[i]]@loci.names,
-			populations=rep('1', nrow(cs2.spp.StructureData.LST[[i]]@matrix)),
-			labels=cs2.spp.StructureData.LST[[i]]@sample.names
-		)
-		curr.nmds <- bayescanr::nmds(
-			curr.spp,
-			metric='gower',
[…]

### 2017-04-06 code/R/analysis/05-benchmark-analysis.R

@@ -1,55 +1,60 @@
-## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/00-initialization.rda')
+## load session
+session::restore.session("data/intermediate/00-initialization.rda")
 
 ## load parameters
-benchmark.params.LST <- parseTOML('code/parameters/benchmark.toml')
-
-# load simulation functions
-source('code/R/functions/simulate.R')
+benchmark_parameters <- "code/parameters/benchmark.toml" %>%
+                        RcppTOML::parseTOML() %>%
+                        `[[`(MODE)
 
 ## compile benchmark parameters
-benchmark.params.DF <- ldply(seq_along(benchmark.params.LST[[MODE]][['number.of.features']]), function(i) {
-  expand.grid(number.features=benchmark.params.LST[[MODE]][['number.of.features']][[i]],
-    number.planning.units=benchmark.params.LST[[MODE]][['number.of.planning.units']][[i]],
-    formulation=benchmark.params.LST[[MODE]][['formulation']][[i]],
-    blm=benchmark.params.LST[[MODE]][['blm']][[i]],
-    replicate=seq_len(benchmark.params.LST[[MODE]][['replicates']]))
-  }
-)
-
-benchmark.DF <- ldply(sample.int(nrow(benchmark.params.DF)), function(i) {
-  # simulate data
-  curr.rd <- simulate.problem.data(
-    number.features=benchmark.params.DF[i,1],
-    number.planning.units=benchmark.params.DF[i,2],
-    amount.target=benchmark.params.LST[[MODE]]$amount.target,
-    space.target=benchmark.params.LST[[MODE]]$space.target,
-    probability.of.occupancy=benchmark.params.LST[[MODE]]$occupancy.probability
-  )
-  # create unsolved problem
-  curr.go <- GurobiOpts(MIPGap=general.params.LST[[MODE]][['MIPGap']],
-    Threads=general.params.LST[[MODE]][['threads']],
-    TimeLimit = benchmark.params.LST[[MODE]]$time.limit)
-  if (benchmark.params.DF[i,3] == "unreliable") {
-    curr.ru <- RapUnsolved(RapUnreliableOpts(BLM=benchmark.params.DF[i,4]), curr.rd)
-  } else {
-    curr.ru <- RapUnsolved(RapReliableOpts(BLM=benchmark.params.DF[i,4]), curr.rd)
[…]

### 2017-04-06 code/R/analysis/06-statistical-analysis.R

@@ -1,69 +1,88 @@
-## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/04-case-study-2.rda')
-load('data/intermediate/03-case-study-1.rda')
-load('data/intermediate/02-simulations.rda')
-load('data/intermediate/05-benchmark-analysis.rda')
+## load session
+session::restore.session("data/intermediate/04-case-study-2.rda")
+load("data/intermediate/03-case-study-1.rda")
+load("data/intermediate/02-simulations.rda")
+load("data/intermediate/05-benchmark-analysis.rda")
 
 ## case-study 2
 # generate inverse distance matrix
-cs2.pu.centroids <- gCentroid(cs2.grid.sub.PLY, byid=TRUE) %>% spTransform(CRS('+init=epsg:4326'))
-cs2.pu.inv.dists <- cs2.pu.centroids@coords %>% rdist.earth
-cs2.pu.inv.dists <- 1/cs2.pu.inv.dists
-diag(cs2.pu.inv.dists) <- 0
+cs2_pu_centroids <- cs2_grid_sub_polygons %>%
+                    rgeos::gCentroid(byid = TRUE) %>%
+                    sp::spTransform(sp::CRS("+init=epsg:4326"))
+cs2_pu_inv_dists <- cs2_pu_centroids@coords %>%
+                    fields::rdist.earth()
+cs2_pu_inv_dists <- 1 / cs2_pu_inv_dists
+diag(cs2_pu_inv_dists) <- 0
 
 # calculate Morans I
-cs1.nmds.MoransI <- llply(
-	unique(cs2.spp.samples.sub.DF$species),
-	function(x) {
-		curr.DF <- select(cs2.grid.sub.DF, contains(paste0(x,'_genetic_d')))
-		return(
-			llply(
-				seq_len(ncol(curr.DF)),
-				function(i) {
-					curr.col <- curr.DF[[i]]
-					curr.pos <- which(!is.na(curr.DF[[i]]))
-					return(
-						Moran.I(curr.DF[[i]][curr.pos], cs2.pu.inv.dists[curr.pos, curr.pos], scale=TRUE)
-					)
-				}
-			)
-		)
-	}
-)
+cs1_nmds_morans_i <- plyr::llply(
+  unique(cs2_spp_samples_sub_data$species),
+  function(x) {
[…]

### 2017-04-06 code/R/analysis/07-exports.R

@@ -1,6 +1,5 @@
- ## load .rda
-checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
-session::restore.session('data/intermediate/06-statistical-analysis.rda')
+ ## load session
+session::restore.session("data/intermediate/06-statistical-analysis.rda")
 
 ## save workspace
-save.session('data/final/results.rda', compress='xz')
+session::save.session("data/final/results.rda", compress = "xz")

### 2017-04-06 code/R/functions/data.R

@@ -4,8 +4,8 @@
 #'
 #' @seealso \code{\link[sp]{CRS}}
 #' @examples
-#' wgs1984 <- CRS('+init=epsg:4326')
-wgs1984 <- CRS('+init=epsg:4326')
+#' wgs1984
+wgs1984 <- sp::CRS("+init=epsg:4326")
 
 #' Europe Lambert Conformal Conic CRS
 #'
@@ -13,8 +13,8 @@ wgs1984 <- CRS('+init=epsg:4326')
 #'
 #' @seealso \code{\link[sp]{CRS}}
 #' @examples
-#' europeEA <- CRS('+init=esri:102014')
-europeEA <- CRS('+init=esri:102014')
+#' europe_ea
+europe_ea <- sp::CRS("+init=esri:102014")
 
 #' Europe Equidistant Conic
 #'
@@ -22,6 +22,5 @@ europeEA <- CRS('+init=esri:102014')
 #'
 #' @seealso \code{\link[sp]{CRS}}
 #' @examples
-#' europeED <- CRS('+init=esri:102031')
-europeED <- CRS('+init=esri:102031')
-
+#' europe_ed
+europe_ed <- sp::CRS("+init=esri:102031")

### 2017-04-06 code/R/functions/misc.R

@@ -1,146 +1,168 @@
-#' @include dependencies.R genetic.surrogates-internal.R
-
-#' Buffered range 
+#' Buffered range
 #'
 #' This function returns the range of a vector that has been buffered by a percentage.
-#' 
+#'
 #' @param x \code{numeric} vector.
+#'
 #' @return \code{numeric} vector of two elements.
+#'
 #' @seealso \code{\link[base]{range}}.
+#'
 #' @examples
 #' buffered.range(c(0,0.5,1), 0.05)
+#'
 #' @export
-buffered.range <- function(x, percent=0.05) {
-	range(x)+(c(-1, 1) * ((diff(range(x))*percent)))
+buffered_range <- function(x, percent = 0.05) {
+  range(x) + (c(-1, 1) * ( (diff(range(x)) * percent)))
 }
 
 #' spFortify
-#' 
-#' This function is a convienience function designed to prepare a \code{Spatial*DataFrame} object
-#' for visualisation using ggplot2.
-#' @param x \code{\link[sp]{SpatialLinesDataFrame}}, \code{\link[sp]{SpatialPointsDataFrame}}, or \code{\link[sp]{SpatialPolygonsDataFrame}} object.
-#' @param columns \code{character} with column names to include in returned object.
-#' @return \code{data.frame}
-#' @seealso \code{\link[ggplot2]{fortify}}, \code{\link[sp]{SpatialLinesDataFrame}}, \code{\link[sp]{SpatialPointsDataFrame}} \code{\link[sp]{SpatialPolygonsDataFrame}}.
-#' @export
-spFortify <- function(x, columns=names(x@data)) {
-	if ('id' %in% names(x))
-		x@data <- rename(x@data, id2=id)
-	x$fid <- seq_len(nrow(x@data))
-	f <- fortify(x, region="fid")
-	f <- merge(f, x@data, by.x='id', by.y='fid')
-	f <- f[,which(names(f) != 'fid')]
-	return(f)
-}
-
-#' pretty.pval
-#' 
-#' This function returns the rounded version of a p-value for use in text along with the correct arithmatic operator.
-#' @param x \code{numeric} p-value
-#' @return \code{character} object.
-#' @example
-#' pretty.pval(0.04)
-#' pretty.pval(0.0000001)
-#' pretty.pval(0.565656)
-#' @export
-pretty.pval <- function(x) {
-	if (x > 1)
-		return('> 1')
[…]

### 2017-04-06 code/R/functions/simulate.R

@@ -24,25 +24,26 @@
 #'
 simulate_data <- function(x, n, model, transform=identity, ...) {
   # assert valid arguments
-  if (!'RandomFields' %in% rownames(installed.packages()))
-    stop('the "RandomFields" package needs to be installed to simulate data')
+  if (!"RandomFields" %in% rownames(installed.packages()))
+    stop("the \"RandomFields\" package needs to be installed to simulate data")
   assertthat::assert_that(
-    inherits(x, 'RasterLayer'),
+    inherits(x, "RasterLayer"),
     assertthat::is.number(n),
-    is.finite(raster::cellStats(x, 'max')),
-    inherits(model, 'RMmodel'),
-    inherits(transform, 'function'))
+    is.finite(raster::cellStats(x, "max")),
+    inherits(model, "RMmodel"),
+    inherits(transform, "function"))
   # generate values for rasters
-  coords <- as(x, 'SpatialPoints')@coords
-  mtx <- RandomFields::RFsimulate(model=model, x=coords[,1], y=coords[,2],
-                                  n=n, spConform=FALSE, ...)
+  coords <- as(x, "SpatialPoints")@coords
+  mtx <- RandomFields::RFsimulate(model = model, x = coords[, 1],
+                                  y = coords[, 2], n = n, spConform = FALSE,
+                                  ...)
   # convert to matrix if not a matrix
-  if (!inherits(mtx, 'matrix'))
-    mtx <- matrix(mtx, ncol=1)
+  if (!inherits(mtx, "matrix"))
+    mtx <- matrix(mtx, ncol = 1)
   # generate populate rasters with values
   stk <- raster::stack(lapply(seq_len(ncol(mtx)), function(i) {
     r <- x
-    r[raster::Which(!is.na(r))] <- transform(mtx[,i])
+    r[raster::Which(!is.na(r))] <- transform(mtx[, i])
     r
   }))
   # return raster stack with simulated distributions
@@ -60,101 +61,100 @@ simulate_data <- function(x, n, model, transform=identity, ...) {
 #'
 #' @seealso \code{\link{simulate_data}}.
 #'
-simulate_species <- function(x, n=1, model=RandomFields::RMgauss(),
+simulate_species <- function(x, n = 1, model = RandomFields::RMgauss(),
                              transform=plogis, ...) {
-  simulate_data(x=x, n=n, model=model, transform=transform, ...)
+  simulate_data(x = x, n = n, model = model, transform = transform, ...)
 }
 
 #' Simulate problem data
[…]

### 2017-04-06 code/rmarkdown/article.tex
r pkgs

--- 2016-07-18 article/article.tex
@@
     \usepackage{mathspec}
-    \usepackage{xltxtra,xunicode}
   \else
@@
   \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
 \fi
@@
 \usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={raptr: Representative and Adequate Prioritization Toolkit in R},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={rapr: Representative and Adequate Prioritisations in R},
-            colorlinks=true,
-            citecolor=blue,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
 \setlength{\parindent}{0pt}
 \setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 \setcounter{secnumdepth}{0}
@@
 \setlength{\droptitle}{-2em}
-  \title{rapr: Representative and Adequate Prioritisations in R}
+  \title{raptr: Representative and Adequate Prioritization Toolkit in R}
   \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-  \author{Jeffrey O. Hanson$^1$, Jonathan R. Rhodes$^2$, Hugh P. Possingham$^1$,
-Richard A. Fuller$^1$\\$^1$School of Biological Sciences, The University
-of Queensland, Brisbane, QLD, Australia\\$^2$School of Geography,
-Planning and Environmental Management, The University of Queensland,
-Brisbane, QLD, Australia\\Correspondance should be addressed to
-\href{mailto:jeffrey.hanson@uqconnect.edu.au}{jeffrey.hanson@uqconnect.edu.au}}
+  \author{Jeffrey O. Hanson\(^1\), Jonathan R. Rhodes\(^2\), Hugh P.
+Possingham\(^{1,3}\), Richard A. Fuller\(^1\)\\
+\(^1\)School of Biological Sciences, The University of Queensland,
+Brisbane, QLD, Australia\\
+\(^2\)School of Earth and Environmental Sciences, The University of
+Queensland, Brisbane, QLD, Australia\\
+\(^3\)The Nature Conservancy, South Brisbane, QLD, Australia\\
+Correspondence should be addressed to
+\href{mailto:jeffrey.hanson@uqconnect.edu.au}{\nolinkurl{jeffrey.hanson@uqconnect.edu.au}}}
   \preauthor{\centering\large\emph}
@@
   \postdate{\par}
-  \date{18 July 2016}
+  \date{06 April 2017}
 
 % load packages
-\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,titling,natbib,pdfpages,lineno,booktabs}
+\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,lineno,booktabs,subfiles}
 \usepackage[T1]{fontenc}
@@
 
-% format toc
-%\renewcommand\cftsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubsubsecfont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftparafont{\normalfont\normalsize\bfseries}
-% \renewcommand\cftsubparafont{\normalfont\normalsize\itshape}
-
 % line numbers
 \linenumbers
+
+% allow breaks in equations
+\allowdisplaybreaks
 
@@
 \begin{document}
-
 \maketitle
 
-\begin{abstract}
+Abstract word count: 347 / 350\\
+Total word count: XXXX / 7000\\
+Number of references: 42\\
+Number of figures: 8\\
+Number of tables: 0\\
+Number of textboxes: 0\\
+Number of equations: 4\\
+Short running title: Short abstract: Prioritize biodiversity processes
+and patterns using raptr \url{https://github.com/jeffreyhanson/raptr}\\
+Keywords: conservation, biodiversity, mixed integer linear programming,
+optimization, protected areas
+
+\clearpage
+
+\section{Summary}\label{summary}
+
 \begin{enumerate}
 \def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
 \item
-  An underlying aim in spatial conservation planning is to maximise the
-  long-term persistence of biodiversity. Yet current approaches
-  overwhelming focus on ``adequacy'', the concept ensuring a target
-  proportion of the distribution of a biodiversity feature (species or
-  ecosystem) is reserved. Ultimately, to achieve long-term persistence,
-  the processes the sustain biodiversity must also be conserved. Many
-  biodiversity processes can be preserved by conserving a representative
-  sample of each biodiversity feature. This idea has for decades been
-  encapsulated in principle of ``representativeness''. For example,
-  representing the full range of genetic diversity or ability to persist
-  in a range of climatic conditions might be crucial for the long -term
-  persistence of a species in a world where environmental change is
-  accelerating. However, fully operationalising the principle of
-  repressiveness in spatial conservation planning has proven
-  methodologically challenging.
+  An underlying aim in conservation planning is to maximize the
+  long-term persistence of biodiversity. To fulfill this aim, the
+  ecological and evolutionary processes that sustain biodiversity must
+  be preserved. One way to conserve such processes at the feature level
+  (eg. species, ecosystem) is to preserve a sample of the feature (eg.
+  individuals, areas) that is representative in terms of the intrinsic
+  or extrinsic physical attributes that underpin the process of
+  interest. For example, by conserving a sample of populations with
+  local adaptations that is representative in terms of the full
+  variation found in the species--physical attributes associated with
+  adaptation--protected areas can foster adaptive processes by ensuring
+  these adaptations are not lost. Despite this, current reserve
+  selection methods overwhelmingly focus on securing an adequate amount
+  of area or habitat for each feature. Little attention has been
+  directed towards capturing a representative sample of each feature.
 \item
-  To address this issue, we developed the
-  \texttt{rapr: Representative and Adequate Pioritisations in R}, a
-  toolkit to guide reserve selection using explicit targets for
-  representing biodiversity processes as well as meeting area protection
-  targets. Using a novel formulation of the reserve selection problem,
-  users set ``space targets'' to secure a representative sample of each
-  species within an attribute space reflecting any biodiversity process
-  (eg. an attribute space expressing variation in genetic
-  characteristics between individuals in an area, or variation in
-  climatic conditions between areas occupied by indivduals). We explored
-  the functionality of this \texttt{R} package using a simulation study
-  and two case studies. In each study, we generated prioritisations that
-  aimed to preserve an amount of habitat for each species using amount
-  target--to represent conventional reserve selection methods--and
-  compared them with prioritisations generated using both habitat and
-  space targets--to represent this method.
+  To address this issue, we developed the \texttt{raptr R} package to
+  help guide reserve selection. Users set ``amount targets''--similar to
+  conventional methods--to ensure that solutions secure a sufficient
+  proportion of area or habitat for each feature. Additionally, users
+  set ``space targets'' to secure a representative sample of variation
+  in ecologically or evolutionary relevant attributes (eg. environmental
+  or genetic variation). We demonstrate the functionality of this
+  package using simulations and two case studies. In these studies, we
+  generated solutions using amount targets--representing conventional
+  methods--and compared them with solutions generated using amount and
+  space targets.
 \item
-  We show that explicitly considering biodiversity processes in reserve
-  selection can substantially change the configuration of the resulting
-  solution. Simulations suggest that including representativeness in
-  conservation planning is particularly important where the biodiversity
-  feature has multi-model distributions in an attribute space. Results
-  from initial case studies show that explicitly considering space-based
-  targets for biodiversity processes can result in prioritisations that
-  do not necessarily require much more area than traditional approaches,
-  but that could be much more effective in achieving long term
-  biodiversity persistence.
+  Our results demonstrate that different solutions emerge when targeting
+  a representative sample of each feature. We show that using these
+  targets is important for features that have multimodal distributions
+  in the process-related attributes (eg. species with multimodal
+  niches). We also found that solutions could conserve a far more
+  representative sample with only a slight increase in reserve size.
 \item
-  The \texttt{rapr R} package provides a unified framework achieving
-  spatial conservation prioritisations that are both adequate and
-  representative, resulting in a greater chance of long term
-  biodiversity persistence that area-based planning alone.
+  The \texttt{raptr R} package provides a toolkit for making
+  prioritizations that secure an adequate and representative sample of
+  each feature. By using solutions that secure a representative sample
+  of each feature, prioritizations may have a greater chance of
+  achieving long-term biodiversity persistence.
 \end{enumerate}
-\end{abstract}
 
-Short running title: Representative and adequate prioritisations
-\newline
-Abstract word count: 370 / 350\\Total word count: XXXXX / 8000\\Number
-of references: 34\\Number of figures: 8\\Number of tables: 0\\Number of
-textboxes: 0\\Number of equations: 8\\\newpage
+\clearpage
+\input{./text.tex}
 
-\section{Introduction}\label{introduction}
-
-Perhaps the most fundamental aim of conservation is to maximise the
-long-term persistence of biodiversity at organisational levels (McNeely
-1994; Margules \& Pressey 2000). To achieve this, conservation actions
-must preserve biodiversity patterns (eg. populations, species,
-ecosystems), but also crucially the processes that sustain them. One of
-the major tangible achievements of modern conservation has been the act
-of setting aside areas for preservation (Sanderson \emph{et al.} 2015).
-Reserve networks buffer species from gross threatening processes and set
-the stage for enhanced management interventions (Gaston \emph{et al.}
-\{2008\}). However, the resources available for conservation action are
-limited, and so reserve networks must be sited in places that satisfy
-conservation objectives for minimum cost (Margules \& Pressey 2000). To
-achieve this, reserve selection is often formulated as an optimisation
-problem and then solved to identify cost-effective candidate reserve
-systems (termed prioritisations; Margules \& Pressey 2000). Many
-decision support tools have been developed to help identify
-cost-effective solutions (eg. \texttt{ ConsNet}, Ciarleglio \emph{et
-al.} 2009;\texttt{ Marxan}, Ball \emph{et al.} 2009;\texttt{ Zonation},
-Moilanen 2007). Typically, these tools are used to deliver a
-prioritisation that preserves a set of species of interest to the
-conservation planner (termed target species or features). Given a set of
-species, decision makers can use these tools to preserve biodiversity
-patterns by generating a prioritisation that secures an adequate
-proportion of each species' range.
-
-However, to achieve effective conservation, reserve networks must do
-more than cover enough area. They must capture the processes that
-sustain biodiversity over the long term; they must capture both
-ecological and evolutionary processes (Crandall \emph{et al.} 2000;
-Margules \& Pressey 2000). Ecological processes, such as predator-prey
-interactions, pollination, and decomposition, are required for
-biodiversity to persist over short time-scales. Typically, they operate
-over small geographic domains--with exceptions such as migration and
-refugial habitats--and can be preserved using suitably large planning
-units (Ciarleglio \emph{et al.} 2009) that each contain a discrete unit
-of habitat (Klein \emph{et al.} 2009). On the other hand, evolutionary
-processes are required for biodiversity to persist over long
-time-scales, and they typically operate over large geographic domains.
-Protected areas must represent adaptive evolutionary processes to foster
-resilience against environmental change (eg. climate change; Pyke \&
-Fischer 2005). Protected areas must also preserve neutral evolutionary
-processes (Moritz 1994, 1999, 2002), arising from restriction of gene
-flow between populations. They are important for maintaining genetic
-diversity, and avoiding inbreeding depression. In recent decades, a
-wealth of data relevant to evolutionary processes has become available
-to conservation planners (eg. bioclimatic, genetic, and trait data;
-Hijmans \emph{et al.} 2005; Raymond \emph{et al.} 2015; Jones \emph{et
-al.} 2009). Yet such data are rarely used to guide conservation planning
-(Hendry \emph{et al.} 2010). This is perhaps in large part due to that
-fact that existing reserve selection tools focus on capturing either
-biodiversity patterns or processes--but not both.
-
-Existing conservation planning tools are generally not well-suited for
-representing biodiversity processes (but see Faith 2003). To preserve
-biodiversity processes, a prioritisation must capture a representative
-sample of each species. For instance, to preserve predator-prey
-interactions, a prioritisation must preserve individuals from each
-predator and prey species in the same area. To preserve adaptive
-evolutionary processes, a prioritisation must preserve the adaptive
-landscape of each species--populations experiencing different selection
-pressures (Moritz 2002). To preserve neutral evolutionary processes, a
-prioritisation must secure individuals descended from each of the
-genetic lineages that comprise each species (Moritz 1994). Previous
-attempts to accommodate such into multi-species conservation planning
-have partitioned species' ranges into different groups, for example at
-habitat discontinuities (e.g. Carvalho \emph{et al.} 2011) as a
-pre-processing step to render a new set of ``pseudo-species''. However,
-this approach assumes that the biodiversity processes that operate
-across a species' range can be readily split into discrete units. Yet
-data on biodiversity processes is often continuous and
-hyper-dimensional, and often cannot be reduced to a few categories
-without significant information loss (Faith \& Walker 1996).
-
-One of the key issues in reserve selection is the lack of a unifying
-decision support tool that can accommodate data on biodiversity patterns
-and processes in a multi-species context. To begin to fill this gap, we
-present the \texttt{rapr R} package. This \texttt{R} package uses novel
-formulations of the reserve selection problem to provide decision makers
-with the tools to generate prioritisations that preserve both
-biodiversity patterns and processes. We aim to define the concepts
-behind the problem formulations. Furthermore, we aim to explore the
-functionality of the \texttt{rapr R} package by applying one of the
-formulations to a set of simulated species and two case studies.
-
-\section{Methods}\label{methods}
-
-\subsection{PROBLEM FORMULATION}\label{problem-formulation}
-
-Biodiversity features are defined as the entities that the
-prioritisation is required to preserve (eg. species, ecosystems).
-Spatial attributes are defined as the variation across the species'
-range that the prioritisation is required to sample. They can be
-intrinsic (eg. genetic or trait variation) or extrinsic (eg.
-environmental variation) to the feature. These attributes should be
-related to the biodiversity processes that the decision maker aims to
-preserve.
-
-Each attribute is conceptualised as a space. This space is termed an
-``attribute space''. Each planning unit is thought to occupy a single
-point inside each space. For example, a decision maker may require a
-prioritisation that represents populations along climatic gradients. To
-achieve this, the decision maker might use an ``climatic'' attribute
-space with dimensions relating to mean annual temperature ($^{\circ}$C)
-and precipitation (mm). Any given combination of temperature and
-precipitation may be conceived as a point in this environmental space.
-By associating planning units with climatic data, they can be mapped
-from geographic space to this environmental attribute space.
-
-Demand points are points that also exist in an attribute space. They are
-designated by the decision maker to indicate regions of the attribute
-space that the decision maker wishes to represent in the prioritisation
-(see below for discussion on how demand points can be generated for
-real-world datasets). The amount of variation in the attribute space
-that a prioritisation secures is a function of the distance between each
-demand point and each selected planning unit in the attribute space. The
-shorter the distances between the demand points and the planning units;
-the better the prioritisation is at securing the variation in the
-spatial attribute. To convert these amounts to a proportion--a
-meaningful unit for a decision maker--the distances between the selected
-planing units and the demand points are scaled by the distances between
-the demand points to the centroid of the demand points. In any attribute
-space there may exist points that are impossible (eg. mean annual
-rainfall -5 mm), or do not occur in the study area (eg. mean annual
-temperature 30$^{\circ}$C in Antarctica). Additionally, there may be
-some regions that are desirable for some features and undesirable for
-others (eg. conditions known to be outside the physiological tolerance
-of certain species). Thus a different set of demand points and weights
-are used for each attribute space and each feature. By placing demand
-points in desirable regions of an attribute space for a given feature,
-the decision maker can ensure that prioritisations secure the feature in
-planning units with spatial attributes that are desirable for that
-feature.
-
-To illustrate these concepts, consider an example conservation planning
-scenario example involving an attribute space and demand points. We wish
-to develop a prioritisation for a single species that has four
-populations in the study area. However, we can only afford to preserve
-three populations. We aim preserve the adaptive landscape of the
-species, and to achieve this, we will preserve populations inhabiting
-different environmental conditions. To describe environmental variation,
-we obtain data on the environmental conditions (rainfall (mm) and
-temperature ($^{\circ}$C)) where each population is found. These
-environmental data are then used to construct a two-dimensional
-environmental attribute space. Next, we generated demand points as
-equi-distant points between the range of values where the populations
-were found. By comparing the distribution of the demand points to the
-distribution of the populations in the attribute space, we can identify
-a suitable prioritisation (Fig. 1). We can see that preserving both
-populations $A$ and $C$, effectively ``doubles-up'' on the same
-environmental characteristics, resulting in considerable redundancy in
-the solution. Instead, a more representative sample of the
-intra-specific variation could be captured by securing populations $A$,
-$B$, and $D$. This example demonstrates how the inclusion of
-biodiversity processes can guide the reserve selection process.
-
-The formulations used to express the reserve selection problem in the
-\texttt{rapr R} package are based on a combination of the
-\texttt{Marxan} reserve selection problem and the uncapacitated facility
-location problems (Cornu{é}jols \emph{et al.} 1990). Although the
-\texttt{rapr R} package provides two novel formulations, for brevity, we
-will define the simpler formulation--referred to as the unreliable
-formulation--below and define the more complex version--the reliable
-formulation--in the Supporting Information S1. These formulations are
-named after the unreliable and reliable facility location formulations
-from which they are based upon (Cui \emph{et al.} 2010). The key
-difference between these two formulations is that the reliable
-formulation explicitly considers the probability that planning units are
-occupied when calculating the proportion of an attribute space sampled
-in a solution. All mathematical terms defined hereafter are described in
-Table S1. For convenience, the cardinality of sets will be denoted using
-the same symbol used to denote the variable.
-
-Define $F$ to be the set of features one wishes to conserve (indexed by
-$f$). Let $J$ be a set of planning units (indexed by $j$). Also, let
-$A_j$ denote the area, and $C_j$ denote the cost of preserving planning
-unit $j \in J$. To assess the extent to which each feature is secured in
-a given prioritisation, let $q_{fj}$ denote the probability of feature
-$f$ occupying planning unit $j$. The level of fragmentation associated
-with a prioritisation is parametrised as the net exposed boundary
-length. Let the shared edges between each planning unit $j \in J$ and
-$k \in J$ be $e_{jk}$.
-
-Let $S$ denote a set of attribute spaces (indexed by $s$). Each
-$j \in J$ is associated with spatially explicit data that represent
-coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a
-set of demand points (indexed by $i$) for each feature $f \in F$ and
-each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting
-for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$
-denote the distance between each demand point $i \in I$ and each
-planning unit $j \in J$ for each feature $f \in F$ and attribute space
-$s \in S$. To describe the inherent variation in the distribution of
-demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote
-the distance between each demand point $i \in I$ and the centroid of the
-demand points. Demand points with greater weight $\lambda_{fsi}$ are
-more important, and the optimal solution will be likely to select
-planning units close to highly weighted demand points. As a consequence,
-the decision maker will need to choose an appropriate weighting for each
-demand point.
-
-Targets are used to ensure that prioritisations adequately preserve each
-species. Amount-based targets are used to ensure that the total amount
-of habitat preserved is sufficient. Let $T_f$ denote the expected amount
-of area that needs to be preserved for each feature $f \in F$.
-Space-based targets ensure that a sufficient proportion of the
-intra-specific variation is secured. Let $\tau_{fs}$ denote the
-space-based targets for feature $f \in F$ and attribute space $a \in A$.
-For convenience, these both types of targets are expressed as
-proportions in the \texttt{R} package.
-
-The control variables for the unreliable formulation are the $B$,
-$T_{s}$, and $\tau_{sa}$ variables.
-
-\begin{align*}
-T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
-%
-\tau_{sa} &= \text{representation target for feature $f$ in attribute space $a$} \tag*{eqn 1b}\\
-%
-B &= \text{boundary length modifier (BLM): penalise fragmented solutions} \tag*{eqn 1c}\\
-\end{align*}
-
-The decision variables are the $X_j$ and $Y_{fsij}$ variables.
-
-\begin{align*}
-X_j
-    &= \begin{cases}
-        1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
-        0, & \parbox{25em}{otherwise} \\
-    \end{cases} \\
-%
-Y_{fsij} &= \begin{cases}
-        1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
-        0, & \parbox{25em}{otherwise} \\
-    \end{cases} \\
-\end{align*}
-
-Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is
-assigned to a selected planning unit $J$ where $X_j = 1$. The weighted
-distance between the demand point and its assigned planning unit
-$\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is
-represented in a given solution. Generally, demand points are assigned
-to the closest selected planning units (unless particularly low
-space-based targets are used).
-
-The unreliable formulation (URAP) is a defined as a multi-objective
-optimisation problem.
-
-\begin{align*}
-& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
-%
-& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
-%
-& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
-& & & & & 0 \leq s \leq S-1\\
-%
-& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
-& & & & & 0 \leq s \leq S-1, \\
-& & & & & 0 \leq i \leq I-1\\
-%
-& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
-& & & & & 0 \leq s \leq S-1, \\
-& & & & & 0 \leq i \leq I-1,\\
-& & & & & 0 \leq j \leq J-1\\
-%
-& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
-& & & & & 0 \leq s \leq S-1,\\
-& & & & & 0 \leq i \leq I-1\\
-%
-\end{align*}
-
-The objective function (eqn 3a) determines the utility of a given
-prioritisation: a combination of the total cost of a prioritisation and
-how fragmented it is. Constraints (eqn 3b) ensure that all the
-amount-based targets are met. Constraints (eqn 3c) ensure that all the
-space-based targets are met for each feature and each attribute space.
-For each feature and attribute space, the total weighted distance
-between the demand points and their closest selected planning units is
-calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted
-distance is then scaled by the inherent variation in the demand points
-($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a
-proportion conceptually similar to the $R^2$ statistic used in $k$-means
-clustering analysis. The constraints ensure that this proportion must be
-equal to or greater than the space-based target. Constraints (eqn 3d)
-ensure that only one planning unit is assigned to each demand point.
-Constraints (eqn 3e) ensure that demand points are only assigned to
-selected planning units. Constraints (eqn 3f) ensure that the $X$ and
-$Y$ variables are binary.
-
-\subsection{OPTIMISATION}\label{optimisation}
-
-The unreliable formulation is non-linear. However, the non-linear
-components can be linearised using existing techniques. The expression
-$X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer
-et al. (2016). Linearised versions of the problems can be solved using
-commercial exact algorithm solvers. The \texttt{rapr} R package provides
-functions to express conservation planning data as optimisation problems
-using linearised versions of the unreliable and reliable formulations.
-These optimisation problems can then be solved to generate
-prioritisations using the commercial \texttt{Gurobi} software suite
-(\url{http://www.gurobi.com}). Presently, academics can obtain a license
-at no cost from the Gurobi website. After installing the \texttt{Gurobi}
-software suite, users will need to install the \texttt{Gurobi R}
-package.
-
-\section{Examples}\label{examples}
-
-To understand the behaviour of the unreliable problem and showcase its
-value, we conducted a simulation study and two case studies. These
-studies involved generating solutions using only amount targets to
-represent prioritisations generated using conventional methods (eg.
-\texttt{Marxan}), and solutions using amount and space targets using the
-unreliable formulation. By comparing these solutions, we can guage the
-benefits of explicitly including space targets in reserve selection. All
-analyses were performed in \texttt{R} (version 3.3.0; {R Core Team}
-2016).
-
-\subsection{SIMULATION STUDY}\label{simulation-study}
-
-\subsubsection{Methods}\label{methods-1}
-
-We simulated a hypothetical study area with planning units arranged in
-$10 \times 10$ square grid (Fig. 2). We then simulated three species
-across this study area. The first species was simulated to represent a
-hyper-generalist--occurring in all planning units with equal probability
-(Fig. 2a). This species' distribution was based on a uniform
-distribution (eqn 8a). The second species was simulated to represent a
-species with an idealised distribution (Fig. 2b). It has a core range
-area, and is less likely to be found in areas that are distant from the
-core area. This species' distribution was simulated using the
-probability density function of a single multivariate normal
-distribution (eqn 8c). The third species represents a species with two
-distinct populations (Fig. 2c). This species' distribution was simulated
-to depict a bimodal distribution, based on the combination of the
-probability density functions of two multivariate normal distributions
-(eqn 8d). The probability that each species inhabited a given planning
-unit was calculated using the $X, Y$ coordinates of the planning unit
-and eqns 8a--8d. We used a geographic attribute space to showcase the
-behavior of the problem. For each species, demand points were generated
-by calculating the centroids of each planning units and their weights
-were set as the probability that the planning units were occupied.
-
-\begin{align*}
-\\
-P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 8a} \\
-%
-f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 8b} \\
-%
-P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 8c} \\
-%
-P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
-f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
-\end{cases} \tag*{eqn 8d} \\
-%
-\end{align*}
-
-We generated four solutions for each species. First to represent
-solutions generated using conventional conservation planning methods, we
-generated solutions using only 20 \% amount targets. Second to show how
-the addition of geographic targets can affect a prioritisation, we
-generated solutions using 20 \% amount targets and 75 \% geographic
-targets. Third to represent solutions using conventional planning
-methods that penalise for fragmentation, we generated solutions using 20
-\% amount targets and a boundary length modifier of 1. Fourth to
-illustrate the combined affect using geographic targets and penalising
-fragmentation, we generated solutions using using 20 \% amount targets
-and 75 \% geographic targets and a boundary length modifier of 1.
-
-\subsubsection{Results}\label{results}
-
-The uniform species was simulated to occur with a constant probability
-of occupancy across the study area (Fig. 2a). The solution generated
-using 20 \% amount-based targets (Fig. 3a) selected 20 planning units
-near the southern end of study area. This configuration is an artifact
-of the method used to solve this particular instance of the reserve
-selection problem. In fact, for this species, all prioritisations
-containing 20 planning units are optimal when considering only 20 \%
-amount targets. In the absence of criteria to guarantee
-representativeness, reserve selection methods may or may not return
-solutions that secure a representative sample of the features. In this
-particular case, the solution does not secure a representative sample of
-the species' range (-23.64 \% sampled).
-
-The addition of a 75 \% geographic target resulted in a solution that
-secured a representative sample of the uniform species' range (90.67 \%
-sampled; Fig. 3b). Since all planning units have equal cost and an equal
-chance of being occupied, this solution has the same number of planning
-units as the solution generated using only amount targets (cf.~Fig 3a).
-Although the use of amount and space based targets has addressed
-adequacy and representativeness objectives (respectively), they have
-resulted in a highly fragmented solution.
-
-The addition of a positive boundary length modifier (BLM) parameter
-resulted in a well-connected solution that secured an adequate
-proportion (20 \% ) and representative sample (75.76 \% sampled) of the
-species geographic distribution (Fig. 3d). This solution contains 20
-planning units--a few more than the previously discussed solutions--to
-ensure that the solution secures a representative proportion of the
-species geographic distribution in a configuration that is not highly
-fragmented. This result suggests that the combination of space targets
-and boundary length modifiers may yield solutions that contain more
-planning units, since the solution generated using amount targets and a
-boundary length modifier (Fig. 3c) contained the sample number of
-planning units as the solution with just the amount targets (Fig. 3a).
-Overall, these results show that under the simplest of conditions, the
-reliance on just amount targets can cause an under-specified reserve
-selection problem that is unlikely to return a suitable solution for
-implementation.
-
-The normally distributed species was simulated to contain a single core
-area where individuals are most prevalent and marginal areas where
-individuals are less likely to occur (Fig. 2b). The solution generated
-using 20 \% amount targets contained 10 planning units and concentrated
-conservation efforts in the core area (Fig. 3e). Since all planning
-units have equal costs and areas, this solution contains the planning
-units with the highest probabilities of occupancy. Whilst this solution
-satisfies the adequacy objective for a prioritisation in a
-cost-effective manner--it fails to fulfill the representative objective
-for a prioritisation (59.09 \% distribution sampled).
-
-The solution generated using 20 \% amount targets and 75 \% geographic
-targets resulted in a solution that is both adequate and representative
-in terms of the uniform species distribution (Fig. 3f). This solution
-used 11 planning units to secure 20.02 \% and sample 76.31 \% of the
-normal species' distribution. By using amount and geographic targets in
-the reserve selection problem, we have obtained a solution that
-concentrates conservation effort in the range core and also the range
-margin. However, similar to the corresponding solution for the uniform
-species (Fig. 3b), this solution is highly fragmented.
-
-By using 20 \% amount targets and 75 \% geographic targets and a
-boundary length modifier parameter ($BLM = 1$), we obtained a solution
-that fulfills adequacy, representativeness and connectivity objectives
-(Fig. 3h). This solution contains 12 planning units. Similar to the
-corresponding solution for the uniform species (Fig. 3d), this solution
-contains more planning units than any other solutions for this species
-(cf.~Figs. 3e--3g). The results for the normally distributed species
-suggest that the space targets can result in solutions that secure a
-more representative sample of the species'--even for species without
-significant structure.
-
-The bimodally distributed species was simulated to represent a species
-with a highly structured distribution, and it contains two distinct
-populations (Fig. 2c). The solution generated using just 20 \% amount
-targets assigned conservation effort to just one of the two populations
-(Fig. 3i). While this solution secured an adequate proportion of the
-species' distribution (20.18 \% secured), it did not sample a
-representative proportion of the range (12 \% sampled). The addition of
-boundary length modifiers to just amount targets resulted in a solution
-which sampled even less of the species distribution (25.19 \% sampled;
-Fig. 3k). However, the addition of geographic space targets resolved
-these issues.
-
-The solution generated using 20 \% amount and 75 \% geographic targets
-included planning units from both populations (Fig. 3j). Although this
-solution required more planning units to fulfill both targets ($n=9$),
-this solution secured a representative sample of the bimodally
-distributed species (80.95 \% range sampled). Similar to corresponding
-solutions generated for both the uniformly (Fig. 3b) and normally
-distributed species (Fig. 3f), this solution was also highly fragmented
-and we could obtain a more well-connected solution at the expense of
-selecting additional planning units ($n=9$; Fig. 3l). The results for
-the bimodally distributed species suggest that species with highly
-structured populations or strong variation between individuals could
-benefit the most from prioritisations that are generated using
-space-based targets.
-
-\subsection{CASE STUDY 1}\label{case-study-1}
-
-\subsubsection{Methods}\label{methods-2}
-
-We investigated how space-based targets can be used in a multi-species
-planning context to generate a prioritisation that sufficiently
-preserves the realised niche for several species. By preserving the
-populations in suitable habitats with different with environmental
-conditions, conservation planners can preserve the species' adaptive
-landscape and foster resilience against environmental change (Moritz
-2002). We selected Queensland, Australia as the study area. This region
-is ideal for exploring the potential of niche-based targets because it
-contains a broad range of different habitats. We obtained data for 19
-bioclimatic variables across the region (at $30^{\prime \prime}$
-resolution from \url{www.worldclim.org}; Hijmans \emph{et al.} 2005) and
-subjected them to a principal components analysis (using ArcMap 10.3.1).
-We used the first two principal components (cumulatively explained 99.5
-\% of the total variation) to characterise the environmental variation
-across the study area (Fig. 4).
-
-We selected four bird species that span a range of different
-evolutionary histories, distributions and ecologies: blue-winged
-kookaburra (\emph{Dacelo leachii}), brown-backed honeyeater
-(\emph{Ramsayornis modestus}), brown falcon (\emph{Falco berigora}), and
-pale headed rosella (\emph{Platycercus adscitus}). We then mapped the
-extent of occurrence for each species (Fig. 5). To do this, we obtained
-observation data from the Atlas of Living Australia across the whole of
-Australia (using the \texttt{ALA4R R} package; Raymond \emph{et al.}
-2015), spatially thinned the data to omit points within 10 km$^2$ of
-each other to ameliorate the effects of sampling bias (using the
-developmental version of the \texttt{spThin R} package;
-\url{www.github.com/mlammens/spThin}; Aiello-Lammens \emph{et al.}
-2015), and fit 85 \% minimum convex polygons (using the
-\texttt{adehabitatHR R} package; Calenge 2006). We used this method
-because it is trivially reproducible using freely available data.
-
-We generated 500 demand points for each species (Fig. S2). To achieve
-this, for each species, we generated random points inside the species
-range and at each point extracted the principal component values at that
-location. We then fit hyperbox kernels to the distribution of these
-points to characterise the realised niche of each species (using a
-manually chosen bandwidth of 0.2 and a 0.5 quantile to map the core
-parts of the species' niches; implemented in the \texttt{hypervolume R}
-package; Blonder \emph{et al.} 2014). We then generated uniformly
-distributed points inside the species' kernels and estimated the density
-of the training points at the uniformly generated points. These
-uniformly distributed points and associated density estimated were used
-as demand point coordinates and weights (respectively).
-
-We generated two solutions. The first solution was generated using 20 \%
-amount-based targets for all species. The second solution was generated
-the same amount-based target with an additional 75 \% niche-based for
-each species.
-
-\subsubsection{Results}\label{results-1}
-
-Generally, the solution generated using amount targets preserved a
-representative sample of each the four bird species' niches--with one
-excpetion. Specifically, this solution sampled over 75 \% of the
-realised niche of blue-winged kookaburra (89.54 \%), brown falcon (92.31
-\%), and the pale-headed rosella (76.86 \%). Yet it failed to achieve
-this for Brown-backed honeyeater(30.93 \%). This result suggests that
-prioritisations generated using conventional methods may preserve a
-representative sample of most species realised niches (Figure S1).
-However, despite this, there may yet be species for which only a small
-fraction of their realised niche is preserved. By explictly using space
-targets, conservation planners can generate prioritisations that are
-guaranteed to capture a representative sample of species' niches.
-
-\subsection{CASE STUDY 2}\label{case-study-2}
-
-\subsubsection{Methods}\label{methods-3}
-
-Here we used space-based targets to generate a prioritisation securing a
-representative sample of a species' intra-specific genetic variation. We
-used species occurrence and multilocus AFLP data collected by the
-international IntraBioDiv project in the European Alps (Meirmans
-\emph{et al.} 2011; see Alvarez \emph{et al.} 2009 for further
-explanation of data collection methods). Althougthis dataset contains
-multiple plant species, we used data for the betony-leaved rampion
-(\textit{Phyteuma betonicifolium}), a wide-spread species with
-significant genetic structure (Aiello-Lammens \emph{et al.} 2015).
-Members of the IntraBioDiv project collected data using a $20^{\prime}$
-longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25
-km; Fig. 7a). They visited every second grid cell, and if the
-betony-leaved rampion was detected in a cell, samples were collected
-from three individuals. Samples were genotyped using amplified fragment
-length polymorphisms (AFLP; Raymond \emph{et al.} 2015), and used to
-construct matrices denoting the presence/absence of polymorphisms at
-loci. In total, 131 individuals were genotyped at 138 markers.
-
-We used non-metric multi-dimensional scaling (NMDS; using Gower
-distances to accommodate sparsity; Gower 1971; implemented the
-\texttt{cluster R} package; Maechler \emph{et al.} 2015) to ordinate the
-presence (or absence) of locus-specific alleles wthin individuals into
-two continuous variables (implemented in the \texttt{vegan R} package;
-Oksanen \emph{et al.} 2015). These continuous variables described the
-main axes of genetic variation within the species. We calculated the
-average of the values associated with individuals in each grid cell.
-These values were used to create a genetic attribute space (Figs. 7c and
-7d). To assess spatial auto-correlation, we calculated Moran's I
-auto-correlation index for each NMDS axis using inverse great circle
-distances based on the grid cells' centroids (using the \texttt{ape R}
-package; Paradis \emph{et al.} 2004).
-
-The grid cells were used as planning units. The grid-cell averaged
-ordinations were used to describe the typical genetic characteristics of
-individuals in the planning unit. Since the number of planning units was
-relatively small, we used the same grid-cell averaged ordinations as
-demand points. To ensure that the solutions did not prioritise
-particularly costly areas, we included opportunity cost data (Fig. 7b).
-We obtained population density data (1 km resolution from the Global
-Rural-Urban Mapping Project; \texttt{GRUMP V1}; CIESEN, Columbia
-University \emph{et al.} 2011) and estimated the total population
-density inside each grid cell.
-
-We generated two solutions. The first solution was generated using an 10
-\% amount-based target. The second solution was generated using the same
-amount-based target and an additional 95 \% genetic-based target.
-
-\subsubsection{Results}\label{results-2}
-
-The binary genetic data was ordinated into a two-dimensional space that
-described the main different between individuals (stress = 0.17). These
-values were then averaged to the planning unit level to describe the
-typical genetic characteristics of individuals in each planning unit
-(Figs. 7c--7d). Planning units located near each were found to contain
-individuals with similar genetic characteristics (Moran's I: NMDS axis
-1, $I = 0.4$, $P < 0.001$; Moran's I: NMDS axis 2, $I = 0.32$,
-$P < 0.001$). In terms of the average genetic characteristics of
-individuals in the planning units, they tended to cluster into two main
-groups, with evidence of within-group structure inside the larger group
-(Fig 8c). This analysis supports previous work by Alvarez \emph{et al.}
-(2009) who also found evidence of genetic structure within this species.
-
-The solution generated using just the amount target failed to preserve a
-representative portion of the species' genetic variation (15.54 \%
-sampled; Figs. 8a and 8c). This solution only sampled planning units
-that contained individuals belonging to one of the main two groups. On
-the other hand, the solution generated using additional genetic targets
-selected planning units belonging to both of the main groups. Note that
-since these solutions were generated using opportunity cost data, the
-solution generated using only the amount target is essentially the
-minimum number of least costly areas needed to fulfill the amount
-target. The only difference between the two solutions is a single
-planning unit. By swapping this single planning unit, the solution
-generated using amount and genetic targets was able to preserve a
-representative portion of the species' genetic variation 15.54 for only
-a minor increase in cost (99.41 total cost compared to 99.66 total
-cost).
-
-\section{Implications and future
-directions}\label{implications-and-future-directions}
-
-The \texttt{rapr} R package provides a unified approach to reserve
-selection. One of the key advantages of this package is that it is
-general enough to incorporate any spatial variable as an attribute
-space. This package can accommodate intrinsic or extrinsic variation to
-the feature(s). For example, adaptation processes could be secured using
-environmental variation or genetic variation among loci under selection.
-Ecological processes, such as predator-prey interactions, could be
-secured by capturing intra-specific trophic variation. Different
-attribute spaces with different targets can be also specified for
-different species. As long as the variation can be described using
-Euclidean distances, the \texttt{R} package can be used to obtain a
-representative sample (note that data may require transformation to
-conform to this assumption). This \textbackslash{}texttt\{R\} package
-provides decision makers with the tools to generate prioritisations that
-secure both biodiversity patterns and the processes that maintain
-biodiversity. Additionally, the package contains functionality to
-accommodate uncertainty in the distribution of features, and also
-identify suitably connected reserves. Both the simulated and case study
-species suggest that conservation planning exercises need to explicitly
-consider biodiversity processes during the reserve selection process to
-ensure they are captured.
-
-The degree to which a prioritisation truly secures a representative
-sample of a feature depends on the attribute spaces and distribution of
-demand points chosen by the decision maker. Ultimately, the space-based
-targets are set as a proportion based on the distribution of the
-demand-points. As a consequence, if the decision maker uses an
-inappropriate set of spatial variables to construct an attribute space,
-or an inappropriate set of demand points, then the optimal solution will
-not be an effective prioritisation. We therefore stress that decision
-makers must carefully consider which biodiversity processes need to be
-reflected in the prioritisation, and which spatial data can be used to
-represent these processes. Note that maximising one process can be
-detrimental to another. For instance, to maximise the geographic spread
-of a prioritisation, reserves need to be further away from each other,
-yet to maximise the connectivity of a prioritisation, reserves need to
-be closer to each other. To assist in the selection of appropriate
-demand points, the \texttt{R} package provides several routines for
-generating demand points (see the \texttt{make.DemandPoints} function).
-These routines essentially use the distribution of a feature in the
-attribute space to define a polygon. Demand points are then generated as
-random points within the polygon. A kernel is then fit to the
-distribution of the feature in the space (using Blonder \emph{et al.}
-2014; Duong 2015), and the demand points are weighted based on the
-estimated density of the feature at the demand points.
-
-The formulation requires spatially comprehensive data to map attribute
-spaces to planning units. For instance, in both case-studies, all
-planning units occupied by the species were associated with
-values/coordinates in the attribute spaces. However, most real-world
-data sets are patchy--some planning units will be occupied by species
-for which attribute space data is not available. To use such patchy data
-with the \textbackslash{}texttt\{rapr R\} package, the gaps in the data
-must first be filled in as a pre-processing step. Spatially explicit
-models could be used to estimate values in locations that are missing
-from a data set (eg. krigging or generalised dissimilarity models;
-Oliver \& Webster 1990; Ferrier 2002). This approach has been
-successfully applied to a range of biological data sets (Thomassen
-\emph{et al.} 2010).
-
-To maximise the long-term persistence of biodiversity, decision makers
-need to identify prioritisations that preserve existing patterns of
-biodiversity and the processes that support them. To achieve this,
-conservation planners need a decision support tool that can explicitly
-accommodate biodiversity patterns and processes. Here, we developed the
-\texttt{rapr R} package to fill this void. By exploring the
-functionality of this package using several simulated species, we found
-that including space-based targets can radically change a prioritisation
-for the simplest of species.
-
-\section{Acknowledgements}\label{acknowledgements}
-
-JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
-has an Australian Research Council Future Fellowship. This work was
-supported by the Centre of Excellence for Environmental Decisions (CEED)
-and the Landscape Ecology and Conservation Group (LEC) at The University
-of Queensland.
-
-\section{Data accessibility}\label{data-accessibility}
-
-All data, code, and results are stored in an online repository
-(\url{www.github.com/paleo13/rapr-manuscript}) to permit replication and
-validation of this study.
-
-\section{References}\label{references}
-
-\bibliography{references}
-
-Aiello-Lammens, M.E., Boria, R.A., Radosavljevic, A., Vilela, B. \&
-Anderson, R.P. (2015). spThin: an R package for spatial thinning of
-species occurrence records for use in ecological niche models.
-\emph{Ecography}, \textbf{38}, 541--545.
-
-Alvarez, N., Thiel-Egenter, C., Tribsch, A., Holderegger, R., Manel, S.,
-Sch{ö}nswetter, P., Taberlet, P., Brodbeck, S., Gaudeul, M., Gielly, L.,
-K{ü}pfer, P., Mansion, G., Negrini, R., Paun, O., Pellecchia, M., Rioux,
-D., Sch{ü}pfer, F., Van Loo, M., Winkler, M., Gugerli, F. \& IntraBioDiv
-Consortium. (2009). History or ecology? Substrate type as a major driver
-of patial genetic structure in Alpine plants. \emph{Ecology Letters},
-\textbf{12}, 632--640.
-
-Ball, I., Possingham, H. \& Watts, M.E. (2009). Marxan and relatives:
-software for spatial conservation prioritisation. \emph{Spatial
-conservation prioritisation: Quantitative methods \& computational
-tools} (eds A. Moilanen, K.A. Wilson \& H. Possingham), pp. 185--189.
-Oxford University Press, Oxford, UK.
-
-Beyer, H.L., Dujardin, Y., Watts, M.E. \& Possingham, H.P. (2016).
-Solving conservation planning problems with integer linear programming.
-\emph{Ecological Modelling}, \textbf{328}, 14--22.
-
-Blonder, B., Lamanna, C., Violle, C. \& Enquist, B.J. (2014). The
-\emph{n}-dimensional hypervolume. \emph{Global Ecology and
-Biogeography}, \textbf{23}, 595--609.
-
-CIESEN, Columbia University, International Food Policy Research
-Institute, The World Bank \& Centro Internacional de Agricultura
-Tropical. (2011). Global rural-urban mapping project, Version 1 (GRUMP
-v1): Urban extents grid.
-
-Calenge, C. (2006). The package adehabitat for the R software: tool for
-the analysis of space and habitat use by animals. \emph{Ecological
-Modelling}, \textbf{197}, 1035.
-
-Carvalho, S.B., Brito, J.C., Crespo, E.J. \& Possingham, H.P. (2011).
-Incorporating evolutionary processes into conservation planning using
-species distribution data: a case study with the western Mediterranean
-herpetofauna. \emph{Diversity \& Distributions}, \textbf{17}, 408--421.
-
-Ciarleglio, M., Wesley Barnes, J. \& Sarkar, S. (2009). ConsNet: new
-software for the selection of conservation area networks with spatial
-and multi-criteria analyses. \emph{Ecography}, \textbf{32}, 205--209.
-
-Cornu{é}jols, G., Nemhauser, G.L. \& Wolsey, L.A. (1990). The
-uncapacitated facility location problem. \emph{Discrete Location Theory}
-(eds P.B. Mirchandani \& R.L. Francis), pp. 119--171. Wiley, New York.
-
-Crandall, K.A., Bininda-Emonds, O.R.P., Mace, G.M. \& Wayne, R.K.
-(2000). Considering evolutionary processes in conservation biology.
-\emph{Trends in Ecology \& Evolution}, \textbf{15}, 290--295.
-
-Cui, T.T., Ouyang, Y.F. \& Shen, Z.J.M. (2010). Reliable facility
-location design under the risk of disruptions. \emph{Operations
-Research}, \textbf{58}, 998--1011.
-
-Duong, T. (2015). \emph{ks: Kernel Smoothing, version 1.9.4}.
-
-Faith, D.P. (2003). Environmental diversity (ED) as surrogate
-information for species-level biodiversity. \emph{Ecography},
-\textbf{26}, 374--379.
-
-Faith, D.P. \& Walker, P.A. (1996). Environmental diversity: on the
-best-possible use of surrogate data for assessing the relative
-biodiversity of sets of areas. \emph{Biodiversity \& Conservation},
-\textbf{5}, 399--415.
-
-Ferrier, S. (2002). Mapping Spatial Pattern in Biodiversity for Regional
-Conservation Planning: Where to from Here? \emph{Systematic Biology},
-\textbf{51}, 331--363.
-
-Gaston, K.J., Jackson, S.E., Cantu-Salazar, L. \& Cruz-Pinon, G.
-(\{2008\}). The Ecological Performance of Protected Areas. \emph{ANNUAL
-REVIEW OF ECOLOGY EVOLUTION AND SYSTEMATICS, Annual Review of Ecology
-Evolution and Systematics}, \textbf{39}, 93--113.
-
-Gower, J.C. (1971). A general coefficient of similarity and some of its
-properties. \emph{Biometrics}, \textbf{27}, 857--871.
-
-Hendry, A.P., Lohmann, L.G., Conti, E., Cracraft, J., Crandall, K.A.,
-Faith, D.P., Hauser, C., Joly, C.A., Kogure, K., Larigauderie, A.,
-Magallon, S., Moritz, C., Tillier, S., Zardoya, R., Prieur-Richard,
-A.H., Walther, B.A., Yahara, T. \& Donoghue, M.J. (2010). Evolutionary
-biology in biodiversity science, conservation, and policy: A call to
-action. \emph{Evolution}, \textbf{64}, 1517--1528.
-
-Hijmans, R.J., Cameron, S.E., Parra, J.L., Jones, P.G. \& Jarvis, A.
-(2005). Very high resolution interpolated climate surfaces for global
-land areas. \emph{International Journal of Climatology}, \textbf{25},
-1965--1978.
-
-Jones, K.E., Bielby, J., Cardillo, M., Fritz, S.A., O'Dell, J., Orme,
-C.D.L., Safi, K., Sechrest, W., Boakes, E.H., Carbone, C., Connolly, C.,
-Cutts, M.J., Foster, J.K., Grenyer, R., Habib, M., Plaster, C.A., Price,
-S.A., Rigby, E.A., Rist, J., Teacher, A., Bininda-Emonds, O.R.P.,
-Gittleman, J.L., Mace, G.M. \& Purvis, A. (2009). PanTHERIA: a
-species-level database of life history, ecology, and geography of extant
-and recently extinct mammals. \emph{Ecology}, \textbf{90}, 2648--2648.
-
-Klein, C., Wilson, K., Watts, M., Stein, J., Berry, S., Carwardine, J.,
-Smith, M.S., Mackey, B. \& Possingham, H. (2009). Incorporating
-ecological and evolutionary processes into continental-scale
-conservation planning. \emph{Ecological applications}, \textbf{19},
-206--217.
-
-Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M. \& Hornik, K.
-(2015). \emph{cluster: Cluster Analysis Basics and Extensions}.
-
-Margules, C.R. \& Pressey, R.L. (2000). Systematic conservation
-planning. \emph{Nature}, \textbf{405}, 243--253.
-
-McNeely, J.A. (1994). Protected areas for the 21st-century: Working to
-provide benefits to society. \emph{Biodiversity and Conservation},
-\textbf{3}, 390--405.
-
-Meirmans, P., Goudet, J., IntraBioDiv Consortium \& Gaggiotti, O.
-(2011). Ecology and life history affect different aspects of the
-population structure of 27 high-alpine plants. \emph{Molecular Ecology},
-\textbf{20}, 3144--3155.
-
-Moilanen, A. (2007). Landscape Zonation, benefit functions and
-target-based planning: unifying reserve selection strategies.
-\emph{Biological Conservation}, \textbf{134}, 571--579.
-
-Moritz, C. (1999). Conservation units and translocations: strategies for
-conserving evolutionary processes. \emph{Hereditas}, \textbf{130},
-217--228.
-
-Moritz, C. (1994). Defining evolutionarily significant units for
-conservation. \emph{Trends in Ecology \& Evolution}, \textbf{9},
-373--375.
-
-Moritz, C. (2002). Strategies to protect biological diversity and the
-evolutionary processes that sustain it. \emph{Systematic Biology},
-\textbf{51}, 238--254.
-
-Oksanen, J., Blanchet, F.G., Kindt, R., Legendre, P., Minchin, P.R.,
-O'Hara, R.B., Simpson, G.L., Solymos, P., Stevens, M.H.H. \& Wagner, H.
-(2015). \emph{vegan: community ecology package}.
-
-Oliver, M.A. \& Webster, R. (1990). Kriging: a method of interpolation
-for geographical information systems. \emph{International Journal of
-Geographical Information Systems}, \textbf{4}, 313--332.
-
-Paradis, E., Claude, J. \& Strimmer, K. (2004). APE: analyses of
-phylogenetics and evolution in R language. \emph{Bioinformatics},
-\textbf{20}, 289--290.
-
-Pyke, C.R. \& Fischer, D.T. (2005). Selection of bioclimatically
-representative biological reserve systems under climate change.
-\emph{Biological Conservation}, \textbf{121}, 429--441.
-
-{R Core Team}. (2016). R: A language and environment for statistical
-computing.
-
-Raymond, B., VanDerWal, J. \& Belbin, L. (2015). \emph{ALA4R: Atlas of
-Living Australia (ALA) Data and Resources in R}.
-
-Sanderson, E.W., Segan, D.B. \& Watson, J.E.M. (2015). Global status of
-and prospects for protection of terrestrial geophysical diversity.
-\emph{Conservation Biology}, \textbf{29}, 649--656.
-
-Thomassen, H.A., Buermann, W., Milá, B., Graham, C.H., Cameron, S.E.,
-Schneider, C.J., Pollinger, J.P., Saatchi, S., Wayne, R.K. \& Smith,
-T.B. (2010). Modeling environmentally associated morphological and
-genetic variation in a rainforest bird, and its application to
-conservation prioritization. \emph{Evolutionary Applications},
-\textbf{3}, 1--16.
+\clearpage
+\input{./figures.tex}
 

### 2017-04-06 code/rmarkdown/figures.tex
r pkgs

\section{Figures}\label{figures}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Attribute space example. This environmental attribute space has
dimensions relating to annual temperature (\(^{\circ}\)C) and rainfall
(mm). Letters denote the environmental conditions associated with the
geographic locations where four hypothetical populations are found.
Points represent demand points. In this space, populations close to each
other inhabit similar environmental conditions.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Distributions of three simulated species. Squares denote
planning units. Colors indicate probability of occupancy.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{Prioritizations for the simulation study. Each panel shows a
prioritization generated for a single species using a set of parameters.
Squares denote planning units. Dark green planning units were selected
for protection. Each row of panels show prioritizations generated for a
different species. Each column of panels corresponds to a different set
of parameters used to generate the prioritization.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{Two main gradients of climatic variation across Queensland,
Australia. Polygons denote planning units.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Distribution of the species used in the first case study.
Polygons denote planning units. Planning units occupied by a given
species are shown in light blue.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Prioritizations for the first case study. Polygons denote
planning units. Dark green planning units were selected for protection.
Panel (a) shows the solution generated when using 20 \% amount targets.
Panel (b) shows the solution when using 20 \% amount targets and -1e+06
\% niche targets.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{Data used for the second case study. Squares denote planning
units. Panel (a) shows all grid cells surveyed by the IntraBioDiv
project. Grid cells occupied by the betony-leaved rampion are shown in
bright blue. The subsequent panels contain only show occupied grid
cells. Panel (b) shows the acquisition cost of each planning unit
(estimated as the total human population density). Panels (c--d) show
the spatial distribution of the ordinations describing genetic
variation. These values describe the typical genetic characteristics of
individuals in each planning unit. Planning units with similar
values/colors contain individuals with similar loci polymorphisms.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{Prioritizations for the second case study. Panels (a--b) show
prioritizations generated using different parameters. Polygons denote
planning units. Dark green planning units were selected for protection.
Panel (a) shows the planning units selected when using 10 \% amount
targets. Panel (b) shows the planning units selected when using \%
amount targets and 85 \% genetic targets. Panels (c--d) show the
solutions in the genetic space. Each point corresponds to a planning
unit. The coordinates of the points represent the typical genetic
characteristics of individuals sampled in that planning unit (based on
an NMDS of the binary loci data). Planning units associated with points
that are closer together contain individuals with more similar genetic
characteristics than planning units that are further apart.}
\end{figure}

### 2017-04-06 code/rmarkdown/supporting-information.tex
r pkgs

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supporting Information},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Supporting Information}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

% load packages
\usepackage{amsmath,amsfonts,float,makecell,titletoc,titlesec,tocloft,lineno,booktabs,subfiles}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[doublespacing]{setspace}

% format captions
\usepackage[labelfont={small,bf}, labelsep=space, font={small}]{caption}

% line numbers
\linenumbers

% allow breaks in equations
\allowdisplaybreaks

% format abstract
\renewcommand{\abstractname}{Summary}
\renewenvironment{abstract}
 {\small
  \begin{center}
  \bfseries \abstractname\vspace{-.5em}\vspace{0pt}
  \end{center}
  \list{} {%
   \setlength{\leftmargin}{2mm}
   \setlength{\rightmargin}{\leftmargin}%
  }%
  \item\relax}
{\endlist}

% format section headers
\titleformat*{\section}{\Large\bfseries}
\titleformat{\subsection}[display]
	{\large\sffamily\lsstyle}
	{\subsectiontitlename\ \thesubsection}
	{0.5em}{}
\titleformat*{\subsubsection}{\large\itshape}

% make figures static
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
	\expandafter\origfigure\expandafter[H]
} {
	\endorigfigure
}

% define struts for tables
\newcommand\T{\rule{0pt}{2.6ex}} % top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % bottom strut

% define command to put new lines in table cells
%\newcommand{\specialcell}[2][c]{%
%  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\begin{document}
\maketitle

\setcounter{figure}{0} \setcounter{table}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\section{Appendix S1}\label{appendix-s1}

The reliable formulation explicitly considers the probability that
planning units are inhabited by features. As a consequence, it may
deliver prioritizations that will sufficiently represent an attribute
space even if the features do not inhabit several of the planning units
when the prioritization is implemented. This behavior is achieved by
siting back-up planning units near selected planning units with low
occupancy probabilities in the attribute space(s). To ensure that
prioritizations are robust against multiple planning units being
uninhabited, the problem assigns planning units at multiple backup
levels.

Backup levels levels are defined as \(r\)-levels (similar to failure
levels in Snyder \& Daskin 2005). The first backup \(r\)-level is used
to calculate the level of representation when all of the selected
planning units are occupied by all \(f \in F\). For this scenario, the
closest selected planning unit to each demand point \(i\) for attribute
space \(s\) is assigned at \(r\)-level\(=0\). This scenario essentially
represents \(Y_{fsij}\) in the unreliable formulation. The second backup
\(r\)-level is used to assess the level of representation when the
closest planning unit to each demand point \(i\) is unoccupied. For this
scenario, the second closest planning units are assigned at
\(r\)-level\(=1\). The third backup \(r\)-level is used to assess
representation when the first two closest planning units are unoccupied.
The third closest planning units are assigned at \(r\)-level\(=2\).
Continuing on, in this manner, the selected planning units in a
prioritization are assigned to each demand point \(i \in I\), attribute
space \(s \in S\), and each feature \(f \in F\) at an \(r\)-level.

A final backup \(r\)-level when \(r=R\) is used to assess the level of
representation when the features \(f \in F\) do not occupy any selected
planning units in a prioritisation. Each demand point \(i \in I\) for
each \(s \in S\) and \(f \in F\) is assigned to an ``imaginary''
planning unit \(j=J\) at \(r=R\). The distance variables associated with
this imaginary planning unit \(d_{fsiJ}\) denote the loss of biological
value associated with failing to secure a representative sample of
feature \(f\) in attribute space \(s\). However, the \(d\) variables are
in distance units which effectively meaningless in this context. Thus
these variables are calculated using a failure multiplier (\(M\)) and
the maximum distance between the planning units and the demand points
for \(f \in F\), \(s \in S\) (eqn 5).

\begin{align*}
& d_{fsiJ} = M \max\limits_{0 \leq i \leq I-1, 0 \leq j \leq J-1} d_{fsij} & \forall & 0 \leq f \leq F-1, \tag*{eqn 5} \\
& & & 0 \leq s \leq S-1\\
\end{align*}

Conservation planning problems often use more than several hundred
planning units. It is currently not be feasible to solve this problem
when considering all possible failure scenarios within a reasonable
amount of time (see Appendix S2). As a consequence, the \(R\) variable
can be any \(1 \leq R \leq J-1\). For instance, when \(R=3\) only 2
backup levels are considered in addition to the final backup level. Cui
\emph{et al.} (2010) found that \(R=5\) yields similar solutions to
\(R >> 5\). However, in most cases the decision maker will likely be
limited to \(R=1\) to obtain prioritizations in a feasible amount of
time.

The control variables for the reliable formulation are the \(B\) (eqn
1a), \(T_{s}\) (eqn 1b), \(\tau_{fs}\) (eqn 1c), \(R\), and \(M\)
variables.

\begin{align*}
R &= \parbox{25em}{number of failure levels} \tag*{eqn 6a} \\
%
M &= \parbox{25em}{failure multiplier} \tag*{eqn 6b}\\
%
\end{align*}

The decision variables are the \(X_j\) (eqn 2a), \(Y_{fsijr}\),
\(P_{fsijr}\) variables.

\begin{align*}
Y_{fsijr} &= \begin{cases}
    1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag*{eqn 7a} \\
    0, & \parbox{25em}{otherwise} \\
  \end{cases} \\
%
P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag*{eqn 7b}\\
%
\end{align*}

The reliable formulation (RRAP) is a multi-objective optimization
problem.

\begin{align*}
& \text{(RRAP)} & \text{Min } & \text{(3a)}\\
%
& & \text{s.t. } & \text{(3b)}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} P_{fsijr} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq T_{fs}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 8a}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 8b}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq r \leq R\\
%
& & & \sum_{r=0}^{R} Y_{fsijr} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 8c}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J\\
%
& & & \sum_{r=0}^{R-1} Y_{fsijr} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 8d}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & Y_{fsiJR} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 8e}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
& & & P_{fsij0} = q_{fj} & \forall & 0 \leq f \leq F-1, \tag*{eqn 8f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J\\
%
& & & P_{fsijr} = \left(1 - \right) \sum_{k=0}^{J-1} \frac{1 - q_k}{q_k} P_{f,s,i,k,r-1} Y_{f,s,i,k,r-1}  & \forall & 0 \leq f \leq F-1, \tag*{eqn 8g}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J,\\
& & & & & 1 \leq r \leq R\\
%
& & & X_j, Y_{fsijr} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 8h}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J,\\
& & & & & 0 \leq r \leq R\\
\end{align*}

The objective function for the reliable formulation is the same as for
the unreliable formation (eqn 3a). Similar to the unreliable
formulation, constraints (eqn 3b) and (eqn 8a) ensure that the
amount-based and space-based targets are met. Constraint (eqns 8b--7c)
ensure that each planning unit is only assigned to one backup
\(r\)-level for \(i \in I\). Constraints (eqn 8d) ensure that only
selected planning units are assigned to demand points \(i \in I\).
Constraints (eqn 8e) ensure that the imaginary planning unit is always
assigned to the highest backup \(r\)-level. Constraints (eqns 8f--8g)
determine the probability that planning unit \(j\) will be used to
sample demand point \(i \in I\) for \(s \in S\) and \(f \in F\) (see Cui
\emph{et al.} 2010 for more information). Constraints (eqn 8h) ensure
that the \(X\) and \(Y\) variables are binary.

The reliable formulation is non-linear. However, the non-linear
components can be linearized. First--as discussed in the main text--the
expression \(X_j X_k\) in (eqn 3a) can be linearized using methods
described by Beyer et al. (2016). Second, the expression
\(P_{fsijr} Y_{fsijr}\) in (eqn 8a) can be linearized using techniques
described by Sherali and Alameddine (1992) as implemented in Cui et al.
(2010).

\section{Appendix S2}\label{appendix-s2}

We conducted a benchmark analysis to understand how much time would be
required to build and solve different sized conservation planning
problems. Additionally, we also aimed to understand how the time
required to execute a problem varied depending on which formulation was
used (unreliable or reliable), and if fragmentation constraints were
used.

\subsection{METHODS}\label{methods}

We generated simulated conservation planning data sets using varying
number of features (c(10, 15, 20),c(10, 15, 20),c(10, 15, 20),c(10, 15,
20)) and planning units (c(10, 20, 30),c(10, 20, 30),c(10, 20, 30),c(10,
20, 30)). For a given number of planning units and features, the data
was simulated using the following steps. First, the planning units were
simulated as squares that were arranged in a grid. All planning unit
costs were set to one. The spatial distribution of features were
generated using binary, spatially autocorrelated random fields (using
the \texttt{RandomFields R} package; Schlather \emph{et al.} 2015,
2016). If features were associated with less than two occupied planning
units, these features were set to occupy two randomly selected planning
units. All occupied planning units were set to have a probability of
occupancy of 0.9. Similar to the simulated data sets used in the main
text, a geographic attribute space was used for each feature. The
features' demand points were set as the centroids of the planning units
that they occupied, and were weighted by the units' probability of
occupancy. All problems used amount-based targets of 20 \% and
space-based targets of 50

We solved each conservation planning data set using the unreliable and
reliable problem formulations (using default failure distance and level
parameter) using no boundary length modifier, and using a boundary
length modifier of 5. To assess variation in execution times, we
generated and solved 4 random data sets per combination of number of
features and planning units. All conservation planning problems were
solved to within 99 \% of optimality using Gurobi (version 7.0.0) on a
system with 64 GB RAM and an Intel(R) Xeon (R) CPU (E5-2630 v3; 2.40
GHz). All problems were solved using a single thread.

\subsection{RESULTS AND DISCUSSION}\label{results-and-discussion}

The benchmark analysis revealed that moderately-sized conservation
problems can be solved in a relatively short period of time (Figure S2).
As predicted, problems were solved much quicker when using the
unreliable formulation than when using the reliable formulation.
Additionally, the solve times varied substantially for different data
sets with the same properties (ie. number of features and planning
units). This result suggests that it may difficult to estimate exactly
how long a given problem will take to solve.

We were not able to explore how long large-sized problems would take to
solve because our computer system did not have enough memory. One method
to reduce the problem size would be to simply use less demand points.
Another method to reduce problem size would be to set a maximum distance
between demand points and planning units such that demand points cannot
be represented by planning units that that are too far away. Both of
these method could be applied using functions provided in the
\texttt{raptr R} package. Future research is needed to explore these
methods and understand how the impact solution quality. Additionally,
future research could involve implementing more advanced methods to
solve the optimization problems (eg. as described in Cui \emph{et al.}
2010; Aboolian \emph{et al.} 2013).

\clearpage

\section{Figures}\label{figures}

\begin{figure}
\centering
\includegraphics{supporting-information_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Attribute spaces used in the first case-study. Each panel shows
a the distribution of a solution in environmental space and how it
samples the realized niche for a different species. The left column of
panels shows the solution generated using amount targets. The right
column of panels shows the solution generated using amount and niche
targets. Each column of panels corresponds to a different species.
Hexagons show the distribution of demand points. The color of each
hexagon denotes the weighted frequency of demand points inside it.
Points represent the environmental conditions associated with planning
units inside the species geographic range that were selected for
preservation in a given solution.}
\end{figure}

\begin{figure}
\centering
\includegraphics{supporting-information_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Time required to solve different sized problems using different
parameters and formulations. Points represent average times and bars
show standard errors. Time is shown on a log\(_{10}\) scale.}
\end{figure}

\clearpage

\section{Tables}\label{tables}

\begin{table}[H]
\caption{Symbols and descriptions of terms used in the formulation of the unreliable representative and adequate prioritization (URAP) problem.\label{term}} 
\begin{center}
\begin{tabular}{cl}
\toprule[1pt]
\multicolumn{1}{c}{Symbol}&\multicolumn{1}{c}{Description}\tabularnewline
\midrule
$F$&\parbox[t][][t]{15cm}{set of biodiversity features (indexed by $f$)}\tabularnewline
$J$&\parbox[t][][t]{15cm}{set of planning units (indexed by $j$)}\tabularnewline
$q_{fj}$&\parbox[t][][t]{15cm}{probability of feature $f$ occupying planning unit $j$}\tabularnewline
$e_{jk}$&\parbox[t][][t]{15cm}{shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.}\tabularnewline
$S$&\parbox[t][][t]{15cm}{set of attribute spaces (indexed by $S$)}\tabularnewline
$I_{fsi}$&\parbox[t][][t]{15cm}{set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$}\tabularnewline
$\lambda_{fsi}$&\parbox[t][][t]{15cm}{set of weights for demand point $i$ for feature $f$ in attribute space $s$}\tabularnewline
$d_{fsij}$&\parbox[t][][t]{15cm}{distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$}\tabularnewline
$\delta_{fsi}$&\parbox[t][][t]{15cm}{the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$}\tabularnewline
$T_f$&\parbox[t][][t]{15cm}{amount target for feature $f$}\tabularnewline
$\tau_{fs}$&\parbox[t][][t]{15cm}{space-based target for feature $f$ in attribute space $s$}\tabularnewline
$X_j$&\parbox[t][][t]{15cm}{binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)}\tabularnewline
$Y_{fsij}$&\parbox[t][][t]{15cm}{binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determining the amount of the attribute space sampled by the selected planning units}\tabularnewline
\bottomrule[1pt]
\end{tabular}\end{center}
\end{table}

\clearpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-r15}{}
Aboolian, R., Cui, T.T. \& Shen, Z.J.M. (2013). An efficient approach
for solving reliable facility location models. \emph{Informs Journal on
Computing}, \textbf{25}, 720--729.

\hypertarget{ref-r426}{}
Beyer, H.L., Dujardin, Y., Watts, M.E. \& Possingham, H.P. (2016).
Solving conservation planning problems with integer linear programming.
\emph{Ecological Modelling}, \textbf{328}, 14--22.

\hypertarget{ref-r16}{}
Cui, T.T., Ouyang, Y.F. \& Shen, Z.J.M. (2010). Reliable facility
location design under the risk of disruptions. \emph{Operations
Research}, \textbf{58}, 998--1011.

\hypertarget{ref-r494}{}
Schlather, M., Malinowski, A., Menck, P.J., Oesting, M. \& Strokorb, K.
(2015). Analysis, simulation and prediction of multivariate random
fields with package RandomFields. \emph{Journal of Statistical
Software}, \textbf{63}, 1--25.

\hypertarget{ref-r493}{}
Schlather, M., Malinowski, A., Oesting, M., Boecker, D., Strokorb, K.,
Engelke, S., Martini, J., Ballani, F., Moreva, O., Auel, J., Menck,
P.J., Gross, S., Ober, U., Christoph Berreth, Burmeister, K., Manitz,
J., Ribeiro, P., Singleton, R., Pfaff, B. \& R Core Team. (2016).
RandomFields: Simulation and analysis of random fields. R package
version 3.1.24.1, \url{https://cran.r-project.org/package=RandomFields}.

\hypertarget{ref-r20}{}
Sherali, H. \& Alameddine, A. (1992). A new reformulation-linearization
technique for bilinear programming problems. \emph{Journal of Global
Optimization}, \textbf{2}, 379--410.

\hypertarget{ref-r18}{}
Snyder, L.V. \& Daskin, M.S. (2005). Reliability models for facility
location: the expected failure cost case. \emph{Transportation Science},
\textbf{39}, 400--416.

\end{document}

### 2017-04-06 code/rmarkdown/text.tex
r pkgs

\section{Introduction}\label{introduction}

Perhaps the most fundamental aim of conservation is to maximize the
long-term persistence of biodiversity (McNeely 1994; Margules \& Pressey
2000). To achieve this, conservation actions must preserve biodiversity
patterns (eg. populations, species, ecosystems), but also crucially the
processes that sustain them. One of the major tangible achievements of
modern conservation has been the act of setting aside areas for
preservation (Watson \emph{et al.} 2014; Sanderson \emph{et al.} 2015).
Reserve networks buffer species from gross threatening processes and set
the stage for enhanced management interventions (Gaston \emph{et al.}
2008). Since the resources available for conservation action are
limited, protected area networks must be sited in places that satisfy
conservation objectives for minimal cost (Margules \& Pressey 2000).

Today, the most widely used conservation planning tools focus on
biodiversity patterns (\texttt{Zonation} and \texttt{Marxan}; Moilanen
2007; Ball \emph{et al.} 2009). Decision makers can use these tools to
obtain solutions that secure a proportion of the geographic range of
each biological feature (populations, species, or ecosystems) of
interest by setting targets. One method to incorporate data on the
ecological and evolutionary processes that sustain biological features
is to conserve a representative sample of the physical attributes that
underpin these processes. To achieve this, current methods typically
involve partitioning each feature into sub-groups based on an attribute
variable that relates to a biodiversity process of interest (Cowling \&
Pressey 2001; Pyke \& Fischer 2005; Klein \emph{et al.} 2009). For
instance, by dividing species distributions into sub-groups according to
habitat discontinuities, and ensuring that each sub-group is represented
in the solution, conservation planners can obtain prioritizations that
promote adaptive processes (Carvalho \emph{et al.} 2011). However, using
this method is challenging because biodiversity often cannot be divided
into operational groups for conservation planning without loss of
information (Orians 1993; Pressey \& Logan 1994; Faith \& Walker 1996).
This limitation has been known of quite some time, and dates back to
some of the earliest reserve selection methods (Orians 1993).

To overcome this limitation, Faith and Walker (1993) developed a
decision support tool (\texttt{DIVERSITY}) to identify prioritizations
that secure a representative sample of the environmental variation found
across a study area. Although this method was originally pitched as an
alternative to species-based conservation planning (Faith 1994, 2003;
Faith \& Walker 1996), it can also identify solutions that conserve a
representative sample of variation across a single species' range.
Recent work has built on this reserve selection method by solving
problems with more advanced optimization algorithms (Engelbrecht
\emph{et al.} 2016). However, existing \texttt{DIVERSITY}-based methods
have limited utility for species-based conservation planning, in part,
because they cannot accommodate multiple biological features and do not
provide options to specify specific targets pertaining to the amount of
habitat required to conserve species (unlike \texttt{Marxan} and
\texttt{Zonation}). As a consequence, such methods are not useful for
most conservation planning exercises (Margules \& Pressey 2000).

Conservation planners lack a decision support tool that lets them set
explicit targets to obtain solutions that (1) secure an adequate amount
of habitat for each feature and (2) a representative sample of the
variation in each feature. To begin to fill this gap, here we unite the
ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new
formulations of the reserve selection problem, and implement them in the
\texttt{raptr R} package. This \texttt{R} package provides decision
makers with the tools to generate prioritizations based on data that
relate to biodiversity patterns and processes. Here, we aim to the
provide an in-depth understanding of this \texttt{R} package and explore
its functionality.

\section{Methods}\label{methods}

\subsection{PROBLEM FORMULATION}\label{problem-formulation}

Biodiversity features are defined as the entities that the
prioritization is required to preserve (eg. species, ecosystems). Here,
attributes are defined as the variation across the features' ranges that
the prioritization is required to sample. These attributes can be
intrinsic (eg. genetic or phenotypic) or extrinsic (eg. environmental
conditions) to the feature. There should, however, be a reasonable
underlying hypothesis that relates the attribute to the biodiversity
processes that needs to be conserved.

A set of attributes is conceptualized as a space (termed an ``attribute
space''). For example, a decision maker may require a prioritization
that represents populations along climatic gradients. To achieve this,
the decision maker might use a ``climatic'' attribute space with
dimensions relating to mean annual temperature (\(^{\circ}\)C) and
precipitation (mm). Any given combination of temperature and
precipitation may be conceived as a point in this climatic space.
Although they exist as polygons, for simplicity, each planning unit may
be thought to exist as a single point inside a given attribute space. By
associating the planning units with climatic data, and calculating a
descriptive statistic for each planning unit (eg. mean), they can be
mapped from geographic space to this climatic attribute space. In
addition to points representing planning units, attribute spaces also
contain demand points.

Demand points (Faith \& Walker 1993, 1996; Faith 2003) are designated by
the decision maker to indicate regions of the attribute space that they
wish to represent in the prioritization (see below for discussion on
generating demand points for real-world data sets). For instance, by
siting demand points throughout an attribute space, planners can obtain
solutions that conserve a representative sample of an attribute space.
Alternatively, by siting demand points in specific regions of an
attribute space, planners can obtain solutions that target specific
samples of an attribute space. For a given set of demand points, the
shorter the distances between the demand points and the planning units
selected for prioritization; the better a solution is at securing the
desired variation in the attribute space. In any attribute space there
may exist points that are impossible (eg. mean annual rainfall -5 mm),
or do not occur in the study area (eg. mean annual temperature
30\(^{\circ}\)C in Antarctica). Additionally, there may be some regions
that are desirable for some features and undesirable for others (eg.
conditions known to be outside the physiological tolerance of certain
species). Thus a different set of demand points may be appropriate for
different attribute spaces and features.

To illustrate these concepts, consider the following example: we wish to
develop a prioritization for a single species that has four populations.
Since we can only afford to preserve three of the four populations, our
objective is to conserve the most representative sample possible. To
achieve this goal, we obtained annual rainfall (mm) and temperature
(\(^{\circ}\)C) data at the location of each population. We used this
data to construct a two-dimensional climatic attribute space. Next, we
generated demand points as equi-distant points inside this space. By
comparing the distribution of the demand points to the distribution of
the populations in the attribute space, we can identify the optimal
solution (Fig. 1). We can see that a solution that prioritizes both
populations \emph{A} and \emph{B}, effectively ``doubles-up'' on the
same climatic characteristics and constitutes considerable redundancy if
both selected. Instead, a more representative sample of the
intra-specific variation could be captured by securing populations
\emph{A} (or \emph{B}), \emph{C}, and \emph{D}. However, if the goal of
the prioritization was to preserve populations living in warmer
temperatures, then instead of siting demand points across the full range
of conditions, we could site demand points in environmental conditions
with temperatures over 30 \(^{\circ}\)C (ie. the top two rows of demand
points in Fig. 1). Given this new set of demand points, populations
\emph{A}, \emph{B}, and \emph{C} would be prioritized. Since demand
points can be sited and weighted in any configuration, they provide a
flexible means to guide the reserve selection process.

The \texttt{raptr R} package utilizes two novel formulations of reserve
selection problem to generate prioritizations. These formulations are
based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and
uncapacitated facility location problems (Cornuéjols \emph{et al.}
1990). Since they are based on the unreliable and reliable facility
location problems (Cui \emph{et al.} 2010), the formulations are
hereafter referred to as the ``unreliable'' and ``reliable''
formulations. The difference between the two formulations is that the
reliable formulation explicitly considers the probability that planning
units are occupied by features when calculating how well a given
solution samples a feature's attribute space. On the other hand, the
unreliable formulation assumes that features occupy all of the planning
units within which they are found with 100 \% certainty when performing
these calculations. For brevity, we will define the simpler
formulation--the unreliable formulation--below and define the more
complex version--the reliable formulation--in Appendix S1. All
mathematical terms defined hereafter are described in Table S1. For
convenience, the cardinality of sets will be denoted using the same
symbol used to denote the set.

Define \(F\) to be the set of features one wishes to conserve (indexed
by \(f\)). Let \(J\) be a set of planning units (indexed by \(j\)).
Also, let \(A_j\) denote the area, and \(C_j\) denote the cost of
preserving planning unit \(j \in J\). To assess the extent to which each
feature is secured in a given prioritization, let \(q_{fj}\) denote the
probability of feature \(f\) occupying planning unit \(j\). The level of
fragmentation associated with a prioritization is parameterized as the
total exposed boundary length (as in \texttt{Marxan}). Let the shared
edges between each planning unit \(j \in J\) and \(k \in J\) be
\(e_{jk}\).

Let \(S\) denote a set of attribute spaces (indexed by \(s\)). Each
\(j \in J\) is associated with spatially explicit data that represent
coordinates for each attribute space \(s \in S\). Let \(I_{fsi}\) denote
a set of demand points (indexed by \(i\)) for each feature \(f \in F\)
and each attribute space \(s \in S\). Let \(\lambda_{fsi}\) denote the
weighting for each demand point \(i \in I\), \(f \in F\) and
\(s \in S\). Let \(d_{fsij}\) denote the distance between each demand
point \(i \in I\) and each planning unit \(j \in J\) for each feature
\(f \in F\) and attribute space \(s \in S\). To describe the inherent
variation in the distribution of demand points for feature \(f\) and
space \(s\), let \(\delta_{fsi}\) denote the distance between each
demand point \(i \in I\) and the centroid of the demand points. Demand
points with greater weight \(\lambda_{fsi}\) are more important, and the
solutions that select planning units close to highly weighted demand
points will be more favorable. As a consequence, the decision maker will
need to choose an appropriate weighting for each demand point.

Targets are used to ensure that prioritizations sufficiently conserve
each feature. Amount-based targets specify the minimum amount of habitat
required for each feature to be adequately conserved (similar to those
used in \texttt{Marxan}). Let \(T_f\) denote the amount of area or
habitat that needs to be preserved for each feature \(f \in F\).
Space-based targets specify the minimum proportion of variation in the
demand points that needs to be secured for each feature to be
sufficiently conserved. Space-based targets are expressed as
proportions--instead of a sum of weighted distances--by scaling the sum
of weighted distances between the demand points and the selected
planning units in a solution relative to the distances between demand
points and the demand points' centroid. This scaling is conceptually
similar to that used in calculating the R\(^2\) statistic for
\emph{K}-means clustering analyses from the within sums of squares and
total sums of squares {[}Greenacre \& Primicerio (2014); page 106{]}.
Let \(\tau_{fs}\) denote the space-based targets for feature \(f \in F\)
and attribute space \(s \in S\). The control variables for the
unreliable formulation are the \(B\), \(T_{s}\), and \(\tau_{fs}\)
variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{fs} &= \parbox{25em}{space target for feature $f$ in attribute space $s$} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM) to penalize fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

The decision variables are the \(X_j\) and \(Y_{fsij}\) variables.

\begin{align*}
X_j
  &= \begin{cases}
    1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
    0, & \parbox{25em}{otherwise} \\
  \end{cases} \\
%
Y_{fsij} &= \begin{cases}
    1, & \parbox{25em}{if demand point $i$ is used to represent planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
    0, & \parbox{25em}{otherwise} \\
  \end{cases} \\
\end{align*}

Each demand point \(i \in I\) for feature \(f \in F\) and space
\(s \in S\) is represented by a single selected planning unit (ie. a
\(j \in J\) where \(X_j = 1\)). The degree to which a demand point \(i\)
is represented by a planning unit \(j\) is determined by the distance
between them (\(d_{fsij}\)). Generally--unless near zero space targets
are used so that the problem is effectively unconstrained by the
target--demand points are represented by their closest selected planning
units. In poorer quality solutions, demand points will be represented by
planning units that have larger distances between them. As a
consequence, the sum of the weighted distances between all of the demand
demand points and the planning units used to represent them will be
larger, and so, the solution will capture less of the variation
described by the demand points.

The unreliable formulation (URAP) is a defined as a multi-objective
optimization problem.

\begin{align*}
& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
%
& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
%
& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
& & & & & 0 \leq s \leq S-1\\
%
& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1\\
%
& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
& & & & & 0 \leq s \leq S-1, \\
& & & & & 0 \leq i \leq I-1,\\
& & & & & 0 \leq j \leq J-1\\
%
& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
& & & & & 0 \leq s \leq S-1,\\
& & & & & 0 \leq i \leq I-1\\
%
\end{align*}

The objective function (eqn 3a) determines the utility of a given
prioritization: a combination of the total cost of a prioritization and
how fragmented it is. Constraints (eqn 3b) ensure that all the
amount-based targets are met. Constraints (eqn 3c) ensure that all the
space-based targets are met for each feature and each attribute space.
For each feature and attribute space, the total weighted distance
between the demand points and their closest selected planning units is
calculated
(\(\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}\)).
This total weighted distance is then scaled by the inherent variation in
the demand points
(\(\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}\)). As previously
mentioned, the resulting fraction is used to calculate a proportion
conceptually similar to the \(R^2\) statistic used in \(k\)-means
clustering analysis. The constraints (eqn 3c) ensure that the proportion
of variation in the demand points secured in the solution must be equal
to or greater than the space-based target. Constraints (eqn 3d) ensure
that only one planning unit is assigned to each demand point.
Constraints (eqn 3e) ensure that demand points are only assigned to
selected planning units. Constraints (eqn 3f) ensure that the \(X\) and
\(Y\) variables are binary.

\subsection{Optimization}\label{optimization}

Although the reserve selection problems presented here are non-linear
(see Appendix S1 for reliable formulation), they can be linearized using
methods described by Beyer \emph{et al.} (2016) and Cui \emph{et al.}
(Cui \emph{et al.} 2010). The \texttt{raptr} R package provides
functions to express conservation planning data as linearized versions
of the optimization problems and solve them using the commercial
\texttt{Gurobi} software suite (\url{www.gurobi.com}). Presently,
academics can obtain a license at no cost.

\section{Examples}\label{examples}

To showcase the behavior of the unreliable formulation, we conducted a
simulation study and two case studies. To understand how long it would
take to solve various sized problems, we also conducted a benchmark
analysis (Appendix S2). We completed the analyses using \texttt{R}
(version 3.3.2; R Core Team 2016) and solved all optimization problems
to within 99 \% of optimality.

\subsection{SIMULATION STUDY}\label{simulation-study}

\subsubsection{Methods}\label{methods-1}

We simulated a hypothetical study area with square planning units
arranged in a \(10 \times 10\) grid (Fig. 2). We then simulated three
species to inhabit this study area. Firstly, we simulated a
hyper-generalist species (hereafter referred to as the ``uniformly
distributed species''). It occupied all planning units with equal
probability (Fig. 2a; eqn 4a). Secondly, we simulated a species with
simple habitat requirements (hereafter referred to as the ``normally
distributed species''; Fig. 2b). This species was most likely to be
found in planning units nearest to the center of the study area. It was
simulated using the density function of a multivariate normal
distribution (represented by \(\mathcal{N}\); eqn 4b). Thirdly, we
simulated a species two distinct populations (hereafter referred to as
the ``bimodally distributed species''; Fig. 2c). It was simulated using
the maximum density of two multivariate normal distributions (eqn 4c).
For a given species, planning unit occupancy was calculated using the
\((X, Y)\) coordinates of the units' centroids and the relevant
equation. We used a geographic attribute space to provide an intuitive
visualization of the solutions. Demand points were set as the planning
units' centroids and were weighted according to the units' probability
of occupancy.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
\mathcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4c} \\
%
\end{align*}

We generated four solutions for each species. Firstly, to represent
solutions generated using conventional methods (eg. \texttt{Marxan}), we
generated solutions using 20 \% amount targets. Secondly, to show how
the addition of space targets can affect solutions, we generated
solutions using 20 \% amount targets and 90 \% space targets to capture
the geographic spread of each species (hereafter referred to as
geographic spread targets). Thirdly, to represent solutions generated
using conventional planning methods that penalize for fragmentation, we
generated solutions using 20 \% amount targets and a boundary length
modifier of 2.5. Fourthly, to illustrate the combined effects of using
amount and space targets as well as penalties for fragmentation, we
generated solutions using 20 \% amount targets, 90 \% geographic spread
targets, and boundary length modifiers of 2.5.

\subsubsection{Uniformly distributed
species}\label{uniformly-distributed-species}

The solution generated for the uniformly distributed species using an
amount target prioritized 20 planning units near the southern end of
study area (Fig. 2a). For this species, all solutions containing this
number of planning units are optimal when using a 20 \% amount target.
We obtained this particular solution due to artifacts in the solver (eg.
seed for the random number generator, order of variables in the problem
matrix). This solution captured a small amount of the geographic spread
of the species (-23.64 \% sampled). In fact, the coverage was so poor
that the proportion was negative because the solution captured a less
representative sample than a solution containing one planning unit in
the center of the species' distribution.

By explicitly targeting representativeness, we obtained a solution that
captured the geographic spread of the uniform species (93.39 \% sampled;
Fig. 3b). Although this solution secured an adequate amount of habitat
and a representative sample of the species' geographic spread, this
solution was highly fragmented. However, by penalizing fragmentation
using a boundary length modifier, we were able to obtain a well
connected solution that met all of the objectives (Fig. 3d). This
solution prioritized the same number of planning units as the previous
solutions even though it is far superior. As we can see, under the
simplest of circumstances, reserve selection methods may not yield
solutions that secure a representative sample of features unless
constraints are used to guarantee this property.

\subsubsection{Normally distributed
species}\label{normally-distributed-species}

The solution generated for the normally distributed species using just
an amount target prioritized planning units in the species' core range
area (Fig. 3e). While this strategy is cost-effective for protecting
habitat, it was ineffective for securing a representative sample of the
species' geographic spread (-6.14 \% sampled) because it did not protect
any planning units along the periphery of the species' distribution. By
using amount and space targets, we were able to conserve an adequate
amount of habitat and also secure a representative sample of the
species' geographic spread (99.06 \% sampled; Fig. 3f). Similar to the
solutions for the uniformly distributed species, this solution was
highly fragmented and we were able to obtain a better connected solution
by specifying fragmentation penalties (Fig. 3h). However, unlike the
solutions for the uniformly distributed species, the solution generated
using an amount target, a space target, and fragmentation penalties
required more planning units than the other solutions. These results
suggest that solutions may need to prioritize more planning units to
meet additional conservation objectives.

\subsubsection{Bimodally distributed
species}\label{bimodally-distributed-species}

The solution generated for the bimodally distributed species using an
amount target only conserved individuals belonging to one of the two
populations (Fig. 3i). As a consequence, this solution did not secure a
representative sample of the species' geographic spread (8.09 \%
sampled). The addition of fragmentation penalties exacerbated this
issue, and resulted in a solution that sampled even less of the species'
geographic spread (8.09 \% sampled; Fig. 3k). However, this issue was
resolved by using a space target to obtain a solution that secured both
populations (99.39 \% geographic spread secured; Fig. 3l). This finding
suggests that species with large intra-specific variation could benefit
the most from prioritizations generated using space-based targets.

\subsection{CASE STUDY 1}\label{case-study-1}

\subsubsection{Methods}\label{methods-2}

We investigated how space-based targets can be used in a multi-species
planning context to generate a prioritization that sufficiently
preserves the species' realized niches. By preserving the populations in
suitable habitats with different environmental conditions, conservation
planners can preserve the species' adaptive landscape and foster
resilience to environmental change (Moritz 2002). We selected
Queensland, Australia as the study area. We obtained data for 19
bioclimatic variables across the region (at \(30^{\prime \prime}\)
resolution from \url{www.worldclim.org}; Hijmans \emph{et al.} 2005) and
subjected them to a principal components analysis (using ArcMap 10.3.1).
We used the scores from the first two principal components to
characterize the environmental variation across the study area
(explaining 99.5 \% of the total variation; Fig. 4).

We used four bird species in this case study: blue-winged kookaburra
(\emph{Dacelo leachii}), brown-backed honeyeater (\emph{Ramsayornis
modestus}), brown falcon (\emph{Falco berigora}), and pale headed
rosella (\emph{Platycercus adscitus}). These species span a range of
different habitat requirements. We mapped the extent of occurrence for
each species (Fig. 5). To do this, we obtained occurrence records from
the Atlas of Living Australia across the whole of Australia (using the
\texttt{ALA4R R} package; Raymond \emph{et al.} 2015), spatially thinned
the data to omit points within 10 km\(^2\) of each other to ameliorate
the effects of sampling bias (using the developmental version of the
\texttt{spThin R} package; \url{https://github.com/mlammens/spThin};
Aiello-Lammens \emph{et al.} 2015), and fit 85 \% minimum convex
polygons (using the \texttt{adehabitatHR R} package; Calenge 2006). We
used this method because it is entirely reproducible using freely
available data.

We generated 20 demand points for each species (Fig. S2). To ensure that
the demand points reflected the core parts of the species' realized
niches, we used the following method. For each species, we generated
random points inside the species' geographic range and at each point
extracted the principal component values at that location. We then fit
hyperbox kernels to the distribution of principal component values to
characterize the realized niche of each species using a manually chosen
bandwidth of 0.2 and a 0.5 quantile to map the core parts of the
species' niches (implemented in the \texttt{hypervolume R} package;
Blonder \emph{et al.} 2014). We then generated uniformly distributed
points inside the species' distribution in environmental space, and
extracted the kernel density at their locations. These uniformly
distributed points and associated density estimations were used as
demand point coordinates and weights (respectively).

We generated two multi-species prioritizations. The first solution was
generated using 20 \% amount targets for each species. The second
solution was generated the same amount targets with an additional
10\^{}\{6\} \% niche-based target for each species.

\subsubsection{Results}\label{results}

Generally, the solution generated using just amount targets preserved a
representative sample of each the four bird species' niches. (Figure
S1). This solution secured over 10\^{}\{6\} \% of the realized niche of
blue-winged kookaburra (84.51 \%), brown falcon (93.51 \%), brown-backed
honeyeater (-29.33 \%), and the pale-headed rosella (64.8 \%). Yet it
failed to achieve this for the . This result demonstrates that although
conventional methods may yield solutions that conserve a representative
sample of some features, only through the use of explicit targets can
planners obtain cost-effective solutions that secure a representative
sample of all features.

\subsection{CASE STUDY 2}\label{case-study-2}

\subsubsection{Methods}\label{methods-3}

Here we used space-based targets to generate a prioritization securing a
representative sample of a species' intra-specific genetic variation. We
used species occurrence and genetic data collected by the international
IntraBioDiv project in the European Alps (Gugerli \emph{et al.} 2008;
Alvarez \emph{et al.} 2009; Meirmans \emph{et al.} 2011). Although this
dataset contains multiple species, we used data for the betony-leaved
rampion (\textit{Phyteuma betonicifolium}) as a simple case study,
because exhibits significant inter-population genetic structure
(Meirmans \emph{et al.} 2011). Members of the IntraBioDiv project
collected data using a \(20^{\prime}\) longitude by \(21^{\prime}\)
latitude grid (approx. 22.3 km \(\times\) 25 km; Fig. 7a). They visited
every second grid cell, and if the species was detected in a cell,
samples were collected from three individuals. Samples were genotyped
using amplified fragment length polymorphisms (AFLP; Vos \emph{et al.}
1995), and used to construct matrices denoting the presence/absence of
polymorphisms at loci. In total, 131 individuals were genotyped at 138
markers.

We used non-metric multi-dimensional scaling (NMDS; using Gower
distances to accommodate sparsity; Gower 1971; implemented using the
\texttt{cluster R} package; Maechler \emph{et al.} 2015) to ordinate the
presence (or absence) of locus-specific alleles within individuals into
two continuous variables (implemented in the \texttt{vegan R} package;
Oksanen \emph{et al.} 2015). These continuous variables described the
main axes of genetic variation within the species. We averaged together
values corresponding to samples collected in the same grid cell, and
used them to create a genetic attribute space (Figs. 7c and 7d). To
assess spatial auto-correlation, we calculated Moran's I
auto-correlation index for each NMDS axis using scaled great circle
distances between the grid cells' centroids (using the \texttt{ape R}
package; Paradis \emph{et al.} 2004).

The grid cells were used as planning units for generating
prioritizations. Since data were not collected in every grid cell, and
the planning units were arranged in a checkerboard pattern. The
grid-cell averaged ordinations were used to describe the typical genetic
characteristics of individuals in the planning unit. Since the number of
planning units was relatively small, we used the same grid-cell averaged
ordinations as demand points. To ensure that the solutions did not
prioritize particularly costly areas, we obtained population density
data (1 km resolution from the Global Rural-Urban Mapping Project;
\texttt{GRUMP V1}; Center for International Earth Science Information
Network (CIESIN) \emph{et al.} 2011) and estimated the total population
density inside each grid cell. We used these values to represent
opportunity cost (Fig. 7b).

We generated two solutions. The first solution was generated using an 10
\% amount-based target. The second solution was generated using the same
amount-based target and an additional 85 \% genetic-based target.

\subsubsection{Results}\label{results-1}

A two-dimensional attribute genetic space was able to capture most of
the variation between individuals (stress = 0.17; Figs. 7c--7d).
Planning units located near each other contained individuals with
similar genetic characteristics (Moran's I: NMDS axis 1, \(I = 0.4\),
\(P < 0.001\); Moran's I: NMDS axis 2, \(I = 0.32\), \(P < 0.001\)). In
terms of the average genetic characteristics of individuals in the
planning units, they tended to cluster into two main groups, with
evidence of within-group structure inside the larger group (Fig 8c).
This analysis supports previous work by Alvarez \emph{et al.} (2009) who
also found evidence of genetic structure within this species.

The solution generated using just the amount target failed to preserve a
representative portion of the species' genetic variation (15.52 \%
sampled; Figs. 8a and 8c). This solution only conserved individuals in
one of the two main genetic groups. In fact, this solution is just a
collection of the cheapest planning units occupied by the species needed
to fullfil the target. Whereas, the solution generated using both amount
and genetic targets conserved individuals in each of the two main
genetic groups (99.96 \% sampled; Figs. 8b and 8d). Although the
solutions only differ by a single planning unit, by securing individuals
in both main genetic groups the second solution was able to secure a
much more representative sample of the species' genetic variation for
only a minor increase in cost (920.08 compared to 99.41 total cost).

\section{Implications and future
directions}\label{implications-and-future-directions}

The \texttt{raptr R} package provides a unified approach to reserve
selection. Conservation planners can use this \texttt{R} package to
generate prioritizations that secure intra- and inter-specific
biodiversity patterns. Both the simulations and case studies show that
explicitly targeting a representative sample of each feature in a
conservation planning exercise can substantially alter prioritizations.
Additionally, we found that prioritizations often needed to secure more
habitat to conserve a representative sample of each feature. Although it
may not be practical to conserve substantially more habitat than the
amount specified using amount-based targets (eg. 10 \% geographic range;
Margules \& Pressey 2000), we did find that even small increases in
reserve size could yield large gains.

One of the key advantages of this package is that it is general enough
to incorporate almost any spatially explicit variable. This package can
accommodate intrinsic or extrinsic variation in the feature(s). For
example, adaptation processes could be secured using environmental
variation (eg. Pyke \& Fischer 2005), and trophic processes could be
conserved by capturing the overlapping distributions of predator and
prey species (eg. Rayfield \emph{et al.} 2009; Chernomor \emph{et al.}
2015). Note that maximizing one process can result in trading off
against another (eg. geographic spread vs.~connectivity). Additionally,
space targets can be used to achieve representation for reasons that are
not related to biodiversity processes, such as obtaining a
geographically representative set of reserves so there is equitable
access to parks by people (eg. Moilanen \emph{et al.} 2013). As long as
the variation can be described using Euclidean distances--which could be
achieved through transformation or dimension reduction--the \texttt{R}
package can be used to obtain a representative sample.

The degree to which a prioritization truly secures a representative
sample of a feature depends on (1) the attribute space(s), (2) the
distribution of demand points chosen by the conservation planner, and
(3) the space target used. The optimal solution will not effectively
represent the decision maker's objectives if an inappropriate set of
spatial variables or demand points are used to represent the attribute
space. Demand points should be distributed across the full range of
variation in an attribute space to obtain solutions that secure a
representative sample of the attribute space (see the
\texttt{make.DemandPoints} function for statistical routines to achieve
this). As with conventional amount-based targets, planners will have to
ensure that space targets are high enough to fulfill conservation
objectives (eg. when setting targets to conserve intra-specific genetic
variation; Frankham \emph{et al.} 2014). We encourage conservation
planners to consult with experts to identify suitable targets.

Attribute data need to be spatially comprehensive to map planning units
to attribute spaces. However, many real-world data sets are patchy. For
example, due to constrained resources, genetic data may only be
available for some planning units. To use such patchy data, planners
could omit the planning units that are missing data, or estimate the
missing data using spatially explicit models (eg. krigging or
generalized dissimilarity models; Oliver \& Webster 1990; Ferrier
\emph{et al.} 2007). This approach has been successfully applied to a
range of biological data sets (eg. Thomassen \emph{et al.} 2010). We
caution that inaccuracy and uncertainty in data can negatively impact
prioritizations (Wilson \emph{et al.} 2005), and recommend that planners
ensure that data are sufficiently reliable or use the reliable
formulation (described in Appendix S1) to accommodate uncertainty into
the reserve selection process.

To maximize the long-term persistence of biodiversity, decision makers
need to identify prioritizations that preserve existing patterns of
biodiversity and the processes that support them. It is becoming
increasingly apparent that simply protecting a large area will fail to
achieve this (Barnes 2015). Here, we developed the \texttt{raptr R}
package to provide conservation planners with the tools to deliver
cost-effective prioritizations that secure an adequate amount of a
representative sample of biodiversity features. By exploring the
functionality of this software package, we found that explicitly
targeting a representative sample of each feature can result in
substantially different solutions.

\section{Acknowledgements}\label{acknowledgements}

JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF
and HPP have Australian Research Council Fellowships. This work was
supported by the Centre for Biodiversity and Conservation Science
(CBCS).

\section{Data accessibility}\label{data-accessibility}

The \texttt{raptr R} package can be downloaded from The Comprehensive R
Archive Network (\url{https://CRAN.R-project.org/package=raptr}). All
data, code, and results are stored in an online repository
(\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit
replication and validation of this study.

\section{References}\label{references}

\bibliography{references}

\hypertarget{refs}{}
\hypertarget{ref-r454}{}
Aiello-Lammens, M.E., Boria, R.A., Radosavljevic, A., Vilela, B. \&
Anderson, R.P. (2015). spThin: an R package for spatial thinning of
species occurrence records for use in ecological niche models.
\emph{Ecography}, \textbf{38}, 541--545.

\hypertarget{ref-r461}{}
Alvarez, N., Thiel-Egenter, C., Tribsch, A., Holderegger, R., Manel, S.,
Schönswetter, P., Taberlet, P., Brodbeck, S., Gaudeul, M., Gielly, L.,
Küpfer, P., Mansion, G., Negrini, R., Paun, O., Pellecchia, M., Rioux,
D., Schüpfer, F., Van Loo, M., Winkler, M., Gugerli, F. \& IntraBioDiv
Consortium. (2009). History or ecology? Substrate type as a major driver
of patial genetic structure in alpine plants. \emph{Ecology Letters},
\textbf{12}, 632--640.

\hypertarget{ref-r22}{}
Ball, I., Possingham, H. \& Watts, M.E. (2009). Marxan and relatives:
software for spatial conservation prioritisation. \emph{Spatial
conservation prioritisation: Quantitative methods \& computational
tools} (eds A. Moilanen, K.A. Wilson \& H. Possingham), pp. 185--189.
Oxford University Press, Oxford, UK.

\hypertarget{ref-r492}{}
Barnes, M. (2015). Aichi targets: Protect biodiversity, not just area.
\emph{Nature}, \textbf{526}, 195--195.

\hypertarget{ref-r426}{}
Beyer, H.L., Dujardin, Y., Watts, M.E. \& Possingham, H.P. (2016).
Solving conservation planning problems with integer linear programming.
\emph{Ecological Modelling}, \textbf{328}, 14--22.

\hypertarget{ref-r231}{}
Blonder, B., Lamanna, C., Violle, C. \& Enquist, B.J. (2014). The
\emph{n}-dimensional hypervolume. \emph{Global Ecology and
Biogeography}, \textbf{23}, 595--609.

\hypertarget{ref-r455}{}
Calenge, C. (2006). The package adehabitat for the R software: tool for
the analysis of space and habitat use by animals. \emph{Ecological
Modelling}, \textbf{197}, 1035.

\hypertarget{ref-r2}{}
Carvalho, S.B., Brito, J.C., Crespo, E.J. \& Possingham, H.P. (2011).
Incorporating evolutionary processes into conservation planning using
species distribution data: A case study with the western mediterranean
herpetofauna. \emph{Diversity \& Distributions}, \textbf{17}, 408--421.

\hypertarget{ref-r456}{}
Center for International Earth Science Information Network (CIESIN),
International Food Policy Research Institute (IFPRI), The World Bank \&
Centro Internacional de Agricultura Tropical. (2011). \emph{Global
rural-urban mapping project, Version 1 (GRUMP V1): Urban extents grid}.
Socioeconomic Data and Applications Center (SEDAC), Palisades, NY,
\url{http://dx.doi.org/10.7927/H4GH9FVG}.

\hypertarget{ref-r485}{}
Chernomor, O., Minh, B.Q., Forest, F., Klaere, S., Ingram, T.,
Henzinger, M. \& Haeseler, A. von. (2015). Split diversity in
constrained conservation prioritization using integer linear
programming. \emph{Methods in Ecology and Evolution}, \textbf{6},
83--91.

\hypertarget{ref-r451}{}
Cornuéjols, G., Nemhauser, G.L. \& Wolsey, L.A. (1990). The
uncapacitated facility location problem. \emph{Discrete location theory}
(eds P.B. Mirchandani \& R.L. Francis), pp. 119--171. Wiley, New York.

\hypertarget{ref-r9}{}
Cowling, R.M. \& Pressey, R.L. (2001). Rapid plant diversification:
Planning for an evolutionary future. \emph{Proceedings of the National
Academy of Sciences}, \textbf{98}, 5452--5457.

\hypertarget{ref-r16}{}
Cui, T.T., Ouyang, Y.F. \& Shen, Z.J.M. (2010). Reliable facility
location design under the risk of disruptions. \emph{Operations
Research}, \textbf{58}, 998--1011.

\hypertarget{ref-r482}{}
Engelbrecht, I., Robertson, M., Stoltz, M. \& Joubert, J.W. (2016).
Reconsidering environmental diversity (ED) as a biodiversity surrogacy
strategy. \emph{Biological Conservation}, \textbf{197}, 171--179.

\hypertarget{ref-r203}{}
Faith, D.P. (2003). Environmental diversity (ED) as surrogate
information for species-level biodiversity. \emph{Ecography},
\textbf{26}, 374--379.

\hypertarget{ref-r481}{}
Faith, D.P. (1994). Phylogenetic pattern and the quantification of
organismal biodiversity. \emph{Philosophical Transactions: Biological
Sciences}, \textbf{345}, 45--58.

\hypertarget{ref-r480}{}
Faith, D.P. \& Walker, P.A. (1993). DIVERSITY: A software package for
sampling phylogenetic and environmental diversity. Reference and user's
guide (version 1.0),
\url{http://australianmuseum.com/uploads/documents/faith\%20diversity.pdf}.

\hypertarget{ref-r218}{}
Faith, D.P. \& Walker, P.A. (1996). Environmental diversity: On the
best-possible use of surrogate data for assessing the relative
biodiversity of sets of areas. \emph{Biodiversity \& Conservation},
\textbf{5}, 399--415.

\hypertarget{ref-r486}{}
Ferrier, S., Manion, G., Elith, J. \& Richardson, K. (2007). Using
generalized dissimilarity modelling to analyse and predict patterns of
beta diversity in regional biodiversity assessment. \emph{Diversity and
Distributions}, \textbf{13}, 252--264.

\hypertarget{ref-r487}{}
Frankham, R., Bradshaw, C.J. \& Brook, B.W. (2014). Genetics in
conservation management: Revised recommendations for the 50/500 rules,
red list criteria and population viability analyses. \emph{Biological
Conservation}, \textbf{170}, 56--63.

\hypertarget{ref-r465}{}
Gaston, K.J., Jackson, S.E., Cantu-Salazar, L. \& Cruz-Pinon, G. (2008).
The ecological performance of protected areas. \emph{Annual Review of
Ecology, Evolution and Systematics}, \textbf{39}, 93--113.

\hypertarget{ref-r459}{}
Gower, J.C. (1971). A general coefficient of similarity and some of its
properties. \emph{Biometrics}, \textbf{27}, 857--871.

\hypertarget{ref-r491}{}
Greenacre, M. \& Primicerio, R. (2014). \emph{Multivariate analysis of
ecological data}. BBVA Foundation, Madrid, Spain.

\hypertarget{ref-r463}{}
Gugerli, F., Englisch, T., Niklfeld, H., Tribsch, A., Mirek, Z.,
Ronikier, M., Zimmermann, N., Holderegger, R. \& Taberlet, P. (2008).
Relationships among levels of biodiversity and the relevance of
intraspecific diversity in conservation - a project synopsis.
\emph{Perspectives in Plant Ecology, Evolution and Systematics},
\textbf{10}, 259--281.

\hypertarget{ref-r33}{}
Hijmans, R.J., Cameron, S.E., Parra, J.L., Jones, P.G. \& Jarvis, A.
(2005). Very high resolution interpolated climate surfaces for global
land areas. \emph{International Journal of Climatology}, \textbf{25},
1965--1978.

\hypertarget{ref-r448}{}
Klein, C., Wilson, K., Watts, M., Stein, J., Berry, S., Carwardine, J.,
Smith, M.S., Mackey, B. \& Possingham, H. (2009). Incorporating
ecological and evolutionary processes into continental-scale
conservation planning. \emph{Ecological applications}, \textbf{19},
206--217.

\hypertarget{ref-r457}{}
Maechler, M., Rousseeuw, P., Struyf, A., Hubert, M. \& Hornik, K.
(2015). Cluster: Cluster analysis basics and extensions. R package
version 2.0.3, \url{http://cran.r-project.org/package=cluster}.

\hypertarget{ref-r255}{}
Margules, C.R. \& Pressey, R.L. (2000). Systematic conservation
planning. \emph{Nature}, \textbf{405}, 243--253.

\hypertarget{ref-r423}{}
McNeely, J.A. (1994). Protected areas for the 21st-century: Working to
provide benefits to society. \emph{Biodiversity and Conservation},
\textbf{3}, 390--405.

\hypertarget{ref-r462}{}
Meirmans, P., Goudet, J., IntraBioDiv Consortium \& Gaggiotti, O.
(2011). Ecology and life history affect different aspects of the
population structure of 27 high-alpine plants. \emph{Molecular Ecology},
\textbf{20}, 3144--3155.

\hypertarget{ref-r429}{}
Moilanen, A. (2007). Landscape Zonation, benefit functions and
target-based planning: unifying reserve selection strategies.
\emph{Biological Conservation}, \textbf{134}, 571--579.

\hypertarget{ref-r483}{}
Moilanen, A., Anderson, B.J., Arponen, A., Pouzols, F.M. \& Thomas, C.D.
(2013). Edge artefacts and lost performance in national versus
continental conservation priority areas. \emph{Diversity and
Distributions}, \textbf{19}, 171--183.

\hypertarget{ref-r63}{}
Moritz, C. (2002). Strategies to protect biological diversity and the
evolutionary processes that sustain it. \emph{Systematic Biology},
\textbf{51}, 238--254.

\hypertarget{ref-r458}{}
Oksanen, J., Blanchet, F.G., Kindt, R., Legendre, P., Minchin, P.R.,
O'Hara, R.B., Simpson, G.L., Solymos, P., Stevens, M.H.H. \& Wagner, H.
(2015). vegan: community ecology package. R package version 2.3-2,
\url{http://cran.r-project.org/package=vegan}.

\hypertarget{ref-r467}{}
Oliver, M.A. \& Webster, R. (1990). Kriging: A method of interpolation
for geographical information systems. \emph{International Journal of
Geographical Information Systems}, \textbf{4}, 313--332.

\hypertarget{ref-r489}{}
Orians, G.H. (1993). Endangered at what level? \emph{Ecological
Applications}, \textbf{3}, 206--208.

\hypertarget{ref-r464}{}
Paradis, E., Claude, J. \& Strimmer, K. (2004). APE: Analyses of
phylogenetics and evolution in r language. \emph{Bioinformatics},
\textbf{20}, 289--290.

\hypertarget{ref-r490}{}
Pressey, R.L. \& Logan, V.S. (1994). Level of geographical subdivision
and its effects on assessments of reserve coverage: A review of regional
studies. \emph{Conservation Biology}, \textbf{8}, 1037--1046.

\hypertarget{ref-r247}{}
Pyke, C.R. \& Fischer, D.T. (2005). Selection of bioclimatically
representative biological reserve systems under climate change.
\emph{Biological Conservation}, \textbf{121}, 429--441.

\hypertarget{ref-r192}{}
R Core Team. (2016). R: A language and environment for statistical
computing. \url{http://www.R-project.org}.

\hypertarget{ref-r471}{}
Rayfield, B., Moilanen, A. \& Fortin, M.-J. (2009). Incorporating
consumer-resource spatial interactions in reserve design.
\emph{Ecological Modelling}, \textbf{220}, 725--733.

\hypertarget{ref-r453}{}
Raymond, B., VanDerWal, J. \& Belbin, L. (2015). ALA4R: Atlas of Living
Australia (ALA) Data and Resources in R. R package version 1.25,
\url{https://github.com/AtlasOfLivingAustralia/ALA4R}.

\hypertarget{ref-r440}{}
Sanderson, E.W., Segan, D.B. \& Watson, J.E.M. (2015). Global status of
and prospects for protection of terrestrial geophysical diversity.
\emph{Conservation Biology}, \textbf{29}, 649--656.

\hypertarget{ref-r468}{}
Thomassen, H.A., Buermann, W., Milá, B., Graham, C.H., Cameron, S.E.,
Schneider, C.J., Pollinger, J.P., Saatchi, S., Wayne, R.K. \& Smith,
T.B. (2010). Modeling environmentally associated morphological and
genetic variation in a rainforest bird, and its application to
conservation prioritization. \emph{Evolutionary Applications},
\textbf{3}, 1--16.

\hypertarget{ref-r472}{}
Vos, P., Hogers, R., Bleeker, M., Reijans, M., Lee, T. van de, Hornes,
M., Friters, A., Pot, J., Paleman, J., Kuiper, M. \& Zabeau, M. (1995).
AFLP: a new technique for DNA fingerprinting. \emph{Nucleic Acids
Research}, \textbf{23}, 4407--4414.

\hypertarget{ref-r484}{}
Watson, J.E., Dudley, N., Segan, D.B. \& Hockings, M. (2014). The
performance and potential of protected areas. \emph{Nature},
\textbf{515}, 67--73.

\hypertarget{ref-r488}{}
Wilson, K.A., Westphal, M.I., Possingham, H.P. \& Elith, J. (2005).
Sensitivity of conservation planning to different approaches to using
predicted species distribution data. \emph{Biological Conservation},
\textbf{122}, 99--112.

### 2017-04-06 code/R/functions/extractResults.R

@@ -1,46 +1,0 @@
-#' Extract results from RapSolved object
-#'
-#' This function extracts results from a RapSolved object
-#'
-#' @param x \code{RapSolved} object.
-#' @export
-extractResults <- function(x) {
-	# extract score
-	score.DF <- data.frame(
-		Solution=rep(seq_along(x@results@summary$Score), each=length(x@data@species$name)),
-		Score=rep(x@results@summary$Score,each=length(x@data@species$name))
-	)
-	# extract amount held
-	amount.held.DF <- ldply(
-		seq_len(nrow(x@results@summary)),
-		function(r) {
-			data.frame(
-			Species=x@data@species$name,
-			amount.held=c(amount.held(x,y=r))
-		)
-	})
-	# extract space names
-	space.names <- sapply(x@data@attribute.spaces, 'slot', 'name')
-	# extract space held
-	space.held.DF <- ldply(
-		seq_len(nrow(x@results@summary)),
-		function(r) {
-		ldply(
-			x@data@species$name, 
-			function(y) {
-			# get space names
-			curr.DF <- as.data.frame(space.held(x, y=r, species=y))
-			names(curr.DF) <- space.names
-			# make data.frame
-			return(curr.DF)
-		})
-	})
-	# return data.frame
-	return(
-		cbind(
-			score.DF,
-			amount.held.DF,
-			space.held.DF
-		)
-	)
-}

### 2017-04-06 code/R/functions/extract_results.R

@@ -1,0 +1,36 @@
+#' Extract results from RapSolved object
+#'
+#' This function extracts results from a RapSolved object
+#'
+#' @param x \code{\link[raptr]{RapSolved}} object.
+#'
+#' @export
+extract_results <- function(x) {
+  # extract score
+  score_data <- data.frame(
+    Solution = rep(seq_along(x@results@summary$Score),
+                   each = length(x@data@species$name)),
+    Score = rep(x@results@summary$Score,
+                each = length(x@data@species$name)))
+  # extract amount held
+  amount_held_data <- plyr::ldply(
+    seq_len(nrow(x@results@summary)),
+    function(r) {
+      data.frame(Species = x@data@species$name,
+                 amount.held = c(raptr::amount.held(x, y = r)))
+  })
+  # extract space names
+  space_names <- sapply(x@data@attribute.spaces, "slot", "name")
+  # extract space held
+  space_held_data <- plyr::ldply(
+    seq_len(nrow(x@results@summary)), function(r) {
+      plyr::ldply(x@data@species$name, function(y) {
+        raptr::space.held(x, y = r, species = y) %>%
+        as.data.frame() %>%
+        magrittr::set_names(space_names[r])
+      })
+  })
+  # return data.frame
+  score_data %>%
+  cbind(amount_held_data, space_held_data)
+}

### 2017-04-06 code/R/functions/make.targets.R

@@ -1,48 +1,0 @@
-#' Make target data using attribute spaces
-#'
-#' This function takes a set of species names and 
-#' attribute spaces and generates a table with targets.
-#'
-#' @param species \code{character} species names.
-#' @param genetic.space \code{AttributeSpace} object with genetic space data.
-#' @param amount.target \code{numeric} amount-based target
-#' @param genetic.target \code{numeric} genetic-based target
-#' @seealso \code{\link[raptr]{AttributeSpace}}.
-#' @export
-make.targets <- function(species, genetic.spaces, amount.target=0.2, genetic.target=0.2) {
-	# amount targets
-	amount.targets.DF <- data.frame(
-		species=seq_along(species),
-		target=rep(0, length(species)),
-		proportion=rep(amount.target, length(species)),
-		name=paste0('amount_',species)
-	)
-	# genetic spaces
-	genetic.targets.DF <- ldply(seq_along(genetic.spaces), .fun=function(i) {
-		# get species position
-		spp.pos <- which(sapply(genetic.spaces[[i]]@demand.points, function(y) {length(y@weights)})>2)
-		# return data.frame
-		spp.DF <- data.frame(
-			species=seq_along(species),
-			target=rep(i, length(species)),
-			proportion=replace(
-				rep(NA_real_, length(species)),
-				spp.pos,
-				genetic.target
-			),
-			name=replace(
-				rep('null',length(species)),
-				spp.pos,
-				paste0('genetic_',species[spp.pos])
-			)
-		)
-		return(spp.DF)
-	})
-	# return
-	return(
-		mutate(rbind(amount.targets.DF, genetic.targets.DF),
-		target=as.integer(target), species=as.integer(species))
-	)
-}
-
-

### 2017-04-06 code/R/functions/make_genetic_attribute_space.R

@@ -1,0 +1,26 @@
+#' Make AttributeSpace using genetic data
+#'
+#' This function creates an \code{AttributeSpace} object with for a single species.
+#'
+#' @param site_data \code{data.frame} with coordinates for planning units.
+#'
+#' @param species_data \code{data.frame} with coordinates of demand points for the species.
+#'
+#' @return \code{AttributeSpace}.
+#'
+#' @seealso \code{\link[raptr]{AttributeSpace}}.
+#'
+#' @export
+make_genetic_attribute_space <- function(site_data, species_data, species) {
+  # extract ids of pus where the species is found
+  curr_ids <- which(rowSums(is.na(site_data)) == 0)
+  # return attribute space
+  raptr::AttributeSpace(
+    planning.unit.points = raptr::PlanningUnitPoints(
+      coords = as.matrix(site_data[curr_ids, , drop = FALSE]),
+      ids = curr_ids),
+    demand.points = raptr::DemandPoints(coords = as.matrix(species_data),
+                                        weights = rep(1, nrow(species_data))),
+    species = as.integer(species)
+  )
+}

### 2017-04-06 code/R/functions/make_targets.R

@@ -1,0 +1,43 @@
+#' Make target data using attribute spaces
+#'
+#' This function takes a set of species names and
+#' attribute spaces and generates a table with targets.
+#'
+#' @param species \code{character} species names.
+#'
+#' @param genetic_space \code{AttributeSpace} object with genetic space data.
+#'
+#' @param amount_target \code{numeric} amount-based target
+#'
+#' @param genetic_target \code{numeric} genetic-based target
+#'
+#' @seealso \code{\link[raptr]{AttributeSpace}}.
+#'
+#' @export
+make_targets <- function(species, genetic_spaces, amount_target = 0.2,
+                          genetic_target = 0.2) {
+  # amount targets
+  amount_targets_data <- data.frame(species = seq_along(species),
+                                    target = rep(0, length(species)),
+                                    proportion = rep(amount_target,
+                                                     length(species)),
+                                    name = paste0("amount_", species))
+  # genetic spaces
+  genetic_targets_data <- plyr::ldply(seq_along(genetic.spaces), function(i) {
+    # get species position
+    spp_pos <- which(sapply(genetic_spaces[[i]]@demand.points,
+                            function(y) length(y@weights)) > 2)
+    # return data.frame
+    spp_data <- data.frame(species = seq_along(species),
+                           target = rep(i, length(species)),
+                           proportion = replace(rep(NA_real_, length(species)),
+                                              spp.pos, genetic.target),
+                           name = replace(rep("null", length(species)),
+                                        spp.pos,
+                                        paste0("genetic_", species[spp.pos])))
+  })
+  # return targets
+  amount_targets_data %>%
+  rbind(genetic_targets_data) %>%
+  dplyr::mutate(target = as.integer(target), species = as.integer(species))
+}

### 2017-04-06 code/R/functions/species.prioritisation.R

@@ -1,20 +1,0 @@
-#' Generate prioritisation using specific targets
-#'
-#' This function is a wrapper for updating targets in a \code{RapUnsolved} object and solving it.
-#'
-#' @param x \code{RapUnsolved} object.
-#' @param amount.targets \code{numeric} amount-based targets to use.
-#' @param genetic.targets \code{numeric} genetic targets.
-#' @param ... arguments passed to solve.
-#' @return \code{RapSolved}
-species.prioritisation <- function(x, amount.targets, genetic.targets, ...) {
-	# init
-	genetic.pos <- grep('^genetic\\_.*$', x@data@targets$name)
-	# update targets
-	x@data@targets[which(x@data@targets$target==0),'proportion'] <- amount.targets
-	x@data@targets[genetic.pos,'proportion'] <- genetic.targets
-	# solve object
-	ret<-solve(x, ...)
-	return(ret)
-}
-

### 2017-04-06 code/R/functions/species_prioritisation.R

@@ -1,0 +1,25 @@
+#' Generate prioritisation using specific targets
+#'
+#' This function is a wrapper for updating targets in a
+#' \code{\link[raptr]{RapUnsolved}} object and solving it.
+#'
+#' @param x \code{RapUnsolved} object.
+#'
+#' @param amount.targets \code{numeric} amount-based targets to use.
+#'
+#' @param genetic.targets \code{numeric} genetic targets.
+#'
+#' @param ... arguments passed to solve.
+#'
+#' @return \code{\link[raptr]{RapSolved}}
+species_prioritisation <- function(x, amount_targets, genetic_targets, ...) {
+  # init
+  genetic_pos <- grep("^genetic\\_.*$", x@data@targets$name)
+  amount_pos <- which(x@data@targets$target == 0)
+  # update targets
+  x@data@targets[amount_pos, "proportion"] <- amount_targets
+  x@data@targets[genetic_pos, "proportion"] <- genetic_targets
+  # solve object
+  ret <- raptr::solve(x, ...)
+  return(ret)
+}

### 2017-04-06 code/R/analysis/00-initialization.R
add code to automatically install gurobi

@@ -15,6 +15,16 @@ library(magrittr)
 library(raster)
 library(sp)
 
+## Install gurobi R package if missing
+if (!requireNamespace("gurobi")) {
+  # find gurobi R package
+  gurobi_path <- dir("/opt", "gurobi", full.names = TRUE)
+  gurobi_path <- paste0(gurobi_path[length(gurobi_path)], "/linux64/R")
+  gurobi_path <- dir(gurobi_path, "gurobi", full.names = TRUE)[1]
+  # install pkgs
+  install.packages(gurobi_path)
+}
+
 ## Load parameters
 general_parameters <- RcppTOML::parseTOML("code/parameters/general.toml") %>%
                       `[[`(MODE)
@@ -22,7 +32,6 @@ general_parameters <- RcppTOML::parseTOML("code/parameters/general.toml") %>%
 
 ## Load functions
 for (x in dir(file.path("code", "R", "functions"), full.names = TRUE)) {
-  print(x)
   source(x)
 }
 

### 2017-04-15 code/rmarkdown/article.Rmd
massive overhaul - packrat to manage r packages - reformat code according to lintr - fix book references
use pngs for figs update makefile update toml file

@@
 ---
+
+```{r, include = FALSE}
+# set wd
+knitr::opts_knit$set(root.dir = normalizePath("../.."))
+```
+
+```{r, include = FALSE}
+## initialization
+# load .rda file
+try(session::restore.session("data/final/results.rda"))
+```
 
@@
 \clearpage
-\input{./text.tex}
 
-\clearpage
-\input{./figures.tex}
+# Introduction
+
+Perhaps the most fundamental aim of conservation is to maximize the long-term persistence of biodiversity [@r423; @r255].
+To achieve this, conservation actions must preserve biodiversity patterns (eg.
+populations, species, ecosystems), but also crucially the processes that sustain them.
+One of the major tangible achievements of modern conservation has been the act of setting aside areas for preservation [@r484; @r440].
+Reserve networks buffer species from gross threatening processes and set the stage for enhanced management interventions [@r465].
+Since the resources available for conservation action are limited, protected area networks must be sited in places that satisfy conservation objectives for minimal cost [@r255].
+
+Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22].
+Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets.
+One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes.
+To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448].
+For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
+However, using this method is challenging because biodiversity often cannot be divided into operational groups for conservation planning without loss of information [@r489; @r490; @r218].
+This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [@r489].
+
+To overcome this limitation, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of the environmental variation found across a study area.
+Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
+Recent work has built on this reserve selection method by solving problems with more advanced optimization algorithms [@r482].
+However, existing \texttt{DIVERSITY}-based methods have limited utility for species-based conservation planning, in part, because they cannot accommodate multiple biological features and do not provide options to specify specific targets pertaining to the amount of habitat required to conserve species (unlike \texttt{Marxan} and \texttt{Zonation}).
+As a consequence, such methods are not useful for most conservation planning exercises [@r255].
+
+Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of the variation in each feature.
+To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
+This \texttt{R} package provides decision makers with the tools to generate prioritizations based on data that relate to biodiversity patterns and processes.
+Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.
+
+# Methods
+## PROBLEM FORMULATION
+Biodiversity features are defined as the entities that the prioritization is required to preserve (eg.
+species, ecosystems).
+Here, attributes are defined as the variation across the features' ranges that the prioritization is required to sample.
+These attributes can be intrinsic (eg.
+genetic or phenotypic) or extrinsic (eg.
+environmental conditions) to the feature.
+There should, however, be a reasonable underlying hypothesis that relates the attribute to the biodiversity processes that needs to be conserved.
+
+A set of attributes is conceptualized as a space (termed an "attribute space").
+For example, a decision maker may require a prioritization that represents populations along climatic gradients.
+To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm).
+Any given combination of temperature and precipitation may be conceived as a point in this climatic space.
+Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given attribute space.
+By associating the planning units with climatic data, and calculating a descriptive statistic for each planning unit (eg.
+mean), they can be mapped from geographic space to this climatic attribute space.
+In addition to points representing planning units, attribute spaces also contain demand points.
+
+Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets).
+For instance, by siting demand points throughout an attribute space, planners can obtain solutions that conserve a representative sample of an attribute space.
+Alternatively, by siting demand points in specific regions of an attribute space, planners can obtain solutions that target specific samples of an attribute space.
+For a given set of demand points, the shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the desired variation in the attribute space.
+In any attribute space there may exist points that are impossible (eg.
+mean annual rainfall -5 mm), or do not occur in the study area (eg.
+mean annual temperature 30$^{\circ}$C in Antarctica).
+Additionally, there may be some regions that are desirable for some features and undesirable for others (eg.
+conditions known to be outside the physiological tolerance of certain species).
+Thus a different set of demand points may be appropriate for different attribute spaces and features.
+
+To illustrate these concepts, consider the following example: we wish to develop a prioritization for a single species that has four populations.
+Since we can only afford to preserve three of the four populations, our objective is to conserve the most representative sample possible.
+To achieve this goal, we obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population.
+We used this data to construct a two-dimensional climatic attribute space.
+Next, we generated demand points as equi-distant points inside this space.
+By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify the optimal solution (Fig.
+1).
+We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same climatic characteristics and constitutes considerable redundancy if both selected.
+Instead, a more representative sample of the intra-specific variation could be captured by securing populations _A_ (or _B_), _C_, and _D_.
+However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, we could site demand points in environmental conditions with temperatures over 30 $^{\circ}$C (ie.
+the top two rows of demand points in Fig.
+1).
+Given this new set of demand points, populations _A_, _B_, and _C_ would be prioritized.
+Since demand points can be sited and weighted in any configuration, they provide a flexible means to guide the reserve selection process.
+
+The \texttt{raptr R} package utilizes two novel formulations of reserve selection problem to generate prioritizations.
+These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
+Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations.
+The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a given solution samples a feature's attribute space.
+On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with 100 % certainty when performing these calculations.
+For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
+All mathematical terms defined hereafter are described in Table S1.
+For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.
+
+Define $F$ to be the set of features one wishes to conserve (indexed by $f$).
+Let $J$ be a set of planning units (indexed by $j$).
+Also, let $A_j$ denote the area, and $C_j$ denote the cost of preserving planning unit $j \in J$.
+To assess the extent to which each feature is secured in a given prioritization, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$.
+The level of fragmentation associated with a prioritization is parameterized as the total exposed boundary length (as in \texttt{Marxan}).
+Let the shared edges between each planning unit $j \in J$ and $k \in J$ be $e_{jk}$.
+
+Let $S$ denote a set of attribute spaces (indexed by $s$).
+Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$.
+Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$.
+Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$.
+Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$.
+To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.
+Demand points with greater weight $\lambda_{fsi}$ are more important, and the solutions that select planning units close to highly weighted demand points will be more favorable.
+As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.
+
+Targets are used to ensure that prioritizations sufficiently conserve each feature.
+Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}).
+Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$.
+Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved.
+Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid.
+This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [@r491; page 106].
+Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
+The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.
+
+\begin{align*}
+T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
+%
+\tau_{fs} &= \parbox{25em}{space target for feature $f$ in attribute space $s$} \tag*{eqn 1b}\\
+%
+B &= \text{boundary length modifier (BLM) to penalize fragmented solutions} \tag*{eqn 1c}\\
+\end{align*}
+
+The decision variables are the $X_j$ and $Y_{fsij}$ variables.
+
+\begin{align*}
+X_j
+  &= \begin{cases}
+    1, & \parbox{25em}{if planning unit $j$ is selected for conservation action} \tag*{eqn 2a} \\
+    0, & \parbox{25em}{otherwise} \\
+  \end{cases} \\
+%
+Y_{fsij} &= \begin{cases}
+    1, & \parbox{25em}{if demand point $i$ is used to represent planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
+    0, & \parbox{25em}{otherwise} \\
+  \end{cases} \\
+\end{align*}
+
+Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie.
+a $j \in J$ where $X_j = 1$).
+The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$).
+Generally--unless near zero space targets are used so that the problem is effectively unconstrained by the target--demand points are represented by their closest selected planning units.
+In poorer quality solutions, demand points will be represented by planning units that have larger distances between them.
+As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.
+
+The unreliable formulation (URAP) is a defined as a multi-objective optimization problem.
+
+\begin{align*}
+& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
+%
+& & \text{s.t. } & \sum_{j=0}^{J-1} A_j q_{fj} \geq T_{f} & \forall & 0 \leq f \leq F-1 \tag*{eqn 3b}\\
+%
+& & & 1 - \frac{\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}}{\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}} \geq \tau_{fs} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3c}\\
+& & & & & 0 \leq s \leq S-1\\
+%
+& & & \sum_{j=0}^{J-1} Y_{fsij} = 1 & \forall & 0 \leq f \leq F-1, \tag*{eqn 3d}\\
+& & & & & 0 \leq s \leq S-1, \\
+& & & & & 0 \leq i \leq I-1\\
+%
+& & & Y_{fsij} \leq X_j & \forall & 0 \leq f \leq F-1, \tag*{eqn 3e}\\
+& & & & & 0 \leq s \leq S-1, \\
+& & & & & 0 \leq i \leq I-1,\\
+& & & & & 0 \leq j \leq J-1\\
+%
+& & & X_j, Y_{fsij} \in {0,1} & \forall & 0 \leq f \leq F-1, \tag*{eqn 3f}\\
+& & & & & 0 \leq s \leq S-1,\\
+& & & & & 0 \leq i \leq I-1\\
+%
+\end{align*}
+
+The objective function (eqn 3a) determines the utility of a given prioritization: a combination of the total cost of a prioritization and how fragmented it is.
+Constraints (eqn 3b) ensure that all the amount-based targets are met.
+Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space.
+For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}$).
+This total weighted distance is then scaled by the inherent variation in the demand points ($\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}$).
+As previously mentioned, the resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
+The constraints (eqn 3c) ensure that the proportion of variation in the demand points secured in the solution must be equal to or greater than the space-based target.
+Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point.
+Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units.
+Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.
+
+## Optimization
+Although the reserve selection problems presented here are non-linear (see Appendix S1 for reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [@r16].
+The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
+Presently, academics can obtain a license at no cost.
+
+# Examples
+
+To showcase the behavior of the unreliable formulation, we conducted a simulation study and two case studies.
+To understand how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2).
+We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality.
+
+## SIMULATION STUDY
+### Methods
+We simulated a hypothetical study area with square planning units arranged in a $`r sqrt(sim_parameters$number_planning_units)` \times `r sqrt(sim_parameters$number_planning_units)`$ grid (Fig.
+2).
+We then simulated three species to inhabit this study area.
+Firstly, we simulated a hyper-generalist species (hereafter referred to as the "uniformly distributed species").
+It occupied all planning units with equal probability (Fig.
+2a; eqn 4a).
+Secondly, we simulated a species with simple habitat requirements (hereafter referred to as the "normally distributed species"; Fig.
+2b).
+This species was most likely to be found in planning units nearest to the center of the study area.
+It was simulated using the density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
+Thirdly, we simulated a species two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig.
+2c).
+It was simulated using the maximum density of two multivariate normal distributions (eqn 4c).
+For a given species, planning unit occupancy was calculated using the $(X, Y)$ coordinates of the units' centroids and the relevant equation.
+We used a geographic attribute space to provide an intuitive visualization of the solutions.
+Demand points were set as the planning units' centroids and were weighted according to the units' probability of occupancy.
+
+\begin{align*}
+\\
+P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
+%
+P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
+%
+P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
+\mathcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
+\end{cases} \tag*{eqn 4c} \\
+%
+\end{align*}
+
+We generated four solutions for each species.
+Firstly, to represent solutions generated using conventional methods (eg.
+\texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets.
+Secondly, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets).
+Thirdly, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`.
+Fourthly, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % geographic spread targets, and boundary length modifiers of `r sim_parameters$blm`.
+
+### Uniformly distributed species
+
+The solution generated for the uniformly distributed species using an amount target prioritized `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` planning units near the southern end of study area (Fig.
+2a).
+For this species, all solutions containing this number of planning units are optimal when using a `r sim_parameters$amount_target*100` % amount target.
+We obtained this particular solution due to artifacts in the solver (eg.
+seed for the random number generator, order of variables in the problem matrix).
+This solution captured a small amount of the geographic spread of the species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled).
+In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the center of the species' distribution.
+
+By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
+3b).
+Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented.
+However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (Fig.
+3d).
+This solution prioritized the same number of planning units as the previous solutions even though it is far superior.
+As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.
+
+### Normally distributed species
+
+The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core range area (Fig.
+3e).
+While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution.
+By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
+3f).
+Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (Fig.
+3h).
+However, unlike the solutions for the uniformly distributed species, the solution generated using an amount target, a space target, and fragmentation penalties required more planning units than the other solutions.
+These results suggest that solutions may need to prioritize more planning units to meet additional conservation objectives.
+
+### Bimodally distributed species
+
+The solution generated for the bimodally distributed species using an amount target only conserved individuals belonging to one of the two populations (Fig.
+3i).
+As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled).
+The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount target\nand BLM")$space_held, 2)` % sampled; Fig.
+3k).
+However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % geographic spread secured; Fig.
+3l).
+This finding suggests that species with large intra-specific variation could benefit the most from prioritizations generated using space-based targets.
+
+## CASE STUDY 1
+### Methods
+We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches.
+By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63].
+We selected Queensland, Australia as the study area.
+We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
+We used the scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig.
+4).
+
+We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_).
+These species span a range of different habitat requirements.
+We mapped the extent of occurrence for each species (Fig.
+5).
+To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+We used this method because it is entirely reproducible using freely available data.
+
+We generated `r cs1_parameters$dp_number` demand points for each species (Fig.
+S2).
+To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method.
+For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location.
+We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1_parameters$dp_bandwidth[1]` and a `r cs1_parameters$dp_quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231].
+We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations.
+These uniformly distributed points and associated density estimations were used as demand point coordinates and weights (respectively).
+
+We generated two multi-species prioritizations.
+The first solution was generated using `r cs1_parameters$amount_target*100` % amount targets for each species.
+The second solution was generated the same amount targets with an additional `r cs1_parameters$space_target*100` % niche-based target for each species.
+
+### Results
+Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches.
+(Figure S1).
+This solution secured over `r cs1_parameters$space_target*100` % of the realized niche of `r parsed_representative_space_held_names`.
+Yet it failed to achieve this for the `r parsed_not_representative_space_held_names`.
+This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.
+
+## CASE STUDY 2
+### Methods
+Here we used space-based targets to generate a prioritization securing a representative sample of a species' intra-specific genetic variation.
+We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
+Although this dataset contains multiple species, we used data for the `r tolower(cs2_parameters$common_name)` (\textit{`r gsub('_', ' ',cs2_parameters$species_name,fixed=TRUE)`}) as a simple case study, because exhibits significant inter-population genetic structure [@r462].
+Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
+22.3 km $\times$ 25 km; Fig.
+7a).
+They visited every second grid cell, and if the species was detected in a cell, samples were collected from three individuals.
+Samples were genotyped using amplified fragment length polymorphisms [AFLP; @r472], and used to construct matrices denoting the presence/absence of polymorphisms at loci.
+In total, `r structurer::n.samples(cs2_spp_structure_data[[1]])` individuals were genotyped at `r structurer::n.loci(cs2_spp_structure_data[[1]])` markers.
+
+We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented using the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458].
+These continuous variables described the main axes of genetic variation within the species.
+We averaged together values corresponding to samples collected in the same grid cell, and used them to create a genetic attribute space (Figs.
+7c and 7d).
+To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled  great circle distances between the grid cells' centroids [using the \texttt{ape R} package; @r464].
+
+The grid cells were used as planning units for generating prioritizations.
+Since data were not collected in every grid cell, and the planning units were arranged in a checkerboard pattern.
+The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
+Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points.
+To ensure that the solutions did not prioritize particularly costly areas, we obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
+We used these values to represent opportunity cost (Fig.
+7b).
+
+We generated two solutions.
+The first solution was generated using an `r cs2_parameters$amount_target*100` % amount-based target.
+The second solution was generated using the same amount-based target and an additional `r cs2_parameters$genetic_target*100` % genetic-based target.
+
+### Results
+
+A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2_spp_nmds[[1]]$stress,2)`; Figs.
+7c--7d).
+Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1_nmds_morans_i[[1]][[1]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1_nmds_morans_i[[1]][[2]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[2]]$p.value)`$).
+In terms of the average genetic characteristics of individuals in the planning units, they tended to cluster into two main groups, with evidence of within-group structure inside the larger group (Fig 8c).
+This analysis supports previous work by Alvarez _et al._ [-@r461] who also found evidence of genetic structure within this species.
+
+The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2_spp_results$genetic[1], 2)` % sampled; Figs.
+8a and 8c).
+This solution only conserved individuals in one of the two main genetic groups.
+In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target.
+Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs.
+8b and 8d).
+Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2_spp_results$Score[2],2)` compared to `r round(cs2_spp_results$Score[1],2)` total cost).
+
+# Implications and future directions
+The \texttt{raptr R} package provides a unified approach to reserve selection.
+Conservation planners can use this \texttt{R} package to generate prioritizations that secure intra- and inter-specific biodiversity patterns.
+Both the simulations and case studies show that explicitly targeting a representative sample of each feature in a conservation planning exercise can substantially alter prioritizations.
+Additionally, we found that prioritizations often needed to secure more habitat to conserve a representative sample of each feature.
+Although it may not be practical to conserve substantially more habitat than the amount specified using amount-based targets [eg.
+10 % geographic range; @r255], we did find that even small increases in reserve size could yield large gains.
+
+One of the key advantages of this package is that it is general enough to incorporate almost any spatially explicit variable.
+This package can accommodate intrinsic or extrinsic variation in the feature(s).
+For example, adaptation processes could be secured using environmental variation [eg.
+@r247], and trophic processes could be conserved by capturing the overlapping distributions of predator and prey species [eg.
+@r471; @r485].
+Note that maximizing one process can result in trading off against another (eg.
+geographic spread vs.
+connectivity).
+Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserves so there is equitable access to parks by people [eg.
+@r483].
+As long as the variation can be described using Euclidean distances--which could be achieved through transformation or dimension reduction--the \texttt{R} package can be used to obtain a representative sample.
+
+The degree to which a prioritization truly secures a representative sample of a feature depends on (1) the attribute space(s), (2) the distribution of demand points chosen by the conservation planner, and (3) the space target used.
+The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space.
+Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this).
+As with conventional amount-based targets, planners will have to ensure that space targets are high enough to fulfill conservation objectives [eg.
+when setting targets to conserve intra-specific genetic variation; @r487].
+We encourage conservation planners to consult with experts to identify suitable targets.
+
+Attribute data need to be spatially comprehensive to map planning units to attribute spaces.
+However, many real-world data sets are patchy.
+For example, due to constrained resources, genetic data may only be available for some planning units.
+To use such patchy data, planners could omit the planning units that are missing data, or estimate the missing data using spatially explicit models [eg.
+krigging or generalized dissimilarity models; @r467; @r486].
+This approach has been successfully applied to a range of biological data sets [eg.
+@r468].
+We caution that inaccuracy and uncertainty in data can negatively impact prioritizations [@r488], and recommend that planners ensure that data are sufficiently reliable or use the reliable formulation (described in Appendix S1) to accommodate uncertainty into the reserve selection process.
+
+To maximize the long-term persistence of biodiversity, decision makers need to identify prioritizations that preserve existing patterns of biodiversity and the processes that support them.
+It is becoming increasingly apparent that simply protecting a large area will fail to achieve this [@r492].
+Here, we developed the \texttt{raptr R} package to provide conservation planners with the tools to deliver cost-effective prioritizations that secure an adequate amount of a representative sample of biodiversity features.
+By exploring the functionality of this software package, we found that explicitly targeting a representative sample of each feature can result in substantially different solutions.
+
+# Acknowledgements
+JOH is funded by an Australian Postgraduate Award (APA) scholarship.
+RAF and HPP have Australian Research Council Fellowships.
+This work was supported by the Centre for Biodiversity and Conservation Science (CBCS).
+
+# Data accessibility
+The \texttt{raptr R} package can be downloaded from The Comprehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
+All data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
+
+# References
+
+\bibliography{references}

### 2017-04-15 code/rmarkdown/figures.Rmd
use pngs for figs update makefile update toml file

@@
     keep_tex: yes
+    includes:
+      in_header: preamble.tex
 ---
@@
 ```{r, include = FALSE}
-# set wd
 knitr::opts_knit$set(root.dir = normalizePath("../.."))
+knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo"), dpi = 500)
 ```
@@
     legend.direction = "horizontal",
-    legend.justification = c(0,1),
-    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
-                                                                  "mm"))) +
+    legend.justification = c(0,1)) +
   ggplot2::coord_cartesian(
@@
     legend.direction = "horizontal",
-    legend.justification = c(0,1),
-    legend.background = ggplot2::element_rect(size = ggplot2::unit(0.25,
-                                                                   "mm"))) +
+    legend.justification = c(0,1)) +
   ggplot2::coord_cartesian(
@@
     strip.text = ggplot2::element_text(color = "white", size = 8),
-    axis.title = ggplot2::element_blank(),
     legend.background = ggplot2::element_rect(color = "black", fill = "white"),

### 2017-04-15 code/rmarkdown/supporting-information.Rmd
use pngs for figs update makefile update toml file

@@
 ```{r, include=FALSE}
-# set wd
 knitr::opts_knit$set(root.dir = normalizePath("../.."))
+knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo"), dpi = 500)
 ```
@@
 
-We conducted a benchmark analysis to understand how much time would be required to build and solve different sized conservation planning problems.
-Additionally, we also aimed to understand how the time required to execute a problem varied depending on which formulation was used (unreliable or reliable), and if fragmentation constraints were used.
+We conducted a benchmark analysis to understand how much time would be required to solve different sized problems using the \texttt{raptr R} package.
 
@@
 
-We generated simulated conservation planning data sets using varying number of features (`r paste(benchmark_parameters$number_of_features, collapse = ",")`) and planning units (`r paste(benchmark_parameters$number_of_planning_units, collapse = ",")`).
-For a given number of planning units and features, the data was simulated using the following steps.
+We simulated conservation planning data sets containing different numbers of features and planning units (see Figure S2 for amounts).
+For a given number of planning units and features, a data set was simulated using the following steps.
 First, the planning units were simulated as squares that were arranged in a grid.
-All planning unit costs were set to one.
-The spatial distribution of features were generated using binary, spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @r494; @r493].
-If features were associated with less than two occupied planning units, these features were set to occupy two randomly selected planning units.
-All occupied planning units were set to have a probability of occupancy of `r benchmark_parameters$occupancy_probability`.
-Similar to the simulated data sets used in the main text, a geographic attribute space was used for each feature.
+Second, the planning unit costs were set to one.
+Third, the spatial distribution of each feature in the planning units were simulated using binary, spatially autocorrelated random fields [using the \texttt{RandomFields R} package; @r494; @r493].
+Fourth, if a feature was found to occupy less than two planning units, the feature was set to occupy two randomly selected planning units.
+All occupied planning units were set to have a probability of occupancy of `r benchmark_parameters$occupancy_probability` for all features.
+Fifth, similar to the simulated data sets described in the main text, a geographic attribute space was generated for each feature.
 The features' demand points were set as the centroids of the planning units that they occupied, and were weighted by the units' probability of occupancy.
-All problems used amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100`
+After simulating the data sets, solutions were generated using the \texttt{raptr R} package and Gurobi (version 7.0.2).
 
-We solved each conservation planning data set using the unreliable and reliable problem formulations (using default failure distance and level parameter) using no boundary length modifier, and using a boundary length modifier of `r max(benchmark_parameters$blm)`.
-To assess variation in execution times, we generated and solved `r benchmark_parameters$replicates` random data sets per combination of number of features and planning units.
-All conservation planning problems were solved to within `r general_parameters$MIPGap * 100` % of optimality using Gurobi (version 7.0.0) on a system with 64 GB RAM and an Intel(R) Xeon (R) CPU (E5-2630 v3; 2.40 GHz).
-All problems were solved using a single thread.
+We generated solutions for the conservation planning data sets using different formulations and parameters.
+Specifically, the solutions were generated using the unreliable and reliable formulation, and several boundary length multipliers (`r paste(unique(benchmark_parameters$blm), sep = "and")`).
+The solutions were generated using amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100`.
+To assess variation in solve times, we generated and solved `r benchmark_parameters$replicates` random data sets per combination of number of features and planning units.
+All problems were solved to within `r general_parameters$MIPGap * 100` % of optimality and were solved using a single thread on a system with 64 GB RAM and an Intel(R) Xeon (R) CPU (E5-2630 v3; 2.40 GHz).
 
@@
 The benchmark analysis revealed that moderately-sized conservation problems can be solved in a relatively short period of time (Figure S2).
-As predicted, problems were solved much quicker when using the unreliable formulation than when using the reliable formulation.
-Additionally, the solve times varied substantially for different data sets with the same properties (ie.
-number of features and planning units).
-This result suggests that it may difficult to estimate exactly how long a given problem will take to solve.
-
-We were not able to explore how long large-sized problems would take to solve because our computer system did not have enough memory.
-One method to reduce the problem size would be to simply use less demand points.
-Another method to reduce problem size would be to set a maximum distance between demand points and planning units such that demand points cannot be represented by planning units that that are too far away.
-Both of these method could be applied using functions provided in the \texttt{raptr R} package.
-Future research is needed to explore these methods and understand how the impact solution quality.
-Additionally, future research could involve implementing more advanced methods to solve the optimization problems [eg.
+Problems were solved much quicker when using the unreliable formulation than when using the reliable formulation.
+The solve times could be reduced by using more threads.
+We were not able to determine how long problems containing over 100 planning units would take to solve using the reliable formulation because the system did not contain enough would take to solve because our computer system did not have enough RAM.
+Future research could involve implementing more advanced methods to solve the optimization problems [eg.
 as described in @r16; @r15].

### 2017-04-24 code/rmarkdown/article.Rmd
assorted fixes

@@
 
-Abstract word count: 347 / 350
-Total word count: XXXX / 7000
-Number of references: 42
-Number of figures: 8
-Number of tables: 0
-Number of textboxes: 0
-Number of equations: 4
-Short running title:
-Short abstract: Prioritize biodiversity processes and patterns using raptr https://github.com/jeffreyhanson/raptr
-Keywords: conservation, biodiversity, mixed integer linear programming, optimization, protected areas
+Abstract word count: 344 / 350 \newline
+Total word count: XXXX / 7000 \newline
+Number of references: 42 \newline
+Number of figures: 8 \newline
+Number of tables: 0 \newline
+Number of textboxes: 0 \newline
+Number of equations: 4 \newline
+Short running title: \texttt{raptr} R package \newline
+Short abstract: Prioritize biodiversity processes and patterns using raptr https://github.com/jeffreyhanson/raptr  \newline
+Keywords: conservation, biodiversity, mixed integer linear programming, optimization, protected areas  \newline
 
@@
 individuals, areas) that is representative in terms of the intrinsic or extrinsic physical attributes that underpin the process of interest.
-For example, by conserving a sample of populations with local adaptations that is representative in terms of the full variation found in the species--physical attributes associated with adaptation--protected areas can foster adaptive processes by ensuring these adaptations are not lost.
+For example, by conserving a sample of populations with local adaptations--physical attributes associated with adaptation--that is representative of the range of adaptations found in the species, protected areas can maintain adaptive processes by ensuring these adaptations are not lost.
 Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature.
-Little attention has been directed towards capturing a representative sample of each feature.
+Little attention has been directed towards capturing a representative sample of the variation within each feature.
 
@@
 Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets.
-One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes.
+One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical internal or external attributes that underpin these processes.
 To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448].
 For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
-However, using this method is challenging because biodiversity often cannot be divided into operational groups for conservation planning without loss of information [@r489; @r490; @r218].
-This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [@r489].
+However, using this method is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490].
+This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [eg.
+@r295].
 
@@
 The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a given solution samples a feature's attribute space.
-On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with 100 % certainty when performing these calculations.
+On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with absolute certainty when performing these calculations.
 For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
@@
 Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid.
-This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [@r491; page 106].
+This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [page 106; @r491].
 Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
@@
 By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63].
-We selected Queensland, Australia as the study area.
+We selected Queensland, Australia as the study area, and used an equal area coordinate system for geosptial analyses (Australian Albers GDA94; \texttt{EPSG:3577}).
+We generated planning units by clipping a `r cs1_parameters$pu_size/1000` km$^2$ grid to the study area.
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
@@
 
-We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_).
+We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale-headed rosella (_Platycercus adscitus_).
 These species span a range of different habitat requirements.
@@
 ### Results
-Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches.
-(Figure S1).
+Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches (Figure S1).
 This solution secured over `r cs1_parameters$space_target*100` % of the realized niche of `r parsed_representative_space_held_names`.
@@
 7c and 7d).
-To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled  great circle distances between the grid cells' centroids [using the \texttt{ape R} package; @r464].
+To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances between the grid cells' centroids as weights [using the \texttt{ape R} package; @r464].
 
@@
 In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target.
-Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs.
+Alternatively, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs.
 8b and 8d).
@@
 Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this).
-As with conventional amount-based targets, planners will have to ensure that space targets are high enough to fulfill conservation objectives [eg.
+As with conventional amount-based targets, planners will need ensure that space targets are high enough to fulfill conservation objectives [eg.
 when setting targets to conserve intra-specific genetic variation; @r487].
+The most appropriate space target will depend on the species of conservation interest [eg.
+threatened species may more of their genetic variation to be conserved; @r495], and how conserving more variation relates to the species' persistence [eg.
+the relationship between inbreeding and extinction has a threshold; @r496].
 We encourage conservation planners to consult with experts to identify suitable targets.

### 2017-04-24 code/rmarkdown/figures.Rmd
assorted fixes

@@
 ---
+title: "Figures"
+output:
+  rmarkdown::pdf_document:
+    toc: false
+    keep_tex: yes
+    fig_caption: yes
+    includes:
+      in_header: preamble.tex
+fontsize: 11pt
+documentclass: article
+subparagraph: true
 bibliography: references.bib
 csl: reference-style.csl
-output:
-  rmarkdown::pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-    includes:
-      in_header: preamble.tex
 ---
@@
 
-```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = "Two main gradients of climatic variation across Queensland, Australia. Polygons denote planning units."}
+```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = "Two main gradients of climatic variation across Queensland, Australia.
+Polygons denote planning units.
+The map is rendered in an equal-area coordinate system (\texttt{EPSG:3577})"}
 # prepare data
@@
 ```{r, echo = FALSE, fig.height = 3.8, fig.width = column_width, fig.cap = "Distribution of the species used in the first case study.
-Polygons denote planning units.
+See Fig.
+4 caption for conventions.
 Planning units occupied by a given species are shown in light blue."}

### 2017-04-24 code/rmarkdown/supporting-information.Rmd
assorted fixes

@@
 Specifically, the solutions were generated using the unreliable and reliable formulation, and several boundary length multipliers (`r paste(unique(benchmark_parameters$blm), sep = "and")`).
-The solutions were generated using amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100`.
+The solutions were generated using amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100` %.
 To assess variation in solve times, we generated and solved `r benchmark_parameters$replicates` random data sets per combination of number of features and planning units.
@@
 Problems were solved much quicker when using the unreliable formulation than when using the reliable formulation.
-The solve times could be reduced by using more threads.
-We were not able to determine how long problems containing over 100 planning units would take to solve using the reliable formulation because the system did not contain enough would take to solve because our computer system did not have enough RAM.
+The solve times could be further reduced by using more threads.
+We were not able to determine how long problems bigger problems would take to solve using the reliable formulation because the system did not have enough RAM.
 Future research could involve implementing more advanced methods to solve the optimization problems [eg.

### 2017-04-24 code/R/analysis/06-statistical-analysis.R

@@ -73,14 +73,14 @@ if (length(amount_not_represented_species_names) > 1) {
      dplyr::last(parsed_not_representative_space_held_names))
 } else if (length(amount_not_represented_species_names) == 1) {
   parsed_not_representative_space_held_names <- paste0(
-    amount_not_represented_species_names, "(", round(dplyr::filter(
+    tolower(amount_not_represented_species_names), " (", round(dplyr::filter(
       cs1_spp_results, Prioritisation == "Amount targets",
       Species == amount_not_represented_species_names)$niche, 2), " %)")
 } else {
   parsed_not_representative_space_held_names <- c()
 }
 
-amount_represented_species_names  %<>% tolower()
+amount_represented_species_names %<>% tolower()
 amount_not_represented_species_names %<>% tolower()
 
 ## save session

### 2017-04-26 code/rmarkdown/article.Rmd
include word count

@@
 
-Abstract word count: 344 / 350 \newline
+Abstract word count: 341 / 350 \newline
 Total word count: XXXX / 7000 \newline
@@
 species, ecosystem) is to preserve a sample of the feature (eg.
-individuals, areas) that is representative in terms of the intrinsic or extrinsic physical attributes that underpin the process of interest.
+individuals, areas) that is representative of the intrinsic or extrinsic physical attributes that underpin the process of interest.
 For example, by conserving a sample of populations with local adaptations--physical attributes associated with adaptation--that is representative of the range of adaptations found in the species, protected areas can maintain adaptive processes by ensuring these adaptations are not lost.
@@
 Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets.
-One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical internal or external attributes that underpin these processes.
+One method to incorporate data on the ecological and evolutionary processes that sustain such features is to conserve a representative sample of the internal or external physical attributes underpinning these processes.
 To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448].
@@
 However, using this method is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490].
-This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [eg.
+This limitation has been known for quite some time, and dates back to some of the earliest reserve selection methods [eg.
 @r295].
@@
 Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of the variation in each feature.
-To begin to fill this gap, here we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
+To begin to fill this gap, we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
 This \texttt{R} package provides decision makers with the tools to generate prioritizations based on data that relate to biodiversity patterns and processes.
@@
 
-The \texttt{raptr R} package utilizes two novel formulations of reserve selection problem to generate prioritizations.
+The \texttt{raptr R} package utilizes two novel formulations of the reserve selection problem to generate prioritizations.
 These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
 Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations.
-The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a given solution samples a feature's attribute space.
-On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with absolute certainty when performing these calculations.
+The difference between the two formulations is that the reliable formulation explicitly accommodates uncertainty in the probability of features occupying the planning units when calculating how well a given solution samples a feature's attribute space.
 For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
@@
 To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.
-Demand points with greater weight $\lambda_{fsi}$ are more important, and the solutions that select planning units close to highly weighted demand points will be more favorable.
+Demand points with greater weight $\lambda_{fsi}$ are more important, so solutions that select planning units close to highly weighted demand points will be more favorable.
 As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.
@@
 To showcase the behavior of the unreliable formulation, we conducted a simulation study and two case studies.
-To understand how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2).
+To assess how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2).
 We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality.
@@
 It was simulated using the density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
-Thirdly, we simulated a species two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig.
+Thirdly, we simulated a species with two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig.
 2c).
@@
 We generated four solutions for each species.
-Firstly, to represent solutions generated using conventional methods (eg.
+First, to represent solutions generated using conventional methods (eg.
 \texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets.
-Secondly, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets).
-Thirdly, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`.
-Fourthly, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % geographic spread targets, and boundary length modifiers of `r sim_parameters$blm`.
+Second, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets).
+Third, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`.
+Finally, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % geographic spread targets, and boundary length modifiers of `r sim_parameters$blm`.
 
@@
 By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63].
-We selected Queensland, Australia as the study area, and used an equal area coordinate system for geosptial analyses (Australian Albers GDA94; \texttt{EPSG:3577}).
+We selected Queensland, Australia as the study area, and used an equal area coordinate system for geospatial analyses (Australian Albers GDA94; \texttt{EPSG:3577}).
 We generated planning units by clipping a `r cs1_parameters$pu_size/1000` km$^2$ grid to the study area.
@@
 S2).
-To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method.
-For each species, we generated random points inside the species' geographic range and at each point extracted the principal component values at that location.
+To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method: for each species we generated random points inside the species' geographic range and at each point extracted the principal component values at that location.
 We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1_parameters$dp_bandwidth[1]` and a `r cs1_parameters$dp_quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231].
@@
 The first solution was generated using `r cs1_parameters$amount_target*100` % amount targets for each species.
-The second solution was generated the same amount targets with an additional `r cs1_parameters$space_target*100` % niche-based target for each species.
+The second solution was generated using the same amount targets with an additional `r cs1_parameters$space_target*100` % niche-based target for each species.
 
@@
 We used species occurrence and genetic data collected by the international IntraBioDiv project in the European Alps [@r463; @r461; @r462].
-Although this dataset contains multiple species, we used data for the `r tolower(cs2_parameters$common_name)` (\textit{`r gsub('_', ' ',cs2_parameters$species_name,fixed=TRUE)`}) as a simple case study, because exhibits significant inter-population genetic structure [@r462].
+Although this dataset contains multiple species, we used data for the `r tolower(cs2_parameters$common_name)` (\textit{`r gsub('_', ' ',cs2_parameters$species_name,fixed=TRUE)`}) as a simple case study, because it exhibits significant inter-population genetic structure [@r462].
 Members of the IntraBioDiv project collected data using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx.
@@
 The grid cells were used as planning units for generating prioritizations.
-Since data were not collected in every grid cell, and the planning units were arranged in a checkerboard pattern.
+Since data were not collected in every grid cell, the planning units were arranged in a checkerboard pattern.
 The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
@@
 This solution only conserved individuals in one of the two main genetic groups.
-In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target.
+In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fulfill the target.
 Alternatively, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs.
@@
 The most appropriate space target will depend on the species of conservation interest [eg.
-threatened species may more of their genetic variation to be conserved; @r495], and how conserving more variation relates to the species' persistence [eg.
+threatened species may require more genetic variation to be conserved; @r495], and how conserving more variation relates to the species' persistence [eg.
 the relationship between inbreeding and extinction has a threshold; @r496].

### 2017-04-26 code/rmarkdown/article.Rmd
include discussion on practicality of preserving larger areas

@@
 The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space.
-Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this).
-As with conventional amount-based targets, planners will need ensure that space targets are high enough to fulfill conservation objectives [eg.
-when setting targets to conserve intra-specific genetic variation; @r487].
+Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample (see the `make.DemandPoints` function for statistical routines to achieve this).
+Although it may not always be practical to conserve a substantially larger area for some species, planners will need ensure that space targets are high enough to fulfill conservation objectives (eg.
+to secure genetic diversity as mandated in the Convention on Biological Diversity; CBD).
 The most appropriate space target will depend on the species of conservation interest [eg.

### 2017-04-27 code/rmarkdown/article.Rmd
updated with Rich's suggestions

@@
 Abstract word count: 341 / 350 \newline
-Total word count: XXXX / 7000 \newline
-Number of references: 42 \newline
+Total word count: 6931 / 7000 \newline
+Number of references: 49 \newline
 Number of figures: 8 \newline
@@
 This \texttt{R} package provides decision makers with the tools to generate prioritizations based on data that relate to biodiversity patterns and processes.
-Here, we aim to the provide an in-depth understanding of this \texttt{R} package and explore its functionality.
+Here, we aim to provide an in-depth understanding of this \texttt{R} package and explore its functionality.
 
@@
 We used this data to construct a two-dimensional climatic attribute space.
-Next, we generated demand points as equi-distant points inside this space.
+Next, we generated demand points as equidistant points inside this space.
 By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify the optimal solution (Fig.
@@
 
-The unreliable formulation (URAP) is a defined as a multi-objective optimization problem.
+The unreliable formulation of the representative and adequate prioritization problem (URAP) is a defined as a multi-objective optimization problem.
 
@@
 This total weighted distance is then scaled by the inherent variation in the demand points ($\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}$).
-As previously mentioned, the resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
+The resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
 The constraints (eqn 3c) ensure that the proportion of variation in the demand points secured in the solution must be equal to or greater than the space-based target.
@@
 ## Optimization
-Although the reserve selection problems presented here are non-linear (see Appendix S1 for reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [@r16].
+Although the reserve selection problems presented here are non-linear (see Appendix S1 for the reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [@r16].
 The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
@@
 We simulated a hypothetical study area with square planning units arranged in a $`r sqrt(sim_parameters$number_planning_units)` \times `r sqrt(sim_parameters$number_planning_units)`$ grid (Fig.
-2).
-We then simulated three species to inhabit this study area.
+2), and three species inhabiting this study area.
 Firstly, we simulated a hyper-generalist species (hereafter referred to as the "uniformly distributed species").
@@
 
-The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core range area (Fig.
+The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core distribution (Fig.
 3e).
@@
 
-The solution generated for the bimodally distributed species using an amount target only conserved individuals belonging to one of the two populations (Fig.
+The solution generated for the bimodally distributed species using just an amount target conserved individuals belonging to one of the two populations (Fig.
 3i).
@@
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
-We used the scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig.
+We used scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig.
 4).
@@
 ### Results
-Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches (Figure S1).
-This solution secured over `r cs1_parameters$space_target*100` % of the realized niche of `r parsed_representative_space_held_names`.
-Yet it failed to achieve this for the `r parsed_not_representative_space_held_names`.
+Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches (Fig.
+6a; left column Fig.
+S1).
+This solution secured a large proportion of the realized niche of `r parsed_representative_space_held_names`, but only a small proportion for the `r parsed_not_representative_space_held_names`.
+On the other hand, the solution generated using niche-based targets secured a large proportion of the realized niche for all four of the species (Fig.
+6b; right column Fig.
+S1).
 This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.
@@
 
-The grid cells were used as planning units for generating prioritizations.
-Since data were not collected in every grid cell, the planning units were arranged in a checkerboard pattern.
+The grid cells were used as planning units for generating prioritizations, noting that data were not collected in every grid cell, and so the planning units were arranged in a checkerboard pattern.
 The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
@@
 We generated two solutions.
-The first solution was generated using an `r cs2_parameters$amount_target*100` % amount-based target.
+The first solution was generated using a `r cs2_parameters$amount_target*100` % amount-based target.
 The second solution was generated using the same amount-based target and an additional `r cs2_parameters$genetic_target*100` % genetic-based target.
@@
 Additionally, we found that prioritizations often needed to secure more habitat to conserve a representative sample of each feature.
-Although it may not be practical to conserve substantially more habitat than the amount specified using amount-based targets [eg.
-10 % geographic range; @r255], we did find that even small increases in reserve size could yield large gains.
+Although it may not be practical to conserve substantially more habitat than the amount specified using amount-based targets, we did find that even small increases in reserve system size could yield large gains.
 
@@
 Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample (see the `make.DemandPoints` function for statistical routines to achieve this).
-Although it may not always be practical to conserve a substantially larger area for some species, planners will need ensure that space targets are high enough to fulfill conservation objectives (eg.
+As with conventional amount-based targets, planners will need to ensure that space targets are high enough to fulfill conservation objectives (eg.
 to secure genetic diversity as mandated in the Convention on Biological Diversity; CBD).
 The most appropriate space target will depend on the species of conservation interest [eg.
-threatened species may require more genetic variation to be conserved; @r495], and how conserving more variation relates to the species' persistence [eg.
+threatened species may need more genetic variation to be conserved; @r495], and how conserving more variation relates to the species' persistence [eg.
 the relationship between inbreeding and extinction has a threshold; @r496].
-We encourage conservation planners to consult with experts to identify suitable targets.
+We encourage conservation planners to consider carefully the relevant biological information to identify suitable targets.
 
@@
 To use such patchy data, planners could omit the planning units that are missing data, or estimate the missing data using spatially explicit models [eg.
-krigging or generalized dissimilarity models; @r467; @r486].
+kriging or generalized dissimilarity models; @r467; @r486].
 This approach has been successfully applied to a range of biological data sets [eg.

### 2017-04-27 code/rmarkdown/figures.Rmd
updated with Rich's suggestions

@@
 Polygons denote planning units.
-The map is rendered in an equal-area coordinate system (\texttt{EPSG:3577})"}
+The map is rendered in an equal-area coordinate system (\\texttt{EPSG:3577})."}
 # prepare data
@@
 These values describe the typical genetic characteristics of individuals in each planning unit.
-Planning units with similar values/colors contain individuals with similar loci polymorphisms.")}
+Planning units with similar values/colors contain individuals with similar loci polymorphisms.
+Note that data were not collected in every grid cell, and the planning units are therefore arranged in a checkerboard pattern.")}
 # prepare full grid data

### 2017-06-12 code/rmarkdown/article.Rmd
assorted improvements based on reviewer's suggestions

@@
 # load .rda file
-try(session::restore.session("data/final/results.rda"))
+try(session::restore.session("data/final/results.rda"), silent = TRUE)
 ```
 
-Abstract word count: 341 / 350 \newline
-Total word count: 6931 / 7000 \newline
+Abstract word count: 346 / 350 \newline
+Total word count: ??? / 7000 \newline
 Number of references: 49 \newline
@@
 individuals, areas) that is representative of the intrinsic or extrinsic physical attributes that underpin the process of interest.
-For example, by conserving a sample of populations with local adaptations--physical attributes associated with adaptation--that is representative of the range of adaptations found in the species, protected areas can maintain adaptive processes by ensuring these adaptations are not lost.
+For example, by conserving a sample of populations with local adaptations---physical attributes associated with adaptation---that is representative of the range of adaptations found in the species, protected areas can maintain adaptive processes by ensuring these adaptations are not lost.
 Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature.
@@
 To address this issue, we developed the \texttt{raptr R} package to help guide reserve selection.
-Users set "amount targets"--similar to conventional methods--to ensure that solutions secure a sufficient proportion of area or habitat for each feature.
+Users set "amount targets"---similar to conventional methods---to ensure that solutions secure a sufficient proportion of area or habitat for each feature.
 Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
@@
 We demonstrate the functionality of this package using simulations and two case studies.
-In these studies, we generated solutions using amount targets--representing conventional methods--and compared them with solutions generated using amount and space targets.
+In these studies, we generated solutions using amount targets---representing conventional methods---and compared them with solutions generated using amount and space targets.
 
@@
 4.
-The \texttt{raptr R} package provides a toolkit for making prioritizations that secure an adequate and representative sample of each feature.
+The \texttt{raptr R} package provides a toolkit for making prioritizations that secure an adequate and representative sample of variation within each feature.
 By using solutions that secure a representative sample of each feature, prioritizations may have a greater chance of achieving long-term biodiversity persistence.
@@
 Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22].
-Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets.
+Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest, by setting targets.
 One method to incorporate data on the ecological and evolutionary processes that sustain such features is to conserve a representative sample of the internal or external physical attributes underpinning these processes.
-To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448].
-For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
-However, using this method is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490].
+To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448] or phylogenetic trees [@r497].
+For instance, by dividing species distributions into sub-groups according to habitat discontinuities and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
+However, using these methods is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490].
 This limitation has been known for quite some time, and dates back to some of the earliest reserve selection methods [eg.
@@
 
-To overcome this limitation, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of the environmental variation found across a study area.
-Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range.
-Recent work has built on this reserve selection method by solving problems with more advanced optimization algorithms [@r482].
-However, existing \texttt{DIVERSITY}-based methods have limited utility for species-based conservation planning, in part, because they cannot accommodate multiple biological features and do not provide options to specify specific targets pertaining to the amount of habitat required to conserve species (unlike \texttt{Marxan} and \texttt{Zonation}).
-As a consequence, such methods are not useful for most conservation planning exercises [@r255].
+To overcome this limitation, Faith and Walker (1996) developed the "environmental diversity" (ED) reserve selection framework to conserve a representative sample of the environmental variation found across a study area.
+Drawing inspiration from the _p_-median problem [@r498], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; [@r204].
+Recent work has built on this method using more advanced optimization algorithms [@r482].
+However, existing environmental diversity based methods are not often used in conservation planning.
+This is, in part, because they seek to find the most representative solution that is within a budget, unlike the more commonly used decision support tools (eg.
+\texttt{Marxan}) that aim to find the cheapest solution subject to a set of targets (or constraints).
 
 Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of the variation in each feature.
-To begin to fill this gap, we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
+To begin to fill this gap, we unite the ideas underpinning environmental diversity and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package.
 This \texttt{R} package provides decision makers with the tools to generate prioritizations based on data that relate to biodiversity patterns and processes.
@@
 species, ecosystems).
-Here, attributes are defined as the variation across the features' ranges that the prioritization is required to sample.
+Each biodiversity feature may have one or more "attributes" which vary across its geographic range as a result of ecological or evolutionary processes.
 These attributes can be intrinsic (eg.
@@
 environmental conditions) to the feature.
-There should, however, be a reasonable underlying hypothesis that relates the attribute to the biodiversity processes that needs to be conserved.
+The overarching goal of the prioritization is to conserve these biodiversity processes, and by representing the variation in these species-specific attributes, the prioritization can make some progress towards achieving this.
+Thus there should be a reasonable underlying hypothesis that relates the attributes used to inform the reserve selection processes to the biodiversity process that needs to be conserved.
 
-A set of attributes is conceptualized as a space (termed an "attribute space").
+The variation in the attributes of a feature can be represented using an _n_-dimensional "attribute space".
 For example, a decision maker may require a prioritization that represents populations along climatic gradients.
@@
 mean), they can be mapped from geographic space to this climatic attribute space.
-In addition to points representing planning units, attribute spaces also contain demand points.
+In addition to points representing planning units, attribute spaces also contain points that describe which attributes should be conserved.
 
-Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets).
-For instance, by siting demand points throughout an attribute space, planners can obtain solutions that conserve a representative sample of an attribute space.
+"Demand points" [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets).
+For instance, by siting demand points throughout an attribute space, planners can obtain solutions that better capture existing variation.
 Alternatively, by siting demand points in specific regions of an attribute space, planners can obtain solutions that target specific samples of an attribute space.
-For a given set of demand points, the shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the desired variation in the attribute space.
-In any attribute space there may exist points that are impossible (eg.
-mean annual rainfall -5 mm), or do not occur in the study area (eg.
-mean annual temperature 30$^{\circ}$C in Antarctica).
+For a given set of demand points, the shorter the sum of the distances between the demand points and the planning units selected for prioritization: the better a solution is at securing the desired variation in the attribute space.
+Demand points are used heavily in the _p_-median and facility location problems [reviewed in @r498].
+Additionally, different features and attribute spaces may require different demand points.
+For instance, points that may be valid in one attribute space (eg.
+mean annual temperature -5 ${\circ}$C) may be invalid for another attribute space (eg.
+mean annual rainfall -5 mm).
 Additionally, there may be some regions that are desirable for some features and undesirable for others (eg.
 conditions known to be outside the physiological tolerance of certain species).
-Thus a different set of demand points may be appropriate for different attribute spaces and features.
 
@@
 1).
-We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same climatic characteristics and constitutes considerable redundancy if both selected.
+We can see that a solution that prioritizes both populations _A_ and _B_ constitutes considerable redundancy by selecting populations that inhabit similar conditions.
 Instead, a more representative sample of the intra-specific variation could be captured by securing populations _A_ (or _B_), _C_, and _D_.
@@
 The \texttt{raptr R} package utilizes two novel formulations of the reserve selection problem to generate prioritizations.
-These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451].
+These formulations are based on ideas that underpin the \texttt{Marxan}, environmental diversity, and uncapacitated facility location problems [@r498].
 Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations.
 The difference between the two formulations is that the reliable formulation explicitly accommodates uncertainty in the probability of features occupying the planning units when calculating how well a given solution samples a feature's attribute space.
-For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1.
+For brevity, we will define the simpler formulation---the unreliable formulation---below and define the more complex version---the reliable formulation---in Appendix S1.
 All mathematical terms defined hereafter are described in Table S1.
-For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.
+For convenience, the cardinality of a given set will be denoted using the same symbol used to denote the set.
 
@@
 To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.
-Demand points with greater weight $\lambda_{fsi}$ are more important, so solutions that select planning units close to highly weighted demand points will be more favorable.
-As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.
+
+Demand points with greater weight $\lambda_{fsi}$ are more important, and so solutions may need to select planning units closer to highly weighted demand points to fulfill space targets.
+Since increasing the number of demand points in the problem also increases the time required to solve it, conservation planners may find it more useful to increase of the weighting of demand points in important regions of an attribute space rather than increasing the number of demand points in the region.
+Decision maker will need to choose an appropriate weighting for each demand point to ensure that the resulting solution reflects their overarching goals.
 
@@
 Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved.
-Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid.
-This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [page 106; @r491].
+Space-based targets are expressed as proportions---instead of a sum of weighted distances---by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid.
+This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means cluster analyses from the within sums of squares and total sums of squares [page 106; @r491].
 Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
@@
 The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$).
-Generally--unless near zero space targets are used so that the problem is effectively unconstrained by the target--demand points are represented by their closest selected planning units.
-In poorer quality solutions, demand points will be represented by planning units that have larger distances between them.
+Generally---unless near zero space targets are used so that the problem is effectively unconstrained by the target---demand points are represented by their closest selected planning units.
+In poorer quality solutions, demand points will be represented by planning units that are further away from them.
 As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.
@@
 Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space.
+By treating targets as constraints, rather than including them in the objective function, all feasible solutions will fulfill the targets.
 For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}$).
@@
 The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
-Presently, academics can obtain a license at no cost.
+Presently, academics can obtain a special license for \texttt{Gurobi} at no cost.
 
@@
 \texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets.
-Second, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets).
+Second, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species.
 Third, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`.
-Finally, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % geographic spread targets, and boundary length modifiers of `r sim_parameters$blm`.
+Finally, to illustrate the combined effects of using amount and space targets as well as fragmentation penalties, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % space targets, and boundary length modifiers of `r sim_parameters$blm`.
 
@@
 
-The solution generated for the uniformly distributed species using an amount target prioritized `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` planning units near the southern end of study area (Fig.
-2a).
-For this species, all solutions containing this number of planning units are optimal when using a `r sim_parameters$amount_target*100` % amount target.
-We obtained this particular solution due to artifacts in the solver (eg.
-seed for the random number generator, order of variables in the problem matrix).
-This solution captured a small amount of the geographic spread of the species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled).
-In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the center of the species' distribution.
+The solution generated for the uniformly distributed species using an amount target prioritized `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` planning units (Fig.
+3a).
+For this scenario, all solutions containing `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` units are optimal because this species has an equal chance of occurring in any given planning unit.
+Although it may seem odd that the solution prioritized planning units along the southern end of the study area, this behavior occurred because there were many optimal solutions and the solver---without any other criteria to compare solutions---returned a solution that reflected the order that the data was encoded in the optimization problem.
+As we can see in later examples, this behavior only manifests when there are many optimal solutions.
+It is unlikely that conservation planners would encounter this behavior when using real-world data, and if they did, it would suggest that they need to obtain more data to inform the reserve selection process.
 
-By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
+The solution generated using just amount targets only captured a small proportion of the variation in the geographic attribute space (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled).
+The coverage was so poor that this "proportion" was negative because the solution captured a less representative sample than a solution containing one planning unit in the center of the species' distribution.
+In other words, the total distance between the demand points and the 20 selected planning units was larger than the total distance between the demand points and a single planning unit in the center of the study area.
+Similarly, $R^2$ statistics can be have negative values when they are describing models that perform worse than a null model.
+
+By explicitly targeting a representative sample of the species' geographic spread, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
 3b).
 Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented.
-However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (Fig.
+However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig.
 3d).
@@
 3e).
-While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution.
-By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
+Although this strategy may seem cost-effective, simply conserving the places where this species is most likely to be found is a poor strategy for securing a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled) because it did not protect any peripheral habitat.
+By using amount and space targets, we obtained a solution that conserved an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig.
 3f).
-Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (Fig.
+Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig.
 3h).
@@
 3k).
-However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % geographic spread secured; Fig.
+However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig.
 3l).
@@
 We selected Queensland, Australia as the study area, and used an equal area coordinate system for geospatial analyses (Australian Albers GDA94; \texttt{EPSG:3577}).
-We generated planning units by clipping a `r cs1_parameters$pu_size/1000` km$^2$ grid to the study area.
+To represent planning units for this case-study, we generated a `r cs1_parameters$pu_size/1000` $\times$ `r cs1_parameters$pu_size/1000` km$^2$ grid and clipped the square polygons to the state boundary.
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
@@
 5).
-To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
 We used this method because it is entirely reproducible using freely available data.
@@
 We generated `r cs1_parameters$dp_number` demand points for each species (Fig.
-S2).
+S1).
 To ensure that the demand points reflected the core parts of the species' realized niches, we used the following method: for each species we generated random points inside the species' geographic range and at each point extracted the principal component values at that location.
-We then fit hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1_parameters$dp_bandwidth[1]` and a `r cs1_parameters$dp_quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231].
+We then fitted hyperbox kernels to the distribution of principal component values to characterize the realized niche of each species using a manually chosen bandwidth of `r cs1_parameters$dp_bandwidth[1]` and a `r cs1_parameters$dp_quantile` quantile to map the core parts of the species' niches [implemented in the \texttt{hypervolume R} package; @r231].
 We then generated uniformly distributed points inside the species' distribution in environmental space, and extracted the kernel density at their locations.
@@
 The first solution was generated using `r cs1_parameters$amount_target*100` % amount targets for each species.
-The second solution was generated using the same amount targets with an additional `r cs1_parameters$space_target*100` % niche-based target for each species.
+The second solution was generated using the same amount targets with additional `r cs1_parameters$space_target*100` % space targets to capture a representative sample of the realized niche for each species.
 
@@
 This solution secured a large proportion of the realized niche of `r parsed_representative_space_held_names`, but only a small proportion for the `r parsed_not_representative_space_held_names`.
-On the other hand, the solution generated using niche-based targets secured a large proportion of the realized niche for all four of the species (Fig.
+On the other hand, the solution generated using space targets secured a large proportion of the realized niche for all four of the species (Fig.
 6b; right column Fig.
@@
 
-The grid cells were used as planning units for generating prioritizations, noting that data were not collected in every grid cell, and so the planning units were arranged in a checkerboard pattern.
-The grid-cell averaged ordinations were used to describe the typical genetic characteristics of individuals in the planning unit.
-Since the number of planning units was relatively small, we used the same grid-cell averaged ordinations as demand points.
+The grid cells were used as planning units for generating prioritizations, noting that data were not collected in every planning unit, and so the planning units were arranged in a checkerboard pattern.
+The spatially averaged ordinations were used to describe the typical genetic characteristics of individuals in each planning unit.
+Since the number of planning units was relatively small, we used the same spatially averaged ordinations as demand points.
 To ensure that the solutions did not prioritize particularly costly areas, we obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
@@
 The first solution was generated using a `r cs2_parameters$amount_target*100` % amount-based target.
-The second solution was generated using the same amount-based target and an additional `r cs2_parameters$genetic_target*100` % genetic-based target.
+The second solution was generated using the same amount-based target and an additional `r cs2_parameters$genetic_target*100` % space target to conserve genetic variation.
 
@@
 
-A two-dimensional attribute genetic space was able to capture most of the variation between individuals (stress = `r round(cs2_spp_nmds[[1]]$stress,2)`; Figs.
-7c--7d).
+A two-dimensional genetic attribute space was able to capture most of the variation between individuals (stress = `r round(cs2_spp_nmds[[1]]$stress,2)`; Figs.
+7c and 7d).
 Planning units located near each other contained individuals with similar genetic characteristics (Moran's I: NMDS axis 1, $I = `r round(cs1_nmds_morans_i[[1]][[1]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[1]]$p.value)`$; Moran's I: NMDS axis 2, $I = `r round(cs1_nmds_morans_i[[1]][[2]]$observed,2)`$, $P `r lazyWeave::pvalString(cs1_nmds_morans_i[[1]][[2]]$p.value)`$).
@@
 In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fulfill the target.
-Alternatively, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs.
+Alternatively, the solution generated using both amount and space targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs.
 8b and 8d).
@@
 @r471; @r485].
-Note that maximizing one process can result in trading off against another (eg.
-geographic spread vs.
-connectivity).
-Additionally, space targets can be used to achieve representation for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserves so there is equitable access to parks by people [eg.
+As is often the case with multi-objective conservation planning problems, trade-offs between different goals may be unavoidable.
+For example, maximizing geographic spread of sites will almost always have an impact on connectivity.
+Additionally, space targets can be used to secure a representative sample of features for reasons that are not related to biodiversity processes, such as obtaining a geographically representative set of reserves so there is equitable access to parks by people [eg.
 @r483].
-As long as the variation can be described using Euclidean distances--which could be achieved through transformation or dimension reduction--the \texttt{R} package can be used to obtain a representative sample.
+As long as the variation can be described using Euclidean distances---which could be achieved through transformation or dimension reduction---the \texttt{R} package can be used to obtain a representative sample.
 

### 2017-06-12 code/rmarkdown/figures.Rmd
assorted improvements based on reviewer's suggestions

@@
 Panel (a) shows the solution generated when using ",cs1_parameters$amount_target*100," % amount targets.
-Panel (b) shows the solution when using ", cs1_parameters$amount_target*100, " % amount targets and ", cs1_parameters$space_target*100, " % niche targets.")}
+Panel (b) shows the solution when using ", cs1_parameters$amount_target*100, " % amount targets and ", cs1_parameters$space_target*100, " % space targets.")}
 # prepare data
@@
 Panel (a) shows the planning units selected when using ",cs2_parameters$amount_target * 100," % amount targets.
-Panel (b) shows the planning units selected when using ",cs2_parameters$amount.target*100," % amount targets and ",cs2_parameters$genetic_target * 100," % genetic targets.
+Panel (b) shows the planning units selected when using ",cs2_parameters$amount.target*100," % amount targets and ",cs2_parameters$genetic_target * 100," % space targets.
 Panels (c--d) show the solutions in the genetic space.

### 2017-06-12 code/rmarkdown/supporting-information.Rmd
assorted improvements based on reviewer's suggestions

@@
 The left column of panels shows the solution generated using amount targets.
-The right column of panels shows the solution generated using amount and niche targets.
+The right column of panels shows the solution generated using amount and space targets.
 Each column of panels corresponds to a different species.
@@
 })
-cs1_pu_niche_data$Prioritisation <- factor(as.character(
-                                      cs1_pu_niche_data$Prioritisation),
-                                      levels = names(cs1_prioritisations))
-cs1_dp_niche_data$Prioritisation <- factor(as.character(
-                                      cs1_dp_niche_data$Prioritisation),
-                                      levels = names(cs1_prioritisations))
+
+cs1_pu_niche_data$Prioritisation[grepl("niche", as.character(cs1_pu_niche_data$Prioritisation))] <- "Amount & space targets"
+
+cs1_dp_niche_data$Prioritisation[grepl("niche", as.character(cs1_dp_niche_data$Prioritisation))] <- "Amount & space targets"
+
+cs1_pu_niche_data$Prioritisation <- factor(cs1_pu_niche_data$Prioritisation, levels = rev(unique(cs1_pu_niche_data$Prioritisation)))
+
+cs1_dp_niche_data$Prioritisation <- factor(cs1_dp_niche_data$Prioritisation, levels = rev(unique(cs1_dp_niche_data$Prioritisation)))
 

### 2017-06-13 code/rmarkdown/article.Rmd
update raptr text

@@
 Abstract word count: 346 / 350 \newline
-Total word count: ??? / 7000 \newline
-Number of references: 49 \newline
+Total word count: 6883 / 7000 \newline
+Number of references: 50 \newline
 Number of figures: 8 \newline
@@
 To overcome this limitation, Faith and Walker (1996) developed the "environmental diversity" (ED) reserve selection framework to conserve a representative sample of the environmental variation found across a study area.
-Drawing inspiration from the _p_-median problem [@r498], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; [@r204].
+Drawing inspiration from the _p_-median problem [@r498], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; @r204].
 Recent work has built on this method using more advanced optimization algorithms [@r482].
@@
 For instance, points that may be valid in one attribute space (eg.
-mean annual temperature -5 ${\circ}$C) may be invalid for another attribute space (eg.
+mean annual temperature $-5^{\circ}$C) may be invalid for another attribute space (eg.
 mean annual rainfall -5 mm).
@@
 
-Demand points with greater weight $\lambda_{fsi}$ are more important, and so solutions may need to select planning units closer to highly weighted demand points to fulfill space targets.
+Demand points with greater weight $\lambda_{fsi}$ are more important, and so solutions may need to select planning units closer to highly weighted demand points.
 Since increasing the number of demand points in the problem also increases the time required to solve it, conservation planners may find it more useful to increase of the weighting of demand points in important regions of an attribute space rather than increasing the number of demand points in the region.
@@
 Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$.
-Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved.
+Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature.
+These targets directly relate to a continuous and multi-dimensional attribute space and the set of demand points within it.
 Space-based targets are expressed as proportions---instead of a sum of weighted distances---by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid.
-This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means cluster analyses from the within sums of squares and total sums of squares [page 106; @r491].
+This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _k_-means cluster analyses from the within sums of squares and total sums of squares [page 106; @r491].
 Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
@@
 This total weighted distance is then scaled by the inherent variation in the demand points ($\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}$).
-The resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis.
-The constraints (eqn 3c) ensure that the proportion of variation in the demand points secured in the solution must be equal to or greater than the space-based target.
+The resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in _k_-means cluster analysis.
 Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point.
@@
 ## Optimization
-Although the reserve selection problems presented here are non-linear (see Appendix S1 for the reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [@r16].
+Although the reserve selection problems presented here are non-linear (see Appendix S1 for the reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [-@r16].
 The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
@@
 Similarly, $R^2$ statistics can be have negative values when they are describing models that perform worse than a null model.
+This solution performed worse than the solution generated using amount targets and a boundary length modifier to reduce fragmentation (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount target\nand BLM")$space_held, 2)` % sampled; Fig.
+3c) because the latter solution sited more planning units closer to center of the study area--reducing the total distance between the selected planning units and the demand points.
 
@@
 S1).
-This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.
+This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of the variation in some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of the variation in all features.
 
@@
 Conservation planners can use this \texttt{R} package to generate prioritizations that secure intra- and inter-specific biodiversity patterns.
-Both the simulations and case studies show that explicitly targeting a representative sample of each feature in a conservation planning exercise can substantially alter prioritizations.
-Additionally, we found that prioritizations often needed to secure more habitat to conserve a representative sample of each feature.
+Both the simulations and case studies show that explicitly targeting a representative sample of the variation within each feature in a conservation planning exercise can substantially alter prioritizations.
+Additionally, we found prioritizations often needed to secure more habitat to conserve a representative sample of the variation within each feature.
 Although it may not be practical to conserve substantially more habitat than the amount specified using amount-based targets, we did find that even small increases in reserve system size could yield large gains.
@@
 Here, we developed the \texttt{raptr R} package to provide conservation planners with the tools to deliver cost-effective prioritizations that secure an adequate amount of a representative sample of biodiversity features.
-By exploring the functionality of this software package, we found that explicitly targeting a representative sample of each feature can result in substantially different solutions.
+By exploring the functionality of this software package, we found that explicitly targeting a representative sample of the variation in each feature can result in substantially different solutions.
 
@@
 This work was supported by the Centre for Biodiversity and Conservation Science (CBCS).
+We are also grateful to Dan Rosauer and an anonymous reviewer for their suggestions that greatly improved the manuscript.
 

### 2017-06-15 code/rmarkdown/article.Rmd
update ms

@@
 Abstract word count: 346 / 350 \newline
-Total word count: 6883 / 7000 \newline
-Number of references: 50 \newline
+Total word count: 6864 / 7000 \newline
+Number of references: 49 \newline
 Number of figures: 8 \newline
@@
 We demonstrate the functionality of this package using simulations and two case studies.
-In these studies, we generated solutions using amount targets---representing conventional methods---and compared them with solutions generated using amount and space targets.
+In these studies, we generated solutions using amount targets---portraying conventional methods---and compared them with solutions generated using amount and space targets.
 
@@
 One method to incorporate data on the ecological and evolutionary processes that sustain such features is to conserve a representative sample of the internal or external physical attributes underpinning these processes.
-To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448] or phylogenetic trees [@r497].
-For instance, by dividing species distributions into sub-groups according to habitat discontinuities and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
+To achieve this, current methods typically involve partitioning features into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r499; @r500] or phylogenetic trees [@r497].
+For instance, by dividing species distributions into sub-groups according to habitat discontinuities and ensuring that each sub-group is conserved in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
 However, using these methods is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490].
@@
 environmental conditions) to the feature.
-The overarching goal of the prioritization is to conserve these biodiversity processes, and by representing the variation in these species-specific attributes, the prioritization can make some progress towards achieving this.
-Thus there should be a reasonable underlying hypothesis that relates the attributes used to inform the reserve selection processes to the biodiversity process that needs to be conserved.
+The overarching goal of the prioritization is to conserve these biodiversity processes, and by capturing the variation in these species-specific attributes, the prioritization can make some progress towards achieving this.
+Thus there should be a reasonable underlying hypothesis that relates the attributes to the biodiversity processes that need to be conserved.
 
-The variation in the attributes of a feature can be represented using an _n_-dimensional "attribute space".
-For example, a decision maker may require a prioritization that represents populations along climatic gradients.
+The variation in the attributes of a feature can be described using an _n_-dimensional "attribute space".
+For example, a decision maker may require a prioritization that captures populations along climatic gradients.
 To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm).
@@
 mean), they can be mapped from geographic space to this climatic attribute space.
-In addition to points representing planning units, attribute spaces also contain points that describe which attributes should be conserved.
 
-"Demand points" [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets).
+"Demand points" [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to conserve in the prioritization (see below for discussion on generating demand points for real-world data sets).
 For instance, by siting demand points throughout an attribute space, planners can obtain solutions that better capture existing variation.
@@
 Define $F$ to be the set of features one wishes to conserve (indexed by $f$).
-Let $J$ be a set of planning units (indexed by $j$).
-Also, let $A_j$ denote the area, and $C_j$ denote the cost of preserving planning unit $j \in J$.
+Let $J$ be a set of planning units (indexed by $j$), and $C_j$ denote the cost of preserving planning unit $j \in J$.
+Also, let $A_j$ denote the area---or some other metric such as habitat quality or amount of habitat---associated with planning unit $j$.
 To assess the extent to which each feature is secured in a given prioritization, let $q_{fj}$ denote the probability of feature $f$ occupying planning unit $j$.
@@
 Let $S$ denote a set of attribute spaces (indexed by $s$).
-Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$.
+Each $j \in J$ is associated with spatially explicit data that serve as coordinates in each attribute space $s \in S$.
 Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$.
@@
 Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}).
-Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$.
+Let $T_f$ denote the area or amount of habitat that needs to be preserved for each feature $f \in F$.
 Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature.
@@
 Y_{fsij} &= \begin{cases}
-    1, & \parbox{25em}{if demand point $i$ is used to represent planning unit $j$ for feature $f$ in space $s$. } \tag*{eqn 2b} \\
+    1, & \parbox{25em}{if demand point $i$ for feature $f$ in space $s$ is conserved by planning unit $j$. } \tag*{eqn 2b} \\
     0, & \parbox{25em}{otherwise} \\
@@
 
-Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie.
+Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is conserved by a single selected planning unit (ie.
 a $j \in J$ where $X_j = 1$).
-The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$).
-Generally---unless near zero space targets are used so that the problem is effectively unconstrained by the target---demand points are represented by their closest selected planning units.
-In poorer quality solutions, demand points will be represented by planning units that are further away from them.
-As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.
+The degree to which a demand point $i$ is conserved by a planning unit $j$ is determined by the distance between them ($d_{fsij}$).
+Generally---unless near zero space targets are used so that the problem is effectively unconstrained by the target---demand points are conserved by their closest selected planning units.
+In poorer quality solutions, demand points will be conserved by planning units that are further away from them.
+As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to conserve them will be larger, and so, the solution will capture less of the variation described by the demand points.
 
@@
 
-The objective function (eqn 3a) determines the utility of a given prioritization: a combination of the total cost of a prioritization and how fragmented it is.
+The objective (eqn 3a) is to minimize the total cost and fragmentation of the solution.
 Constraints (eqn 3b) ensure that all the amount-based targets are met.
@@
 The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}).
-Presently, academics can obtain a special license for \texttt{Gurobi} at no cost.
+Presently, academics can obtain a \texttt{Gurobi} license at no cost.
 
@@
 This species was most likely to be found in planning units nearest to the center of the study area.
-It was simulated using the density function of a multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b).
+It was simulated using the density function of a multivariate normal distribution (denoted by $\mathcal{N}$; eqn 4b).
 Thirdly, we simulated a species with two distinct populations (hereafter referred to as the "bimodally distributed species"; Fig.
@@
 We generated four solutions for each species.
-First, to represent solutions generated using conventional methods (eg.
+First, to portray solutions generated using conventional methods (eg.
 \texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets.
 Second, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species.
-Third, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`.
+Third, to portray solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`.
 Finally, to illustrate the combined effects of using amount and space targets as well as fragmentation penalties, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % space targets, and boundary length modifiers of `r sim_parameters$blm`.
@@
 We selected Queensland, Australia as the study area, and used an equal area coordinate system for geospatial analyses (Australian Albers GDA94; \texttt{EPSG:3577}).
-To represent planning units for this case-study, we generated a `r cs1_parameters$pu_size/1000` $\times$ `r cs1_parameters$pu_size/1000` km$^2$ grid and clipped the square polygons to the state boundary.
+We used a `r cs1_parameters$pu_size/1000` $\times$ `r cs1_parameters$pu_size/1000` km grid within the state boundary as planning units for this case study.
 We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1).
@@
 To ensure that the solutions did not prioritize particularly costly areas, we obtained population density data [1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r456] and estimated the total population density inside each grid cell.
-We used these values to represent opportunity cost (Fig.
+We used these values to denote opportunity cost (Fig.
 7b).
@@
 For example, adaptation processes could be secured using environmental variation [eg.
-@r247], and trophic processes could be conserved by capturing the overlapping distributions of predator and prey species [eg.
+@r2], and trophic processes could be conserved by capturing the overlapping distributions of predator and prey species [eg.
 @r471; @r485].
@@
 The degree to which a prioritization truly secures a representative sample of a feature depends on (1) the attribute space(s), (2) the distribution of demand points chosen by the conservation planner, and (3) the space target used.
-The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space.
+The optimal solution will not effectively capture the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to construct the attribute space(s).
 Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample (see the `make.DemandPoints` function for statistical routines to achieve this).

### 2017-06-15 code/rmarkdown/figures.Rmd
update ms

@@
 Letters denote the environmental conditions associated with the geographic locations where four hypothetical populations are found.
-Points represent demand points.
+Points denote demand points.
 In this space, populations close to each other inhabit similar environmental conditions."}
@@
 Each point corresponds to a planning unit.
-The coordinates of the points represent the typical genetic characteristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data).
+The coordinates of the points denote the typical genetic characteristics of individuals sampled in that planning unit (based on an NMDS of the binary loci data).
 Planning units associated with points that are closer together contain individuals with more similar genetic characteristics than planning units that are further apart.")}

### 2017-06-15 code/rmarkdown/supporting-information.Rmd
update ms

@@
 # Appendix S1
+
 The reliable formulation explicitly considers the probability that planning units are inhabited by features.
-As a consequence, it may deliver prioritizations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritization is implemented.
+As a consequence, it may deliver prioritizations that sufficiently conserve an attribute space even if the features do not inhabit several of the planning units when the prioritization is implemented.
 This behavior is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s).
-To ensure that prioritizations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.
+To ensure that the solution is robust against multiple planning units being uninhabited, demand points are treated as being conserved by different planning units at different backup levels.
 
-Backup levels levels are defined as $r$-levels [similar to failure levels in @r18].
-The first backup $r$-level is used to calculate the level of representation when all of the selected planning units are occupied by all $f \in F$.
-For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level$=0$.
-This scenario essentially represents $Y_{fsij}$ in the unreliable formulation.
-The second backup $r$-level is used to assess the level of representation when the closest planning unit to each demand point $i$ is unoccupied.
-For this scenario, the second closest planning units are assigned at $r$-level$=1$.
-The third backup $r$-level is used to assess representation when the first two closest planning units are unoccupied.
-The third closest planning units are assigned at $r$-level$=2$.
+Backup levels are defined as $r$-levels [similar to failure levels in @r18].
+The first backup $r$-level is used to calculate how well the demand points are conserved assuming that all the selected planning units are occupied by all $f \in F$.
+For this scenario, the closest selected planning unit to each demand point $i$ for attribute space $s$ is assigned at $r$-level = 0.
+This scenario essentially serves as $Y_{fsij}$ in the unreliable formulation.
+The second backup $r$-level is used to assess how well each demand point is conserved when the closest planning unit to each demand point $i$ is unoccupied.
+For this scenario, the second closest planning units are assigned at $r$-level = 1.
+The third backup $r$-level is used to assess how well demand points are captured when the first two closest planning units are unoccupied.
+The third closest planning units are assigned at $r$-level = 2 .
 Continuing on, in this manner, the selected planning units in a prioritization are assigned to each demand point $i \in I$, attribute space $s \in S$, and each feature $f \in F$ at an $r$-level.
 
-A final backup $r$-level when $r=R$ is used to assess the level of representation when the features $f \in F$ do not occupy any selected planning units in a prioritisation.
+A final backup $r$-level ($r=R$) is used to describe how well demand points are captured by a solution when the features $f \in F$ do not occupy any selected planning units.
 Each demand point $i \in I$ for each $s \in S$ and $f \in F$ is assigned to an "imaginary" planning unit $j=J$ at $r=R$.
@@
 
-Conservation planning problems often use more than several hundred planning units.
-It is currently not be feasible to solve this problem when considering all possible failure scenarios within a reasonable amount of time (see Appendix S2).
-As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$.
+Conservation planning problems often involve several hundred or more planning units.
+It is currently not be feasible to solve problems this large when considering all possible failure scenarios (see Appendix S2).
+To reduce computational burden, users can set a smaller number to denote the maximum number of backup levels ($R$).
 For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level.
-Cui _et al._ [-@r16] found that $R=5$ yields similar solutions to $R >> 5$.
-However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritizations in a feasible amount of time.
+Cui _et al._ [-@r16] found that $R=5$ yields comparable solutions to when $R$ is much larger.
+However, in most cases the decision maker will be limited to $R=1$ to obtain prioritizations in a feasible amount of time.
 
@@
 \begin{align*}
-R &= \parbox{25em}{number of failure levels} \tag*{eqn 6a} \\
+R &= \parbox{25em}{number of backup levels} \tag*{eqn 6a} \\
 %
@@
 Y_{fsijr} &= \begin{cases}
-    1, & \parbox{25em}{if demand point $i$ is assigned to planning unit $j$ for feature $f$ in space $s$ at back-up level $r$. } \tag*{eqn 7a} \\
+    1, & \parbox{25em}{if demand point $i$ for feature $f$ in space $s$ is conserved by planning unit $j$ at back-up level $r$. } \tag*{eqn 7a} \\
     0, & \parbox{25em}{otherwise} \\
@@
 %
-P_{fsijr} &= \parbox{25em}{probability that demand point $i$ is assigned to planning unit $j$ at back-up level $r$ for feature $f$ and space $s$} \tag*{eqn 7b}\\
+P_{fsijr} &= \parbox{25em}{probability that demand point $i$ for feature $f$ in space $s$ will need to be conserved by planning unit $j$ at back-up level $r$} \tag*{eqn 7b}\\
 %
@@
 However, the non-linear components can be linearized.
-First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al.
+First---as discussed in the main text---the expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al.
 [-@r426].
@@
 The features' demand points were set as the centroids of the planning units that they occupied, and were weighted by the units' probability of occupancy.
-After simulating the data sets, solutions were generated using the \texttt{raptr R} package and Gurobi (version 7.0.2).
+After simulating the data sets, solutions were generated using the \texttt{raptr R} package and \texttt{Gurobi} (version 7.0.2).
 
 We generated solutions for the conservation planning data sets using different formulations and parameters.
-Specifically, the solutions were generated using the unreliable and reliable formulation, and several boundary length multipliers (`r paste(unique(benchmark_parameters$blm), sep = "and")`).
+Specifically, the solutions were generated using the unreliable and reliable formulation, and two different boundary length multipliers (`r paste(unique(benchmark_parameters$blm), collapse = " and ")`).
 The solutions were generated using amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100` %.
@@
 The solve times could be further reduced by using more threads.
-We were not able to determine how long problems bigger problems would take to solve using the reliable formulation because the system did not have enough RAM.
-Future research could involve implementing more advanced methods to solve the optimization problems [eg.
+We were not able to determine how long bigger problems would take to solve using the reliable formulation because the system did not have enough RAM.
+Future research could involve implementing more advanced methods to solve the problems [eg.
 as described in @r16; @r15].
@@
 The color of each hexagon denotes the weighted frequency of demand points inside it.
-Points represent the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution."}
+Points denote the environmental conditions associated with planning units inside the species geographic range that were selected for preservation in a given solution."}
 ## prepare data
@@
 ```{r, echo = FALSE, message = FALSE, fig.height = 5.0, fig.width = 4.5, fig.cap = "Time required to solve different sized problems using different parameters and formulations.
-Points represent average times and bars show standard errors.
+Points denote average times and bars show standard errors.
 Time is shown on a log$_{10}$ scale."}
@@
 terms <- c(
+  "$A_j$", "area of planning unit $j$---note this can also be some other measure of conservation utility",
+  "$B$", "boundary length modifier (BLM)",
+  "$d_{fsij}$", "distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$",
+  "$e_{jk}$", "shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.",
   "$F$", "set of biodiversity features (indexed by $f$)",
+  "$I_{fsi}$", "set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$",
   "$J$", "set of planning units (indexed by $j$)",
+  "$M$", "failure multiplier",
+  "$P_{fsijr}$", "probability that demand point $i$ for feature $f$ in space $s$ will need to be conserved by planning unit $j$ at back-up level $r$",
   "$q_{fj}$", "probability of feature $f$ occupying planning unit $j$",
-  "$e_{jk}$", "shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrize exposed edges with no neighbours.",
+  "$R$", "number of backup levels (indexed by $r$)",
   "$S$", "set of attribute spaces (indexed by $S$)",
-  "$I_{fsi}$", "set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$",
+  "$T_f$", "amount target for feature $f$",
+  "$X_j$", "binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)",
+  "$Y_{fsij}$", "binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in attribute space $s$ when determining the amount of the attribute space sampled by the selected planning units (unreliable version)",
+  "$Y_{fsijr}$", "binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in attribute space $s$ at $r$-level $r$ when determining the amount of the attribute space sampled by the selected planning units (reliable version)",
+  "$\\delta_{fsi}$", "the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$",
   "$\\lambda_{fsi}$", "set of weights for demand point $i$ for feature $f$ in attribute space $s$",
-  "$d_{fsij}$", "distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$",
-  "$\\delta_{fsi}$", "the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$",
-  "$T_f$", "amount target for feature $f$",
-  "$\\tau_{fs}$", "space-based target for feature $f$ in attribute space $s$",
-  "$X_j$", "binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)",
-  "$Y_{fsij}$", "binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determining the amount of the attribute space sampled by the selected planning units"
+  "$\\tau_{fs}$", "space-based target for feature $f$ in attribute space $s$"
 )
@@
                                      "representative and adequate ",
-                                     "prioritization (URAP) problem."))
+                                     "prioritization (URAP) problem and reliable version (RRAP)."))
 curr_file %>%

### 2017-07-06 code/rmarkdown/article.Rmd
small fixes to typos

@@
 
-Abstract word count: 346 / 350 \newline
-Total word count: 6864 / 7000 \newline
+Abstract word count: 349 / 350 \newline
+Total word count: 6855 / 7000 \newline
 Number of references: 49 \newline
@@
 Users set "amount targets"---similar to conventional methods---to ensure that solutions secure a sufficient proportion of area or habitat for each feature.
-Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg.
+Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionarily relevant attributes (eg.
 environmental or genetic variation).
 We demonstrate the functionality of this package using simulations and two case studies.
-In these studies, we generated solutions using amount targets---portraying conventional methods---and compared them with solutions generated using amount and space targets.
+In these studies, we generated solutions using amount targets---similar to conventional methods---and compared them with solutions generated using amount and space targets.
 
 3.
-Our results demonstrate that different solutions emerge when targeting a representative sample of each feature.
+Our results demonstrate that markedly different solutions emerge when targeting a representative sample of each feature.
 We show that using these targets is important for features that have multimodal distributions in the process-related attributes (eg.
 species with multimodal niches).
-We also found that solutions could conserve a far more representative sample with only a slight increase in reserve size.
+We also found that solutions could conserve a far more representative sample with only a slight increase in reserve system size.
 
@@
 Additionally, different features and attribute spaces may require different demand points.
-For instance, points that may be valid in one attribute space (eg.
+For instance, values that might be valid in one attribute space (eg.
 mean annual temperature $-5^{\circ}$C) may be invalid for another attribute space (eg.
-mean annual rainfall -5 mm).
+mean annual rainfall $-5$ mm).
 Additionally, there may be some regions that are desirable for some features and undesirable for others (eg.
@@
 Space-based targets are expressed as proportions---instead of a sum of weighted distances---by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid.
-This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _k_-means cluster analyses from the within sums of squares and total sums of squares [page 106; @r491].
+This scaling is conceptually similar to that used in calculating the $R^2$ statistic for _k_-means cluster analyses from the within sums of squares and total sums of squares [page 106; @r491].
 Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$.
-The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.
+The control variables for the unreliable formulation are $B$, $T_{s}$, and $\tau_{fs}$.
 
@@
 
-The decision variables are the $X_j$ and $Y_{fsij}$ variables.
+The decision variables are $X_j$ and $Y_{fsij}$.
 
@@
 The objective (eqn 3a) is to minimize the total cost and fragmentation of the solution.
-Constraints (eqn 3b) ensure that all the amount-based targets are met.
-Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space.
+Constraints ensure that all amount-based targets are met (eqn 3b), and all space-based targets are met for each feature and each attribute space (eqn 3c).
 By treating targets as constraints, rather than including them in the objective function, all feasible solutions will fulfill the targets.
@@
 For this scenario, all solutions containing `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` units are optimal because this species has an equal chance of occurring in any given planning unit.
-Although it may seem odd that the solution prioritized planning units along the southern end of the study area, this behavior occurred because there were many optimal solutions and the solver---without any other criteria to compare solutions---returned a solution that reflected the order that the data was encoded in the optimization problem.
+Although it may seem odd that the solution prioritized planning units along the southern end of the study area, this behavior occurred because there were many optimal solutions and the solver---without any other criteria to compare solutions---returned a solution that reflected the order that data were encoded in the optimization problem.
 As we can see in later examples, this behavior only manifests when there are many optimal solutions.
@@
 However, unlike the solutions for the uniformly distributed species, the solution generated using an amount target, a space target, and fragmentation penalties required more planning units than the other solutions.
-These results suggest that solutions may need to prioritize more planning units to meet additional conservation objectives.
+These results suggest that solutions may need to select more planning units to meet additional conservation objectives.
 
@@
 5).
-To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km of each other to ameliorate the effects of sampling bias [using a modified version of the \texttt{spThin R} package; \url{https://github.com/jeffreyhanson/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
 We used this method because it is entirely reproducible using freely available data.
@@
 
+# Author contributions
+
+JOH performed the analysis and drafted the manuscript. All authors developed the study and edited the manuscript.
+
 # Data accessibility

### 2017-07-06 code/rmarkdown/figures.Rmd
small fixes to typos

@@
 Panel (a) shows the planning units selected when using ",cs2_parameters$amount_target * 100," % amount targets.
-Panel (b) shows the planning units selected when using ",cs2_parameters$amount.target*100," % amount targets and ",cs2_parameters$genetic_target * 100," % space targets.
+Panel (b) shows the planning units selected when using ",cs2_parameters$amount_target * 100," % amount targets and ",cs2_parameters$genetic_target * 100," % space targets.
 Panels (c--d) show the solutions in the genetic space.

### 2017-07-06 code/rmarkdown/article.Rmd
add doi

@@
 The \texttt{raptr R} package can be downloaded from The Comprehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
-All data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
+To permit replication and validation of this study, all data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) and are available at the DOI: \texttt{10.5281/zenodo.823479}.
 

### 2017-07-06 code/rmarkdown/article.Rmd
add zenodo doi

--- 2017-07-06 code/rmarkdown/article.Rmd
@@
 The \texttt{raptr R} package can be downloaded from The Comprehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
-All data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
+To permit replication and validation of this study, all data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) and are available at the DOI: \texttt{10.5281/zenodo.823768}.
 

### 2017-08-22 code/rmarkdown/article.Rmd
fix minor typos and diversity ref

@@
 For instance, by dividing species distributions into sub-groups according to habitat discontinuities and ensuring that each sub-group is conserved in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2].
-However, using these methods is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490].
+However, using these methods is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups can substantially alter solutions [@r490].
 This limitation has been known for quite some time, and dates back to some of the earliest reserve selection methods [eg.
@@
 
-To overcome this limitation, Faith and Walker (1996) developed the "environmental diversity" (ED) reserve selection framework to conserve a representative sample of the environmental variation found across a study area.
-Drawing inspiration from the _p_-median problem [@r498], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; @r204].
+To overcome this limitation, Faith and Walker (1994) developed the "environmental diversity" (ED) reserve selection framework to conserve a representative sample of the environmental variation found across a study area.
+Drawing inspiration from the _p_-median problem [@r480], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; @r204].
 Recent work has built on this method using more advanced optimization algorithms [@r482].
@@
 Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$.
-To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.
+To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points in space $s$.
 
@@
 \begin{align*}
-& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) + & & \tag*{eqn 3a} \\
+& \text{(URAP)} & \text{Min } & \sum_{j=0}^{J-1} \left( X_j C_j \right) + \sum_{j=0}^{J-1} \sum_{k=j}^{J-1} X_j \left( 1-X_k \right) \left( B e_{jk} \right) & & \tag*{eqn 3a} \\
 %
@@
 To assess how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2).
-We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality.
+We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality using \texttt{Gurobi} (version 7.0.2).
 
@@
 \\
-P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
+P \left( \text{uniformly distributed species} | \left( X, Y \right) \right) &= 0.5 \tag*{eqn 4a} \\
 %
-P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
+P \left( \text{normally distributed species} | \left( X, Y \right) \right) &= \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
 %
-P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
+P \left( \text{bimodally distributed species} | \left( X, Y \right) \right) &= \text{Max} \begin{cases}
 \mathcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\mathcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
@@
 3d).
-This solution prioritized the same number of planning units as the previous solutions even though it is far superior.
+This solution prioritized a similar number of planning units as the previous solutions even though it is far superior.
 As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.
@@
 5).
-To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km of each other to ameliorate the effects of sampling bias [using a modified version of the \texttt{spThin R} package; \url{https://github.com/jeffreyhanson/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
+To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km of each other to ameliorate the effects of sampling bias [using a modified version of the \texttt{spThin R} package; \url{www.github.com/jeffreyhanson/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455].
 We used this method because it is entirely reproducible using freely available data.
@@
 # Acknowledgements
-JOH is funded by an Australian Postgraduate Award (APA) scholarship.
+JOH is supported by an Australian Government Research Training Program (RTP) Scholarship.
 RAF and HPP have Australian Research Council Fellowships.

### 2017-08-22 code/rmarkdown/figures.Rmd
figures and readme tweaks
fix minor typos and diversity ref

@@
 Each row of panels show prioritizations generated for a different species.
-Each column of panels corresponds to a different set of parameters used to generate the prioritization.")}
+Each column of panels corresponds to a different set of parameters used to generate the prioritizations.")}
 # prepare data
@@
                         ggplot2::aes(x = long, y = lat, group = group),
-                        fill = "grey85", color = "grey70") +
+                        fill = "grey85", color = "grey70", size = 0.05) +
   ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
@@
                         ggplot2::aes(x = long, y = lat, group = group),
-                        fill="grey85", color="grey70") +
+                        fill="grey85", color="grey70", size = 0.05) +
   ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
@@
                         ggplot2::aes(x = long, y = lat, group = group),
-                        fill = "grey85", color = "grey70") +
+                        fill = "grey85", color = "grey70", size = 0.05) +
   ggplot2::geom_polygon(data = cs2_spp_grid_fpolygons,
@@
                         ggplot2::aes(x =long, y = lat, group = group),
-                        fill = "grey85", color = "grey70") +
+                        fill = "grey85", color = "grey70", size = 0.05) +
   ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
@@
                         ggplot2::aes(x = long, y = lat, group = group),
-                        fill = "grey85", color = "grey70") +
+                        fill = "grey85", color = "grey70", size = 0.05) +
   ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
@@
                         ggplot2::aes(x = long, y = lat, group = group),
-                        fill="grey85", color="grey70") +
+                        fill="grey85", color="grey70", size = 0.05) +
   ggplot2::geom_polygon(data = cs2_spp_grid_sub_fpolygons,
@@
                         ggplot2::aes(x = long, y = lat, group = group),
-                        fill = "grey85", color = "grey70") +
+                        fill = "grey85", color = "grey70", size = 0.05) +
   ggplot2::geom_polygon(data = cs2_sel_grid_sub_fpolygons,
@@
                         ggplot2::aes(x=long, y=lat, group=group),
-                        fill="grey85", color="grey70") +
+                        fill="grey85", color="grey70", size = 0.05) +
   ggplot2::geom_polygon(data = cs2_sel_grid_sub_fpolygons,

### 2017-08-22 code/rmarkdown/article.Rmd
update doi

@@
 The \texttt{raptr R} package can be downloaded from The Comprehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
-To permit replication and validation of this study, all data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) and are available at the DOI: \texttt{10.5281/zenodo.823768}.
+To permit replication and validation of this study, all data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) and are available at the DOI: \texttt{10.5281/zenodo.846477}.
 

### 2017-08-23 code/rmarkdown/article.Rmd
add gurobi version

@@
 To assess how long it would take to solve various sized problems, we also conducted a benchmark analysis (Appendix S2).
-We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality using \texttt{Gurobi} (version 7.0.2).
+We completed the analyses using \texttt{R} [version `r getRversion()`; @r192] and \texttt{raptr} (version 0.0.3).
+We solved all optimization problems to within `r general_parameters$MIPGap * 100` % of optimality using \texttt{Gurobi} (version 7.0.2).
 
@@
 The \texttt{raptr R} package can be downloaded from The Comprehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}).
-To permit replication and validation of this study, all data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) and are available at the DOI: \texttt{10.5281/zenodo.846477}.
+To permit replication and validation of this study, all data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) and are available at the digital object identifier (DOI) \texttt{10.5281/zenodo.847016}.
 
